---
title: "R Notebook"
output: html_document
---


## NOTE: this analysis is performed on O2 cluster instead of locally


```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/01a_Ped/", 
                     "02_ClonalOverexpression/")

library(Seurat)
library(ggplot2)
library(ggpubr)
library(Rtreemix)

figure_dir_tmp<- "figures/"
analysis_dir_tmp<- "analysis/"

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(nmf_helper)
clonalOverexpression<- paste0(base_dir, script_dir, "ClonalVsNonClonal_helper.R")
source(clonalOverexpression)

```


## Load in preprocessed list of clonal/nonclonal seurat objects
## preprocessed in 01_ClonalVsNonClonal_Preprocessing.Rmd
```{r}
rank_use<-8
sharedBy<- "union" 
analysis_append<- ""

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}


output_dir<- paste0(analysis_dir, "WithBackground100x")
seurat_list<- readRDS(paste0(analysis_dir, "ped_cd8_seurat_cnc_preprocessed", analysis_append, ".Rds"))
analysis_name<- "ped_cd8_all"
n_iterations<-100
```



## Expression score:
1. take each clonotype in each sample and select 10 sets of non-clonal cells. Each set has the same number of cells as the clonotype
2. Get mean expression of all genes across ALL clonal cells and across ALL randomly selected non-clonal cells
3. subtract clonal mean exp - nonclonal mean exp

## Distance measurement:
1. Within each clonotype, calculate the L1 distance between ALL cells within that clonotype, for each gene separately. Performed for all possible combinations within that clonotype
2. Calculate average distance for each gene across all clonotypes (across all samples)
3. Divide this mean L1 value by that gene's mean expression (from above). This is the CV value
4. Repeat this for nonclonal- using all combinations of cells within each of the random sets


```{r}
##########################
## With correct labels ##
#########################

## Calculate clonal/nonclonal expression
## Cycle through all samples, calculating mean expression for each clonotype/matched nonclonal cells
## Also save list of clonal/nonclonal cells with clonotype info retained for CV analysis
expr_list<- Exp_WrapperForMultipleSamples(seurat_list, min_clonotype_size = 1)

## Calculate mean clonal/nonclonal expression, merge, and calculate zscores
expr_CNC_mean<- Exp_ProcessCounts(expr_list$counts)

## Check expression of genes of interest
expr_CNC_mean[c("KLRB1", "GZMB", "NKG7", "IL7R", "CCR7", "CCL5", "KLRK1", "GPR56", "CCL4", "ITGAE", "KLRD1"),]

## Clonal CV
# Merge clonotypes- each clonotype should be unique, and there can be cells from different samples in the same clonotype
#cv_list<- CV_MergeClonotypes(expr_list)
cv_list<- expr_list

## Cycle through each clonotype, create list of all unique cell combinations, and calculate distance between all genes for all combinations
cv_list<-CV_CalculateL1Dist(cv_list, "clonal", seurat_list, tpm_or_norm = "norm")

## Average distance for all genes
CV_clonal_res<- CV_ProcessCV(cv_list$CVres_l1Dist, "clonal")
CV_clonal_res[c("KLRB1", "GZMB", "NKG7", "IL7R", "CCR7", "CCL5", "KLRK1", "GPR56", "CCL4", "ITGAE", "KLRD1"),]

## NonClonal CV
## For nonclonal, need to add Set# (1:10) to cell name
cv_list<- CV_AddSetToNonclonalCells(cv_list)

## Cycle through each clonotype, create list of all unique cell combinations, and calculate distance between all genes for all combinations
cv_list<-CV_CalculateL1Dist(cv_list, "nonclonal", seurat_list, verbose = FALSE, tpm_or_norm = "norm")

## Average distance for all genes
CV_nonclonal_res<- CV_ProcessCV(cv_list$CVres_l1Dist, "nonclonal")

## Merge counts/L1 distances, calculate CV, calculate CV zscore
Expr_CV_final<- MergeAndProcess_Counts.L1Dist(CV_clonal_res, CV_nonclonal_res, expr_CNC_mean)
actual_Expr_CV_final<- Expr_CV_final; rm(Expr_CV_final)

##########################
## With shuffled labels ##
##########################

## merge clonal/nonclonal
seurat_merge<- merge(seurat_list$clonal, seurat_list$nonclonal)
seurat_shuffle<- seurat_merge
cm<- seurat_merge@assays$RNA@counts
cm_shuffle<- cm

## list for adding expr/CV differences for shuffled datasets
shuffled_exprCV_diff<- list()

test<- list(seurat_shuffle=list(),
            cm_shuffle=list(),
            seurat_switch=list(),
            expr_list=list(),
            expr_CNC_mean=list(),
            cv_list=list(),
            CV_clonal_res=list(),
            CV_nonclonal_res=list())


for(i in 1:n_iterations){
  print(paste0("iteration number ", i))
  ## shuffle gene expression values for each gene
  cm_shuffle<- as.data.frame(t(apply(cm_shuffle, 1, function(x){sample(x)})))
  colnames(cm_shuffle)<- colnames(cm); cm_shuffle<- as.matrix(cm_shuffle)
  seurat_shuffle@assays$RNA@counts<- cm_shuffle
  test$cm_shuffle[[i]]<-cm_shuffle
  #test$seurat_shuffle[[i]]<- seurat_shuffle
  
  ## split back to clonal/nonclonal
  seurat_switch<- list(clonal=subset(seurat_shuffle, clonal=="clonal"),
                       nonclonal=subset(seurat_shuffle, clonal=="nonclonal"))
  #test$seurat_switch[[i]]<-seurat_switch
  
  ## Calculate mean expression values, merge clonal/nonclonal, calculate zscores
  expr_list<- Exp_WrapperForMultipleSamples(seurat_switch, min_clonotype_size = 1)
  expr_CNC_mean<- Exp_ProcessCounts(expr_list$counts)
  #test$expr_list[[i]]<- expr_list 
  
  ## Clonal: merge clonotypes, Calculate L1 distances, calculate zscores
  cv_list<- CV_MergeClonotypes(expr_list)
  cv_list<-CV_CalculateL1Dist(cv_list, "clonal", seurat_switch, verbose = FALSE, tpm_or_norm = "norm")
  CV_clonal_res<- CV_ProcessCV(cv_list$CVres_l1Dist, "clonal")
  #test$cv_list[[i]]<- cv_list 
  
  ## Nonclonal: add to "clonotype" name, calculate L1 distances, calculate zscores
  cv_list<- CV_AddSetToNonclonalCells(cv_list)
  cv_list<-CV_CalculateL1Dist(cv_list, "nonclonal", seurat_switch, verbose=FALSE, tpm_or_norm = "norm")
  CV_nonclonal_res<- CV_ProcessCV(cv_list$CVres_l1Dist, "nonclonal")
  
  ## Merge counts/L1 distances, calculate CV, calculate CV zscore
  Expr_CV_final<- MergeAndProcess_Counts.L1Dist(CV_clonal_res, CV_nonclonal_res, expr_CNC_mean)
  
  ## There may be a few genes that are not included in shuffled results
  ## if all values for a gene are shuffled so that they're 0 for clonal or nonclonal, they will be excluded from results
  ## Should just be a couple genes, may be 0 genes
  ## Remove any here
  genes_remove<- rownames(Expr_CV_final)[!rownames(Expr_CV_final) %in% rownames(actual_Expr_CV_final)]
  print(paste0("genes removed: ", paste(genes_remove, collapse=", ")))
  Expr_CV_final<-Expr_CV_final[!rownames(Expr_CV_final) %in% genes_remove,]
  
  ## add iteration number
  Expr_CV_final$iteration<- i
  
  ## add gene id as column
  Expr_CV_final$gene<- rownames(Expr_CV_final)
  
  shuffled_exprCV_diff[[i]]<- Expr_CV_final
}

## merge into single df
shuffled_exprCV_diff_df<- do.call("rbind", shuffled_exprCV_diff)

## calculate zscore comparing actual expr differences vs background distribution of shuffled expr differences
## for each gene: actual gene expr diff - mean(background gene expr diff) / sd(background gene expr diff)
expr_zscores<- sapply(rownames(actual_Expr_CV_final),function(gene){
  actual_expr_diff<- actual_Expr_CV_final[gene, "Expr_CminusNC"]
  background_expr_diff<- shuffled_exprCV_diff_df[shuffled_exprCV_diff_df$gene %in% gene, "Expr_CminusNC"]
  expr_zscore<- (actual_expr_diff - mean(background_expr_diff))/sd(background_expr_diff)
  return(expr_zscore)
})
zscores_df<- as.data.frame(expr_zscores)

## calculate zscore comparing actual CV differences vs background distribution of shuffled CV differences
## for each gene: actual gene CV diff - mean(background gene CV diff) / sd(background gene CV diff)
cv_zscores<- sapply(rownames(actual_Expr_CV_final),function(gene){
  actual_cv_diff<- actual_Expr_CV_final[gene, "CV_CminusNC"]
  background_cv_diff<- shuffled_exprCV_diff_df[shuffled_exprCV_diff_df$gene %in% gene, "CV_CminusNC"]
  cv_zscore<- (actual_cv_diff - mean(background_cv_diff))/sd(background_cv_diff)
  return(cv_zscore)
})
cv_zscores_df<- as.data.frame(cv_zscores)
zscores_df<- merge(zscores_df, cv_zscores_df, by=0)
rownames(zscores_df)<- zscores_df$Row.names; zscores_df<- zscores_df[,-1]
colnames(zscores_df)<- c("Expr_background_zscore", "CV_background_zscore")

## merge with actual results
actual_Expr_CV_final<- merge(actual_Expr_CV_final, zscores_df, by=0)
rownames(actual_Expr_CV_final)<- actual_Expr_CV_final$Row.names; actual_Expr_CV_final<- actual_Expr_CV_final[,-1]

## Replace NAs with 0- or remove
#actual_Expr_CV_final[is.na(actual_Expr_CV_final)]<- 0
actual_Expr_CV_final<- actual_Expr_CV_final[!is.na(actual_Expr_CV_final$CV_background_zscore),]

saveRDS(actual_Expr_CV_final, 
     file=paste0( "WithBackgroundx", n_iterations, "_ClonalVsNonclonal_Expression.CV_Zscores_", analysis_name, ".Rds"))
write.csv(actual_Expr_CV_final, file=paste0("WithBackgroundx", n_iterations, "_ClonalVsNonclonal_Expression.CV_Zscores_", 
                                            analysis_name, ".csv"))

saveRDS(shuffled_exprCV_diff_df, 
     file=paste0("Backgroundx", n_iterations, "_AllIterations_", analysis_name, ".Rds"))
```

