---
title: "ImmuneRegev_ependymoma_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/BySubtype/Ependymoma/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
stat_helper = paste0(base_dir, script_dir, "common_stat_test.R")
source(preprocessing)
source(stat_helper)

liger_analysis_folder = "analysis/liger/"
seurat_analysis_folder = "analysis/seurat/"
seurat_fig_folder = "figures/seurat/"
for (folder in c(seurat_analysis_folder, seurat_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}
```

## Load counts and metadata for initiating seurat object
```{r load-data}
input_dir<- "../../Preprocessing/SplitBySubtype/analysis/"
load(paste0(input_dir, "epn_cm_list.Robj"))
samples = readRDS(paste0(input_dir,"epn_samples.rds"))
tumor_stat = readRDS(paste0(input_dir,"epn_call_tumor.rds"))
cm_list<-epn_cm_list

sum(colnames(cm_list$raw_data)==names(samples)); length(samples)
sum(names(samples)==names(tumor_stat))
```

## Subset cm to remove noise/normal samples
```{r subset-tumor-cells}
cm = cm_list$raw_data; rm(cm_list); gc()
cm = cm[,tumor_stat=="Malignant"]
samples = samples[tumor_stat=="Malignant"]

any(colnames(cm) != names(samples))
```

## Initiate seurat obj and add metadata 
```{r seurat_obj_meta_data}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", min.cells=0, min.genes=0,  scale.factor=1E5, do.scale=F, do.center=T)

## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples

subtypes<- read.csv("../Subtypes.csv")
sample_df<- as.data.frame(samples); sample_df$cell<- rownames(sample_df)
subtype_df<- merge(sample_df, subtypes, by="samples")
seurat_obj@meta.data$Subtype<- subtype_df$Diagnosis

```

## OPTIONAL: plot highly variable genes
```{r}
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## PCA and PC selection
```{r PCA_selection}
## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)
```

## Harmony data integration 
```{r run_harmony}
library(harmony)
library(magrittr)
library(cowplot)

## Run Harmony
seurat_obj = RunHarmony(seurat_obj, "sample", theta = 2, 
                        max.iter.harmony = 50, plot_convergence = TRUE)

## Simple dim and vlnplot to examine integration 
options(repr.plot.height = 6, repr.plot.width = 12)

p1 <- DimPlot(object = seurat_obj, reduction = "pca", pt.size = .1, group.by = "sample")
p2<- DimPlot(object = seurat_obj, reduction = "harmony", pt.size = .1, group.by = "sample")
p3 <- VlnPlot(object = seurat_obj, features = "PC_1", pt.size = .1, group.by = "sample")
p4 <- VlnPlot(object = seurat_obj, features = "harmony_1", pt.size = .1, group.by = "sample")

plot_grid(p1,p2, p3, p4)
ggsave("Harmony_diagnostics_theta2.jpg", path=seurat_fig_folder, width=12, height=8)


## DimHeatmap
DimHeatmap(object = seurat_obj, reduction = "harmony", cells = 500, dims = 1:6, balanced = TRUE)
```

## Clustering and UMAP
```{r clustering_tSNE}
## Run UMAP and clustering using reduced dimensions from Harmony
## You may play around with resolution parameter to tune the clustering result
resolution = 1
reduction_type = "harmony"
dims = 20

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = reduction_type, dims = 1:dims)


seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = reduction_type,
                            dims = 1:dims,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = resolution)


r0.8<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5) + ggtitle("Resolution_0.8")
r0.5<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5)+ ggtitle("Resolution_0.5")
r1.0<-DimPlot(seurat_obj, reduction = "umap", pt.size = 2.5)+ ggtitle("Resolution_1")

r0.5+r0.8+r1.0 #resolution of 0.8 may be best
ggsave("UMAP_clustering_VaryResolution.jpg", path=seurat_fig_folder, width=15, height=6)

## Clustering results 
table(Idents(seurat_obj))


## Save the seurat object
save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```

## UMAP plot colored by clustering and samples
## Can color tSNE with other metadata features (e.g. primary/recurrence, treatment)
```{r UMAP_plots}
## By clustering
DimPlot(object = seurat_obj, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering.jpg", path=seurat_fig_folder, width=8, height=8)

## By samples
DimPlot(object = seurat_obj, reduction="umap", group.by = "sample", label = FALSE, 
        pt.size = 2, label.size = 8) + NoAxes() 
ggsave("UMAP_sample.jpg", path=seurat_fig_folder, width=8, height=8)

## By additional subtype
DimPlot(object = seurat_obj, reduction="umap", group.by = "Subtype", label = FALSE, 
        pt.size = 2, label.size = 8) + NoAxes() 
ggsave("UMAP_subtype.jpg", path=seurat_fig_folder, width=8, height=8)
```

## Plot breakdown of each cluster by sample
```{r}
## By sample
cluster_by_sample<- table(seurat_obj@meta.data$sample, seurat_obj@meta.data$seurat_clusters)
cluster_by_sample<- melt(cluster_by_sample)
colnames(cluster_by_sample)<- c("Sample", "Cluster", "NCells")
cluster_by_sample$Cluster<- paste0("Cluster", cluster_by_sample$Cluster)

ggplot(cluster_by_sample, aes(x=Cluster, y=NCells, fill=Sample))+ 
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Seurat clustering: \nbreakdown of samples per cluster")
ggsave(paste0(seurat_fig_folder, "SeuratClusterSampleDistr.png"), width=3, height=4)

## By subtype
cluster_by_subtype<- table(seurat_obj@meta.data$Subtype, seurat_obj@meta.data$seurat_clusters)
cluster_by_subtype<- melt(cluster_by_subtype)
colnames(cluster_by_subtype)<- c("Subtype", "Cluster", "NCells")
cluster_by_subtype$Cluster<- paste0("Cluster", cluster_by_subtype$Cluster)

ggplot(cluster_by_subtype, aes(x=Cluster, y=NCells, fill=Subtype))+ 
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Seurat clustering: \nbreakdown of subtype per cluster")
ggsave(paste0(seurat_fig_folder, "SeuratClusterSubtypeDistr.png"), width=3, height=4)
```


## DEGs for each cluster 
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.5)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05 & all_markers$avg_logFC > 0.5, ]


## Subset the top 30 DEGs for each cluster
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC) %>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(seurat_analysis_folder, "all_markers.Robj"))
write.csv(top_marker_genes, file=paste0(seurat_analysis_folder, "Top30Markers.csv"))

## Heatmap for initial inspection
DoHeatmap(seurat_obj, 
          features=all_markers_filtered$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs.jpg", path=seurat_fig_folder)
```

## Quick annotation by OPC/OC/AC marker genes 
```{r}
## OPC-like
FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OPC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## OC-like cells
FeaturePlot(seurat_obj,
            c("BCAS1", "SIRT2", "RGR", 
              "GPR17", "NFASC", "MBP",
              "TF", "MOG", "PLP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## AC-like cells 
FeaturePlot(seurat_obj,
            c("APOE", "AQP4", "AGT", 
              "SPARCL1","ATP1A2", "GFAP",
              "S100B", "ALDOC", "CLU"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("AC_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## Neuronal-like cells
FeaturePlot(seurat_obj,
            c("STMN2", "STMN1", "STMN4", 
              "TUBB3", "INA", "ELAVL4",
              "SNAP25", "GABRG2", "ST18"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("Neuron_markers.jpg", path=seurat_fig_folder, width=9, height=9)


##CLEC2D
FeaturePlot(seurat_obj,
            c("CLEC2D"),
            cols = c("lightgrey", "red"),
            ncol = 1)
ggsave("CLEC2D_markers.jpg", path=seurat_fig_folder, width=7, height=7)


##CLEC2D
FeaturePlot(seurat_obj,
            c("CLEC2D", "CD274", "PDCD1LG2", "LGALS9", "HFREP1"),
            cols = c("lightgrey", "red"),
            ncol = 4)
ggsave("Ligand_markers.jpg", path=seurat_fig_folder, width=18, height=4.5)
```

## ORA test 
```{r over-representation-test}
library(clusterProfiler)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
allOE_genes <- as.character(gene_symbol_to_ensembl_id(rownames(seurat_obj@assays$RNA@counts),dataset)$ensembl_gene_id)


## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
go_result = list()
for (i in unique(all_markers_filtered$cluster)){
  sigOE_genes = as.character(gene_symbol_to_ensembl_id(subset(all_markers_filtered, cluster==i,gene)[,], dataset)$ensembl_gene_id)
  go_result[[paste0("c", i)]] = go_analysis(sigOE_genes, allOE_genes)
} 


## Plot top pathways for all clusters, display on one plot
clusters = paste0("c", sort(levels(seurat_obj@active.ident)))
allDotPlots_20<-list()
for (c in 1:length(levels(seurat_obj@active.ident))){
  cluster=clusters[c]
  print(cluster)
  if (nrow(go_result[[cluster]]$cluster_summary)>0){
    print(cluster)
    allDotPlots_20[[c]]<-dotplot(go_result[[cluster]]$ego, 
                                showCategory=20, 
                                font.size = 15, 
                                title = cluster, 
                                label_format=10) +
      scale_y_discrete(labels=function(x) str_wrap(x, width=60))+
      theme(plot.title = element_text(hjust = 0.5, face = "bold", 
                                      color="black", size = 28),
            axis.title = element_text(face = "bold", color="black"), 
            axis.text.x = element_text(angle = 45, hjust = 1, color="black"),
            axis.text.y = element_text( color="black",face = "bold"))
  }
  
}

allDotPlots_20<- allDotPlots_20[!(unlist(lapply(allDotPlots_20, function(x){is.null(x)})))]
cowplot::plot_grid(plotlist = allDotPlots_20, ncol=2)
ggsave(paste0("GO_OVA_AllClusters_top20.jpg"), path=seurat_fig_folder, width=18, height=40)
save(go_result, file=paste0(seurat_analysis_folder,  "go_result.Robj"))
```

## ORA with ependymoma genesets
## Compare with metaprograms from previously defined metaprograms from ependymoma
```{r}
all_markers_list<-split(all_markers_filtered, all_markers_filtered$cluster )
all_markers_list<-lapply(all_markers_list, function(x){x$gene})
names(all_markers_list)<-paste0("Cluster", names(all_markers_list))

epn_genesets<- read.csv("../TableS2_epn2020Paper_Metaprograms_AllMerged.csv")
epn_genesets<- as.list(epn_genesets)


tmp2 = sapply(all_markers_list, function(x){
    sapply(epn_genesets, function(y){
        fisher_test(x, y, unlist(union(epn_genesets, all_markers_list)))
    })
})
tmp2 = data.frame(tmp2)
tmp3<- apply(tmp2, 2, function(x) {p.adjust(x) < 0.1})
tmp4<-apply(tmp2, 2, function(x) {p.adjust(x)})

write.table(tmp4, paste0(seurat_analysis_folder,
                         "ScoreSeuratClustersUsingPreviousNMF_padj.txt"))

```



# Add NMF results to the seurat object

## Map NMF scores to seurat object 
```{r add-NMF-score}
NameOfMetaprogramList<-"Merge2"

nmf_score = readRDS(paste0("analysis/nmf/Merge/nmf_score_", NameOfMetaprogramList, ".rds"))
seurat_obj = AddMetaData(seurat_obj, nmf_score)
load(paste0("analysis/nmf/Merge/nmf_marker_genes_final_", NameOfMetaprogramList, ".Robj"))

DimPlot(object = seurat_obj, group.by = "signature_1", label = TRUE, 
         pt.size = 2, label.size = 6, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_nmf_signature_",NameOfMetaprogramList, ".jpg"), 
       path=seurat_fig_folder, 
       width=10, height=8)

#DoHeatmap(seurat_obj, 
          #features=unlist(nmf_marker_genes_final)) + NoLegend()

#ggsave("nmf_marker_genes.jpg", path=seurat_fig_folder, width=12, height=8)
```

## Map LIGER iNMF scores to seurat object 
```{r add-NMF-score}
nmf_score = readRDS(paste0(liger_analysis_folder, "nmf_score_AllFactors_SubtypeInt.Rds"))
seurat_obj = AddMetaData(seurat_obj, nmf_score)
#seurat_obj@meta.data$Factor15<- nmf_score

DimPlot(object = seurat_obj, group.by = "signature_1", label = FALSE, 
         pt.size = 2, label.size = 8, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_nmf_signature_LIGER", ".jpg"), 
       path=seurat_fig_folder, 
       width=14, height=12)

## Set idents to NMF factors, then find markers between desired factors
seurat_obj_liger<-seurat_obj
Idents(seurat_obj_liger)<- "signature_1"
f18_v_f8<- FindMarkers(seurat_obj_liger, 
                       ident.1="factor18",
                       ident.2="factor8")
f18_v_f8_filtered<- f18_v_f8[f18_v_f8$p_val_adj<0.05 &
                                abs(f18_v_f8$avg_logFC)>1,]
write.csv(f18_v_f8_filtered, paste0(seurat_analysis_folder, "LIGER_factor18_vs_factor8.csv"))

#DoHeatmap(seurat_obj, 
          #features=unlist(nmf_marker_genes_final)) + NoLegend()

#ggsave("nmf_marker_genes.jpg", path=seurat_fig_folder, width=12, height=8)
```


## Add a second metagene classification without cell cycle programs
```{r}
nmf_score = readRDS(paste0("analysis/nmf/all_subtypes/new/nmf_score_updated", NameOfMetaprogramList,  ".rds"))
seurat_obj@meta.data$signaure_1_without_cc = nmf_score$signature_1
```

## Make pseudobulk
```{r}
metagene_program_names = sort(unique(nmf_score$signature_1))
pseudobulk = NULL
aggr_method = "mean"
for (metagene in metagene_program_names){
  cm = seurat_obj@assays$RNA@counts[,nmf_score$signature_1 == metagene & nmf_score$score_1 >= 1]
  if (aggr_method == "mean"){
    pseudobulk = cbind(pseudobulk, rowMeans(cm))
  }
  if (aggr_method == "median"){
    pseudobulk = cbind(pseudobulk, rowMedians(cm))
  }
}
colnames(pseudobulk) = metagene_program_names
rownames(pseudobulk) = rownames(seurat_obj@assays$RNA@counts)
saveRDS(pseudobulk, file=paste0(seurat_analysis_folder, 
                                "pseudobulk_cm_unique_only_", aggr_method, NameOfMetaprogramList,".RDS"))
```

## Metagene program specific genes as overall DEGs
```{r}
seurat_obj = SetIdent(seurat_obj, value = seurat_obj@meta.data$signaure_1_without_cc)

all_markers_nmf = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_nmf = all_markers[all_markers$p_val_adj < 0.05, ]
saveRDS(all_markers, paste0(seurat_analysis_folder, "deg.rds"))
##sapply(names(nmf_marker_genes_final), function(x) table(nmf_marker_genes_final[[x]] %in% gene_signatures[[x]]))
signature_genes = unique(all_markers$gene)
saveRDS(signature_genes, paste0(seurat_analysis_folder, "signature_genes.rds"))
```

## Manual cluster annotation
```{r}
clusters<-as.character(unique(seurat_obj@active.ident))
cluster_annot<- c("ImmuneReactive", "Ependymal", "Metabolic", "SP-Variable_Metabolic",
                  "ST-RELA-Variable", "Cycling", "AC-like", "RadialGlia-like",
                  "ST-Midline-Ribosomal")
seurat_obj@meta.data$ManualClusterAnnot<- plyr::mapvalues(seurat_obj@active.ident,
                                                          from=clusters,
                                                          to=cluster_annot)
DimPlot(object = seurat_obj, group.by = "ManualClusterAnnot", label = TRUE, 
         pt.size = 2, label.size = 8, reduction="umap", cols = col_vector[10:28]) + NoAxes() + NoLegend()
ggsave(paste0("UMAP_ManualClusterAnnot.jpg"), 
       path=seurat_fig_folder, 
       width=8, height=8)

## Plot breakdown of each cluster
## By sample
cluster_by_sample<- table(seurat_obj@meta.data$sample, seurat_obj@meta.data$ManualClusterAnnot)
cluster_by_sample<- melt(cluster_by_sample)
colnames(cluster_by_sample)<- c("Sample", "Cluster", "NCells")
cluster_by_sample$Cluster<- paste0("Cluster", cluster_by_sample$Cluster)

ggplot(cluster_by_sample, aes(x=Cluster, y=NCells, fill=Sample))+ 
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Seurat clustering: \nbreakdown of samples per cluster")
ggsave(paste0(seurat_fig_folder, "SeuratClusterSampleDistr_Annot.png"), width=3, height=4)

## By subtype
cluster_by_subtype<- table(seurat_obj@meta.data$Subtype, seurat_obj@meta.data$ManualClusterAnnot)
cluster_by_subtype<- melt(cluster_by_subtype)
colnames(cluster_by_subtype)<- c("Subtype", "Cluster", "NCells")
cluster_by_subtype$Cluster<- paste0("Cluster", cluster_by_subtype$Cluster)

ggplot(cluster_by_subtype, aes(x=Cluster, y=NCells, fill=Subtype))+ 
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Seurat clustering: \nbreakdown of subtype per cluster")
ggsave(paste0(seurat_fig_folder, "SeuratClusterSubtypeDistr_Annot.png"), width=3, height=4)
```

## Split adult sample(s) into separate seurat object, remove from other samples
```{r}
## Read in seurat object with all samples
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj.Robj")
load(seurat_filename)

## Set adult sample(s)
adult_samples<-"MUV063"
ped_samples<- unique(seurat_obj$sample[!seurat_obj$sample %in% adult_samples])

## Create adult/ped seurat object
adult_seurat<- subset(seurat_obj, sample %in%adult_samples)
seurat_obj<- subset(seurat_obj, sample %in% ped_samples)

## Save- ped will be used for downstream, adult only for minimal comparison purposes
save(seurat_obj, file=paste0(seurat_analysis_folder, "seurat_obj.Robj"))
save(adult_seurat, file=paste0(seurat_analysis_folder, "seurat_obj_adultSamples.Robj"))
```

