---
title: "LIGER iNMF"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(liger)
library(ggplot2)

base_dir = "/Users/jlabelle/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/BySubtype/Ependymoma/")
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"

liger_analysis_folder = "analysis/liger/"
liger_fig_folder = "figures/liger/"
for (folder in c(liger_analysis_folder, liger_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}
nmf_analysis_folder = "/analysis/nmf/Merge/"

nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
stat_helper = paste0(base_dir, script_dir, "common_stat_test.R")
preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(nmf_helper)
source(stat_helper)
source(preprocessing)

```

## Read in data, then split by subtype
```{r}
## Read in cm list, sample list, and tumor/normal annotation
input_dir<- "../../Preprocessing/SplitBySubtype/analysis/"
load(paste0(input_dir, "epn_cm_list.Robj"))
samples = readRDS(paste0(input_dir,"epn_samples.rds"))
tumor_stat = readRDS(paste0(input_dir,"epn_call_tumor.rds"))
cm_list<-epn_cm_list

## Keep maligant cells
cm<- cm_list$raw_data
cm = cm_list$raw_data[, tumor_stat=="Malignant"]
samples_tumor = samples[tumor_stat=="Malignant"]
length(samples_tumor); sum(colnames(cm)==names(samples_tumor))

## Read in subtypes
subtypes<- read.csv("../Subtypes.csv")
subtypes<-subtypes[subtypes$shorthand=="epn",]

subtype_vector<- plyr::mapvalues(x = samples_tumor, from = subtypes$samples, to = subtypes$Diagnosis)


## Split cm into list of cm, 1 for each subtype
split_cm<- list()
for (s in unique(subtype_vector)){
  cells<- subtype_vector[subtype_vector %in% s]
  cm_sample<- cm[,colnames(cm) %in% names(cells)]
  split_cm[[s]]<- cm_sample
}
```

## Create liger object
```{r}
liger_obj <- createLiger(split_cm)
```

## Normalize, identify variable expressed genes, scale data
```{r}
liger_obj <- liger::normalize(liger_obj)
liger_obj <- selectGenes(liger_obj)
liger_obj <- scaleNotCenter(liger_obj)
```


## Optional: Use the suggestK function to determine the appropriate number of factors to use.
## Optional: Use the suggestLambda function to find the smallest lambda for which the alignment metric stabilizes.
## These steps take approximately ~1 hour
```{r}
suggestK(liger_obj) ## default 20
suggestLambda(liger_obj, k = 30, num.cores = 4) ## default 5
```

## Joint Matrix Facotorization
Important parameters are as follows:

1. k. Integer value specifying the inner dimension of factorization, or number of factors. Higher k is recommended for datasets with more substructure. We find that a value of k in the range 20 - 40 works well for most datasets. Because this is an unsupervised, exploratory analysis, there is no single “right” value for k, and in practice, users choose k from a combination of biological prior knowledge and other information.

2. lambda. This is a regularization parameter. Larger values penalize dataset-specific effects more strongly, causing the datasets to be better aligned, but possibly at the cost of higher reconstruction error. The default value is 5. We recommend using this value for most analyses, but find that it can be lowered to 1 in cases where the dataset differences are expected to be relatively small, such as scRNA-seq data from the same tissue but different individuals.

3. thresh. This sets the convergence threshold. Lower values cause the algorithm to run longer. The default is 1e-6.

4. max.iters. This variable sets the maximum number of iterations to perform. The default value is 30.

## This step takes approximately ~10 minutes
```{r}
liger_obj <- optimizeALS(liger_obj, k = 20, lambda = 5)

## cluster cells, perform quantile normalization by dataset, factor, and cluster
## All performed in quantile_norm function
liger_obj <- quantile_norm(liger_obj)

## Save liger object
save(liger_obj, file=paste0(liger_analysis_folder, "liger_obj_PreUMAP_SubtypeInt.Robj"))
```

## Run UMAP
## Can run Louvain first, or skip that step and retain all factors
```{r}
## Optional: Louvain clustering
liger_obj_louvain <- louvainCluster(liger_obj, resolution = 0.25)

## Run UMAP
liger_obj <- liger::runUMAP(liger_obj, distance = 'cosine', n_neighbors = 30, min_dist = 0.3)

save(liger_obj, file=paste0(liger_analysis_folder, "liger_obj_PostUMAP_NoLouvain.Robj"))
```


## Plot on UMAP- colored by factors + sample
```{r}
all.plots <- plotByDatasetAndCluster(liger_obj, axis.labels = c('UMAP 1', 'UMAP 2'),
                                     return.plots = T, pt.size = 1)
all.plots[[1]]
ggsave(paste0(liger_fig_folder, "UMAP_Sample_SubtypeInt.png"), width=7,height=7)


all.plots[[2]]
ggsave(paste0(liger_fig_folder, "UMAP_AllFactors_SubtypeInt.png"), width=7,height=7)

```

## Export cell assignments for use with seurat clustering
```{r}
liger_factorAssignments<- as.factor(as.character(liger_obj@clusters))

save(liger_factorAssignments, file=paste0(liger_analysis_folder, "ligerFactor_Assignment.Robj"))

```

## Plot factor loaded directly onto UMAP
## Limitation here: can only compare 1 dataset vs 1 other
```{r}
## Set 2 datasets to compare
dataset1="MUV079"
dataset2="MUV092"

## saves all gene loadings plots to pdf
pdf(paste0(liger_fig_folder, "gene_loadings_", dataset1, "vs", dataset2, ".pdf"))
gene_loadings <- plotGeneLoadings(liger_obj, 
                                  dataset1 = dataset1, 
                                  dataset2=dataset2,
                                  num.genes.show = 12, ## default 12
                                  do.spec.plot = TRUE, 
                                  return.plots = FALSE,
                                  do.title = TRUE,
                                  pval.thresh = 0.05, ## default 0.05, wilcoxon
                                  log.fc.thresh = 1) ## default 1
dev.off()
```

## Identify markers for each factor
## Only identifies markers for factors that are the top factor for at least 1 cell
```{r}
cluster.results <- runWilcoxon(liger_obj, compare.method = "clusters")
cluster.results_filtered<-cluster.results[cluster.results$padj<0.05 &
                                            cluster.results$logFC>2,]

save(cluster.results_filtered, file=paste0(liger_analysis_folder,
                                           "AllFactors_FilteredMarkers_SubtypeInt_logFC2.Robj"))
```

## Identify markers between each dataset, for each factor
## These are dataset-specific markers for each factor
## Only identifies markers for factors that are the top factor for at least 1 cell
```{r}
cluster.results_dataset <- runWilcoxon(liger_obj, compare.method = "datasets")
cluster.results_dataset_filtered<-cluster.results_dataset[cluster.results_dataset$padj<0.05 &
                                            cluster.results_dataset$logFC>3,]
```


## Visualize expression of top gene across samples
```{r}
geneToTest<- "ABHD3"

PRF1 <- plotGene(liger_obj, geneToTest, 
                 axis.labels = c('UMAP 1', 'UMAP 2'), 
                 return.plots = T,
                 plot.by="dataset", 
                 pt.size = 1)

cowplot::plot_grid(plotlist=PRF1)
ggsave(paste0(liger_fig_folder, "plotGene_", geneToTest, ".png"), width=10, height=10)
```

## There are 2 factors here that are shared between all 4 samples: factor 15 and factor 5
## Score all cells for these 2 factors, see which have reasonably high score
```{r}
sharedFactors<- c("15", "5")
sharedFactors_markers<-cluster.results_filtered[cluster.results_filtered$group %in% sharedFactors,]

## Split by factor, then Get top 50 markers, based on p value
sharedFactors_markers_list<- split(sharedFactors_markers, f= sharedFactors_markers$group)
sharedFactors_markers_list<- lapply(sharedFactors_markers_list, function(x){
  tmp<- x[order(x$padj),]
  tmp[1:50,]
})
names(sharedFactors_markers_list)<- paste0("factor", names(sharedFactors_markers_list))

## Get list of top genes for each factor
gene_list<- lapply(sharedFactors_markers_list, function(x){na.omit(x$feature)})

## center counts and get mean gene expression
cm_center<-cm_list$center_data
cm_center<- cm_center[,colnames(cm_center) %in% colnames(cm)]
cm_mean<- log2(rowMeans(cm)+1)

## Score each cell for factor gene set
nmf_score = scoreNmfGenes(cm_center, cm_mean, gene_list)
nmf_score = t(nmf_score) 

num_meta_program = length(gene_list)
nmf_score_final_t = data.frame(nmf_score)

ggplot(nmf_score_final_t, aes(x=factor15))+ geom_histogram()

## assign cells with factor 5 score >3 as that cluster (ependymal)
factor5Cells<- nmf_score_final_t$factor5>=2
names(factor5Cells)<- rownames(nmf_score_final_t)
factor5Cells<- gsub("TRUE", "Factor5", factor5Cells)
saveRDS(factor5Cells, file=paste0(liger_analysis_folder, "Factor5Cells_Liger_SubtypeInt_ScoreAbove3.Rds"))

## assign cells with factor 15 score >0.5 as that cluster (ependymal)
factor15Cells<- nmf_score_final_t$factor15>=0.5
names(factor15Cells)<- rownames(nmf_score_final_t)
factor15Cells<- gsub("TRUE", "Factor15", factor15Cells)
saveRDS(factor15Cells, file=paste0(liger_analysis_folder, "Factor15Cells_Liger_SubtypeInt_ScoreAbove3.Rds"))
```


## Alternative approach for annotating cells with factor
## Score each cell using top marker genes for each factor- using scoreNmfGenes function
```{r}

factor_markers<-cluster.results_filtered

## Split by factor, then Get top 50 markers, based on p value
factor_markers_list<- split(factor_markers, f= factor_markers$group)
factor_markers_list<- lapply(factor_markers_list, function(x){
  tmp<- x[order(x$padj),]
  tmp[1:50,]
})
names(factor_markers_list)<- paste0("factor", names(factor_markers_list))

## Get list of top genes for each factor
gene_list<- lapply(factor_markers_list, function(x){na.omit(x$feature)})

## center counts and get mean gene expression
cm_center<-cm_list$center_data
cm_center<- cm_center[,colnames(cm_center) %in% colnames(cm)]
cm_mean<- log2(rowMeans(cm)+1)

## Score each cell for factor gene set
nmf_score = scoreNmfGenes(cm_center, cm_mean, gene_list)
nmf_score = t(nmf_score) 

num_meta_program = length(gene_list)
nmf_score_final_t = data.frame(nmf_score)

## Get top 3 factors for each cell
for (i in 1:3){
  nmf_score_final_t = metagene_score_signature(nmf_score_final_t, num_meta_program, i)
}

saveRDS(nmf_score_final_t, 
        file=paste0(liger_analysis_folder, "nmf_score_AllFactors_SubtypeInt", ".Rds"))
```


## ORA with ependymoma genesets
## Compare with factors from previously defined metaprograms from hemispheric samples
```{r}
all_markers_list<-split(cluster.results_filtered, cluster.results_filtered$group )
all_markers_list<-lapply(all_markers_list, function(x){x$feature})
names(all_markers_list)<-paste0("Factor", names(all_markers_list))

epn_genesets<- read.csv("../TableS2_epn2020Paper_Metaprograms_AllMerged.csv")
epn_genesets<- as.list(epn_genesets)

tmp2 = sapply(gene_list, function(x){
    sapply(epn_genesets, function(y){
        fisher_test(x, y, unlist(union(epn_genesets, gene_list)))
    })
})
tmp2 = data.frame(tmp2)
tmp3<- apply(tmp2, 2, function(x) {p.adjust(x) < 0.1})
tmp4<-apply(tmp2, 2, function(x) {p.adjust(x)})

write.table(tmp4, paste0(liger_analysis_folder,
                        "ScoreFactors_UsingPreviousNMF_padj.txt"))

```



## Compare sample intergration using LIGER iNMF vs previous NMF
```{r}
## Make df of liger annotations, removing factors with no cells
liger_table_annots<-table(liger_obj@cell.data$dataset, liger_obj@clusters)
liger_table_annots<- liger_table_annots[,colSums(liger_table_annots)>0]

liger_table_annots<- melt(liger_table_annots)
colnames(liger_table_annots)<- c("Sample", "Factor", "NCells")
liger_table_annots$Factor<- as.character(paste0("Factor", liger_table_annots$Factor))

## Plot distribution of samples across factors
liger_distr<-ggplot(liger_table_annots, aes(x=Factor, y=NCells, fill=Sample))+ 
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("LIGER NMF: \nbreakdown of samples per factor")


## Make df of previous NMF annotations

## Read in previous scores
NameOfMetaprogramList<-"Merge2"
nmf_score = readRDS(paste0("analysis/nmf/Merge/nmf_score_", NameOfMetaprogramList, ".rds"))

## Add column denoting sample
nmf_score$sample<- gsub("\\..*", "", rownames(nmf_score) )
nmf_table_annots<-table(nmf_score$sample, nmf_score$signature_1)
nmf_table_annots<- melt(nmf_table_annots)
colnames(nmf_table_annots)<- c("Sample", "Program", "NCells")

## Plot distribution of samples across programs
nmf_distr<- ggplot(nmf_table_annots, aes(x=Program, y=NCells, fill=Sample))+ 
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Previous NMF: \nbreakdown of samples per program")

liger_distr + nmf_distr
ggsave(paste0(liger_fig_folder, "LigerSampleDistr_vs_NMFSampleDistr.png"))
```

```{r}
sessionInfo()
```



