---
title: "R Notebook"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/Immune/", 
                     "TumorCellAnnotation/BySubtype/Ependymoma/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
stat_helper = paste0(base_dir, script_dir, "common_stat_test.R")
source(preprocessing)
source(stat_helper)

nmf_analysis_folder = "analysis/nmf/Merge/"
seurat_analysis_folder = "analysis/seurat/"
liger_analysis_folder<- paste0(working_dir, "analysis/liger/")

figure_dir<- "figures/Projection_AllEpn_ToImmuneEpn"
analysis_dir<- "analysis/Projection_AllEpn_ToImmuneEpn"



## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
library(harmony)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

subtypes<- read.csv("../../../Subtypes.csv")
```

## read in ependymoma seurat object- just immune samples
## remove any samples that were run in the 2020 ependymoma paper- will just use their annotation
## Subset to each subtype, rerun seurat pipeline
```{r}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}
immune_seurat_obj<- seurat_obj
immune_samples<-unique(immune_seurat_obj$sample)

## Load in previously published datasets
published_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/PublishedDatasets/FilbinPublishedDatasets/SeuratV3/")
load(paste0(published_dir, "seurat_obj_EPN_PF.Robj"))
PD_seurat_pf<- seurat_objV3; rm(seurat_objV3)

load(paste0(published_dir, "seurat_obj_EPN_SP.Robj"))
PD_seurat_sp<- seurat_objV3; rm(seurat_objV3)

load(paste0(published_dir, "seurat_obj_EPN_ST.Robj"))
PD_seurat_st<- seurat_objV3

## subset to cells that have annotations in one of the above objects
PD_cells<- unique(c(colnames(PD_seurat_pf),
                    colnames(PD_seurat_sp),
                    colnames(PD_seurat_st)))
PD_immune_seurat<- subset(immune_seurat_obj, cells=PD_cells) 
dim(immune_seurat_obj); dim(PD_immune_seurat)

## re-run seurat pipeline on just these cells
cm<- PD_immune_seurat@assays$RNA@counts
meta<-PD_immune_seurat@meta.data
PD_immune_seurat_new<-RunFullSeurat(cm=cm, RunHarmony=TRUE, samples=meta$sample)


## For samples in published dataset: extract best annotation for each cell
## Merge metadata across all 3 subtypes, then subset to just cells in immune
meta_pf<- PD_seurat_pf@meta.data[,c("subtype", "score_1", "signature_1")]
meta_sp<- PD_seurat_sp@meta.data[,c("subtype",  "score_1", "signature_1")]
meta_st<- PD_seurat_st@meta.data[,c("subtype", "score_1", "signature_1")]

meta_PD<- rbind(meta_pf, meta_sp); meta_PD<-rbind(meta_PD, meta_st)
meta_PD<-meta_PD[rownames(meta_PD) %in% colnames(PD_immune_seurat),]
meta_PD<- meta_PD[order(colnames(PD_immune_seurat_new)),]

sum(rownames(meta_PD) == colnames (PD_immune_seurat_new))

## add PD annotation to immune PD object
annot_PD<- meta_PD$signature_1; names(annot_PD)<- rownames(meta_PD)
annot_PD<- annot_PD[order(match(names(annot_PD), colnames(PD_immune_seurat_new)))]
sum(names(annot_PD)==colnames(PD_immune_seurat_new))

PD_immune_seurat_new@meta.data$annot_PD<- annot_PD

DimPlot(PD_immune_seurat_new, group.by = "annot_PD")
ggsave(paste0(figure_dir, "AnnotatedByPrevious/PF_SP_Samples/PF_SP_AnnotsFromPublishedEpendymoma.png"), width=8,height=8)

DimPlot(PD_immune_seurat_new, group.by = "sample")
ggsave(paste0(figure_dir, "AnnotatedByPrevious/PF_SP_Samples/PF_SP_sample.png"), width=8,height=8)
```

# Subset to cells that are NOT in published dataset
## split by subtype, rerun seurat, then project
```{r}
## subset to samples that are not in PD
PD_samples<- unique(c(PD_seurat_pf$sample, PD_seurat_sp$sample, PD_seurat_st$sample))
noPD_samples<- immune_samples[!immune_samples %in% PD_samples] 
noPD_immune_seurat<-subset(immune_seurat_obj, sample %in% noPD_samples)

## Re-run seurat pipeline for these 2 samples- separately
bt1733<- subset(noPD_immune_seurat, sample=="BT1733")
bt1717<- subset(noPD_immune_seurat, sample=="BT1717")

bt1733_seurat<- RunFullSeurat(cm=bt1733@assays$RNA@counts, RunHarmony=FALSE, samples=bt1733@meta.data$sample)
bt1717_seurat<- RunFullSeurat(cm=bt1717@assays$RNA@counts, RunHarmony=FALSE, samples=bt1717@meta.data$sample)


## Projection for ST sample (BT1717)
meta<-PD_seurat_st@meta.data
PD_seurat_st<- RunFullSeurat(cm=PD_seurat_st@assays$RNA@counts, RunHarmony = FALSE, samples=PD_seurat_st$sample)
PD_seurat_st$signature_1<- meta$signature_1
seurat_list<- list(immune_st=bt1717_seurat, PD_st=PD_seurat_st)

transferAnchors<-FindTransferAnchors(reference = seurat_list$PD_st, query=seurat_list$immune_st)

predictions <- TransferData(anchorset = transferAnchors, 
                            refdata = PD_seurat_st$signature_1)

  ## add predictions to immune seurat object
  prediction_seurat<- predictions$predicted.id; names(prediction_seurat)<- rownames(predictions)
  sum(names(prediction_seurat)==colnames(bt1717_seurat)); dim(bt1717_seurat)
  
  bt1717_seurat$ProjectedPredictions<- prediction_seurat
  DimPlot(bt1717_seurat, group.by="ProjectedPredictions", label=TRUE)
  ggsave(paste0(figure_dir, "UPDATED_STRela_ProjectSTepn_to_BT1717.png"))

## Projection for PF sample (BT1733)
meta<-PD_seurat_pf@meta.data
PD_seurat_pf<- RunFullSeurat(cm=PD_seurat_pf@assays$RNA@counts, RunHarmony = FALSE, samples=PD_seurat_pf$sample)
PD_seurat_pf$signature_1<- meta$signature_1
seurat_list<- list(immune_st=bt1733_seurat, PD_pf=PD_seurat_pf)

transferAnchors<-FindTransferAnchors(reference = seurat_list$PD_pf, query=seurat_list$immune_st)

predictions <- TransferData(anchorset = transferAnchors, 
                            refdata = PD_seurat_pf$signature_1)

  ## add predictions to immune seurat object
  prediction_seurat<- predictions$predicted.id; names(prediction_seurat)<- rownames(predictions)
  sum(names(prediction_seurat)==colnames(bt1733_seurat)); dim(bt1733_seurat)
  
  bt1733_seurat$ProjectedPredictions<- prediction_seurat
  DimPlot(bt1733_seurat, group.by="ProjectedPredictions", label=TRUE)
  ggsave(paste0(figure_dir, "UPDATED_STRela_ProjectPFepn_to_BT1733.png"))


```

## NO LONGER IN USE
## Subset to cells that are NOT in published dataset
## split by subtype, rerun seurat, then project
## old analysis- both samples considered ST-RELA, but BT1873 is actually PF
## BT1873 should be projected from PF
```{r}
## subset to samples that are not in PD
PD_samples<- unique(c(PD_seurat_pf$sample, PD_seurat_sp$sample, PD_seurat_st$sample))
noPD_samples<- immune_samples[!immune_samples %in% PD_samples] 
noPD_immune_seurat<-subset(immune_seurat_obj, sample %in% noPD_samples)

## Re-run seurat pipeline for these 2 samples
cm<-noPD_immune_seurat@assays$RNA@counts
meta<-noPD_immune_seurat@meta.data
noPD_immune_seurat_new<- RunFullSeurat(cm=cm, RunHarmony=TRUE, samples=meta$sample)

## Run PCA/UMAP/Clustering for the previous seurat object
cm<- PD_seurat_st@assays$RNA@counts
meta<-PD_seurat_st@meta.data
PD_seurat_st_new<-RunFullSeurat(cm=cm, RunHarmony=TRUE, samples=meta$sample)
PD_seurat_st_new@meta.data<- meta
DimPlot(PD_seurat_st_new, group.by="signature_1")

## Project previous ST seurat object onto these 2 samples
seurat_list<- list(immune_st=noPD_immune_seurat_new, PD_st=PD_seurat_st_new)

transferAnchors<-FindTransferAnchors(reference = seurat_list$PD_st, query=seurat_list$immune_st)

predictions <- TransferData(anchorset = transferAnchors, 
                            refdata = PD_seurat_st_new$signature_1)

## add predictions to immune seurat object
prediction_seurat<- predictions$predicted.id; names(prediction_seurat)<- rownames(predictions)
sum(names(prediction_seurat)==colnames(noPD_immune_seurat_new)); dim(noPD_immune_seurat_new)

noPD_immune_seurat_new$ProjectedPredictions<- prediction_seurat
DimPlot(noPD_immune_seurat_new, group.by="ProjectedPredictions", label=TRUE)
ggsave(paste0(figure_dir, "AnnotatedByPrevious/ST_RELA_Samples/STRela_ProjectedAnnots_FromPublishedEpendymoma.png"))

DimPlot(noPD_immune_seurat_new, group.by="sample")
ggsave(paste0(figure_dir, "AnnotatedByPrevious/ST_RELA_Samples/STRela_Sample_FromPublishedEpendymoma.png"))
```

```{r}
save(noPD_immune_seurat_new, file=paste0(analysis_dir, "/STRela_ProjectedAnnotsFromPublished_seurat.Robj"))
save(PD_immune_seurat_new, file=paste0(analysis_dir, "/PF_SP_AnnotsFromPublished_seurat.Robj"))
```

## Merge 3 seurat objects- direct label transfer for most samples, BT1733 projections, BT1717 projections
## Will use same annotations obtained separately, merged into 1 seurat object
```{r}
## Load in seurat object containing all samples
load("analysis/seurat/seurat_obj.Robj")

## Add updated cell annot to object
  ## Merge metadata for updated cell annots
  BT1733_meta<- dplyr::rename(bt1733_seurat@meta.data,
                            CellAnnot=ProjectedPredictions)
  
  BT1717_meta<- dplyr::rename(bt1717_seurat@meta.data,
                          CellAnnot=ProjectedPredictions)
  
  PD_meta<- dplyr::rename(PD_immune_seurat_new@meta.data,
                          CellAnnot=annot_PD)
  
  meta_merged<- rbind(PD_meta, BT1717_meta)
  meta_merged<- rbind(meta_merged, BT1733_meta)
  meta_merged$Cell<- rownames(meta_merged)
  
  ## Remove cells from full seurat not in updated meta
  ## Only 18 cells
  seurat_obj<- subset(seurat_obj, 
                      cells=colnames(seurat_obj)[colnames(seurat_obj) %in%
                                                   meta_merged$Cell])
  sum(!colnames(seurat_obj) %in% meta_merged$Cell)
  
  ## add updated cell annot to full seurat_obj
  seurat_obj$Cell<- colnames(seurat_obj)
  seurat_obj$UpdatedAnnot<- plyr::mapvalues(seurat_obj$Cell,
                                            from=meta_merged$Cell,
                                            to=meta_merged$CellAnnot)
save(seurat_obj, file=paste0(analysis_dir, "UPDATED_seurat_obj_ProjectedAnnots.Robj"))

```



## NOT IN USE- OUT OF DATE
## Merge 2 seurat objects
## Currently running cellchat separately for 2 objects/annotations
## But want to visualize ligand/receptor expression simultaneously across all epn samples
## Will use same annotations obtained separately, merged into 1 seurat object
```{r}
## Load in seurat object containing all samples
load("analysis/seurat/seurat_obj.Robj")

## Add updated cell annot to object
  ## Merge metadata for updated cell annots
  noPD_meta<- dplyr::rename(noPD_immune_seurat_new@meta.data,
                            CellAnnot=ProjectedPredictions)
  
  PD_meta<- dplyr::rename(PD_immune_seurat_new@meta.data,
                          CellAnnot=annot_PD)
  meta_merged<- rbind(noPD_meta, PD_meta)
  meta_merged$Cell<- rownames(meta_merged)
  
  ## Remove cells from full seurat not in updated meta
  ## Only 18 cells
  seurat_obj<- subset(seurat_obj, 
                      cells=colnames(seurat_obj)[colnames(seurat_obj) %in%
                                                   meta_merged$Cell])
  sum(!colnames(seurat_obj) %in% meta_merged$Cell)
  
  ## add updated cell annot to full seurat_obj
  seurat_obj$Cell<- colnames(seurat_obj)
  seurat_obj$UpdatedAnnot<- plyr::mapvalues(seurat_obj$Cell,
                                            from=meta_merged$Cell,
                                            to=meta_merged$CellAnnot)
save(seurat_obj, file=paste0(analysis_dir, "seurat_obj_ProjectedAnnots.Robj"))

```


## Get marker genes for all previous ependymoma programs
```{r}
## Load in previous seurat objects
load("../../../PublishedDatasets/SeuratV3Objects/seurat_obj_EPN_PF.Robj")
PD_seurat_pf<- seurat_objV3; rm(seurat_objV3)

load("../../../PublishedDatasets/SeuratV3Objects/seurat_obj_EPN_SP.Robj")
PD_seurat_sp<- seurat_objV3; rm(seurat_objV3)

load("../../../PublishedDatasets/SeuratV3Objects/seurat_obj_EPN_ST.Robj")
PD_seurat_st<- seurat_objV3

## Set idents to annotations
PD_seurat_pf<- SetIdent(PD_seurat_pf, value=PD_seurat_pf$signature_1)
PD_seurat_sp<- SetIdent(PD_seurat_sp, value=PD_seurat_sp$signature_1)
PD_seurat_st<- SetIdent(PD_seurat_st, value=PD_seurat_st$signature_1)

## Find markers
pf_markers<-FindAllMarkers(PD_seurat_pf, min.pct = .5, logfc.threshold = .2)
pf_Markers_filt<-pf_markers[pf_markers$p_val_adj<0.05 & pf_markers$avg_logFC >0.2 ,]
save(pf_Markers_filt, file=paste0(analysis_dir, "epn_pf_Markers.Robj"))

sp_markers<-FindAllMarkers(PD_seurat_sp, min.pct = .5, logfc.threshold = .2)
sp_Markers_filt<-sp_markers[sp_markers$p_val_adj<0.05 & sp_markers$avg_logFC >0.2,]
save(sp_Markers_filt, file=paste0(analysis_dir, "epn_sp_Markers.Robj"))

st_markers<-FindAllMarkers(PD_seurat_st, min.pct = .5, logfc.threshold = .2)
st_Markers_filt<-st_markers[st_markers$p_val_adj<0.05 & st_markers$avg_logFC >0.2,]
save(st_Markers_filt, file=paste0(analysis_dir, "epn_st_Markers.Robj"))
```



## Updated annotation: same annotations used for CellChat
```{r}
load(paste0(analysis_dir, "UPDATED_seurat_obj_ProjectedAnnots.Robj"))

## Change subtype of CPDM0785 from PF-A to ST-Midline
seurat_obj$Subtype<- ifelse(seurat_obj$sample=="CPDM0785", "ST-Midline", seurat_obj$Subtype)

## Change subtype of BT1733 from ST-RELA to PF-A
seurat_obj$Subtype<- ifelse(seurat_obj$sample=="BT1733", "PF-A", seurat_obj$Subtype)

## make cell annotation column for each- name needs to match
cellAnnots<- c("PF_Ependymal_like", "SP_Ependymal_like", "ST_Ependymal_like",
               "PF_Astroependymal_like", "PF_Metabolic", "ST_Metabolic",
               "PF_Immune_Reactive", "SP_Immune_Reactive",
               "PF_Astroglial_like","ST_Metabolic", "ST_Ependymal_like","ST_Interferon_Response",
               "PF_G2M_Phase","ST_G2M_Phase", "PF_S_Phase", "ST_S_Phase")
names(cellAnnots)<- c(rep("Ependymal_like", 4), rep("Metabolic", 2), rep("Immune_Reactive",2), "AC-like",
                      "Metabolic", "Ependymal_like", "Immune_Reactive", rep("G2M",2),rep("SPhase",2))

seurat_obj$CellChatAnnot<- plyr::mapvalues(seurat_obj$UpdatedAnnot,
                                             cellAnnots, names(cellAnnots))


```

## Subset to pf/st/sp and rerun pipeline
```{r}
subtypes<- list(PF=c("PF-A", "PF-SE"),
                SP=c("SP-MPE"),
                ST=c("ST-RELA", "ST-Midline"))

epn_seurat_split<- lapply(subtypes, function(i){
  print(i)
  seurat_tmp<- subset(seurat_obj, Subtype %in% unlist(i))
  meta_tmp<- seurat_tmp@meta.data
  
  seurat_tmp<- preprocessSeuratObject(seurat_tmp@assays$RNA@counts, project="")
  seurat_tmp <- RunPCA(seurat_tmp, features = seurat_tmp@assays$RNA@var.features, npcs = 100)
  seurat_tmp$sample<- meta_tmp$sample
  seurat_tmp$CellAnnot<- meta_tmp$CellChatAnnot
  
  seurat_tmp = RunHarmony(seurat_tmp, "sample", theta = 2, 
                        max.iter.harmony = 50, plot_convergence = TRUE)
  
  seurat_tmp <- RunUMAP(seurat_tmp, reduction = "harmony", dims = 1:20)


  seurat_tmp <- FindNeighbors(seurat_tmp,
                              reduction = "harmony",
                              dims = 1:20,
                              force.recalc = TRUE)   %>% 
                FindClusters(resolution = 0.8)
  
  return(seurat_tmp)
})
names(epn_seurat_split)<- names(subtypes)
save(epn_seurat_split, file=paste0(analysis_dir, "UPDATED_ProjectedAnnot_AllEPN_SplitBySubtype_seurat.Robj"))

```

## UMAP of each epn subtype, same annots as for CellChat
## Breakdown of number of cells per sample
```{r}
## Set colors so they're shared across all subtypes
load("../../cellTypeColors.Robj")
cellTypes<- unique(seurat_obj$CellChatAnnot)
newCellTypes<- cellTypes[c(1:2,4:5,9,7, 11:14,8,3)]
newColors<- c(brewer.pal(8, "Dark2"), "grey", "darksalmon", "indianred", "gold2")
names(newColors)<- newCellTypes
colorsUse<- c(colorsUse, newColors)
save(colorsUse, file="../../../cellTypeColors.Robj")


## UMAP
load(paste0(analysis_dir, "UPDATED_ProjectedAnnot_AllEPN_SplitBySubtype_seurat.Robj"))
all_plots<- list()
for(subtype in names(epn_seurat_split)){
  epn_seurat<- epn_seurat_split[[subtype]]
  
  ## Remove any cell types with less than 10 cells
  cellTypeKeep<- table(epn_seurat$CellAnnot); cellTypeKeep<- names(cellTypeKeep)[cellTypeKeep>10]
  epn_seurat<- subset(epn_seurat, CellAnnot %in% cellTypeKeep)
  p<- DimPlot(object = epn_seurat, reduction="umap", group.by = "CellAnnot", 
        label = TRUE, pt.size = 2, label.size = 6) + NoAxes() +
    ggtitle(subtype)+
    scale_color_manual(values=colorsUse)
  all_plots[[subtype]]<- p
}
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave("UPDATED_UMAP_clustering_CellChatAnnot_EPNsubtypesSplit_Label.jpg",
         path=figure_dir, width=25, height=6)


## Breakdown of cell annots
for(subtype in names(epn_seurat_split)){
  epn_seurat<- epn_seurat_split[[subtype]]
  
  annotBreakdown<- as.data.frame(table(epn_seurat$sample, epn_seurat$CellAnnot))
  colnames(annotBreakdown)<- c("Sample", "CellType", "NCells")
  
  plot_list<-list()
  for(i in unique(annotBreakdown$Sample)){
    df<- annotBreakdown[annotBreakdown$Sample==i,]
    p<-ggplot(df, aes(x="", y=NCells, fill=CellType))+
      geom_bar(stat="identity", color="white")+
      coord_polar("y", start=0)+
      theme_void()+
      ggtitle(i)+theme(plot.title = element_text(size=15, face="bold", hjust=0.5))+
      scale_fill_manual(values=colorsUse)
    plot_list[[i]]<- p
  }
  cowplot::plot_grid(plotlist=plot_list, ncol=length(plot_list))
  ggsave(paste0(fig_dir, "UPDATED_CellBreakdown_", subtype,".png"), 
         width=length(plot_list)*4, height=4)
}

```

