---
title: "Compare.Adult.Ped"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Compare ped/adult de novo Tcell programs- shared? specific?

## Approach:
Correlate scores for ped/adult programs with one another + ped cd4/cd8 to help annotate. Various visualizations (venn diagram, heatmap, umap)

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/01_compare_NMF/")

analysis_dir<- paste0(working_dir, "/analysis/correlate_deNovoPed.AdultTcell/")
figure_dir<- paste0(working_dir, "/figures/correlate_deNovoPed.AdultTcell/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```




## Read in merged ped/adult tcell seurat object. Merged in Merge_AssessIntegration.Rmd, processed + annotated in AnnotateCD4.CD8_Ped.Adult
```{r}
cohort<- "pedOnly_nomuv63"
adult_withMUV<- "FALSE"

## ped/adult tcell seurat
tcell_seurat<- readRDS(paste0("../../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/Tcells_Ped.Adult_Merge/",
                              "adult_ped_tcell_merged_seurat_harmony.Rds"))

## Read in ped cd4/cd8 derived programs
cd8_markers<- read.csv(paste0("../../01a_Ped/01_NMF/ARCHIVE/CD8/analysis/", cohort, "/top_DE_genes_annotated_NMF_rank4.csv"))
cd4_markers<- read.csv(paste0( "../../01a_Ped/01_NMF/ARCHIVE/CD4/analysis/", cohort, "/top_DE_genes_annotated_NMF_rank4.csv"))

cd8_markers<- split(cd8_markers, f=cd8_markers$cluster); cd8_markers<- lapply(cd8_markers, function(x){x$gene})
cd4_markers<- split(cd4_markers, f=cd4_markers$cluster); cd4_markers<- lapply(cd4_markers, function(x){x$gene})

names(cd8_markers)<- paste0("CD8_", names(cd8_markers))
names(cd4_markers)<- paste0("CD4_", names(cd4_markers))
cd4.cd8_markers<- c(cd8_markers, cd4_markers)

```


## Score for ped cd4/cd8 programs + ped/adult tcell programs
```{r}
## Norm/center current cm
cm_list<- NormCenter(tcell_seurat@assays$RNA@counts)
cm_mean<- log2(rowMeans(cm_list$raw_data)+1)

## Score for cd4/cd8 ped
cd4.cd8_scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, c(cd8_markers, cd4_markers))))


## Read in ped and adult tcell derived programs for all ranks. Score all
ranks<- 7:9

  ## Ped markers
  all_ped_tcell_markers<- lapply(ranks, function(x){
    ped_tcell_markers<- read.csv(paste0( "../../01a_Ped/NMF/analysis/deNovoNMF/", cohort, "/top_DE_genes_NMF_rank", x,".csv"))
    ped_tcell_markers<- split(ped_tcell_markers, f=ped_tcell_markers$cluster) 
    ped_tcell_markers<- lapply(ped_tcell_markers, function(x){x$gene})
    names(ped_tcell_markers)<-paste0("PedTcell_", names(ped_tcell_markers))
    return(ped_tcell_markers)
  })
  names(all_ped_tcell_markers)<-paste0("rank", ranks)


  ## Read in adult tcell derived programs
  all_adult_tcell_markers<-lapply(ranks, function(x){
    adult_tcell_markers<- read.csv(paste0( "../../01b_Adult/NMF/analysis/deNovoNMF/", "/top_DE_genes_NMF_rank",x, ".csv"))
    adult_tcell_markers<- split(adult_tcell_markers, f=adult_tcell_markers$cluster) 
    adult_tcell_markers<- lapply(adult_tcell_markers, function(x){x$gene})
    names(adult_tcell_markers)<- paste0("AdultTcell_", names(adult_tcell_markers))
    return(adult_tcell_markers)
  })
  names(all_adult_tcell_markers)<- paste0("rank", ranks)
  
  
  ranks<- 7:9
  all_scores<- list()
  for(r in ranks){
    print(r)
    ped_markers<- all_ped_tcell_markers[[paste0("rank", r)]]
    adult_markers<- all_adult_tcell_markers[[paste0("rank", r)]]
    
    
    ## score for ped cd8/cd4 programs, adult tcell, ped tcell
    scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, c(ped_markers, adult_markers))))
    all_scores[[paste0("rank", r)]]<- scores
  }

saveRDS(all_scores, file=paste0(analysis_dir, "scores_ped.adult.tcell_ranks", ranks[1], "-",ranks[length(ranks)], ".Rds"))
saveRDS(cd4.cd8_scores, file=paste0(analysis_dir, "scores_mergedTcells_for_pedCD4.CD8.Rds"))
```



## Correlate with ped cd4/cd8 programs to help annotate tcell programs
```{r}
## Read in ped/adult tcell scores
ranks<- 7:9
all_scores<- readRDS(paste0(analysis_dir, "scores_ped.adult.tcell_ranks", ranks[1], "-",ranks[length(ranks)], ".Rds"))

## read in scores for ped cd4/cd8 programs
cd4.cd8_scores<- readRDS(paste0(analysis_dir, "scores_mergedTcells_for_pedCD4.CD8.Rds"))

## For each rank, correlate with cd4.cd8 scores, plot correlation heatmap
for(r in ranks){
  tcell_scores<- all_scores[[paste0("rank", r)]]
  scores<- merge(tcell_scores, cd4.cd8_scores, by=0)
  rownames(scores)<- scores$Row.names; scores$Row.names<-NULL
  
  
  ## Correlate and plot heatmap
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  pheatmap(scores_factor_cor, color = hm_colors(100), 
           cluster_rows = F, cluster_cols = F, 
           annotation_names_row = F, annotation_names_col =T,
           show_rownames = T, show_colnames = F,
           filename = paste0(figure_dir, "Cor_PedCD4.CD8rank4_PedTcellrank", r, "_AdultTcellrank", r,".png"),
           width = 8, height = 4)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir, "HC_PedCD4.CD8rank4_PedTcellrank", r, "_AdultTcellrank", r,".png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
  
  
}

```


## Correlation between just ped/adult tcell programs
```{r}
## Read in ped/adult tcell scores
ranks<- 7:9
all_scores<- readRDS(paste0(analysis_dir, "scores_ped.adult.tcell_ranks", ranks[1], "-",ranks[length(ranks)], ".Rds"))


## For each rank, correlate with cd4.cd8 scores, plot correlation heatmap
for(r in ranks){
  scores<- all_scores[[paste0("rank", r)]]
  
  
  ## Correlate and plot heatmap
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  pheatmap(scores_factor_cor, color = hm_colors(100), 
           cluster_rows = F, cluster_cols = F, 
           annotation_names_row = F, annotation_names_col =T,
           show_rownames = T, show_colnames = F,
           filename = paste0(figure_dir, "Cor_PedTcellrank", r, "_AdultTcellrank", r,".png"),
           width = 8, height = 4)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir, "HC_PedTcellrank", r, "_AdultTcellrank", r,".png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
  
  
}
  
```

## Annotate gene lists based on correlation (+ marker genes for Treg/Stress annotation)
```{r}
rank_use<-9
cohort<- "pedOnly_nomuv63"

## Read in tcell derived programs
ped_tcell_markers<- read.csv(paste0( "../../01a_Ped/01_NMF/analysis/deNovoNMF/", cohort, 
                                     "/top_DE_genes_NMF_rank", rank_use,".csv"))
adult_tcell_markers<- read.csv(paste0( "../../01b_Adult/01_NMF/analysis/deNovoNMF/", 
                                       "/top_DE_genes_NMF_rank",rank_use, ".csv"))

## Annotate programs based on correlation with ped cd4/cd8 programs (+ based on marker genes for Treg/Stress)
## Note: for ped rank8, only 7 programs actually have any marker genes
adult_annotate<- list(rank7=c("Memory", "Inhibitory", "Treg", "Cytotoxic.CD4", "EffectorMemory", "NK-like", "Stress"),
                      rank8=c("Inhibitory.CD8", "EffectorMemory", "Treg", "Stress", "NK-like", "Inhibitory.CD4", "Memory", "Cytotoxic.CD4"),
                      rank9=c("EffectorMemory", "Stress", "Memory", "Inhibitory.CD8", "Inhibitory.CD4", "Treg",
                              "NK-like", "Cytotoxic.CD4", "NMF9"))
adult_annotate<- adult_annotate[[paste0("rank", rank_use)]]
names(adult_annotate)<- unique(adult_tcell_markers$cluster)

ped_annotate<- list(rank7=c("Memory", "Cytotoxic.CD4", "NK-like", "Inhibitory.CD8", "Cytotoxic.CD8", "Inhibitory.CD4", "EffectorMemory"),
                    rank8=c("Cytotoxic.CD4", "Cytotoxic.CD8", "Memory", "EffectorMemory", 
                            "NK-like", "Inhibitory.CD8", "Inhibitory.CD4"),
                    rank9=c("Cytotoxic.CD4", "Cytotoxic.CD8", "Inhibitory.CD4","NK-like", "EffectorMemory2", "Memory",
                            "RiboActive", "Inhibitory.CD8", "EffectorMemory"))
ped_annotate<- ped_annotate[[paste0("rank", rank_use)]]
names(ped_annotate)<-unique(ped_tcell_markers$cluster)

## Add annotations to marker genes
adult_tcell_markers$cluster<- plyr::mapvalues(adult_tcell_markers$cluster, names(adult_annotate), adult_annotate)
ped_tcell_markers$cluster<- plyr::mapvalues(ped_tcell_markers$cluster, names(ped_annotate), ped_annotate)

## Save annotated marker lists
paste0( "../../01a_Ped/NMF/analysis/deNovoNMF/", cohort, 
                                     "/top_DE_genes_annotated_NMF_rank", x,".csv")
write.csv(ped_tcell_markers, 
          file=paste0( "../../01a_Ped/NMF/analysis/deNovoNMF/", cohort, 
                                     "/top_DE_genes_annotated_NMF_rank", x,".csv"))
write.csv(adult_tcell_markers, 
          file=paste0( "../../01b_Adult/NMF/analysis/deNovoNMF/", 
                                       "/top_DE_genes_annotated_NMF_rank",x, ".csv"))

## Save annotations
ped.adult_annotations<- list(adult=adult_annotate, ped=ped_annotate)
saveRDS(ped.adult_annotations, file=paste0(analysis_dir, "ped.adult.programs_annotations_rank",rank_use, ".Rds"))
```



## Compare gene lists between similar ped/adult programs
```{r}
library(ggVennDiagram)
rank_use<-9
option1<- FALSE

## Read in annotated gene lists (see above)
ped_tcell_markers<- read.csv(paste0( "../../01a_Ped/NMF/analysis/deNovoNMF/", cohort, 
                                     "/top_DE_genes_annotated_NMF_rank", x,".csv"))
adult_tcell_markers<- read.csv(paste0( "../../01b_Adult/NMF/analysis/deNovoNMF/", 
                                       "/top_DE_genes_annotated_NMF_rank",x, ".csv"))

## Split into list
ped_tcell_markers<- split(ped_tcell_markers, f=ped_tcell_markers$cluster) 
ped_tcell_markers<- lapply(ped_tcell_markers, function(x){x$gene})

adult_tcell_markers<- split(adult_tcell_markers, f=adult_tcell_markers$cluster) 
adult_tcell_markers<- lapply(adult_tcell_markers, function(x){x$gene})


## Venn diagram comparing overlap for genesets
  ## Optional: Show overlap between adult inhibitory/cytotoxic + ped cd4/cd8 inhibitory separately
  ## Only used for rank 7
  if(option1){
    adult_tcell_markers_tmp<- adult_tcell_markers
    adult_tcell_markers_tmp$Inhibitory.CD4<- adult_tcell_markers_tmp$Inhibitory
    adult_tcell_markers_tmp$Inhibitory.CD8<- adult_tcell_markers_tmp$Inhibitory
    adult_tcell_markers_tmp$Inhibitory<-NULL
    
    adult_tcell_markers_tmp$Cytotoxic.CD8<- adult_tcell_markers_tmp$Cytotoxic.CD4
    
    figure_append<- "_CytoCD8.InhCD4.added"
    
  }else{
    adult_tcell_markers_tmp<- adult_tcell_markers
    figure_append<- ""
  }
  
  ## Generate list of comparisons between programs named the same
  comparisons<- names(adult_tcell_markers_tmp)[names(adult_tcell_markers_tmp) %in% names(ped_tcell_markers)]
  comparisons<- comparisons[order(comparisons)]
  comparison_list<- lapply(comparisons, function(x){
    list(Adult=adult_tcell_markers_tmp[[x]], Ped=ped_tcell_markers[[x]])
  })
  names(comparison_list)<- comparisons
  
  all_plots<- lapply(names(comparison_list), function(x){
    tmp_list<- comparison_list[[x]]
    ggVennDiagram(tmp_list, label="count")+
      scale_fill_gradient(low="white", high="red")+
      scale_color_manual(values=rep("black", 3))+
      ggtitle(x)
  })
  cowplot::plot_grid(plotlist=all_plots, ncol=3)
  ggsave(paste0(figure_dir, "VennDiagram_ped.adult_geneOverlaps_rank", rank_use, ".png"), width=9, height=7)
```

## Add NMF annotations (directly from NMF) to merged tcell seurat object
NMF annotations from ped-only analysis / adult-only analysis read in, added to merged tcell seurat object in single column
```{r}
rank_use<-9
variable_name<- paste0("NMF_rank", rank_use, "_annot")
map_from_variable<- paste0("NMF_rank", rank_use)

## read in annotations for NMF programs
annotations<- readRDS(paste0(analysis_dir, "ped.adult.programs_annotations_rank", rank_use, ".Rds"))

## Read in seurat objects where annotations are stored
ped_tcell_seurat<- readRDS(paste0("../../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/", cohort, "/tcell_seurat.Rds"))
adult_tcell_seurat<- readRDS(paste0("../../01b_Adult/01_NMF/analysis/deNovoNMF/tcell_seurat.Rds"))

## Add on annotations to ped/adult seurat objects
ped_tcell_seurat@meta.data[[variable_name]]<- plyr::mapvalues(ped_tcell_seurat@meta.data[[map_from_variable]], 
                                                              names(annotations$ped), annotations$ped)
adult_tcell_seurat@meta.data[[variable_name]]<- plyr::mapvalues(adult_tcell_seurat@meta.data[[map_from_variable]], 
                                                                names(annotations$adult), annotations$adult)

## Add new column to merged ped/adult seurat object with these annotations
all_annots<- as.character(c(ped_tcell_seurat@meta.data[[variable_name]], adult_tcell_seurat@meta.data[[variable_name]]))
names(all_annots)<-c(colnames(ped_tcell_seurat), colnames(adult_tcell_seurat))
tcell_seurat@meta.data[[variable_name]]<- as.character(plyr::mapvalues(colnames(tcell_seurat), 
                                                                       names(all_annots), all_annots,
                                                                       warn_missing = FALSE))

## Save all
saveRDS(ped_tcell_seurat, 
        file=paste0("../../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/", cohort, "/tcell_seurat.Rds"))
saveRDS(adult_tcell_seurat, file=paste0("../../01b_Adult/01_NMF/analysis/deNovoNMF/tcell_seurat.Rds"))
saveRDS(tcell_seurat, file=paste0("../../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/Tcells_Ped.Adult_Merge/",
                              "adult_ped_tcell_merged_seurat_harmony.Rds"))
```


## Heatmap of top genes for ped/adult tcell NMF (annotated as above)
```{r}
rank_use<-8
pvalue_logFC<-"pvalue" ## Select top genes by pvalue or logFC
n_top<-5


## Read in annotated gene lists
ped_tcell_markers<- read.csv(paste0( "../../01a_Ped/NMF/analysis/deNovoNMF/", cohort, 
                                     "/top_DE_genes_annotated_NMF_rank", x,".csv"))
adult_tcell_markers<- read.csv(paste0( "../../01b_Adult/NMF/analysis/deNovoNMF/", 
                                       "/top_DE_genes_annotated_NMF_rank",x, ".csv"))

## merge into single df, adding "Ped" "Adult" suffix
ped_tcell_markers$cluster<- paste0(ped_tcell_markers$cluster, "_Ped")
adult_tcell_markers$cluster<- paste0(adult_tcell_markers$cluster,"_Adult")
tcell_markers<- rbind(ped_tcell_markers, adult_tcell_markers)
tcell_markers<- tcell_markers[order(tcell_markers$cluster),]


## Get top genes
if(pvalue_logFC=="pvalue"){
    top_genes<- tcell_markers %>% group_by(cluster) %>% top_n(n=n_top, wt=-p_val_adj) %>% as.data.frame()
}else if(pvalue_logFC=="logFC"){
  top_genes<- tcell_markers %>% group_by(cluster) %>% top_n(n=n_top, wt=avg_log2FC) %>% as.data.frame()
}

## plot on heatmap using ped/adult seurat object
backup<- ScaleData(tcell_seurat, do.scale = FALSE, do.center=TRUE, features=rownames(tcell_seurat))
DoHeatmap(backup, features=top_genes$gene, group.by = paste0("NMF_rank", rank_use, "_annot") )+
  scale_fill_gradient2(low="blue", mid="white", high="red")
ggsave(paste0(figure_dir, "Heatmap_AllTcells_top",n_top, "_NMFrank",rank_use, "genes_by",pvalue_logFC, ".png"), width = 12, height=12)

```

## UMAP of all tcells by NMF annot
```{r}
rank_use<-9
variable_name<- paste0("NMF_rank", rank_use, "_annot")

## With cyto/inhibitory split
colors_use<-c(`Cytotoxic.CD4`="orange", `Cytotoxic.CD8`="red",EffectorMemory="olivedrab", EffectorMemory2="palegreen",Inhibitory="blue", 
              `Inhibitory.CD4`="turquoise", `Inhibitory.CD8`="navy",Memory="purple", `NK-like`="green", 
              Stress="yellow", Treg="grey22", `8`="white", RiboActive="pink",NMF9="white")

DimPlot(tcell_seurat, group.by=variable_name)+
  scale_color_manual(values=colors_use)
ggsave(paste0(figure_dir, "UMAP_AllTcells_byNMFrank", rank_use, ".png"), width=7, height=5)

  ## Plot each individually
  all_plots<- lapply(unique(tcell_seurat@meta.data[[variable_name]]), function(x){
    tcell_seurat$tmp<- ifelse(tcell_seurat@meta.data[[variable_name]]==x, x, "Other")
    DimPlot(tcell_seurat, group.by="tmp")+
      scale_color_manual(values=c(colors_use, Other="grey"))+NoLegend()+
      ggtitle(x)
  })
  cowplot::plot_grid(plotlist=all_plots, ncol=4)
  ggsave(paste0(figure_dir, "UMAP_AllTcells_byNMFrank", rank_use, "_indv.png"), width=12, height=9)

## With cyto/inhibitory types together
colors_use<- c(Cytotoxic="red",EffectorMemory="olivedrab", `Inhibitory`="turquoise",
               Memory="purple", `NK-like`="green", Stress="yellow", Treg="grey22", `8`="white",
               RiboActive="pink",NMF9="white", EffectorMemory2="palegreen")
tcell_seurat@meta.data[[paste0(variable_name, "_merge")]]<- gsub(".CD4", "", gsub(".CD8", "", tcell_seurat@meta.data[[variable_name]]))
DimPlot(tcell_seurat, group.by=paste0(variable_name, "_merge"))+
  scale_color_manual(values=colors_use)
ggsave(paste0(figure_dir, "UMAP_AllTcells_byNMFrank",rank_use, ".mergeCytoInh.png"), width=7, height=5)

  ## Plot each individually
  all_plots<- lapply(unique(tcell_seurat@meta.data[[paste0(variable_name, "_merge")]]), function(x){
    tcell_seurat$tmp<- ifelse(tcell_seurat@meta.data[[paste0(variable_name, "_merge")]]==x, x, "Other")
    DimPlot(tcell_seurat, group.by="tmp")+
      scale_color_manual(values=c(colors_use, Other="grey"))+NoLegend()+
      ggtitle(x)
  })
  cowplot::plot_grid(plotlist=all_plots, ncol=4)
  ggsave(paste0(figure_dir, "UMAP_AllTcells_byNMFrank",rank_use, ".mergeCytoInh_indv.png"), width=12, height=9)
```

## Expression of expected marker genes based on NMF annotation
```{r}
rank_use<-8
variable_name<- paste0("NMF_rank", rank_use, "_annot")

DotPlot(tcell_seurat, group.by = variable_name, 
        features=c("GZMK", "CTLA4","LAG3", "TIGIT","PDCD1", "HAVCR2", "TOP2A", "CCR7", "GNLY", "HSPA1A", "FOXP3"))+
  xlab("") + ylab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave(paste0(figure_dir, "DotPlot_expectedMarkerGenes_forNMFrank7.png"), width=6, height=4)

## Split all into ped/adult
tcell_seurat$tmp<- paste0(tcell_seurat@meta.data[[variable_name]], "_", tcell_seurat$Age)
DotPlot(tcell_seurat, group.by = "tmp", 
        features=c("GZMK", "CTLA4","LAG3", "TIGIT","PDCD1", "HAVCR2", "TOP2A", "CCR7", "GNLY", "HSPA1A", "FOXP3"))+
  xlab("") + ylab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave(paste0(figure_dir, "DotPlot_expectedMarkerGenes_forNMFrank", rank_use,".byAge.png"), width=8, height=4)
```

## Exploratory- can delete
```{r}
adult_tregs<- subset(tcell_seurat, NMF_rank7_annot=="Treg")
test<- as.data.frame(t(adult_tregs@assays$RNA@data[c("CTLA4", "TIGIT","HAVCR2", "FOXP3"),]))
test<- test[rowSums(test)>0,]
hist(test$FOXP3)
test$lowfoxp3<- test$FOXP3<0.5 ## 430 tregs with no/low foxp3
test<- melt(test, id="lowfoxp3")
colnames(test)<-c("FOXP3_expr", "Gene", "logTPM")
ggplot(test, aes(x=Gene, y=logTPM, fill=FOXP3_expr))+
  geom_boxplot()


ped<- subset(tcell_seurat, Age=="Ped")
foxp3<- as.data.frame(t(ped@assays$RNA@data[c("FOXP3", "TIGIT"),]))
foxp3$nmf<- plyr::mapvalues(rownames(foxp3), colnames(ped), ped$NMF_rank7_annot)
foxp3$potential_tregs<- foxp3$FOXP3>2
table(foxp3$potential_tregs, foxp3$nmf)
```

