---
title: "GenerateNewMarkers"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Using shared/specific ped/adult tcell programs, merge gene lists then generate new annotations/genes. This is an alternative to using marker genes directly from NMF, and allows for the use of identical gene lists between ped/adult.

## Approach:
Generate union/intersection between shared ped/adult tcell programs. Score cells for these new programs, assign max, then use these assignments to generate new markers. Create similar plots as in deNovoPedTcell.v.deNovoAdultTcell.Rmd- UMAP, heatmap, etc

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/01_compare_NMF/")

analysis_dir<- paste0(working_dir, "/analysis/newMarkers_deNovoPed.AdultTcell/")
figure_dir<- paste0(working_dir, "/figures/newMarkers_deNovoPed.AdultTcell/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

# Read in merged ped/adult tcell seurat object. Merged in Merge_AssessIntegration.Rmd, processed + annotated in AnnotateCD4.CD8_Ped.Adult
```{r}
cohort<- "pedOnly_nomuv63"

## ped/adult tcell seurat
tcell_seurat<- readRDS(paste0("../../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/Tcells_Ped.Adult_Merge/",
                              "adult_ped_tcell_merged_seurat_harmony.Rds"))

cm_list<- NormCenter(tcell_seurat@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data + 1))
```

## Read in annotated genes (based on deNovoPedTcell.v.deNovoAdultTcell.Rmd)
```{r}
rank_use<-8
merge_cd4.cd8<- FALSE
programs_remove<-c("EffectorMemory2", "NMF9")

## Read in markers
ped_tcell_markers<- read.csv(paste0( "../../01a_Ped/01_NMF/analysis/deNovoNMF/", cohort, 
                                     "/top_DE_genes_annotated_NMF_rank", rank_use,".csv"))
adult_tcell_markers<- read.csv(paste0( "../../01b_Adult/01_NMF/analysis/deNovoNMF/", 
                                       "/top_DE_genes_annotated_NMF_rank",rank_use, ".csv"))

## Rename markers as needed to reflect desired merges- for now, all cyto/inhibitory are merged together
## This may change with higher rank
if(merge_cd4.cd8){
  adult_tcell_markers$cluster<- gsub(".CD4", "", gsub(".CD8", "", adult_tcell_markers$cluster))
  ped_tcell_markers$cluster<- gsub(".CD4", "", gsub(".CD8", "", ped_tcell_markers$cluster))
}


## Split into list
ped_tcell_markers<- split(ped_tcell_markers, f=ped_tcell_markers$cluster) 
ped_tcell_markers<- lapply(ped_tcell_markers, function(x){x$gene})

adult_tcell_markers<- split(adult_tcell_markers, f=adult_tcell_markers$cluster) 
adult_tcell_markers<- lapply(adult_tcell_markers, function(x){x$gene})

## Remove any programs
ped_tcell_markers<- ped_tcell_markers[!names(ped_tcell_markers) %in% programs_remove]
adult_tcell_markers<- adult_tcell_markers[!names(adult_tcell_markers) %in% programs_remove]

## Remove any duplicate genes (as a result of merging CD4/CD8 cyto/inhibitory)
ped_tcell_markers<- lapply(ped_tcell_markers, function(x){x[!duplicated(x)]})

```

## Get union and intersection of "shared" programs
```{r}
## Programs to merge
merge_programs<- names(adult_tcell_markers)[names(adult_tcell_markers) %in% names(ped_tcell_markers)]

## Option 1: union
union_gene_list<-lapply(merge_programs, function(x){
  c(adult_tcell_markers[[x]], ped_tcell_markers[[x]])
})
names(union_gene_list)<- merge_programs

## Option 2: intersection
intersect_gene_list<-lapply(merge_programs, function(x){
  print(x)
  adult<- adult_tcell_markers[[x]]
  ped<-ped_tcell_markers[[x]]
  merged<- adult[adult %in% ped]
  print(length(merged))
  return(merged)
})
names(intersect_gene_list)<- merge_programs

## Add on "specific" programs
specific_program_names<- c(names(adult_tcell_markers)[!names(adult_tcell_markers) %in% names(ped_tcell_markers)],
                           names(ped_tcell_markers)[!names(ped_tcell_markers) %in% names(adult_tcell_markers)])
specific_gene_list<- c(adult_tcell_markers[names(adult_tcell_markers) %in% specific_program_names],
                       ped_tcell_markers[names(ped_tcell_markers) %in% specific_program_names])

## Generate list of new genesets
## Note: these are NOT the genesets used for downstream analysis- used as baseline to assign programs --> generate new markers
## These markers don't have any pvalue/logFC associated, so difficult to use in some downstream
new_programs<- list(union=c(union_gene_list, specific_gene_list),
                    intersect=c(intersect_gene_list, specific_gene_list))
saveRDS(new_programs, file=paste0(analysis_dir, "merged.ped.adultRank",rank_use, ".genes_specific_shared.union.intersect.Rds"))
```

## Plot overlap of genes / correlation of scores between "shared" programs
To show that "shared" programs in T cells are quite similar to one another (in contrast to myeloid)
```{r}
## Programs to merge
merge_programs<- names(adult_tcell_markers)[names(adult_tcell_markers) %in% names(ped_tcell_markers)]

## Score for annotated programs- pre-merging/re-generating
names(ped_tcell_markers)<-paste0("Ped_" ,names(ped_tcell_markers))
names(adult_tcell_markers)<- paste0("Adult_", names(adult_tcell_markers))

scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean,c(ped_tcell_markers, adult_tcell_markers) )))
scores_factor_hc = clusterNmfFactors(scores)
scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
cor_shared_programs<- sapply(merge_programs, function(program){
  scores_factor_cor[paste0("Ped_", program), paste0("Adult_", program)]
})

## Get overlap of "shared" programs
gene_overlaps<- sapply(merge_programs, function(program){
  sum(unlist(unname(ped_tcell_markers[grepl(program, names(ped_tcell_markers))])) %in% 
        unlist(unname(adult_tcell_markers[grepl(program, names(adult_tcell_markers))])))
})

## df with correlation/overlaps
overlaps<- data.frame(program=merge_programs,
                nOverlap=as.numeric(plyr::mapvalues(merge_programs, names(gene_overlaps), gene_overlaps)),
                PearsonCorrelation=as.numeric(plyr::mapvalues(merge_programs, names(cor_shared_programs), cor_shared_programs)))

## Plot numeric overlaps
overlaps$dummy<-"x"
ggplot(overlaps, aes(y=program, x=dummy, color=PearsonCorrelation, size=nOverlap))+
  geom_point()+
  scale_color_gradient2(low="yellow",mid="orange", high="red", midpoint=0.7)+
  theme_classic()+
  theme(axis.text.y = element_text(face="bold"),
        axis.text.x = element_blank())+
  ylab("")+
  xlab("")
ggsave(paste0(figure_dir, "Overlap_sharedPedAdult_Rank", rank_use, ".png"), width=4, height=3)

## save df for plotting simultaneously with myeloid
write.csv(overlaps, file = paste0(analysis_dir, "sharedPedAdult_rank", rank_use, "_nGeneOverlap_PearsonCorr.csv"))
```

## Score all cells for "new" programs, assign max program 
intersect: 65 cells score <0, 301<.2. Spread evenly across programs
```{r}
rank_use<- 8
new_programs<- readRDS(paste0(analysis_dir, "merged.ped.adultRank",rank_use, ".genes_specific_shared.union.intersect.Rds"))
union_intersect<-"intersect"
gene_list<- new_programs[[union_intersect]]

## Score
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, gene_list)))

## Assign max program/max score
scores$MaxProgram<- apply(scores, 1, function(x){names(x)[which.max(x)]})
scores$MaxScore<- apply(scores[,colnames(scores) != "MaxProgram"], 1, function(x){x[which.max(x)]})

saveRDS(scores, file=paste0(analysis_dir, "scores_merged.ped.adultRank", rank_use, "_specific.sharedBy", union_intersect, ".Rds"))
```


## Compare new assignments to original (confusion matrix)- which is more similar to original?
Use this to inform the intersect vs union choice
Separately for ped/adult bc of ped/adult specific programs (treg, stress, etc)
```{r}
rank_use<-8
intersect_name<- paste0("NMF_rank", rank_use, "_annot_mergedByintersect")
union_name<- paste0("NMF_rank", rank_use, "_annot_mergedByunion")
original_name<- paste0("NMF_rank", rank_use, "_annot")
ped_adult<- "Adult"

seurat_tmp<- subset(tcell_seurat, Age==ped_adult)
meta_tmp<- seurat_tmp@meta.data[,c(intersect_name, union_name, original_name)]
colnames(meta_tmp)<-c("byIntersect", "byUnion", "Original")

## Confusion matrix for annotations
intersect_con.mat<-as.data.frame.matrix(table(meta_tmp$Original,
                          meta_tmp$byIntersect))

union_con.mat<- as.data.frame.matrix(table(meta_tmp$Original,
                          meta_tmp$byUnion))

## Remove NMF8 (ped)- no markers, no use downstream
intersect_con.mat<- intersect_con.mat[rownames(intersect_con.mat)!= "8",]
union_con.mat<- union_con.mat[rownames(union_con.mat)!= "8",]
meta_tmp<- meta_tmp[meta_tmp$Original != "8",]

## Melt for plotting
intersect_con.mat$Original<- rownames(intersect_con.mat)
union_con.mat$Original<- rownames(union_con.mat)

intersect_melt<- melt(intersect_con.mat, id="Original")
union_melt<- melt(union_con.mat, id="Original")

colnames(intersect_melt)<- c("Original", "New_byIntersect", "NCells")
colnames(union_melt)<- c("Original", "New_byUnion", "NCells")

## Convert N cells to proportion
intersect_melt<- intersect_melt %>% dplyr::group_by(Original) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()
union_melt<- union_melt %>% group_by(Original) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()

## Plot confusion matrix
p1<-ggplot(intersect_melt, aes(x=Original, y=New_byIntersect, fill=PerCells))+
  geom_tile()+
  scale_fill_gradient(low="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Original vs New by intersect")

p2<-ggplot(union_melt, aes(x=Original, y=New_byUnion, fill=PerCells))+
  geom_tile()+
  scale_fill_gradient(low="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Original vs New by union")
p1+p2
ggsave(paste0(figure_dir, "ConfusionMat_",ped_adult, "_NMFrank", rank_use, "_union.v.intersect.png"), width=12, height=5)

## Plot number of cells mismatched in each + total mismatched
  ## Get number of cells matched/mismatched for intersection/union vs original
  mismatch<- meta_tmp %>% group_by(Original) %>% dplyr::summarise(intersect_mismatch=sum(Original != byIntersect),
                                                           intersect_match= sum(Original == byIntersect),
                                                           union_mismatch=sum(Original != byUnion),
                                                           union_match= sum(Original == byUnion)) %>% 
    as.data.frame()
  
 ## Remove NMF8, add on "total"
  mismatch<- mismatch[mismatch$Original != "8",]
  totals<- as.data.frame(t(as.data.frame(apply(mismatch[,colnames(mismatch) != "Original"], 2, function(x){sum(x)}))))
  rownames(totals)<- NULL
  totals$Original<- "Total"
  mismatch<- rbind(mismatch, totals[,colnames(mismatch)])

  
  ## melt for plotting
  mismatch_melt<- melt(mismatch, id="Original")
  colnames(mismatch_melt)<-c("Original", "variable", "NCells")
  
  ## Add on union.intersection/matching columns (splitting melt variable)
  mismatch_melt$UnionIntersect<- sapply(as.character(mismatch_melt$variable),function(x){unlist(strsplit(x, split="_"))[1]})
  mismatch_melt$Match<- sapply(as.character(mismatch_melt$variable),function(x){unlist(strsplit(x, split="_"))[2]})
  
  ## Plot
  mismatch_melt$Original<-gsub("\\.", "\n", gsub("Effector", "Effector\n", mismatch_melt$Original))
  ggplot(mismatch_melt, aes(x=UnionIntersect, y=NCells, fill=Match))+
    geom_bar(stat="identity", position="stack")+
    theme_bw()+
    facet_grid(cols=vars(Original))+
    theme(axis.text.x = element_text(angle=45, hjust=1))+
    xlab("")+ylab("Number of cells")
  
  ggsave(paste0(figure_dir, "MismatchRate_",ped_adult, "_NMFrank", rank_use, "_union.v.intersect.png"), width=7, height=3)

```



## Use assignments to generate new gene list
```{r}
rank_use<- 8
union_intersect<-"intersect"
scores<- readRDS(paste0(analysis_dir, "scores_merged.ped.adultRank", rank_use, "_specific.sharedBy", union_intersect, ".Rds"))
variable_name<- paste0("NMF_rank", rank_use, "_annot_mergedBy", union_intersect)

## Add max program to seurat
tcell_seurat@meta.data[[variable_name]]<- plyr::mapvalues(colnames(tcell_seurat), rownames(scores), scores$MaxProgram)


## Generate new gene lists
tcell_seurat<- SetIdent(tcell_seurat, value=tcell_seurat@meta.data[[variable_name]])
all_markers<- FindAllMarkers(tcell_seurat, min.pct = 0.2)
all_markers_filtered<- all_markers[all_markers$p_val_adj<0.05 & all_markers$avg_log2FC>0.5,]
top_markers<- all_markers_filtered %>% group_by(cluster) %>% top_n(n=50, wt=avg_log2FC) %>% as.data.frame()

saveRDS(all_markers_filtered, file=paste0(analysis_dir, "new_markers_NMFrank", rank_use, 
                                          "_bySeurat_sharedBy", union_intersect, ".Rds"))
write.csv(top_markers, file=paste0(analysis_dir, "top_markers_NMFrank", rank_use, "bySeurat_sharedBy", union_intersect, ".csv"))
save(tcell_seurat, file=paste0("../../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/Tcells_Ped.Adult_Merge/",
                              "adult_ped_tcell_merged_seurat_harmony.Rds"))
```



## Compare gene lists between original/by seurat markers
```{r}
library(ggVennDiagram)

## Read in annotated gene lists (see above)
union_intersect<-"intersect"
rank_use<- 8
new_markers<- readRDS(paste0(analysis_dir, "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", union_intersect, ".Rds"))
old_markers<-  readRDS(paste0(analysis_dir, "merged.ped.adultRank", rank_use, ".genes_specific_shared.union.intersect.Rds"))
old_markers<- old_markers[[union_intersect]]


## Split new markers into list
new_markers<- split(new_markers, f=new_markers$cluster) 
new_markers<- lapply(new_markers, function(x){x$gene})

## Venn diagram comparing overlap for genesets
comparisons<- names(new_markers)
comparison_list<- lapply(comparisons, function(x){
  list(Original=old_markers[[x]], New=new_markers[[x]])
})
names(comparison_list)<- comparisons

all_plots<- lapply(names(comparison_list), function(x){
  tmp_list<- comparison_list[[x]]
  ggVennDiagram(tmp_list, label="count")+
    scale_fill_gradient(low="white", high="red")+
    scale_color_manual(values=rep("black", 3))+
    ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "VennDiagram_bySeurat_vs_original.sharedBy",union_intersect, "_NMFrank", rank_use, ".png"), 
       width=9, height=7)
```



## Heatmap of top genes for new, by seurat gene lists
```{r}
## Read in annotated gene lists
union_intersect<-"intersect"
rank_use<-9
n_top<-5
tcell_markers<- readRDS(paste0(analysis_dir, "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", union_intersect, ".Rds"))
variable_name<- paste0("NMF_rank", rank_use, "_annot_mergedBy", union_intersect)

## Get top genes
top_genes_lf<- tcell_markers %>% group_by(cluster) %>% top_n(n=n_top, wt=avg_log2FC) %>% as.data.frame()
top_genes_lf<- top_genes_lf[order(as.character(top_genes_lf$cluster)),]
  

## plot on heatmap using ped/adult seurat object
backup<- ScaleData(tcell_seurat, do.scale = FALSE, do.center=TRUE, features=rownames(tcell_seurat))
DoHeatmap(backup, features=top_genes_lf$gene, group.by = variable_name)+
  scale_fill_gradient2(low="blue", mid="white", high="red")
ggsave(paste0(figure_dir, "Heatmap_AllTcells_NMFrank", rank_use, "bySeuratTopMarkers.sharedBy", union_intersect, ".png"), 
       width = 8, height=12)

```

## UMAP of all tcells by NMF annot
```{r}
rank_use<-9
union_intersect<-"intersect"
variable_name<- paste0("NMF_rank", rank_use, "_annot_mergedBy", union_intersect)

## With cyto/inhibitory split
colors_use<-c(Cytotoxic.CD4="red",Cytotoxic.CD8="orange", EffectorMemory="olivedrab", `Inhibitory.CD4`="turquoise",
              Inhibitory.CD8="blue", Memory="purple", `NK-like`="green", Stress="yellow", Treg="grey22",
              RiboActive="pink")

DimPlot(tcell_seurat, group.by=variable_name)+
  scale_color_manual(values=colors_use)
ggsave(paste0(figure_dir, "UMAP_AllTcells_NMFrank", rank_use, "bySeuratTopMarkers.sharedBy", union_intersect, ".png"), 
       width=7, height=5)

  ## Plot each individually
  all_plots<- lapply(unique(tcell_seurat@meta.data[[variable_name]]), function(x){
    tcell_seurat$tmp<- ifelse(tcell_seurat@meta.data[[variable_name]]==x, x, "Other")
    DimPlot(tcell_seurat, group.by="tmp")+
      scale_color_manual(values=c(colors_use, Other="grey"))+NoLegend()+
      ggtitle(x)
  })
  cowplot::plot_grid(plotlist=all_plots, ncol=4)
  ggsave(paste0(figure_dir, "UMAP_AllTcells_NMFrank", rank_use, "indv_bySeuratTopMarkers.sharedBy", union_intersect, ".png"),
         width=12, height=9)
  

```

## Breakdown of each program by age
```{r}
rank_use<-8
union_intersect<-"intersect"
variable_name<- paste0("NMF_rank", rank_use, "_annot_mergedBy", union_intersect)

## Number of ped/adult cells by program
df_bd<- as.data.frame(table(tcell_seurat$Age, tcell_seurat@meta.data[[variable_name]]))
colnames(df_bd)<-c("Age", "Program", "Ncells")

## Normalize cell counts to account for difference in cohort size- so ped/adult are weighted the same
nCells<- table(tcell_seurat$Age)
normFactor<- max(nCells)/nCells
df_bd$NormFactor<- as.numeric(as.character(plyr::mapvalues(df_bd$Age, names(normFactor), normFactor)))
df_bd$Ncells_norm<- df_bd$Ncells * df_bd$NormFactor

## Proportion of normalized ped/adult counts per program
df_bd<- df_bd %>% group_by(Program) %>% mutate(perCells_norm=Ncells_norm/sum(Ncells_norm)) %>% as.data.frame()

ggplot(df_bd, aes(x=Program, y=perCells_norm, fill=Age))+
  geom_bar(stat="identity", position="stack")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("")+ylab("Proportion of cells\nNormalized for cohort size")+
  scale_fill_manual(values=c(Ped="red", Adult="blue"))+
  geom_hline(yintercept = 0.5, linetype="dashed")
ggsave(paste0(figure_dir, "NormalizedCellProportion_byAge_byProgram.rank", rank_use, "_mergedBy", union_intersect, ".png"), width=5, height=3)
```


## Expression of expected marker genes based on NMF annotation
```{r}
rank_use<-8
union_intersect<-"union"
variable_name<- paste0("NMF_rank", rank_use, "_annot_mergedBy", union_intersect)

DotPlot(tcell_seurat, group.by = variable_name, 
        features=c("CD4", "CD8A", "GZMK", "KLRB1","PTGER2","CTLA4","LAG3", "TIGIT","PDCD1", 
                   "HAVCR2", "TOP2A", "CCR7", "GNLY", "HSPA1A", "FOXP3"))+
  xlab("") + ylab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave(paste0(figure_dir, "DotPlot_expectedMarkerGenes_forNMFrank", rank_use, "_bySeuratTopMarkers.sharedBy", 
              union_intersect, ".png"), 
       width=7, height=4)

## Split all into ped/adult
tcell_seurat$tmp<- paste0(tcell_seurat@meta.data[[variable_name]], "_", tcell_seurat$Age)
DotPlot(tcell_seurat, group.by = "tmp", 
        features=c("CD4", "CD8A", "GZMK","KLRB1","PTGER2", "CTLA4","LAG3", "TIGIT","PDCD1", 
                   "HAVCR2", "TOP2A", "CCR7", "GNLY", "HSPA1A", "FOXP3"))+
  xlab("") + ylab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave(paste0(figure_dir, "DotPlot_expectedMarkerGenes_forNMFrank", rank_use, "_bySeuratTopMarkers.sharedBy", 
              union_intersect, ".byAge.png"), 
       width=8, height=5)

```


## Add on annotations to individual ped/adult tcell seurat objects
```{r}
cohort<- "pedOnly_nomuv63"
ped_tcell<- readRDS(paste0("../../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",
                           cohort, "/tcell_seurat_harmony.Rds" ))

adult_tcell<- readRDS(paste0("../../../02a_Tcells/01b_Adult/01_NMF/analysis/deNovoNMF/tcell_seurat.Rds"))

## Map annotations on
## A few (13 cells) not in original- removed due to annotation/clustering issue. Rename NMF as "Unclear"; so few cells, issue not worth revisitng
rank_use<- 8
sharedBy<- "union"
variable_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy)

ped_tcell@meta.data[[variable_name]]<-ifelse(colnames(ped_tcell) %in% colnames(tcell_seurat), 
                                             plyr::mapvalues(colnames(ped_tcell), colnames(tcell_seurat),
                                                       tcell_seurat@meta.data[[variable_name]],
                                                       warn_missing = FALSE),
                                             "Unclear")
adult_tcell@meta.data[[variable_name]]<- ifelse(colnames(adult_tcell) %in% colnames(tcell_seurat), 
                                             plyr::mapvalues(colnames(adult_tcell), colnames(tcell_seurat),
                                                       tcell_seurat@meta.data[[variable_name]],
                                                       warn_missing = FALSE),
                                             "Unclear")


saveRDS(ped_tcell, file=paste0("../../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",
                           cohort, "/tcell_seurat_harmony.Rds" ) )
saveRDS(adult_tcell, file=paste0("../../../02a_Tcells/01b_Adult/01_NMF/analysis/deNovoNMF/tcell_seurat.Rds"))
```



## Exploratory- can delete
```{r}
adult_tregs<- subset(tcell_seurat, NMF_rank7_annot=="Treg")
test<- as.data.frame(t(adult_tregs@assays$RNA@data[c("CTLA4", "TIGIT","HAVCR2", "FOXP3"),]))
test<- test[rowSums(test)>0,]
hist(test$FOXP3)
test$lowfoxp3<- test$FOXP3<0.5 ## 430 tregs with no/low foxp3
test<- melt(test, id="lowfoxp3")
colnames(test)<-c("FOXP3_expr", "Gene", "logTPM")
ggplot(test, aes(x=Gene, y=logTPM, fill=FOXP3_expr))+
  geom_boxplot()


ped<- subset(tcell_seurat, Age=="Ped")
foxp3<- as.data.frame(t(ped@assays$RNA@data[c("FOXP3", "TIGIT"),]))
foxp3$nmf<- plyr::mapvalues(rownames(foxp3), colnames(ped), ped$NMF_rank7_annot)
foxp3$potential_tregs<- foxp3$FOXP3>2
table(foxp3$potential_tregs, foxp3$nmf)

inh_expressors<- as.data.frame(t(tcell_seurat@assays$RNA@data[c("TIGIT", "HAVCR2", "CTLA4", "FOXP3"),]))
inh_expressors<- inh_expressors[rowSums(inh_expressors)!=0,]
inh_expressors$mean_inh<- rowMeans(inh_expressors[,c("TIGIT", "HAVCR2", "CTLA4")])
inh_expressors$highFOXP3<- inh_expressors$FOXP3>2
inh_expressors$lowFox.highInh<- inh_expressors$mean_inh>2 & !inh_expressors$highFOXP3
inh_expressors$program<- plyr::mapvalues(rownames(inh_expressors), colnames(tcell_seurat), tcell_seurat$NMF_rank7_annot_mergedByintersect,
                                         warn_missing = FALSE)
table(inh_expressors$program, inh_expressors$lowFox.highInh)

test<- subset(tcell_seurat, NMF_rank9_annot_mergedByunion %in% c("Inhibitory.CD4", "Inhibitory.CD8",
                                                                 "Cytotoxic.CD4", "Cytotoxic.CD8"))
test_meta<- test@meta.data
test<- RunFullSeurat(test@assays$RNA@counts, RunHarmony = TRUE, samples=test_meta$sampleid)
test<- AddMetaData(test, test_meta[,!colnames(test_meta) %in% colnames(test@meta.data)])
DimPlot(test, group.by="NMF_rank9_annot_mergedByunion")
```

