---
title: "R Notebook"
output: html_document
---

```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/Preprocessing/inferCNV/")
```

## Load in tumor calling info from original and new (with 2 new samples)
## Includes CNV calling + whether cell clustered with a normal cell
```{r}
originalClass<- readRDS(paste0("analysis/inferCNV/seurat/", "FullCNVCallingInfo.Rds"))
newClass<- readRDS(paste0("WithNewSamples/analysis/inferCNV/seurat/", 
                          "FullCNVCallingInfo.Rds"))

## Remove new samples
newClass<- newClass[!newClass$sample %in% c("BT1857", "BT1873vent"),]
sum(rownames(newClass)==rownames(originalClass)); nrow(newClass);nrow(originalClass)
```

## Merge into 1 df
```{r}
originalClass<- originalClass[,c("cnv", "sample", "ClusteredWithNormal", "Classification")]
newClass<- newClass[,c("cnv", "ClusteredWithNormal", "Classification")]

colnames(originalClass)<- c("Original_CNV", "sample", "Original_ClusteredWithNormal",
                            "Original_Classification")
colnames(newClass)<- c("New_CNV", "New_ClusteredWithNormal",
                            "New_Classification")

all_class<- cbind(originalClass, newClass)

## How many cells mismatch in terms of final classification?
mismatchCells<- all_class[all_class$Original_Classification!=all_class$New_Classification,]
## 77 cells

## How many cells mismatch in terms of clustering with normal cells?
sum(all_class$Original_ClusteredWithNormal != all_class$New_ClusteredWithNormal)
## only 1 cell: MUV086.P3.D04 
## Noise->Malig, both called as having CNV, but clustered with normal cells in original, not in new
## However- it's in cluster 16 and belongs to MUV86, so it was later removed as "normal" anyways
## in the end then, it agrees for both

## How many cells mismatch in terms of CNV calling?
sum(all_class$Original_CNV != all_class$New_CNV)
## 76 cells
```

## Plotting change in final annotation
```{r}
## All Cells
all_class$Change<- paste0(all_class$Original_Classification, "->",
                          all_class$New_Classification)
df<- as.data.frame(table(all_class$Change, all_class$sample))
colnames(df)<- c("ChangeInClassification", "sample", "NCells")

ggplot(df, aes(x=reorder(ChangeInClassification,NCells, mean), 
               y=NCells, 
               fill=sample))+
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("")

## Only mismatch cells
mismatchCells<- all_class[all_class$Original_Classification!=all_class$New_Classification,]

df<- as.data.frame(table(mismatchCells$Change, mismatchCells$sample))
colnames(df)<- c("ChangeInClassification", "sample", "NCells")

ggplot(df, aes(x=reorder(ChangeInClassification,NCells, mean), 
               y=NCells, 
               fill=sample))+
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("")
```

## Of the Noise--> malignant and malignant --> noise: are any cells in the midline histone mutant cells that were removed downstream?
These cells are in cluster 6 of the midline histone mutant clustering analysis (in both original and current)
They were removed downstream since they seemed to be clearly normal
Can any discrepancy here be removed due to this? 
i.e., a cell was called "Noisy" in original and "malig" in current, but was removed as "normal" anyways downstream?
```{r}
current_cluster6<- readRDS("../../Annotation/BySubtype/Midline/Midline_HistoneMutant/WithNewSample/analysis/SeuratCluster6_PossiblyNormal.Rds")

original_cluster6<-readRDS("../../Annotation/BySubtype/Midline/Midline_HistoneMutant/analysis/SeuratCluster6_PossiblyNormal.Rds")

sum(rownames(current_cluster6) %in% rownames(original_cluster6))
## All but 2 cells matches: 
## "MUV086.P3.D04" in current only, "MUV086.P3.A07" in original only

## How many "mismatch" cells are in either cluster 6?
sum(rownames(mismatchCells) %in% rownames(current_cluster6))
## just 1: "MUV086.P3.D04"

sum(rownames(mismatchCells) %in% rownames(original_cluster6))
## None

## so: the whole "odd midline histone mutant normal cluster" removes discrepancy for 1 cell:
## MUV086.P3.A07
## Was called noisy in original, malig in current, but removed as normal in current anyways
## NOTE: MUV086.P3.A07 was called malig in both, BUT removed in normal cluster 6 downstream for original ONLY. That means that this cell really goes Normal --> malig
```

