---
title: "_inferCNV"
output: html_document
---

```{r load_packages}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and functions
```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/Preprocessing/inferCNV/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
inferCNV = paste0(base_dir, script_dir, "inferCNV_helper_functions.R")
source(preprocessing)
source(inferCNV)

gene_order_file = paste0(base_dir, "/Dropbox (Partners HealthCare)/FilbinLab/data_analysis/genome/gencode_v19_gene_pos.txt")
ext_ctrl = paste0(base_dir, "/Dropbox (Partners HealthCare)/FilbinLab/data_analysis/ctrl_data/")


## Create output files
inferCNV_analysis_folder = "analysis/inferCNV/"
inferCNV_fig_folder = "figures/inferCNV/"
for (folder in c(inferCNV_analysis_folder, inferCNV_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder, recursive = TRUE)
  }
}

inferCNV_input_folder<- "analysis/inferCNV/"
```

## Load input data (observations + control) and gene order file
## Including: cm, samples, CNV values, 
## and labels for obsevations (obs) and references (ref)
```{r load_data}
cm = readRDS(paste0(inferCNV_input_folder, 
                    "cm_exp_ctrl.rds"))
samples = readRDS(paste0(inferCNV_input_folder, 
                         "samples_exp_ctrl.rds"))
all_cnv_values = readRDS(paste0(inferCNV_input_folder,
                                "all_cnv_values_All.rds"))
obs_ref = readRDS(paste0(inferCNV_input_folder,
                         "obs_ref_All.rds"))

gene_order_file = paste0(base_dir, 
                         "/Dropbox (Partners HealthCare)/FilbinLab/data_analysis/genome/gencode_v19_gene_pos.txt")
gene_order = read.table(gene_order_file, header = F, sep = "\t", row.names = 1)
```

## In each sample, weighted pairwise correlation of cells + references 
## Hierarchical clustering using correlatiton coefficient 
## Cutree 
```{r hierarchical_clustering}
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend)
library(weights)

cnv_hc_each_sample = list()
cnv_cutree_each_sample = list()
weights_list = list()
nTree = 4
samples<-samples

## For all samples
for (sample in sort(unique(samples))){
  ## Subset each sample + (mg and od)
  to_subset = (samples == sample | samples == "mg" | samples == 'od')
  message("======================================")
  message("Processing sample: ", sample)
  message("Start time: ", Sys.time())
  ## Compute weights
  weights = apply(all_cnv_values[,to_subset], 1, var)
  weights_list[[sample]] = weights
  ## Compute pairwise correlations and distances
  pairwise_cor = wtd.cors(all_cnv_values[,to_subset], weight = weights) 
  pairwise_dist = 1 - pairwise_cor
  ## Hierarchical clustering and cutree 
  hc = hclust(as.dist(pairwise_dist), method="ward.D2")
  cnv_hc_each_sample[[sample]] = hc
  cnv_cutree_each_sample[[sample]] = cutree(hc, nTree)
  message("End time: ", Sys.time())
}

saveRDS(cnv_cutree_each_sample, paste0(inferCNV_analysis_folder,
                                       "cnv_cutree_each_sample.rds"))
saveRDS(cnv_hc_each_sample, paste0(inferCNV_analysis_folder,
                                   "cnv_hc_each_sample.rds"))

```


## Print out cutree results for each sample 
```{r examine_clustering_results}
cnv_cutree_table = list()
for (sample in sort(unique(samples))){
  print("======================")
  print(sample)
  print(table(samples[names(cnv_cutree_each_sample[[sample]])], cnv_cutree_each_sample[[sample]]))
  cnv_cutree_table[[sample]] = table(samples[names(cnv_cutree_each_sample[[sample]])],
                                     cnv_cutree_each_sample[[sample]])
}
names(cnv_cutree_table) = sort(unique(samples))

cnv_cutree_table_cut = list()
for (sample in sort(unique(samples))){
  cnv_cutree_table_cut[[sample]] = table(samples[names(cnv_cutree_each_sample[[sample]])],
                                     cnv_cutree_each_sample[[sample]])
}
names(cnv_cutree_table_cut) = sort(unique(samples))

## write out table- renaming rows so that it's obvious what sample each row corresponds to
newRownames<-mapply(function(x,y) paste(rownames(x), y, sep="_Sample:"), cnv_cutree_table, names(cnv_hc_each_sample))
all_cnv_cutree_table<- as.data.frame(do.call(rbind, cnv_cutree_table))
rownames(all_cnv_cutree_table)<-unname(unlist(newRownames))
write.table(all_cnv_cutree_table, paste0(inferCNV_analysis_folder, "AllCNVcutreeResults.txt"))

```

## Repeat inferCNV for target sample with references 
## Cells in the sample grouped based on previously computed hierarchical clustering
## and cutree results 
## May inspect resulting heatmaps to see if cutree results successfully separate 
## cells +/- CNV and subclones
```{r heatmap_for_normal_tumor_cells}
## For double-check samples
  normal_type = c("mg", "od")
  cutree_res = cnv_cutree_each_sample
  
  for (i in unique(samples)){
    sample = i
    folder = paste0("analysis/inferCNV/", sample, "/")
    plot_subcluster_10X(folder, sample, cm, samples, 
                        normal_type, cutree_res, gene_order_file, cutoff=1)
  }

```


## Assign CNV status to each based on the number of normal cells in its cluster 
```{r call-cnv}
normal_cutoff = 5
## Store normal clusters in a list
normal_clusters = lapply(cnv_cutree_table, 
                         function(x) which((x["mg",] + x["od",])>=normal_cutoff))
try (if(any(names(cnv_cutree_each_sample) != names(normal_clusters))) stop("cutree results list and normal clusters list should have identical names for elements"))

## Reassign normal clusters for BT1478- no  major CNVs present across the entire sample
normal_clusters$BT1478<- c(nTree+1); names(normal_clusters$BT1478)<- normal_clusters$BT1478


## Classify each cell as Normal (F) vs Tumor (T) based on normal_clusters list 
call_cnv_res = NULL
for (i in seq(length(normal_clusters))){
  sample = names(normal_clusters)[i]
  print(sample)
  normal_clust = normal_clusters[[i]]
  print(normal_clust)
  hc_clust_res = cnv_cutree_each_sample[[i]]
  print(head(hc_clust_res))
  sample_names = sapply(names(hc_clust_res), function(x) unlist(strsplit(x, split = ".", fixed = T))[1])
  print(head(sample_names))
  hc_clust_res = hc_clust_res[sample_names == sample]
  tmp = ifelse(hc_clust_res %in% normal_clust, F, T)
  names(tmp) = names(hc_clust_res)[sample_names == sample]
  call_cnv_res = c(call_cnv_res, tmp)
}

## Generate a named vector for mg and od, and label them as normal 
mg_names = names(samples)[samples == "mg"]
mg_res = rep(F, length(mg_names))
names(mg_res) = mg_names
od_names = names(samples)[samples == "od"]
od_res = rep(F, length(od_names))
names(od_res) = od_names

## Concatenate CNV calling for all samples and normal control 
call_cnv_res_complete = c(call_cnv_res, mg_res, od_res)
call_cnv_res_complete = call_cnv_res_complete[names(samples)]
call_cnv_res = call_cnv_res_complete[names(samples)[obs_ref == "obs"]]

## Store CNV calling for all samples alone and all samples + normal control 
saveRDS(call_cnv_res, paste0(inferCNV_analysis_folder, "call_cnv.rds"))
saveRDS(call_cnv_res_complete, paste0(inferCNV_analysis_folder, "call_cnv_w_ctrl.rds"))

  ## Compare cnv_res to other analysis
  ## replicating results vs original
  ## Both do not have 2 new samples (BT1873/BT1857)
  original_cnv_res<- readRDS("analysis/inferCNV/HC/call_cnv.rds")
  identical(original_cnv_res, call_cnv_res)
  
  ## analyses match exactly
```

## OPTIONAL: Save cutree results
```{r}
hc_res = sapply(unique(samples[obs_ref == "obs"]), function(x){
  tmp = cnv_cutree_each_sample[[x]]
  tmp = tmp[names(tmp) %in% names(obs_ref)[obs_ref == "obs"]]
})

hc_res = unlist(hc_res)
tmp = lapply(names(hc_res), 
                           function(x) unlist(strsplit(x, split = ".", fixed = T))[2:4])
tmp2 = sapply(tmp, function(x) paste(x[1], x[2], x[3], sep = "."))
names(hc_res) = tmp2


saveRDS(hc_res, paste0(inferCNV_analysis_folder, "cnv_hc_res.rds"))
```