---
title: "_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/Immune/Annotation/BySubtype/Hemispheric/BCOR/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)

liger_fig_dir<- paste0(working_dir, "figures/Liger/")
liger_analysis_folder<- paste0(working_dir, "analysis/liger/")
seurat_analysis_folder = "analysis/seurat/"
seurat_fig_folder = "figures/seurat/"
for (folder in c(seurat_analysis_folder, seurat_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj_annot.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}

seurat_obj<-seurat_obj
```

## Load counts and metadata for initiating seurat object
```{r load-data}
load("../../../../SplitBySubtype/analysis/hemiBCOR_cm_list.Robj")
samples = readRDS("../../../../SplitBySubtype/analysis/hemiBCOR_samples.rds")
tumor_stat = readRDS("../../../../SplitBySubtype/analysis/hemiBCOR_call_tumor.rds")

cm_list<-hemiBCOR_cm_list
sum(colnames(cm_list$raw_data)==names(samples)); length(samples)
sum(names(samples)==names(tumor_stat))
```

## Subset cm to remove noise/normal samples
```{r subset-tumor-cells}
cm = cm_list$raw_data; rm(cm_list); gc()
cm = cm[,tumor_stat=="Malignant"]
samples = samples[tumor_stat=="Malignant"]

any(colnames(cm) != names(samples))
```

## Initiate seurat obj and add metadata 
```{r seurat_obj_meta_data}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", min.cells=0, min.genes=0,
                                    scale.factor=1E5, do.scale=F, do.center=T)

## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples
```

## OPTIONAL: plot highly variable genes
```{r}
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## PCA and PC selection
```{r PCA_selection}
## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)

```

## Harmony data integration 
```{r run_harmony}
library(harmony)
library(magrittr)
library(cowplot)

## Run Harmony
seurat_obj = RunHarmony(seurat_obj, "sample", theta = 2, 
                        max.iter.harmony = 50, plot_convergence = TRUE)

## Simple dim and vlnplot to examine integration 
options(repr.plot.height = 6, repr.plot.width = 12)

p1 <- DimPlot(object = seurat_obj, reduction = "pca", pt.size = .1, group.by = "sample")
p2<- DimPlot(object = seurat_obj, reduction = "harmony", pt.size = .1, group.by = "sample")
p3 <- VlnPlot(object = seurat_obj, features = "PC_1", pt.size = .1, group.by = "sample")
p4 <- VlnPlot(object = seurat_obj, features = "harmony_1", pt.size = .1, group.by = "sample")

plot_grid(p1,p2, p3, p4)
ggsave("Harmony_diagnostics_theta2_NoBT1478.jpg", path=seurat_fig_folder, width=12, height=8)


## DimHeatmap
DimHeatmap(object = seurat_obj, reduction = "harmony", cells = 500, dims = 1:6, balanced = TRUE)
```

## Clustering and UMAP
```{r clustering_tSNE}
## Run UMAP and clustering using reduced dimensions from Harmony
## You may play around with resolution parameter to tune the clustering result
resolution = .8
reduction_type = "harmony"
dims = 20

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = reduction_type, dims = 1:dims)


seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = reduction_type,
                            dims = 1:dims,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = resolution)


r0.8<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5) + ggtitle("Resolution_0.8")
r0.6<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5)+ ggtitle("Resolution_0.6")
r1.0<-DimPlot(seurat_obj, reduction = "umap", pt.size = 2.5)+ ggtitle("Resolution_1")

r0.6+r0.8+r1.0 #resolution of 0.8 may be best
ggsave("UMAP_clustering_VaryResolution_NoBT1478.jpg", path=seurat_fig_folder, width=15, height=6)

## Clustering results 
table(Idents(seurat_obj))


## Save the seurat object
save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```

## UMAP plot colored by clustering and samples
## Can color tSNE with other metadata features (e.g. primary/recurrence, treatment)
```{r UMAP_plots}
## By clustering
DimPlot(object = seurat_obj, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 3, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering.jpg", path=seurat_fig_folder, width=8, height=8)

## By samples
DimPlot(object = seurat_obj, reduction="umap", group.by = "sample", label = TRUE, 
        pt.size = 3, label.size = 5) + NoAxes() 
ggsave("UMAP_sample.jpg", path=seurat_fig_folder, width=10, height=8)

```
## Plot breakdown of each cluster by sample
```{r}
cluster_by_sample<- table(seurat_obj@meta.data$sample, seurat_obj@meta.data$seurat_clusters)
cluster_by_sample<- melt(cluster_by_sample)
colnames(cluster_by_sample)<- c("Sample", "Cluster", "NCells")
cluster_by_sample$Cluster<- paste0("Cluster", cluster_by_sample$Cluster)

ggplot(cluster_by_sample, aes(x=Cluster, y=NCells, fill=Sample))+ 
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Seurat clustering: \nbreakdown of samples per cluster")
ggsave(paste0(seurat_fig_folder, "SeuratClusterSampleDistr.png"), width=5, height=3)
```


## Optional: Remove sample BT1478 --> reprocess
```{r}
seurat_obj_noBT1478<- subset(seurat_obj, sample!="BT1478")
cm_noBT1478<- seurat_obj_noBT1478@assays$RNA
samples_noBT1478<- samples[samples %in% seurat_obj_noBT1478@meta.data$sample]

## Create seurat project
seurat_obj_noBT1478 = preprocessSeuratObject(cm_noBT1478, project="", min.cells=0, min.genes=0,
                                    scale.factor=1E5, do.scale=F, do.center=T)

## Add samples and subtype as meta.data 
seurat_obj_noBT1478@meta.data$sample = samples_noBT1478

## PCA
seurat_obj_noBT1478 <- RunPCA(
  object = seurat_obj_noBT1478, 
  features = seurat_obj_noBT1478@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)

## Run Harmony
seurat_obj_noBT1478 = RunHarmony(seurat_obj_noBT1478, "sample", theta = 2, 
                        max.iter.harmony = 50, plot_convergence = TRUE)

## Run UMAP
seurat_obj_noBT1478 <- RunUMAP(seurat_obj_noBT1478, reduction = "harmony", dims = 1:20)


seurat_obj_noBT1478 <- FindNeighbors(seurat_obj_noBT1478,
                            reduction = "harmony",
                            dims = 1:20,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = 0.8)

## By clustering
DimPlot(object = seurat_obj_noBT1478, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 3, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_NoBT1478.jpg", path=seurat_fig_folder, width=8, height=8)

## By samples
DimPlot(object = seurat_obj_noBT1478, reduction="umap", group.by = "sample", label = TRUE, 
        pt.size = 3, label.size = 5) + NoAxes() 
ggsave("UMAP_sample.jpg_NoBT1478.jpg", path=seurat_fig_folder, width=10, height=8)

save(seurat_obj_noBT1478, file = paste0(seurat_analysis_folder, "seurat_obj_NoBT1478.Robj"))

## Set seurat_obj to seurat_obj_NoBT1478 for downstream analysis
seurat_obj<-seurat_obj_noBT1478
```

## DEGs for each cluster 
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.5)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05 & all_markers$avg_logFC > 0.5, ]

seurat_obj@meta.data$original_cluster<- seurat_obj@meta.data$seurat_clusters

## cluster 0 vs cluster 1
c0_v_c1<- FindMarkers(seurat_obj, ident.1 = 0, ident.2 = 1)
c0_v_c1_filtered<- c0_v_c1[c0_v_c1$p_val_adj<0.05,]
write.csv(c0_v_c1_filtered, paste0(seurat_analysis_folder, "DEGs_Cluster1vCluster0.csv"))

## Merge clusters 1/2/5, then re-find all markers
newIdents<- gsub("0", "1",   seurat_obj@meta.data$original_cluster)
names(newIdents)<-colnames(seurat_obj)
seurat_obj@meta.data$new_clusters<- newIdents
seurat_obj<- SetIdent(seurat_obj, value=newIdents)

DimPlot(object = seurat_obj, reduction="umap", 
        label = TRUE, pt.size = 3, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_Clusters0.1Merged.jpg", path=seurat_fig_folder, width=8, height=8)

all_markers_mergedClusters<- FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.5)
all_markers_filtered_mergedClusters = all_markers_mergedClusters[all_markers_mergedClusters$p_val_adj < 0.05 &
                                                    all_markers_mergedClusters$avg_logFC > .5, ]
all_markers_filtered<- all_markers_filtered_mergedClusters

mergedc0.1_v_c5<- FindMarkers(seurat_obj, ident.1=1, ident.2 = 5)
mergedc0.1_v_c5_filtered<-mergedc0.1_v_c5[mergedc0.1_v_c5$p_val_adj<0.05 &
                                            abs(mergedc0.1_v_c5$avg_logFC)>1,]
write.csv(mergedc0.1_v_c5_filtered, paste0(seurat_analysis_folder, "DEGs_MergedCluster1.0vCluster5.csv"))

## Subset the top 30 DEGs for each cluster
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC) %>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(seurat_analysis_folder, "all_markers.Robj"))
write.csv(top_marker_genes, file=paste0(seurat_analysis_folder, "Top30Markers_NoBT1478_Clusters01Merged.csv"))

## Heatmap for initial inspection
DoHeatmap(seurat_obj, 
          features=all_markers_filtered_mergedClusters$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs_theta2_NoBT1478_Clusters014Merged.jpg", path=seurat_fig_folder)
```

## Quick annotation by OPC/OC/AC marker genes 
```{r}
## OPC-like
FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OPC_markers_NoBT1478.jpg", path=seurat_fig_folder, width = 9, height=9)




## OC-like cells
FeaturePlot(seurat_obj,
            c("BCAS1", "SIRT2", "RGR", 
              "GPR17", "NFASC", "MBP",
              "TF", "MOG", "PLP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OC_markers_NoBT1478.jpg", path=seurat_fig_folder, width = 9, height=9)

## AC-like cells 
FeaturePlot(seurat_obj,
            c("APOE", "AQP4", "AGT", 
              "SPARCL1","ATP1A2", "GFAP",
              "S100B", "ALDOC", "CLU"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("AC_markers_NoBT1478.jpg", path=seurat_fig_folder, width=9, height=9)

## Neuronal-like cells
FeaturePlot(seurat_obj,
            c("STMN2", "STMN1", "STMN4", 
              "TUBB3", "INA", "ELAVL4",
              "SNAP25", "GABRG2", "ST18"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("Neuron_markers_NoBT1478.jpg", path=seurat_fig_folder, width=9, height=9)



##CLEC2D
FeaturePlot(seurat_obj,
            c("CLEC2D", "WNT5A", "PVR", "CDH1", "SPP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("InhibitoryLigand_markers.jpg", path=seurat_fig_folder, width=9, height=6)
```
## Vln plots
```{r}
VlnPlot(seurat_obj, 
       c("CLEC2D", "WNT5A", "PVR", "CDH1", "SPP1"),
        pt.size = 0, sort = TRUE)
ggsave("InhibitoryLigand_markers_ViolinPlot.jpg", path=seurat_fig_folder, width=9, height=6)

VlnPlot(seurat_obj, group.by = "sample",
       c("CLEC2D", "WNT5A", "PVR", "CDH1", "SPP1"),
        pt.size = 0, sort = TRUE)
ggsave("InhibitoryLigand_markers_ViolinPlot_samples.jpg", path=seurat_fig_folder, width=9, height=6)
```

## ORA test 
```{r over-representation-test}
library(clusterProfiler)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
allOE_genes <- as.character(gene_symbol_to_ensembl_id(rownames(seurat_obj@assays$RNA@counts),dataset)$ensembl_gene_id)


## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
go_result = list()
for (i in unique(all_markers_filtered$cluster)){
  sigOE_genes = as.character(gene_symbol_to_ensembl_id(subset(all_markers_filtered, cluster==i,
                                                              gene)[,], dataset)$ensembl_gene_id)
  go_result[[paste0("c", i)]] = go_analysis(sigOE_genes, allOE_genes)
} 

## Additionally: DEGs between merged clusters0/1 and cluster 5
cluster5_DEGs<- mergedc0.1_v_c5_filtered[mergedc0.1_v_c5_filtered$avg_logFC<0,]
cluster0.1_DEGs<- mergedc0.1_v_c5_filtered[mergedc0.1_v_c5_filtered$avg_logFC>0,]

go_result = list()
sigOE_genes_c5 = as.character(gene_symbol_to_ensembl_id(rownames(cluster5_DEGs), dataset)$ensembl_gene_id)
go_result[["c5"]] = go_analysis(sigOE_genes_c5, allOE_genes)

sigOE_genes_c1 = as.character(gene_symbol_to_ensembl_id(rownames(cluster0.1_DEGs), dataset)$ensembl_gene_id)
go_result[["c1"]] = go_analysis(sigOE_genes_c1, allOE_genes)
go_result<- go_result[names(go_result) %in% c("c1", "c5")] 

## Plot top pathways for all clusters, display on one plot
clusters<-names(go_result)
allDotPlots_20<-list()
for (c in 1:length(clusters)){
  cluster=clusters[c]
  print(cluster)
  if (nrow(go_result[[cluster]]$cluster_summary)>0){
    print(cluster)
    allDotPlots_20[[c]]<-dotplot(go_result[[cluster]]$ego, 
                                showCategory=10, 
                                font.size = 15, 
                                title = cluster, 
                                label_format=10) +
      scale_y_discrete(labels=function(x) str_wrap(x, width=60))+
      theme(plot.title = element_text(hjust = 0.5, face = "bold", 
                                      color="black", size = 28),
            axis.title = element_text(face = "bold", color="black"), 
            axis.text.x = element_text(angle = 45, hjust = 1, color="black"),
            axis.text.y = element_text( color="black",face = "bold"))
  }
  
}

allDotPlots_20<- allDotPlots_20[!(unlist(lapply(allDotPlots_20, function(x){is.null(x)})))]
cowplot::plot_grid(plotlist = allDotPlots_20, ncol=2)
ggsave(paste0("GO_OVA_Cluster5vMergedCluster0.1_top10.jpg"), path=seurat_fig_folder, width=18, height=5)
save(go_result, file=paste0(seurat_analysis_folder,  "go_result_Cluster5vMergedCluster0..Robj"))
```

# Add NMF results to the seurat object

## Map NMF scores to seurat object 
```{r add-NMF-score}
NameOfMetaprogramList<-"Merge1"

nmf_score = readRDS(paste0("analysis/nmf/Merge/nmf_score", NameOfMetaprogramList, ".rds"))
seurat_obj = AddMetaData(seurat_obj, nmf_score)
load(paste0("analysis/nmf/Merge/nmf_marker_genes_final", NameOfMetaprogramList, ".Robj"))

DimPlot(object = seurat_obj, group.by = "signature_1", label = TRUE, 
         pt.size = 3, label.size = 5, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_nmf_signature_Theta2",NameOfMetaprogramList, ".jpg"), 
       path=seurat_fig_folder, 
       width=14, height=12)

#DoHeatmap(seurat_obj, 
          #features=unlist(nmf_marker_genes_final)) + NoLegend()

#ggsave("nmf_marker_genes.jpg", path=seurat_fig_folder, width=12, height=8)
save(seurat_obj, file=paste0(seurat_analysis_folder, "seurat_obj_NoBT1478.Robj"))
```

```{r}
save(seurat_obj, file=paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```


## Map liger NMF score to seurat object
```{r}
## From LIGER
load(paste0(liger_analysis_folder, "ligerFactor_Assignment.Robj"))

## From scoreNmfGenes of top marker genes
liger_factorAssignments<-readRDS(paste0(liger_analysis_folder, "nmf_score_AllFactors", ".Rds"))
liger_factorAssignments<-liger_factorAssignments$signature_1

seurat_obj@meta.data$TopLigerFactor<- liger_factorAssignments


DimPlot(object = seurat_obj, group.by = "TopLigerFactor", label = TRUE, 
         pt.size = 2, label.size = 5, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_LigerFactors_fromScoreNmfGenes", ".jpg"), 
       path=seurat_fig_folder, 
       width=7, height=7)

#DoHeatmap(seurat_obj, 
          #features=unlist(nmf_marker_genes_final)) + NoLegend()

#ggsave("nmf_marker_genes.jpg", path=seurat_fig_folder, width=12, height=8)
save(seurat_obj, file="analysis/seurat/seurat_obj.Robj")
```


## Add "Shared" factor- only some cells with have shared factors
```{r}
load("analysis/liger/TopLigerFactor_SharedScoreAbove2.Robj")
sharedScore_df<- semiShared_HighScores[semiShared_HighScores$cell %in% colnames(seurat_obj),]
sharedScore<- sharedScore_df$signature_1
names(sharedScore)<-sharedScore_df$cell

noSharedScore_cells<- colnames(seurat_obj)[!(colnames(seurat_obj) %in% names(sharedScore))]
noSharedScore<- rep("-", length(noSharedScore_cells))
names(noSharedScore)<- noSharedScore_cells

sharedScore_forSeurat<- c(sharedScore, noSharedScore)
sharedScore_forSeurat<- sharedScore_forSeurat[order(names(sharedScore_forSeurat))]

seurat_obj@meta.data$SharedFactor<-sharedScore_forSeurat
DimPlot(seurat_obj, group.by="SharedFactor", cols=c("grey", "magenta", "blue", "green", "orange", "purple"), pt.size=2)
ggsave(paste0(seurat_fig_folder, "SharedPrograms_ScoreAbove2.png"))

```


## Differential expression of ligands between more progenitor (factors 3/8/9)
## And more differentiated (factors 10/13/14)
```{r}
prog_v_diff<- paste0("factor", c(3,8,9,10,13,14))
names(prog_v_diff)<- c(rep("progenitor", 3), rep("differentiated", 3))


seurat_prog_v_diff<- subset(seurat_obj, TopLigerFactor %in% prog_v_diff)
seurat_prog_v_diff@meta.data$prog_v_diff<- plyr::mapvalues(seurat_prog_v_diff@meta.data$TopLigerFactor,
                                                           from=prog_v_diff, to=names(prog_v_diff))
Idents(seurat_prog_v_diff)<- seurat_prog_v_diff@meta.data$prog_v_diff

FindMarkers(seurat_prog_v_diff, ident.1 = "differentiated", ident.2 = "progenitor", 
            features = "CLEC2D", logfc.threshold = 0)

```

## Add a second metagene classification without cell cycle programs
```{r}
nmf_score = readRDS(paste0("analysis/nmf/all_subtypes/new/nmf_score_updated", NameOfMetaprogramList,  ".rds"))
seurat_obj@meta.data$signaure_1_without_cc = nmf_score$signature_1
```

## Make pseudobulk
```{r}
metagene_program_names = sort(unique(nmf_score$signature_1))
pseudobulk = NULL
aggr_method = "mean"
for (metagene in metagene_program_names){
  cm = seurat_obj@assays$RNA@counts[,nmf_score$signature_1 == metagene & nmf_score$score_1 >= 1]
  if (aggr_method == "mean"){
    pseudobulk = cbind(pseudobulk, rowMeans(cm))
  }
  if (aggr_method == "median"){
    pseudobulk = cbind(pseudobulk, rowMedians(cm))
  }
}
colnames(pseudobulk) = metagene_program_names
rownames(pseudobulk) = rownames(seurat_obj@assays$RNA@counts)
saveRDS(pseudobulk, file=paste0(seurat_analysis_folder, 
                                "pseudobulk_cm_unique_only_", aggr_method, NameOfMetaprogramList,".RDS"))

## All cell types together- single pseudobulk for each sample
## Sum of counts across each sample
pseudobulk<- data.frame(gene=rownames(seurat_obj))
mySamples<- unique(seurat_obj$sample)
for (i in mySamples){
  seurat_tmp<- subset(seurat_obj, sample==i)
  cm_tmp<-seurat_tmp@assays$RNA@counts
  sumGene<- as.data.frame(rowSums(cm_tmp))
  colnames(sumGene)<- i
  sumGene$gene<- rownames(sumGene)
  pseudobulk<- merge(pseudobulk, sumGene, by="gene")
}

rownames(pseudobulk)<- pseudobulk$gene
pseudobulk<- pseudobulk[,-1]

save(pseudobulk, file="../../data/PseudobulkCounts/HemiBCOR_Pseudobulk.Robj")
```

## Metagene program specific genes as overall DEGs
```{r}
seurat_obj = SetIdent(seurat_obj, value = seurat_obj@meta.data$signaure_1_without_cc)

all_markers_nmf = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_nmf = all_markers[all_markers$p_val_adj < 0.05, ]
saveRDS(all_markers, paste0(seurat_analysis_folder, "deg.rds"))
##sapply(names(nmf_marker_genes_final), function(x) table(nmf_marker_genes_final[[x]] %in% gene_signatures[[x]]))
signature_genes = unique(all_markers$gene)
saveRDS(signature_genes, paste0(seurat_analysis_folder, "signature_genes.rds"))
```

## Manual annotation- combining NMF/graph based annotations
```{r}
## Make named list of nmf annotations
nmf_annot<-seurat_obj@meta.data$TopLigerFactor
names(nmf_annot)<- rownames(seurat_obj@meta.data)

## Make named vector of seurat annotations, but only for cluster 5
seurat_annot<- seurat_obj@meta.data$seurat_clusters
names(seurat_annot)<- rownames(seurat_obj@meta.data)
tmp<- seurat_annot[seurat_annot=="5"]
seurat_annot_5<- paste0("Cluster", tmp)
names(seurat_annot_5)<-names(tmp)

## Seurat annot takes priority over nmf- remove nmf annotation for cells with seurat annot
merged_annot<-nmf_annot[!(names(nmf_annot) %in% names(seurat_annot_5))]

## Combine annots
merged_annot<-c(merged_annot, seurat_annot_5)

merged_annot<- gsub("Cluster5", "OPC-like_Ribosomal",
                    gsub("factor10", "Neuronal",
                         gsub("factor13", "MES/AC-like",
                              gsub("factor14", "AC-like",
                                   gsub("factor3", "MES2",
                                        gsub("factor6", "Sphase", 
                                             gsub("factor7", "G2M",
                                                  gsub("factor8", "OPC-like",
                                                       gsub("factor9", "NPC2",merged_annot )))))))))

merged_annot<-merged_annot[order(names(merged_annot))]

seurat_obj@meta.data$ManualAnnotation<-merged_annot

DimPlot(object = seurat_obj, reduction="umap", group.by = "ManualAnnotation", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_ManualAnnotation.jpg", path=seurat_fig_folder, width=8, height=8)

```

```{r}
seurat_obj<- SetIdent(seurat_obj, value="ManualAnnotation" )
save(seurat_obj, file="analysis/seurat/seurat_obj_annot.Robj")
```

## Export marker genes for all cell types
```{r}
ManualCellAnnot_Markers<-FindAllMarkers(seurat_obj, min.pct=.2)
ManualCellAnnot_Markers_filt<-ManualCellAnnot_Markers[ManualCellAnnot_Markers$p_val_adj<0.05
                                                      & ManualCellAnnot_Markers$avg_logFC >0.2,]
save(ManualCellAnnot_Markers_filt, file="analysis/ManualCellAnnot_Markers.Robj")
```


## Differential expression of ligands between more progenitor (factors 3/8/9)
## And more differentiated (factors 10/13/14)
```{r}
prog_v_diff<- paste0("factor", c(3,8,9,10,13,14))
names(prog_v_diff)<- c(rep("progenitor", 3), rep("differentiated", 3))


seurat_prog_v_diff<- subset(seurat_obj, TopLigerFactor %in% prog_v_diff)
seurat_prog_v_diff@meta.data$prog_v_diff<- plyr::mapvalues(seurat_prog_v_diff@meta.data$TopLigerFactor,
                                                           from=prog_v_diff, to=names(prog_v_diff))
Idents(seurat_prog_v_diff)<- seurat_prog_v_diff@meta.data$prog_v_diff

ligand_DE<- FindMarkers(seurat_prog_v_diff, ident.1 = "differentiated", ident.2 = "progenitor", 
            features = c("CLEC2D", "CD274", "PDCD1LG2", "LGALS9"), logfc.threshold = 0,
            min.pct=0)
write.csv(ligand_DE, paste0(seurat_analysis_folder, "ligand_DE.csv"))

DimPlot(seurat_prog_v_diff, cols = c("grey", "red"), pt.size=2)
ggsave(paste0(seurat_fig_folder, "ProgenitorVsDiff.png"))
```

## Updated annotation: same annotations used for CellChat
## UMAP and sample breakdown
```{r}
seurat_obj$CellChatAnnot<- gsub("MES/AC-like", "AC-like",
                                    gsub("NPC2", "Neuronal-like",
                                    gsub("Neuronal", "Neuronal-like",
                                    gsub("OPC-like.*", "OPC-like",
                                         gsub("Sphase", "S-Phase",
                                         Idents(seurat_obj))))))

## Set colors so they're shared across all subtypes
load("../../../cellTypeColors.Robj")
cellTypes<- unique(seurat_obj$CellChatAnnot)
colorsUse<- c(colorsUse, `Neuronal-like`="turquoise3", MES2="mediumpurple")
save(colorsUse, file="../../../cellTypeColors.Robj")


DimPlot(object = seurat_obj, reduction="umap", group.by = "CellChatAnnot", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend() + scale_color_manual(values=colorsUse)
ggsave("UMAP_clustering_CellChatAnnot.jpg", path=seurat_fig_folder, width=8, height=8)


annotBreakdown<- as.data.frame(table(seurat_obj$sample, seurat_obj$CellChatAnnot))
colnames(annotBreakdown)<- c("Sample", "CellType", "NCells")

plot_list<-list()
for(i in unique(annotBreakdown$Sample)){
  df<- annotBreakdown[annotBreakdown$Sample==i,]
  p<-ggplot(df, aes(x="", y=NCells, fill=CellType))+
    geom_bar(stat="identity", color="white")+
    coord_polar("y", start=0)+
    theme_void()+
    ggtitle(i)+theme(plot.title = element_text(size=15, face="bold", hjust=0.5))+
    scale_fill_manual(values=colorsUse)
  plot_list[[i]]<- p
}
cowplot::plot_grid(plotlist=plot_list, ncol=4)
ggsave(paste0(seurat_fig_folder, "CellBreakdown.png"), width=12, height=3)
```

