---
title: "_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "C:/Users/jenna/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/BySubtype/Midline/Midline_All/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)


seurat_analysis_folder = "analysis/seurat/"
seurat_fig_folder = "figures/seurat/"
for (folder in c(seurat_analysis_folder, seurat_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

subtypes<-read.csv("../../NotUploaded/Subtypes.csv")
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}
```

## Load counts and metadata for initiating seurat object
```{r load-data}
input_dir<- "../../../Preprocessing/SplitBySubtype/analysis/"
load(paste0(input_dir, "midline_cm_list.Robj"))
samples = readRDS(paste0(input_dir,"midline_samples.rds"))
tumor_stat = readRDS(paste0(input_dir,"midline_call_tumor.rds"))
cm_list<-midline_cm_list

sum(colnames(cm_list$raw_data) == names(tumor_stat)); length(tumor_stat)
sum(names(samples)==names(tumor_stat))
```

## Subset cm to remove noise/normal samples
```{r subset-tumor-cells}
cm = cm_list$raw_data; rm(cm_list); gc()
cm = cm[,tumor_stat=="Malignant"]
samples = samples[tumor_stat=="Malignant"]

any(colnames(cm) != names(samples))
```

## Initiate seurat obj and add metadata 
```{r seurat_obj_meta_data}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", min.cells=0, min.genes=0, scale.factor=1E5, do.scale=F, do.center=T)


## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples

sample_df<- as.data.frame(samples); sample_df$cell<- rownames(sample_df)
subtype_df<- merge(sample_df, subtypes, by="samples")

seurat_obj@meta.data$subtype<- subtype_df$subtype
seurat_obj@meta.data$AdditionalSubtype<- subtype_df$Diagnosis
```

## OPTIONAL: plot highly variable genes
```{r}
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## PCA and PC selection
```{r PCA_selection}
## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)


```

## Harmony data integration 
```{r run_harmony}
library(harmony)
library(magrittr)
library(cowplot)

## Run Harmony
seurat_obj = RunHarmony(seurat_obj, "sample", theta = 2, 
                        max.iter.harmony = 50, plot_convergence = TRUE)

## Simple dim and vlnplot to examine integration 
options(repr.plot.height = 6, repr.plot.width = 12)

p1 <- DimPlot(object = seurat_obj, reduction = "pca", pt.size = .1, group.by = "sample")
p2<- DimPlot(object = seurat_obj, reduction = "harmony", pt.size = .1, group.by = "sample")
p3 <- VlnPlot(object = seurat_obj, features = "PC_1", pt.size = .1, group.by = "sample")
p4 <- VlnPlot(object = seurat_obj, features = "harmony_1", pt.size = .1, group.by = "sample")

plot_grid(p1,p2, p3, p4)
ggsave("Harmony_diagnostics_theta2.jpg", path=seurat_fig_folder, width=12, height=8)


## DimHeatmap
DimHeatmap(object = seurat_obj, reduction = "harmony", cells = 500, dims = 1:6, balanced = TRUE)
```

## Clustering and UMAP
```{r clustering_tSNE}
## Run UMAP and clustering using reduced dimensions from Harmony
## You may play around with resolution parameter to tune the clustering result
resolution = .8
reduction_type = "harmony"
dims = 20

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = reduction_type, dims = 1:dims)


seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = reduction_type,
                            dims = 1:dims,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = resolution)


r0.8<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5) + ggtitle("Resolution_0.8")
r0.5<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5)+ ggtitle("Resolution_0.5")
r1.0<-DimPlot(seurat_obj, reduction = "umap", pt.size = 2.5)+ ggtitle("Resolution_1")

r0.5+r0.8+r1.0 #resolution of 0.8 may be best
ggsave("UMAP_clustering_VaryResolution.jpg", path=seurat_fig_folder, width=15, height=6)

## Clustering results 
table(Idents(seurat_obj))


## Save the seurat object
save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```

## UMAP plot colored by clustering and samples
## Can color tSNE with other metadata features (e.g. primary/recurrence, treatment)
```{r UMAP_plots}
## By clustering
DimPlot(object = seurat_obj, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering.jpg", path=seurat_fig_folder, width=8, height=8)

## By samples
DimPlot(object = seurat_obj, reduction="umap", group.by = "sample", label = FALSE, 
        pt.size = 2, label.size = 5) + NoAxes() 
ggsave("UMAP_sample.jpg", path=seurat_fig_folder, width=10, height=8)


## By subtype
DimPlot(object = seurat_obj, reduction="umap", group.by = "subtype", label = FALSE, 
        pt.size = 2, label.size = 5) + NoAxes() 
ggsave("UMAP_subtype.jpg", path=seurat_fig_folder, width=10, height=8)

## By diagnosis
DimPlot(object = seurat_obj, reduction="umap", group.by = "AdditionalSubtype", label = FALSE, 
        pt.size = 2, label.size = 5) + NoAxes() 
ggsave("UMAP_Diagnosis.jpg", path=seurat_fig_folder, width=10, height=8)
```

## DEGs for each cluster 
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.5)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05 & all_markers$avg_logFC > 0.5, ]

## Combine clusters that are next to each other and have no DEGs- likely too similar to each other
## No longer merging clusters 0/1- plenty of DEGs using logfc>0.5 instead of logfc>1
seurat_obj<- AddMetaData(seurat_obj, seurat_obj@active.ident, col.name = "OriginalClusters")
newIdents<-gsub("0", "1", seurat_obj@active.ident); names(newIdents)<- names(seurat_obj@active.ident)
newIdents<-gsub("4", "1", seurat_obj@active.ident); names(newIdents)<- names(seurat_obj@active.ident)
seurat_obj<-SetIdent(seurat_obj, value=newIdents)
seurat_obj<-AddMetaData(seurat_obj, seurat_obj@active.ident, col.name = "MergedClusters")

## Rerun FindAllMarkers with new cluster idents- optionally lower thresholds for DEGs
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold =.25, min.pct = 0.50)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05 & all_markers$avg_logFC>1, ]
all_markers_filtered$cluster<-as.integer(all_markers_filtered$cluster)
all_markers_filtered<- all_markers_filtered[order(all_markers_filtered$cluster),]
all_markers_filtered$cluster<-as.character(all_markers_filtered$cluster)


## Subset the top 30 DEGs for each cluster
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC) %>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(seurat_analysis_folder, "all_markers.Robj"))
write.csv(top_marker_genes, file=paste0(seurat_analysis_folder, "Top30Markers_Theta1.csv"))

## Heatmap for initial inspection
DoHeatmap(seurat_obj, 
          features=all_markers_filtered$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs_theta1.jpg", path=seurat_fig_folder)
```

## Quick annotation by OPC/OC/AC marker genes 
```{r}
## OPC-like
AugmentPlot(FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
            cols = c("lightgrey", "red"),
            ncol = 3)) + ggtitle("OPC-like")
ggsave("OPC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## OC-like cells
AugmentPlot(FeaturePlot(seurat_obj,
            c("BCAS1", "SIRT2", "RGR", 
              "GPR17", "NFASC", "MBP",
              "TF", "MOG", "PLP1"),
            cols = c("lightgrey", "red"),
            ncol = 3))+ ggtitle("OC-like")
ggsave("OC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## AC-like cells 
AugmentPlot(FeaturePlot(seurat_obj,
            c("APOE", "AQP4", "AGT", 
              "SPARCL1","ATP1A2", "GFAP",
              "S100B", "ALDOC", "CLU"),
            cols = c("lightgrey", "red"),
            ncol = 3))+ ggtitle("AC-like")
ggsave("AC_markers.jpg", path=seurat_fig_folder, width=12, height=9)

## Neuronal-like cells
AugmentPlot(FeaturePlot(seurat_obj,
            c("STMN2", "STMN1", "STMN4", 
              "TUBB3", "INA", "ELAVL4",
              "SNAP25", "GABRG2", "ST18"),
            cols = c("lightgrey", "red"),
            ncol = 3))+ ggtitle("Neuronal-like")
ggsave("Neuron_markers.jpg", path=seurat_fig_folder, width=9, height=9)
```

## ORA test 
```{r over-representation-test}
library(clusterProfiler)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
allOE_genes <- as.character(gene_symbol_to_ensembl_id(rownames(seurat_obj@assays$RNA@counts),dataset)$ensembl_gene_id)


## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
go_result = list()
for (i in unique(all_markers_filtered$cluster)){
  sigOE_genes = as.character(gene_symbol_to_ensembl_id(subset(all_markers_filtered, cluster==i,
                                                              gene)[,], dataset)$ensembl_gene_id)
  go_result[[paste0("c", i)]] = go_analysis(sigOE_genes, allOE_genes)
} 

## Optional: remove biosynthetic/metabolic pathwasy
RemoveMetabolicBiosynthetic<- lapply(go_result, function(x){
  BS_MB<- x$ego[grep("metabolic", x$ego@result$Description, invert=TRUE),asis=TRUE]
  BS_MB<- BS_MB[grep("biosynthetic", BS_MB$Description, invert=TRUE),asis=TRUE]
})



## Plot top pathways for all clusters, display on one plot
clusters = paste0("c", sort(levels(seurat_obj@active.ident)))
allDotPlots_20_NoBCMB<-list()
for (c in 1:length(levels(seurat_obj@active.ident))){
  cluster=clusters[c]
  print(cluster)
  allDotPlots_20_NoBCMB[[c]]<-dotplot(RemoveMetabolicBiosynthetic[[cluster]], 
                                      showCategory=20, font.size = 12, title = cluster)
}

cowplot::plot_grid(plotlist = allDotPlots_20_NoBCMB, ncol=2)
ggsave("GO_OVA_AllClusters_top20_NoBiosyntheticMetabolicPathways.jpg", path=seurat_fig_folder, width=16, height=20)
save(go_result, file=paste0(seurat_analysis_folder, "go_result_metagene_DEGs_Clusters1_0Separate_LowerThreshold.Robj"))
```

# Add NMF results to the seurat object

## Map NMF scores to seurat object 
```{r add-NMF-score}
NameOfMetaprogramList<-"Merge4"

nmf_score = readRDS(paste0("analysis/nmf/all_subtypes/new/nmf_score", NameOfMetaprogramList, ".rds"))
seurat_obj = AddMetaData(seurat_obj, nmf_score)
load(paste0("analysis/nmf/all_subtypes/new/nmf_marker_genes_final", NameOfMetaprogramList, ".Robj"))

DimPlot(object = seurat_obj, group.by = "signature_1", label = TRUE, 
         pt.size = 3, label.size = 5, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_nmf_signature_Theta1",NameOfMetaprogramList, ".jpg"), 
       path=seurat_fig_folder, 
       width=14, height=12)

#DoHeatmap(seurat_obj, 
          #features=unlist(nmf_marker_genes_final)) + NoLegend()

#ggsave("nmf_marker_genes.jpg", path=seurat_fig_folder, width=12, height=8)
```

## Clusters 2-3: proportion of the 3 MUV084 programs that make up each cluster
```{r}
cluster2_3<- seurat_obj@meta.data[seurat_obj@meta.data$signature_1 %in% c("MUV084_nmf_2", "MUV084_nmf_3", "MUV084_nmf_5"),
                                  colnames(seurat_obj@meta.data) %in% c("signature_1", "MergedClusters")]

cluster2_3_table<-table(cluster2_3$MergedClusters,cluster2_3$signature_1)[c(5:6),]
df<-melt(cluster2_3_table) 
df<-split(df, f=df$Var1)

c2<-ggplot(df$`2`, aes(x="", y=value, fill=Var2))+
  geom_bar(width=1, stat="identity")+
  coord_polar("y", start=0)+
  ggtitle("Cluster2")
c3<-ggplot(df$`3`, aes(x="", y=value, fill=Var2))+
  geom_bar(width=1, stat="identity")+
  coord_polar("y", start=0)+
  ggtitle("Cluster3")

c2+c3
ggsave(paste0(seurat_fig_folder, "ProportionOfPrograms_Clusters2.3.png"))
```
## Plot subsets of cells to see patterns more clearly:
```{r}
#BT1516Programs<-c("BT1516_nmf_3","BT1516_nmf_2","BT1516_nmf_1", "BT1516_nmf_5")
#MUV065.66Programs<-c("MUV065_nmf_3", "MUV065_nmf_4" ,"MUV065_nmf_1", "MUV066_nmf_2", "MUV066_nmf_5","MUV066_nmf_1")
#MUV065.66Programs<- c("X065nmf1_066nmf5","X065nmf4_066nmf2", "MES2", "MUV065_nmf_3")
#BT1516Programs<- c("BT1516_nmf_1.3" ,  "BT1516_nmf_5","BT1516_nmf_2" )
#MUV065.66Programs<-c( "MUV066_nmf_1" ,    "MUV065_nmf_3" ,"X065nmf4_066nmf2" ,"X065nmf1_066nmf5" )

BT1516<-DimPlot(seurat_obj, reduction="umap", 
        cells = rownames(seurat_obj@meta.data[seurat_obj@meta.data$sample == "BT1516",]),
        group.by="signature_1",
        label=TRUE) 

MUV065.66<-DimPlot(seurat_obj, reduction="umap", 
        cells = rownames(seurat_obj@meta.data[seurat_obj@meta.data$sample %in% c("MUV065", "MUV066"),]),
        group.by="signature_1",
        label=TRUE) 

BT1516+ MUV065.66
ggsave(paste0(seurat_fig_folder, "SplitMUV6566.BT1516_Theta1", NameOfMetaprogramList, ".png"), height=4, width=12)
```

## Split up seurat object by sample, then re-run clustering and UMAP, then look where NMF programs fall
```{r}
## MUV065/MUV066:
MUV06566_cells<-rownames(seurat_obj@meta.data[seurat_obj@meta.data$sample=="MUV065" | seurat_obj@meta.data$sample=="MUV066",])
seurat_MUV06566<-subset(seurat_obj, cells= MUV06566_cells)

seurat_MUV06566<-FindVariableFeatures(seurat_MUV06566) %>% ScaleData(do.scale=FALSE, do.center=TRUE) %>% RunPCA(pcs.compute=100) %>%
                 RunUMAP(reduction="pca", dims=1:20) %>% FindNeighbors() %>% FindClusters(resolution=0.5)



## BT1516:
BT1516_cells<-rownames(seurat_obj@meta.data[seurat_obj@meta.data$sample=="BT1516",])
seurat_BT1516<-subset(seurat_obj, cells= BT1516_cells)

seurat_BT1516<-FindVariableFeatures(seurat_BT1516) %>% ScaleData(do.scale=FALSE, do.center=TRUE) %>% RunPCA(pcs.compute=100) %>%
                 RunUMAP(reduction="pca", dims=1:20) %>% FindNeighbors() %>% FindClusters(resolution=0.5)

## MUV084:
MUV084_cells<-rownames(seurat_obj@meta.data[seurat_obj@meta.data$sample=="MUV084",])
seurat_MUV084<-subset(seurat_obj, cells= MUV084_cells)

seurat_MUV084<-FindVariableFeatures(seurat_MUV084) %>% ScaleData(do.scale=FALSE, do.center=TRUE) %>% RunPCA(pcs.compute=100) %>%
                 RunUMAP(reduction="pca", dims=1:20) %>% FindNeighbors() %>% FindClusters(resolution=0.5)

## BT1478:
BT1478_cells<-rownames(seurat_obj@meta.data[seurat_obj@meta.data$sample=="BT1478",])
seurat_BT1478<-subset(seurat_obj, cells= BT1478_cells)

seurat_BT1478<-FindVariableFeatures(seurat_BT1478) %>% ScaleData(do.scale=FALSE, do.center=TRUE) %>% RunPCA(pcs.compute=100) %>%
                 RunUMAP(reduction="pca", dims=1:20) %>% FindNeighbors() %>% FindClusters(resolution=0.5)

(DimPlot(seurat_MUV06566, reduction="umap", group.by = "signature_1")+ ggtitle("MUV065_MUV066"))+
(DimPlot(seurat_BT1516, reduction="umap", group.by = "signature_1")+ ggtitle("BT1516"))+
(DimPlot(seurat_MUV084, reduction="umap", group.by = "signature_1")+ggtitle("MUV084"))+
(DimPlot(seurat_BT1478, reduction="umap", group.by = "signature_1")+ggtitle("BT1478"))
ggsave(paste0(seurat_fig_folder, "UMAPSplitBySample_Programs.jpg"), width=12, height=10)


(DimPlot(seurat_MUV06566, reduction="umap")+ ggtitle("MUV065_MUV066"))+
(DimPlot(seurat_BT1516, reduction="umap")+ ggtitle("BT1516"))+
(DimPlot(seurat_MUV084, reduction="umap")+ggtitle("MUV084"))+
(DimPlot(seurat_BT1478, reduction="umap")+ggtitle("BT1478"))
ggsave(paste0(seurat_fig_folder, "UMAPSplitBySample_Clusters.jpg"), width=12, height=10)

```


## Add a second metagene classification without cell cycle programs
```{r}
nmf_score = readRDS(paste0("analysis/nmf/all_subtypes/new/nmf_score_updated", NameOfMetaprogramList,  ".rds"))
seurat_obj@meta.data$signaure_1_without_cc = nmf_score$signature_1
```

## Make pseudobulk
```{r}
metagene_program_names = sort(unique(nmf_score$signature_1))
pseudobulk = NULL
aggr_method = "mean"
for (metagene in metagene_program_names){
  cm = seurat_obj@assays$RNA@counts[,nmf_score$signature_1 == metagene & nmf_score$score_1 >= 1]
  if (aggr_method == "mean"){
    pseudobulk = cbind(pseudobulk, rowMeans(cm))
  }
  if (aggr_method == "median"){
    pseudobulk = cbind(pseudobulk, rowMedians(cm))
  }
}
colnames(pseudobulk) = metagene_program_names
rownames(pseudobulk) = rownames(seurat_obj@assays$RNA@counts)
saveRDS(pseudobulk, file=paste0(seurat_analysis_folder, 
                                "pseudobulk_cm_unique_only_", aggr_method, NameOfMetaprogramList,".RDS"))
```

## Metagene program specific genes as overall DEGs
```{r}
seurat_obj = SetIdent(seurat_obj, value = seurat_obj@meta.data$signaure_1_without_cc)

all_markers_nmf = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_nmf = all_markers[all_markers$p_val_adj < 0.05, ]
saveRDS(all_markers, paste0(seurat_analysis_folder, "deg.rds"))
##sapply(names(nmf_marker_genes_final), function(x) table(nmf_marker_genes_final[[x]] %in% gene_signatures[[x]]))
signature_genes = unique(all_markers$gene)
saveRDS(signature_genes, paste0(seurat_analysis_folder, "signature_genes.rds"))
```

## Obsolete
## Pairwise metagene program specific genes 
```{r}
# logfc.threshold = 0.5
# test.use = "wilcox"
# min.pct = 0.3
# min.diff.pct = 0
# only.pos = TRUE 
# 
# gene_signatures = list()
# 
# print(Sys.time())
# for (target in unique(seurat_obj@meta.data$signature_1_without_cc)){
#   gene_signatures_tmp = rownames(seurat_obj@raw.data)
#   for (metagene in unique(seurat_obj@meta.data$signature_1_without_cc)){
#     tmp = NULL
#     if (metagene != target){
#       tmp = FindMarkers(seurat_obj, ident.1 = target, ident.2 = metagene, only.pos = only.pos,
#                         logfc.threshold = logfc.threshold, test.use = test.use,
#                         min.pct = min.pct, min.diff.pct = min.diff.pct)
#       tmp = tmp[tmp$p_val_adj < 0.05, ]
#       gene_signatures_tmp = intersect(gene_signatures_tmp, rownames(tmp))
#     }
#   }
#   gene_signatures[[target]] = gene_signatures_tmp
# }
# print(Sys.time())
```

