---
title: "Merge_AssessIntegration.Rmd"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Assess how well adult/ped myeloid cells integrate- do they form separate clusters?

## Approach:
Merge ped/adult (gbm and IDHmut) into a single seurat object, look at integration on UMAP

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(ggpubr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02b_Myeloid/03_CompareToAdult/")

integration<- ""
analysis_dir_tmp<- paste0(working_dir, "/analysis/Merge_AssessIntegration/", integration)
figure_dir_tmp<- paste0(working_dir, "/figures/Merge_AssessIntegration/", integration)
data_dir_tmp<- paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",integration)

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp)};if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in seurat objects
```{r}
cohort<- "pedOnly_nomuv63"
adult_withMUV<- FALSE ## adult cohort with just published datasetes (FALSE) or with our 4 adult samples (TRUE)

figure_dir<- paste0(figure_dir_tmp, cohort, "/", paste0("adult_withMUV.", adult_withMUV), "/")
analysis_dir<- paste0(analysis_dir_tmp, cohort, "/", paste0("adult_withMUV.", adult_withMUV), "/")
data_dir<- paste0(data_dir_tmp, cohort, "/")


if(!dir.exists(figure_dir)){dir.create(figure_dir)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

## ped seurat object
use_harmony_forVis<- ""
ped_seurat<-readRDS(paste0(data_dir, "myeloid_seurat_noDC.B",use_harmony_forVis, ".Rds"))

## adult (GBM and IDHmut; smartseq2)
if(adult_withMUV){
  adult_seurat<- readRDS(paste0("../../01_Preprocessing/03_PreprocessAdultDatasets/", 
                                "analysis/Preprocess_AdultMyeloid/",
                                "adult_myeloid_seurat_PublishedGBM.IDHmut.Filbin.Rds"))
}else{
  load(paste0("../../01_Preprocessing/03_PreprocessAdultDatasets/", 
                                "analysis/Preprocess_AdultMyeloid/", "GBMNeftel_IDHmut2018_myeloid_merged_seurat.Robj"))
  #adult_seurat<- gbm_idhmut_myeloid_seurat; rm(gbm_idhmut_myeloid_seurat)
  adult_meta$NiceSubtype<- adult_meta$TumorType
}

```

## Merge into single seurat object
```{r}
## Create merged metadata
ped_meta<- ped_seurat@meta.data; adult_meta<- adult_seurat@meta.data
ped_meta$Age<- "Ped"; adult_meta$Age<- "Adult"
ped_meta<- ped_meta[,colnames(ped_meta) %in% colnames(adult_meta)];adult_meta<- adult_meta[,colnames(adult_meta) %in% colnames(ped_meta)]
meta<- rbind(ped_meta, adult_meta)

## Merge count matrices
ped_cm<- ped_seurat@assays$RNA@counts; adult_cm<- adult_seurat@assays$RNA@counts
ped_cm<- ped_cm[rownames(adult_cm),]
sum(rownames(ped_cm) == rownames(adult_cm)); nrow(ped_cm); nrow(adult_cm)
cm<- cbind(ped_cm, adult_cm)
sum(colnames(cm)==rownames(meta)); ncol(cm); nrow(meta)

## Run seurat pipeline (for harmony integration analysis, run both with and without integration. Want to know if cells still don't cluster by age even without any integration)
if(integration=="HarmonyInt/"){
  library(harmony)
  seurat_obj<- RunFullSeurat(cm=cm,RunHarmony=TRUE,  samples = meta$sampleid)
  seurat_obj_noInt<- RunFullSeurat(cm=cm,RunHarmony=FALSE,  samples = meta$sampleid)
}else{
  seurat_obj<- RunFullSeurat_Immune(cm=cm, samples = meta$sampleid)
}

## Add on metadata
if(exists("seurat_obj_noInt")){
  seurat_obj_noInt<- AddMetaData(seurat_obj_noInt, meta[,!colnames(meta) %in% colnames(seurat_obj_noInt@meta.data)])
}
seurat_obj<- AddMetaData(seurat_obj, meta[,!colnames(meta) %in% colnames(seurat_obj@meta.data)])

if(exists("seurat_obj_noInt")){
  saveRDS(seurat_obj_noInt, file=paste0(analysis_dir, "myeloid_seurat_Ped.AdultGBM.IDHMut_merged_noInt.Rds"))
}

saveRDS(seurat_obj, file=paste0(analysis_dir, "myeloid_seurat_Ped.AdultGBM.IDHMut_merged.Rds"))
```

## Plot to assess integration
```{r}
seurat_obj<- readRDS(paste0(analysis_dir, "myeloid_seurat_Ped.AdultGBM.IDHMut_merged.Rds"))
figure_append<- ""

## reduce resolution- aim for 6-8 clusters to reflect heterogeneity found in NMF/Liger
if(integration=="HarmonyInt/"){
  seurat_obj<-FindClusters(seurat_obj, resolution = 0.5) ## .32 for without integration
}else{
  seurat_obj<- FindClusters(seurat_obj, resolution = .7, algorithm = 4, random.seed = 100)
}


## UMAP- age/clusters
DimPlot(seurat_obj, group.by = "Age")+
  scale_color_manual(values=c(Adult="navy", Ped="red"))
ggsave(paste0(figure_dir, "UMAP_ByAge", figure_append, ".png"), width=6, height=5)

DimPlot(seurat_obj, label=TRUE)+NoLegend()
ggsave(paste0(figure_dir, "UMAP_ByCluster_res0.5", figure_append, ".png"), width=6, height=6)

```

## composition of clusters by age
```{r}
cluster_bd<- as.data.frame(table(seurat_obj$Age, seurat_obj$seurat_clusters))
colnames(cluster_bd)<- c("Age", "Cluster", "NCells")

## Add total adult/ped frequency for references
total_bd<- as.data.frame(table(seurat_obj$Age))
colnames(total_bd)<- c("Age", "NCells")
total_bd$Cluster<- "Total"
total_bd<- total_bd[,c("Age", "Cluster", "NCells")]
cluster_bd<- rbind(total_bd, cluster_bd)

## Convert NCells to percentage
cluster_bd<- cluster_bd %>% group_by(Cluster) %>% mutate(CellPercentage=NCells/sum(NCells)) %>% as.data.frame()

## plot
#cluster_bd$Cluster<- factor(cluster_bd$Cluster, levels=c("Total", variable_levels))
ggplot(cluster_bd, aes(x=Cluster, y=CellPercentage, fill=Age))+
  geom_bar(position="stack", stat="identity")+
  scale_fill_manual(values=c(Adult="navy", Ped="red"))+
  theme_classic()+
  theme(axis.text = element_text(face="bold", size=12, color="black"),
        axis.title = element_text(face="bold", size=12,color="black"))+
  ylab("Percentage of adult/pediatric cells \nper cluster")+
  xlab("")+
  geom_hline(yintercept = cluster_bd[cluster_bd$Cluster=="Total" & cluster_bd$Age=="Ped", "CellPercentage"],
             linetype="dashed", size=1)
ggsave(paste0(figure_dir, "ProportionOfAdultPed_PerCluster.png"), width=6, height=4)

## alternative plot
cluster_bdByAge<- cluster_bd[cluster_bd$Cluster!="Total",] ## cluster_bd from previous chunk
cluster_bdByAge<- cluster_bdByAge %>% group_by(Age) %>% mutate(CellPercentage_ByAge=NCells/sum(NCells)) %>% as.data.frame()


#cluster_bdByAge$Cluster<- factor(as.character(cluster_bdByAge$Cluster), levels=c(variable_levels))
ggplot(cluster_bdByAge, aes(x=Cluster, y=CellPercentage_ByAge, fill=Age))+
  geom_bar(position="dodge", stat="identity")+
  scale_fill_manual(values=c(Adult="navy", Ped="red"))+
  theme_classic()+
  theme(axis.text = element_text(face="bold", size=12, color="black"),
        axis.title = element_text(face="bold", size=12,color="black"))+
  ylab("Percentage of adult/pediatric cells \nper cluster")+
  xlab("")
ggsave(paste0(figure_dir, "ProportionOfAdultPed_PerCluster_SideBySide.png"), width=6, height=4)


```
results: no age-specific clusters are formed. suggests the possibility that while there are differences in PROPORTION of cell states, there are not any totally age-specific programs. This informs how we compare programs in adult/ped- don't expect age-specific programs, so can either use the ped programs directly, or run NMF on ped/adult together and use those.

To help determine whether to use ped programs directly OR run NMF on ped adult together:
Look at where ped programs fall in this UMAP/on these clusters. How well do they line up?
```{r}
rank_use<- 7

## Subset to ped, keeping structure/clusters from merged ped/adult seurat
ped_seurat_merged<-subset(seurat_obj, Age=="Ped")

## Add on NMF programs
ped_seurat_merged$PedNMF_Program<- plyr::mapvalues(colnames(ped_seurat_merged), colnames(ped_seurat), 
                                                   as.character(ped_seurat@meta.data[[paste0("NMF_rank", rank_use, "_annotByAdult10X")]]))

## Plot- UMAP + breakdown of each cluster by program
colors_use<-c("orange", "turquoise3", "red", "olivedrab4",  "purple", "navy" ,"grey22")
names(colors_use)<- unique(ped_seurat_merged$PedNMF_Program)
p1<- DimPlot(ped_seurat_merged, group.by = "PedNMF_Program")+scale_color_manual(values=colors_use)

bd<- as.data.frame(table(ped_seurat_merged$PedNMF_Program, ped_seurat_merged$seurat_clusters))
colnames(bd)<- c("Program", "Cluster", "Ncells")
bd <- bd %>% group_by(Cluster) %>% mutate(PerCells=Ncells/sum(Ncells)) %>% as.data.frame()
p2<- ggplot(bd, aes(x=Cluster, y=PerCells, fill=Program))+
  geom_bar(stat="identity", position="stack")+
  scale_fill_manual(values=colors_use)
p1+p2
ggsave(paste0(figure_dir, "PedMyeloid_ByPedNMFProgram.png"), width=12, height=5)

```


