---
title: "CompareProgramProportions.Rmd"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Compare proportion of cells assigned to ped NMF programs in adult vs ped myeloid cells

## Approach:
Score adult for ped myeloid programs. Assign max, then compare proportions


```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(ggpubr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02b_Myeloid/03_CompareToAdult/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/CompareProgramProportions_PedNMF/")
figure_dir_tmp<- paste0(working_dir, "/figures/CompareProgramProportions_PedNMF/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp)};if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in seurat object + marker genes
Seurat object from Merge_AssessIntegration (merged adult/ped myeloid)
```{r}
cohort<- "pedOnly_nomuv63"
figure_dir<- paste0(figure_dir_tmp, cohort, "/"); analysis_dir<- paste0(analysis_dir_tmp, cohort, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

ped_rank<- 6
seurat_obj<- readRDS(paste0("analysis/Merge_AssessIntegration/", cohort, "/myeloid_seurat_Ped.AdultGBM.IDHMut_merged.Rds"))

## ped marker genes
ped_markers<- read.csv(paste0("../01a_NMF/analysis/de.novoNMF/", cohort,  "/top_DE_genes_10Xannotated_NMF_rank", ped_rank, ".csv"))
ped_markers<- ped_markers[,!grepl("X", colnames(ped_markers))]
```



## Score ped/adult myeloid cells for ped/adult programs, add to seurat
```{r}
## Norm/Center
cm_list<-NormCenter(seurat_obj@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data + 1))

## create marker list
marker_list<- split(ped_markers, f=ped_markers$cluster); marker_list<- lapply(marker_list, function(x){x$gene})

## score for marker list
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, marker_list)))

## Determine max program/score
scores$MaxProgram<- apply(scores, 1, function(x){names(x)[which.max(x)]})
scores$MaxScore<- apply(scores[,colnames(scores) != "MaxProgram"], 1, function(x){x[which.max(x)]})

## Add to seurat object
seurat_obj<- AddMetaData(seurat_obj, scores)
```

## Plot: max program, spread of scores in ped vs adult
```{r}
colors_use<-c("orange", "turquoise3", "red", "olivedrab4",  "purple", "navy" )
names(colors_use)<- unique(seurat_obj$MaxProgram)

## UMAP of max program
p1<-DimPlot(seurat_obj, group.by = "MaxProgram")+scale_color_manual(values=colors_use)
seurat_obj$MaxProgram_HighScore<- ifelse(seurat_obj$MaxScore>0.25, seurat_obj$MaxProgram, "ScoreTooLow")
p2<-DimPlot(seurat_obj, group.by = "MaxProgram_HighScore")+scale_color_manual(values=c(colors_use, ScoreTooLow="grey"))
p1+p2
ggsave(paste0(figure_dir,"UMAP_MaxPedProgram.png" ), width=12, height=5)

## Spread of scores in ped vs adult
df<- seurat_obj@meta.data
ggplot(df, aes(x=MaxScore, fill=Age))+
  geom_histogram()+
  scale_fill_manual(values=c(Adult="navy", Ped="red"))
ggsave(paste0(figure_dir, "Histogram_scoresForMaxProgram_adult.vs.ped.png"), width=5, height=4)

## Proportion of scores <.1, <.2, <.5, <1 in adult vs ped
df_byGroup<- df %>% group_by(Age) %>% summarise(LessThan_0.1=sum(MaxScore<0.1),
                                          LessThan_0.2=sum(MaxScore<0.2),
                                          LessThan_0.5=sum(MaxScore<0.5),
                                          LessThan_1=sum(MaxScore<1),
                                          GreaterThan_1=sum(MaxScore>1)) %>% as.data.frame()
df_byGroup<- as.data.frame(t(df_byGroup))
colnames(df_byGroup)<- df_byGroup[1,]; df_byGroup<- df_byGroup[-1,]
df_byGroup$Group<- rownames(df_byGroup)
df_byGroup<- melt(df_byGroup, id="Group"); colnames(df_byGroup)<- c("Group", "Age", "nCells")
df_byGroup$nCells<- as.numeric(df_byGroup$nCells)
df_byGroup<- df_byGroup %>% group_by(Age) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()

df_byGroup$Group<- factor(df_byGroup$Group, 
                          levels=c("GreaterThan_1", "LessThan_1", "LessThan_0.5", "LessThan_0.2", "LessThan_0.1"))
ggplot(df_byGroup, aes(x=Age, y=perCells, fill=Group))+
  geom_bar(stat="identity", position="stack")+
    scale_fill_manual(values=c(GreaterThan_1= "green",
                             LessThan_1="pink",
                             LessThan_0.1="red",
                             LessThan_0.2="purple",
                             LessThan_0.5="magenta"))
ggsave(paste0(figure_dir, "Grouping_scoresForMaxProgram_adult.vs.ped.png"), width=5, height=4)
```
Results: scoring for ped programs is too disparate between adult/ped- while most adult cells score above 1, their overall spread of scores is much lower than ped. Likely too big of a confounder. Will run NMF on merged adult/ped instead.

