---
title: "CompareProgramProportions.Rmd"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Compare proportion of cells assigned to ped NMF programs in adult vs ped myeloid cells

## Approach:
Score adult for ped myeloid programs. Assign max, then compare proportions


```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(ggpubr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02b_Myeloid/03_CompareToAdult/")

integration<- "HarmonyInt/"
analysis_dir_tmp<- paste0(working_dir, "/analysis/Macrophage.Microglia/", integration)
figure_dir_tmp<- paste0(working_dir, "/figures/Macrophage.Microglia/", integration)
data_dir_tmp<- paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp)};if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in seurat objects + marker genes
Seurat object from Merge_AssessIntegration (merged adult/ped myeloid)
```{r}
cohort<- "pedOnly_nomuv63"
figure_dir<- paste0(figure_dir_tmp, cohort, "/"); analysis_dir<- paste0(analysis_dir_tmp, cohort, "/")
data_dir<- paste0(data_dir_tmp, cohort, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

## adult
load(paste0("../../01_Preprocessing/03_PreprocessAdultDatasets/", 
                                "analysis/Preprocess_AdultMyeloid/", "GBMNeftel_IDHmut2018_myeloid_merged_seurat.Robj"))
adult_seurat<- gbm_idhmut_myeloid_seurat; rm(gbm_idhmut_myeloid_seurat)

## ped
ped_seurat<-readRDS(paste0(data_dir, "myeloid_seurat_noDC.B.Rds"))

## macrophage/myeloid marker genes
gene_set_dir<- "../../../../../../Marker_genes/"
microglia_genes = read.table(paste0(gene_set_dir, "Microglia_markers.txt"))
macrophage_genes = read.table(paste0(gene_set_dir, "Macrophage_markers.txt"))
tam_genes = list("microglia" = microglia_genes$V1, "macrophage" = macrophage_genes$V1)

## motam/mgtam
gene_set_dir<- "../../../../../../Marker_genes/"
gene_sets = read.csv(paste0(gene_set_dir, "Nature2021Movahedi_MoTAM.vs.MgTAM.csv"))
gene_list<- split(gene_sets, f=gene_sets$cluster)
gene_list<- lapply(gene_list, function(x){
  tmp<- x[order(x$avg_logFC,decreasing = TRUE),]
  return(tmp$gene[1:25])
})

## set gene set to use downstream
tam_genes<- tam_genes
analysis_name<- "Macrophage.Microglia"

```

## Score adult
```{r}
## Norm and center counts
cm = GetAssayData(adult_seurat, slot = "counts")
cm_norm = log2(cm/10+1)
cm_mean = log2(rowMeans(cm)+1)
cm_center = cm_norm-rowMeans(cm_norm)


## Score for microglia/macrophage markers
tam_score = scoreNmfGenes(cm_center, cm_mean, tam_genes)
tam_score = data.frame(t(tam_score))
```

## Add score and annotation to adult seurat obj
```{r}
adult_seurat = AddMetaData(adult_seurat, tam_score$microglia, "microglia_score")
adult_seurat = AddMetaData(adult_seurat, tam_score$macrophage, "macrophage_score")

#adult_seurat = AddMetaData(adult_seurat, tam_score$mgTAM, col.name="Mg_TAM")
#adult_seurat = AddMetaData(adult_seurat, tam_score$moTAM, "Mo_TAM")

##  assign as macrophage/microglia if score is high enough
mg_annot = ifelse(adult_seurat$microglia_score < 0 & adult_seurat$macrophage_score < 0,
                  "undetermined",
                  ifelse(adult_seurat$microglia_score > adult_seurat$macrophage_score,
                        "microglia", "macrophage"))
adult_seurat = AddMetaData(adult_seurat, mg_annot, "MM_annot")

## alternative: assign microglia/macrophage if scores are high
score_thresh<- 0.5
adult_seurat$AltMM_annot<- ifelse(adult_seurat$microglia_score>score_thresh & 
                                    adult_seurat$macrophage_score >score_thresh, "HighBoth",
                                ifelse(adult_seurat$microglia_score>score_thresh, "Microglia",
                                       ifelse(adult_seurat$macrophage_score>score_thresh, "Macrophage", "NotHighForEither")))

save(adult_seurat, file=paste0("../../01_Preprocessing/03_PreprocessAdultDatasets/", 
                                "analysis/Preprocess_AdultMyeloid/", "GBMNeftel_IDHmut2018_myeloid_merged_seurat.Robj"))
```

## Compare proportion of microglia/macrophage in adult vs ped
```{r}
all_scores<- data.frame(macrophage_score=c(ped_seurat$macrophage_score, adult_seurat$macrophage_score),
                        microglia_score=c(ped_seurat$microglia_score, adult_seurat$microglia_score),
                        MM_annot=c(ped_seurat$MM_annot, adult_seurat$MM_annot),
                        AltMM_annot=c(ped_seurat$AltMM_annot, adult_seurat$AltMM_annot),
                        sample=c(ped_seurat$sampleid, adult_seurat$sampleid),
                        Age=c(rep("Ped", ncol(ped_seurat)), rep("Adult", ncol(adult_seurat))))

## Plot adult vs ped, all samples grouped together
  ## Create df with number of macrophages/microglia for adult and ped
  df<- as.data.frame(table(all_scores$Age, all_scores$MM_annot))
  colnames(df)<- c("Age", "CellType", "NCells")
  
  ## Normalize by cohort size
  df<- df %>% group_by(Age) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()
  
  ggplot(df, aes(x=Age, y=PerCells, fill=CellType))+
    geom_bar(stat="identity", position="stack")+
    theme_classic()+
    scale_fill_manual(values=c(microglia="red", macrophage="blue", undetermined="grey"))
  ggsave(paste0(figure_dir, "Adult.vs.Ped_nMacrophage.Microglia.png"), width=3, height=3)

## Plot adult vs ped, by sample
  
  ## Create df with number of macrophages/microglia for adult and ped
  df<- as.data.frame(table(all_scores$sample, all_scores$MM_annot))
  colnames(df)<- c("Sample", "CellType", "NCells")
  df$Age<- plyr::mapvalues(df$Sample, all_scores$sample, all_scores$Age, warn_missing = FALSE)
  
  ## Normalize by cohort size
  df<- df %>% group_by(Sample) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()
  
  ## order by proportion of macrophages
  tmp<- df[df$CellType=="macrophage",]
  tmp<- tmp[order(tmp$PerCells),]
  tmp$Order<- 1:nrow(tmp)
  df$Order<- as.numeric(as.character(plyr::mapvalues(df$Sample, tmp$Sample, tmp$Order)))
  
  df$CellType<- factor(df$CellType, levels=rev(c("macrophage", "microglia", "undetermined")))
  ggplot(df, aes(x=reorder(Sample, Order), y=PerCells, fill=CellType))+
    geom_bar(stat="identity", position="stack")+
    theme_classic()+
    scale_fill_manual(values=c(microglia="red", macrophage="blue", undetermined="grey"))+
    facet_grid(cols = vars(Age), scale="free_x", space="free_x")+
    theme(axis.text.x = element_text(angle=45, hjust=1))+
    xlab("")
  ggsave(paste0(figure_dir, "Adult.vs.Ped_nMacrophage.Microglia_bySample.png"), width=8, height=3)
  
## By scores
  ggplot(all_scores, aes(x=macrophage_score, y=microglia_score, color=Age))+
    geom_point(aes(alpha=0.9))+
    theme_classic()+
    scale_color_manual(values=c(Ped="red", Adult="blue"))+
    geom_hline(yintercept = 0)+geom_vline(xintercept = 0)+
    facet_grid(cols = vars(Age))
  ggsave(paste0(figure_dir, "Adult.vs.Ped_MM_scores_dotplot.png"), width=6, height=3)
  
## By change in scores
  all_scores$mm_change<- all_scores$macrophage_score-all_scores$microglia_score
  ggplot(all_scores, aes(x=mm_change, fill=Age))+
    geom_histogram()+
    theme_classic()+
    scale_fill_manual(values=c(Ped="red", Adult="blue"))+
    facet_grid(cols = vars(Age))+
    geom_vline(xintercept = 0)+
    xlab("negative = high microglia\npositive = high macrophage")
   ggsave(paste0(figure_dir, "Adult.vs.Ped_MM_scores_histogram.png"), width=6, height=3)
```

## Plot heatmap of macrophage/microglia scores (based on 2017 Science IDHmut paper, figure 5C)
```{r}
## Make dataframe of scores based on ped/adult seurat objects
all_scores<- data.frame(macrophage_score=c(ped_seurat$macrophage_score, adult_seurat$macrophage_score),
                        microglia_score=c(ped_seurat$microglia_score, adult_seurat$microglia_score),
                        sample=c(ped_seurat$sampleid, adult_seurat$sampleid),
                        Age=c(rep("Ped", ncol(ped_seurat)), rep("Adult", ncol(adult_seurat))),
                        Subtype=c(ped_seurat$NiceSubtype, adult_seurat$TumorType))

## Add on score represetning both macrophage/microglia score
all_scores$MM_score<- all_scores$macrophage_score-all_scores$microglia_score

## Set number of bins to use
nbins<- 50
bins<- seq(from=min(all_scores$MM_score), to=max(all_scores$MM_score), length.out=nbins)

## For each subtype, calculate % of cells that fall into each bin
all_bin_scores<- lapply(unique(all_scores$Subtype), function(x){
  print(x)
  scores<- all_scores[all_scores$Subtype==x,]
  nCells<- nrow(scores)
  
  ## Cycle through each score bin, determine how many cells in this subtype fall into each bin
  bin_scores<- lapply(1:length(bins), function(y){
    current_bin<- bins[y]
    if(y==1){previous_bin<- -100}else{
      previous_bin<- bins[y-1]
    }
    sum(scores$MM_score>previous_bin & scores$MM_score<=current_bin)/nCells *100
  })
  
  ## Merge % of cells for each bin into single df
  names(bin_scores)<- paste0("bin", 1:25)
  bin_scores<- as.data.frame.matrix(do.call("rbind", bin_scores))
  
  ## Add in column with actual bins, subtype
  bin_scores$MM_score<- bins
  bin_scores$Subtype<- x
  colnames(bin_scores)<- c("PercentCellsInBin", "Bin", "Subtype")
  print(sum(bin_scores$PercentCellsInBin))
  return(bin_scores)
  })
all_bins_df<- do.call("rbind", all_bin_scores)

## Add on Age
all_bins_df$Age<- plyr::mapvalues(all_bins_df$Subtype, all_scores$Subtype, all_scores$Age,
                                  warn_missing = FALSE)
## colors
color_bin<- round(length(unique(all_bins_df$PercentCellsInBin))/5, 0)

yellow_red <- colorRampPalette(c("yellow", "red"))

## order subtypes
all_bins_df$Subtype<- factor(all_bins_df$Subtype, levels=rev(c("IDHmut", "GBM",
                                                               "Midline-H3K27M",
                                                           "Midline-HistoneWT", "Ependymoma", "Hemispheric-BCOR",
                                                           "Hemispheric-H3G34R/V",
                                                           "Hemispheric-HistoneWT")))
ggplot(all_bins_df, aes(x=Bin, y=Subtype))+
 #geom_tile(aes(fill = as.factor(PercentCellsInBin)), colour = "white")+
  geom_raster(aes(fill = as.factor(PercentCellsInBin)), interpolate = TRUE)+
  #scale_fill_gradient2(low="white",  mid="yellow", high="red", midpoint = 5)+
  scale_fill_manual(values=c(rep("white", color_bin),yellow_red(color_bin *4)))+
  #facet_grid(rows=vars(Age), scale="free_y", space="free_y")+
  theme_classic()+
  theme(legend.position = "none",
        axis.text = element_text(color="black", face="bold"))+
  xlab("Microglia-like    - to -    Macrophage-like")+
  ylab("")
ggsave(paste0(figure_dir, "Heatmap_macrophage.to.microglia_gradient.png"), width=7, height=2.5)
```


