---
title: "inferCNV_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and functions
```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/Immune",
                     "/TumorCellAnnotation/Preprocessing/inferCNV/WithNewSamples/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)

inferCNV_input_folder<- "analysis/inferCNV/"

inferCNV_analysis_folder = "analysis/inferCNV/seurat/"
inferCNV_fig_folder = "figures/inferCNV/seurat/"
for (folder in c(inferCNV_analysis_folder, inferCNV_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```

## Load data
```{r load-data}
## Load cm, samples, and qc data
cm = readRDS(paste0(inferCNV_input_folder, "cm_exp_ctrl.rds"))
samples = readRDS(paste0(inferCNV_input_folder, "samples_exp_ctrl.rds"))
any(colnames(cm) != rownames(samples))
qc_filtered_data<- readRDS("../../QC/WithNewSamples/analysis/qc_filtered_data.rds")

## If seurat_obj already generated 
seurat_filename = paste0(inferCNV_analysis_folder,
                         "seurat_obj.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}

## Load cnv status inferred by inferCNV
cnv_stat = readRDS(paste0(inferCNV_input_folder, "HC/", "call_cnv_w_ctrl.rds"))
obs_ref = readRDS(paste0(inferCNV_input_folder, "obs_ref.rds"))

```

## Preprocessing with Seurat
```{r seurat_clustering}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, 
                                    project="", 
                                    min.cells=0, 
                                    min.genes=0,
                                    scale.factor=1E5, 
                                    do.scale=F, 
                                    do.center=T)

## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples
seurat_obj@meta.data$type = obs_ref
seurat_obj@meta.data$cnv = cnv_stat

## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 50,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)


## Clustering
numPC = 28
resolution = .6
seurat_obj <- FindNeighbors(object=seurat_obj,
                            dims=1:numPC,
                            reduction="pca")
seurat_obj<- FindClusters(object=seurat_obj, 
                          resolution=resolution)


## UMAP
table(seurat_obj@active.ident)
seurat_obj <- RunUMAP(
  object = seurat_obj, 
  dims= 1:numPC, 
  do.fast = TRUE
)

save(seurat_obj, file = paste0(inferCNV_analysis_folder, "seurat_obj.Robj"))
```

## UMAP plots
```{r tsne_plots}
## Colored by clustering
DimPlot(object = seurat_obj, label = T, pt.size = 2, label.size = 8) + NoLegend()
ggsave("UMAP_clustering.jpg", path=inferCNV_fig_folder, width = 8, height = 6, dpi = 300)

## Colored by samples
DimPlot(object = seurat_obj, label = T, pt.size = 2, label.size = 4, group.by = "sample") + NoLegend()
ggsave("UMAP_sample.jpg", path=inferCNV_fig_folder, width = 8, height = 6, dpi = 300)

## Colored by obs vs ref
DimPlot(object = seurat_obj, label = T, pt.size = 2, label.size = 8, group.by = "type") + NoLegend()
ggsave("UMAP_type.jpg", path=inferCNV_fig_folder, width = 8, height = 5, dpi = 300)

## Colored by CNV status
DimPlot(object = seurat_obj, label = T, pt.size = 2, label.size = 8, group.by = "cnv") + NoLegend()
ggsave("UMAP_cnv.jpg", path=inferCNV_fig_folder, width = 8, height = 8, dpi = 300)
```


## Take a second look at cluster16- clusters close to oligodendrocytes, with od markers, but classified as having a cnv
## Subset to cells in cluster16 and export for inferCNV
```{r}
## Subset to cells of interest
cluster16<- seurat_obj@meta.data[seurat_obj@meta.data$seurat_clusters=="16",]

## subset cm_list and samples to cells of interest
load("../../QC/WithNewSamples/analysis/cm_list.Robj")
cm_list_cluster16<- lapply(cm_list, function(x){x[,colnames(x) %in% rownames(cluster16)]})
samples_cluster16<- cluster16$sample; names(samples_cluster16)<-rownames(cluster16)
sum(colnames(cm_list_cluster16$raw_data)==names(samples_cluster16))

## Save for input into inferCNV if desired
save(cm_list_cluster16, file="../../QC/WithNewSamples/analysis/cm_list_cluster16.Robj")
saveRDS(samples_cluster16, file="../../QC/WithNewSamples/analysis/samples_cluster17.rds")

## Zoom in on cluster 6/16
seurat_c6.16<- subset(seurat_obj, seurat_clusters %in% c("6", "16"))

mapped_colors = col_vector[10:(length(unique(samples))+10)]
p1<- DimPlot(seurat_c6.16, group.by = "sample", cols = mapped_colors)
p2<- DimPlot(seurat_c6.16, cols=mapped_colors)
p1+p2
ggsave(paste0(inferCNV_fig_folder, "ZoomOnClusters6.16_16PotentiallyNormal.png"),
       height=5, width=10)

## Only MUV79 and MUV86 are substantially present in cluster 17
## MUV79 clusters clearly away from normal cluster- combined with clear CNV, def malig
## MUV86 is a little unclear- only 26 cells
## For now: consider cluster 17 "normal", EXCEPT for MUV79 cells
```

## There are 77 cells that mismatch in tumor calling between this analysis and original
## Ran inferCNV on these separately ("MismatchCells")
## 6 clustered with normal cells- may be normal
## The rest were unclear- split into 2 clusters (2/4)
## Cluster 4 had a couple of mg cells- could be normal?
## Cluster 2 expressed normal OC markers- could be normal?
## Where do they fall here?
```{r}
current_seurat<- seurat_obj
load("MismatchCells/analysis/inferCNV/seurat/seurat_obj.Robj")
mismatch_seurat<- seurat_obj

## Annotate current seurat with clusters from mismatch seurat
mismatch_seurat<- subset(mismatch_seurat, type=="obs")
toAnnotate<- paste0("MismatchCluster", as.character(mismatch_seurat$seurat_clusters))
names(toAnnotate)<- rownames(mismatch_seurat@meta.data)

notMismatch<- subset(current_seurat, 
                     cells=colnames(current_seurat)[!colnames(current_seurat) %in% names(toAnnotate)])
notMismatch_vec<-rep("NotMismatch", length(colnames(notMismatch)))
names(notMismatch_vec)<-colnames(notMismatch)

all_toAnnotate<-c(toAnnotate, notMismatch_vec)

current_seurat$Mismatch<- plyr::mapvalues(rownames(current_seurat@meta.data),
                                          from=names(all_toAnnotate),
                                          to=all_toAnnotate)

DimPlot(current_seurat, group.by="Mismatch")
ggsave(paste0(inferCNV_fig_folder, "CellsThatMismatchBetweenOriginalCurrent_Cluster.png"))
```


## DEGs
```{r differential_gene_expressions}
#seurat_obj <- SetIdent(seurat_obj, ident.use = seurat_obj@meta.data$seurat_clustering)
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05, ]
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC)%>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(inferCNV_analysis_folder, "all_markers.Robj"))

## Heatmap
DoHeatmap(seurat_obj, 
          features=all_markers_filtered$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs.jpg", path=inferCNV_fig_folder)
```

## Expresssions of common normal cell marker
```{r deg_for_normal_cells}
FeaturePlot(seurat_obj, 
            features = c("MBP", "PLP1", "TF", ## Oligodendrocyte
                         "OLIG1", "PDGFRA", "APOD", ## OPC
                         "CTGF", "IFITM1", "MGP", ## Endothelial cells,
                         "CD14", "CD3E", "CD4" ## Immune cells
                         ), #CD45 marker for all immune cells-distinguish neural from immune
            cols = c("lightgrey", "red"), ncol=3,
            slot="data")
ggsave("normal_marker_genes_umap_data.jpg", path=inferCNV_fig_folder, width=9, height=12)
```

## Map gene counts to UMAP
```{r}
FeaturePlot(seurat_obj, "nFeature_RNA", cols = c("lightgrey", "purple"))
ggsave("nFeature_RNA_UMAP.jpg", path=inferCNV_fig_folder, width=9, height=12)
```

## Map mito counts to UMAp
```{r}
mito.genes <- grep(pattern = "^MT-", x = rownames(seurat_obj@assays$RNA@data), 
                   value = TRUE)
percent.mito <- Matrix::colSums(seurat_obj@assays$RNA@counts[mito.genes,])/
  Matrix::colSums(seurat_obj@assays$RNA@counts)
seurat_obj <- AddMetaData(seurat_obj, 
                          metadata = percent.mito, 
                          col.name = "percent_mito")
FeaturePlot(seurat_obj, "percent_mito", cols = c("lightgrey", "red"))
ggsave("mito_genes.jpg", path=inferCNV_fig_folder, width=6, height=6)
```

## Map alignment rate onto tSNE
```{r}
qc_raw_data = readRDS("../../QC/WithNewSamples/analysis/qc_raw_data.rds")
seurat_obj@meta.data$alignment_rate = qc_raw_data[colnames(seurat_obj), 
                                                  "align_rate"]
tmp = is.na(seurat_obj@meta.data$alignment_rate)
seurat_obj@meta.data$alignment_rate[tmp] = 0.5
FeaturePlot(seurat_obj, "alignment_rate", cols = c("lightgrey", "red"))
ggsave("align_rate.jpg", path=inferCNV_fig_folder, width=6, height=6)
```

## OPTIONAL: Additional QC on CNV
```{r}
## cluster vs cnv: which cluster has the highest % of cells without CNV 
tmp = table(seurat_obj@meta.data$seurat_clusters, seurat_obj@meta.data$cnv)
tmp2 = round(tmp/rowSums(tmp),2)
tmp2[order(tmp2[,1], decreasing = T),]

## sample vs clustering: which sample belongs to normal clusters
## 16 set as malig here- a few cells removed downstream in annotation from histone mut midline samples since they cluster clearly away and are normal. 
## These cells are virtually entirely from the "upper half" of cluster 16- the lower half appears to really be malignant
## For simplicity's sake: cluster 16 considered "malignant" here, and the effectively the upper half removed as "normal" downstream
normal_clust = c( "6", "23") 
tmp3 = table(seurat_obj@meta.data$sample, seurat_obj@meta.data$seurat_cluster)
tmp4 = round(tmp3/rowSums(tmp3),2)
tmp4[,normal_clust]
rowSums(tmp4[,normal_clust])

## sample vs cnv: which sample has the highest % fo CNV negative cells 
tmp5 = table(seurat_obj@meta.data$sample, seurat_obj@meta.data$cnv)
tmp6 = round(tmp5/rowSums(tmp5),2)

## Compare % of cells for each sample in normal clusters vs CNV negative cells
rowSums(tmp4[,normal_clust])
tmp6[,1]
```

## Check intrapatient batch effect
```{r}
## subset seurat object to just obs samples, then add plate metadata
seurat_onlyObs<-subset(seurat_obj, type=="obs" )
plate<-qc_filtered_data$plate; names(plate)<- rownames(qc_filtered_data)
seurat_onlyObs@meta.data$plate<- plate


## Iteratively plot the spread of cells across all plates for each sample individually
plots<-list()
for (sample in unique(samples)){
  plots[[sample]]<-DimPlot(seurat_onlyObs, group.by="plate", reduction="umap",
         cells = colnames(seurat_onlyObs)[seurat_onlyObs@meta.data$sample == sample], 
         label.size = 8, pt.size = 2, label = T)
}

## Save each plot
for (i in 1:length(plots)){
  filename=paste0(inferCNV_fig_folder, "IntrpatientBatchEffect_", unique(samples)[i], ".jpg")
  jpeg(filename,width=600,height=600)
  print(plots[[i]])
  dev.off()
}
```

## Malignant vs non-malignant cell calling 
```{r call_tumor}
#############Strategy###############
## All cells in normal clusters and without CNV = Normal
## All cells in samples without cnv are Malignant
## All cells NOT in normal clusters and with CNV = Malignant 
## The rest = Low-quality 
####################################
normal_clusters = normal_clust
samples_without_cnv = c()
call_tumor <- function(x){
  ## cells in normal cluster are normal cells 
  if (seurat_obj@meta.data$seurat_cluster[x] %in% normal_clusters &
      !seurat_obj@meta.data$cnv[x]){
    return("Normal")
  }else if (seurat_obj@meta.data$sample[x] %in% samples_without_cnv){
    return("Malignant")
  ## All others classified based on having CNV or not 
  }else if (!(seurat_obj@meta.data$seurat_cluster[x] %in% normal_clusters) &
            seurat_obj@meta.data$cnv[x]){
     return("Malignant") 
  }else{
    return("Noise")
  }
}
tumor_stat_w_ctrl = sapply(seq(length(colnames(seurat_obj))), call_tumor)
names(tumor_stat_w_ctrl) = colnames(seurat_obj)
any(names(cnv_stat) != names(tumor_stat_w_ctrl))


## Remove normal controls 
normalSamples<- seurat_obj@meta.data[seurat_obj$sample %in% c("mg", "od"),]
tumor_stat = tumor_stat_w_ctrl[!names(tumor_stat_w_ctrl) %in% rownames(normalSamples)]
saveRDS(tumor_stat, file = paste0(inferCNV_analysis_folder, "call_tumor.rds"))

## UMAP colored by malignancy
seurat_obj@meta.data$tumor = tumor_stat_w_ctrl
DimPlot(object = seurat_obj, label = F, pt.size = 1, 
        label.size = 8, group.by = "tumor", reduction="umap",
        cols = c("lightgrey", gg_color_hue(2))) + NoAxes()

ggsave("UMAP_tumor.jpg", path=inferCNV_fig_folder, width = 8, height = 8, dpi = 300)
save(seurat_obj, file = paste0(inferCNV_analysis_folder, "seurat_obj.Robj"))
```

## Export CNV calling info for each cell. Include:
1. Whether CNV was detected
2. If it clustered with a "normal" cluster
3. Final annotation (malig/normal/noise)
```{r}
df<- seurat_obj@meta.data[,c("cnv", "type", "sample", "seurat_clusters")]
df<- df[df$type=="obs",]
df$ClusteredWithNormal<- ifelse(df$seurat_clusters %in% normal_clust,
                                yes="Yes", no="No")
df$Classification<- ifelse(df$cnv==TRUE & df$ClusteredWithNormal=="No",
                           yes="Malignant",
                           no=ifelse(df$cnv==FALSE & df$ClusteredWithNormal=="Yes",
                                     yes="Normal", no="Noise"))
saveRDS(df, file=paste0(inferCNV_analysis_folder, "FullCNVCallingInfo.Rds"))
```


## Annotation 
## Adapt this code based on normal cells identified in your dataset
```{r}
annotation = seurat_obj@meta.data$tumor
cluster = seurat_obj@meta.data$seurat_cluster


## Cells in normal cluster and labeled Normal are assigned corresponding cell type
## Update corresponding cluster number for each normal cell type 
annot_fun <- function(x){
    if (cluster[x] == 6 & annotation[x] == "Normal"){
        return("Oligodendrocyte")
    } else if ((cluster[x] == 21) & annotation[x] == "Normal"){
        return("ImmuneCell")
    } else{
        return(annotation[x])
    }
}
annotation = sapply(1:length(annotation), annot_fun)

## Added to seurat_obj and plot UMAP (removing normal controls and low-quality ones)
seurat_obj@meta.data$annotation = annotation 

## Remove normal controls 
annot_stat = annotation[(seurat_obj@meta.data$sample != "mg" & 
                         seurat_obj@meta.data$sample != "od")]
saveRDS(annot_stat, file = paste0(inferCNV_analysis_folder, "annotation.rds"))
```

## UMAP with only tumor samples and high quality cells
```{r}
## Generate df for only tumor samples
df = cbind.data.frame(seurat_obj@reductions$umap@cell.embeddings,
                      sample=as.character(seurat_obj@meta.data$sample),
                      annotation=as.character(seurat_obj@meta.data$annotation))

df = df[seurat_obj@meta.data$type == "obs", ]

## Remove low quality cells
df = df[df$annotation != "Noise",]
df$annotation = ifelse(df$annotation == "Malignant", df$sample, df$annotation)

mapped_colors = col_vector[1:(length(unique(samples))+2)]
names(mapped_colors)<- c(unique(samples), "Oligodendrocyte", "ImmuneCell")

ggplot(df, aes(x=UMAP_1, y=UMAP_2, color=annotation)) + 
    geom_point(size=2, alpha=0.5) +
    ggtitle("") + xlab("UMAP-1") + ylab("UMAP-2") + labs(color="") +
    scale_color_manual(values=mapped_colors) +
    theme(axis.title = element_text(face="bold", size=24),
          axis.text.x = element_text(size=24),
          axis.text.y = element_text(size=24),
          legend.text = element_text(size=16))


ggsave("UMAP_annotation_final.png", path = inferCNV_fig_folder, width = 10, height = 8)
```


## Compare results between origianl analysis and analysis with new samples
```{r}
original_call_tumor<- readRDS("../analysis/inferCNV/seurat/call_tumor.rds")
original_call_tumor<- original_call_tumor[order(names(original_call_tumor))]

## remove new samples from this analysis
new_call_tumor<- tumor_stat[grep(paste(c("BT1873", "BT1857"), collapse="|"), 
                                 names(tumor_stat), invert=TRUE)]
new_call_tumor<- new_call_tumor[order(names(new_call_tumor))]

sum(names(original_call_tumor) == names(new_call_tumor)); length(new_call_tumor);length(original_call_tumor)


## Make df with annotations from original/new
merge_call_tumor<- as.data.frame(original_call_tumor)
merge_call_tumor$new_call_tumor<- plyr::mapvalues(rownames(merge_call_tumor),
                                                  from=names(new_call_tumor),
                                                  to=new_call_tumor)
merge_call_tumor$sample<- plyr::mapvalues(rownames(merge_call_tumor),
                                          from=names(samples),
                                          to=samples, warn_missing = FALSE)

CellsThatDiffer<- merge_call_tumor[merge_call_tumor$original_call_tumor!=merge_call_tumor$new_call_tumor,] 


## Are the cells that differ here the result of different inferCNV clustering or seurat?
## Read in comparison of old/new inferCNV HC results
load("analysis/cnv_match_withWithoutNewSamples.Robj")

sum(rownames(CellsThatDiffer) %in% rownames(cnv_match_WithWithoutNewSamples))
nrow(CellsThatDiffer)

## all differences due to inferCNV HC differences, not to seurat clustering
## Majority (80%) were called malignant in original, noise or normal in the new analysis
```





## Compare tumor_stat results with all samples vs all samples except for BT1857/BT1873
## These samples were added later
## CHecking that results are highly similar- if so, can just keep "old" tumor_stat for all samples except for BT1857/BT1873
```{r}
newTumor_stat<- readRDS(paste0(inferCNV_analysis_folder, "call_tumor.rds"))
oldTumor_stat<- readRDS("../analysis/inferCNV/seurat/call_tumor.rds")
samples = readRDS(paste0(inferCNV_input_folder, "samples_exp_ctrl.rds"))

## Remove new samples (any cells in new not in old)
newTumor_stat<- newTumor_stat[names(newTumor_stat) %in% names(oldTumor_stat)]

## COmbine into 1 df
all_tumorStat<- as.data.frame(newTumor_stat)
all_tumorStat<- merge(all_tumorStat, as.data.frame(oldTumor_stat), by=0, all.x=TRUE)
all_tumorStat$sample<- plyr::mapvalues(all_tumorStat$Row.names,
                                       from=names(samples), to=samples,
                                       warn_missing = FALSE)
all_tumorStat$Match<- ifelse(all_tumorStat$newTumor_stat==all_tumorStat$oldTumor_stat,
                             yes="Match", no="NoMatch")

## Add column denoting how sample ID changed. Use for plotting
all_tumorStat$Change<- paste0(all_tumorStat$oldTumor_stat,
                              " --> ", all_tumorStat$newTumor_stat)
df<- as.data.frame(table(all_tumorStat$Change, all_tumorStat$sample))
colnames(df)<- c("ChangeInTumorCall","Sample", "NumberCells")
ggplot(df, aes(x=reorder(ChangeInTumorCall,NumberCells, mean), 
               y=NumberCells, 
               fill=Sample))+
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black", face="bold"))+
  xlab("")
ggsave(paste0(inferCNV_fig_folder, "Original_vs_WithNewSamples_MaligCellCalling.png"),
       width=7, height=5)

## Remove cells whose annotations matched, plot
tmp<- all_tumorStat[all_tumorStat$Match=="NoMatch",]
df2<- as.data.frame(table(tmp$Change, tmp$sample))
colnames(df2)<- c("ChangeInTumorCall","Sample", "NumberCells")
df2<- df2[df2$NumberCells>0,]
ggplot(df2, aes(x=reorder(ChangeInTumorCall,NumberCells, mean), 
               y=NumberCells, 
               fill=Sample))+
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black", face="bold"))+
  xlab("")
ggsave(paste0(inferCNV_fig_folder, "Original_vs_WithNewSamples_MaligCellCalling_MismatchCells.png"),
       width=7, height=5)

## Export cells that mismatch (noise-> malig or malig-> noise) for inferCNV
## Will run inferCNV on just these to determine if they have CNVs
mismatchCells<- all_tumorStat[all_tumorStat$Change %in% c("Malignant --> Noise",
                                                          "Noise --> Malignant"),]
save(mismatchCells, file=paste0(inferCNV_analysis_folder,
                                "CellsThatMismatchBetweenOriginalNewAnalysis.Robj"))
```

