---
title: "ImmuneRegev_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/TumorCellAnnotation/AllSubtypesTogether/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)


seurat_analysis_folder = "analysis/seurat/"
seurat_fig_folder = "figures/seurat/"
for (folder in c(seurat_analysis_folder, seurat_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}

subtypes<-read.csv("../../Subtypes.csv")

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
subtype<-"All_New"
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj_", subtype, ".Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}

```


## Load counts and metadata for initiating seurat object
```{r load-data}
load("../Preprocessing/QC/analysis/cm_list.Robj")
tumor_stat = readRDS("../Preprocessing/inferCNV/analysis/inferCNV/seurat/call_tumor.rds")
samples = readRDS("../Preprocessing/QC/analysis/samples.rds")

sum(colnames(cm_list$raw_data) == names(tumor_stat)); length(tumor_stat)
sum(names(samples)==names(tumor_stat))
```

## Subset cm to remove noise/normal samples
```{r subset-tumor-cells}
cm = cm_list$raw_data; rm(cm_list); gc()
cm = cm[,tumor_stat=="Malignant"]
samples = samples[tumor_stat=="Malignant"]

any(colnames(cm) != names(samples))
```


## Initiate seurat obj and add metadata 
```{r seurat_obj_meta_data}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", min.cells=0, min.genes=0, scale.factor=1E5, do.scale=F, do.center=T)

## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples

sample_df<- as.data.frame(samples); sample_df$cell<- rownames(sample_df)
subtype_df<- merge(sample_df, subtypes, by="samples")

seurat_obj@meta.data$subtype<- subtype_df$subtype
seurat_obj@meta.data$AdditionalSubtype<- subtype_df$Diagnosis #for epn only
```

## OPTIONAL: plot highly variable genes
```{r}
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## PCA and PC selection
```{r PCA_selection}
## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)


```

## Harmony data integration 
```{r run_harmony}
library(harmony)
library(magrittr)
library(cowplot)

## Run Harmony
seurat_obj = RunHarmony(seurat_obj, "sample", theta = 3, 
                        max.iter.harmony = 50, plot_convergence = TRUE)

## Simple dim and vlnplot to examine integration 
options(repr.plot.height = 6, repr.plot.width = 12)

p1 <- DimPlot(object = seurat_obj, reduction = "pca", pt.size = .1, group.by = "sample")
p2<- DimPlot(object = seurat_obj, reduction = "harmony", pt.size = .1, group.by = "sample")
p3 <- VlnPlot(object = seurat_obj, features = "PC_1", pt.size = .1, group.by = "sample")
p4 <- VlnPlot(object = seurat_obj, features = "harmony_1", pt.size = .1, group.by = "sample")

plot_grid(p1,p2, p3, p4)
ggsave("Harmony_diagnostics.jpg", path=seurat_fig_folder, width=12, height=8)


## DimHeatmap
DimHeatmap(object = seurat_obj, reduction = "harmony", cells = 500, dims = 1:6, balanced = TRUE)
```

## Clustering and UMAP
```{r clustering_tSNE}
## Run UMAP and clustering using reduced dimensions from Harmony
## You may play around with resolution parameter to tune the clustering result
resolution = .8
reduction_type = "harmony"
dims = 20

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = reduction_type, dims = 1:dims)


seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = reduction_type,
                            dims = 1:dims,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = resolution)


## Clustering results 
table(Idents(seurat_obj))


## Save the seurat object
save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj", ".Robj"))
```

## UMAP plot colored by clustering and samples
## Can color tSNE with other metadata features (e.g. primary/recurrence, treatment)
```{r UMAP_plots}
## By clustering
DimPlot(object = seurat_obj, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 3, label.size = 8) + NoAxes() + NoLegend()
ggsave(paste0("UMAP_clustering.jpg"), path=seurat_fig_folder, width=8, height=8)

## By samples
DimPlot(object = seurat_obj, reduction="umap", group.by = "sample", label = FALSE, 
        pt.size = 2, label.size = 5) + NoAxes() 
ggsave(paste0("UMAP_sample.jpg"), path=seurat_fig_folder, width=10, height=8)

## By subtype
DimPlot(object = seurat_obj, reduction="umap", group.by = "subtype", label =FALSE, 
        pt.size = 2, label.size = 5) + NoAxes()
ggsave(paste0("UMAP_subtype.jpg"), path=seurat_fig_folder, width=10, height=8)

## By additional subtype 
DimPlot(object = seurat_obj, reduction="umap", group.by = "AdditionalSubtype", 
        label = TRUE, pt.size = 3, label.size = 5) + NoAxes() 
ggsave(paste0("UMAP_AdditionalSubtype.jpg"), path=seurat_fig_folder, width=10, height=8)
```

## DEGs for each cluster 
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.5)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05 & all_markers$avg_logFC > 0.5, ]



## Subset the top 30 DEGs for each cluster
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC) %>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(seurat_analysis_folder, 
                                       "all_markers_", subtype, ".Robj"))
write.csv(top_marker_genes, file=paste0(seurat_analysis_folder, 
                                        "Top30Markers_", subtype, ".csv"))

## Heatmap for initial inspection
#DoHeatmap(seurat_obj, 
          #features=all_markers_filtered$gene) + NoLegend() + NoAxes()
#ggsave(paste0("Heatmap_all_degs_", subtype, ".jpg"), path=seurat_fig_folder)
```

## Comparing between clusters
```{r}
hemi_cluster1v2<-FindMarkers(seurat_obj, ident.1 = 1, ident.2 = 2, 
                             min.pct = .50, logfc.threshold = .25)
hemi_cluster1v2<- hemi_cluster1v2[hemi_cluster1v2$p_val_adj<0.05 ,]

midline_cluster5v6<-FindMarkers(seurat_obj, ident.1 = 5, ident.2 = 6, 
                             min.pct = .5, logfc.threshold = .25)
midline_cluster5v6<- midline_cluster5v6[midline_cluster5v6$p_val_adj<0.05,]

midline_cluster1v2<-FindMarkers(seurat_obj, ident.1 = 1, ident.2 = 2, 
                             min.pct = .5, logfc.threshold = .25)
midline_cluster1v2<- midline_cluster5v6[midline_cluster1v2$p_val_adj<0.05,]
```

## Quick annotation by OPC/OC/AC marker genes 
```{r}
## OPC-like
AugmentPlot(FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
            cols = c("lightgrey", "red"),
            ncol = 3)) + ggtitle("OPC-like")
ggsave(paste0(subtype, "_OPC_markers.jpg"), path=seurat_fig_folder, width = 9, height=9)

## OC-like cells
AugmentPlot(FeaturePlot(seurat_obj,
            c("BCAS1", "SIRT2", "RGR", 
              "GPR17", "NFASC", "MBP",
              "TF", "MOG", "PLP1"),
            cols = c("lightgrey", "red"),
            ncol = 3))+ ggtitle("OC-like")
ggsave(paste0(subtype, "_OC_markers.jpg"), path=seurat_fig_folder, width = 9, height=9)

## AC-like cells 
AugmentPlot(FeaturePlot(seurat_obj,
            c("APOE", "AQP4", "AGT", 
              "SPARCL1","ATP1A2", "GFAP",
              "S100B", "ALDOC", "CLU"),
            cols = c("lightgrey", "red"),
            ncol = 3))+ ggtitle("AC-like")
ggsave(paste0(subtype, "_AC_markers.jpg"), path=seurat_fig_folder, width=9, height=9)

## Neuronal-like cells
AugmentPlot(FeaturePlot(seurat_obj,
            c("STMN2", "STMN1", "STMN4", 
              "TUBB3", "INA", "ELAVL4",
              "SNAP25", "GABRG2", "ST18"),
            cols = c("lightgrey", "red"),
            ncol = 3))+ ggtitle("Neuronal-like")
ggsave(paste0(subtype, "_Neuron_markers.jpg"), path=seurat_fig_folder, width=9, height=9)

##CLEC2D
FeaturePlot(seurat_obj,
            c("CLEC2D", "KLRB1"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("CLEC2D_KLRB1_markers.jpg", path=seurat_fig_folder, width=9, height=4.5)
```

## ORA test 
```{r over-representation-test}
library(clusterProfiler)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
allOE_genes <- as.character(gene_symbol_to_ensembl_id(rownames(seurat_obj@assays$RNA@counts),dataset)$ensembl_gene_id)


## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
go_result = list()
for (i in unique(all_markers_filtered$cluster)){
  sigOE_genes = as.character(gene_symbol_to_ensembl_id(subset(all_markers_filtered, cluster==i,
                                                              gene)[,], dataset)$ensembl_gene_id)
  go_result[[paste0("c", i)]] = go_analysis(sigOE_genes, allOE_genes)
} 



## Plot top pathways for all clusters, display on one plot
clusters = paste0("c", sort(levels(seurat_obj@active.ident)))
allDotPlots_20<-list()
for (c in 1:length(levels(seurat_obj@active.ident))){
  cluster=clusters[c]
  print(cluster)
  if (nrow(go_result[[cluster]]$cluster_summary)>0){
    print(cluster)
    allDotPlots_20[[c]]<-dotplot(go_result[[cluster]]$ego, 
                                showCategory=20, 
                                font.size = 15, 
                                title = cluster, 
                                label_format=10) +
      scale_y_discrete(labels=function(x) str_wrap(x, width=30))+
      theme(plot.title = element_text(hjust = 0.5, face = "bold", 
                                      color="black", size = 28),
            axis.title = element_text(face = "bold", color="black"), 
            axis.text.x = element_text(angle = 45, hjust = 1, color="black"),
            axis.text.y = element_text( color="black",face = "bold"))
  }
  
}

allDotPlots_20<- allDotPlots_20[!(unlist(lapply(allDotPlots_20, function(x){is.null(x)})))]
cowplot::plot_grid(plotlist = allDotPlots_20, ncol=2)
ggsave(paste0(subtype, "_Resolution1.2_GO_OVA_AllClusters_top20.jpg"), path=seurat_fig_folder, width=15, height=35)
save(go_result, file=paste0(seurat_analysis_folder, subtype, "_go_result.Robj"))
```

# Add NMF results to the seurat object

## Map NMF scores to seurat object 
```{r add-NMF-score}
NameOfMetaprogramList<-"MidlineAllAgesSeparate_Merge8"

nmf_score = readRDS(paste0("analysis/nmf/all_subtypes/new/nmf_score_", NameOfMetaprogramList, ".rds"))
seurat_obj = AddMetaData(seurat_obj, nmf_score)
load(paste0("analysis/nmf/all_subtypes/new/nmf_marker_genes_final_", NameOfMetaprogramList, ".Robj"))

DimPlot(object = seurat_obj, group.by = "signature_1", label = TRUE, 
         pt.size = 3, label.size = 10, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_nmf_signature_", subtype, "_", NameOfMetaprogramList, ".jpg"), 
       path=seurat_fig_folder, 
       width=12, height=25)

#DoHeatmap(seurat_obj, 
          #features=unlist(nmf_marker_genes_final)) + NoLegend()

#ggsave("nmf_marker_genes.jpg", path=seurat_fig_folder, width=12, height=8)
```

## For hemispheric: comparing NMF4/5
```{r}
## Subset to cluster 0, then interactively plot NMF4 vs NMF5
AllMeta<-seurat_obj@meta.data[seurat_obj@meta.data$seurat_clusters=="0",]
AllMeta$Cell<-rownames(AllMeta)
p<-ggplot(AllMeta, aes(x=NMF4, y=NMF5)) + geom_point()
ggsave(paste0(seurat_fig_folder, "SpreadofNMF4_vs_NMF5_Cluster0Only.png"))

RunShiny(AllMeta,p,"NMF4", "NMF5")

## Read in results from interactive plot- high NMF4 high NMF5
highNMF4<- read.delim(paste0(seurat_analysis_folder, "Cluster0_HighNMF4.txt"))
highNMF5<- read.delim(paste0(seurat_analysis_folder, "Cluster0_HighNMF5.txt"))

## Create df listing each cell as high NMF4, high NMF5, or both
MidBoth<- AllMeta[!(AllMeta$Cell %in% c(highNMF4$Cell, highNMF5$Cell)),]
NMF4_NMF5_group<-c(rep("HighNMF4", nrow(highNMF4)),
                   rep("HighNMF5", nrow(highNMF5)),
                   rep("MidBoth", nrow(MidBoth)))
names(NMF4_NMF5_group)<- c(highNMF4$Cell, highNMF5$Cell, MidBoth$Cell)

## Add to cluster 0 seurat object and dimplot
cluster0_seurat<-subset(seurat_obj, subset=seurat_clusters=="0")
cluster0_seurat@meta.data$NMF4_NMF5_group<- NMF4_NMF5_group

DimPlot(cluster0_seurat, group.by = "NMF4_NMF5_group")
```


## Add a second metagene classification without cell cycle programs
```{r}
nmf_score = readRDS(paste0("analysis/nmf/all_subtypes/new/nmf_score_updated", NameOfMetaprogramList,  ".rds"))
seurat_obj@meta.data$signaure_1_without_cc = nmf_score$signature_1
```

## Make pseudobulk
```{r}
metagene_program_names = sort(unique(nmf_score$signature_1))
pseudobulk = NULL
aggr_method = "median"
for (metagene in metagene_program_names){
  cm = seurat_obj@assays$RNA@counts[,nmf_score$signature_1 == metagene & nmf_score$score_1 >= 1]
  if (aggr_method == "mean"){
    pseudobulk = cbind(pseudobulk, rowMeans(cm))
  }
  if (aggr_method == "median"){
    pseudobulk = cbind(pseudobulk, rowMedians(cm))
  }
}
colnames(pseudobulk) = metagene_program_names
rownames(pseudobulk) = rownames(seurat_obj@assays$RNA@counts)
saveRDS(pseudobulk, file=paste0(seurat_analysis_folder, 
                                "pseudobulk_cm_unique_only_", aggr_method,
                                NameOfMetaprogramList,".RDS"))

## All cell types together- single pseudobulk for each sample
## Sum of counts across each sample
pseudobulk<- data.frame(gene=rownames(seurat_obj))
mySamples<- unique(seurat_obj$sample)
for (i in mySamples){
  print(i)
  seurat_tmp<- subset(seurat_obj, sample==i)
  cm_tmp<-seurat_tmp@assays$RNA@counts
  sumGene<- as.data.frame(rowSums(cm_tmp))
  colnames(sumGene)<- i
  sumGene$gene<- rownames(sumGene)
  pseudobulk<- merge(pseudobulk, sumGene, by="gene")
}

rownames(pseudobulk)<- pseudobulk$gene
pseudobulk<- pseudobulk[,-1]

save(pseudobulk, file="data/PseudobulkCounts/AllSamples_Pseudobulk.Robj")
```

## Create final cluster annotation- editing NMF annotation as needed
```{r}
## hemispheric
OriginalAnnotation<-seurat_obj@meta.data$signature_1
names(OriginalAnnotation)<-colnames(seurat_obj)

cluster6_cells<-rownames(seurat_obj@meta.data[seurat_obj@meta.data$seurat_clusters=="6",])
cluster6_annotation<-rep("MES1", length(cluster6_cells))
names(cluster6_annotation)<-cluster6_cells

EditedAnnotation<-OriginalAnnotation[!(names(OriginalAnnotation) %in% cluster6_cells)]
EditedAnnotation<- c(EditedAnnotation, cluster6_annotation)
EditedAnnotation<- EditedAnnotation[order(names(EditedAnnotation))]

seurat_obj@meta.data$EditedAnnotation<- EditedAnnotation

DimPlot(object = seurat_obj, reduction="umap", group.by = "EditedAnnotation", 
        label = TRUE, pt.size = 3, label.size = 8) + NoAxes() + NoLegend()
ggsave(paste0(subtype, "_UMAP_FinalEditedAnnotation.jpg"), path=seurat_fig_folder, width=10, height=8)

## midline
OriginalAnnotation<- seurat_obj@meta.data$signature_1
names(OriginalAnnotation)<-colnames(seurat_obj)

cluster1_cells<-rownames(seurat_obj@meta.data[seurat_obj@meta.data$seurat_clusters=="1",])
cluster1_annotation<-rep("OPC-like", length(cluster1_cells))
names(cluster1_annotation)<-cluster1_cells

cluster2_cells<-rownames(seurat_obj@meta.data[seurat_obj@meta.data$seurat_clusters=="2",])
cluster2_annotation<-rep("OPC-like_Ribosomal", length(cluster2_cells))
names(cluster2_annotation)<-cluster2_cells

EditedAnnotation<-OriginalAnnotation[!(names(OriginalAnnotation) %in% names(cluster1_annotation))] 
EditedAnnotation<-EditedAnnotation[!(names(EditedAnnotation) %in% names(cluster2_annotation))]
EditedAnnotation<- c(EditedAnnotation, cluster1_annotation, cluster2_annotation)
EditedAnnotation<- EditedAnnotation[order(names(EditedAnnotation))]

seurat_obj@meta.data$EditedAnnotation<- EditedAnnotation

DimPlot(object = seurat_obj, reduction="umap", group.by = "EditedAnnotation", 
        label = TRUE, pt.size = 3, label.size = 8) + NoAxes() 
ggsave(paste0(subtype, "_UMAP_FinalEditedAnnotation.jpg"), path=seurat_fig_folder, width=10, height=8)
```




## Rename clusters, plot and export
```{r}
## Ependymoma
seurat_obj<- RenameIdents(seurat_obj, 
                          `0` = "Ependymal", 
                          `1` = "ST-RELASpecific_Neuronal-like", 
                          `2` = "SP-Variable_OPC-like",
                          `3` = "AC-like",
                          `4` = "ST-Midline_RibosomeActive",
                          `5` = "Metabolic",
                          `6` = "Cycling",
                          `7` = "ImmuneReactive",
                          `8` = "OPC-like")
DimPlot(object = seurat_obj, label = TRUE,
         pt.size = 3, label.size = 9, reduction="umap", cols = col_vector[10:28]) +
  NoAxes()
ggsave(paste0("UMAP_Annotated_",subtype, ".jpg"), path=seurat_fig_folder, 
       width=12, height=10)

## Hemispheric
seurat_obj<- RenameIdents(seurat_obj, 
                          `0` = "AC_like", 
                          `1` = "MES2", 
                          `2` = "MES2_Alt",
                          `3` = "NPC2",
                          `4` = "CellCycle",
                          `5` = "OC_like",
                          `6` = "MES1",
                          `7` = "OC_like")
DimPlot(object = seurat_obj, label = TRUE,
         pt.size = 3, label.size = 10, reduction="umap", cols = col_vector[10:28]) +
  NoAxes()
ggsave(paste0("UMAP_Annotated_",subtype, ".jpg"), path=seurat_fig_folder, 
       width=14, height=12)

## Midline
seurat_obj<- RenameIdents(seurat_obj, 
                          `0` = "Immune", 
                          `1` = "Cluster1", 
                          `2` = "OPC-like",
                          `3` = "CellCycle",
                          `4` = "AC-like",
                          `5` = "Neuronal-like",
                          `6` = "OPC-like_Alt")
DimPlot(object = seurat_obj, label = TRUE,
         pt.size = 3, label.size = 8, reduction="umap", cols = col_vector[10:28]) +
  NoAxes()
ggsave(paste0("UMAP_Annotated_",subtype, ".jpg"), path=seurat_fig_folder, 
       width=14, height=12)
```


## Compare NMF annotations vs seurat annotations
```{r}
SeuratAnnotation<- gsub( 
                          "0" , "AC_like", gsub(
                          "1" , "MES2",gsub( 
                          "2" ,"MES2_Alt",gsub( 
                          "3" , "NPC2",gsub( 
                          "4" , "CellCycle",gsub( 
                          "5" , "OC_like",gsub( 
                          "6" , "MES1",gsub( 
                          "7" , "OC_like", seurat_obj@meta.data$seurat_clusters))))))))
table(seurat_obj@meta.data$signature_1, seurat_obj@meta.data$SeuratAnnotation)
```

## ORA with DIPG and GBM genesets
## Compare with metaprograms from previously defined metaprograms from DIPG/GBM
```{r}
subtype<-"epn_C67Removed"
load(paste0(seurat_analysis_folder, "all_markers_", subtype, ".Robj"))
all_markers_list<-split(all_markers_filtered, all_markers_filtered$cluster )
all_markers_list<-lapply(all_markers_list, function(x){x$gene})
names(all_markers_list)<-paste0("Cluster", names(all_markers_list))

epn_genesets<- read.csv("TableS2_epn2020Paper_Metaprograms_AllMerged.csv")
epn_genesets<- as.list(epn_genesets)
hemi_previousPrograms<-readRDS(paste0(nmf_analysis_folder, "hemisphericPreviousPrograms_2019_OC.rds"))
midline_previousprograms<-readRDS(paste0(nmf_analysis_folder, "midlinePreviousPrograms_OC.AC.OPC.MES1.rds"))

geneset<-epn_genesets


tmp2 = sapply(all_markers_list, function(x){
    sapply(geneset, function(y){
        fisher_test(x, y, unlist(union(geneset, all_markers_list)))
    })
})
tmp2 = data.frame(tmp2)
tmp3<- apply(tmp2, 2, function(x) {p.adjust(x) < 0.1})
tmp4<-apply(tmp2, 2, function(x) {p.adjust(x)})

write.table(tmp4, paste0(seurat_analysis_folder,
                         "ScoreSeuratClustersUsingPreviousNMF_padj", 
                       subtype, ".txt"))

```

## Metagene program specific genes as overall DEGs
```{r}
seurat_obj = SetIdent(seurat_obj, value = seurat_obj@meta.data$signaure_1_without_cc)

all_markers_nmf = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_nmf = all_markers_nmf[all_markers_nmf$p_val_adj < 0.05, ]
saveRDS(all_markers_nmf, paste0(seurat_analysis_folder, NameOfMetaprogramList,"_deg.rds"))
##sapply(names(nmf_marker_genes_final), function(x) table(nmf_marker_genes_final[[x]] %in% gene_signatures[[x]]))
signature_genes = unique(all_markers_nmf$gene)
saveRDS(signature_genes, paste0(seurat_analysis_folder, NameOfMetaprogramList, 
                                "_signature_genes.rds"))
```

## Obsolete
## Pairwise metagene program specific genes 
```{r}
# logfc.threshold = 0.5
# test.use = "wilcox"
# min.pct = 0.3
# min.diff.pct = 0
# only.pos = TRUE 
# 
# gene_signatures = list()
# 
# print(Sys.time())
# for (target in unique(seurat_obj@meta.data$signature_1_without_cc)){
#   gene_signatures_tmp = rownames(seurat_obj@raw.data)
#   for (metagene in unique(seurat_obj@meta.data$signature_1_without_cc)){
#     tmp = NULL
#     if (metagene != target){
#       tmp = FindMarkers(seurat_obj, ident.1 = target, ident.2 = metagene, only.pos = only.pos,
#                         logfc.threshold = logfc.threshold, test.use = test.use,
#                         min.pct = min.pct, min.diff.pct = min.diff.pct)
#       tmp = tmp[tmp$p_val_adj < 0.05, ]
#       gene_signatures_tmp = intersect(gene_signatures_tmp, rownames(tmp))
#     }
#   }
#   gene_signatures[[target]] = gene_signatures_tmp
# }
# print(Sys.time())
```

