---
title: "Kaplan-Meier survival analysis"
output: html_document
---

This template shows how to run kaplan-meier survival analysis. Bulk RNAseq data is split into 2 or more groups- based on high/low geneset score, cell type proportion, etc

Survival analysis is often run for multiple cohorts, but cohorts kept separate due to batch effects. This template allows for running multiple cohorts in parallel to streamline analysis

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(survminer)
library(factoextra)
library(viridis)
library(reshape2)
library(survival)
library(gridExtra)
library(RColorBrewer)

## Set working/script directory
base_dir = "/Users/jlabelle/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/04b_DeconvoluteBulk/")
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Scripts/"


## load helper functions
helper_functions<- paste0(base_dir, script_dir,"Survival_HelperFunctions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(helper_functions)
source(nmf_helper)
source(preprocessing)


## Create analysis/figure directories
figure_dir<- paste0("figures/KM_Survival/")
analysis_dir<- paste0("analysis/KM_Survival/")
data_dir<- paste0(working_dir, "/analysis/bulk_deconvolute/")

if(!dir.exists(figure_dir)){dir.create(figure_dir)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

```


## Load in processed data (03_DeconvoluteBulke_Preprocess.Rmd)
separate dataframes for each cohort- survival + high/low norm prop, separate for each major cell type
```{r}
files<- list.files(data_dir, pattern="HighLowPropSplit")

surv_obj_list_allCellTypes<- lapply(files, function(x){
  tmp<- readRDS(paste0(data_dir, x))
  return(tmp)
})
names(surv_obj_list_allCellTypes)<- gsub("HighLowPropSplit_clinical_", "", gsub(".Rds", "", files))
```


##############################
## Run KM survival analysis ##
##############################


## Run survival analysis: CD8
```{r}
genesets_test<- c("Memory", "Predysfunctional", "Cytotoxic_NKreceptor", "IFN_Response", "Stress", "Effector_Memory", "Cycling")
surv_obj_list<- lapply(surv_obj_list_allCellTypes, function(x){x$cd8})

surv_res_list<- lapply(surv_obj_list, function(FinalMeta){
  all_genesets_surv<- lapply(genesets_test, function(geneset){
    surv_res<- run_mySurvAnalysis(FinalMeta, DepColumnName = geneset)
    return(surv_res)
  })
  names(all_genesets_surv)<- genesets_test
  return(all_genesets_surv)
})


## Plot results separately for each cohort
for(cohort in names(surv_res_list)){
  print(cohort)
  
  ## Set output dir- separate for each cohort
  figure_dir_use<- paste0(figure_dir, cohort, "/")
  if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}
  
  ## Set input/km results
  input<- surv_obj_list[[cohort]]
  survRes<- surv_res_list[[cohort]]
  
  ## Plot
  all_plots<- lapply(genesets_test, function(geneset){
    plot_mySurvAnalysis(survRes[[geneset]]$fit,input, DepColumnName = geneset)
  })
  p<-arrange_ggsurvplots(all_plots, ncol=3, nrow=ceiling(length(all_plots)/3))
  ggsave(plot=p, paste0(figure_dir_use, "CD8_SurvRes_allGenesets.png"), width=9, height=9)
}
```



## Run survival analysis: CD4
```{r}
genesets_test<- c("Memory", "Treg", "Inhibitory", "IFN_Response", "Stress", "Effector_Memory", "Cycling")
surv_obj_list<- lapply(surv_obj_list_allCellTypes, function(x){x$cd4})

surv_res_list<- lapply(surv_obj_list, function(FinalMeta){
  all_genesets_surv<- lapply(genesets_test, function(geneset){
    surv_res<- run_mySurvAnalysis(FinalMeta, DepColumnName = geneset)
    return(surv_res)
  })
  names(all_genesets_surv)<- genesets_test
  return(all_genesets_surv)
})


## Plot results separately for each cohort
for(cohort in names(surv_res_list)){
  print(cohort)
  
  ## Set output dir- separate for each cohort
  figure_dir_use<- paste0(figure_dir, cohort, "/")
  if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}
  
  ## Set input/km results
  input<- surv_obj_list[[cohort]]
  survRes<- surv_res_list[[cohort]]
  
  ## Plot
  all_plots<- lapply(genesets_test, function(geneset){
    plot_mySurvAnalysis(survRes[[geneset]]$fit,input, DepColumnName = geneset, showPvalue = FALSE)
  })
  p<-arrange_ggsurvplots(all_plots, ncol=3, nrow=ceiling(length(all_plots)/3))
  ggsave(plot=p, paste0(figure_dir_use, "CD4_SurvRes_allGenesets_noPvalue.png"), width=9, height=9)
}
```

## Run survival analysis: Myeloid
```{r}
genesets_test<-  c("IFN_TAM", "Monocytes", "Inflammatory", "SEPP1_Mo_TAM", "Hypoxic_TAM", "Stress_Response")
surv_obj_list<- lapply(surv_obj_list_allCellTypes, function(x){x$myeloid})

surv_res_list<- lapply(surv_obj_list, function(FinalMeta){
  ## Remove any programs that are "low" for all samples (i.e., 0% for all)
  program_0per<- apply(FinalMeta[,genesets_test], 2, function(x){dim(table(x))==1})
  FinalMeta<- FinalMeta[,!colnames(FinalMeta) %in% names(program_0per)[program_0per]]
  genesets_test_use<- genesets_test[!genesets_test %in% names(program_0per)[program_0per]]
  
  all_genesets_surv<- lapply(genesets_test_use, function(geneset){
    surv_res<- run_mySurvAnalysis(FinalMeta, DepColumnName = geneset)
    return(surv_res)
  })
  names(all_genesets_surv)<- genesets_test_use
  return(all_genesets_surv)
})


## Plot results separately for each cohort
for(cohort in names(surv_res_list)){
  print(cohort)
  
  ## Set output dir- separate for each cohort
  figure_dir_use<- paste0(figure_dir, cohort, "/")
  if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}
  
  ## Set input/km results
  input<- surv_obj_list[[cohort]]
  survRes<- surv_res_list[[cohort]]
  
  ## Plot
  all_plots<- lapply(names(survRes), function(geneset){
    plot_mySurvAnalysis(survRes[[geneset]]$fit,input, DepColumnName = geneset)
  })
  p<-arrange_ggsurvplots(all_plots, ncol=3, nrow=ceiling(length(all_plots)/3))
  ggsave(plot=p, paste0(figure_dir_use, "Myeloid_SurvRes_allGenesets.png"), width=9, height=6)
}
```

## Run survival analysis: GOI
```{r}
genesets_test<- c("KLRB1", "PTGER2", "SELPLG", "GNLY", "PDCD1")

surv_obj_list<- lapply(surv_obj_list_allCellTypes, function(x){x$GOI})
surv_obj_list<-surv_obj_list[unlist(lapply(surv_obj_list, function(x){!is.null(x)}))]

surv_res_list<- lapply(surv_obj_list, function(FinalMeta){
  all_genesets_surv<- lapply(genesets_test, function(geneset){
    surv_res<- run_mySurvAnalysis(FinalMeta, DepColumnName = geneset)
    return(surv_res)
  })
  names(all_genesets_surv)<- genesets_test
  return(all_genesets_surv)
})


## Plot results separately for each cohort
for(cohort in names(surv_res_list)){
  print(cohort)
  
  ## Set output dir- separate for each cohort
  figure_dir_use<- paste0(figure_dir, cohort, "/")
  if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}
  
  ## Set input/km results
  input<- surv_obj_list[[cohort]]
  survRes<- surv_res_list[[cohort]]
  
  ## Plot
  all_plots<- lapply(genesets_test, function(geneset){
    plot_mySurvAnalysis(survRes[[geneset]]$fit,input, DepColumnName = geneset)
  })
  p<-arrange_ggsurvplots(all_plots, ncol=3, nrow=ceiling(length(all_plots)/3))
  ggsave(plot=p, paste0(figure_dir_use, "GOI_SurvRes_allGenesets.png"), width=9, height=6)
}
```


