---
title: "Multivariate Cox regression analysis"
output: html_document
---

This template used to run multivariate cox regression analysis based on gene set score. This allows for controlling for confounders like tumor type, age, etc.

## Overall steps:

  1. Determine which clinical variables to control for- run univariate cox to identify variables that may have an effect on survival
  2. Determine baseline level of effect of gene set score on survival- univariate cox. Can compare to kaplan-meier survival results (should be very similar)
  3. Determine effect of gene set score on survival when clinical variables are included in model- multivariate cox
  
```{r}
library(survival)
library(survminer)
library(dplyr)

## Set working/script directory
base_dir = "/Users/jlabelle/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/04b_DeconvoluteBulk/")
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Scripts/"

## Load helper functions
helper_functions<- paste0(base_dir, script_dir,"Survival_HelperFunctions.R")
source(helper_functions)


## Set directories
data_dir<- paste0("analysis/bulk_deconvolute/")
analysis_dir<- paste0("analysis/coxRegression/")
figure_dir<- paste0("figures/coxRegression/")
if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

```

## Load in data
Input:
list of "FinalMeta" from SurvivalAnalysis_template.Rmd

Minimum columns required: 
  a. SampleID ("submitter_id") 
  b. time to death ("time") 
  c. alive/dead ("censored"; alive= F, dead=T). Should be set as logical. 
  d. At least 1 method for splitting samples. Can be 1) raw scores 2) high/low scores 3) raw proportion 4) high/low proportion. Each geneset/program should have its own column

2.Clinical metadata. Contains variables you want to control for- age/sex/diagnosis/treatment/etc. First column should be sample ID. 
Note: should remove any variables with NA values for any sample, or remove those samples.
```{r}
## Load in processed data (03_DeconvoluteBulke_Preprocess.Rmd)
files<- list.files(data_dir, pattern="HighLowPropSplit")

## Subset to just CBTN- 2 different cohorts, clinical matches exactly
## For other cohorts, will have to run separately for each cohort or match up the metatdata as much as possible
files<- files[grepl("CBTN", files)]

surv_obj_list_allCellTypes<- lapply(files, function(x){
  tmp<- readRDS(paste0(data_dir, x))
  return(tmp)
})
names(surv_obj_list_allCellTypes)<- gsub("HighLowPropSplit_clinical_", "", gsub(".Rds", "", files))

## Add cohort/cell type info to each
final_meta_list<- lapply(names(surv_obj_list_allCellTypes), function(cohort){
  surv_obj_allCellTypes<- surv_obj_list_allCellTypes[[cohort]]
  finalMeta_allCellTypes<- lapply(names(surv_obj_allCellTypes), function(cellType){
    finalMeta<- surv_obj_allCellTypes[[cellType]]
    finalMeta$cohort<- cohort
    finalMeta$cellType<- cellType
    return(finalMeta)
  })
  names(finalMeta_allCellTypes)<- names(surv_obj_allCellTypes)
  return(finalMeta_allCellTypes)
  })
names(final_meta_list)<-names(surv_obj_list_allCellTypes)

## Flatten into 1 finalMeta df for each celltype
finalMeta_allCellTypes<- lapply(c("cd8", "cd4", "myeloid"), function(cellType){
  tmp<- lapply(final_meta_list, function(x){x[[cellType]]})
  tmp<- do.call("rbind", tmp)
  return(tmp)
})
names(finalMeta_allCellTypes)<- c("cd8", "cd4", "myeloid")


```


## Set variables to potentially control for
## These are variables from your clinical metadata that you think may have an effect on survival.
## Can also just use all variables
```{r}
## List with variables to control for (if running across multiple cohorts, may want to include "Cohort")
ToControlFor<- c("AGE", "CANCER_TYPE", "ETHNICITY", "TUMOR_TYPE", "cohort")

```


## Optional: 
1. Create additional variable, merging similar diagnoses together. This may not apply to all data.
2. Subset by diagnosis- bc glioblastoma diagnosis has such a large effect, may want to remove it from consideration
```{r}
## Optional: add column to meta- astro/oligo/mixed/glio
FinalMeta$MyDiagnosis<-gsub("Mixed glioma", "MixedGlioma", gsub("Astrocytoma, NOS", "Astrocytoma", 
                         gsub("Oligodendroglioma, anaplastic", "Oligodendroglioma", 
                              gsub("Oligodendroglioma, NOS", "Oligodendroglioma", 
                                   gsub("Astrocytoma, anaplastic", "Astrocytoma", 
                                        gsub("Glioblastoma", "Glioblastoma",
                                             FinalMeta$primary_diagnosis))))))
FinalMeta$MyDiagnosis<-as.factor(FinalMeta$MyDiagnosis)


## Optional: Subset to remove gbm or just look at 1 type
FinalMeta_NoGBM<- FinalMeta[FinalMeta$project_id=="TCGA-LGG",]
FinalMeta_Mixed<-FinalMeta[FinalMeta$primary_diagnosis=="Mixed glioma",]
```

                      
## Run univariate cox to determine which variables to control for
## RunCoxWrapper function input: 
geneSet_groups = vector of genesets names to apply wrapper to
FinalMeta = "FinalMeta"- contains time, censored, and all genesets in geneSet_groups
controlFor = whether or not this will be univariate (FALSE) or multivariate (TRUE)
mySubset = which samples used in analysis- only used for saving analysis. useful when trying out multiple subsets
exportName = analysis name to use when saving data
exportCsv = whether or not to export csv file of results
subsetToHighLow = when splitting 25/50/25, subset to just 25/25

#########
## CD8 ##
#########
```{r}
## Run univariate cox for all potential controlling variables
## set exportCsv to FALSE if you want to examine results before exporting
allSamples_diagnosis<- RunCoxWrapper(geneSet_groups = ToControlFor, AllMeta= finalMeta_allCellTypes$cd8, mySubset="AllSamples", 
                                     controlFor=FALSE, 
                                     exportName = "CD8_univariate_variables", exportCsv=FALSE, resultsFolder = analysis_dir)

## select just variables that had significant hazard
cox_df<-na.omit(allSamples_diagnosis$cox_df)
sig<-cox_df[cox_df$`Pr(>|z|)` < 0.1,]
sig_variables<-unique(sig$Variable)

ToBeControlled<- sig_variables
```


## Run univariate cox for genesets
## Univariate
```{r}
genesets_test<- c("Memory", "Predysfunctional", "Cytotoxic_NKreceptor", "IFN_Response", "Stress", "Effector_Memory", "Cycling")

## Run univariate for all samples
allSamples<- RunCoxWrapper(genesets_test, finalMeta_allCellTypes$cd8, mySubset="AllSamples", 
                                    controlFor=FALSE, exportName = "CD8_univariate",
                                    exportCsv=TRUE, subsetToHighLow = TRUE, resultsFolder = analysis_dir)
```

## Run multivariate cox, for all gene sets, controlling for variables

```{r}
## Control for all sig clinical confounders- can also specify only a subset of these
allSamples_diagnosis<- RunCoxWrapper(genesets_test, finalMeta_allCellTypes$cd8, mySubset="AllSamples",
                                     controlFor=ToControlFor,
                                     exportName="CD8_ControlForAll", exportCsv = TRUE, resultsFolder = analysis_dir)


```

#########
## CD4 ##
#########
```{r}
## Run univariate cox for all potential controlling variables
## set exportCsv to FALSE if you want to examine results before exporting
allSamples_diagnosis<- RunCoxWrapper(geneSet_groups = ToControlFor, AllMeta= finalMeta_allCellTypes$cd4, mySubset="AllSamples", 
                                     controlFor=FALSE, 
                                     exportName = "CD4_univariate_variables", exportCsv=FALSE, resultsFolder = analysis_dir)

## select just variables that had significant hazard
cox_df<-na.omit(allSamples_diagnosis$cox_df)
sig<-cox_df[cox_df$`Pr(>|z|)` < 0.1,]
sig_variables<-unique(sig$Variable)

ToBeControlled<- sig_variables
```


## Run univariate cox for genesets
## Univariate
```{r}
genesets_test<- c("Memory", "Treg", "Inhibitory", "IFN_Response", "Stress", "Effector_Memory", "Cycling")

## Run univariate for all samples
allSamples<- RunCoxWrapper(genesets_test, finalMeta_allCellTypes$cd4, mySubset="AllSamples", 
                                    controlFor=FALSE, exportName = "CD4_univariate",
                                    exportCsv=TRUE, subsetToHighLow = TRUE, resultsFolder = analysis_dir)
```

## Run multivariate cox, for all gene sets, controlling for variables

```{r}
## Control for all sig clinical confounders- can also specify only a subset of these
allSamples_diagnosis<- RunCoxWrapper(genesets_test, finalMeta_allCellTypes$cd4, mySubset="AllSamples",
                                     controlFor=ToControlFor,
                                     exportName="CD4_ControlForAll", exportCsv = TRUE, resultsFolder = analysis_dir)


```


## Testing proportionality assumption
## Check this for at least one of the above models
## Tests whether it's appropriate to apply regression model to these variables
## If p<0.05, may need to stratify by that variable rather than controlling for it
```{r}
lapply(allSamples_diagnosis$cox_objs, function(x){cox.zph(x)})
```


#############
## Myeloid ##
#############
```{r}
## Run univariate cox for all potential controlling variables
## set exportCsv to FALSE if you want to examine results before exporting
allSamples_diagnosis<- RunCoxWrapper(geneSet_groups = ToControlFor, AllMeta= finalMeta_allCellTypes$myeloid, mySubset="AllSamples", 
                                     controlFor=FALSE, 
                                     exportName = "Myeloid_univariate_variables", exportCsv=FALSE, resultsFolder = analysis_dir)

## select just variables that had significant hazard
cox_df<-na.omit(allSamples_diagnosis$cox_df)
sig<-cox_df[cox_df$`Pr(>|z|)` < 0.1,]
sig_variables<-unique(sig$Variable)

ToBeControlled<- sig_variables
```


## Run univariate cox for genesets
## Univariate
```{r}
genesets_test<-  c("IFN_TAM", "Monocytes", "Inflammatory", "SEPP1_Mo_TAM", "Hypoxic_TAM", "Stress_Response")


## Run univariate for all samples
allSamples<- RunCoxWrapper(genesets_test, finalMeta_allCellTypes$myeloid, mySubset="AllSamples", 
                                    controlFor=FALSE, exportName = "Myeloid_univariate",
                                    exportCsv=TRUE, subsetToHighLow = TRUE, resultsFolder = analysis_dir)
```

## Run multivariate cox, for all gene sets, controlling for variables

```{r}
## Control for all sig clinical confounders- can also specify only a subset of these
allSamples_diagnosis<- RunCoxWrapper(genesets_test, finalMeta_allCellTypes$myeloid, mySubset="AllSamples",
                                     controlFor=ToControlFor,
                                     exportName="Myeloid_ControlForAll", exportCsv = TRUE, resultsFolder = analysis_dir)


```


## Testing proportionality assumption
## Check this for at least one of the above models
## Tests whether it's appropriate to apply regression model to these variables
## If p<0.05, may need to stratify by that variable rather than controlling for it
```{r}
lapply(allSamples_diagnosis$cox_objs, function(x){cox.zph(x)})
```













## Optional: merge results together into a single df in order to summarize in a plot
Useful when you run multiple cohorts separately, but want to get an overview of all cohorts together
May need to modify for your data
```{r}
genesets_test<-  c("IFN_TAM", "Monocytes", "Inflammatory", "SEPP1_Mo_TAM", "Hypoxic_TAM", "Stress_Response",
                   "Memory", "Treg", "Inhibitory", "IFN_Response", "Stress", "Effector_Memory", "Cycling",
                   "Memory", "Predysfunctional", "Cytotoxic_NKreceptor", "IFN_Response", "Stress", "Effector_Memory", "Cycling")

## Get all results files
multi_files<-list.files(analysis_dir, pattern="ControlFor")
multi_files<- multi_files[grep("AllSamples",multi_files)]
uni_files<-list.files(analysis_dir, pattern="univariate")
uni_files<- uni_files[!grepl("variables_", uni_files)]

## Initiate dataframe with univariate file:
  ## Read in first file
  AllCox<- read.csv(paste0(analysis_dir, uni_files[1]))
  
  #Subset to variables of interest, rename
  AllCox<-AllCox[,colnames(AllCox) %in% c("X", "exp.coef.", "Pr...z..")]
  colnames(AllCox)<- c("Variable", "HazardRatio", "pvalue")
  
  ## Add "ControlFor" column- "None" for univariate
  AllCox$ControlFor<- "None"
  
  ## Add major cell type
  AllCox$CellType<- gsub("_.*", "" ,uni_files[1])

## cycle through rest of univariate, adding to univariate df
  for (i in uni_files[2:length(uni_files)]){
    df<- read.csv(paste0(analysis_dir, i))
    df<-df[,colnames(df) %in% c("X", "exp.coef.", "Pr...z..")]
    colnames(df)<- c("Variable", "HazardRatio", "pvalue")
    
    ## Add "ControlFor" column- "None" for univariate
    df$ControlFor<-"None"
    
    ## Add major cell type
    df$CellType<- gsub("_.*", "" ,i)
    
    AllCox<- rbind(AllCox, df)
  }
  
  
## cycle through rest of cox (multivariate), adding to univariate df
  for (i in multi_files){
    df<- read.csv(paste0(analysis_dir, i))
    df<-df[,colnames(df) %in% c("X", "exp.coef.", "Pr...z..")]
    colnames(df)<- c("Variable", "HazardRatio", "pvalue")
    
    #Get name of variable being controlled for, add to dataframe
    ControlFor<- gsub("_AllSamples_results.csv", "",i )
    ControlFor<-gsub("ControlFor", "", ControlFor)
    df$ControlFor<-ControlFor
    
    ## Remove any rows that are the variable being controlled for- not of interest
    df2<- df[df$Variable %in% paste0(genesets_test, "Low"),]

    
    ## Add major cell type
    df2$CellType<- gsub("_.*", "" ,i)
    
    AllCox<- rbind(AllCox, df2)
  }


write.csv(AllCox,paste0(analysis_dir, "allCoxResults_CBTN.WT.K27M.csv"))
```

## Plot results
```{r}
df<- AllCox

df$tmp<-df$pvalue<0.05
df$tmp2<- df$pvalue<0.1
df$Sig<-rowSums(df[,c("tmp", "tmp2")])
df$Sig<-gsub("0", "NotSig", gsub("1", "AlmostSig", gsub("2", "Sig", df$Sig)))
df$Sig<-factor(df$Sig, levels=c("NotSig", "AlmostSig", "Sig"))

ggplot(df, aes(x=HazardRatio, y=ControlFor,color=Variable ,alpha=Sig))+
  geom_point(size=8)+
  geom_vline(xintercept=1)+
  xlab("Hazard Ratio")+
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(face="bold", color="black", size="12"))+
  facet_grid(rows=vars(CellType), space="free_y", scale="free_y")
```




