---
title: "Functional markers"
output: html_document
---

## Purpose:
Identify alternative and/or supporting programs for myeloid cells

## Approach:
Score myeloid cells for programs Li shared with me on 10/27/21- "functional" markers. Plot by max score, compare to nmf programs, etc
These annotations not really used downstream at all

```{r}
library(Seurat) 

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02b_Myeloid/02_AdditionalAnnotation/")

analysis_dir<- paste0(working_dir, "/analysis/FunctionalMarkers/")
figure_dir<- paste0(working_dir, "/figures/FunctionalMarkers/")
data_dir_tmp<- paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",integration)

if(!dir.exists(figure_dir)){dir.create(figure_dir)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)


```

## Read in ped myeloid seurat object
```{r}
cohort<- "pedOnly_nomuv63"

data_dir<- paste0(data_dir_tmp, cohort, "/")
analysis_dir<- paste0(analysis_dir, cohort, "/")
figure_dir<- paste0(figure_dir, cohort, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

## load ped myeloid seurat object- without DC or B cells
seurat_obj<- readRDS(paste0(data_dir, "myeloid_seurat_noDC.B.Rds"))

## Norm/center counts
cm_list<-NormCenter(seurat_obj@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data + 1))
```

## Read in myeloid markers- "functional" markers that Li shared with me on 10/27/21. Unsure of source.
```{r}
markers<- readRDS("../../../../../../Marker_genes/myeloid_functional_programs_Li10.27.21.rds")
```

## Score for functional programs
```{r}
## Score
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data,cm_mean, markers)))

## Assign max, add to seurat obj
scores$MaxProgram<- apply(scores, 1, function(x){names(x)[which.max(x)]})
scores$MaxScore<- apply(scores[,colnames(scores) != "MaxProgram"], 1, function(x){x[which.max(x)]})
seurat_obj$functionalMarkers_maxProgram<- plyr::mapvalues(colnames(seurat_obj), rownames(scores), scores$MaxProgram)
seurat_obj$functionalMarkers_maxScore<- plyr::mapvalues(colnames(seurat_obj), rownames(scores), scores$MaxScore)

## save updated seurat object/scores
saveRDS(seurat_obj, paste0(data_dir, "myeloid_seurat_noDC.B.Rds"))
saveRDS(scores, file=paste0(analysis_dir, "scores.Rds"))
```

## Plot max programs
```{r}
program_colors<- c("dodgerblue", "green", "olivedrab", "darkgreen", "burlywood4", "magenta",  "yellow2", 
                          "red", "orange", "cyan", "plum4", "purple")
names(program_colors)<- names(markers)

## UMAP of top programs
DimPlot(seurat_obj,group.by = "functionalMarkers_maxProgram")+scale_color_manual(values=program_colors)
ggsave(paste0(figure_dir, "UMAP_TopFunctionalProgram.png"), width=6, height=4)

## UMAP of top programs, with scoreTooLow < 0.5
seurat_obj$functionalMarkers_maxProgram_STL<- ifelse(seurat_obj$functionalMarkers_maxScore>0.5,
                                                     seurat_obj$functionalMarkers_maxProgram, "ScoreTooLow")
DimPlot(seurat_obj,group.by = "functionalMarkers_maxProgram_STL")+scale_color_manual(values=c(program_colors, ScoreTooLow="grey"))
ggsave(paste0(figure_dir, "UMAP_TopFunctionalProgram_STL.png"), width=6, height=4)

## Breakdown by NMF program. NMF programs annotated by correlation with adult 10x dataset (see AdultGBM10X_AnnotatePed.Rmd)
df<- seurat_obj@meta.data
bd<- as.data.frame(table(df$functionalMarkers_maxProgram, df$NMF_rank6_annotByAdult10X))
colnames(bd)<- c("functional_program", "Adult10XProgram", "NCells")
bd<- bd %>% group_by(Adult10XProgram) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()

ggplot(bd, aes(x=Adult10XProgram, y=PerCells, fill=functional_program))+
  geom_bar(stat="identity", position="stack")+
  scale_fill_manual(values = program_colors)+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("")
ggsave(paste0(figure_dir, "Breakdown_NMFprogram_byFunctionalProgram.png"), width=6, height=4)
```

## Correlate scores for NMF programs with scores for functional programs
```{r}
## read in scores
functional_scores<- readRDS(paste0(analysis_dir, "scores.Rds"))
rank_use<-6
nmf_scores<- readRDS(paste0("../01a_NMF/analysis/de.novoNMF/", cohort, "/NMF_scores_annotByAdult10X_rank6-6.Rds"))
nmf_scores<- nmf_scores[[paste0("rank", rank_use)]]

## convert NMF programs to annotated
#nmf_annots<- readRDS(paste0("analysis/GBM.10X_AnnotatePed/PedNMFrank", cohort, "/", rank_use, "_RenamedByGBM10XCorrelation.Rds"))
#colnames(nmf_scores)<- plyr::mapvalues(colnames(nmf_scores),nmf_annots, names(nmf_annots))

## Merge
all_scores<- cbind(nmf_scores, functional_scores)
all_scores<- all_scores[,!grepl("Max", colnames(all_scores))]

## Correlate and plot
scores_factor_hc = clusterNmfFactors(all_scores)
scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]

  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  pheatmap(scores_factor_cor, color = hm_colors(100), 
           cluster_rows = F, cluster_cols = F, 
           annotation_names_row = F, annotation_names_col =T,
           show_rownames = T, show_colnames = F,
           filename = paste0(figure_dir, "Corr_PedNMFrank", rank_use, "FunctionalPrograms.png"),
           width = 8, height = 4)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0( figure_dir, "HCC_PedNMFrank", rank_use, "FunctionalPrograms.png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
```

