---
title: "Microglia, macrophage annotation"
output: html_document
---

## Purpose:
Identify potential microglia/macrophage cells in myeloid cell datasets. To be used as additional layer of annotation on top of NMF

## Approach:
Score myeloid cells for microglia/macrophage genesets, visualize spread of scores, high scorers, etc. 

```{r}
library(Seurat) 

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02b_Myeloid/02_AdditionalAnnotation/")

integration<- "" ## HarmonyInt/ or ""
analysis_dir_tmp<- paste0(working_dir,"/analysis/Microglia.Macrophage/", integration)
figure_dir_tmp<- paste0(working_dir, "/figures/Microglia.Macrophage/", integration)
data_dir_tmp<- paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",integration)

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)


```

## Read in ped myeloid seurat object
```{r}
cohort<- "pedOnly_nomuv63"
data_dir<- paste0(data_dir, cohort, "/")

figure_dir<- paste0(figure_dir_tmp, cohort, "/")
analysis_dir<- paste0(analysis_dir_tmp, cohort, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

use_harmony_forVis<- "_harmony"
if(use_harmony_forVis=="_harmony"){
  figure_dir<- paste0(figure_dir, "Harmony/"); if(!dir.exists(figure_dir)){dir.create(figure_dir)}
  }

## load ped myeloid seurat object- without DC or B cells
seurat_obj<- readRDS(paste0(data_dir, "myeloid_seurat_noDC.B",use_harmony_forVis, ".Rds"))
```

 
## Marker genes
```{r}
## Mg: "CLEC12A", "FXYD5"; Mo: "SALL1",
FeaturePlot(seurat_obj,
            c("P2RY12","TMEM119",
              "CD163", "TGFBI"),
            cols = c("lightgrey", "red"),
            pt.size = 1,
            ncol = 2) & NoAxes() & NoLegend()
ggsave("mo_mg_genes_feature.png", path=figure_dir,
       width=6, height=6)
```

## Score macrophage and microglia program
```{r}
## Norm and center counts
cm = GetAssayData(seurat_obj, slot = "counts")
cm_norm = log2(cm/10+1)
cm_mean = log2(rowMeans(cm)+1)
cm_center = cm_norm-rowMeans(cm_norm)

## Read in microglia and macrophage markers
gene_set_dir<- "../../../../../../Marker_genes/"
microglia_genes = read.table(paste0(gene_set_dir, "Microglia_markers.txt"))
macrophage_genes = read.table(paste0(gene_set_dir, "Macrophage_markers.txt"))
tam_genes = list("microglia" = microglia_genes$V1, "macrophage" = macrophage_genes$V1)

## Score for microglia/macrophage markers
tam_score = scoreNmfGenes(cm_center, cm_mean, tam_genes)
tam_score = data.frame(t(tam_score))
```
 
## Correlation between macrophage and microglia scores
```{r}
ggplot(tam_score, aes(x=microglia, y=macrophage)) + geom_point() +
  labs(x="Microglia score", y="Macrophage score") +
  theme(axis.title=element_text(size=20),
        axis.text=element_text(size=20))
ggsave("microglia_macrophage_correlation.png", path=figure_dir,
       width=6, height=6)
```
 
## Add score and annotation to seurat obj
```{r}
seurat_obj = AddMetaData(seurat_obj, tam_score$microglia, "microglia_score")
seurat_obj = AddMetaData(seurat_obj, tam_score$macrophage, "macrophage_score")

## Assign as macrophage/microglia
mg_annot = ifelse(seurat_obj$microglia_score > seurat_obj$macrophage_score,
                  "microglia", "macrophage")

## Alternative: only assign as macrophage/microglia if score is high enough
mg_annot = ifelse(seurat_obj$microglia_score < 0 & seurat_obj$macrophage_score < 0,
                  "undetermined",
                  ifelse(seurat_obj$microglia_score > seurat_obj$macrophage_score,
                        "microglia", "macrophage"))
seurat_obj = AddMetaData(seurat_obj, mg_annot, "MM_annot")

## alternative: assign microglia/macrophage if scores are high
score_thresh<- 0.5
seurat_obj$AltMM_annot<- ifelse(seurat_obj$microglia_score>score_thresh & seurat_obj$macrophage_score >score_thresh, "HighBoth",
                                ifelse(seurat_obj$microglia_score>score_thresh, "Microglia",
                                       ifelse(seurat_obj$macrophage_score>score_thresh, "Macrophage", "NotHighForEither")))

saveRDS(seurat_obj, file=paste0(data_dir, "myeloid_seurat_noDC.B", use_harmony_forVis,".Rds"))
```
 
## Plot microglia/macrophage classification and score
```{r}
## Plot annotations- only "undetermined" if score for both is too low
DimPlot(object = seurat_obj,
         group.by = "MM_annot",
         label = FALSE,
         pt.size = 2,
         label.size = 8) + NoAxes()+
  scale_color_manual(values=c(macrophage="red", microglia="blue", undetermined="grey50"))
ggsave("UMAP_mg_classifications.png", path=figure_dir, width=6, height=4)

## Plot annotations- 
DimPlot(object = seurat_obj,
         group.by = "AltMM_annot",
         label = FALSE,
         pt.size = 2,
         label.size = 8) + NoAxes()+
  scale_color_manual(values=c(Macrophage="red", Microglia="blue", NotHighForEither="grey",
                              HighBoth="green"))
ggsave(paste0("UMAP_mg_classifications_ScoreThresh", score_thresh, ".png"), path=figure_dir, width=6, height=4)

## Plot numeric scores for microglia/macrophage
all_plots<- lapply(c("microglia_score", "macrophage_score"), function(x){
  FeaturePlot(seurat_obj,
            features = x )+
    scale_color_gradient(low="white", high="red")
})
cowplot::plot_grid(plotlist=all_plots, ncol=2)

ggsave("tam_scores.png", path=figure_dir, width=9, height=4)
```

## Plot gradient of macrophage/microglia scores
```{r}
seurat_obj$MM_gradient<- seurat_obj$macrophage_score - seurat_obj$microglia_score
FeaturePlot(seurat_obj, features="MM_gradient")+
  scale_color_gradient2(low="blue", mid="white", high="red")+
  ggtitle("> 0 = macrophage / < 0 = microglia")
ggsave(paste0(figure_dir, "microglia.macrophage_gradient.png"), width=6, height=5)
```


## Do scores for NMF programs correlate with microglia/macrophage scores?
```{r}
## Read in scores for all NMF programs, for all ranks
rank_use<-6
all_scores<- readRDS(paste0("../01a_NMF/analysis/de.novoNMF/",integration,
                            cohort, "/NMF_scores_annotByAdult10X_rank6-6.Rds"))
scores<- all_scores[[paste0("rank", rank_use)]]

## Read in annotations- these annotations are based on correlation with adult GBM 10X (see AdultGBM10X_AnnotatePed.Rmd) EXCEPT for Inflammatory and Stress_response. These 2 programs do not exist in adult 10X dataset, and are based on correlation with "original" myeloid programs (see NMF/Myeloid/NMF.Rmd)
#annotations<- readRDS(paste0("analysis/GBM.10X_AnnotatePed/", "PedNMFrank", rank_use, "_RenamedByGBM10XCorrelation.Rds"))

## Rename scores with annotations from 10x
#colnames(scores)<- plyr::mapvalues(colnames(scores), annotations, names(annotations))
#scores$MaxProgram<- plyr::mapvalues(scores$MaxProgram, annotations, names(annotations))

## Add on microglia/macrophage score
scores$macrophage_score<- as.numeric(plyr::mapvalues(rownames(scores), colnames(seurat_obj), seurat_obj$macrophage_score))
scores$microglia_score<- as.numeric(plyr::mapvalues(rownames(scores), colnames(seurat_obj), seurat_obj$microglia_score))

## Cycle through each program, plotting correlation with macrophage and microglia scores
all_plots<-lapply(colnames(all_scores[[paste0("rank", rank_use)]]), function(x){
  tmp<- scores[,c(x, "macrophage_score", "microglia_score")]
  colnames(tmp)<- c("program_score", "macrophage_score", "microglia_score")
  p1<- ggplot(tmp, aes(x=program_score, y=macrophage_score))+geom_point(color="red")+ggtitle(paste0("Macrophage vs ", x))
  p2<- ggplot(tmp, aes(x=program_score, y=microglia_score))+geom_point(color="blue")+ggtitle(paste0("Microglia vs ", x))
  p1.2<-cowplot::plot_grid(plotlist=list(p1, p2), ncol = 2)
  return(p1.2)
})
cowplot::plot_grid(plotlist = all_plots, ncol = 1)
ggsave(paste0(figure_dir, "macrophage.microgliaScores_vs_nmfScores.png"), width=8, height=20)


## Alternative- standard correlation heatmap
scores_tmp<- scores[,!colnames(scores) %in%c("MaxProgram", "MaxScore")]

scores_factor_hc = clusterNmfFactors(scores_tmp)
scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]

  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  pheatmap(scores_factor_cor, color = hm_colors(100), 
           cluster_rows = F, cluster_cols = F, 
           annotation_names_row = F, annotation_names_col =T,
           show_rownames = T, show_colnames = F,
           filename = paste0(figure_dir, "Corr_PedNMFrank", rank_use, "Microglia.Macrophage.png"),
           width = 8, height = 4)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0( figure_dir, "HCC_PedNMFrank", rank_use, "Microglia.Macrophage.png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
```

## For each program, subset to cells scoring highest for that program. Then plot their Mg_TAM vs their Mo_TAM scores
```{r}
rank_use<-7
colors_use<-c("purple","turquoise3", "red", "orange", "navy", "olivedrab4", "grey22"  )
names(colors_use)<- as.character(unique(myeloid_seurat@meta.data[[paste0("NMF_rank", rank_use, "_annotByAdult10X")]]))


## Read in scores for all NMF programs, for all ranks
all_scores<- readRDS(paste0("../01a_NMF/analysis/de.novoNMF/",integration, cohort, "/NMF_scores_annotByAdult10X_rank6-6.Rds"))
scores<- all_scores[[paste0("rank", rank_use)]]
programs<-colnames(scores)[!grepl("Max", colnames(scores))]


## Add on microglia/macrophage score
scores$microglia_score<- as.numeric(plyr::mapvalues(rownames(scores), colnames(seurat_obj), seurat_obj$microglia_score))
scores$macrophage_score<- as.numeric(plyr::mapvalues(rownames(scores), colnames(seurat_obj), seurat_obj$macrophage_score))

## cycle through each program, subset to those cells, plot moTAM vs MgTAM
all_plots<- lapply(programs, function(x){
  df<- scores[scores$MaxProgram==x, c("macrophage_score", "microglia_score")]
  ggplot(df, aes(x=macrophage_score, y=microglia_score))+
    geom_point(color=colors_use[x])+
    ggtitle(x)+
    ylim(-3,3)+xlim(-2,3)+
    geom_hline(yintercept = 0)+geom_vline(xintercept = 0)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "Macrophage.vs.Microglia_scores_ByMaxNMFProgram.png"), width=9, height=9)
```

## Histogram of program scores, colored by mg_tam/mo_tam distinction
```{r}
## Read in scores for programs
rank_use<- 6
all_scores<- readRDS(paste0("../01a_NMF/analysis/de.novoNMF/",cohort,
                            "/NMF_scores_annotByAdult10X_rank6-6", ".Rds"))
scores<- all_scores[[paste0("rank", rank_use)]]
programs<- colnames(scores[!grepl("Max", colnames(scores))])

## Add on max mo_tam/mg_tam
scores$MM_annot<- plyr::mapvalues(rownames(scores), colnames(seurat_obj), seurat_obj$MM_annot)



## plot
all_plots<- lapply(programs, function(x){
  df<- scores[,c(x, "MM_annot")]
  colnames(df)<- c("ProgramScore", "MM_annot")
  p<- ggplot(df, aes(x=ProgramScore, fill=MM_annot))+
    geom_histogram()+
    scale_fill_manual(values=c(microglia="blue", macrophage="red", undetermined="grey"))+
    ggtitle(x)
    return(p)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "Histogram_ProgramScores.png"), width=15, height=6)
```


## Density plots
```{r}
## Read in scores for programs
rank_use<- 7
all_scores<- readRDS(paste0("../../NMF/Myeloid/analysis/de.novoNMF/",integration,cohort,
                            "/NMF_scores_annotByAdult10X_rank6-NA", ".Rds"))
scores<- all_scores[[paste0("rank", rank_use)]]
programs<- colnames(scores[!grepl("Max", colnames(scores))])


## Create df for plotting
myeloid_df<-seurat_obj@meta.data[,c("sampleid", "MM_annot", paste0("NMF_rank", rank_use, "_annotByAdult10X"))]
sum(rownames(scores)==rownames(myeloid_df)); nrow(scores); nrow(myeloid_df)
myeloid_df<- cbind(myeloid_df, scores[,!grepl("Max", colnames(scores))] )

  ## optional- remove undetermined
  myeloid_df<- myeloid_df[myeloid_df$MM_annot!="undetermined",
                          ]
colors_use<- c(microglia="blue", macrophage="red", undetermined="grey")


myeloid_df$MM_annot<- factor(myeloid_df$MM_annot, levels=c("undetermined","macrophage", "microglia"))

## Plot each nmf program
all_plots<- list()
for(i in as.character(unique(seurat_obj@meta.data[[paste0("NMF_rank", rank_use, "_annotByAdult10X")]]))){
  df<- myeloid_df[,c(as.character(i), "MM_annot")]
  colnames(df)<- c("score", "CellType")
  p<-ggplot(df, aes(x=score, fill=CellType))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.position = "none",
          plot.title = element_text(size=20, face="bold", hjust=0.5))

  all_plots[[i]]<-p
}

## add legend separately
legend<-ggplot(df, aes(x=score, fill=CellType))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.key.size = unit(1.5, 'cm'),
          legend.text = element_text(size=20))
all_plots$legend<-  as_ggplot(get_legend(legend))
all_plots<- all_plots[c(1:3,8,4:7)]
cowplot::plot_grid(plotlist = all_plots, ncol=4)

ggsave(paste0(figure_dir, "DensityOfNMFByMacrophage.Microglia_noUndetermined_rank", rank_use,".png"), width=20, height=8)

```



