---
title: "R Notebook"
output: html_document
---

## Purpose:
Compare proportions/scores for myeloid, tcell, and tumor programs. Do certain programs/cell types correlate with others? Here, perform necessary preprocessing (generating proportions by sample)

## Approach:
Load in seurat objects + scores for myeloid, tcell, and tumor ped samples. Generate proportions by sample for each program/cell type within each major cell type.

```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/", 
                     "04a_Integrate_Immune.Tumor/Correlate_Immune.Tumor_CellType.NMF/")

library(Seurat)
library(ggplot2)
library(ggpubr)
library(randomcoloR)

analysis_dir_tmp<- paste0("analysis/")
data_dir_tmp<- paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/")
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)
source(paste0(working_dir, "HelperFunctions.R"))
```



## Load in seurat objects + scores
```{r}
rank_use<- 8; sharedBy<- "union"; cohort<- "pedOnly_nomuv63"
analysis_dir<- paste0(analysis_dir_tmp, cohort, "/")
if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

tcell_data_dir<- paste0("../../02a_Tcells/02_Ped.Adult/03_Visualize.Reannotate_programs/analysis/Reannotate_programs/",
                         "rank", rank_use, "_sharedBy", sharedBy, "/")
myeloid_data_dir<- paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/", cohort, "/")
tumor_data_dir<- "../../../TumorCellAnnotation/AllSubtypesTogether/analysis/Merge_allSubtypes/"

## Load in seurat objects
tcell_seurat<- readRDS(paste0(tcell_data_dir, "ped_seurat_tcell_harmony.Rds"))
cd4_seurat<- readRDS(paste0(tcell_data_dir, "ped_seurat_cd4_harmony.Rds"))
cd8_seurat<- readRDS(paste0(tcell_data_dir, "ped_seurat_cd8_harmony.Rds"))
myeloid_seurat<- readRDS(paste0(myeloid_data_dir, "myeloid_seurat_noDC.B_harmony.Rds"))
load(paste0(tumor_data_dir, "AllTumorSeuratObj.Robj"))
tumor_seurat<- tumor_merged; rm(tumor_merged)

## Load in scores- currently don't have tumor cell scores, but may add this in
myeloid_scores<- readRDS(paste0("../../02b_Myeloid/01a_NMF/analysis/de.novoNMF/", 
                                cohort, "/NMF_scores_annotByAdult10X_rank6-6", ".Rds"))
myeloid_scores<- myeloid_scores$rank6
cd4_scores<- readRDS(paste0("../../02a_Tcells/02_Ped.Adult/02_detailed_annot/analysis/identifyCD4.CD8_byNMF/",
                              "rank", rank_use, "_sharedBy", sharedBy, "/ped_cd4_program_scores.Rds"))
cd8_scores<- readRDS(paste0("../../02a_Tcells/02_Ped.Adult/02_detailed_annot/analysis/identifyCD4.CD8_byNMF/",
                                "rank", rank_use, "_sharedBy", sharedBy,"/ped_cd8_program_scores.Rds"))

all_scores<- list(myeloid=myeloid_scores, cd4=cd4_scores, cd8=cd8_scores)
```

## Create sample metadata df- metadata for all samples (with at least 1 tcell or myeloid cell)
```{r}
## Use tcell metadata arbitrarly- same for all immune cell seurat objects
sample_meta<- tcell_seurat@meta.data
sample_meta<- sample_meta[!duplicated(sample_meta$sampleid),]
sample_meta<- sample_meta[,c("sampleid", "age_numeric",  "NiceSubtype")]
  
## add in 1 myeloid sample
sample_meta_myeloid<- myeloid_seurat@meta.data[!myeloid_seurat$sampleid %in% sample_meta$sampleid, ]
sample_meta_myeloid$age_numeric<- gsub("yr", "", sample_meta_myeloid$age)
sample_meta<- rbind(sample_meta, sample_meta_myeloid[1, c("sampleid", "age_numeric",  "NiceSubtype")])
rownames(sample_meta)<- sample_meta$sampleid

## Reformat variables to match
sample_meta<- sample_meta %>% dplyr::rename("Sample"="sampleid",
                                            "Subtype"="NiceSubtype") %>% as.data.frame()
sample_meta$age_numeric<- as.numeric(sample_meta$age_numeric)
sample_meta$Subtype<- factor(sample_meta$Subtype, levels=unique(sample_meta$Subtype))

saveRDS(sample_meta, file=paste0(analysis_dir, "all_sample_metadata.Rds"))
```

## Calculate proportion of each program by sample present in myeloid, cd4, cd8, and tumor cells
```{r}
## Clonal CD8
clonal_cd8_prop<- calculate_prop(cd8_seurat, programColumn="clonal")
clonal_cd8_prop<- clonal_cd8_prop[clonal_cd8_prop$Program=="clonal",]

## Myeloid
myeloid_prop<- calculate_prop(myeloid_seurat, programColumn = "NMF_rank6_annotByAdult10X")

## CD8
cd8_prop<- calculate_prop(cd8_seurat, programColumn ="NMF_rank8_annot_mergedByunion_reAnnot" )

## CD4
cd4_prop<- calculate_prop(cd4_seurat, programColumn ="NMF_rank8_annot_mergedByunion_reAnnot" )

## Tumor
tumor_seurat$CellAnnot<- gsub("cycling", "Cycling", tumor_seurat$CellAnnot)
tumor_prop<- calculate_prop(tumor_seurat, sampleColumn = "sample", programColumn = "CellAnnot")

## Tumor without EPN
tumor_seurat_noEPN<- subset(tumor_seurat, subtype!="epn")
tumor_prop_noEPN<- calculate_prop(tumor_seurat_noEPN, sampleColumn = "sample", programColumn = "CellAnnot")

## Create list with all proportions, save
prop_list<- list(clonal_cd8=clonal_cd8_prop, cd8=cd8_prop, cd4=cd4_prop,
                 myeloid=myeloid_prop, tumor=tumor_prop, tumor_noEPN=tumor_prop_noEPN)
saveRDS(prop_list, file=paste0(analysis_dir, "program_proportions_tumor.myeloid.cd4.cd8.cd8Clonal.Rds"))
```


## Process scores for myeloid/tcell programs for use downstream
```{r}
## Create list of scores
all_scores<- list(myeloid=myeloid_scores, cd4=cd4_scores, cd8=cd8_scores)

## Add on sample
all_scores<- lapply(all_scores, function(x){
  x$Sample<- sapply(strsplit(rownames(x), split="CD"), function(y){y[1]})
  return(x)
})

## Remove max program/max score column
all_scores<- lapply(all_scores, function(x){x[,!grepl("Max", colnames(x))]})

## Calculate mean score per sample per program
mean_scores<- lapply(all_scores, function(x){
  x<- x %>% group_by(Sample) %>% summarise(across(everything(), list(mean))) %>% as.data.frame()
  colnames(x)<- gsub("_1", "", colnames(x))
  return(x)
})

## Melt by sample (1 column for all programs)
mean_scores<- lapply(mean_scores, function(x){
  x<- melt(x, id="Sample")
  colnames(x)<- c("Sample", "Program", "MeanScore")
  return(x)
})

## Save
saveRDS(mean_scores, file=paste0(analysis_dir, "program_scores_myeloid.cd4.cd8.Rds"))
```




