---
title: "R Notebook"
output: html_document
---

## Purpose:
Compare proportions for myeloid, tcell, and tumor programs. Do certain programs/cell types correlate with others? 

## Approach:
Using program proportions by sample processed in 01_Preprocess_proportion.scores.Rmd, compare the following using linear regression and pearson correlation:
1. Proportion of cd8 clonal cells vs cd8/cd4/myeloid/tumor programs 
2. Proportion of cd8/cd4/myeloid/tumor vs cd8/cd4/myeloid/tumor programs



```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/", 
                     "04a_Integrate_Immune.Tumor/Correlate_Immune.Tumor_CellType.NMF/")

library(Seurat)
library(ggplot2)
library(ggpubr)
library(randomcoloR)

figure_dir_tmp<- paste0("figures/")
data_dir_tmp<- paste0("analysis/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp, recursive = TRUE)}
if(!dir.exists(data_dir_tmp)){dir.create(data_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)
source(paste0(working_dir, "HelperFunctions.R"))
```

# May need to remove BCOR tumors- not published yet. Set whether to do so (new figure/analysis folder)
```{r}
removeBCOR<- TRUE

figure_dir_tmp<- ifelse(removeBCOR, paste0(figure_dir_tmp, "removeBCOR/"), figure_dir_tmp)
analysis_dir_tmp<- ifelse(removeBCOR, paste0(data_dir_tmp, "removeBCOR/"), data_dir_tmp)

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp, recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}
```


## Set analysis/figure directories. Read in processed files with program proportions, mean scores, and sample metadata
processed in 01_Preprocess_proportionFiles.Rmd
```{r}
cohort<- "pedOnly_nomuv63"
figure_dir<- paste0(figure_dir_tmp, cohort, "/")
data_dir<-paste0(data_dir_tmp, cohort, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
if(!dir.exists(data_dir)){dir.create(analysis_dir)}

## Proportion list
prop_list<- readRDS(paste0(data_dir, "program_proportions_tumor.myeloid.cd4.cd8.cd8Clonal.Rds"))


## Sample metadata
sample_meta<- readRDS(paste0(data_dir, "all_sample_metadata.Rds"))


```

## 1. Proportion of cd8 clonal cells vs cd8/cd4/myeloid/tumor programs 
```{r}
figure_dir_use<- paste0(figure_dir, "clonal/")
if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}

## Input
ind_data<- "myeloid"
dep_data<- "clonal_cd8"
colors_use<- c("cyan", "olivedrab", "navy", "violetred", "orange", "gold", "purple", "grey") ## Colors- for independent variable
confounders_test<- c("age_numeric", "Subtype")

## Merge together
ind_df<- prop_list[[ind_data]]
dep_df<- prop_list[[dep_data]]
ind_df<- ind_df %>% dplyr::rename(indProgram=Program, indPerCells=perCells) %>% dplyr::select(-Ncells)
dep_df<- dep_df %>% dplyr::rename(depProgram=Program, depPerCells=perCells) %>% dplyr::select(-Ncells)
df<- merge(ind_df, dep_df, by="Sample")

## Add on sample metadata
df<- merge(df, sample_meta, by="Sample")


## Determine potential confounders
## RESULTS: subtype and age sig (or close), control for this downstream.
test_covariates<- lm_checkForConfounders(df,confounder_columns = confounders_test ,
                                         dep_value_column = "depPerCells")
pvalues<- lapply(test_covariates, function(x){x$min_pvalue})
controlFor=c("Subtype", "age_numeric")

## Remove any cell types with no cells in any sample used in this analysis
nCells<- df %>% dplyr::group_by_at("indProgram") %>% dplyr::summarise(nCells=sum(indPerCells)) %>% as.data.frame()
cellType_remove<- as.character(nCells$indProgram[nCells$nCells==0])
df<- df[!df$indProgram %in% cellType_remove, ]

## Set colors
ind_programs<-unique(df$indProgram)
names(colors_use)<-ind_programs

## linear regression model for each comparison of interest- return pvalue and r2 for each cell type
all_lm<- linearRegression_multiplePrograms(df, 
                                           dep_value_column="depPerCells",
                                           ind_value_column="indPerCells",
                                           dep_splitter_column="depProgram",
                                           ind_splitter_column="indProgram",
                                           controlFor=controlFor)


all_plots<- plot_Proportions_withLm.res(df, programs=ind_programs, program_variable_name = "indProgram",
                                        colors_use=colors_use, lm_res=all_lm$model_res,
                                        independent_variable = "indPerCells",
                                        dependent_variable = "depPerCells",
                                        ind_variable_plot_name = ind_data,
                                        dep_variable_plot_name = dep_data,
                                        pvalue.x=15,r.x=18)
cowplot::plot_grid(plotlist=all_plots, ncol=1)
ggsave(paste0(figure_dir_use,ind_data, ".vs.", dep_data, "_modelWith", 
             paste(controlFor, collapse = "."), ".png"), width=5, height=12)

```

## 2. Proportion of cd8/cd4/myeloid/tumor vs cd8/cd4/myeloid/tumor programs
```{r}
figure_dir_use<- paste0(figure_dir, "byProportion/")
if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}

## Input
ind_data<- "tumor_noEPN"
dep_data<- "myeloid"
colors_use<- c("cyan", "olivedrab", "navy", "violetred", "orange", "gold", "purple", "grey") ## Colors- for independent variable
confounders_test<- c("age_numeric", "Subtype")

## Merge together
ind_df<- prop_list[[ind_data]]
dep_df<- prop_list[[dep_data]]
ind_df<- ind_df %>% dplyr::rename(indProgram=Program, indPerCells=perCells) %>% dplyr::select(-Ncells)
dep_df<- dep_df %>% dplyr::rename(depProgram=Program, depPerCells=perCells) %>% dplyr::select(-Ncells)
df<- merge(ind_df, dep_df, by="Sample")

## Add on sample metadata
df<- merge(df, sample_meta, by="Sample")

## Remove any cell types with no cells in any sample used in this analysis
nCells<- df %>% dplyr::group_by_at("indProgram") %>% dplyr::summarise(nCells=sum(indPerCells)) %>% as.data.frame()
cellType_remove<- as.character(nCells$indProgram[nCells$nCells==0])
df<- df[!df$indProgram %in% cellType_remove, ]

nCells<- df %>% dplyr::group_by_at("depProgram") %>% dplyr::summarise(nCells=sum(depPerCells)) %>% as.data.frame()
cellType_remove<- as.character(nCells$depProgram[nCells$nCells==0])
df<- df[!df$depProgram %in% cellType_remove, ]

## Run pearson corrleation for all combinations
cor_df<- correlateProportions(df=df, dep_value_column = "depPerCells", ind_value_column = "indPerCells",
                              dep_splitter_column = "depProgram",ind_splitter_column = "indProgram")



## Linear regression for all combinations
  ## Determine which variables to control for
  ## RESULTS: according to this, need to control for age/subtype (also sex, but only sig for 1, so decided not to)
  dependent_programs<-as.character(unique(df$depProgram))
  test_covariates<- lapply(dependent_programs, function(x){
    message(x)
    tmp_input_df<- df[df$depProgram==x,]
    model_res<-lm_checkForConfounders(tmp_input_df, 
                                      confounder_columns =  c("age_numeric", "Subtype"),
                                      dep_value_column =  "depPerCells")
    pvalues<- lapply(model_res, function(y){y$min_pvalue})
    return(unlist(pvalues))
  } ); names(test_covariates)<-dependent_programs
  control_for<- c( "Subtype",  "age_numeric")
  
  ## Run lm for all comparisons, controling for age/sex/subtype
  all_lm<- linearRegression_multiplePrograms(df=df, dep_value_column="depPerCells",ind_value_column="indPerCells",
                                             dep_splitter_column="depProgram", ind_splitter_column="indProgram",
                                             controlFor = control_for)
  all_lm_res<- all_lm$model_res


## Merge pearson correlation and lm results
  ## Add unique id to correlation/lm for merging
  cor_df$ind_v_dep<- paste0(cor_df$ind_variable, "_vs_", cor_df$dep_variable)
  all_lm_res$ind_v_dep<- paste0(all_lm_res$ind_variable, "_vs_", all_lm_res$dep_variable)
  
  ## Order the same way by unique id
  cor_df<- cor_df[order(cor_df$ind_v_dep),]
  all_lm_res<- all_lm_res[order(all_lm_res$ind_v_dep),]
  sum(all_lm_res$ind_v_dep == cor_df$ind_v_dep); length(unique(all_lm_res$ind_v_dep)); length(unique(cor_df$ind_v_dep))
  cor_lm_results<- merge(cor_df[,c("PearsonCorrelation", "ind_v_dep")],
                         all_lm_res[,c("pvalue", "r2", "ind_v_dep")],
                         by="ind_v_dep")
  
  ## Split unique id back into indepedent/dependent
  cor_lm_results$independent_variable<- sapply(strsplit(cor_lm_results$ind_v_dep, split = "_vs_"), function(x){x[1]})
  cor_lm_results$dependent_variable<- sapply(strsplit(cor_lm_results$ind_v_dep, split = "_vs_"), function(x){x[2]})

  ## Add on cutoff for pvalue- will change shape if <0.1
  cor_lm_results$`sig(<0.1)`<- ifelse(cor_lm_results$pvalue<=0.1, "sig", "notSig")  
  

  
## Plot
ggplot(cor_lm_results, aes(x=independent_variable, y=dependent_variable, color=PearsonCorrelation, size=pvalue, shape=`sig(<0.1)`))+
  geom_point()+
  scale_color_gradient2(low="blue", mid="white", high="red", midpoint = 0)+
  scale_size(trans = 'reverse', breaks=c(0, .05, .1, .25, .5, .75), labels=c(0, .05, .1, .25, .5, .75))+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text=element_text(color="black", face="bold"))+
  ylab("")+ xlab("")+
  ggtitle(paste0(ind_data, " vs ", dep_data, " program %"))
ggsave(paste0(figure_dir_use, ind_data, ".vs.", dep_data, "_modelWith",
              paste(control_for, collapse = "."),
              ".png"), width=5, height=4)
                        

```

## 2b: Alternative plot- heatmap by pearson correlation for multiple cell types
```{r}
figure_dir_use<- paste0(figure_dir, "byProportion/")
if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}

## Input- use single independent and all dependent
ind_data<- "cd8" 
dep_data<-names(prop_list)[names(prop_list) != ind_data & names(prop_list) !="tumor"]

## Merge together
ind_df<- prop_list[[ind_data]]
dep_list<- prop_list[names(prop_list) %in% dep_data]

  ## Add major cell type to dep data
  dep_list<- lapply(names(dep_list), function(x){tmp<- dep_list[[x]]; tmp$Program=paste0(x, "_program", tmp$Program); return(tmp)})
  
  ## Merge dep into single df
  dep_df<- do.call("rbind", dep_list)
  
  ## rename columns
  ind_df<- ind_df %>% dplyr::rename(indProgram=Program, indPerCells=perCells) %>% dplyr::select(-Ncells)
  dep_df<- dep_df %>% dplyr::rename(depProgram=Program, depPerCells=perCells) %>% dplyr::select(-Ncells)
  
  ## Merge
  df<- merge(ind_df, dep_df, by="Sample")

## Add on sample metadata
df<- merge(df, sample_meta, by="Sample")

## Remove any cell types with no cells in any sample used in this analysis
nCells<- df %>% dplyr::group_by_at("indProgram") %>% dplyr::summarise(nCells=sum(indPerCells)) %>% as.data.frame()
cellType_remove<- as.character(nCells$indProgram[nCells$nCells==0])
df<- df[!df$indProgram %in% cellType_remove, ]

nCells<- df %>% dplyr::group_by_at("depProgram") %>% dplyr::summarise(nCells=sum(depPerCells)) %>% as.data.frame()
cellType_remove<- as.character(nCells$depProgram[nCells$nCells==0])
df<- df[!df$depProgram %in% cellType_remove, ]

## Run pearson corrleation for all combinations
cor_df<- correlateProportions(df=df, dep_value_column = "depPerCells", ind_value_column = "indPerCells",
                              dep_splitter_column = "depProgram",ind_splitter_column = "indProgram")

## Split dependent variable into program + major cell type
cor_df$dep_majorCellType<- apply(cor_df, 1, function(x){unlist(strsplit(x["dep_variable"], split="_program"))[1]})
cor_df$dep_variable<- apply(cor_df, 1, function(x){unlist(strsplit(x["dep_variable"], split="_program"))[2]})
  
## Make major cell type names nicer
cor_df$dep_majorCellType<- gsub("cd4", "CD4+", gsub("clonal_", "",
                           gsub("cd8", "CD8+", 
                           gsub("tumor_noEPN", "Tumor", 
                           gsub("myeloid", "Myeloid",cor_df$dep_majorCellType)))))

ggplot(cor_df, aes(x=ind_variable, y=dep_variable, fill=PearsonCorrelation))+
  geom_tile()+
      theme_bw()+
  facet_grid(rows=vars(dep_majorCellType), scale="free_y", space="free_y")+
  scale_fill_gradient2(low="blue",mid="white", high = "red")+
   theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text=element_text(color="black", face="bold"),
        axis.title=element_text(color="black", face="bold"),
        strip.text = element_text(color="black", face="bold"))+
  xlab(paste0(gsub("_noEPN", "", ind_data), " cell types"))+ ylab("")

ggsave(paste0(figure_dir_use, "Heatmap_PearsonCor_", ind_data, ".vs.",paste(dep_data, collapse = "."), ".png"), width=5, height=6)
```




