---
title: "R Notebook"
output: html_document
---

## Purpose:
Compare scores for myeloid, tcell, and tumor programs. Do certain programs/cell types correlate with others? 

## Approach:
Using program proportions/scores by sample processed in 01_Preprocess_proportionFiles.Rmd, compare mean scores by sample for cd8/cd4 vs myeloid programs  using linear regression and pearson correlation


```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/", 
                     "04a_Integrate_Immune.Tumor/Correlate_Immune.Tumor_CellType.NMF/")

library(Seurat)
library(ggplot2)
library(ggpubr)
library(randomcoloR)

figure_dir_tmp<- paste0("figures/")
data_dir_tmp<- paste0("analysis/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp, recursive = TRUE)}
if(!dir.exists(data_dir_tmp)){dir.create(data_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)
source(paste0(working_dir, "HelperFunctions.R"))
```

## May need to remove BCOR tumors- not published yet. Set whether to do so (new figure/analysis folder)
```{r}
removeBCOR<- TRUE

figure_dir_tmp<- ifelse(removeBCOR, paste0(figure_dir_tmp, "removeBCOR/"), figure_dir_tmp)
analysis_dir_tmp<- ifelse(removeBCOR, paste0(data_dir_tmp, "removeBCOR/"), data_dir_tmp)

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp, recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}
```

## Set analysis/figure directories. Read in processed files with  mean scores and sample metadata
processed in 01_Preprocess_proportionFiles.Rmd
```{r}
cohort<- "pedOnly_nomuv63"
figure_dir<- paste0(figure_dir_tmp, cohort, "/byScores/")
analysis_dir<-paste0(analysis_dir_tmp, cohort, "/")
data_dir<-paste0(data_dir_tmp, cohort, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
if(!dir.exists(data_dir)){dir.create(analysis_dir)}


## Mean scores by sample
score_list<- readRDS(paste0(data_dir, "program_scores_myeloid.cd4.cd8.tumor.Rds"))


## Sample metadata
sample_meta<- readRDS(paste0(data_dir, "all_sample_metadata.Rds"))
```

## Remove BCOR samples if needed
```{r}
if(removeBCOR){
  sample_meta<- sample_meta[sample_meta$Subtype != "Hemispheric-BCOR",]
  score_list<- lapply(score_list, function(x){x[x$Sample %in% sample_meta$Sample,]})
}
```

## Reformat scores into single data frame
```{r}
## Add on broad cell type
score_list_withCellType<- lapply(names(score_list), function(x){
  tmp<- score_list[[x]]
  tmp$cellType<-x
  tmp$Program<- paste0(x, "_", tmp$Program)
  return(tmp)
})
names(score_list_withCellType)<- names(score_list)

## Merge into single df
score_df<- do.call("rbind", score_list_withCellType)
score_df$Program<- as.character(score_df$Program)

## remove "-" (issues with lm downstream)
score_df$Program<- gsub("-", "_", score_df$Program)

```


## Generate df of scores for each pair of programs
```{r}
## Get all program pairs
programs<- unique(score_df$Program)
program_pairs<- apply(combn(programs,2),2,paste,collapse='__')

## manually add "pairs" for same program
#program_pairs<- c(program_pairs, unname(sapply(programs, function(x){paste0(x, "__", x)})))

## Cycle through all pairs, get samples with both programs, reformat into single df for each pair
all_pair_score_df<- lapply(program_pairs, function(x){
  ## Subset to pair
  pair<- unlist(strsplit(x, split="__"))
  score_tmp<- score_df[score_df$Program %in% pair, ]
  
  ## remove rows without 2 samples (i.e., found in both programs in pair)
  nSamples<- table(score_tmp$Sample)
  score_tmp<- score_tmp[score_tmp$Sample %in% names(nSamples)[nSamples==2],]
  
  ## reformat: 1 df with columns for both
  score_split<- split(score_tmp, f=score_tmp$Program)
  score_split<- lapply(score_split, function(x){
    tmp<-x
    colnames(tmp)<-c("Sample", "Program", unique(tmp$Program), "cellType")
    tmp$Program<-NULL; tmp$cellType<-NULL
    return(tmp)
  })
  score_tmp<- merge(score_split[[1]], score_split[[2]], by="Sample")
  return(score_tmp)
})
names(all_pair_score_df)<- program_pairs

saveRDS(all_pair_score_df, file=paste0(analysis_dir, "programScores_splitByPairs.Rds"))
```

## Pearson correlation for each pair of programs
```{r}
all_pair_score_df<- readRDS(paste0(analysis_dir, "programScores_splitByPairs.Rds"))
all_pearson<- sapply(all_pair_score_df, function(x){
  cor(x[,2], x[,3], method="pearson")
})
```


## linear regression for each pair of programs
```{r}
all_pair_score_df<- readRDS(paste0(analysis_dir, "programScores_splitByPairs.Rds"))
all_lm<- lapply(all_pair_score_df, function(x){
  #print(colnames(x))
  runLm<- paste0("lm(", colnames(x)[2], "~", colnames(x)[3], ", x)")
  lm_res<-summary(eval(parse(text=runLm)))
  return(list(model=lm_res,
              r2=lm_res$r.squared,
              pvalue=lm_res$coefficients[colnames(x)[3], "Pr(>|t|)"]))
    
})

## vector of pvalues
all_pvalues<- sapply(all_lm, function(x){x$pvalue})
all_r2<- sapply(all_lm, function(x){x$r2})
```

## Plot heatmap of pearson correlation
```{r}
## reformat into dataframe
pearson_df<- as.data.frame(all_pearson)
pearson_df$p1<- sapply(rownames(pearson_df), function(x){unlist(strsplit(x, split="__"))[1]})
pearson_df$p2<- sapply(rownames(pearson_df), function(x){unlist(strsplit(x, split="__"))[2]})

## add on broad cell type
pearson_df$broad1<- gsub("_.*", "", pearson_df$p1)
pearson_df$broad2<- gsub("_.*", "", pearson_df$p2)

## remove broad cell type from program name
pearson_df$p1<- gsub("cd4_", "", gsub("cd8_", "", gsub("myeloid_", "", gsub("tumor_", "", pearson_df$p1))))
pearson_df$p2<- gsub("cd4_", "", gsub("cd8_", "", gsub("myeloid_", "", gsub("tumor_", "", pearson_df$p2))))

## Optional- remove correlations between same broad cell types
pearson_df<- pearson_df[pearson_df$broad1 != pearson_df$broad2,]


## Order programs- reverse p2 so that it aligns with p1 in heatmap
#pearson_df$p1<- factor(pearson_df$p1, levels=unique(pearson_df$p1))
#pearson_df$p2<- factor(pearson_df$p2, levels=rev(unique(pearson_df$p1)))

## Add on sig from linear regression (used to add box on plot)
pearson_df$pvalue<- as.numeric(plyr::mapvalues(rownames(pearson_df), names(all_pvalues), all_pvalues, warn_missing = FALSE))
pearson_df$sig<- ifelse(pearson_df$pvalue<0.01, "verySig",ifelse(pearson_df$pvalue<0.05, "sig", "notSig")) 


## Change cell type to upper case for nicer plotting
pearson_df$broad1<- gsub("cd4", "CD4", gsub("cd8","CD8", gsub("myeloid", "Myeloid", gsub("tumor", "Tumor", pearson_df$broad1))))
pearson_df$broad2<- gsub("cd4", "CD4", gsub("cd8","CD8", gsub("myeloid", "Myeloid", gsub("tumor", "Tumor", pearson_df$broad2))))

## Order by cell type 
pearson_df$broad1<- factor(pearson_df$broad1, levels=c( "Myeloid","CD4", "CD8"))
pearson_df$broad2<- factor(pearson_df$broad2, levels=rev(c( "Tumor", "CD8", "CD4")))

pearson_df<- pearson_df %>% dplyr::rename(PearsonCorrelation=all_pearson) %>% as.data.frame()

## Plot
ggplot(pearson_df, aes(x=p1, y=p2, fill=PearsonCorrelation))+
  geom_tile()+
  theme_bw()+
  scale_fill_gradient2(low="blue", mid="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"),
        strip.text = element_text(face="bold", color="black", size=12))+
  facet_grid(cols=vars(broad1), rows=vars(broad2), scales="free", space="free")+
  geom_tile(data = pearson_df[pearson_df$sig=="sig",], fill = NA, color = "black", size = .5) +
    geom_tile(data = pearson_df[pearson_df$sig=="verySig",], fill = NA, color = "black", size = 1.5) +
  xlab("") + ylab("")
ggsave(paste0(figure_dir, "Heatmap_pearsonCorr_lmForSig_cd4.cd8.myeloid.tumor.Scores.pdf"),device="pdf", width=7.5, height=5)
```


## 3. Mean scores by sample for cd8/cd4 vs myeloid programs 
```{r}
## Input
ind_data<- "tumor"
dep_data<- "myeloid"
colors_use<- c("cyan", "olivedrab", "navy", "violetred", "orange", "gold", "purple", "grey") ## Colors- for independent variable
confounders_test<- c("age_numeric", "Subtype")

## Merge together
ind_df<- score_list[[ind_data]]
dep_df<- score_list[[dep_data]]
ind_df<- ind_df %>% dplyr::rename(indProgram=Program, indMeanScore=MeanScore) 
dep_df<- dep_df %>% dplyr::rename(depProgram=Program, depMeanScore=MeanScore) 
df<- merge(ind_df, dep_df, by="Sample")

## Add on sample metadata
df<- merge(df, sample_meta, by="Sample")

## Run pearson corrleation for all combinations
cor_df<- correlateProportions(df=df, dep_value_column="depMeanScore",ind_value_column="indMeanScore",
                                dep_splitter_column="depProgram",ind_splitter_column="indProgram")

## then do lm test
  ## Determine which variables to control for
  ## RESULTS: according to this, need to control for age/subtype (also sex, but only sig for 1, so decided not to)
  dependent_programs<-as.character(unique(df$depProgram))
  test_covariates<- lapply(dependent_programs, function(x){
    message(x)
    tmp_input_df<- df[df$depProgram==x,]
    model_res<-lm_checkForConfounders(tmp_input_df, 
                                      confounder_columns =  c("age_numeric","Subtype"),
                                      dep_value_column =  "depMeanScore")
    pvalues<- lapply(model_res, function(y){y$min_pvalue})
    return(unlist(pvalues))
  } ); names(test_covariates)<-dependent_programs
  control_for<- c( "Subtype", "age_numeric")

  ## Run lm for all comparisons, controling for age/sex/subtype
  all_lm<- linearRegression_multiplePrograms(df=df, dep_value_column="depMeanScore",ind_value_column="indMeanScore",
                                             dep_splitter_column="depProgram",ind_splitter_column="indProgram",
                                             controlFor = control_for)
  all_lm_res<- all_lm$model_res
                        
  
## Merge pearson correlation and lm results
  ## Add unique id to correlation/lm for merging
  cor_df$ind_v_dep<- paste0(cor_df$ind_variable, "_vs_", cor_df$dep_variable)
  all_lm_res$ind_v_dep<- paste0(all_lm_res$ind_variable, "_vs_", all_lm_res$dep_variable)
  
  ## Order the same way by unique id
  cor_df<- cor_df[order(cor_df$ind_v_dep),]
  all_lm_res<- all_lm_res[order(all_lm_res$ind_v_dep),]
  sum(all_lm_res$ind_v_dep == cor_df$ind_v_dep); length(unique(all_lm_res$ind_v_dep)); length(unique(cor_df$ind_v_dep))
  cor_lm_results<- merge(cor_df[,c("PearsonCorrelation", "ind_v_dep")],
                         all_lm_res[,c("pvalue", "r2", "ind_v_dep")],
                         by="ind_v_dep")
  
  ## Split unique id back into indepedent/dependent
  cor_lm_results$independent_variable<- sapply(strsplit(cor_lm_results$ind_v_dep, split = "_vs_"), function(x){x[1]})
  cor_lm_results$dependent_variable<- sapply(strsplit(cor_lm_results$ind_v_dep, split = "_vs_"), function(x){x[2]})

  ## Add on cutoff for pvalue- will change shape if <0.1
  cor_lm_results$`sig(<0.1)`<- ifelse(cor_lm_results$pvalue<=0.1, "sig", "notSig")

## Plot
ggplot(cor_lm_results, aes(x=independent_variable, y=dependent_variable, color=PearsonCorrelation, size=pvalue, shape=`sig(<0.1)`))+
  geom_point()+
  scale_color_gradient2(low="blue", mid="white", high="red", midpoint = 0)+
  scale_size(trans = 'reverse', breaks=c(0, .05, .1, .25, .5, .75), labels=c(0, .05, .1, .25, .5, .75))+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text=element_text(color="black", face="bold"))+
  ylab("")+ xlab("")+
  ggtitle(paste0(ind_data, " vs ", dep_data))
ggsave(paste0(figure_dir,ind_data,".vs.", dep_data, "_modelWith",
              paste(control_for, collapse = "."), ".png"), width=5, height=4)
```


