---
title: "NMF.Rmd"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Identify metaprograms for myeloid cells

## Approach:
Run de novo NMF (10x repeats) for pediatric myeloid. 

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02b_Myeloid/01a_NMF/")

integration<- "HarmonyInt/" ## HarmonyInt/ or ""
analysis_dir_tmp<- paste0(working_dir, "/analysis/de.novoNMF/", integration, "/")
figure_dir_tmp<- paste0(working_dir, "/figures/de.novoNMF/", integration, "/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp)};if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

library(reticulate) 
use_python("/Users/jlabelle/miniconda3/bin/python", required = T)  # specify python in my conda env for scrublet
use_condaenv("base", conda = "/Users/jlabelle/miniconda3/condabin/conda")  # reticulate library to load the conda environnment for running
library(Seurat)

source(paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/", "/Scripts/plotutils.R"))
source(paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/", "/Scripts/seurat3utils.R"))
source(paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/", "/Scripts/color.R"))

tests <- c("wilcox", "t", "roc", "MAST", "bimod")
test <- tests[2]
```

## Read in seurat object + original markers + original annotations
```{r}
cohort<-"pedOnly_nomuv63"

## Current seurat object
myeloid_seurat<-readRDS(paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",integration, "/",
                               cohort, "/myeloid_seurat_noDC.B.Rds"))

## Original seurat object (with cell annotations)
backup<- myeloid_seurat
load("../../../Preprocessing_ByOrr/Preprocessing/analysis/WithNewSamples/Myeloid_seurat_obj.Robj")
original_myeloid_seurat<- myeloid_seurat
myeloid_seurat<- backup; rm(backup)

## Original myeloid NMF marker genes
original_markers<- read.csv("../../../Preprocessing_ByOrr/data/pedbrain_immune_Myeloid_top_DE_genes_NMF_rank6.csv",
                            row.names = "X")

analysis_dir<- paste0(analysis_dir_tmp, cohort, "/")
figure_dir<- paste0(figure_dir_tmp, cohort, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}
```

## Clustering using non-negative matrix factorization (NMF)
Cluster data using NMF with a range of ranks. The rank is the number of metagenes and the number of clusters. The choice of the rank is heuristic. For each clustering, we also calculate DE genes.
```{r NMF, eval = F}
# Vector of ranks for which to calculate NMF.
rank <- 5:7

# Identify variable genes within this cell subset.
var.genes <- SelectIntegrationFeatures(SplitObject(myeloid_seurat, split.by = "sampleid"), nfeatures = 2000, verbose = TRUE,
                                       fvf.nfeatures = 2000, selection.method = "vst")

VariableFeatures(myeloid_seurat) <- var.genes

# Number of variable genes
print(length(VariableFeatures(myeloid_seurat)))

# Perform 10 NMF runs for each value of rank r.
x <- as.matrix(myeloid_seurat[['RNA']][VariableFeatures(myeloid_seurat), ])
x <- x[which(rowSums(x) > 0), ]  # remove any rows filled with 0 (just in case these exist)

estim.r <- nmf(x, rank = rank, method = "brunet", nrun = 50, seed = 123456, .options = "tvP")
png(paste0(figure_dir, "_NMF_plot_rank.png"), width = 2000, height = 1000, res = 100)
plot(estim.r)
dev.off()

# Iterate over different ranks and look at clustering.
for (i in seq(rank)) {
  r <- rank[i]
  
  # Plot coefficient values of metagenes for each cell.
  png(paste0(figure_dir, "_NMF_coefmap_rank", r, ".png"), width = 2000, height = 1000, res = 100)
  coefmap(estim.r$fit[[i]], Colv=T)
  dev.off()

  # Show genes contributing the most to each metagene (factor). 
  # Remember, this is limited to the variable genes if that was the input to NMF.
  for (j in seq(r)) {
    top.features <- names(sort(estim.r$fit[[i]]@fit@W[, j], decreasing = T)[1:10])
    print(paste(top.features, collapse = " "))
  }
  s <- extractFeatures(estim.r$fit[[i]], 10L)
  for (j in seq(r)) {
    top.features <- rownames(estim.r$fit[[i]]@fit@W)[s[[j]]]
    print(paste(top.features, collapse = " "))
  }

  # Plot NMF residuals.
  png(paste0(figure_dir, "_NMF_residuals_rank", r, ".png"), width = 2000, height = 1000, res = 100)
  plot(estim.r$fit[[i]])
  dev.off()

  # Check coefficient matrix has cell names in same order as Seurat object myeloid_seurat.
  H <- estim.r$fit[[i]]@fit@H
  all.equal(as.integer(as.character(predict(estim.r$fit[[i]], what = "samples"))), as.integer(apply(H, 2, which.max)))
  all.equal(colnames(H), colnames(myeloid_seurat))
  all.equal(rownames(estim.r$fit[[i]]@fit@W), VariableFeatures(myeloid_seurat))
  
  # Cluster cells based on maximum coefficient for metagene.
  myeloid_seurat[[paste0("NMF_rank", r)]] <- predict(estim.r$fit[[i]], what = "samples")
  Idents(myeloid_seurat) <- paste0("NMF_rank", r)

  # UMAP visualization and myeloid and myeloid expression across clusters and find DE genes.
  DimPlot(myeloid_seurat, reduction = "umap", cols = material.heat(n_distinct(Idents(myeloid_seurat))), label = T)
  ggsave(paste0(figure_dir,  "UMAP_NMF_rank", r, ".png"), width = 6, height = 6)
  VlnPlot(myeloid_seurat, features = c("CSF1R", "CD3G"), cols = material.heat(n_distinct(Idents(myeloid_seurat))), 
          ncol = 1, pt.size = 0.5) 
  ggsave(paste0(figure_dir, "vln_NMF_rank", r, ".png"), width = 8, height = 8)
  
  # Find differentially expressed genes for each cluster.
  myeloid_seurat.markers <- FindAllMarkers(myeloid_seurat, min.pct = 0.1, logfc.threshold = 0.5, 
                                           max.cells.per.ident = 1500, test.use = test, 
                                           return.thresh = 0.1 / nrow(myeloid_seurat), verbose = T)
  top10genes <- myeloid_seurat.markers %>% filter(avg_log2FC > 0) %>% group_by(cluster) %>% top_n(10, avg_log2FC)
  top100genes <- myeloid_seurat.markers %>% filter(avg_log2FC > 0) %>% group_by(cluster) %>% top_n(100, avg_log2FC)
  
  # Write cluster markers to csv file.
  write.csv(top100genes, file = paste0(analysis_dir,  "top_DE_genes_NMF_rank", r, ".csv"))
  myeloid_seurat <- ScaleData(myeloid_seurat, features = c(top10genes$gene), do.center = T, do.scale = F)
  DoHeatmap(myeloid_seurat, features = top10genes$gene) + scale_fill_distiller(palette = "RdYlBu")
  ggsave(paste0(figure_dir,  "DE_genes_expression_rank", r, ".png"), width = 18, height = 24)
}


saveRDS(estim.r, file=paste0(analysis_dir, "estimR_rank5-7.Rds"))
saveRDS(myeloid_seurat, file=paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",integration, "/",
                               cohort, "/myeloid_seurat_noDC.B.Rds"))
```

## Correlate with original programs to help annotate new programs (not used downstream)
```{r}
## Norm/center current cm
cm_list<- NormCenter(myeloid_seurat@assays$RNA@counts)
cm_mean<- log2(rowMeans(cm_list$raw_data)+1)

## Rename previous NMF programs --> annotations
original_markers<- split(original_markers, f=original_markers$cluster)
original_markers<- lapply(original_markers, function(x){x$gene})
nmf_annots<- c(1:6)
names(nmf_annots)<- c("MES_Mono", "RiboActive", "Stress_Response", "IFN_Response", "Phago_Lipid_Mg", "Inflammatory")
names(original_markers)<- plyr::mapvalues(names(original_markers), nmf_annots, names(nmf_annots))

## Try for rank 5-7
ranks<- 5:7
all_scores<-list()
for(rank in ranks){
  print(rank)
  
  ## Read in current marker genes- rank 6. Convert to list
  current_markers<- read.csv(paste0(analysis_dir, "top_DE_genes_NMF_rank", rank, ".csv"))
  current_markers<- split(current_markers, f=current_markers$cluster)
  current_markers<- lapply(current_markers, function(x){x$gene})
  names(current_markers)<- paste0("NewNMF_", names(current_markers))
  
  ## Score current cm for both
  all_markers<- c(current_markers, original_markers)
  scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, all_markers)))
  all_scores[[paste0("rank", rank)]]<- scores
  
  ## Correlate and plot heatmap
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  pheatmap(scores_factor_cor, color = hm_colors(100), 
           cluster_rows = F, cluster_cols = F, 
           annotation_names_row = F, annotation_names_col =T,
           show_rownames = T, show_colnames = F,
           filename = paste0(figure_dir, "Cor_NewNMFrank", rank, "_OriginalPrograms.png"),
           width = 8, height = 4)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir, "nmf_factors_hc", rank, ".png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
  
}


saveRDS(all_scores, file=paste0(analysis_dir, "scores_NewNMF.OriginalPrograms_rank5-7.Rds"))
```

## Assign each new program to highest correlated previous program
## However- the original analysis included Bcells and DC. This seems to change annotation significantly
## These annotations may not be used downstream.
Current approach: use original programs to annotate ONLY Stress_Response and Inflammatory. Leave others as NMF1/NMF2, etc
Then correlate with adult GBM 10x (in AdultGBM10X_AnnotatePed.Rmd) to annotate remaining programs

## Annotate based on correlation with original (for Inflammatory and Stress_Response)
```{r}
ranks<-6:7

  ## pedOnly
  annot_by_rank<- list(rank5=c("NMF1", "Inflammatory", "NMF3", "Stress_Response", "NMF5"),
                       rank6=c("NMF1", "Inflammatory", "Stress_Response", "NMF4", "NMF5", "NMF6"),
                       rank7=c("NMF1", "Inflammatory", "NMF3","Stress_Response", "NMF5", "NMF6", "NMF7"),
                       rank8=c("NMF1", "NMF2","Inflammatory", "MES_Mono", "NMF5", "Stress_Response","NMF7", "IFN_Response"),
                       rank9=c("IFN_Response", "NMF2", "Stress_Response", "NMF4","Inflammatory", "NMF6", "MES_Mono", "NMF8", "RiboActive"))
  
  ## pedOnly_nomuv63.withmuv91
  annot_by_rank<- list(rank6=c("Inflammatory", "NMF2", "NMF3", "Stress_Response", "NMF5", "NMF6"),
                       rank7=c("NMF1", "NMF2", "NMF3", "Stress_Response", "NMF5", "NMF6", "Inflammatory"))
  
  ## pedOnly_nomuv63
  annot_by_rank<- list(rank6=c("NMF1", "NMF2", "NMF3", "Stress_Response", "NMF5", "Inflammatory"),
                       rank7=c("Stress_Response", "NMF2", "NMF3", "NMF4", "Inflammatory", "NMF6", "NMF7"))
  
  ## pedOnly_novmuv63 + harmony integration
  annot_by_rank<- list(rank6=c("NMF1", "NMF2", "NMF3", "NMF4", "Stress_Response", "NMF6"),
                       rank7=c("NMF1", "NMF2", "Inflammatory", "NMF4", "NMF5", "NMF6", "Stress_Response"))
  

for(rank in ranks){
  print(rank)
  
  ## Read in NMF marker genes
  marker_genes<- read.csv(paste0(analysis_dir,  "top_DE_genes_NMF_rank", rank, ".csv"))
  
  ## Annotate
  annots<- annot_by_rank[[paste0("rank", rank)]]
  names(annots)<- paste0(1:rank)
  
  marker_genes$cluster<- plyr::mapvalues(marker_genes$cluster,  names(annots),annots)
  myeloid_seurat@meta.data[[paste0("NMF_rank", rank, "_annot")]]<- plyr::mapvalues(myeloid_seurat@meta.data[[paste0("NMF_rank",
                                                                                                                    rank)]],
                                                                                    names(annots),annots)
  
  ## Save marker genes with annotations
  write.csv(marker_genes, file=paste0(analysis_dir,  "top_DE_genes_annotated_NMF_rank", rank, ".csv"))
}


saveRDS(myeloid_seurat, file=paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",integration, "/",
                               cohort, "/myeloid_seurat_noDC.B.Rds"))

## Save annotations
annot_by_rank_named<- lapply(annot_by_rank, function(x){names(x)<-paste0("NMF", 1:length(x)); return(x)})
saveRDS(annot_by_rank_named, file=paste0(analysis_dir, "NMF_annotations_rank", ranks[1], "-", ranks[length(ranks)], ".Rds"))

```



## Alternative to using NMF to directly assign programs to cell: score cells for NMF marker genes, then assign max program
This may be preferrable to using NMF assignments directly, as downstream comparisons to adult/etc will involve scoring cells
Score cells for each rank, assign max, add to seurat
```{r}
ranks<- 6:7

cm_list<-NormCenter(myeloid_seurat@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data + 1))
all_scores<- list()
for(rank in ranks){
  print(rank)

  ## Read in marker genes with annotations
  marker_genes<- read.csv(paste0(analysis_dir,  "top_DE_genes_annotated_NMF_rank", rank, ".csv"))
  marker_genes<- split(marker_genes, f=marker_genes$cluster); marker_genes<- lapply(marker_genes, function(x){x$gene})
  
  ## score cells
  scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, marker_genes)))
  
  ## Identify max program/score
  scores$MaxProgram<- apply(scores, 1, function(x){names(x)[which.max(x)]})
  scores$MaxScore<- apply(scores[,colnames(scores)!="MaxProgram"], 1, function(x){x[which.max(x)]})
  all_scores[[paste0("rank", rank)]]<- scores
  
  ## Add to seurat object
  myeloid_seurat@meta.data[[paste0("NMF_rank", rank, "byScore_annot")]]<- plyr::mapvalues(colnames(myeloid_seurat),
                                                                               rownames(scores), scores$MaxProgram)
  myeloid_seurat@meta.data[[paste0("NMF_rank", rank, "byScore_score")]]<- plyr::mapvalues(colnames(myeloid_seurat),
                                                                              rownames(scores), scores$MaxScore)
  
}

saveRDS(myeloid_seurat, file=paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",integration, "/",
                               cohort, "/myeloid_seurat_noDC.B.Rds"))
saveRDS(all_scores, file=paste0(analysis_dir, "NMF_scores_rank",ranks[1], "-", ranks[length(ranks)],".Rds"))
```



## For non-harmony integrated ONLY: integrate with harmony, then add on non-harmony-integrated annotations
```{r}
noInt_seurat<- readRDS(paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",
                               cohort, "/myeloid_seurat_noDC.B.Rds"))

## Run harmony integration
noInt_meta<- noInt_seurat@meta.data
harmony_seurat<- RunFullSeurat(cm=noInt_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=noInt_meta$sampleid)
harmony_seurat<- AddMetaData(harmony_seurat, noInt_meta[,!colnames(noInt_meta) %in% colnames(harmony_seurat@meta.data)])

saveRDS(harmony_seurat, file=paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",
                               cohort, "/myeloid_seurat_noDC.B_harmony.Rds"))
```

