---
title: "NMF with cycling set as program"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Set cycling cells as a separate "program". Re-generate all program data with this updated program list (seurat object, scores, markers)

## Approach:
Proliferating myeloid cells were already identified by correlating with adult GBM 10X data, but the subset was too small to be represented by NMF program. Take these cells and assign to cycling "program". Set as ident, then identify markers, score, add to seurat.


```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/NMF/Myeloid/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/NMF_withCycling/")
figure_dir_tmp<- paste0(working_dir, "/figures/NMF_withCycling/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp)};if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)



source(paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/", "/Scripts/plotutils.R"))
source(paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/", "/Scripts/seurat3utils.R"))
source(paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/", "/Scripts/color.R"))

tests <- c("wilcox", "t", "roc", "MAST", "bimod")
test <- tests[2]
```

## Read in seurat object + original markers + original annotations
```{r}
cohort<-"pedOnly_nomuv63"

## Current seurat object
myeloid_seurat<-readRDS(paste0("../../Preprocessing/Counts_Preprocess/analysis/detailed_annotation/",
                               cohort, "/myeloid_seurat_noDC.B.Rds"))


analysis_dir<- paste0(analysis_dir_tmp, cohort, "/")
figure_dir<- paste0(figure_dir_tmp, cohort, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}
```

## Create new variable in metadata with cycling cells included
```{r}
rank_use<-6
variable_name<- paste0("NMF_rank", rank_use, "byScore_annotByAdult10X")
new_variable_name<-  paste0("NMF_rank", rank_use, "_annotByAdult10X_withCycling")

myeloid_seurat@meta.data[[new_variable_name]]<-ifelse(myeloid_seurat$prol_tam=="highScore_prolTAM", "Cycling",
                                                                         myeloid_seurat@meta.data[[variable_name]])

saveRDS(myeloid_seurat, file=paste0("../../Preprocessing/Counts_Preprocess/analysis/detailed_annotation/",
                               cohort, "/myeloid_seurat_noDC.B.Rds"))
```

## Plot cycling genes to support cycling annotation
```{r}
all_plots<- lapply(c("cycling_score", "TOP2A"), function(x){
  FeaturePlot(myeloid_seurat, features=x)+scale_color_gradient(low="grey", high="red")
})
cowplot::plot_grid(plotlist=all_plots, ncol=2)
ggsave(paste0(figure_dir, "CyclingScore_TOP2A.png"), width=10, height=4.5)
```

## Identify new marker genes based on NMF + cycling
```{r}
rank_use<-6
variable_name<- paste0("NMF_rank", rank_use, "_annotByAdult10X_withCycling")

## Set ident to programs with cycling
myeloid_seurat<-SetIdent(myeloid_seurat, value = myeloid_seurat@meta.data[[variable_name]])

## Identify marker genes
all_markers<- FindAllMarkers(myeloid_seurat, min.pct = 0.1, logfc.threshold = 0.5, 
                                           max.cells.per.ident = 1500, test.use = test, 
                                           return.thresh = 0.1 / nrow(myeloid_seurat), verbose = T)
top100genes <-all_markers %>% filter(avg_log2FC > 0) %>% group_by(cluster) %>% top_n(100, avg_log2FC)

write.csv(top100genes, file = paste0(analysis_dir,  "top_DE_genes_10Xannotated_NMF_rank", rank_use, ".csv"))
```

## Score cells for new marker genes, assign max
```{r}
rank_use<- 6
marker_genes<- read.csv(paste0(analysis_dir,  "top_DE_genes_10Xannotated_NMF_rank", rank_use, ".csv"))
marker_list<- split(marker_genes, f=marker_genes$cluster); marker_list<- lapply(marker_list, function(x){x$gene})

variable_name<-paste0("NMF_rank", rank_use, "_annotByAdult10X_withCycling")
new_variable_name<- paste0("NMF_rank", rank_use, "byScore_annotByAdult10X_withCycling")

myeloid_seurat<- readRDS(paste0("../../Preprocessing/Counts_Preprocess/analysis/detailed_annotation/",
                               cohort, "/myeloid_seurat_noDC.B.Rds"))

## Score for programs with cycling
cm_list<- NormCenter(myeloid_seurat@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data + 1))
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, marker_list)))
scores$MaxProgram<- apply(scores, 1, function(x){names(x)[which.max(x)]})
scores$MaxScore<- apply(scores[,colnames(scores) != "MaxProgram"], 1, function(x){x[which.max(x)]})

## Add max program to seurat
myeloid_seurat@meta.data[[new_variable_name]]<- plyr::mapvalues(colnames(myeloid_seurat), rownames(scores), scores$MaxProgram)

saveRDS(scores, file=paste0(analysis_dir, "NMF_scores_rank",rank_use,".Rds"))
saveRDS(myeloid_seurat, file=paste0("../../Preprocessing/Counts_Preprocess/analysis/detailed_annotation/",
                               cohort, "/myeloid_seurat_noDC.B.Rds"))
```



## colors for all follow up
```{r}
## pedOnly_nomuv63
color_annots<- list(rank6=c(SEPP1_Mo_TAM ="orange", Inflammatory="red", IFN_Mg_TAM="purple", Stress_Response="olivedrab4",
                            Monocytes="turquoise3", Mo_TAM="navy", Cycling="grey22"))
saveRDS(color_annots, paste0(analysis_dir, "annot_colors.Rds"))
```

## Plot heatmaps of top marker gene expression
```{r}
rank_use<- "6"
add_lines<- TRUE
variable_name<- paste0("NMF_rank", rank_use, "byScore_annotByAdult10X_withCycling")

## Read in NMF marker genes
marker_genes<- read.csv(paste0(analysis_dir,  "top_DE_genes_10Xannotated_NMF_rank", rank, ".csv"))
marker_genes$cluster<- gsub("_", "\n", marker_genes$cluster)

## Subset to top markers for each nmf- top 10
## If any genes are top markers for multiple clusters- remove, then find top 10 again
top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
if(sum(duplicated(top_markers$gene))!=0){
  duplicated<- top_markers$gene[duplicated(top_markers$gene)]
  print(paste0("Shared top markers: ", paste(duplicated, collapse=",")))
  marker_genes<- marker_genes[!marker_genes$gene %in% duplicated,]
  top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
}

  ## optional: manually add in SPP1 to show where it falls
  top_markers<- rbind(top_markers,marker_genes[marker_genes$gene=="SPP1" & marker_genes$cluster=="IFN\nMg\nTAM",])
  geneRemove<- top_markers[top_markers$cluster=="IFN\nMg\nTAM", "gene"][10]
  top_markers<- top_markers[top_markers$gene!=geneRemove,]
  top_markers<- top_markers[order(top_markers$cluster),]

## Pseudobulk
myeloid_seurat$ShortSubtype<- gsub("\\\n", "-", myeloid_seurat$ShortSubtype)
myeloid_seurat@meta.data[[variable_name]]<- gsub("\\.", "\n",gsub("_", "\n", myeloid_seurat@meta.data[[variable_name]]))
myeloid_seurat$PB_variable<- paste0(myeloid_seurat$ShortSubtype, "_", 
                                    myeloid_seurat@meta.data[[variable_name]])
pb<- pseudobulk_byVariable(myeloid_seurat@assays$RNA@counts, myeloid_seurat@meta.data, "PB_variable")

## center and score
pb_list<- NormCenter(pb)
pb_center<- pb_list$center_data

## Plot
p<- myHeatmap(pb_center, GOI=top_markers$gene, min.value = -3, max.value = 3, 
              orderFactors = unique(top_markers$cluster))+
  scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")

## Add horizontal lines around each set of nmf genes
if(add_lines){
  tmp<- split(top_markers, f=top_markers$cluster)
  lastGene<- lapply(tmp, function(x){
    x$tmp<- 1:nrow(x)
    x<- x[x$gene %in% rownames(pb), ]
    return(x[max(x$tmp), "gene"])
  })
  linePositions<- as.data.frame(pb_center[top_markers$gene,])
  linePositions$tmp<- 1:nrow(linePositions)
  linePositions<- linePositions[unlist(lastGene),"tmp"]
  linePositions<- linePositions[1:length(linePositions)-1]+ 0.5
  
  p+geom_hline(yintercept = linePositions, linetype="dashed",size=1)
}else{p}

  
ggsave(paste0(figure_dir, "Heatmap_TopNMFgenes_annotBy10X_rank", rank, "_withSPP1.png"), width=12, height=12)

```

## UMAP with annotations
```{r}
rank_use<-6
variable_name<- paste0("NMF_rank", rank_use, "byScore_annotByAdult10X_withCycling")
colors_use<-color_annots[[paste0("rank", rank_use)]]
myeloid_seurat<-readRDS(paste0("../../Preprocessing/Counts_Preprocess/analysis/detailed_annotation/", cohort, "/myeloid_seurat_noDC.B.Rds"))

DimPlot(myeloid_seurat, group.by =variable_name)+scale_color_manual(values=colors_use)
ggsave(paste0(figure_dir, "UMAP_byAdult10XAnnot_rank", rank_use, ".png"), width=6, height=4)

## Plot each program one at a time
seurat_tmp<- myeloid_seurat
all_plots<- lapply(as.character(unique(myeloid_seurat@meta.data[[variable_name]])), function(x){
  seurat_tmp$tmp<- ifelse(seurat_tmp@meta.data[[variable_name]]==x,
                          x, "Other")
  p<-DimPlot(seurat_tmp, group.by = "tmp")+scale_color_manual(values=c(colors_use, Other="grey"))+ggtitle(x)+NoLegend()
  return(p)
})
cowplot::plot_grid(plotlist=all_plots, ncol = 2)
ggsave(paste0(figure_dir, "UMAP_indv_byAdult10XAnnot_rank", rank_use, ".png"), width=8, height=14)
```

## Density plots
```{r}
## Read in scores for programs
rank_use<-6
scores<-readRDS(paste0(analysis_dir, "NMF_scores_rank",rank_use,".Rds"))
variable_name<- paste0("NMF_rank", rank_use, "byScore_annotByAdult10X_withCycling")

## Create df for plotting
myeloid_df<-myeloid_seurat@meta.data[,c("sampleid", "NiceSubtype", variable_name)]
sum(rownames(scores)==rownames(myeloid_df)); nrow(scores); nrow(myeloid_df)
myeloid_df<- cbind(myeloid_df, scores[,!grepl("Max", colnames(scores))] )

## get number of cells per subtype, add to df
nCells<- as.data.frame(table(myeloid_seurat$NiceSubtype))
myeloid_df$NCells<- plyr::mapvalues(myeloid_seurat$NiceSubtype, nCells$Var1, nCells$Freq)
myeloid_df$JustSubtype<- myeloid_df$NiceSubtype
myeloid_df$NiceSubtype<- paste0(myeloid_df$NiceSubtype, "\n", "nCells=",myeloid_df$NCells)

colors_use<- c(Ependymoma="magenta3",
           `Hemispheric-BCOR`="deepskyblue1",
           `Hemispheric-H3G34R/V`="blue2",
           `Hemispheric-HistoneWT`="lightsteelblue4",
           `Midline-H3K27M`="chartreuse3",
           `Midline-HistoneWT`="grey78")
names(colors_use)<- plyr::mapvalues(names(colors_use), myeloid_df$JustSubtype, myeloid_df$NiceSubtype, warn_missing=FALSE)



## Plot each nmf program
all_plots<- list()
for(i in as.character(unique(myeloid_seurat@meta.data[[variable_name]]))){
  df<- myeloid_df[,c(i, "NiceSubtype")]
  colnames(df)<- c("score", "NiceSubtype")
  p<-ggplot(df, aes(x=score, fill=NiceSubtype))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.position = "none",
          plot.title = element_text(size=20, face="bold", hjust=0.5))

  all_plots[[i]]<-p
}

## add legend separately
legend<-ggplot(df, aes(x=score, fill=NiceSubtype))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.key.size = unit(1.5, 'cm'),
          legend.text = element_text(size=20))
all_plots$legend<-  as_ggplot(get_legend(legend))
all_plots<- all_plots[c(1:3,8,4:7)]
cowplot::plot_grid(plotlist = all_plots, ncol=4)

ggsave(paste0(figure_dir, "Myeloid_DensityOfNMFBySubtype.png"), width=20, height=8)

```

## Histogram of scores- bimodal or gradient?
```{r}
rank_use<-6
colors_use<-color_annots[[paste0("rank", rank_use)]]


## Read in scores for programs
scores<-readRDS(paste0(analysis_dir, "NMF_scores_rank",rank_use,".Rds"))

all_plots<- lapply(colnames(scores[!grepl("Max", colnames(scores))]), function(x){
  df<- scores[,c(x, "MaxProgram")]; colnames(df)<- c("ProgramScore", "MaxProgram")
  ggplot(df, aes(x=ProgramScore, fill=MaxProgram))+
    geom_histogram()+
    scale_fill_manual(values=colors_use)+
    ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "Histogram_ProgramScores.png"), width=15, height=9)
```

## Breakdown percentage of programs in subtypes
```{r}
rank_use<- 6
colors_use<-color_annots[[paste0("rank", rank_use)]]
variable_name<- paste0("NMF_rank", rank_use, "byScore_annotByAdult10X_withCycling")

## Create df for plotting
myeloid_df<-myeloid_seurat@meta.data[,c("sampleid", "NiceSubtype",variable_name)]
myeloid_bd<- as.data.frame(table(myeloid_df$NiceSubtype, myeloid_df[[ variable_name]]))
colnames(myeloid_bd)<- c("Subtype", "Program", "NCells")

## Convert n cells to percentage
myeloid_bd<- myeloid_bd %>% group_by(Subtype) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()

## Add on broad subtype
myeloid_bd$BroadSubtype<- plyr::mapvalues(myeloid_bd$Subtype, myeloid_seurat$NiceSubtype, myeloid_seurat$NiceSubtype_Merge,warn_missing = FALSE)
myeloid_bd$Subtype<- gsub("Hemispheric-", "", gsub("Midline-", "", myeloid_bd$Subtype))

## Plot
ggplot(myeloid_bd, aes(x=Subtype, y=PerCells, fill=Program))+
  geom_bar(position = "stack", stat="identity")+
  scale_fill_manual(values=colors_use)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black", face="bold"),
        axis.text.y = element_text(color="black", face="bold"),
        axis.title.y = element_text(color="black", face="bold"))+
  ylab("Proportion of cells in program")+xlab("")+
  facet_grid(cols=vars(BroadSubtype), scale="free_x", space="free_x")
ggsave(paste0(figure_dir, "ProgramBreakdown_BySubtype_rank", rank_use, ".png"), width=6.2, height=4)  


```

## Visualize how sample-specific each program is
## Normalize based on ncells in each sample, plot program breakdown by sample
```{r}
rank<- 6
variable_name<- paste0("NMF_rank", rank_use, "byScore_annotByAdult10X_withCycling")

bd<- as.data.frame(table(myeloid_seurat$sampleid, myeloid_seurat@meta.data[[variable_name]]))
colnames(bd)<- c("Sample", "Program", "Ncells")

## Add on number of total cells per sample
nCells_perSample<- table(myeloid_seurat$sampleid)
bd$Ncells_perSample<- as.numeric(as.character(plyr::mapvalues(bd$Sample, names(nCells_perSample), nCells_perSample)))

## Normalize Ncells per program per sample by number of total cells per sample
## Add on norm factor, then multiple nCells per program per sample by norm factor
bd$NormFactor<- max(bd$Ncells_perSample)/bd$Ncells_perSample
bd$Ncells_normalized<-bd$Ncells * bd$NormFactor
bd %>% group_by(Sample) %>% summarise(sum(Ncells_normalized))

## Normalize ncells per program by program size
bd <- bd %>% group_by(Program) %>% mutate(Percells_normalized=Ncells_normalized/sum(Ncells_normalized)) %>% as.data.frame()
bd <- bd %>% group_by(Program) %>% mutate(Percells=Ncells/sum(Ncells)) %>% as.data.frame()

## Plot
colors_use<- distinctColorPalette(length(unique(bd$Sample)))
names(colors_use)<-unique(bd$Sample)

  ## without normalization
  ggplot(bd, aes(x=Program, y=Percells, fill=Sample))+
    geom_bar(stat="identity", position="stack")+
    theme_classic()+
    scale_fill_manual(values=colors_use)+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title.y = element_text(face="bold", color="black"))+
    xlab("")+ylab("Proportion of program by sample")
  ggsave(paste0(figure_dir, "ProgramBreakdown_bySample_rank", rank, ".png"), width=6, height=5)

  ## with normalization
  ggplot(bd, aes(x=Program, y=Percells_normalized, fill=Sample))+
    geom_bar(stat="identity", position="stack")+
    theme_classic()+
    scale_fill_manual(values=colors_use)+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title.y = element_text(face="bold", color="black"))+
    xlab("")+ylab("Proportion of program by sample\nnormalized by sample cell count")
  ggsave(paste0(figure_dir, "ProgramBreakdown_bySample_normalizedBySampleSize_rank", rank, ".png"), width=6, height=5)
```

## Breakdowns for scCoda- scCoda doesn't work for this cohort due to small cell sizes across subtypes
```{r}
sccoda_dir<- paste0("../../scCoda/data/",cohort, "/")
if(!dir.exists(sccoda_dir)){dir.create(sccoda_dir)}

## ped tcells
meta<- myeloid_seurat@meta.data
rank<- 6
variable_name<- paste0("NMF_rank", rank_use, "byScore_annotByAdult10X_withCycling")

## Program by subtype
bd<- as.data.frame.matrix(table(meta$sampleid, meta[[variable_name]]))
bd$Subtype<- plyr::mapvalues(rownames(bd), meta$sampleid, meta$NiceSubtype, warn_missing = FALSE)
bd$Sample<- rownames(bd)
bd$Subtype<- gsub("-", "_", bd$Subtype)
write.csv(bd, paste0(sccoda_dir,"Ped_Myeloid_NMFProgramsWithCycling.csv"), row.names = FALSE)

```