---
title: "NMF visualizations"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Visualize de novo NMF programs

## Approach:
Using de novo NMF programs generated in NMF.Rmd and annotated in AdultGBM10X_AnnotatePed.Rmd, create various plots to help visualize and annotate programs:
heatmap, UMAP, density plots, GO terms

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(ggpubr)
library(randomcoloR)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02b_Myeloid/01a_NMF/")

integration<- "" ## HarmonyInt/ or ""
analysis_dir_tmp<- paste0(working_dir, "/analysis/de.novoNMF/", integration)
figure_dir_tmp<- paste0(working_dir, "/figures/de.novoNMF_plots/", integration)

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp)};if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in seurat object
```{r}
cohort<- "pedOnly_nomuv63"

## Current seurat object
myeloid_seurat<-readRDS(paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",integration, "/",
                               cohort, "/myeloid_seurat_noDC.B_harmony.Rds"))

figure_dir<- paste0(figure_dir_tmp, cohort, "/"); analysis_dir<- paste0(analysis_dir_tmp, cohort, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

annot_colors<- readRDS(paste0(analysis_dir, "annot_colors.Rds"))
```

# Need to re-annotate BT1478 to non-BCOR; hemispheric, maybe GBM? Set to GBM for now, may update
```{r}
updateBCOR<-TRUE

figure_dir<- ifelse(updateBCOR, paste0(figure_dir, "updateBCOR/"), figure_dir_tmp)
#analysis_dir<- ifelse(removeBCOR, paste0(analysis_dir, "updateBCOR/"), data_dir_tmp)

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
#if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

if(updateBCOR){
  myeloid_seurat$NiceSubtype<- ifelse(myeloid_seurat$sample=="BT1478", "Hemispheric-HistoneWT", myeloid_seurat$NiceSubtype)
  myeloid_seurat$ShortSubtype<- ifelse(myeloid_seurat$sample=="BT1478", "Hemi-HistoneWT", myeloid_seurat$ShortSubtype)
}
```


## Plot heatmaps of top marker gene expression
```{r}
rank<- "6"
add_lines<- TRUE
annotation<- "10X" ## "10X" annotations or "" for original

## Read in NMF marker genes
marker_genes<- read.csv(paste0(analysis_dir,  "top_DE_genes_", annotation, "annotated_NMF_rank", rank, ".csv"))
marker_genes$cluster<- gsub("_", "\n", marker_genes$cluster)

## Subset to top markers for each nmf- top 10
## If any genes are top markers for multiple clusters- remove, then find top 10 again
top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
if(sum(duplicated(top_markers$gene))!=0){
  duplicated<- top_markers$gene[duplicated(top_markers$gene)]
  print(paste0("Shared top markers: ", paste(duplicated, collapse=",")))
  marker_genes<- marker_genes[!marker_genes$gene %in% duplicated,]
  top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
}

  ## optional: manually add in SPP1 to show where it falls
  top_markers<- rbind(top_markers,marker_genes[marker_genes$gene=="SPP1" & marker_genes$cluster=="Hypoxic\nTAM",])
  geneRemove<- top_markers[top_markers$cluster=="Hypoxic\nTAM", "gene"][10]
  top_markers<- top_markers[top_markers$gene!=geneRemove,]
  top_markers<- top_markers[order(top_markers$cluster),]

  
## Pseudobulk
myeloid_seurat$ShortSubtype<- gsub("\\\n", "-", myeloid_seurat$ShortSubtype)
myeloid_seurat@meta.data[[paste0("NMF_rank", rank, "_annotByAdult10X")]]<- gsub("\\.", "\n",
                                                                      gsub("_", "\n",
                                                                      myeloid_seurat@meta.data[[paste0("NMF_rank", 
                                                                                                       rank, "_annotByAdult10X")]]))
myeloid_seurat$PB_variable<- paste0(myeloid_seurat$ShortSubtype, "_", 
                                    myeloid_seurat@meta.data[[paste0("NMF_rank", rank, "_annotByAdult10X")]])
pb<- pseudobulk_byVariable(myeloid_seurat@assays$RNA@counts, myeloid_seurat@meta.data, "PB_variable")

## center and score
pb_list<- NormCenter(pb)
pb_center<- pb_list$center_data

## Plot
p<- myHeatmap(pb_center, GOI=top_markers$gene, min.value = -3, max.value = 3, 
              orderFactors = unique(top_markers$cluster))+
  scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")

## Add horizontal lines around each set of nmf genes
if(add_lines){
  tmp<- split(top_markers, f=top_markers$cluster)
  lastGene<- lapply(tmp, function(x){
    x$tmp<- 1:nrow(x)
    x<- x[x$gene %in% rownames(pb), ]
    return(x[max(x$tmp), "gene"])
  })
  linePositions<- as.data.frame(pb_center[top_markers$gene,])
  linePositions$tmp<- 1:nrow(linePositions)
  linePositions<- linePositions[unlist(lastGene),"tmp"]
  linePositions<- linePositions[1:length(linePositions)-1]+ 0.5
  
  p+geom_hline(yintercept = linePositions, linetype="dashed",size=1)
}else{p}

  
ggsave(paste0(figure_dir, "Heatmap_TopNMFgenes_annotBy", annotation, "_rank", rank, ".png"), width=12, height=12)
ggsave(paste0(figure_dir, "Heatmap_TopNMFgenes_annotBy", annotation, "_rank", rank, ".pdf"), device="pdf", width=12, height=12)


```

## Plot by pseudobulk, centering each sample independently
```{r}
rank<- "6"
add_lines<- TRUE
annotation<- "10X" ## "10X" annotations or "" for original

## Read in NMF marker genes
marker_genes<- read.csv(paste0(analysis_dir,  "top_DE_genes_", annotation, "annotated_NMF_rank", rank, ".csv"))
marker_genes$cluster<- gsub("_", "\n", marker_genes$cluster)

## Subset to top markers for each nmf- top 10
## If any genes are top markers for multiple clusters- remove, then find top 10 again
top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
if(sum(duplicated(top_markers$gene))!=0){
  duplicated<- top_markers$gene[duplicated(top_markers$gene)]
  print(paste0("Shared top markers: ", paste(duplicated, collapse=",")))
  marker_genes<- marker_genes[!marker_genes$gene %in% duplicated,]
  top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
}

add_lines<-TRUE
n_markers<-25


## Pseudobulk
myeloid_seurat$PB_variable<- paste0(myeloid_seurat$ShortSubtype, "_", 
                                    myeloid_seurat@meta.data[[paste0("NMF_rank", rank, "_annotByAdult10X")]])
pb<- pseudobulk_byVariable(myeloid_seurat@assays$RNA@counts, myeloid_seurat@meta.data, "PB_variable")

## center for each sample separately
center_bySample<- lapply(unique(myeloid_seurat$ShortSubtype), function(sample){
  pb_tmp<- pb[,grepl(sample, colnames(pb))]
  pb_list_tmp<- NormCenter(pb_tmp)
  pb_center_tmp<- pb_list_tmp$center_data
  return(pb_center_tmp)
})
center_bySample<- do.call("cbind", center_bySample)


## Plot
p<- myHeatmap(center_bySample, GOI=top_markers$gene, min.value = -3, max.value = 3, 
              orderFactors = unique(top_markers$cluster))+
  scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")

## Add horizontal lines around each set of nmf genes
if(add_lines){
  tmp<- split(top_markers, f=top_markers$cluster)
  lastGene<- lapply(tmp, function(x){
    x$tmp<- 1:nrow(x)
    x<- x[x$gene %in% rownames(pb), ]
    return(x[max(x$tmp), "gene"])
  })
  linePositions<- as.data.frame(pb_center[top_markers$gene,])
  linePositions$tmp<- 1:nrow(linePositions)
  linePositions<- linePositions[unlist(lastGene),"tmp"]
  linePositions<- linePositions[1:length(linePositions)-1]+ 0.5
  
  p+geom_hline(yintercept = linePositions, linetype="dashed",size=1)
}else{p}

  
ggsave(paste0(figure_dir, "Heatmap_TopNMFgenes_annotBy", annotation, "_rank", rank, "_centerBySubtype.png"), width=12, height=12)
ggsave(paste0(figure_dir, "Heatmap_TopNMFgenes_annotBy", annotation, "_rank", rank, "_centerBySubtype.pdf"), device="pdf", width=12, height=12)
```
## UMAP with annotations
```{r}
## Option: can use seurat with harmony integration JUST for visualization purposes. 
## Setting this to "_harmony" will load in seurat object with harmony integration, but using all annotation/NMF/etc from non-integrated analysis
use_harmony_forVis<- "_harmony" ## "_harmony" to use, "" to not use
if(use_harmony_forVis=="_harmony"){
  figure_dir_tmp<- paste0(figure_dir, "HarmonyInt/"); if(!dir.exists(figure_dir)){dir.create(figure_dir)}
  }else{figure_dir_tmp<- figure_dir}

myeloid_seurat<-readRDS(paste0("../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/",integration, "/",
                               cohort, "/myeloid_seurat_noDC.B", use_harmony_forVis, ".Rds"))
rank_use<-6
colors_use<-annot_colors[[paste0("rank", rank_use)]]

DimPlot(myeloid_seurat, group.by = paste0("NMF_rank", rank_use, "_annotByAdult10X"))+scale_color_manual(values=colors_use)
ggsave(paste0(figure_dir_tmp, "UMAP_byAdult10XAnnot_rank", rank_use, ".png"), width=6, height=4)
ggsave(paste0(figure_dir_tmp, "UMAP_byAdult10XAnnot_rank", rank_use, ".pdf"), device="pdf", width=6, height=4)

## Plot each program one at a time
seurat_tmp<- myeloid_seurat
all_plots<- lapply(as.character(unique(myeloid_seurat@meta.data[[paste0("NMF_rank", rank_use, "_annotByAdult10X")]])), function(x){
  seurat_tmp$tmp<- ifelse(seurat_tmp@meta.data[[ paste0("NMF_rank", rank_use, "_annotByAdult10X")]]==x,
                          x, "Other")
  p<-DimPlot(seurat_tmp, group.by = "tmp")+scale_color_manual(values=c(colors_use, Other="grey"))+ggtitle(x)+NoLegend()
  return(p)
})
cowplot::plot_grid(plotlist=all_plots, ncol = 2)
ggsave(paste0(figure_dir_tmp, "UMAP_indv_byAdult10XAnnot_rank", rank_use, ".png"), width=6, height=9)
```

## GO terms for NMF programs
```{r}
rank<- 6

## read in markers
marker_genes<- read.csv(paste0(analysis_dir,  "top_DE_genes_10Xannotated_NMF_rank", rank, ".csv"))

## GO term ORA
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Read in df for converting symbol --> ensembl
all_genes_Ens.Sym<- readRDS(paste0("/Users/jlabelle/Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Marker_genes/", 
                                   "allGenes_Ensembl.Symbol.Rds"))

## Convert input genes to ensembl. Not all genes found in all_genes_Ens.Sym bc not all Symbols match Ensembl
## These genes removed
allOE_genes<- rownames(myeloid_seurat@assays$RNA@counts)[rownames(myeloid_seurat@assays$RNA@counts) %in%
                                                       all_genes_Ens.Sym$hgnc_symbol]
allOE_genes<- plyr::mapvalues(allOE_genes,all_genes_Ens.Sym$hgnc_symbol, all_genes_Ens.Sym$ensembl_gene_id, warn_missing = FALSE)
  
## Run GO: Convert marker genes to ensembl IDs too, then run OR analysis (GO)
go_result = list()
for (i in unique(marker_genes$cluster)){
  print(i)
  genes<- marker_genes[marker_genes$cluster==i, "gene"]
  if(length(genes)!=1){
    sigOE_genes = genes[genes %in% all_genes_Ens.Sym$hgnc_symbol]
    sigOE_genes<- plyr::mapvalues(sigOE_genes, all_genes_Ens.Sym$hgnc_symbol, 
                                  all_genes_Ens.Sym$ensembl_gene_id, warn_missing = FALSE)
    go_result[[i]] = go_analysis(sigOE_genes, allOE_genes)
    }
} 

## Plot and save
n_terms<- 10
go_plots<- plotGO(go_result = go_result, n_terms = n_terms)
cowplot::plot_grid(plotlist=go_plots, ncol=2)
ggsave(paste0(figure_dir, "GO_NMFrank", rank, "_10xAnnotated_nTerms",n_terms, ".pdf"),device="pdf", width=16, height=18)

saveRDS(go_result, file=paste0(analysis_dir,"GO_NMFrank", rank, "_10xAnnotated_nTerms.Rds") )

## RAW DATA EXPORT
cluster_summary<- lapply(go_result, function(x){x$cluster_summary})
cluster_summary<- lapply(names(cluster_summary), function(program){tmp<- cluster_summary[[program]]; tmp$program<- program; return(tmp)})
cluster_summary<- do.call("rbind", cluster_summary)
write.csv(cluster_summary, file="../../../notUploaded/ImmuneDraft_Rebecca/raw_results_files/myeloid_program_GO_terms.csv")
```

## Density plots
```{r}
## Read in scores for programs
rank_use<-6
all_scores<- readRDS(paste0(analysis_dir, "NMF_scores_annotByAdult10X_rank6-6", ".Rds"))
scores<- all_scores[[paste0("rank", rank_use)]]

## Create df for plotting
myeloid_df<-myeloid_seurat@meta.data[,c("sampleid", "NiceSubtype", 
                                        paste0("NMF_rank", rank_use, "_annotByAdult10X"))]
sum(rownames(scores)==rownames(myeloid_df)); nrow(scores); nrow(myeloid_df)
myeloid_df<- cbind(myeloid_df, scores[,!grepl("Max", colnames(scores))] )

## get number of cells per subtype, add to df
nCells<- as.data.frame(table(myeloid_seurat$NiceSubtype))
myeloid_df$NCells<- plyr::mapvalues(myeloid_seurat$NiceSubtype, nCells$Var1, nCells$Freq)
myeloid_df$JustSubtype<- myeloid_df$NiceSubtype
myeloid_df$NiceSubtype<- paste0(myeloid_df$NiceSubtype, "\n", "nCells=",myeloid_df$NCells)

colors_use<- c(Ependymoma="magenta3",
           `Hemispheric-BCOR`="deepskyblue1",
           `Hemispheric-H3G34R/V`="blue2",
           `Hemispheric-HistoneWT`="lightsteelblue4",
           `Midline-H3K27M`="chartreuse3",
           `Midline-HistoneWT`="grey78")
names(colors_use)<- plyr::mapvalues(names(colors_use), myeloid_df$JustSubtype, myeloid_df$NiceSubtype, warn_missing=FALSE)



## Plot each nmf program
all_plots<- list()
for(i in as.character(unique(myeloid_seurat@meta.data[[paste0("NMF_rank", rank_use, "_annotByAdult10X")]] ))){
  df<- myeloid_df[,c(i, "NiceSubtype")]
  colnames(df)<- c("score", "NiceSubtype")
  p<-ggplot(df, aes(x=score, fill=NiceSubtype))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.position = "none",
          plot.title = element_text(size=20, face="bold", hjust=0.5))

  all_plots[[i]]<-p
}

## add legend separately
legend<-ggplot(df, aes(x=score, fill=NiceSubtype))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.key.size = unit(1.5, 'cm'),
          legend.text = element_text(size=20))
all_plots$legend<-  as_ggplot(get_legend(legend))
all_plots<- all_plots[c(1:3,8,4:7)]
cowplot::plot_grid(plotlist = all_plots, ncol=4)

ggsave(paste0(figure_dir, "Myeloid_DensityOfNMFBySubtype.pdf"),device="pdf", width=20, height=8)

```

## Histogram of scores- bimodal or gradient?
```{r}
rank_use<-7
colors_use<-annot_colors[[paste0("rank", rank_use)]]

## Read in scores for programs
all_scores<- readRDS(paste0(analysis_dir, "NMF_scores_annotByAdult10X_rank6-NA", ".Rds"))
scores<- all_scores[[paste0("rank", rank_use)]]

all_plots<- lapply(colnames(scores[!grepl("Max", colnames(scores))]), function(x){
  df<- scores[,c(x, "MaxProgram")]; colnames(df)<- c("ProgramScore", "MaxProgram")
  ggplot(df, aes(x=ProgramScore, fill=MaxProgram))+
    geom_histogram()+
    scale_fill_manual(values=colors_use)+
    ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "Histogram_ProgramScores.png"), width=15, height=9)
```

## Breakdown of which programs cycling cells are found in
```{r}
rank_use<- 7
colors_use<-annot_colors[[paste0("rank", rank_use)]]

cycling_seurat<- subset(myeloid_seurat, prol_tam=="highScore_prolTAM" )

## df for plotting
df<- as.data.frame(table(cycling_seurat@meta.data[[paste0("NMF_rank", rank_use, "_annotByAdult10X")]]))
colnames(df)<- c("Program", "Ncells")

## Convert to proportion
df$PerCells<- df$Ncells/sum(df$Ncells)

## Plot
ggplot(df, aes(x="", y=PerCells, fill=Program))+
  geom_bar(width=1, stat="identity", color="white")+
  coord_polar("y", start=0)+
  scale_fill_manual(values=colors_use)+
  theme_void()+
  ylab("Proportion of cells in program")+xlab("")
ggsave(paste0(figure_dir, "ProgramBreakdown_WithinCycling_rank", rank, ".png"), width=4, height=3)  
```


## Breakdown percentage of programs in subtypes
```{r}
rank_use<- 6
colors_use<-annot_colors[[paste0("rank", rank_use)]]

## Create df for plotting
myeloid_df<-myeloid_seurat@meta.data[,c("sampleid", "NiceSubtype", paste0("NMF_rank", rank_use, "_annotByAdult10X"))]
myeloid_bd<- as.data.frame(table(myeloid_df$NiceSubtype, myeloid_df[[ paste0("NMF_rank", rank_use, "_annotByAdult10X")]]))
colnames(myeloid_bd)<- c("Subtype", "Program", "NCells")

## Convert n cells to percentage
myeloid_bd<- myeloid_bd %>% group_by(Subtype) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()

## Add on broad subtype
myeloid_bd$BroadSubtype<- plyr::mapvalues(myeloid_bd$Subtype, myeloid_seurat$NiceSubtype, myeloid_seurat$NiceSubtype_Merge,warn_missing = FALSE)
myeloid_bd$Subtype<- gsub("Hemispheric-", "", gsub("Midline-", "", myeloid_bd$Subtype))

## Plot
ggplot(myeloid_bd, aes(x=Subtype, y=PerCells, fill=Program))+
  geom_bar(position = "stack", stat="identity")+
  scale_fill_manual(values=colors_use)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black", face="bold"),
        axis.text.y = element_text(color="black", face="bold"),
        axis.title.y = element_text(color="black", face="bold"))+
  ylab("Proportion of cells in program")+xlab("")+
  facet_grid(cols=vars(BroadSubtype), scale="free_x", space="free_x")
ggsave(paste0(figure_dir, "ProgramBreakdown_BySubtype_rank", rank, ".png"), width=6.2, height=4)  

## Breakdown by sample within subtype
  ## Create df for plotting
  myeloid_df<-myeloid_seurat@meta.data[,c("sampleid", "NiceSubtype", paste0("NMF_rank", rank_use, "_annotByAdult10X"))]
  myeloid_bd<- as.data.frame(table(myeloid_df$sampleid, myeloid_df[[ paste0("NMF_rank", rank_use, "_annotByAdult10X")]]))
  colnames(myeloid_bd)<- c("Sample", "Program", "NCells")
  
  ## Convert n cells to percentage
  myeloid_bd<- myeloid_bd %>% group_by(Sample) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()
  
  ## Add on broad subtype
  myeloid_bd$Subtype<- plyr::mapvalues(myeloid_bd$Sample, 
                                       myeloid_seurat$sampleid, myeloid_seurat$NiceSubtype,warn_missing = FALSE)
  myeloid_bd$Subtype<- gsub("-", "\n", myeloid_bd$Subtype)
  
  ## Plot
  ggplot(myeloid_bd, aes(x=Sample, y=PerCells, fill=Program))+
    geom_bar(position = "stack", stat="identity")+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1, color="black", face="bold"),
          axis.text.y = element_text(color="black", face="bold"),
          axis.title.y = element_text(color="black", face="bold"))+
    ylab("Proportion of cells in program")+xlab("")+
    facet_grid(cols=vars(Subtype), scale="free_x", space="free_x")
  ggsave(paste0(figure_dir, "ProgramBreakdown_BySample_rank", rank_use, ".png"), width=12, height=4)  

```

## Visualize how sample-specific each program is
## Normalize based on ncells in each sample, plot program breakdown by sample
```{r}
rank<- 7

bd<- as.data.frame(table(myeloid_seurat$sampleid, myeloid_seurat@meta.data[[paste0("NMF_rank", rank, "_annotByAdult10X")]]))
colnames(bd)<- c("Sample", "Program", "Ncells")

## Add on number of total cells per sample
nCells_perSample<- table(myeloid_seurat$sampleid)
bd$Ncells_perSample<- as.numeric(as.character(plyr::mapvalues(bd$Sample, names(nCells_perSample), nCells_perSample)))

## Normalize Ncells per program per sample by number of total cells per sample
## Add on norm factor, then multiple nCells per program per sample by norm factor
bd$NormFactor<- max(bd$Ncells_perSample)/bd$Ncells_perSample
bd$Ncells_normalized<-bd$Ncells * bd$NormFactor
bd %>% group_by(Sample) %>% summarise(sum(Ncells_normalized))

## Normalize ncells per program by program size
bd <- bd %>% group_by(Program) %>% mutate(Percells_normalized=Ncells_normalized/sum(Ncells_normalized)) %>% as.data.frame()
bd <- bd %>% group_by(Program) %>% mutate(Percells=Ncells/sum(Ncells)) %>% as.data.frame()

## Plot
colors_use<- distinctColorPalette(length(unique(bd$Sample)))
names(colors_use)<-unique(bd$Sample)

  ## without normalization
  ggplot(bd, aes(x=Program, y=Percells, fill=Sample))+
    geom_bar(stat="identity", position="stack")+
    theme_classic()+
    scale_fill_manual(values=colors_use)+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title.y = element_text(face="bold", color="black"))+
    xlab("")+ylab("Proportion of program by sample")
  ggsave(paste0(figure_dir, "ProgramBreakdown_bySample_rank", rank, ".png"), width=6, height=5)

  ## with normalization
  ggplot(bd, aes(x=Program, y=Percells_normalized, fill=Sample))+
    geom_bar(stat="identity", position="stack")+
    theme_classic()+
    scale_fill_manual(values=colors_use)+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title.y = element_text(face="bold", color="black"))+
    xlab("")+ylab("Proportion of program by sample\nnormalized by sample cell count")
  ggsave(paste0(figure_dir, "ProgramBreakdown_bySample_normalizedBySampleSize_rank", rank, ".png"), width=6, height=5)
```


## Breakdowns for scCoda- scCoda doesn't work for this cohort due to small cell sizes across subtypes
```{r}
sccoda_dir<- paste0("../../scCoda/data/",paste0(gsub("/", "", integration), "_", cohort), "/")
if(!dir.exists(sccoda_dir)){dir.create(sccoda_dir)}

## ped tcells
meta<- myeloid_seurat@meta.data

## Program by subtype
bd<- as.data.frame.matrix(table(meta$sampleid, meta$NMF_rank6_annotByAdult10X))
bd$Subtype<- plyr::mapvalues(rownames(bd), meta$sampleid, meta$NiceSubtype, warn_missing = FALSE)
bd$Sample<- rownames(bd)
bd$Subtype<- gsub("-", "_", bd$Subtype)
write.csv(bd, paste0(sccoda_dir,"Ped_Myeloid_NMFPrograms.csv"), row.names = FALSE)

```

## RAW DATA EXPORT
```{r}
## Load in file to be edited- ped cell wise annotations
df<- read.csv("../../../notUploaded/ImmuneDraft_Rebecca/raw_results_files/ped_cellwise_annotations.csv", row.names = "X")

## Add program
df$Program<- ifelse(df$broad_annot=="Myeloid" & rownames(df) %in% colnames(myeloid_seurat),
                    as.character(plyr::mapvalues(rownames(df),
                                                 colnames(myeloid_seurat),
                                                 as.character(myeloid_seurat$NMF_rank6_annotByAdult10X))),
                    ifelse(df$broad_annot=="Tcell", "Tcell_Hold", "Myeloid_Removed"))
df$Program<- gsub("Stress_Response", "Microglia", df$Program)

## Add mg/mo-tam distinction- separate column only for myeloid
myeloid_seurat$MgMoMonocyte_annot<- gsub("Adult_", "",gsub("Monocytes", "Monocyte", myeloid_seurat$MgMoMonocyte_annot))
df$Microglia_Macrophage_annot<- ifelse(df$broad_annot=="Myeloid" & rownames(df) %in% colnames(myeloid_seurat),
                           as.character(plyr::mapvalues(rownames(df),
                                                 colnames(myeloid_seurat),
                                                 as.character(myeloid_seurat$MgMoMonocyte_annot))),
                           "NA")
df$Microglia_Macrophage_Score<- ifelse(df$broad_annot=="Myeloid" & rownames(df) %in% colnames(myeloid_seurat),
                           as.character(plyr::mapvalues(rownames(df),
                                                 colnames(myeloid_seurat),
                                                 as.character(myeloid_seurat$MgMoMonocyte_score))),
                           "NA")

write.csv(df, file="../../../notUploaded/ImmuneDraft_Rebecca/raw_results_files/ped_cellwise_annotations.csv")
```

