---
title: "Visualize_programs"
output: html_document
---

## Purpose:
After re-annotating programs based on markers, GO, reference genesets, etc in 01_Visualize.Reannotate_programs, visualize updated programs (CD8 only) here


## Approach:
Using updated annotations, plot nice plots for heatmap, UMAP, density, etc


```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(harmony)
library(ggpubr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/01a_Ped/", 
                     "02a_NMFplots/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/AllTcells/")
figure_dir_tmp<- paste0(working_dir, "/figures/AllTcells/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in ped tcell seurat object. Contains NMF assignments (annots added in deNovoPedTcell.v.deNovoAdultTcell, based on correlation with adult T cell + correlation with ped cd4/cd8 programs)
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../../02_Ped.Adult/03_Visualize.Reannotate_programs/analysis/Reannotate_programs/", 
                  "rank",rank_use, "_sharedBy", sharedBy, "/" )

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
  
## ped/adult tcell seurat objects
tcell_seurat<- readRDS(paste0(data_dir, "ped_seurat_tcell_harmony.Rds"))

## Remove "unclear" cells
tcell_seurat<- subset(tcell_seurat, detailed_annot_byNMFrank8_mergedByunion != "Unclear")


## Read in marker genes
program_markers<- readRDS(paste0("../../02_Ped.Adult/01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                                 "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", sharedBy, "_reAnnot.Rds"))



## Read in colors to use
program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, "_reAnnot.Rds"))
celltype_colors<- readRDS(paste0("../../../plot_colors/Tcell.celltypes.Rds"))
subtype_colors<-  readRDS("../../../plot_colors/Subtype.Rds")
```

## Optional: remove inhibitory (may not be real CD8 program)
No longer necessary- removed at source
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
tcell_seurat$tmp<- paste0(tcell_seurat$detailed_annot_byNMFrank8_mergedByunion,  tcell_seurat@meta.data[[program_name]])
tcell_seurat<- subset(tcell_seurat, tmp !="CD8Inhibitory")


figure_dir<- paste0(figure_dir, "noInhibitory/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)}
```


## Subset to top markers for each nmf- top 10
```{r}
marker_genes<- program_markers

## If any genes are top markers for multiple clusters- remove, then find top 10 again
top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
if(sum(duplicated(top_markers$gene))!=0){
  duplicated<- top_markers$gene[duplicated(top_markers$gene)]
  print(paste0("Shared top markers: ", paste(duplicated, collapse=",")))
  marker_genes<- marker_genes[!marker_genes$gene %in% duplicated,]
  top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
}
all_markers<- top_markers
all_markers$cluster<- as.character(all_markers$cluster)
```



## Plot heatmaps of top marker gene expression
```{r}
rank<- "8"
add_lines<- TRUE
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
grouping_variable<- "detailed_annot" ## variable to pseudobulk by. "Subtype", "Age", "detailed_annot"

top_markers<-all_markers
program_order<- c("Cytotoxic\nNKreceptor", "Predysfunctional", "Inhibitory", "Treg", "Cycling", "IFN\nResponse", "Memory","Effector\nMemory", "Stress")

## Order markers
top_markers$cluster<- factor(top_markers$cluster, levels=gsub("\n", "\\.", program_order))
top_markers<- top_markers[order(top_markers$cluster),]
top_markers$cluster<- as.character(top_markers$cluster)

## Pseudobulk
if(grouping_variable=="Subtype"){
  tcell_seurat$SubtypeEdit<- gsub("Hemispheric-BCOR", "BCOR", gsub("Hemispheric-H3G34R/V", "H3G34R/V", gsub("Midline-H3K27M", "H3K27M",
                           gsub("GBM", "Adult.GBM",gsub("IDH-G", "Adult.IDH-G", tcell_seurat$NiceSubtype)))))
  tcell_seurat$PB_variable<- paste0(tcell_seurat$SubtypeEdit, "_",
                                gsub("\\.", "\n", tcell_seurat@meta.data[[program_name]]))
  figure_append<- "_bySubtype"
}else if(grouping_variable=="Age"){
  tcell_seurat$PB_variable<- paste0(tcell_seurat$Age, "_", 
                                gsub("\\.", "\n", tcell_seurat@meta.data[[program_name]]))
  figure_append<- "_byAge"
} else if(grouping_variable== "detailed_annot"){
  tmp<- paste0(grouping_variable, "_byNMFrank", rank_use, "_mergedBy", sharedBy)
  tcell_seurat$PB_variable<- paste0(tcell_seurat@meta.data[[tmp]], "_", 
                                gsub("\\.", "\n", tcell_seurat@meta.data[[program_name]]))
  figure_append<- "_byDetailedAnnot"
}

## Pseudobulk
pb<- pseudobulk_byVariable(tcell_seurat@assays$RNA@counts, tcell_seurat@meta.data, "PB_variable")

## Remove "Unclear" (for detailed_annot)
pb<-pb[,!grepl("Unclear", colnames(pb))]

## center and score
pb_list<- NormCenter(pb)
pb_center<- pb_list$center_data


## Plot
p<- myHeatmap(pb_center, GOI=top_markers$gene, min.value = -3, max.value = 3, 
              orderFactors =  gsub("\\.", "\n", program_order)) +
  scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")

## Add horizontal lines around each set of nmf genes
if(add_lines){
  tmp<- split(top_markers, f=top_markers$cluster)
  lastGene<- lapply(tmp, function(x){
    x$tmp<- 1:nrow(x)
    x<- x[x$gene %in% rownames(pb), ]
    return(x[max(x$tmp), "gene"])
  })
  linePositions<- as.data.frame(pb_center[top_markers$gene,])
  linePositions$tmp<- 1:nrow(linePositions)
  linePositions<- linePositions[unlist(lastGene),"tmp"]
  linePositions<- linePositions[1:length(linePositions)-1]+ 0.5
  
  p+geom_hline(yintercept = linePositions, linetype="dashed",size=1)
}else{p}

  
ggsave(paste0(figure_dir, "AllTcells_Heatmap_TopNMFgenes_annot_rank", rank,figure_append, ".png"), width=16, height=14)


```

## UMAP with annotations
```{r}
rank<- "8"
sharedBy<- "union"
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
program_colors_use<- program_colors[names(program_colors) %in% unique(tcell_seurat@meta.data[[program_name]])]

DimPlot(tcell_seurat, group.by = program_name)+
  scale_color_manual(values=program_colors_use)+
  ggtitle("")
ggsave(paste0(figure_dir, "CD8_UMAP_annot.png"), width=6, height=4)

## Plot each program one at a time
seurat_tmp<- tcell_seurat
all_plots<- lapply(as.character(unique(tcell_seurat@meta.data[[program_name]])), function(x){
  seurat_tmp$tmp<- ifelse(seurat_tmp@meta.data[[program_name]]==x,
                          x, "Other")
  p<-DimPlot(seurat_tmp, group.by = "tmp")+scale_color_manual(values=c(program_colors_use, Other="grey"))+ggtitle(x)+NoLegend()
  return(p)
})
cowplot::plot_grid(plotlist=all_plots, ncol = 2)
ggsave(paste0(figure_dir, "CD8_UMAP_indv.png"), width=6, height=10)
```



## Programs by subtype
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
tcell_seurat$tmp<- tcell_seurat@meta.data[[program_name]]

## remove GBM? Very few cells
removeGBM<- TRUE

if(removeGBM){
  tcell_seurat_use<- subset(tcell_seurat, NiceSubtype != "Hemispheric-HistoneWT")
}else{
  tcell_seurat_use<- tcell_seurat
}

## By subtype
df<- tcell_seurat_use@meta.data %>% group_by(tmp, NiceSubtype) %>% summarise(nCells=n()) %>% 
  group_by(NiceSubtype) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()

ggplot(df, aes(x=NiceSubtype, y=perCells, fill=tmp))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=program_colors[names(program_colors) %in% unique(df$tmp)])+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"),
        axis.title = element_text(face="bold", color="black"))+
  ylab("Proportion of cells")+xlab("")
filename<- ifelse(removeGBM,"Breakdown_SubtypeByProgram_noGBM", "Density_NMFScoreBySubtype" )
ggsave(paste0(figure_dir,filename, ".png"), width=5, height=4)


## By Age- under 4, 4-25
tcell_seurat_use$age_numeric<- as.numeric(tcell_seurat_use$age_numeric)
tcell_seurat_use$Age_group<- ifelse(tcell_seurat_use$age_numeric<4, "under4",
                              ifelse(tcell_seurat_use$age_numeric>=4 & tcell_seurat_use$age_numeric <= 25, "age4-25", "over25"))
df<- tcell_seurat_use@meta.data %>% group_by(tmp, Age_group) %>% summarise(nCells=n()) %>% 
  group_by(Age_group) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()
df$Age_group<- factor(df$Age_group, levels=c("under4", "age4-25"))

ggplot(df, aes(x=Age_group, y=perCells, fill=tmp))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=program_colors[names(program_colors) %in% unique(df$tmp)])+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"),
        axis.title = element_text(face="bold", color="black"))+
  ylab("Proportion of cells")+xlab("")
filename<- ifelse(removeGBM,"Breakdown_AgeByProgram_noGBM", "Density_NMFScoreBySubtype" )
ggsave(paste0(figure_dir,filename, ".png"), width=5, height=4)

```









