---
title: "_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/PD1_Treatment/Tumor/")


preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
stat_helper = paste0(base_dir, script_dir, "common_stat_test.R")
source(preprocessing)
source(stat_helper)


seurat_analysis_folder = "analysis/seurat/"
seurat_fig_folder = "figures/seurat/"
for (folder in c(seurat_analysis_folder, seurat_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}
```

## Load counts and metadata for initiating seurat object
```{r load-data}
load("analysis/cm_list.Robj")
samples = readRDS("analysis/samples.rds")
tumor_stat = readRDS("analysis/inferCNV/seurat/call_tumor.rds")
```

## Subset cm to remove noise/normal samples
```{r subset-tumor-cells}
cm = cm_list$raw_data; rm(cm_list); gc()
cm = cm[,tumor_stat=="Malignant"]
samples = samples[tumor_stat=="Malignant"]

any(colnames(cm) != names(samples))
```

## Initiate seurat obj and add metadata 
```{r seurat_obj_meta_data}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", 
                                    min.cells=0, 
                                    min.genes=0, 
                                    scale.factor=1E5, 
                                    do.scale=F, 
                                    do.center=T)

## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples
```

## OPTIONAL: plot highly variable genes
```{r}
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## PCA and PC selection
```{r PCA_selection}
## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 97,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)

## PC selection- only if not integrating via harmony later
pcs_to_test = 50

seurat_obj <- JackStraw(
  object = seurat_obj, 
  num.pc = pcs_to_test,
  num.replicate = 100, 
  display.progress = TRUE
)
JackStrawPlot(object = seurat_obj, PCs = 1:pcs_to_test)
PCElbowPlot(object = seurat_obj, num.pc = pcs_to_test)
```

## Harmony data integration 
```{r run_harmony}
library(harmony)
library(magrittr)
library(cowplot)

## Run Harmony
seurat_obj = RunHarmony(seurat_obj, "sample", theta = 2, 
                        max.iter.harmony = 50, plot_convergence = TRUE)

## Simple dim and vlnplot to examine integration 
options(repr.plot.height = 6, repr.plot.width = 12)

p1 <- DimPlot(object = seurat_obj, reduction = "pca", pt.size = .1, group.by = "sample")
p2<- DimPlot(object = seurat_obj, reduction = "harmony", pt.size = .1, group.by = "sample")
p3 <- VlnPlot(object = seurat_obj, features = "PC_1", pt.size = .1, group.by = "sample")
p4 <- VlnPlot(object = seurat_obj, features = "harmony_1", pt.size = .1, group.by = "sample")

plot_grid(p1,p2, p3, p4)
ggsave("Harmony_diagnostics_theta2.jpg", path=seurat_fig_folder, width=12, height=8)


## DimHeatmap
DimHeatmap(object = seurat_obj, reduction = "harmony", cells = 500, dims = 1:6, balanced = TRUE)
```

## Clustering and UMAP
```{r clustering_tSNE}
## Run UMAP and clustering using reduced dimensions from Harmony
resolution = .8
reduction_type = "harmony"
dims = 20

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = reduction_type, dims = 1:dims)


seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = reduction_type,
                            dims = 1:dims,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = resolution)


r0.8<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5) + ggtitle("Resolution_0.8")
r1.2<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5)+ ggtitle("Resolution_1.2")
r1.0<-DimPlot(seurat_obj, reduction = "umap", pt.size = 2.5)+ ggtitle("Resolution_1")

r0.8+r1.0+r1.2 #resolution of 0.8 may be best
ggsave("UMAP_clustering_VaryResolution.jpg", path=seurat_fig_folder, width=15, height=6)

## Clustering results 
table(Idents(seurat_obj))

## Save the seurat object
save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```

## UMAP plot colored by clustering and samples
## Can color tSNE with other metadata features (e.g. primary/recurrence, treatment)
```{r UMAP_plots}
## By clustering
DimPlot(object = seurat_obj, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 3, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_WithInt_Res0.8.jpg", path=seurat_fig_folder, width=8, height=8)

## By samples
DimPlot(object = seurat_obj, reduction="umap", group.by = "sample", label = FALSE, 
        pt.size = 3, label.size = 5) + NoAxes() 
ggsave("UMAP_sample_WithInt_Res0.8.jpg", path=seurat_fig_folder, width=10, height=8)

```

## DEGs for each cluster 
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.5)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05 & all_markers$avg_log2FC > 0.5, ]

## Subset the top 30 DEGs for each cluster
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_log2FC) %>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(seurat_analysis_folder, "all_markers_Res0.8.Robj"))
write.csv(top_marker_genes, file=paste0(seurat_analysis_folder, "Top30Markers_Res0.8.csv"))

## Heatmap for initial inspection
DoHeatmap(seurat_obj, 
          features=all_markers_filtered$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs_Res0.8.jpg", path=seurat_fig_folder)

## For res 0.1- compare clusters 0/1
c0_v_c1_seurat<- subset(seurat_obj, seurat_clusters %in% c(0,1))
c0_v_c1<- FindAllMarkers(c0_v_c1_seurat)
save(c0_v_c1, file=paste0(seurat_analysis_folder, "Res1_C0.v.C1_Allgenes.Robj"))
c0_v_c1_filt<- c0_v_c1[c0_v_c1$p_val_adj<0.1 & c0_v_c1$avg_log2FC>0.5,]
```

## Quick annotation by OPC/OC/AC marker genes 
```{r}
## OPC-like
FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OPC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## OC-like cells
FeaturePlot(seurat_obj,
            c("BCAS1", "SIRT2", "RGR", 
              "GPR17", "NFASC", "MBP",
              "TF", "MOG", "PLP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## AC-like cells 
FeaturePlot(seurat_obj,
            c("APOE", "AQP4", "AGT", 
              "SPARCL1", "GFAP", ## ATP1A2 removed
              "S100B", "ALDOC", "CLU"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("AC_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## Neuronal-like cells
FeaturePlot(seurat_obj,
            c("STMN2", "STMN1", "STMN4", 
              "TUBB3", "INA", "ELAVL4",
              "SNAP25", "GABRG2", "ST18"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("Neuron_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## Radial-glia like
## Not expressed in radial glia: AQP4, ID3, and HES5
FeaturePlot(seurat_obj,
            c("NAAA", "NKAIN4", "SOX9", 
              "HES1", "SOX2", "PAX6", 
              "ATP1A2"), ## NAAA and NKAIN4 are identified from mice, so not necessarily apply to human
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("RadialGlia_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## Radial-glia like: vRG
FeaturePlot(seurat_obj,
            c("FBXO32", "CTGF", "CYR61"), 
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("RadialGlia_vRG_markers.jpg", path=seurat_fig_folder, width=9, height=3)

## Radial-glia like: oRG
FeaturePlot(seurat_obj,
            c("MOXD1", "HOPX", "FAM107A", 
              "MT3"), 
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("RadialGlia_oRG_markers.jpg", path=seurat_fig_folder, width=9, height=6)

## Radial-glia like: tRG
FeaturePlot(seurat_obj,
            c("CRYAB", "NR4A1", "FOXJ1"), 
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("RadialGlia_tRG_markers.jpg", path=seurat_fig_folder, width=9, height=3)

## Cycling cells
FeaturePlot(seurat_obj,
            c("CDK4", "TOP2A"), 
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("Cycling_markers.jpg", path=seurat_fig_folder, width=9, height=3)

## Inhibitory ligands
FeaturePlot(seurat_obj,
            c("PROS1", "CLEC2D", "CXCL16", 
              "ANGTPL2", "CDH1", "CCL5", 
              "PVR","SPP1"), 
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("Inhibitory_markers.jpg", path=seurat_fig_folder, width=9, height=6)

VlnPlot(seurat_obj,
            c("PROS1", "CLEC2D", "CXCL16", 
              "ANGTPL2", "CDH1", "CCL5", 
              "PVR","SPP1", "PDL1"), sort = TRUE)
```

## ORA test 
```{r over-representation-test}
library(clusterProfiler)
library(biomaRt)
library(stringr)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
ensembl = useMart("ensembl",dataset="hsapiens_gene_ensembl")
allOE_genes<-getBM(attributes='ensembl_gene_id', 
                   filters = 'hgnc_symbol', 
                   values = rownames(c0_v_c1_seurat@assays$RNA@counts), 
                   mart = ensembl)


## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
all_markers_filtered<- c0_v_c1_filt
go_result = list()
for (i in unique(all_markers_filtered$cluster)){
  print(i)
  genes<- all_markers_filtered[all_markers_filtered$cluster==i, "gene"]
  if(length(genes)!=1){
    sigOE_genes = getBM(attributes='ensembl_gene_id', 
                        filters = 'hgnc_symbol', 
                        values = genes,
                        mart = ensembl)$ensembl_gene_id
    go_result[[paste0("c", i)]] = go_analysis(sigOE_genes, allOE_genes$ensembl_gene_id)
    }
} 



## Plot top pathways for all clusters, display on one plot
clusters = names(go_result)
allDotPlots_20<-list()
for (c in 1:length(clusters)){
  cluster=clusters[c]
  print(cluster)
  if (nrow(go_result[[cluster]]$cluster_summary)>0){
    print(cluster)
    allDotPlots_20[[c]]<-dotplot(go_result[[cluster]]$ego, 
                                showCategory=15, 
                                font.size = 15, 
                                title = cluster, 
                                label_format=10) +
      scale_y_discrete(labels=function(x) str_wrap(x, width=60))+
      theme(plot.title = element_text(hjust = 0.5, face = "bold", 
                                      color="black", size = 28),
            axis.title = element_text(face = "bold", color="black"), 
            axis.text.x = element_text(angle = 45, hjust = 1, color="black"),
            axis.text.y = element_text( color="black",face = "bold"))
  }
  
}


cowplot::plot_grid(plotlist = allDotPlots_20, ncol=2)
ggsave("GO_OVA_AllClusters_top15_Res1_C0vC1.jpg", path=seurat_fig_folder, width=18, height=7)
save(go_result, file=paste0(seurat_analysis_folder, "go_result_Res0.8.Robj"))
```


## ORA with previously published datasets
## Compare with metaprograms from previously defined metaprograms
```{r}
all_markers_list<-split(all_markers_filtered, all_markers_filtered$cluster )
all_markers_list<-lapply(all_markers_list, function(x){x$gene})
names(all_markers_list)<-paste0("Cluster", names(all_markers_list))

hemi_genesets<-readRDS(paste0(working_dir,
                                 "../../../PublishedDatasets/FilbinPublishedDatasets/Metaprograms/",
                                 "hemisphericPreviousPrograms_2019_OC.rds"))
midline_genesets<- readRDS(paste0(working_dir,
                                 "../../../PublishedDatasets/FilbinPublishedDatasets/Metaprograms/",
                                 "midlinePreviousPrograms_OC.AC.OPC.MES1.rds"))

gene_list<- hemi_genesets

tmp2 = sapply(all_markers_list, function(x){
    sapply(gene_list, function(y){
        fisher_test(x, y, unlist(union(gene_list, all_markers_list)))
    })
})
tmp2 = data.frame(tmp2)
tmp3<- apply(tmp2, 2, function(x) {p.adjust(x) < 0.1})
tmp4<-apply(tmp2, 2, function(x) {p.adjust(x)})

write.table(tmp4, paste0(seurat_analysis_folder,
                        "ScoreSeuratClustersUsingHemi_Res0.8.txt"))

## Alternatively: fisher test with de novo NMF programs
NMF_denovo<- read.delim("analysis/nmf/Merge/metaprogram_genesMerge4_cluster6CellsNormal.txt")
NMF_denovo<-as.list(NMF_denovo)

tmp2 = sapply(all_markers_list, function(x){
    sapply(NMF_denovo, function(y){
        fisher_test(x, y, unlist(union(NMF_denovo, all_markers_list)))
    })
})
tmp2 = data.frame(tmp2)
tmp3<- apply(tmp2, 2, function(x) {p.adjust(x) < 0.1})
tmp4<-apply(tmp2, 2, function(x) {p.adjust(x)})

write.table(tmp4, paste0(seurat_analysis_folder,
                        "ScoreSeuratClustersUsingDenovoNMF_Theta0.9_padj.txt"))
```


# Add NMF results to the seurat object

## Map NMF scores to seurat object 
```{r add-NMF-score}
NameOfMetaprogramList<-"Merge4"
nmf_dir<- "analysis/nmf/all_subtypes/"

nmf_score = readRDS(paste0(nmf_dir, "Merge/nmf_score", NameOfMetaprogramList, ".rds"))
seurat_obj = AddMetaData(seurat_obj, nmf_score)
load(paste0(nmf_dir, "Merge/nmf_marker_genes_final", NameOfMetaprogramList, ".Robj"))

DimPlot(object = seurat_obj, group.by = "signature_1", label = TRUE, 
         pt.size = 3, label.size = 5, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_nmf_signature",NameOfMetaprogramList, ".jpg"), 
       path=seurat_fig_folder, 
       width=14, height=12)

DoHeatmap(seurat_obj, 
          features=unlist(nmf_marker_genes_final)) + NoLegend()

ggsave("nmf_marker_genes.jpg", path=seurat_fig_folder, width=12, height=8)
```


## Add a second metagene classification without cell cycle programs
```{r}
nmf_score = readRDS(paste0("analysis/nmf/all_subtypes/new/nmf_score_updated", NameOfMetaprogramList,  ".rds"))
seurat_obj@meta.data$signaure_1_without_cc = nmf_score$signature_1
```

## Make pseudobulk
```{r}
metagene_program_names = sort(unique(nmf_score$signature_1))
pseudobulk = NULL
aggr_method = "mean"
for (metagene in metagene_program_names){
  cm = seurat_obj@assays$RNA@counts[,nmf_score$signature_1 == metagene & nmf_score$score_1 >= 1]
  if (aggr_method == "mean"){
    pseudobulk = cbind(pseudobulk, rowMeans(cm))
  }
  if (aggr_method == "median"){
    pseudobulk = cbind(pseudobulk, rowMedians(cm))
  }
}
colnames(pseudobulk) = metagene_program_names
rownames(pseudobulk) = rownames(seurat_obj@assays$RNA@counts)
saveRDS(pseudobulk, file=paste0(seurat_analysis_folder, 
                                "pseudobulk_cm_unique_only_", aggr_method, NameOfMetaprogramList,".RDS"))
```

## Metagene program specific genes as overall DEGs
```{r}
seurat_obj = SetIdent(seurat_obj, value = seurat_obj@meta.data$signaure_1_without_cc)

all_markers_nmf = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_nmf = all_markers[all_markers$p_val_adj < 0.05, ]
saveRDS(all_markers, paste0(seurat_analysis_folder, "deg.rds"))
##sapply(names(nmf_marker_genes_final), function(x) table(nmf_marker_genes_final[[x]] %in% gene_signatures[[x]]))
signature_genes = unique(all_markers$gene)
saveRDS(signature_genes, paste0(seurat_analysis_folder, "signature_genes.rds"))
```

## Add annotations
```{r}
ManualAnnot<- as.character(unique(seurat_obj$seurat_clusters))
names(ManualAnnot)<- c("Neuronal-like", "Cycling", "MES-like") # etc
seurat_obj$ManualAnnot<- plyr::mapvalues(seurat_obj$seurat_clusters, ManualAnnot, names(ManualAnnot))

DimPlot(seurat_obj, group.by = "ManualAnnot", label=TRUE, pt.size = 2, label.size = 8)+NoLegend()
ggsave(paste0(seurat_fig_folder, "UMAP_ManualAnnot_Res0.8.png"), width=6, height=6)
```

## Plot cell types per sample
```{r}
annotBreakdown<- as.data.frame(table(seurat_obj$sample, seurat_obj$ManualAnnot))
colnames(annotBreakdown)<- c("Sample", "CellType", "NCells")

plot_list<-list()
for(i in unique(annotBreakdown$Sample)){
  df<- annotBreakdown[annotBreakdown$Sample==i,]
  p<-ggplot(df, aes(x="", y=NCells, fill=CellType))+
    geom_bar(stat="identity", color="white")+
    coord_polar("y", start=0)+
    theme_void()+
    ggtitle(i)+theme(plot.title = element_text(size=15, face="bold", hjust=0.5))
  plot_list[[i]]<- p
}
cowplot::plot_grid(plotlist=plot_list, ncol=2)
ggsave(paste0(seurat_fig_folder, "CellBreakdown_Res0.8.png"), width=6, height=3)

```

## Compare expression of inhibitory ligands in these samples vs other hemispheric samples
```{r}
ligands<- c("PROS1", "CLEC2D", "CXCL16", "ANGPTL2", "CDH1", "CCL5", "PVR","CD274")

## Read in other hemispheric samples (maybe non-BCOR?)
load("../ImmuneIntegration/Preprocessing/analysis/AllTumorSeuratObj.Robj")
hemi_seurat<- subset(tumor_merged, subtype %in% c("hemiHistoneMut", "hemiHistoneWT", "hemiBCOR"))

## Merge with current seurat object
pd1_seurat<- seurat_obj; pd1_seurat$subtype<- "PD1_Treat"
all_seurat<- merge(hemi_seurat, pd1_seurat)


## Plot by subtype
VlnPlot(all_seurat, features=ligands, group.by = "subtype", sort=TRUE, ncol = 4)
ggsave(paste0(seurat_fig_folder, "InhibitoryLigands_PD1Treat_vs_otherHemi.png"), width=10, height=7)

## Plot just PD1 by samples
VlnPlot(pd1_seurat, features = ligands, group.by="sample", sort=TRUE, ncol=4)
ggsave(paste0(seurat_fig_folder, "InhibitoryLigands_BySample.png"), width=8, height =5)

## DE between samples
pd1_seurat<- SetIdent(pd1_seurat,value= pd1_seurat$sample)
all_markers_bySample<- FindAllMarkers(pd1_seurat, min.pct = 0.25)
filt_markers_bySample<-all_markers_bySample[all_markers_bySample$p_val_adj<0.1 & all_markers_bySample$avg_log2FC>0,]
```

