---
title: "AnnotateByNMF"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Using shared/specific Tcell programs, annotate each program separately as CD4/CD8

## Approach:
Split seurat object by program, recluster. Annotate cluster-wise when possible, perform re-clustering and cell-wise annotation when needed. Also compare to "actual" annotations (by expression) to assess accuracy

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(harmony)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/02_detailed_annot/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/identifyCD4.CD8_byNMF/")
figure_dir_tmp<- paste0(working_dir, "/figures/identifyCD4.CD8_byNMF/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

immune_colors<- readRDS("../../../plot_colors/ImmuneCell.celltypes.Rds")
```

## Read in ped tcell seurat object. Contains NMF assignments (annots added in deNovoPedTcell.v.deNovoAdultTcell, based on correlation with adult T cell + correlation with ped cd4/cd8 programs)
```{r}
## ped/adult tcell seurat
tcell_seurat<- readRDS(paste0("../../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/Tcells_Ped.Adult_Merge/",
                              "adult_ped_tcell_merged_seurat_harmony.Rds"))

## figure/analysis dir based on rank + how shared programs are merged
rank_use<-8
union_intersect<- "union" 
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", union_intersect, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", union_intersect, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}


```


## For initial estimate of CD4/CD8: classify each cell cell-wise, plot proportions by program
```{r}
variable_name<- paste0("NMF_rank", rank_use, "_annot_mergedBy", union_intersect)

## Make df with CD4/CD8 expression
cd4.cd8_expr<- as.data.frame(t(tcell_seurat@assays$RNA@data[c("CD8A", "CD8B", "CD4"),]))
cd4.cd8_expr$CD8<- apply(cd4.cd8_expr, 1, function(x){max(x["CD8A"], x["CD8B"])})

## Remove cells that are 0 for both
cd4.cd8_expr<- cd4.cd8_expr[rowSums(cd4.cd8_expr)>0,] ## 7,703 cells (77%)

## Assign each cell based on whether CD4 or CD8 expression is higher
cd4.cd8_expr$annotate<- ifelse(cd4.cd8_expr$CD4>cd4.cd8_expr$CD8, "CD4",
                               ifelse(cd4.cd8_expr$CD8 > cd4.cd8_expr$CD4, "CD8", "issue"))

## Add on program info
cd4.cd8_expr$program<- plyr::mapvalues(rownames(cd4.cd8_expr), colnames(tcell_seurat),
                                       tcell_seurat@meta.data[[variable_name]],
                                       warn_missing = FALSE)

## Get proportion of cell types by program
df<- as.data.frame(table(cd4.cd8_expr$program, cd4.cd8_expr$annotate))
colnames(df)<-c("Program", "CellType", "Ncells")
df<-df %>% group_by(Program) %>% mutate(perCells=Ncells/sum(Ncells)) %>% as.data.frame()

## Order
df$CellType<- factor(df$CellType, levels=c(""))

## Plot
ggplot(df, aes(x=Program, y=perCells, fill=CellType))+
  geom_bar(stat="identity", position="stack")+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("") + ylab("Proportion of cells")+
  geom_hline(yintercept = 0.5, linetype="dashed")
ggsave(paste0(figure_dir, "CD4.CD8.predictions_byexpression_NMFrank", rank_use, ".png"), width=5, height=3)

saveRDS(cd4.cd8_expr, file=paste0(analysis_dir, "cd4.cd8_expression_values.Rds"))
```

## Split by NMF, re-run clustering for each
```{r}
rank_use<-8
union_intersect<- "union" ## Doesn't actually matter for this, only for genesets
variable_name<- paste0("NMF_rank", rank_use, "_annot_mergedBy", union_intersect)

tcell_seurat$tmp<- tcell_seurat@meta.data[[variable_name]]
all_seurat<- lapply(unique(tcell_seurat$tmp), function(x){
  print(x)
  seurat_tmp<- subset(tcell_seurat, tmp==x)
  meta_tmp<- seurat_tmp@meta.data
  seurat_tmp<- RunFullSeurat(cm=seurat_tmp@assays$RNA@counts, RunHarmony = TRUE, samples=meta_tmp$sampleid)
  seurat_tmp<- AddMetaData(seurat_tmp, meta_tmp[,!colnames(meta_tmp) %in% colnames(seurat_tmp@meta.data)])
  return(seurat_tmp)
})
names(all_seurat)<- unique(tcell_seurat$tmp)



## Plot CD4/CD8 expression for each
all_plots<-lapply(names(all_seurat), function(x){
  seurat_tmp<- all_seurat[[x]]
  p1<-FeaturePlot(seurat_tmp, features=c("CD4", "CD8A"))
  p2<- DimPlot(seurat_tmp, label=TRUE)+NoLegend()+ggtitle(x)
  p3<- VlnPlot(seurat_tmp, features=c("CD4", "CD8A"))
  cowplot::plot_grid(plotlist=list(p2, p1, p3), ncol=3)
})
cowplot::plot_grid(plotlist=all_plots, ncol=1)
ggsave(paste0(figure_dir, "CD4.CD8_expr_byNMFProgram_rank", rank_use, ".png"), width=15, height=25)

saveRDS(all_seurat, file=paste0(analysis_dir, "seurat_list_splitByNMFrank", rank_use, ".Rds"))
```


## Split by NMF, re-run clustering for each
This approach only used for rank 7
Merge the inhibitory programs into 1, same for cytotoxic
```{r}
## Merge inhibitory and cytotoxic programs into 1
tcell_seurat$NMF_rank7_annot_merge<- gsub(".CD4", "", gsub(".CD8", "", tcell_seurat$NMF_rank7_annot))

all_seurat<- lapply(unique(tcell_seurat$NMF_rank7_annot_merge), function(x){
  print(x)
  seurat_tmp<- subset(tcell_seurat, NMF_rank7_annot_merge==x)
  meta_tmp<- seurat_tmp@meta.data
  seurat_tmp<- RunFullSeurat(cm=seurat_tmp@assays$RNA@counts, RunHarmony = TRUE, samples=meta_tmp$sampleid)
  seurat_tmp<- AddMetaData(seurat_tmp, meta_tmp[,!colnames(meta_tmp) %in% colnames(seurat_tmp@meta.data)])
  return(seurat_tmp)
})
names(all_seurat)<- unique(tcell_seurat$NMF_rank7_annot_merge)



## Plot CD4/CD8 expression for each
all_plots<-lapply(names(all_seurat), function(x){
  seurat_tmp<- all_seurat[[x]]
  p1<-FeaturePlot(seurat_tmp, features=c("CD4", "CD8A"))
  p2<- DimPlot(seurat_tmp, label=TRUE)+NoLegend()+ggtitle(x)
  p3<- VlnPlot(seurat_tmp, features=c("CD4", "CD8A"))
  cowplot::plot_grid(plotlist=list(p2, p1, p3), ncol=3)
})
cowplot::plot_grid(plotlist=all_plots, ncol=1)
ggsave(paste0(figure_dir, "CD4.CD8_expr_byNMFProgram.CytoInhMerge.png"), width=15, height=15)

saveRDS(all_seurat, file=paste0(analysis_dir, "seurat_list_splitByNMF.CytoInhMerge.Rds"))
```


## Annotate cells for each program cluster-wise
Three main types of programs:
1. For some programs (NK, Tregs), all cells are clearly either CD4/CD8
2. For some programs, the clustering above clearly splits cells into CD4/CD8
3. For some programs, while there is clearly CD4/CD8 patterning, need to set the resolution manually to split cluster-wise
4. For some programs, need to do some sub-clustering

## Read in seurat object list (above)- 1 seurat object for each program
```{r}
rank_use<- 8


seurat_list<- readRDS(paste0(analysis_dir, "seurat_list_splitByNMFrank", rank_use, ".Rds"))

```

## Option 1: all cells either CD4/CD8
```{r}
programs<- c("Treg", "Cytotoxic.CD4", "NK-like")
names(programs)<-c("CD4", "CD8", "CD8")
for(i in 1:length(programs)){
  program<- programs[i]
  print(program)
  tmp<- seurat_list[[program]]
  tmp$detailed_annot_byProgram<- names(program)
  seurat_list[[program]]<- tmp
}

```

## Option 2: program contains CD4/CD8, current resolution splits cleanly
```{r}
## Programs that can be annotated at current resolution- annotations based on figure "CD4.CD8_expr_byNMFProgram_rank8.png"
programs<-list(Inhibitory.CD4=list(CD4=c(0:9), CD8=c(10)))

all_plots<-list()
for(program in names(programs)){
  print(program)
  tmp<- seurat_list[[program]]
  program_annot<- programs[[program]]
  tmp$detailed_annot_byProgram<- ifelse(tmp$seurat_clusters %in% program_annot$CD4, "CD4", "CD8")
  p<- DimPlot(tmp, group.by = "detailed_annot_byProgram") + VlnPlot(tmp,
                                                                    group.by = "detailed_annot_byProgram", 
                                                                    features=c("CD4", "CD8A") )
  all_plots[[program]]<-p
  seurat_list[[program]]<- tmp
}
cowplot::plot_grid(plotlist=all_plots)
```

## Option 3: program needs to be manually re-clustered
```{r}
## Memory
seurat_tmp<- seurat_list[["Memory"]]
seurat_tmp<- FindClusters(seurat_tmp, resolution = 2.6)
cd4_clusters<- c(0:2, 4:11, 13, 16:19, 21,23:27)
cd8_clusters<-c(3, 12,14,15,20,22)
seurat_tmp$detailed_annot_byProgram<- ifelse(seurat_tmp$seurat_clusters %in% cd4_clusters, "CD4",
                                             ifelse(seurat_tmp$seurat_clusters %in% cd8_clusters, "CD8",
                                                    "issue"))
DimPlot(seurat_tmp, group.by = "detailed_annot_byProgram") + VlnPlot(seurat_tmp,
                                                                  group.by = "detailed_annot_byProgram", 
                                                                  features=c("CD4", "CD8A") )
seurat_list$Memory<- seurat_tmp

## Stress
seurat_tmp<- seurat_list$Stress

  ## Remove hsp genes- masking cd4/cd8 differences
  hsp_genes<- rownames(seurat_tmp)[grepl("^HSP", rownames(seurat_tmp))]
  seurat_no_hsp<- RunFullSeurat(seurat_tmp@assays$RNA@counts[!rownames(seurat_tmp) %in% hsp_genes,],
                             RunHarmony = TRUE, samples = seurat_tmp$sampleid)
  seurat_no_hsp<- FindClusters(seurat_no_hsp, resolution =.7)
  p1<- FeaturePlot(seurat_no_hsp, features=c("CD4", "CD8A")) 
  p2<- VlnPlot(seurat_no_hsp, features=c("CD4", "CD8A"))
  p3<- DimPlot(seurat_no_hsp, label=TRUE) +NoLegend()
  cowplot::plot_grid(plotlist=list(p1, p2, p3), ncol=1)
  ggsave(paste0(figure_dir, "Stress_NoHSPgenes_annotation.png"), width=6, height=9)
  
  cd4_clusters<-  c(2,3,5,6,8)
  cd8_clusters<-c(0,1,4,7)
  
  seurat_no_hsp$detailed_annot_byProgram<- ifelse(seurat_no_hsp$seurat_clusters %in% cd4_clusters, "CD4",
                                         ifelse(seurat_no_hsp$seurat_clusters %in% cd8_clusters, "CD8",
                                                "issue"))
  
  
  ## Add these annotations back to original seurat object (with hsp genes)
  seurat_tmp$detailed_annot_byProgram<- plyr::mapvalues(colnames(seurat_tmp), colnames(seurat_no_hsp),
                                                        seurat_no_hsp$detailed_annot_byProgram)  
  DimPlot(seurat_tmp, group.by = "detailed_annot_byProgram") + VlnPlot(seurat_tmp,
                                                                group.by = "detailed_annot_byProgram", 
                                                                features=c("CD4", "CD8A") )
  
  seurat_list$Stress<- seurat_tmp
  
```

## Option 4: sub-clustering required
```{r}
## Inhibitory.CD8
seurat_tmp<- seurat_list$Inhibitory.CD8
seurat_tmp<- FindClusters(seurat_tmp, resolution = 2)

p1<- FeaturePlot(seurat_tmp, features=c("CD4", "CD8A")) 
p2<- VlnPlot(seurat_tmp, features=c("CD4", "CD8A"))
p3<- DimPlot(seurat_tmp, label=TRUE) +NoLegend()
cowplot::plot_grid(plotlist=list(p1, p2, p3), ncol=1)

cd4_clusters<- c(0,1,3,6)
cd8_clusters<- c(2,4)
unclear_clusters<- c(5)

seurat_tmp$detailed_annot_byProgram<- ifelse(seurat_tmp$seurat_clusters %in% unclear_clusters, "CellWise",
                                              ifelse(seurat_tmp$seurat_clusters %in% cd4_clusters, "CD4",
                                              ifelse(seurat_tmp$seurat_clusters %in% cd8_clusters, "CD8",
                                              "issue")))

  ## Cell-wise for unclear (all cells show cd4 and/or cd8 expression)
  cellwise_seurat<- subset(seurat_tmp, detailed_annot_byProgram == "CellWise")
  cd4.cd8_expr<- as.data.frame(t(cellwise_seurat@assays$RNA@data[c("CD4", "CD8A", "CD8B"),]))
  cd4.cd8_expr$CD8<- apply(cd4.cd8_expr, 1, function(x){max(x["CD8A"], x["CD8B"])})
  cd4.cd8_expr<- cd4.cd8_expr[rowSums(cd4.cd8_expr)>0,] ## all cells PF
  cd4.cd8_expr$annotate<- ifelse(cd4.cd8_expr$CD4>cd4.cd8_expr$CD8, "CD4",
                                 ifelse(cd4.cd8_expr$CD8>cd4.cd8_expr$CD4, "CD8", "issue"))
  
  ## Add onto rest of inhibitory CD8 annotations
  seurat_tmp$detailed_annot_byProgram<- ifelse(colnames(seurat_tmp) %in% rownames(cd4.cd8_expr),
                                                   plyr::mapvalues(colnames(seurat_tmp), 
                                                                   rownames(cd4.cd8_expr),
                                                                   cd4.cd8_expr$annotate),
                                                   seurat_tmp$detailed_annot_byProgram)
  
 DimPlot(seurat_tmp, group.by = "detailed_annot_byProgram") + VlnPlot(seurat_tmp,
                                                                group.by = "detailed_annot_byProgram", 
                                                                features=c("CD4", "CD8A") )
  
seurat_list$Inhibitory.CD8<- seurat_tmp
    
  


## Effector memory
  seurat_tmp<- seurat_list$EffectorMemory
  seurat_tmp<-FindClusters(seurat_tmp, resolution = 2)
  
  p1<- FeaturePlot(seurat_tmp, features=c("CD4", "CD8A")) 
  p2<- VlnPlot(seurat_tmp, features=c("CD4", "CD8A"))
  p3<- DimPlot(seurat_tmp, label=TRUE) +NoLegend()
  cowplot::plot_grid(plotlist=list(p1, p2, p3), ncol=1)
  
  cd4_clusters<- c(1,4:15,19:20)
  cd8_clusters<- c(0,2,16,18)
  unclear_clusters<- c(3,17)
  
  seurat_tmp$detailed_annot_byProgram<- ifelse(seurat_tmp$seurat_clusters %in% cd4_clusters, "CD4",
                                        ifelse(seurat_tmp$seurat_clusters %in% cd8_clusters, "CD8",
                                        ifelse(seurat_tmp$seurat_clusters %in% unclear_clusters, "unclear",
                                        "issue")))
  
  ## Subset to unclear clusters, rerun clustering
  seurat_unclear<- subset(seurat_tmp, detailed_annot_byProgram=="unclear")
  seurat_unclear<- RunFullSeurat(seurat_unclear@assays$RNA@counts, RunHarmony = TRUE,
                                 samples=seurat_unclear$sample, pca_dims = 100)
  seurat_unclear<-FindClusters(seurat_unclear, resolution = 1.8)
  p1<- FeaturePlot(seurat_unclear, features=c("CD4", "CD8A")) 
  p2<- VlnPlot(seurat_unclear, features=c("CD4", "CD8A"))
  p3<- DimPlot(seurat_unclear, label=TRUE) +NoLegend()
  cowplot::plot_grid(plotlist=list(p1, p2, p3), ncol=1)
  
  cd4_subclusters<- c(1)
  cd8_subclusters<-c(2:5)
  unclear_subclusters<-c(0)
  
    ## Assign 1 cluster cell wise
    seurat_c1<- subset(seurat_unclear, seurat_clusters %in% unclear_subclusters)
    cd4.cd8_expr<- as.data.frame(t(seurat_c1@assays$RNA@counts[c("CD4", "CD8A", "CD8B"),]))
    cd4.cd8_expr$CD8<- apply(cd4.cd8_expr, 1, function(x){max(x["CD8A"], x["CD8B"])})
    cd4.cd8_expr$annotate<- ifelse(rowSums(cd4.cd8_expr) ==0, "Unclear",
                            ifelse(cd4.cd8_expr$CD4>cd4.cd8_expr$CD8, "CD4",
                            ifelse(cd4.cd8_expr$CD8>cd4.cd8_expr$CD4, "CD8", "issue")))
    
    ## Assign others cluster-wise
    seurat_unclear$detailed_annot_byProgram<- ifelse(seurat_unclear$seurat_clusters %in% cd4_subclusters, "CD4",
                                              ifelse(seurat_unclear$seurat_clusters %in% cd8_subclusters, "CD8",
                                              ifelse(seurat_unclear$seurat_clusters %in% unclear_subclusters,
                                                     plyr::mapvalues(colnames(seurat_unclear),
                                                                     rownames(cd4.cd8_expr),
                                                                     cd4.cd8_expr$annotate),
                                                     "issue")))
    ## add on to full effector memory annotations
    seurat_tmp$detailed_annot_byProgram<- ifelse(seurat_tmp$seurat_clusters %in% unclear_clusters,
                                                 plyr::mapvalues(colnames(seurat_tmp),
                                                                 colnames(seurat_unclear),
                                                                 seurat_unclear$detailed_annot_byProgram),
                                                 seurat_tmp$detailed_annot_byProgram)
    
    DimPlot(seurat_tmp, group.by = "detailed_annot_byProgram") + VlnPlot(seurat_tmp,
                                                                group.by = "detailed_annot_byProgram", 
                                                                features=c("CD4", "CD8A") )
    
    seurat_list$EffectorMemory<- seurat_tmp

    
## Cytotoxic CD8
seurat_tmp<- seurat_list$Cytotoxic.CD8

seurat_tmp<-FindClusters(seurat_tmp, resolution = 2.5)
  
p1<- FeaturePlot(seurat_tmp, features=c("CD4", "CD8A")) 
p2<- VlnPlot(seurat_tmp, features=c("CD4", "CD8A"))
p3<- DimPlot(seurat_tmp, label=TRUE) +NoLegend()
cowplot::plot_grid(plotlist=list(p1, p2, p3), ncol=1)

cd8_clusters<- c(0,2,8,10)
unclear_clusters<- c(1,3:7, 9,11)
seurat_tmp$detailed_annot_byProgram<- ifelse(seurat_tmp$seurat_clusters %in% cd8_clusters, "CD8",
                                      ifelse(seurat_tmp$seurat_clusters %in% unclear_clusters, "Unclear",
                                      "issue"))

  ## subcluster unclear clusters
  seurat_subcluster<- subset(seurat_tmp, detailed_annot_byProgram=="Unclear")
  seurat_subcluster<- RunFullSeurat(seurat_subcluster@assays$RNA@counts,
                                    RunHarmony = TRUE, samples=seurat_subcluster$sampleid,
                                    pca_dims = 100)
  seurat_subcluster<-FindClusters(seurat_subcluster, resolution=2)
  
  p1<- FeaturePlot(seurat_subcluster, features=c("CD4", "CD8A")) 
  p2<- VlnPlot(seurat_subcluster, features=c("CD4", "CD8A"))
  p3<- DimPlot(seurat_subcluster, label=TRUE) +NoLegend()
  cowplot::plot_grid(plotlist=list(p1, p2, p3), ncol=1)
  
    ## Assign cd4/cd8/unclear subclusters
    cd4_subclusters<- c(0,1,4:6)
    cd8_subclusters<-c(3)
    cell_wise_subcluster<- c(2)
    
    ## perform cell wise annotation for 1 cluster
    cell_wise_annot<- as.data.frame(t(seurat_subcluster@assays$RNA@counts[c("CD4", "CD8A", "CD8B"),
                                                                        seurat_subcluster$seurat_clusters==cell_wise_subcluster]))
    cell_wise_annot$CD8<- apply(cell_wise_annot, 1, function(x){max(x["CD8A"], x["CD8B"])})
    cell_wise_annot$annotate<- ifelse(cell_wise_annot$CD4 == 0 & cell_wise_annot$CD8 ==0, "Unclear",
                               ifelse(cell_wise_annot$CD4 > cell_wise_annot$CD8, "CD4",
                               ifelse(cell_wise_annot$CD8 > cell_wise_annot$CD4, "CD8",
                                      "issue")))
    
    ## add on annotations for all "unclear" cells
    seurat_subcluster$detailed_annotByProgram<- ifelse(seurat_subcluster$seurat_clusters %in% cd4_subclusters, "CD4",
                                                ifelse(seurat_subcluster$seurat_clusters %in% cd8_subclusters, "CD8",
                                                ifelse(seurat_subcluster$seurat_clusters %in% cell_wise_subcluster,
                                                       plyr::mapvalues(colnames(seurat_subcluster),
                                                                       rownames(cell_wise_annot),
                                                                       cell_wise_annot$annotate),
                                                       "issue")))
    
  ## add on annotations to full cyto cd8 seurat object
  seurat_tmp$detailed_annot_byProgram<- ifelse(seurat_tmp$seurat_clusters %in% unclear_clusters,
                                               plyr::mapvalues(colnames(seurat_tmp), colnames(seurat_subcluster),
                                                               seurat_subcluster$detailed_annotByProgram),
                                               seurat_tmp$detailed_annot_byProgram)
  seurat_list$Cytotoxic.CD8<- seurat_tmp
  
## Save list with updated annotations
saveRDS(seurat_list, file=paste0(analysis_dir, 
                                 "seurat_list_splitByNMFrank", rank_use, "_updated.detailed.annot.Rds"))


```

## Plot total cd4/cd8 annot by program- compare to estimation at beginning of code
```{r}
rank_use<-8
seurat_list<- readRDS(paste0(analysis_dir, 
                                 "seurat_list_splitByNMFrank", rank_use, "_updated.detailed.annot.Rds"))

all_meta<- lapply(names(seurat_list), function(x){
  seurat_tmp<- seurat_list[[x]]
  tmp<-data.frame(annot=seurat_tmp$detailed_annot_byProgram, Program=x)
  return(tmp)
})
all_meta_withUnclear<- do.call("rbind", all_meta)

## remove "unclear" cells 
all_meta<- all_meta_withUnclear[all_meta_withUnclear$annot != "Unclear",]

## Breakdown of percent of cd4/cd8 cells per program
df_bd<- as.data.frame(table(all_meta$annot, all_meta$Program))
colnames(df_bd)<-c("CellType", "Program", "NCells")
df_bd<- df_bd %>% group_by(Program) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()

## Plot
ggplot(df_bd, aes(x=Program, y=PerCells, fill=CellType))+
  geom_bar(stat="identity", position="stack")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("") + ylab("Proportion of cells")+
  geom_hline(yintercept = 0.5, linetype="dashed")
ggsave(paste0(figure_dir, "CD4.CD8.annotationBy_NMFrank", rank_use, ".png"), width=5, height=3)

saveRDS(all_meta_withUnclear, file=paste0(analysis_dir, "all_annotations_byNMFProgram.Rds"))
```

## Using cells with CD4/CD8 expression as "ground truth", what % of cells are "correctly annotated"
Comparing to:
1. NMF-based cluster-wise annotation (this code)
2. cluster-wise annotation across all ped/adult tcells (in tcell_seurat object, processed in deNovoPedTcell.v.deNovoAdultTcell.Rmd)
```{r}
## Read in "ground truth" (processed in first chunk)
cd4.cd8_expr<- readRDS(paste0(analysis_dir, "cd4.cd8_expression_values.Rds"))

## Read in annotations based on NMF (this code)
nmf_annots<- readRDS(paste0(analysis_dir, "all_annotations_byNMFProgram.Rds"))
nmf_annots<- nmf_annots[nmf_annots$annot!="Unclear",]


## Read in tcell seurat (contains annots based on ped/adult tcells all together; comparison #2)
tcell_seurat<- readRDS(paste0("../../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/Tcells_Ped.Adult_Merge/",
                              "adult_ped_tcell_merged_seurat_harmony.Rds"))

tcell_seurat$cell<- colnames(tcell_seurat)
tcell_meta<- tcell_seurat@meta.data[tcell_seurat$detailed_annot != "TooLow", c("detailed_annot", "cell")]

## Subset 3 annotations to common cells
all_annots<- data.frame(cells=rownames(cd4.cd8_expr), byExpression=cd4.cd8_expr$annotate)
all_annots<- all_annots[all_annots$cells %in% rownames(nmf_annots) &
                          all_annots$cells %in% tcell_meta$cell,]

## Merge all annotations into single df
all_annots$byNMF<- plyr::mapvalues(all_annots$cells, rownames(nmf_annots), nmf_annots$annot,warn_missing = FALSE)
all_annots$byAllTcells<- plyr::mapvalues(all_annots$cells, tcell_meta$cell, tcell_meta$detailed_annot, warn_missing = FALSE)

## annotation by NMF slightly outperforms annotation by all tcells
sum(all_annots$byExpression==all_annots$byNMF) ## 6899 (90%)
sum(all_annots$byExpression==all_annots$byAllTcells) ## 6494 (84%)

## Add on program
all_annots$Program<- plyr::mapvalues(all_annots$cells, rownames(nmf_annots), nmf_annots$Program, warn_missing = FALSE)

## Number of mismatches between expression/NMF annotation by program
byNMF_mismatch<- all_annots %>% group_by(Program) %>% summarise(mismatch=sum(byExpression != byNMF)) %>% as.data.frame()
total_cells<- as.data.frame(table(all_annots$Program))
byNMF_mismatch$nCells<- as.numeric(as.character(plyr::mapvalues(byNMF_mismatch$Program, total_cells$Var1, total_cells$Freq)))
byNMF_mismatch$perCells_mismatchByNMF<- byNMF_mismatch$mismatch/byNMF_mismatch$nCells

## Number of mismatches between expression/NMF annotation by program
byAllTcells_mismatch<- all_annots %>% group_by(Program) %>% summarise(mismatch=sum(byExpression != byAllTcells)) %>% as.data.frame()
byAllTcells_mismatch$nCells<- as.numeric(as.character(plyr::mapvalues(byAllTcells_mismatch$Program, 
                                                                       total_cells$Var1, total_cells$Freq)))
byAllTcells_mismatch$perCells_mismatchByAllTcells<- byAllTcells_mismatch$mismatch/byAllTcells_mismatch$nCells

## Merge together
sum(byNMF_mismatch$Program==byAllTcells_mismatch$Program); nrow(byNMF_mismatch)
all_mismatch<- data.frame(program=byNMF_mismatch$Program,
                          byNMF=byNMF_mismatch$perCells_mismatchByNMF,
                          byAllTcells=byAllTcells_mismatch$perCells_mismatchByAllTcells)
all_mismatch<- melt(all_mismatch, id="program")
colnames(all_mismatch)<-c("Program", "AnnotationApproach", "ProportionMismatched")

ggplot(all_mismatch, aes(x=Program, y=ProportionMismatched, fill=AnnotationApproach))+
  geom_bar(stat="identity", position="dodge")+
  theme(axis.text.x = element_text(angle=45, hjust=1))
#ggsave(paste0(figure_dir, "CD4.CD8.annot.approaches_byNMF.vs.allTcellsTogether_comparedToByExpr.png"), width=5, height=3)
```

## Add CD4/CD8 annotation onto full tcell seurat object
```{r}
nmf_annotations<- readRDS(paste0(analysis_dir, "all_annotations_byNMFProgram.Rds"))
variable_name<- paste0("detailed_annot_byNMFrank", rank_use, "_mergedBy", union_intersect)

## Add to tcell seurat object
tcell_seurat@meta.data[[variable_name]]<- plyr::mapvalues(colnames(tcell_seurat), rownames(nmf_annotations), nmf_annotations$annot)


saveRDS(tcell_seurat, file=paste0("../../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/Tcells_Ped.Adult_Merge/",
                              "adult_ped_tcell_merged_seurat_harmony.Rds"))
```

## Split into cd4/cd8, re-run clustering
```{r}
variable_name<- paste0("detailed_annot_byNMFrank", rank_use, "_mergedBy", union_intersect)
tcell_seurat$tmp<- tcell_seurat@meta.data[[variable_name]]

## CD4
  ## Subset to CD4
  cd4_seurat<-subset(tcell_seurat, tmp=="CD4")
  cd4_meta<- cd4_seurat@meta.data
  
  ## Rerun clustering
  cd4_seurat<- RunFullSeurat(cd4_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=cd4_meta$sampleid)
  cd4_seurat<- AddMetaData(cd4_seurat, cd4_meta[,!colnames(cd4_meta) %in% colnames(cd4_seurat@meta.data)])
  
  ## Quick vis to check for any clear CD8- some in cluster 2. sub-clustering shows small cd8 cluster, mostly in stress, but don't cluster in stress seurat. kept.
  DimPlot(cd4_seurat, label=TRUE, group.by = "seurat_clusters")
  VlnPlot(cd4_seurat, features=c("CD4", "CD8A"), ncol=1)
  FeaturePlot(cd4_seurat, features=c("CD4", "CD8A"), ncol=1) + DimPlot(cd4_seurat, label=TRUE)
  
  
## CD8
  ## Subset to CD4
  cd8_seurat<-subset(tcell_seurat, tmp=="CD8")
  cd8_meta<- cd8_seurat@meta.data
  
  ## Rerun clustering
  cd8_seurat<- RunFullSeurat(cd8_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=cd8_meta$sampleid)
  cd8_seurat<- AddMetaData(cd8_seurat, cd8_meta[,!colnames(cd8_meta) %in% colnames(cd8_seurat@meta.data)])
  
  ## Quick vis to check for any clear CD8- some in cluster 2. sub-clustering shows small cd8 cluster, mostly in stress, but don't cluster in stress seurat. kept.
  DimPlot(cd8_seurat, label=TRUE, group.by = "seurat_clusters")
  VlnPlot(cd8_seurat, features=c("CD4", "CD8A"), ncol=1)
  FeaturePlot(cd8_seurat, features=c("CD4", "CD8A"), ncol=1) + DimPlot(cd8_seurat, label=TRUE) 


saveRDS(cd4_seurat, file=paste0(analysis_dir, "ped.adult_seurat_cd4_harmony.Rds"))
saveRDS(cd8_seurat, file=paste0(analysis_dir, "ped.adult_seurat_cd8_harmony.Rds"))
```

## Quick visualizations to confirm
```{r}
FeaturePlot(cd4_seurat, features=c("CD4", "CD8A", "GZMK", "TOP2A", "CCR7", "FOXP3"), ncol=3)
ggsave(paste0(figure_dir, "CD4_MarkerGeneExpression.png"), width=9, height=6)

FeaturePlot(cd8_seurat, features=c("CD4", "CD8A", "GZMK", "TOP2A", "CCR7", "FOXP3"), ncol=3)
ggsave(paste0(figure_dir, "CD8_MarkerGeneExpression.png"), width=9, height=6)
```


## Add finalized annotations to full immune seurat object (for ped only)
```{r}
cohort<- "pedOnly_nomuv63"
immune_seurat<- readRDS(paste0("../../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/", 
                               cohort, "/tcell.myeloid_seurat_harmony.Rds"))
immune_seurat$detailed_annot_byNMF<- ifelse(colnames(immune_seurat) %in% colnames(tcell_seurat),
                                            plyr::mapvalues(colnames(immune_seurat), colnames(tcell_seurat),
                                                            tcell_seurat$detailed_annot_byNMFrank8_mergedByunion,
                                                            warn_missing = FALSE),
                                            immune_seurat$detailed_annot)
immune_seurat$detailed_annot_byNMF<- gsub("Tcell_UnclearCD4.CD8", "Unclear", fixed = TRUE, immune_seurat$detailed_annot_byNMF)
saveRDS(immune_seurat, file=paste0("../../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/", 
                               cohort, "/tcell.myeloid_seurat_harmony.Rds"))
```

## Split cells by cd4/cd8, re-score for programs
```{r}
## Read in cd4/cd8 seurat objects
cd4_seurat<- readRDS(paste0(analysis_dir, "ped.adult_seurat_cd4_harmony.Rds"))
cd8_seurat<- readRDS(paste0(analysis_dir, "ped.adult_seurat_cd8_harmony.Rds"))

## Read in marker genes
rank_use<-8; sharedBy<- "union"
program_markers<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                                 "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", sharedBy, "_reAnnot.Rds"))
program_markers<- split(program_markers, f=program_markers$cluster)
program_markers<- lapply(program_markers, function(x){x$gene})

## Subset marker genes to cd4/cd8 specific
cd4_markers<- program_markers[!names(program_markers) %in% c("Cytotoxic", "Predysfunctional")]
cd8_markers<- program_markers[!names(program_markers) %in% c("Treg")]

## Ped/adult together
cm_list<- NormCenter(cd4_seurat@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data+1))
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, cd4_markers)))
saveRDS(scores, file=paste0(analysis_dir, "ped.adult_cd4_program_scores.Rds"))

cm_list<- NormCenter(cd8_seurat@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data+1))
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, cd8_markers)))
saveRDS(scores, file=paste0(analysis_dir, "ped.adult_cd8_program_scores.Rds"))

## Ped only
cd4_tmp<- subset(cd4_seurat, Age=="Ped")
cd8_tmp<- subset(cd8_seurat, Age=="Ped")

cm_list<- NormCenter(cd4_tmp@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data+1))
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, cd4_markers)))
saveRDS(scores, file=paste0(analysis_dir, "ped_cd4_program_scores.Rds"))

cm_list<- NormCenter(cd8_tmp@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data+1))
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, cd8_markers)))
saveRDS(scores, file=paste0(analysis_dir, "ped_cd8_program_scores.Rds"))

## Adult only
cd4_tmp<- subset(cd4_seurat, Age=="Adult")
cd8_tmp<- subset(cd8_seurat, Age=="Adult")

cm_list<- NormCenter(cd4_tmp@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data+1))
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, cd4_markers)))
saveRDS(scores, file=paste0(analysis_dir, "adult_cd4_program_scores.Rds"))

cm_list<- NormCenter(cd8_tmp@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data+1))
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, cd8_markers)))
saveRDS(scores, file=paste0(analysis_dir, "adult_cd8_program_scores.Rds"))

```

