---
title: "Slingshot"
output: html_document
---

## Purpose:
Use Slingshot for pseudotime analysis to try to determine whether there is any trajectory within T cells, particularly for cytotoxic/dysfunctional


## Approach:
Run Slingshot on all programs to start, consider removing programs that don't contribute


```{r}
library(Seurat) 
library(monocle)
library(dplyr)
library(plyr)


base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/", 
                     "04a_Trajectory/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/Monocle/")
figure_dir_tmp<- paste0(working_dir, "/figures/Monocle/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)

```

## Read in ped tcell seurat object. Contains NMF assignments (annots added in deNovoPedTcell.v.deNovoAdultTcell, based on correlation with adult T cell + correlation with ped cd4/cd8 programs)
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../02_detailed_annot/analysis/identifyCD4.CD8_byNMF/", "rank",rank_use, "_sharedBy", sharedBy, "/" )

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
  
## ped/adult tcell seurat objects
#cd4_seurat<- readRDS(paste0(data_dir, "ped.adult_seurat_cd4_harmony.Rds"))
cd8_seurat<-readRDS(paste0(data_dir, "ped.adult_seurat_cd8_harmony.Rds"))

## Read in colors to use
program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, "_reAnnot.Rds"))

variable_name<- paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
cd8_seurat$CellAnnot<- cd8_seurat@meta.data[[variable_name]]

```

```{r}
cd8_seurat<- subset(cd8_seurat, CellAnnot !=  c("Inhibitory"))
```



```{r}
## Create CellDataSet object
pd<-new("AnnotatedDataFrame", data = cd8_seurat@meta.data)
fd<- as.data.frame(rownames(cd8_seurat))
colnames(fd)<-c("gene_short_name")
rownames(fd)<- fd$gene_short_name
fd <- new("AnnotatedDataFrame", data =fd)

HSMM <- newCellDataSet(cd8_seurat@assays$RNA@counts,
                phenoData = pd,
                featureData = fd,
                lowerDetectionLimit = 0.1,
                expressionFamily = tobit(Lower = 0.1))

## Converting TPM/FPKM values into mRNA counts
rpc_matrix <- relative2abs(HSMM, method = "num_genes")
HSMM <- newCellDataSet(as(as.matrix(rpc_matrix), "sparseMatrix"),
                phenoData = pd,
                featureData = fd,
                lowerDetectionLimit = 0.5,
                expressionFamily = negbinomial.size())

## Estimate size factors and dispersions
HSMM <- estimateSizeFactors(HSMM)
HSMM <- estimateDispersions(HSMM)

## Determine expressed genes
HSMM <- detectGenes(HSMM, min_expr = 0.1)
expressed_genes <- row.names(subset(fData(HSMM),
    num_cells_expressed >= 10))

## Remove low recovery cells and potential doublets
pData(HSMM)$Total_mRNAs <- Matrix::colSums(exprs(HSMM))

HSMM <- HSMM[,pData(HSMM)$Total_mRNAs < 1e6]

upper_bound <- 10^(mean(log10(pData(HSMM)$Total_mRNAs)) +
            2*sd(log10(pData(HSMM)$Total_mRNAs)))
lower_bound <- 10^(mean(log10(pData(HSMM)$Total_mRNAs)) -
            2*sd(log10(pData(HSMM)$Total_mRNAs)))

qplot(Total_mRNAs, data = pData(HSMM), color = CellAnnot, geom =
"density") +
geom_vline(xintercept = lower_bound) +
geom_vline(xintercept = upper_bound)
HSMM <- HSMM[,pData(HSMM)$Total_mRNAs > lower_bound &
      pData(HSMM)$Total_mRNAs < upper_bound]
HSMM <- detectGenes(HSMM, min_expr = 0.1)

  ## Confirm filtering looks good- expect roughly lognormal distribution
    # Log-transform each value in the expression matrix.
    L <- log(exprs(HSMM[expressed_genes,]))
    
    # Standardize each gene, so that they are all on the same scale,
    # Then melt the data with plyr so we can plot it easily
    melted_dens_df <- melt(Matrix::t(scale(Matrix::t(L))))
    
    # Plot the distribution of the standardized gene expression values.
    qplot(value, geom = "density", data = melted_dens_df) +
    stat_function(fun = dnorm, size = 0.5, color = 'red') +
    xlab("Standardized log(FPKM)") +
    ylab("Density")
    
    
    
## Reduce Dims --> tsne (not necessary for monocle, just used for quick vis)
HSMM <- reduceDimension(HSMM, max_components = 2, num_dim = 6,
                reduction_method = 'tSNE', verbose = T)
HSMM <- clusterCells(HSMM, num_clusters = 2)
plot_cell_clusters(HSMM, 1, 2, color = "CellAnnot")


## Trajectory step 1: choose genes that define a cell's progress
  ## Select superset of feature genes as genes expressed in at least 5% of all the cells
  HSMM <- detectGenes(HSMM, min_expr = 0.1)
  fData(HSMM)$use_for_ordering <-
      fData(HSMM)$num_cells_expressed > 0.05 * ncol(HSMM)
  
  ## Perform PCA to determine which PCs to use
  plot_pc_variance_explained(HSMM, return_all = F)
  
  ## Run Tsne on those top PCs
  HSMM <- reduceDimension(HSMM,
                              max_components = 2,
                              norm_method = 'log',
                              num_dim = 4,
                              reduction_method = 'tSNE',
                              verbose = T)
  
  ## Use to help decide on cell cluster threshold
  plot_rho_delta(HSMM, rho_threshold = 2, delta_threshold = 4 )
  
  ## Identify clusters using densityPeak algorithm
  HSMM <- clusterCells(HSMM,
                 rho_threshold = 2,
                 delta_threshold = 15,## lower = more clusters
                 skip_rho_sigma = T, ## may have to set this to true when clustering isn't already run
                 verbose = F)
  plot_cell_clusters(HSMM, color_by = 'as.factor(Cluster)')
  plot_cell_clusters(HSMM, color_by = 'as.factor(CellAnnot)')
  
  
## Perform DE
clustering_DEG_genes <-differentialGeneTest(HSMM[expressed_genes,],
                                            fullModelFormulaStr = '~Cluster',
                                            cores = 1)

## Select top sig genes as the ordering genes
HSMM_ordering_genes <-row.names(clustering_DEG_genes)[order(clustering_DEG_genes$qval)][1:1000]

HSMM <-setOrderingFilter(HSMM,ordering_genes = HSMM_ordering_genes)

## Trajectory step 2: reduce data dimensionality
HSMM <-reduceDimension(HSMM, method = 'DDRTree')

## Trajectory step 3: order cells along the trajectory
HSMM <-orderCells(HSMM)

  ## visualize the trajectory in the reduced dimensional space.
  plot_cell_trajectory(HSMM, color_by = "CellAnnot")
  ggsave(paste0(figure_dir, "trajectory_byCellAnnot_allPrograms.png"), width=6, height=6)
  plot_cell_trajectory(HSMM, color_by = "State")
  plot_cell_trajectory(HSMM, color_by = "Pseudotime")
  
  plot_cell_trajectory(HSMM, color_by = "State") +
    facet_wrap(~State, nrow = 1)
  
  plot_genes_jitter(HSMM[c("GNLY", "PDCD1", "GZMK", "TIGIT"),],
    grouping = "State",
    min_expr = 0.1)
  
  
  saveRDS(HSMM, file=paste0(analysis_dir, "monocle_obj_allPrograms.Rds"))
```

