---
title: "Slingshot"
output: html_document
---

## Purpose:
Use Slingshot for pseudotime analysis to try to determine whether there is any trajectory within T cells, particularly for cytotoxic/dysfunctional


## Approach:
Run Slingshot on all programs to start, consider removing programs that don't contribute


```{r}
library(Seurat) 
library(slingshot)
library(BUSpaRse)
library(tidyverse)
library(tidymodels)
library(scales)
library(viridis)
library(Matrix)
library(uwot)
library(mclust, quietly = TRUE)
library(grDevices)
library(tradeSeq)


base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/", 
                     "04_Trajectory/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/Slingshot/")
figure_dir_tmp<- paste0(working_dir, "/figures/Slingshot/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)

```

```{r}
## Function for color palette
cell_pal <- function(cell_vars, pal_fun,...) {
  if (is.numeric(cell_vars)) {
    pal <- pal_fun(100, ...)
    return(pal[cut(cell_vars, breaks = 100)])
  } else {
    categories <- sort(unique(cell_vars))
    pal <- setNames(pal_fun(length(categories), ...), categories)
    return(pal[cell_vars])
  }
}
```

## Read in ped tcell seurat object. Contains NMF assignments (annots added in deNovoPedTcell.v.deNovoAdultTcell, based on correlation with adult T cell + correlation with ped cd4/cd8 programs)
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../02_detailed_annot/analysis/identifyCD4.CD8_byNMF/", "rank",rank_use, "_sharedBy", sharedBy, "/" )

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
  
## ped/adult tcell seurat objects
#cd4_seurat<- readRDS(paste0(data_dir, "ped.adult_seurat_cd4_harmony.Rds"))
cd8_seurat<-readRDS(paste0(data_dir, "ped.adult_seurat_cd8_harmony.Rds"))

## Read in colors to use
program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, "_reAnnot.Rds"))

variable_name<- paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
cd8_seurat$CellAnnot<- cd8_seurat@meta.data[[variable_name]]

```

## Settings
```{r}
analysis_name<- "CD8_NoStress.Memory"


## Filtering
programs_keep<- unique(cd8_seurat$CellAnnot)[!unique(cd8_seurat$CellAnnot) %in% c("Stress", "Memory")]
seurat_use<- subset(cd8_seurat, CellAnnot %in% programs_keep)
```


## Create singleCellExperiment object, filter, norm, and run PCA/UMAP
```{r}
sce <- SingleCellExperiment(assays = List(counts = seurat_use@assays$RNA@counts))

## Filter
geneFilter <- apply(assays(sce)$counts,1,function(x){
    sum(x >= 3) >= 10
})
sce <- sce[geneFilter, ]

## Norm
FQnorm <- function(counts){
    rk <- apply(counts,2,rank,ties.method='min')
    counts.sort <- apply(counts,2,sort)
    refdist <- apply(counts.sort,1,median)
    norm <- apply(rk,2,function(r){ refdist[r] })
    rownames(norm) <- rownames(counts)
    return(norm)
}
assays(sce)$norm <- FQnorm(assays(sce)$counts)

## PCA
pca <- prcomp(t(log1p(assays(sce)$norm)), scale. = FALSE)
rd1 <- pca$x[,1:2]

## UMAP
rd2 <- uwot::umap(t(log1p(assays(sce)$norm)))
colnames(rd2) <- c('UMAP1', 'UMAP2')

## Add reduced dimensions to sce
reducedDims(sce) <- SimpleList(PCA = rd1, UMAP = rd2)

## Add cluster labels
colData(sce)$NMF_program <- seurat_use$CellAnnot

## Save
saveRDS(sce, file=paste0(analysis_dir, "sce_", analysis_name, ".Rds"))
```


## Plot PCA/UMAP
```{r}
plot(rd1, col = rgb(0,0,0,.5), pch=16, asp = 1)
plot(rd2, col = rgb(0,0,0,.5), pch=16, asp = 1)
```


## Run slignshot
two-step process composed of identifying the global lineage structure with a cluster-based minimum spanning tree (MST) and fitting simultaneous principal curves to describe each lineage
```{r}
## Run slingshot
sce <- slingshot(sce, clusterLabels = "NMF_program", reducedDim = 'PCA') ## ,start.clus="Memory"

summary(sce$slingPseudotime_1)

saveRDS(sce, file=paste0(analysis_dir, "sce_", analysis_name, ".Rds"))
```


## Plot results
```{r}
library(ggpubr)

cell_colors <- cell_pal(sce$NMF_program, brewer_pal("qual", "Set2"))

## Legend to show which color corresponds to which program
cell_legend<- data.frame(color=cell_colors, program=names(cell_colors))
cell_legend$ID<- paste0(cell_legend$color, cell_legend$program)
cell_legend<- cell_legend[!duplicated(cell_legend$ID),]
cell_legend$dummy<- 1
colors_use<- cell_legend$color; names(colors_use)<- cell_legend$program
p<-ggplot(cell_legend, aes(x=program, y=dummy, fill=program))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=colors_use)+
  theme(axis.text.x = element_text(angle=45, hjust=1))
as_ggplot(get_legend(p))
ggsave(paste0(figure_dir, "SS_PCA_", analysis_name, "_legend.png"), width=3, height=4)

## Plot in PCA space
png(filename=paste0(figure_dir, "SS_PCA_", analysis_name, ".png"), width=1000, height=1000)
plot(reducedDims(sce)$PCA, col = cell_colors, pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, type = 'lineages', col = 'black')
dev.off()
```

## Alternative to using NMF to label: run clustering
```{r}
## Clustering: gaussian mixture
cl1 <- Mclust(rd1)$classification
colData(sce)$GMM <- cl1

plot(rd1, col = brewer.pal(9,"Set1")[cl1], pch=16, asp = 1)

## Clustering: k means
cl2 <- kmeans(rd1, centers = 4)$cluster
colData(sce)$kmeans <- cl2

plot(rd1, col = brewer.pal(9,"Set1")[cl2], pch=16, asp = 1)
```

## Identify temporally dynamic genes
For each gene, we will fit a general additive model (GAM) using a negative binomial noise distribution to model the (potentially nonlinear) relationshipships between gene expression and pseudotime. We will then test for significant associations between expression and pseudotime using the associationTest.
```{r}
# fit negative binomial GAM
sce <- fitGAM(sce) ## runs for 2 hours!!!

# test for dynamic expression
ATres <- associationTest(sce)

## Plot top genes
topgenes <- rownames(ATres[order(ATres$pvalue), ])[1:250]
pst.ord <- order(sce$slingPseudotime_1, na.last = NA)
heatdata <- assays(sce)$counts[topgenes, pst.ord]
heatclus <- sce$GMM[pst.ord]

heatmap(log1p(heatdata), Colv = NA,
        ColSideColors = brewer.pal(9,"Set1")[heatclus])
```




