---
title: "Stream preprocessing"
output: html_document
---

## Preprocess for input into stream (python)
End of code: read in stream results for plotting

```{r}
library(Seurat) 
library(SeuratDisk)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/", 
                     "04a_Trajectory/")

analysis_dir<- paste0(working_dir, "/analysis/STREAM/")

if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
```

## Read in ped tcell seurat object. Contains NMF assignments (annots added in deNovoPedTcell.v.deNovoAdultTcell, based on correlation with adult T cell + correlation with ped cd4/cd8 programs)
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../03_Visualize.Reannotate_programs/analysis/Reannotate_programs/", 
                  "rank",rank_use, "_sharedBy", sharedBy, "/" )

cd4_seurat<- readRDS(paste0(data_dir, "ped.adult_seurat_cd4_harmony.Rds"))
cd8_seurat<-readRDS(paste0(data_dir, "ped.adult_seurat_cd8_harmony.Rds"))

## Remove norm/scaled data- want X to be raw data
cd8_seurat<- DietSeurat(cd8_seurat, scale.data = FALSE)

## subset to predys, cyto, cycling
cd8_subset<- subset(cd8_seurat, Final_Annot %in% c("Predysfunctional"))

SaveH5Seurat(cd4_seurat, filename = paste0(analysis_dir, "cd4.h5Seurat"))
Convert( paste0(analysis_dir, "cd4.h5Seurat"), dest = "h5ad")

SaveH5Seurat(cd8_seurat, filename = paste0(analysis_dir, "cd8.h5Seurat"))
Convert( paste0(analysis_dir, "cd8.h5Seurat"), dest = "h5ad")

SaveH5Seurat(cd8_subset, filename = paste0(analysis_dir, "cd8_subset_Cyt.Predys.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_subset_Cyt.Predys.h5Seurat"), dest = "h5ad")

SaveH5Seurat(cd8_subset, filename = paste0(analysis_dir, "cd8_subset_Cyt.Predys.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_subset_Cyt.Predys.h5Seurat"), dest = "h5ad")

SaveH5Seurat(cd8_subset, filename = paste0(analysis_dir, "cd8_subset_Predys.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_subset_Predys.h5Seurat"), dest = "h5ad")
```


## As reference- melanoma and/or nsclc CD8 cells
```{r}
nsclc_cd8<- readRDS("../04b_ComparisonToOtherTumor/analysis/NSCLC/cd8_seurat_nsclc.Rds")
melanoma_cd8<- readRDS("../04b_ComparisonToOtherTumor/analysis/Melanoma/cd8_seurat_melanoma.Rds")
cd8_seurat<- readRDS("../04b_ComparisonToOtherTumor/analysis/NSCLC.melanoma/cd8_seurat_nsclc.glioma.melanoma.Rds")

## Add on cd8 cell state from merged analysis
nsclc_cd8$maxGeneset<- as.character(plyr::mapvalues(colnames(nsclc_cd8), colnames(cd8_seurat), cd8_seurat$maxGeneset, warn_missing = FALSE))
melanoma_cd8$maxGeneset<- as.character(plyr::mapvalues(colnames(melanoma_cd8), colnames(cd8_seurat), cd8_seurat$maxGeneset, warn_missing = FALSE))
nsclc_cd8$maxGeneset<- gsub("[0-9]", "", nsclc_cd8$maxGeneset)
melanoma_cd8$maxGeneset<- gsub("[0-9]", "", melanoma_cd8$maxGeneset)

nsclc_cd8<- DietSeurat(nsclc_cd8, scale.data = FALSE)
melanoma_cd8<- DietSeurat(melanoma_cd8, scale.data = FALSE)

SaveH5Seurat(nsclc_cd8, filename = paste0(analysis_dir, "cd8_nsclc.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_nsclc.h5Seurat"), dest = "h5ad")

SaveH5Seurat(melanoma_cd8, filename = paste0(analysis_dir, "cd8_melanoma.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_melanoma.h5Seurat"), dest = "h5ad")

## cytotoxic nsclc cells
cyto_nsclc<- subset(nsclc_cd8, majorCluster_annot=="Cytotoxic")

SaveH5Seurat(cyto_nsclc, filename = paste0(analysis_dir, "cd8_nsclc_cyto.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_nsclc_cyto.h5Seurat"), dest = "h5ad")
```


## glioma + melanoma
```{r}
cd8_seurat<- readRDS("../04b_ComparisonToOtherTumor/analysis/NSCLC.melanoma/cd8_seurat_nsclc.glioma.melanoma.Rds")

## Add on metadata
cd8_seurat$Age<- ifelse(cd8_seurat$TumorType=="Glioma", ifelse(cd8_seurat$age_numeric<25, "Ped", "Adult"), "Adult")
cd8_seurat$TumorType_Age<- ifelse(cd8_seurat$TumorType=="Glioma", paste0("Glioma_", cd8_seurat$Age), cd8_seurat$TumorType)
cd8_seurat$maxGeneset<- gsub("[0-9]", "", cd8_seurat$maxGeneset)
cd8_seurat$CytoDysScore<- cd8_seurat$cytotoxic1-cd8_seurat$dysfunctional4

## Split  by tumor type
cd8_list<- SplitObject(cd8_seurat, split.by="TumorType")


## Merge melanoma/cytotoxic together
gm_cd8<- merge(cd8_list$Glioma, cd8_list$Melanoma)
gm_cd8_meta<- gm_cd8@meta.data
gm_cd8<- RunFullSeurat(gm_cd8@assays$RNA@counts, RunHarmony = FALSE, samples=gm_cd8$sample)
gm_cd8<- AddMetaData(gm_cd8, gm_cd8_meta[,!colnames(gm_cd8_meta)  %in% colnames(gm_cd8@meta.data)])

## Remove norm/scaled data- want X to be raw data
gm_cd8<- DietSeurat(gm_cd8, scale.data=FALSE)

SaveH5Seurat(gm_cd8, filename = paste0(analysis_dir, "cd8_glioma.melanoma.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_glioma.melanoma.h5Seurat"), dest = "h5ad")
```

## For mapping onto nsclc: merged melanoma and glioma CD8 cells
```{r}
cd8_seurat<- readRDS("../04b_ComparisonToOtherTumor/analysis/NSCLC.melanoma/cd8_seurat_nsclc.glioma.melanoma.Rds")

## remove nsclc
cd8_seurat<- subset(cd8_seurat, TumorType != "NSCLC")

## add on annotation
cd8_seurat$Annot<- ifelse(cd8_seurat$TumorType=="Glioma", cd8_seurat$cellAnnot, cd8_seurat$maxGeneset)

## Add on age for glioma
cd8_seurat$Tumor_Age<- ifelse(cd8_seurat$TumorType=="Glioma", ifelse(as.numeric(cd8_seurat$age_numeric)<=25, "Glioma_Ped", "Glioma_Adult"), "Melanoma")
cd8_seurat<- DietSeurat(cd8_seurat, scale.data = FALSE)

SaveH5Seurat(cd8_seurat, filename = paste0(analysis_dir, "cd8_glioma.melanoma.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_glioma.melanoma.h5Seurat"), dest = "h5ad")
```

## glioma, nsclc, and melanoma all together
```{r}
cd8_seurat<- readRDS("../04b_ComparisonToOtherTumor/analysis/NSCLC.melanoma/cd8_seurat_nsclc.glioma.melanoma.Rds")

## edit metadata
cd8_seurat$Tumor_Age<- ifelse(cd8_seurat$TumorType=="Glioma", 
                              ifelse(as.numeric(cd8_seurat$age_numeric)<=25, "Glioma_Ped", "Glioma_Adult"), cd8_seurat$TumorType)
cd8_seurat@meta.data<- cd8_seurat@meta.data %>% dplyr::rename(cytotoxic=cytotoxic1,
                                                              naive=naive2,
                                                              predysfunctional=predysfunctional3,
                                                              dysfunctional=dysfunctional4) %>% as.data.frame()
cd8_seurat$maxGeneset<- gsub("[1-4]", "", cd8_seurat$maxGeneset)


cd8_seurat<- DietSeurat(cd8_seurat, scale.data = FALSE)

SaveH5Seurat(cd8_seurat, filename = paste0(analysis_dir, "cd8_glioma.melanoma.nsclc.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_glioma.melanoma.nsclc.h5Seurat"), dest = "h5ad")
```

## Downsampled glioma, nsclc, and melanoma
```{r}
cd8_seurat<- readRDS("../04b_ComparisonToOtherTumor/analysis/NSCLC.melanoma/cd8_seurat_nsclc.glioma.melanoma.Rds")
cd8_seurat$Age<- ifelse(cd8_seurat$TumorType=="Glioma", ifelse(cd8_seurat$age_numeric<25, "Ped", "Adult"), "Adult")
cd8_seurat$TumorType_Age<- ifelse(cd8_seurat$TumorType=="Glioma", paste0("Glioma_", cd8_seurat$Age), cd8_seurat$TumorType)

## Split  by tumor type
cd8_list<- SplitObject(cd8_seurat, split.by="TumorType")

## Downsample all to same size
nDs<- min(unlist(lapply(cd8_list, ncol)))
cd8_list<- lapply(cd8_list, function(x){
  x[, sample(colnames(x), size =nDs, replace=F)]
})


## merge melanoma/glioma
gm_ss<- merge(cd8_list$Glioma, cd8_list$Melanoma)
gm_ss$Annot<- ifelse(gm_ss$TumorType=="Glioma", gm_ss$cellAnnot, gm_ss$maxGeneset)
gm_ss_meta<- gm_ss@meta.data
gm_ss<- RunFullSeurat(gm_ss@assays$RNA@counts, RunHarmony = FALSE, samples=gm_ss$sample)
gm_ss<- AddMetaData(gm_ss, gm_ss_meta[,!colnames(gm_ss_meta)  %in% colnames(gm_ss@meta.data)])


## Remove scaled data
nsclc_ss<- DietSeurat(cd8_list$NSCLC, scale.data = FALSE)
melanoma_glioma_ss<- DietSeurat(gm_ss, scale.data = FALSE)

## save
SaveH5Seurat(nsclc_ss, filename = paste0(analysis_dir, "cd8_nsclc_downsample.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_nsclc_downsample.h5Seurat"), dest = "h5ad")

SaveH5Seurat(melanoma_glioma_ss, filename = paste0(analysis_dir, "cd8_melanoma.glioma_downsample.h5Seurat"))
Convert( paste0(analysis_dir, "cd8_melanoma.glioma_downsample.h5Seurat"), dest = "h5ad")
```

## Read in stream results for some quick plots
Exploratory for now
May move this to its own file if too much plotting
```{r}
## read in adata.obs from stream results
stream_res<- read.csv("stream_result_mappedNSCLC.Glioma.Melanoma.csv", row.names = "X")

## Read in seurat
cd8_seurat<- readRDS("../04b_ComparisonToOtherTumor/analysis/NSCLC.melanoma/cd8_seurat_nsclc.glioma.melanoma.Rds")

## Add stream res to seurat
rownames(stream_res)<- gsub("-ref", "", gsub("-new", "", rownames(stream_res)))
cd8_seurat<- AddMetaData(cd8_seurat, stream_res[,!colnames(stream_res) %in% colnames(cd8_seurat@meta.data)])

## Annotate stream branches
branch_annot<- unique(cd8_seurat$branch_id_alias)
names(branch_annot)<- c("predysfunctional", "cytotoxic_1", "Memory", "cytotoxic_2", "cytotoxic_3", "dysfunctional", "predysfunctional_2")
cd8_seurat$branch_annot<- plyr::mapvalues(cd8_seurat$branch_id_alias, branch_annot, names(branch_annot))

## Plot proportion of each branch by tumor type
  ## Number of each branch by subtype
  df<- as.data.frame(table(stream_res$branch_id_alias, stream_res$Tumor_Age))
  df$Var1<- gsub("\\(", "", gsub(")", "", gsub("\\'", "", gsub(", ", "_", df$Var1))))
  colnames(df)<-c("Branch", "TumorType", "NumberOfCells")
  
  ## proportion by subtype
  df<- df %>% group_by(TumorType) %>% mutate(perCells=NumberOfCells/sum(NumberOfCells)) %>% as.data.frame()
  
  ## Plot
  ggplot(df, aes(x=TumorType, y=perCells, fill=Branch))+
    geom_bar(stat="identity")
  
## Plot GOI by branch assignment
  adult<- DotPlot(subset(cd8_seurat, Tumor_Age=="Glioma_Adult"), group.by = "branch_annot", 
                  features=c("CCR7", "GZMB", "PRF1", "GNLY", "PDCD1", "CTLA4", "KLRB1", "PTGER2", "SELPLG"))+
    scale_color_gradient2(low="blue", mid="white", high="red")+
    theme(axis.text.x = element_text(angle=45, hjust=1))
  
  ped<- DotPlot(subset(cd8_seurat, Tumor_Age=="Glioma_Ped"), group.by = "branch_annot", 
                features=c("CCR7", "GZMB", "PRF1", "GNLY", "PDCD1", "CTLA4", "KLRB1", "PTGER2", "SELPLG"))+
  scale_color_gradient2(low="blue", mid="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1))
  
  adult+ped
```

