---
title: "Preprocessing before Clonal.vs.Nonclonal analysis"
output: html_document
---

## Purpose:
Based on Cell 2021 paper, compare expression patterns in clonal compared to non-clonal CD8 cells. Simple DE is not as robust (according to their paper)- this approach essentially creates background distribution of randomly shuffled clonal/nonclonal labels that you compare to actual clonal/nonclonal labels for gene-wise expression + CV. This analysis is run on O2 (see 02_ClonalVsNonClonal_Analysis.Rmd). Here, necessary preprocessing is done, including subsetting the data by age/subtype.

## Approach:
Using cd8 annotations based on program-wise, adult+adult approach, split seurat into clonal non-clonal. Subset by age/subtype/etc as desired before this, then save. These files used directly on O2.

```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/01b_Adult/", 
                     "02_ClonalOverexpression/")

library(Seurat)
library(ggplot2)
library(ggpubr)
library(Rtreemix)

figure_dir_tmp<- "figures/"
analysis_dir_tmp<- "analysis/"

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(nmf_helper)
clonalOverexpression<- paste0(base_dir, script_dir, "CellChat_HelperFunctions.R")
source(clonalOverexpression)

```


## Set figure dir based on annotation approach (rank/sharedBy changes the final cd4/cd8 annotations)
```{r}
rank_use<-8
sharedBy<- "union" 

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
```

## Read in seurat object
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../../02_Ped.Adult/02_detailed_annot/analysis/identifyCD4.CD8_byNMF/", "rank",rank_use, "_sharedBy", sharedBy, "/" )

## Read in seurat object
cd8_seurat<- readRDS(paste0( data_dir, "adult_seurat_cd8_harmony.Rds"))

## rename clonal
cd8_seurat$clonal<-  gsub("not clonal", "nonclonal", cd8_seurat$clonal)

## rename clonotype
cd8_seurat$clonal_group<-cd8_seurat$clonotype
```


###############################################
## Run preprocessing for multiple conditions ##
###############################################


## All CD8
```{r}
analysis_append<-"" ## append to analysis to reflect subsetting by age/subtype
cd8_seurat_use<-cd8_seurat

## Split into clonal/nonclonal
adult_clonal_seurat<- subset(cd8_seurat_use, clonal=="clonal")
adult_nonclonal_seurat<- subset(cd8_seurat_use, clonal=="nonclonal")

## Some genes are excluded- if have less than 0.1 mean(ln(TP100K + 1)) in both clonal and nonclonal
## For clonal/nonclonal, determine genes that don't pass this filter
## If they don't pass in both, remove
  ## convert tpm --> tp100K (divide by 10)
  clonal_cm<- as.data.frame(adult_clonal_seurat@assays$RNA@counts) / 10
  nonclonal_cm<- as.data.frame(adult_nonclonal_seurat@assays$RNA@counts) / 10
  
  ## mean ln(TPK100 + 1) for each gene
  clonal_ln<- log(clonal_cm + 1); clonal_mean_ln<- rowMeans(clonal_ln)
  nonclonal_ln<- log(nonclonal_cm + 1); nonclonal_mean_ln<- rowMeans(nonclonal_ln)
  
  ## genes that dont PF in clonal/nonclonal
  clonal_fail<- names(clonal_mean_ln)[clonal_mean_ln<0.1] 
  nonclonal_fail<- names(nonclonal_mean_ln)[nonclonal_mean_ln<0.1] 
  fail_both<- clonal_fail[clonal_fail %in% nonclonal_fail] 
  pass_either<- rownames(cd8_seurat)[!rownames(cd8_seurat) %in% fail_both] 
  
  ## subset seurat objects to passing genes
  adult_clonal_seurat<- subset(adult_clonal_seurat, features=pass_either)
  adult_nonclonal_seurat<- subset(adult_nonclonal_seurat, features=pass_either)
  
  
## save
adult_seurat_cnc<- list(clonal=adult_clonal_seurat, nonclonal=adult_nonclonal_seurat)
saveRDS(adult_seurat_cnc, file=paste0(analysis_dir, "adult_cd8_seurat_cnc_preprocessed", analysis_append, ".Rds"))
```



## By Subtype
```{r}
subtypes<- unique(as.character(cd8_seurat$NiceSubtype))

for(subtype in subtypes){
  print(subtype)
  analysis_append<-paste0("_", gsub("/","", subtype)) ## append to analysis to reflect subsetting by age/subtype
  
  cd8_seurat_use<- subset(cd8_seurat, NiceSubtype==subtype)
  
  ## Skip any subtypes without enough clonal/nonclonal cells
  if(length(unique(cd8_seurat_use$clonal)) != 2){
    next
  }
  
  ## Split into clonal/nonclonal
  adult_clonal_seurat<- subset(cd8_seurat_use, clonal=="clonal")
  adult_nonclonal_seurat<- subset(cd8_seurat_use, clonal=="nonclonal")
  
  ## Some genes are excluded- if have less than 0.1 mean(ln(TP100K + 1)) in both clonal and nonclonal
  ## For clonal/nonclonal, determine genes that don't pass this filter
  ## If they don't pass in both, remove
    ## convert tpm --> tp100K (divide by 10)
    clonal_cm<- as.data.frame(adult_clonal_seurat@assays$RNA@counts) / 10
    nonclonal_cm<- as.data.frame(adult_nonclonal_seurat@assays$RNA@counts) / 10
    
    ## mean ln(TPK100 + 1) for each gene
    clonal_ln<- log(clonal_cm + 1); clonal_mean_ln<- rowMeans(clonal_ln)
    nonclonal_ln<- log(nonclonal_cm + 1); nonclonal_mean_ln<- rowMeans(nonclonal_ln)
    
    ## genes that dont PF in clonal/nonclonal
    clonal_fail<- names(clonal_mean_ln)[clonal_mean_ln<0.1] 
    nonclonal_fail<- names(nonclonal_mean_ln)[nonclonal_mean_ln<0.1] 
    fail_both<- clonal_fail[clonal_fail %in% nonclonal_fail] 
    pass_either<- rownames(cd8_seurat)[!rownames(cd8_seurat) %in% fail_both] 
    
    ## subset seurat objects to passing genes
    adult_clonal_seurat<- subset(adult_clonal_seurat, features=pass_either)
    adult_nonclonal_seurat<- subset(adult_nonclonal_seurat, features=pass_either)
    
  
  ## save
  adult_seurat_cnc<- list(clonal=adult_clonal_seurat, nonclonal=adult_nonclonal_seurat)
  saveRDS(adult_seurat_cnc, file=paste0(analysis_dir, "adult_cd8_seurat_cnc_preprocessed", analysis_append, ".Rds"))
  
}


```

