---
title: "R Notebook"
output: html_document
---

## Purpose:
Overall goal is to compare tcells in NSCLC (known to have reasonably good response to checkpoint therapies) to tcells in gliomas. Here, merge processed NSCLC tumor cd8 with glioma, compare GOI expression patterns, breakdown of cell types

## Approach:
Merge processed NSCLC tumor cd8 seurat object with glioma cd8 seurat object. Plot GOI in equivalent programs, plot breakdown of cell types

```{r}
library(Seurat) 
library(harmony)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/", 
                     "02a_Tcells/02_Ped.Adult/04b_ComparisonToOtherTumor/")

analysis_dir<- paste0(working_dir, "/analysis/NSCLC/")
figure_dir<- paste0(working_dir, "/figures/Merge_NSCLC.Glioma/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in seurat objects + marker genes
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../02_detailed_annot/analysis/identifyCD4.CD8_byNMF/", "rank",rank_use, "_sharedBy", sharedBy, "/" )

## Read in seurat objects
nsclc_cd8<- readRDS(paste0(analysis_dir, "cd8_seurat_nsclc.Rds"))
glioma_cd8<- readRDS(paste0(data_dir, "ped.adult_seurat_cd8_harmony.Rds")) 

nsclc_cd4<- readRDS(paste0(analysis_dir, "cd4_seurat_nsclc.Rds"))
glioma_cd4<- readRDS(paste0(data_dir, "ped.adult_seurat_cd4_harmony.Rds")) 

## Read in marker genes for programs
nsclc_cd8_markers<- read.csv(paste0(analysis_dir, "nsclc_cd8_cluster_markers_annot.csv"))
glioma_markers<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                                 "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", sharedBy, "_reAnnot.Rds"))

nsclc_cd4_markers<- read.csv(paste0(analysis_dir, "nsclc_cd4_cluster_markers_annot.csv"))

## Read in colors to use- add on nsclc-specific programs
program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, "_reAnnot.Rds"))
program_colors<- c(program_colors, CD28="pink", Naive="skyblue4", MAIT="brown", Predysfunctional.2="lightsalmon1",
                   Dysfunctional="violetred")
program_colors<-c(program_colors, CD69="pink", GZMA="brown",ANXA1="lightsalmon3", ActivatedTreg="gold")
```


## Merge seurat objects
```{r}
## CD8
  ## Check that scale is equivalent
  head(colSums(nsclc_cd8@assays$RNA@counts))/head(colSums(glioma_cd8@assays$RNA@counts))
  
  ## Match metadata
  nsclc_cd8$cellAnnot<- nsclc_cd8$majorCluster_annot
  glioma_cd8$cellAnnot<- glioma_cd8@meta.data[[paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")]]
  nsclc_cd8$NiceSubtype<- "NSCLC"
  nsclc_cd8$TumorType<- "NSCLC"
  glioma_cd8$TumorType<- "Glioma"
  
  glioma_cd8@meta.data<- glioma_cd8@meta.data[,colnames(glioma_cd8@meta.data) %in% colnames(nsclc_cd8@meta.data)]
  nsclc_cd8@meta.data<- nsclc_cd8@meta.data[,colnames(nsclc_cd8@meta.data) %in% colnames(glioma_cd8@meta.data)]
  
  ## Merge seurat
  merge_cd8<- merge(nsclc_cd8, glioma_cd8)
  
  ## Rerun clustering
  merge_cd8_meta<- merge_cd8@meta.data
  merge_cd8<- RunFullSeurat(merge_cd8@assays$RNA@counts, RunHarmony = TRUE, samples = merge_cd8_meta$sample)
  merge_cd8<- AddMetaData(merge_cd8, merge_cd8_meta[,!colnames(merge_cd8_meta) %in% colnames(merge_cd8@meta.data)])
  
  saveRDS(merge_cd8, file=paste0(analysis_dir, "cd8_seurat_nsclc.glioma.Rds"))
  
## CD4
  ## Check that scale is equivalent
  head(colSums(nsclc_cd4@assays$RNA@counts))/head(colSums(glioma_cd4@assays$RNA@counts))
  
  ## Match metadata
  nsclc_cd4$cellAnnot<- nsclc_cd4$majorCluster_annot
  glioma_cd4$cellAnnot<- glioma_cd4@meta.data[[paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")]]
  nsclc_cd4$NiceSubtype<- "NSCLC"
  nsclc_cd4$TumorType<- "NSCLC"
  glioma_cd4$TumorType<- "Glioma"
  
  glioma_cd4@meta.data<- glioma_cd4@meta.data[,colnames(glioma_cd4@meta.data) %in% colnames(nsclc_cd4@meta.data)]
  nsclc_cd4@meta.data<- nsclc_cd4@meta.data[,colnames(nsclc_cd4@meta.data) %in% colnames(glioma_cd4@meta.data)]
  
  ## Merge seurat
  merge_cd4<- merge(nsclc_cd4, glioma_cd4)
  
  ## Rerun clustering
  merge_cd4_meta<- merge_cd4@meta.data
  merge_cd4<- RunFullSeurat(merge_cd4@assays$RNA@counts, RunHarmony = TRUE, samples = merge_cd4_meta$sample)
  merge_cd4<- AddMetaData(merge_cd4, merge_cd4_meta[,!colnames(merge_cd4_meta) %in% colnames(merge_cd4@meta.data)])
  
  saveRDS(merge_cd4, file=paste0(analysis_dir, "cd4_seurat_nsclc.glioma.Rds"))
```

## Basic plots
```{r}
merge_cd8<- readRDS(paste0(analysis_dir, "cd8_seurat_nsclc.glioma.Rds"))
merge_cd4<- readRDS(paste0(analysis_dir, "cd4_seurat_nsclc.glioma.Rds"))
  
## CD8
  ## By glioma/nsclc
  DimPlot(merge_cd8, group.by = "TumorType", label=TRUE)+NoLegend()
  ggsave(paste0(figure_dir, "UMAP_byTumorType.png"), width=5, height=5)
  
  ## By cell type
  DimPlot(merge_cd8, group.by = "cellAnnot")+
    scale_color_manual(values=program_colors)
  ggsave(paste0(figure_dir, "UMAP_byCellType.png"), width=10, height=9)
  
  ## Dotplot of GOI by cell type
  merge_cd8$cellAnnot_byTumor<- paste0( merge_cd8$cellAnnot, "_", merge_cd8$TumorType)
  merge_cd8$cellAnnot_byTumor<- factor(merge_cd8$cellAnnot_byTumor, levels=rev(c("Inhibitory_Glioma" 
                                                                             ,"CD28_NSCLC","MAIT_NSCLC","Stress_Glioma","IFN.Response_Glioma","Effector.Memory_Glioma" ,
                                                                             "Naive_NSCLC","Memory_Glioma",
                                                                             "Cytotoxic_NSCLC", "Cytotoxic_Glioma",
                                                                             "Predysfunctional_NSCLC", "Predysfunctional.2_NSCLC", "Predysfunctional_Glioma",
                                                                             "Dysfunctional_NSCLC", "Inhibitory.Cycling_Glioma" )))
  DotPlot(merge_cd8, group.by = "cellAnnot_byTumor", 
            features=c("CD28","CXCR1", "PRF1", "GNLY","PDCD1", "HAVCR2", "TIGIT", "CTLA4",
                       "IFIT1","TOP2A","SLC4A10", "SELL", "CCR7", "GZMK", "ITGAE", "HSPA6","CD8A", "CD4"))+
      theme(axis.text.x = element_text(angle=45, hjust=1))
  ggsave(paste0(figure_dir, "Dotplot_CD8_GOI_byCellType.png"), width=10, height=6)
  
  ## Dotplot of inhibitory GOI by cell type
  DotPlot(merge_cd8, group.by = "cellAnnot_byTumor", 
            features=c("PDCD1", "HAVCR2", "TIGIT", "CTLA4", "GZMK", "ITGAE"))+
      theme(axis.text.x = element_text(angle=45, hjust=1))
  ggsave(paste0(figure_dir, "Dotplot_CD8_InhibitoryGOI_byCellType.png"), width=6, height=6)
  
  
  ## Cell type breakdown
  merge_cd8$TumorType_byAge<- ifelse(merge_cd8$NiceSubtype %in% c("GBM", "IDH-G"), "Glioma_Adult",
                                     ifelse(merge_cd8$TumorType=="Glioma", "Glioma_Ped", "NSCLC"))
  bd<- merge_cd8@meta.data %>% group_by(TumorType_byAge, cellAnnot) %>% 
    summarise(nCells=n()) %>%
    group_by(TumorType_byAge) %>%
    mutate(perCells=nCells/sum(nCells)) %>%
    as.data.frame()
  
  ggplot(bd, aes(x=TumorType_byAge, y=perCells, fill=cellAnnot))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=program_colors[names(program_colors) %in% unique(merge_cd8$cellAnnot)])+
    theme_classic()+
    ylab("Proportion of cells")+xlab("")+
    theme(axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))
  ggsave(paste0(figure_dir, "cellTypes_byTumorType.png"), width=5, height=4)
  
## CD4
  ## By glioma/nsclc
  DimPlot(merge_cd4, group.by = "TumorType", label=TRUE)+NoLegend()
  ggsave(paste0(figure_dir, "UMAP_CD4_byTumorType.png"), width=5, height=5)
  
  ## By cell type
  DimPlot(merge_cd4, group.by = "cellAnnot")+
    scale_color_manual(values=program_colors[names(program_colors) %in% unique(merge_cd4$cellAnnot)])
  ggsave(paste0(figure_dir, "UMAP_CD4_byCellType.png"), width=10, height=9)
  
  ## Dotplot of GOI
  merge_cd4$cellAnnot_byTumor<- paste0( merge_cd4$cellAnnot, "_", merge_cd4$TumorType)
  DotPlot(merge_cd4, group.by = "cellAnnot_byTumor", 
            features=c("CCR7", "FOXP3", "EOMES", "ANXA1", "CD69",
                     "CXCR1", "PRF1", "GNLY",
                     "GZMK", "ITGAE", "PDCD1", "HAVCR2", "TIGIT", "CTLA4"))+
      theme(axis.text.x = element_text(angle=45, hjust=1))
  ggsave(paste0(figure_dir, "Dotplot_CD4_GOI_byCellType.png"), width=10, height=6)
  
  ## Cell type breakdown
  merge_cd4$TumorType_byAge<- ifelse(merge_cd4$NiceSubtype %in% c("GBM", "IDH-G"), "Glioma_Adult",
                                     ifelse(merge_cd4$TumorType=="Glioma", "Glioma_Ped", "NSCLC"))
  bd<- merge_cd4@meta.data %>% group_by(TumorType_byAge, cellAnnot) %>% 
    summarise(nCells=n()) %>%
    group_by(TumorType_byAge) %>%
    mutate(perCells=nCells/sum(nCells)) %>%
    as.data.frame()
  
  ggplot(bd, aes(x=TumorType_byAge, y=perCells, fill=cellAnnot))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=program_colors[names(program_colors) %in% unique(merge_cd4$cellAnnot)])+
    theme_classic()+
    ylab("Proportion of cells")+xlab("")+
    theme(axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))
  ggsave(paste0(figure_dir, "cellTypes_byTumorType.png"), width=5, height=4)

```

## Score all cells for glioma and nsclc gene sets, correlate scores
```{r}
## CD8
  ## Remove Tregs and Inhibitory- cd4 specifc
  glioma_markers<- glioma_markers[!glioma_markers$cluster %in% c("Treg", "Inhibitory"),]
  
  ## Subset markers to top 100 (may be fewer than this) for each program
  glioma_top_markers<- glioma_markers %>% group_by(cluster) %>% top_n(n=100, wt=avg_log2FC) %>% as.data.frame()
  nsclc_top_markers<- nsclc_cd8_markers %>% group_by(Cluster) %>% top_n(n=100, wt=F) %>% as.data.frame()
  
  ## Split into list
  glioma_marker_list<- split(glioma_top_markers, f=glioma_top_markers$cluster)
  glioma_marker_list<- lapply(glioma_marker_list, function(x){x$gene})
  names(glioma_marker_list)<- paste0("Glioma_", names(glioma_marker_list))
  
  nsclc_marker_list<- split(nsclc_top_markers, f=nsclc_top_markers$Cluster)
  nsclc_marker_list<- lapply(nsclc_marker_list, function(x){x$Gene.Symbol})
  names(nsclc_marker_list)<- paste0("NSCLC_", names(nsclc_marker_list))
  all_markers<-c(glioma_marker_list, nsclc_marker_list)
  
  ## Norm/center
  cm_list<-NormCenter(merge_cd8@assays$RNA@counts)
  cm_mean<- rowMeans(log2(cm_list$raw_data+1))
  
  ## Score
  scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, all_markers)))
  
  ## Correlate and plot heatmap
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  pheatmap(scores_factor_cor, color = hm_colors(100), 
           cluster_rows = F, cluster_cols = F, 
           annotation_names_row = F, annotation_names_col =T,
           show_rownames = T, show_colnames = F,
           filename = paste0(figure_dir, "Cor_CD8programs_Glioma.vs.NSCLC.png"),
           width = 8, height = 4)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir, "HC_CD8programs_Glioma.vs.NSCLC.png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
  
## CD4
  ## Remove cd8-specific markers
  glioma_markers<- glioma_markers[!glioma_markers$cluster %in% c("Cytotoxic","Predysfunctional"),]
  
  ## Subset markers to top 100 (may be fewer than this) for each program
  glioma_top_markers<- glioma_markers %>% group_by(cluster) %>% top_n(n=100, wt=avg_log2FC) %>% as.data.frame()
  nsclc_top_markers<- nsclc_cd4_markers %>% group_by(Cluster) %>% top_n(n=100, wt=F) %>% as.data.frame()
  
  ## Split into list
  glioma_marker_list<- split(glioma_top_markers, f=glioma_top_markers$cluster)
  glioma_marker_list<- lapply(glioma_marker_list, function(x){x$gene})
  names(glioma_marker_list)<- paste0("Glioma_", names(glioma_marker_list))
  
  nsclc_marker_list<- split(nsclc_top_markers, f=nsclc_top_markers$Cluster)
  nsclc_marker_list<- lapply(nsclc_marker_list, function(x){x$Gene.Symbol})
  names(nsclc_marker_list)<- paste0("NSCLC_", names(nsclc_marker_list))
  all_markers<-c(glioma_marker_list, nsclc_marker_list)
  
  ## Norm/center
  cm_list<-NormCenter(merge_cd4@assays$RNA@counts)
  cm_mean<- rowMeans(log2(cm_list$raw_data+1))
  
  ## Score
  scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, all_markers)))
  
  ## Correlate and plot heatmap
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  pheatmap(scores_factor_cor, color = hm_colors(100), 
           cluster_rows = F, cluster_cols = F, 
           annotation_names_row = F, annotation_names_col =T,
           show_rownames = T, show_colnames = F,
           filename = paste0(figure_dir, "Cor_CD4programs_Glioma.vs.NSCLC.png"),
           width = 8, height = 4)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir, "HC_CD4programs_Glioma.vs.NSCLC.png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
  
```

