---
title: "R Notebook"
output: html_document
---

## Purpose:
Overall goal is to compare tcells in melanoma and NSCLC (known to have reasonably good response to checkpoint therapies) to tcells in gliomas. Here, merge processed melanoma/NSCLC tumor cd8 with glioma, compare GOI expression patterns, breakdown of cell types

## Approach:
Merge processed melanoma/NSCLC tumor cd8 seurat object with glioma cd8 seurat object. Plot GOI in equivalent programs, plot breakdown of cell types

```{r}
library(Seurat) 
library(harmony)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/", 
                     "02a_Tcells/02_Ped.Adult/04b_ComparisonToOtherTumor/")

analysis_dir<- paste0(working_dir, "/analysis/NSCLC.melanoma/")
figure_dir<- paste0(working_dir, "/figures/Merge_NSCLC.Glioma.melanoma/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

```{r}
umap_theme<- function(p, colors){
  p+NoLegend()+
  scale_color_manual(values=colors)+
  theme(legend.position = "bottom",
        legend.text = element_text(face="bold"),
      plot.title = element_blank(),
      axis.text = element_text(face="bold", colour = "black"),
      axis.title = element_text(face="bold", colour = "black"))+
  ylab("UMAP2")+xlab("UMAP1")
}

bargraph_bd_theme<- function(){
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black"),
        axis.text = element_text(color="black", face="bold"),
        axis.title =  element_text(color="black", face="bold"),
        axis.text.y = element_text(color="black"),
        legend.position = "bottom",
        legend.title=element_blank())
  }
```

## Read in seurat objects + marker genes
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../03_Visualize.Reannotate_programs/analysis/Reannotate_programs/", "rank",rank_use, "_sharedBy", sharedBy, "/" )

## Read in seurat objects
nsclc_cd8<- readRDS("analysis/NSCLC/cd8_seurat_nsclc.Rds")
glioma_cd8<- readRDS(paste0(data_dir, "ped.adult_seurat_cd8_harmony.Rds")) 
melanoma_cd8<- readRDS("analysis/melanoma/cd8_seurat_melanoma.Rds")

nsclc_cd4<- readRDS("analysis/NSCLC/cd4_seurat_nsclc.Rds")
glioma_cd4<- readRDS(paste0(data_dir, "ped.adult_seurat_cd4_harmony.Rds")) 
melanoma_cd4<- readRDS("analysis/melanoma/cd4_seurat_melanoma.Rds")

## May remove this- no programs for melanoma
  ## Read in marker genes for programs
  nsclc_cd8_markers<- read.csv("analysis/NSCLC/nsclc_cd8_cluster_markers_annot.csv")
  glioma_markers<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                                   "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", sharedBy, "_reAnnot.Rds"))
  
  nsclc_cd4_markers<- read.csv("analysis/NSCLC/nsclc_cd4_cluster_markers_annot.csv")
  
## Read in colors to use- add on nsclc-specific programs
  program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, "_reAnnot.Rds"))
  program_colors<- c(program_colors, CD28="pink", Naive="skyblue4", MAIT="brown", Predysfunctional.2="lightsalmon1",
                     Dysfunctional="violetred", Cytotoxic="yellowgreen")
  program_colors<-c(program_colors, CD69="pink", GZMA="brown",ANXA1="lightsalmon3", ActivatedTreg="gold")
  program_colors<- c(program_colors, M0="grey", M1="grey", M2="grey", M3="grey", M4="grey", M5="grey")
```


## Merge seurat objects
```{r}
## Optional: downsample to minimum cells in 1 object to reduce sample size effects
downSize<- TRUE

## CD8
  ## Check that scale is equivalent
  head(colSums(nsclc_cd8@assays$RNA@counts))/head(colSums(glioma_cd8@assays$RNA@counts))
  head(colSums(melanoma_cd8@assays$RNA@counts))/head(colSums(glioma_cd8@assays$RNA@counts))
  
  ## Optional: downsample
  if(downSize){
    minSize<- min(ncol(nsclc_cd8), ncol(glioma_cd8), ncol(melanoma_cd8)) -1 
    nsclc_cd8<- nsclc_cd8[, sample(colnames(nsclc_cd8), size =minSize, replace=F)]
    melanoma_cd8<- melanoma_cd8[, sample(colnames(melanoma_cd8), size =minSize, replace=F)]
    glioma_cd8<- glioma_cd8[, sample(colnames(glioma_cd8), size =minSize, replace=F)]
  }

  ## Match metadata
  nsclc_cd8$cellAnnot<- nsclc_cd8$majorCluster_annot
  glioma_cd8$cellAnnot<- glioma_cd8@meta.data[[paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")]]
  melanoma_cd8$cellAnnot<- paste0("M", melanoma_cd8$seurat_clusters) ## clusters
  nsclc_cd8$NiceSubtype<- "NSCLC"
  melanoma_cd8$NiceSubtype<- "Melanoma"
  nsclc_cd8$TumorType<- "NSCLC"
  glioma_cd8$TumorType<- "Glioma"
  melanoma_cd8$TumorType<- "Melanoma"
  
  glioma_cd8@meta.data<- glioma_cd8@meta.data[,colnames(glioma_cd8@meta.data) %in% colnames(nsclc_cd8@meta.data)]
  nsclc_cd8@meta.data<- nsclc_cd8@meta.data[,colnames(nsclc_cd8@meta.data) %in% colnames(glioma_cd8@meta.data)]
  melanoma_cd8@meta.data<- melanoma_cd8@meta.data[,colnames(melanoma_cd8@meta.data) %in% colnames(glioma_cd8@meta.data)]
  
  ## Merge seurat
  merge_cd8<- merge(nsclc_cd8, c(glioma_cd8, melanoma_cd8))
  
  ## Rerun clustering
  merge_cd8_meta<- merge_cd8@meta.data
  merge_cd8<- RunFullSeurat(merge_cd8@assays$RNA@counts, RunHarmony = TRUE, samples = merge_cd8_meta$sample)
  merge_cd8<- AddMetaData(merge_cd8, merge_cd8_meta[,!colnames(merge_cd8_meta) %in% colnames(merge_cd8@meta.data)])
  
  filename<- ifelse(downSize,"cd8_seurat_nsclc.glioma.melanoma_downsample.Rds", "cd8_seurat_nsclc.glioma.melanoma.Rds" )
  saveRDS(merge_cd8, file=paste0(analysis_dir, filename))
  
## CD4
  ## Check that scale is equivalent
  head(colSums(nsclc_cd4@assays$RNA@counts))/head(colSums(glioma_cd4@assays$RNA@counts))
  head(colSums(melanoma_cd4@assays$RNA@counts))/head(colSums(glioma_cd4@assays$RNA@counts))
  
  ## Optional: downsample
  if(downSize){
    minSize<- min(ncol(nsclc_cd4), ncol(glioma_cd4), ncol(melanoma_cd4)) -1 
    nsclc_cd4<- nsclc_cd4[, sample(colnames(nsclc_cd4), size =minSize, replace=F)]
    melanoma_cd4<- melanoma_cd4[, sample(colnames(melanoma_cd4), size =minSize, replace=F)]
    glioma_cd4<- glioma_cd4[, sample(colnames(glioma_cd4), size =minSize, replace=F)]
  }
  
  ## Match metadata
  nsclc_cd4$cellAnnot<- nsclc_cd4$majorCluster_annot
  glioma_cd4$cellAnnot<- glioma_cd4@meta.data[[paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")]]
  melanoma_cd4$cellAnnot<- paste0("M", melanoma_cd4$seurat_clusters) ## just seurat clusters
  nsclc_cd4$NiceSubtype<- "NSCLC"
  melanoma_cd4$NiceSubtype<- "Melanoma"
  nsclc_cd4$TumorType<- "NSCLC"
  glioma_cd4$TumorType<- "Glioma"
  melanoma_cd4$TumorType<- "Melanoma"
  
  glioma_cd4@meta.data<- glioma_cd4@meta.data[,colnames(glioma_cd4@meta.data) %in% colnames(nsclc_cd4@meta.data)]
  nsclc_cd4@meta.data<- nsclc_cd4@meta.data[,colnames(nsclc_cd4@meta.data) %in% colnames(glioma_cd4@meta.data)]
  melanoma_cd4@meta.data<- melanoma_cd4@meta.data[,colnames(melanoma_cd4@meta.data) %in% colnames(glioma_cd4@meta.data)]
  
  ## Merge seurat
  merge_cd4<- merge(nsclc_cd4, c(glioma_cd4, melanoma_cd4))
  
  ## Rerun clustering
  merge_cd4_meta<- merge_cd4@meta.data
  merge_cd4<- RunFullSeurat(merge_cd4@assays$RNA@counts, RunHarmony = TRUE, samples = merge_cd4_meta$sample)
  merge_cd4<- AddMetaData(merge_cd4, merge_cd4_meta[,!colnames(merge_cd4_meta) %in% colnames(merge_cd4@meta.data)])
  
  filename<- ifelse(downSize,"cd4_seurat_nsclc.glioma.melanoma_downsample.Rds", "cd4_seurat_nsclc.glioma.melanoma.Rds" )
  saveRDS(merge_cd4, file=paste0(analysis_dir, filename))
```

## For cd8: Add score for naive, cyto, predys, dys
```{r}
downSize<-TRUE
filename<- ifelse(downSize,"cd8_seurat_nsclc.glioma.melanoma_downsample.Rds", "cd8_seurat_nsclc.glioma.melanoma.Rds" )
merge_cd8<- readRDS(paste0(analysis_dir, filename))


all_markers<- list(cytotoxic= c("CX3CR1", "PRF1", "GZMA", "GZMB", "GZMH",  "FGFBP2","KLRG1", "FCGR3A"),
                   naive=c("TCF7", "CCR7", "SELL", "LEF1"),
                   predysfunctional=c("GZMK", "PDCD1", "CXCR3"),
                   dysfunctional= c("HAVCR2", "GZMB","IFNG", "CXCL13", "TIGIT", "LAYN", "LAG3", 
                                    "PDCD1", "CTLA4", "ITGAE", "ENTPD1")) 
merge_cd8<- AddModuleScore(merge_cd8, all_markers, name = names(all_markers))

## Add max program/maxscore
scores<- merge_cd8@meta.data[,c("cytotoxic1", "naive2", "predysfunctional3", "dysfunctional4")]
scores$max<- apply(scores, 1, function(x){names(x)[which.max(x)]})
scores$maxScore<- apply(scores[,colnames(scores) != "max"], 1, function(x){x[which.max(x)]})
merge_cd8$maxGeneset<- scores$max
merge_cd8$maxGeneset_score<- scores$maxScore
merge_cd8$maxGeneset_lowScore<- ifelse(merge_cd8$maxGeneset_score<0.2, "ScoreTooLow", merge_cd8$maxGeneset)

## Remove numbers
merge_cd8$maxGeneset_lowScore<- gsub("[0-9]", "", merge_cd8$maxGeneset_lowScore)

saveRDS(merge_cd8, file=paste0(analysis_dir, filename))

## RAW DATA EXPORT
df<- merge_cd8@meta.data[,c("sample", "TumorType", "cytotoxic1", "naive2", "predysfunctional3", "dysfunctional4", "maxGeneset")]
colnames(df)<- gsub("[0-9]$", "_score", colnames(df))
df<- df %>% dplyr::rename(Program=maxGeneset) %>% as.data.frame()
write.csv(df, file="../../../../notUploaded/ImmuneDraft_Rebecca/raw_results_files/glioma.melanoma.nsclc_cd8_cellwise_scores.annotations.csv")
```

## Plots- CD8
```{r}
downSize<-FALSE
filename<- ifelse(downSize,"cd8_seurat_nsclc.glioma.melanoma_downsample.Rds", "cd8_seurat_nsclc.glioma.melanoma.Rds" )
merge_cd8<- readRDS(paste0(analysis_dir, filename))

## Set figure directory- subfolder for downsampled object
figure_dir_use<- ifelse(downSize, paste0(figure_dir, "Downsample/"), figure_dir)
if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}

## Set colors
program_colors<- c(cytotoxic="greenyellow", predysfunctional="lightsalmon2", naive="skyblue4", dysfunctional="red",ScoreTooLow="grey")
tumor_colors<- c(Glioma="purple", Melanoma="orange", NSCLC="grey22") ## colors updated 12/12/22

## By max geneset
p<-DimPlot(merge_cd8, group.by = "maxGeneset_lowScore")
umap_theme(p, colors=program_colors)
ggsave(paste0(figure_dir_use, "CD8_UMAP_maxNature2020Score.png"), width=6, height=6)

## Proportion by tumor type
bd<- as.data.frame(table(merge_cd8$maxGeneset_lowScore, merge_cd8$TumorType))
colnames(bd)<- c("Geneset", "TumorType", "NCells")
bd<- bd %>% group_by(TumorType) %>% mutate(perCells=NCells/sum(NCells)) %>% as.data.frame()
ggplot(bd, aes(x=TumorType,y=perCells, fill=Geneset))+
  geom_bar(stat="identity")+
  theme_classic()+
  scale_fill_manual(values=program_colors)+
  bargraph_bd_theme()+
  ylab("Proportion of cells")+ xlab("")+
  guides(fill=guide_legend(nrow=2,byrow=TRUE))
ggsave(paste0(figure_dir_use, "CD8_ProportionNature2020_byTumorType.png"), width=4, height=5)

## Proportion without ScoreTooLow
bd<- bd[bd$Geneset!="ScoreTooLow",]
bd<- bd %>% group_by(TumorType) %>% mutate(perCells=NCells/sum(NCells)) %>% as.data.frame()
ggplot(bd, aes(x=TumorType,y=perCells, fill=Geneset))+
  geom_bar(stat="identity")+
  theme_classic()+
  scale_fill_manual(values=program_colors[names(program_colors) %in% bd$Geneset])+
  bargraph_bd_theme()+
  ylab("Proportion of cells")+ xlab("")+
  guides(fill=guide_legend(nrow=2,byrow=TRUE))
ggsave(paste0(figure_dir_use, "CD8_ProportionNature2020_byTumorType_noScoreTooLow.pdf"), device="pdf", width=4, height=5)


## Expression of GOI by tumor type + max geneset
merge_cd8$maxGeneset_byTumor<- paste0(merge_cd8$maxGeneset_lowScore, "_", merge_cd8$TumorType)
DotPlot(merge_cd8, group.by = "maxGeneset_byTumor", 
        features=c("CD8A","CD3G", 
                   "PDCD1","CTLA4", "HAVCR2", "ITGAE",
                   "PRF1", "GNLY", "GZMA", 
                   "SELPLG", "KLRB1", "PTGER2"))+
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave(paste0(figure_dir_use, "CD8_Dotplot_GOI_byNature2020MaxProgram.png"), width=8, height = 6)

## By glioma/nsclc/melanoma
p<-DimPlot(merge_cd8, group.by = "TumorType", label=FALSE)
umap_theme(p, tumor_colors)
ggsave(paste0(figure_dir_use, "CD8_UMAP_byTumorType.pdf"), device="pdf", width=6, height=6)

## By cell type
DimPlot(merge_cd8, group.by = "cellAnnot")
ggsave(paste0(figure_dir_use, "CD8_UMAP_byProgram.png"),  width=11, height=9)


## Dotplot of inhibitory GOI by tumor type
DotPlot(merge_cd8, group.by = "TumorType", 
          features=c("PDCD1", "HAVCR2", "TIGIT", "CTLA4", "GZMK", "ITGAE", "PRF1", "GNLY", "KLRB1"), cols=c("grey","red"))+
    theme(axis.text.x = element_text(angle=45, hjust=1))+
  ylab("")+xlab("")+
  scale_color_gradient2(low="blue", mid="white", high="red")+
  coord_flip()
ggsave(paste0(figure_dir_use, "CD8_Dotplot_InhibitoryGOI_byTumorType_colorCentered.pdf"),device="pdf", width=4, height=5)

## Dotplot of  GOI by tumor type
DotPlot(merge_cd8, group.by = "TumorType", 
          features=c("CD8A", "SELPLG", "PTGER2", "KLRB1"), scale=FALSE, cols=c("grey", "red"))+
    theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave(paste0(figure_dir_use, "CD8_Dotplot_GOI_byTumorType.png"), width=3, height=4)

## DEGs between cyto subsets
cyto_seurat<- subset(merge_cd8, maxGeneset_lowScore=="cytotoxic")
cyto_seurat<- SetIdent(cyto_seurat, value=cyto_seurat$TumorType)

all_markers<- FindAllMarkers(cyto_seurat, min.pct = 0.2)
all_markers_filtered<- all_markers[all_markers$p_val_adj<0.05 & all_markers$avg_log2FC>0,]

predys_seurat<- subset(merge_cd8, maxGeneset_lowScore=="predysfunctional")
predys_seurat<- SetIdent(predys_seurat, value=predys_seurat$TumorType)

all_markers_predys<- FindAllMarkers(predys_seurat, min.pct = 0.2)
all_markers_filtered_predys<- all_markers_predys[all_markers_predys$p_val_adj<0.05 & all_markers_predys$avg_log2FC>0,]
```

## Plots- CD4
```{r}
downSize<-TRUE
filename<- ifelse(downSize,"cd4_seurat_nsclc.glioma.melanoma_downsample.Rds", "cd4_seurat_nsclc.glioma.melanoma.Rds" )
merge_cd4<- readRDS(paste0(analysis_dir, filename))

## Set figure directory- subfolder for downsampled object
figure_dir_use<- ifelse(downSize, paste0(figure_dir, "Downsample/"), figure_dir)
if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}

## Set colors
program_colors<- c(cytotoxic="greenyellow", predysfunctional="lightsalmon2", naive="skyblue4", dysfunctional="red",ScoreTooLow="grey")
tumor_colors<- c(Glioma="olivedrab", Melanoma="navy", NSCLC="violetred")

## By glioma/nsclc/melanoma
p<-DimPlot(merge_cd4, group.by = "TumorType", label=FALSE)
umap_theme(p, tumor_colors)
ggsave(paste0(figure_dir_use, "CD4_UMAP_byTumorType.png"), width=6, height=6)


## By cell type
DimPlot(merge_cd4, group.by = "cellAnnot")
ggsave(paste0(figure_dir_use, "CD4_UMAP_byCellType.png"), width=10, height=9)

## Dotplot of inhibitory GOI by tumor type
DotPlot(merge_cd4, group.by = "TumorType", 
          features=c("PDCD1", "HAVCR2", "TIGIT", "CTLA4", "GZMK", "ITGAE", "PRF1", "GNLY", "KLRB1"))+
    theme(axis.text.x = element_text(angle=45, hjust=1))+
  ylab("")+xlab("")
ggsave(paste0(figure_dir_use, "CD4_Dotplot_InhibitoryGOI_byTumorType.png"), width=6, height=4)

## Dotplot of  GOI by tumor type
DotPlot(merge_cd4, group.by = "TumorType", 
          features=c("CD4", "SELPLG", "PTGER2", "KLRB1"), scale=FALSE)+
    theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave(paste0(figure_dir_use, "CD4_Dotplot_GOI_byTumorType.png"), width=5, height=4)
  
```
## Plots- CD4 and CD8
```{r}
downSize<-FALSE
filename<- ifelse(downSize,"cd4_seurat_nsclc.glioma.melanoma_downsample.Rds", "cd4_seurat_nsclc.glioma.melanoma.Rds" )
merge_cd4<- readRDS(paste0(analysis_dir, filename))
filename<- ifelse(downSize,"cd8_seurat_nsclc.glioma.melanoma_downsample.Rds", "cd8_seurat_nsclc.glioma.melanoma.Rds" )
merge_cd8<- readRDS(paste0(analysis_dir, filename))

## merge together
merge_cd4$detailed_annot<- "CD4"; merge_cd8$detailed_annot<- "CD8"
merge_tcell<- merge(merge_cd4, merge_cd8)

## Add on age for glioma
merge_tcell$age_numeric<- as.numeric(merge_tcell$age_numeric)
merge_tcell$TumorType_age<- ifelse(merge_tcell$TumorType=="Glioma",
                               ifelse(merge_tcell$age_numeric<25, "Glioma_Ped", "Glioma_Adult"),
                               merge_tcell$TumorType)

## Set figure directory- subfolder for downsampled object
figure_dir_use<- ifelse(downSize, paste0(figure_dir, "Downsample/"), figure_dir)
if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}

## Set colors
program_colors<- c(cytotoxic="greenyellow", predysfunctional="lightsalmon2", naive="skyblue4", dysfunctional="red",ScoreTooLow="grey")
tumor_colors<- c(Glioma="olivedrab", Melanoma="navy", NSCLC="violetred")


## Dotplot of inhibitory GOI by tumor type
DotPlot(merge_tcell, group.by = "TumorType_age", 
          features=c( "PDCD1", "HAVCR2", "TIGIT", "CTLA4", "GZMK", "ITGAE", "PRF1", "GNLY","NKG7", "KLRB1"))+
    theme(axis.text.x = element_text(angle=45, hjust=1))+
  ylab("")+xlab("")
ggsave(paste0(figure_dir_use, "Tcell_Dotplot_InhibitoryGOI_byTumorType.png"), width=7, height=3)

merge_tcell<-ScaleData(merge_tcell, do.scale = FALSE, do.center=TRUE, features=rownames(merge_tcell))
DoHeatmap(merge_tcell, group.by = "TumorType_age", slot="scale",
          features=c( "PDCD1", "HAVCR2", "TIGIT", "CTLA4", "GZMK", "ITGAE", "PRF1", "GNLY","NKG7", "KLRB1"))+
    theme(axis.text = element_text(face="bold", color="black"))+
    scale_fill_gradient2(low="blue", mid="white", high="red")
ggsave(paste0(figure_dir_use, "Tcell_Heatmap_InhibitoryGOI_byTumorType.png"), width=6, height=4)

## Dotplot of  GOI by tumor type
DotPlot(merge_tcell, group.by = "TumorType", 
          features=c("CD4","CD8", "SELPLG", "PTGER2", "KLRB1"), scale=FALSE)+
    theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave(paste0(figure_dir_use, "Tcell_Dotplot_GOI_byTumorType.png"), width=5, height=4)
  
```

## Generate new marker genes based on Nature2020 assignment (for CD8 only)
```{r}
downSize<-TRUE
filename<- ifelse(downSize,"cd8_seurat_nsclc.glioma.melanoma_downsample.Rds", "cd8_seurat_nsclc.glioma.melanoma.Rds" )
merge_cd8<- readRDS(paste0(analysis_dir, filename))

## Set figure directory- subfolder for downsampled object
figure_dir_use<- ifelse(downSize, paste0(figure_dir, "Downsample/"), figure_dir)
if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}


## Split by tumor type
cd8_list<- SplitObject(merge_cd8, split.by = "TumorType")

## For each tumor type, remove "ScoreTooLow" and find markers for remaining 4 Nature 2020 genesets
marker_list<- lapply(cd8_list, function(x){
  print(unique(x$TumorType))
  tmp<- subset(x, maxGeneset_lowScore != "ScoreTooLow")
  tmp<- SetIdent(tmp, value=tmp$maxGeneset_lowScore)
  markers<- FindAllMarkers(tmp, min.pct = 0.2)
  return(markers)
})

## Filter
marker_list_filtered<- lapply(marker_list, function(x){
  x[x$p_val_adj<0.05 & x$avg_log2FC>0.5,]
})

## Save
filename<- ifelse(downSize, "markers_forNature2020Annot_byTumorType_downSample.Rds","markers_forNature2020Annot_byTumorType.Rds")
saveRDS(marker_list_filtered, file=paste0(analysis_dir, filename))

## RAW DATA EXPORT
  ## Subset to top 50
  marker_list<- lapply(marker_list_filtered, function(x){
    tmp<- x[order(x$cluster, -x$avg_log2FC),]
    tmp<- tmp %>% group_by(cluster) %>% top_n(n=50, wt=avg_log2FC) %>% as.data.frame()
    return(tmp)
  })
  
  ## Add on tumor type
  marker_list<- lapply(names(marker_list), function(tumortype){tmp<- marker_list[[tumortype]]; tmp$tumortype<- tumortype;return(tmp)})
  
  ## collapse to df
  marker_df<- do.call("rbind", marker_list)
  write.csv(marker_df, file="../../../../notUploaded/ImmuneDraft_Rebecca/raw_results_files/program_markers_cd8_glioma.melanoma.nsclc.csv")
```


## Score all cells for glioma, melanoma, and nsclc gene sets, correlate scores (for CD8 only)
```{r}
## Read in markers generated above
downSize<-TRUE
filename<- ifelse(downSize, "markers_forNature2020Annot_byTumorType_downSample.Rds","markers_forNature2020Annot_byTumorType.Rds")
marker_list<- readRDS(paste0(analysis_dir, filename))

## Subset to top 50
marker_list<- lapply(marker_list, function(x){
  tmp<- x[order(x$cluster, -x$avg_log2FC),]
  tmp<- tmp %>% group_by(cluster) %>% top_n(n=50, wt=avg_log2FC) %>% as.data.frame()
  return(tmp)
})

## Add on tumor type to cluster name
marker_list<- lapply(names(marker_list), function(x){
  tmp<- marker_list[[x]]
  tmp$cluster<- paste0(x,"_", gsub("[0-9]", "" ,tmp$cluster))
  return(tmp)
})

## Reformat into single list- 1 entry per geneset/tumor type
marker_list_gene<- lapply(marker_list, function(x){
  tmp<-split(x, f=x$cluster)
  tmp2<- lapply(tmp, function(x){x$gene})
  names(tmp2)<- names(tmp)
  return(tmp2)
  })
marker_list_gene<- do.call(c, marker_list_gene)


## Norm/center
cm_list<-NormCenter(merge_cd8@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data+1))

## Score
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, marker_list_gene)))

## Correlate and plot heatmap
scores_factor_hc = clusterNmfFactors(scores)
scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]

## Heatmap of correlations
hm_colors = rev((brewer.pal(n=9, name="RdBu")))
hm_colors = colorRampPalette(colors = hm_colors)
pheatmap(scores_factor_cor, color = hm_colors(100), 
         cluster_rows = F, cluster_cols = F, 
         annotation_names_row = F, annotation_names_col =T,
         show_rownames = T, show_colnames = F,
         filename = paste0(figure_dir_use, "Cor_CD8programs_Glioma.vs.NSCLC.vs.melanoma.pdf"),
         width = 8, height = 4)

## Dendrogram of hierarchical clustering 
jpeg(filename=paste0(figure_dir_use, "HC_CD8programs_Glioma.vs.NSCLC.vs.Melanoma.pdf"), width = 400, height = 600)
plot(scores_factor_hc$hc_obj)
dev.off()

## Mean pearson correlation for each geneset
lapply(c("cytotoxic", "dysfunctional", "predysfunctional", "naive"), function(x){
  print(x)
  tmp<- scores_factor_cor[grepl(paste0("_", x), colnames(scores_factor_cor)), 
                          grepl(paste0("_", x), rownames(scores_factor_cor))]
  print(mean(as.numeric(tmp)[as.numeric(tmp) != 1]))
  
})

write.csv(scores, file=paste0(analysis_dir, "scores_CD8programs_Glioma.vs.NSCLC.vs.melanoma.csv"))
```


## GO term overrepresentation
```{r}
## Read in markers generated above
downSize<-FALSE
filename<- ifelse(downSize, "markers_forNature2020Annot_byTumorType_downSample.Rds","markers_forNature2020Annot_byTumorType.Rds")
marker_list<- readRDS(paste0(analysis_dir, filename))

## Rename by tumor type + flatten to single df
marker_list<- lapply(names(marker_list), function(x){
  tmp<- marker_list[[x]]
  tmp$cluster<- paste0(x, "_", gsub("[0-9]", "", tmp$cluster))
  return(tmp)
})
marker_genes<- do.call("rbind", marker_list)

OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
ensembl = useMart("ensembl",dataset="hsapiens_gene_ensembl")
allOE_genes<-getBM(attributes='ensembl_gene_id', 
                   filters = 'hgnc_symbol', 
                   values = unique(c(rownames(glioma_cd8@assays$RNA@counts),
                                     rownames(melanoma_cd8@assays$RNA@counts),
                                     rownames(nsclc_cd8@assays$RNA@counts))), 
                  mart = ensembl)

## Subset to top markers
marker_genes<-marker_genes %>% group_by(cluster) %>% top_n(n=50, wt=avg_log2FC) %>% as.data.frame()

## Run GO: Convert marker genes to ensembl IDs too, then run OR analysis (GO)
go_result = list()
for (i in unique(marker_genes$cluster)){
  print(i)
  genes<- marker_genes[marker_genes$cluster==i, "gene"]
  if(length(genes)!=1){
    sigOE_genes = getBM(attributes='ensembl_gene_id', 
                        filters = 'hgnc_symbol', 
                        values=genes,
                        mart = ensembl)$ensembl_gene_id
    go_result[[i]] = go_analysis(sigOE_genes, allOE_genes)
    }
} 

## Plot and save
go_plots<- plotGO(go_result = go_result, n_terms = 15)
cowplot::plot_grid(plotlist=go_plots, ncol=2)
ggsave(paste0(figure_dir , "GO_NMF.png"), width=18, height=35)

saveRDS(go_result, file=paste0(analysis_dir, "go_result.Rds"))
```


## sccoda
```{r}
data_dir<- "../../../scCoda/data/"

downSize<-FALSE
filename<- ifelse(downSize,"cd8_seurat_nsclc.glioma.melanoma_downsample.Rds", "cd8_seurat_nsclc.glioma.melanoma.Rds" )
merge_cd8<- readRDS(paste0(analysis_dir, filename))

## Breakdown of number of cells per program by sample
bd<- as.data.frame.matrix(table(merge_cd8$sample, merge_cd8$maxGeneset_lowScore))
bd<- bd[,colnames(bd) != "ScoreTooLow"]
bd$Sample<- rownames(bd)

## Add on age
bd$Subtype<- as.character(plyr::mapvalues(bd$Sample,merge_cd8$sample, merge_cd8$TumorType,warn_missing = FALSE))

## Output
write.csv(bd, file=paste0(data_dir, "pedOnly_nomuv63/Glioma.NSCLC.Melanoma_CD8_byNature2020Signature.csv"), row.names=FALSE)
```

