---
title: "_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/BySubtype/Midline/Midline_HistoneWT/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

seurat_analysis_folder = "analysis/seurat/MUV090Only/"
seurat_fig_folder = "figures/seurat/MUV090Only/"
for (folder in c(seurat_analysis_folder, seurat_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

subtypes<-read.csv("../../Subtypes.csv")
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}
```

## Load counts and metadata for initiating seurat object
```{r load-data}
input_dir<- "../../../Preprocessing/SplitBySubtype/analysis/"
load(paste0(input_dir, "midlineWT_cm_list.Robj"))
samples = readRDS(paste0(input_dir,"midlineWT_samples.rds"))
tumor_stat = readRDS(paste0(input_dir,"midlineWTt_cluster6CellsNormal_call_tumor.rds"))
cm_list<-midlineWT_cm_list


samples<- samples[grep("MUV090", samples)]
cm_list<- lapply(cm_list, function(x){x[,grep("MUV090", colnames(x))]})
tumor_stat<- tumor_stat[grep("MUV090", names(tumor_stat))]

sum(colnames(cm_list$raw_data) == names(tumor_stat)); length(tumor_stat)
sum(names(samples)==names(tumor_stat))
```

## Subset cm to remove noise/normal samples
```{r subset-tumor-cells}
cm = cm_list$raw_data; rm(cm_list); gc()
cm = cm[,tumor_stat=="Malignant"]
samples = samples[tumor_stat=="Malignant"]

any(colnames(cm) != names(samples))
```

## Initiate seurat obj and add metadata 
```{r seurat_obj_meta_data}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", min.cells=0, min.genes=0, scale.factor=1E5, do.scale=F, do.center=T)


## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples

sample_df<- as.data.frame(samples); sample_df$cell<- rownames(sample_df)
subtype_df<- merge(sample_df, subtypes, by="samples")

seurat_obj@meta.data$subtype<- subtype_df$subtype
seurat_obj@meta.data$AdditionalSubtype<- subtype_df$Diagnosis
```

## OPTIONAL: plot highly variable genes
```{r}
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## PCA and PC selection
```{r PCA_selection}
## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)


```



## Clustering and UMAP
```{r clustering_tSNE}
## Run UMAP and clustering using reduced dimensions from Harmony
## You may play around with resolution parameter to tune the clustering result
resolution = .8
reduction_type = "pca"
dims = 20

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = reduction_type, dims = 1:dims)


seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = reduction_type,
                            dims = 1:dims,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = resolution)


r0.8<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5) + ggtitle("Resolution_0.8")
r0.5<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5)+ ggtitle("Resolution_0.5")
r1.0<-DimPlot(seurat_obj, reduction = "umap", pt.size = 2.5)+ ggtitle("Resolution_1")

r0.5+r0.8+r1.0 #resolution of 0.8 may be best
ggsave("UMAP_clustering_VaryResolution.jpg", path=seurat_fig_folder, width=15, height=6)

## Clustering results 
table(Idents(seurat_obj))


## Save the seurat object
save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```

## UMAP plot colored by clustering and samples
## Can color tSNE with other metadata features (e.g. primary/recurrence, treatment)
```{r UMAP_plots}
## By clustering
DimPlot(object = seurat_obj, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering.jpg", path=seurat_fig_folder, width=4, height=4)


```

## DEGs for each cluster 
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.25)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05 & all_markers$avg_logFC > 0.5, ]


## Subset the top 30 DEGs for each cluster
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC) %>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(seurat_analysis_folder, "all_markers.Robj"))
write.csv(top_marker_genes, file=paste0(seurat_analysis_folder, "Top30Markers_Theta1.csv"))

## Heatmap for initial inspection
DoHeatmap(seurat_obj, 
          features=all_markers_filtered$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs_theta1.jpg", path=seurat_fig_folder)
```

## Quick annotation by OPC/OC/AC marker genes 
```{r}
## OPC-like
FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OPC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## OC-like cells
FeaturePlot(seurat_obj,
            c("BCAS1", "SIRT2", "RGR", 
              "GPR17", "NFASC", "MBP",
              "TF", "MOG", "PLP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## AC-like cells 
FeaturePlot(seurat_obj,
            c("APOE", "AQP4", "AGT", 
              "SPARCL1","ATP1A2", "GFAP",
              "S100B", "ALDOC", "CLU"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("AC_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## Neuronal-like cells
FeaturePlot(seurat_obj,
            c("STMN2", "STMN1", "STMN4", 
              "TUBB3", "INA", "ELAVL4",
              "SNAP25", "GABRG2", "ST18"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("Neuron_markers.jpg", path=seurat_fig_folder, width=9, height=9)

##CLEC2D
FeaturePlot(seurat_obj,
            c("CLEC2D"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("CLEC2D_KLRB1_markers.jpg", path=seurat_fig_folder, width=9, height=4.5)
```

## ORA test 
```{r over-representation-test}
library(clusterProfiler)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
allOE_genes <- as.character(gene_symbol_to_ensembl_id(rownames(seurat_obj@assays$RNA@counts),dataset)$ensembl_gene_id)


## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
go_result = list()
for (i in unique(all_markers_filtered$cluster)){
  sigOE_genes = as.character(gene_symbol_to_ensembl_id(subset(all_markers_filtered, cluster==i,
                                                              gene)[,], dataset)$ensembl_gene_id)
  go_result[[paste0("c", i)]] = go_analysis(sigOE_genes, allOE_genes)
} 

## Optional: remove biosynthetic/metabolic pathwasy
RemoveMetabolicBiosynthetic<- lapply(go_result, function(x){
  BS_MB<- x$ego[grep("metabolic", x$ego@result$Description, invert=TRUE),asis=TRUE]
  BS_MB<- BS_MB[grep("biosynthetic", BS_MB$Description, invert=TRUE),asis=TRUE]
})



## Plot top pathways for all clusters, display on one plot
clusters = paste0("c", sort(levels(seurat_obj@active.ident)))
allDotPlots_20_NoBCMB<-list()
for (c in 1:length(levels(seurat_obj@active.ident))){
  cluster=clusters[c]
  print(cluster)
  allDotPlots_20_NoBCMB[[c]]<-dotplot(RemoveMetabolicBiosynthetic[[cluster]], 
                                      showCategory=20, font.size = 12, title = cluster)
}

cowplot::plot_grid(plotlist = allDotPlots_20_NoBCMB, ncol=2)
ggsave("GO_OVA_AllClusters_top20_NoBiosyntheticMetabolicPathways.jpg", path=seurat_fig_folder, width=16, height=20)
save(go_result, file=paste0(seurat_analysis_folder, "go_result_metagene_DEGs_Clusters1_0Separate_LowerThreshold.Robj"))
```

# Score cells based on previous midline histone WT samples

```{r}
## read in seurat object of previous programs
backup_seurat<-seurat_obj
load("analysis/seurat_obj_MidlineHistoneWT_FromLi_V2.Robj")
previousWT_seurat<- seurat_obj
seurat_obj<- backup_seurat

## Convert to v3
previousWT_seuratv3<- UpdateSeuratObject(previousWT_seurat)
save(previousWT_seuratv3, file="analysis/seurat_obj_MidlineHistoneWT_FromLi_V3.Robj")
load("analysis/seurat_obj_MidlineHistoneWT_FromLi_V3.Robj")

## Subset to remove U1/U2/U3- not really informative
previousWT_seuratv3<- subset(previousWT_seuratv3, signature_1 %in% c("AC-MES-like",
                                                                     "Cycling",
                                                                     "OPC-like",
                                                                     "AC-like"))

## set ident to signature 1, then find top 50 markers
previousWT_sig1<-SetIdent(previousWT_seuratv3, value="signature_1")
allSig1Markers<-FindAllMarkers(previousWT_sig1, min.pct = .50, 
                               logfc.threshold = .25, only.pos = TRUE)
allSig1Markers_filtered<-allSig1Markers[allSig1Markers$p_val_adj<0.05 &
                                          allSig1Markers$avg_logFC > .75,]
top50Markers<- allSig1Markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(50, avg_logFC) %>%
                   as.data.frame()


## split into gene list
gene_list<- split(top50Markers, top50Markers$cluster)
gene_list<- lapply(gene_list, function(x){x$gene})

## Norm and center my data
cm_center<- NormCenter(seurat_obj@assays$RNA@counts)$center_data
cm_mean<- log2(rowMeans(seurat_obj@assays$RNA@counts)+1)

## score my data for each signature
scores<- scoreNmfGenes(cm_center, cm_mean, gene_list)
scores = t(scores)

topScore<- apply(scores[,1:length(gene_list)], 1, 
                  function(x) sort(x, decreasing = T)[1])
topSig<-apply(scores[,1:length(gene_list)], 1, 
                  function(x) names(sort(x, decreasing = T))[1])

## Add to seurat metadata, plot
seurat_obj@meta.data$TopSigPrevious<- topSig
DimPlot(seurat_obj, group.by="TopSigPrevious", pt.size = 2) + NoAxes()
ggsave(paste0(seurat_fig_folder, "ScoredBasedOnPreviousHistoneWTMidline_NoU.png"))
save(seurat_obj, file="analysis/seurat/seurat_obj.Robj")

## Compare this annotation to annotation also using BT1478 (other midline WT sample)
muv90OnlyMeta<- seurat_obj@meta.data
load("analysis/seurat/seurat_obj.Robj")
BothMeta<- seurat_obj@meta.data
BothMeta<- BothMeta[BothMeta$sample=="MUV090",]
sum(rownames(muv90OnlyMeta)==rownames(BothMeta)); nrow(muv90OnlyMeta)
muv90OnlyMeta$BothAnnot<- BothMeta$TopSigPrevious
```






## Make pseudobulk
```{r}
metagene_program_names = sort(unique(nmf_score$signature_1))
pseudobulk = NULL
aggr_method = "mean"
for (metagene in metagene_program_names){
  cm = seurat_obj@assays$RNA@counts[,nmf_score$signature_1 == metagene & nmf_score$score_1 >= 1]
  if (aggr_method == "mean"){
    pseudobulk = cbind(pseudobulk, rowMeans(cm))
  }
  if (aggr_method == "median"){
    pseudobulk = cbind(pseudobulk, rowMedians(cm))
  }
}
colnames(pseudobulk) = metagene_program_names
rownames(pseudobulk) = rownames(seurat_obj@assays$RNA@counts)
saveRDS(pseudobulk, file=paste0(seurat_analysis_folder, 
                                "pseudobulk_cm_unique_only_", aggr_method, NameOfMetaprogramList,".RDS"))
```

## Metagene program specific genes as overall DEGs
```{r}
seurat_obj = SetIdent(seurat_obj, value = seurat_obj@meta.data$signaure_1_without_cc)

all_markers_nmf = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_nmf = all_markers[all_markers$p_val_adj < 0.05, ]
saveRDS(all_markers, paste0(seurat_analysis_folder, "deg.rds"))
##sapply(names(nmf_marker_genes_final), function(x) table(nmf_marker_genes_final[[x]] %in% gene_signatures[[x]]))
signature_genes = unique(all_markers$gene)
saveRDS(signature_genes, paste0(seurat_analysis_folder, "signature_genes.rds"))
```

## Obsolete
## Pairwise metagene program specific genes 
```{r}
# logfc.threshold = 0.5
# test.use = "wilcox"
# min.pct = 0.3
# min.diff.pct = 0
# only.pos = TRUE 
# 
# gene_signatures = list()
# 
# print(Sys.time())
# for (target in unique(seurat_obj@meta.data$signature_1_without_cc)){
#   gene_signatures_tmp = rownames(seurat_obj@raw.data)
#   for (metagene in unique(seurat_obj@meta.data$signature_1_without_cc)){
#     tmp = NULL
#     if (metagene != target){
#       tmp = FindMarkers(seurat_obj, ident.1 = target, ident.2 = metagene, only.pos = only.pos,
#                         logfc.threshold = logfc.threshold, test.use = test.use,
#                         min.pct = min.pct, min.diff.pct = min.diff.pct)
#       tmp = tmp[tmp$p_val_adj < 0.05, ]
#       gene_signatures_tmp = intersect(gene_signatures_tmp, rownames(tmp))
#     }
#   }
#   gene_signatures[[target]] = gene_signatures_tmp
# }
# print(Sys.time())
```

