---
title: "TCR"
author: "Jenna LaBelle"
date: "10/07/22"
output: html_document
---

## Plot TCR results, clonotype expansion. Add to seurat object

```{r}
library(Seurat) 

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/PD1_Treat/01b_tenX/")

analysis_dir<- paste0(working_dir, "/analysis/TCR/")
figure_dir<- paste0(working_dir, "/figures/TCR/")
full_cohort_dir<- "../../ImmuneCellAnalysis/"

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Helper functions
```{r}
## Helper function to plot v/j gene usage
my_plotVJgeneUsage<- function(seurat, plot_v_or_j, n_genes_plot){
  ## Get df with all cells with V/J gene
  tmp<- seurat@meta.data
  tmp<- tmp[rownames(tmp)!=tmp[plot_v_or_j],]
  tmp<- as.data.frame(table(tmp[plot_v_or_j]))
  
  ## Freq of V/J gene
  tmp<- tmp[order(tmp$Freq, decreasing = TRUE),]
  
  ## Subset to top n V/J genes, order
  tmp<- tmp[1:n_genes_plot,]
  tmp$Var1<- factor(tmp$Var1, levels=tmp$Var1)
  
  ## Plot
  ggplot(tmp, aes(x=Var1, y=Freq))+
    geom_bar(stat="identity")+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"))+
    xlab(plot_v_or_j)+
    ylab("Occurence of gene")
}

## Helper function to get # of clonotypes per program, normalized by total number of cells per program
my_getNormNClonotype<- function(seurat, program_column="Program_projected", clonotype_id_column="raw_clonotype_id"){
  tmp<- seurat@meta.data
  tmp<- tmp[rownames(tmp)!=tmp[clonotype_id_column],] ## getting rid of cells without tcr info
  tmp<-as.data.frame(table(tmp[,clonotype_id_column], tmp[,program_column]))
  colnames(tmp)<- c("Clonotype", "Program", "Ncells")
  
  ## Number of unique clonotypes present in each program
  n_clonotypes_perProgram<- sapply(unique(as.character(tmp$Program)), function(program){
    program_tmp<- tmp[tmp$Program==program & tmp$Ncells>0,]
    return(length(unique(program_tmp$Clonotype)))
  })
  names(n_clonotypes_perProgram)<- unique(as.character(tmp$Program))
  
  
  ## Normalize number of clonotypes by number of cells
  n_CC_perProgram<- data.frame(Program=names(n_clonotypes_perProgram),
                               n_clonotypes=n_clonotypes_perProgram,
                               n_cells= as.numeric(table(seurat$Program_projected)))
  n_CC_perProgram$n_clonotypes_norm<- n_CC_perProgram$n_clonotypes/n_CC_perProgram$n_cells
  return(n_CC_perProgram)
}

## Helper function to get largest clonotype per program
my_getMaxClonotype<- function(seurat, program_column="Program_projected", clonotype_id_column="raw_clonotype_id"){
  tmp<- seurat@meta.data
  tmp<- tmp[rownames(tmp)!=tmp[clonotype_id_column],] ## getting rid of cells without tcr info
  tmp<-as.data.frame(table(tmp[,clonotype_id_column], tmp[,program_column]))
  colnames(tmp)<- c("Clonotype", "Program", "Ncells")

  
  ## get largest clonotype for each program
  max_clonotype<- tmp %>% group_by(Program) %>% summarise(maxClonotypeSize=max(Ncells)) %>% as.data.frame()
  return(max_clonotype)
}
```

## Load data: tcell seurat objects, TCR results, colors
NOTE: vdj to be run on BT2080, will incorporate when finished (+ hopefully after rerunning 10x seq on bt2080)
```{r}
## Seurat
seurat_obj<- readRDS("analysis/preprocessing/immune_seurat.Rds")

## TCR
samples<- c("BT2062", "BT2080_new") 
tcr_list<- lapply(samples, function(x){
  tcr<-read.csv(paste0("data/",x,"/multi/per_sample_outs/", x, "/vdj_t/filtered_contig_annotations.csv"))
  tcr$sample<- x
  tcr$barcode<- paste0(x, "_", gsub("-","_",  tcr$barcode))
  tcr$raw_clonotype_id<- paste0(tcr$sample, "_", tcr$raw_clonotype_id)
  return(tcr)
})
tcr<- do.call("rbind", tcr_list)
tcr$sample<- gsub("_new", "", tcr$sample)


## Colors 
tcell_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Tcell.programs_rank8sharedByunion_reAnnot.Rds"))
myeloid_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Myeloid_program_colors.Rds"))
myeloid_colors<- myeloid_colors$rank6
```


## Add VDJ info to seurat objects
```{r}
## Add clonotype id/vj gene
seurat_obj$raw_clonotype_id<-ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                                    as.character(plyr::mapvalues(colnames(seurat_obj), 
                                                                 tcr$barcode, tcr$raw_clonotype_id, warn_missing = FALSE)),"NA")
seurat_obj$v_gene<- ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                           as.character(plyr::mapvalues(colnames(seurat_obj), 
                                                        tcr$barcode, tcr$v_gene, warn_missing = FALSE)), "NA")
seurat_obj$j_gene<- ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                           as.character(plyr::mapvalues(colnames(seurat_obj), 
                                                        tcr$barcode, tcr$j_gene, warn_missing = FALSE)),"NA")



## Add clonotype size + whether clonally expanded
c_size<- as.data.frame(table(tcr$raw_clonotype_id))

seurat_obj$clonotype_size<- ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                                   as.numeric(as.character(plyr::mapvalues(seurat_obj$raw_clonotype_id, 
                                                                           c_size$Var1, c_size$Freq, warn_missing = FALSE))), "NA")
seurat_obj$clonal<- ifelse(seurat_obj$clonotype_size != "NA", 
                           ifelse(as.numeric(seurat_obj$clonotype_size>1), "clonal",
                                             "nonclonal"),  "NA")
seurat_obj$clonotype_size_max20<- ifelse(seurat_obj$clonotype_size>20 & seurat_obj$clonotype_size != "NA", 20, seurat_obj$clonotype_size)

saveRDS(seurat_obj, file="analysis/preprocessing/immune_seurat.Rds")
```

## Plot V/J gene usage
```{r}
seurat_obj<- readRDS("analysis/preprocessing/immune_seurat.Rds")


my_plotVJgeneUsage(subset(seurat_obj,clonotype_size != "NA"), "v_gene", 20) + ggtitle("top V gene usage")
ggsave(paste0(figure_dir, "TopV_geneUsage.png"), width=10, height=4)

my_plotVJgeneUsage(subset(seurat_obj,clonotype_size != "NA"), "j_gene", 20)+ ggtitle("top J gene usage")
ggsave(paste0(figure_dir, "TopJ_geneUsage.png"), width=10, height=4)

```


THIS NOT RUN
Kept for reference for later- will do these plots after program projection
## Plot number of clonotypes per program
```{r}
## Get number of clonotypes per program, normalized by total number of cells per program
norm_clonotypes_perProgram<- lapply(tenx_seurat_list, function(seurat){
  my_getNormNClonotype(seurat)
})

## Plot raw # clonotypes per program
cowplot::plot_grid(plotlist=lapply(names(norm_clonotypes_perProgram), function(x){
  df<- norm_clonotypes_perProgram[[x]]
  ggplot(df, aes(x=Program, y=n_clonotypes, fill=Program))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=tcell_colors)+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"),
          legend.position = "none")+
    ggtitle(paste0("Number of clonotypes per program: ", x))+
    xlab("")
}))
ggsave(paste0(figure_dir, "NumberRawClonotypes_perProgram.png"), width=8, height=4)


## Plot normalized clonotypes per program
cowplot::plot_grid(plotlist=lapply(names(norm_clonotypes_perProgram), function(x){
  df<- norm_clonotypes_perProgram[[x]]
  ggplot(df, aes(x=Program, y=n_clonotypes_norm, fill=Program))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=tcell_colors)+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"),
          legend.position = "none")+
    ggtitle(paste0("Normalized clonotypes per program: ", x))+
    xlab("")+
    ylab("Number clonotypes / Number cells")
}))
ggsave(paste0(figure_dir, "NumberNormClonotypes_perProgram.png"), width=8, height=4)

```

## Plot largest clonotype (most expanded) per program
```{r}
## Get max clonotype, normalized by program size
max_clonotype_size<- lapply(names(tenx_seurat_list), function(x){
  max_clonotype<- my_getMaxClonotype(tenx_seurat_list[[x]])
  
  norm_clonotype<- norm_clonotypes_perProgram[[x]]
  max_clonotype<- merge(max_clonotype, norm_clonotype,by="Program" )
  max_clonotype$maxClonotypeSize_norm<- max_clonotype$maxClonotypeSize/max_clonotype$n_cells
  return(max_clonotype)
  })
names(max_clonotype_size)<- names(tenx_seurat_list)

## Plot raw largest clonotype per program
cowplot::plot_grid(plotlist=lapply(names(max_clonotype_size), function(x){
  df<- max_clonotype_size[[x]]
  ggplot(df, aes(x=Program, y=maxClonotypeSize, fill=Program))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=tcell_colors)+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"),
          legend.position = "none")+
    ggtitle(paste0("Largest clonotype per program: ", x))+
    xlab("")
}))
ggsave(paste0(figure_dir, "RawLargestClonotype_perProgram.png"), width=8, height=4)


## Plot normalized largest clonotype per program
cowplot::plot_grid(plotlist=lapply(names(max_clonotype_size), function(x){
  df<- max_clonotype_size[[x]]
  ggplot(df, aes(x=Program, y=maxClonotypeSize_norm, fill=Program))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=tcell_colors)+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"),
          legend.position = "none")+
    ggtitle(paste0("Largest clonotype\nper program, normalized: ", x))+
    xlab("")+
    ylab("largest clonotype / N cells")+
    ylim(0,0.4)
}))
ggsave(paste0(figure_dir, "NormLargestClonotype_perProgram.png"), width=8, height=4)
```




