---
title: "Visualize program proportion changes"
output: html_document
---

After merging all gbm (ss2 and 10x), visualize change in program proportions. Note that these samples integrate poorly- program annotations are done separately for ss2/10x. The results are just visualized in 1 plot here for nicer plotting
```{r}
knitr::opts_chunk$set(echo = TRUE)

library(harmony)

base_dir = "/Users/jlabelle/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/PD1_Treat/02_Merge_SS2.tenX/")
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Scripts/"


preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
source(paste0(base_dir, script_dir, "seurat_transferAnchors_helper.R"))


analysis_dir<- "analysis/PlotPrograms/"
figure_dir<- "figures/PlotPrograms/"
full_cohort_dir<- "../../ImmuneCellAnalysis/"
data_dir<- "analysis/Merge/"

if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}

```


## Load in seurat objects
```{r}
## Load in tenx/ss2 merged seurat objects
files<- list.files(data_dir)
seurat_list<- lapply(files, function(f){
   readRDS(paste0(data_dir, f))
})
names(seurat_list)<- gsub("_seurat.Rds", "", files)


## Colors 
tcell_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Tcell.programs_rank8sharedByunion_reAnnot.Rds"))
myeloid_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Myeloid_program_colors.Rds"))
myeloid_colors<- myeloid_colors$rank6
immune_colors<- c(tcell_colors, myeloid_colors)

colors_treatment<- c(PD1="orange", Untreated="purple")
colors_samples<-c( BT1857_SS2="skyblue", BT2062="gold", BT2080_SS2="salmon",
                   BT1478="purple", BT1745="blue", BT1857_TenX="cyan",
                    BT1910="orange", BT1935="khaki", BT2080_TenX="salmon3")
colors_sequencing<- c(SS2="magenta", TenX="green")

## Full cohort marker genes
tcell_markers<- readRDS(paste0(full_cohort_dir, 
                             "02a_Tcells/02_Ped.Adult/01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                             "new_markers_NMFrank8_bySeurat_sharedByunion_reAnnot.Rds"))
myeloid_markers<- read.csv(paste0(full_cohort_dir,
                               "02b_Myeloid/01a_NMF/analysis/de.novoNMF/pedOnly_nomuv63/", 
                               "top_DE_genes_10Xannotated_NMF_rank6.csv"))

  ## Top markers per program
  top_tcell_markers<- tcell_markers %>% group_by(cluster) %>% top_n(n=5, wt=avg_log2FC) %>% as.data.frame()
  top_myeloid_markers<- myeloid_markers %>% group_by(cluster) %>% top_n(n=5, wt=avg_log2FC) %>% as.data.frame()
```



## UMAPs of projected programs
```{r}
all_plots<- lapply(names(seurat_list), function(x){
  tmp<- seurat_list[[x]]
  DimPlot(tmp, group.by="Program_projected", pt.size = 1, label=FALSE, label.size = 6)+ 
    scale_color_manual(values=immune_colors[names(immune_colors) %in% tmp$Program_projected])+
    ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "UMAP_PredictedProgram.png"), width=18, height=5)

```

## Barchart of program proportions
```{r}

all_plots<- lapply(names(seurat_list), function(x){
  tmp<- seurat_list[[x]]
  my_barchart_programProp(tmp,grouping_variable = "sample", 
                          coloring_variable = "Program_projected",
                          wrapping_variable = "Treatment", colors=immune_colors)+ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "tenx_Barchart_PredictedProgram.pdf"), device="pdf", width=18, height=4)

```

## Pseudobulked heatmap of expression of marker genes
```{r}
marker_list<- list(cd4=top_tcell_markers[!top_tcell_markers$cluster %in% c("Predysfunctional", "Cytotoxic.NKreceptor"),], 
                   cd8=top_tcell_markers[!top_tcell_markers$cluster %in% c("Treg", "Inhibitory"),], 
                   myeloid=top_myeloid_markers)


## BT1857_TenX/BT1857_SS2: replace _ with . (messes up facet wrapping in heatmap)
seurat_list<- lapply(seurat_list, function(x){
  x$sample<- gsub("BT1857_", "BT1857.", gsub("BT2080_", "BT2080.", x$sample))
  return(x)
})


samples<-unique(seurat_list$myeloid$sample)
all_plots<- lapply(names(seurat_list), function(x){
  tmp<- seurat_list[[x]]
  markers_use<-marker_list[[x]]
  markers_use<- markers_use[markers_use$cluster %in% unique(tmp$Program_projected),]
  markers_use<- markers_use[markers_use$gene %in% rownames(tmp),]
  
  samples_use<-samples[samples %in% unique(tmp$sample)]
  
  ## Create pseudobulking variable
  tmp$pb<- paste0(tmp$sample, "_", gsub("\\.", "\n", gsub("_", "\n", tmp$Program_projected)))
  pb<- pseudobulk_byVariable(tmp, tmp@meta.data, "pb")
  
  
  ## center for each sample separately
  center_bySample<- lapply(unique(tmp$sample), function(sample){
    pb_tmp<- pb[,grepl(sample, colnames(pb))]
    if(class(pb_tmp)=="numeric"){
      pb_tmp<- as.data.frame(pb_tmp)
      colnames(pb_tmp)<- colnames(pb)[grepl(sample, colnames(pb))]
    }
    pb_list_tmp<- NormCenter(pb_tmp)
    pb_center_tmp<- pb_list_tmp$center_data
    return(pb_center_tmp)
  })
  center_bySample<- do.call("cbind", center_bySample)
  
  
  ## Plot
  myHeatmap(center_bySample, GOI=markers_use$gene, min.value = -3, max.value = 3, 
                orderFactors =  gsub("\\.", "\n",gsub("_", "\n", unique(markers_use$cluster))),
            orderSubtypes =samples_use) +
    scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")+
    ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "Heatmap_ExpectedMarkerGenes_All.png"), width=40, height=9)
```

## Breakdowns for scCoda-
```{r}
cohort<- "pedOnly_nomuv63"
sccoda_dir<- paste0("../../ImmuneCellAnalysis/scCoda/data/",cohort, "/")


## CD8
meta<- seurat_list$cd8@meta.data

  ## Program by Age
  bd<- as.data.frame.matrix(table(meta$sample, meta$Program_projected))
  
  ## In a sample, can't have more than 50% of cell types be 0
  #nPrograms<- ncol(bd)
  #bd$nPrograms<- apply(bd, 1, function(x){sum(x>1)})
  #bd$PF<- bd$nPrograms>(nPrograms/2)
  #bd<- bd[bd$PF,]; bd$PF<-NULL; bd$nPrograms<-NULL
  
  ## Add on age group to compare proportions
  bd$Subtype<- plyr::mapvalues(rownames(bd), meta$sample, meta$Treatment, warn_missing = FALSE)
  bd$Sample<- rownames(bd)
  
  write.csv(bd, paste0(sccoda_dir,"PD1.Untreated_CD8_Programs.csv"), row.names = FALSE)
  
## CD4
meta<- seurat_list$cd4@meta.data

  ## Program by Age
  bd<- as.data.frame.matrix(table(meta$sample, meta$Program_projected))
  
  ## In a sample, can't have more than 50% of cell types be 0
  #nPrograms<- ncol(bd)
  #bd$nPrograms<- apply(bd, 1, function(x){sum(x>1)})
  #bd$PF<- bd$nPrograms>(nPrograms/2)
  #bd<- bd[bd$PF,]; bd$PF<-NULL; bd$nPrograms<-NULL
  
  ## Add on age group to compare proportions
  bd$Subtype<- plyr::mapvalues(rownames(bd), meta$sample, meta$Treatment, warn_missing = FALSE)
  bd$Sample<- rownames(bd)
  bd$Subtype<- gsub("-", "_", bd$Subtype)
  
  write.csv(bd, paste0(sccoda_dir,"PD1.Untreated_CD4_Programs.csv"), row.names = FALSE)
  
## Myeloid
meta<- seurat_list$myeloid@meta.data

  ## Program by Age
  bd<- as.data.frame.matrix(table(meta$sample, meta$Program_projected))
  
  ## In a sample, can't have more than 50% of cell types be 0
  nPrograms<- ncol(bd)
  bd$nPrograms<- apply(bd, 1, function(x){sum(x>1)})
  bd$PF<- bd$nPrograms>(nPrograms/2)
  bd<- bd[bd$PF,]; bd$PF<-NULL; bd$nPrograms<-NULL
  
  ## Add on age group to compare proportions
  bd$Subtype<- plyr::mapvalues(rownames(bd), meta$sample, meta$Treatment, warn_missing = FALSE)
  bd$Sample<- rownames(bd)
  bd$Subtype<- gsub("-", "_", bd$Subtype)
  
  write.csv(bd, paste0(sccoda_dir,"PD1.Untreated_Myeloid_Programs.csv"), row.names = FALSE)

```