---
title: "Identify shared DEGs"
output: html_document
---

DE ran separately for ss2 and 10x, comparing pd1 vs untreated for each program. These results imported and probed for any genes that are DE in the same direction for both SS2 and 10X

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(harmony)

base_dir = "/Users/jlabelle/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/PD1_Treat/02_Merge_SS2.tenX/")
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Scripts/"


preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
source(paste0(base_dir, script_dir, "seurat_transferAnchors_helper.R"))


analysis_dir<- "analysis/SharedDEGs/"
figure_dir<- "figures/SharedDEGs/"
full_cohort_dir<- "../../ImmuneCellAnalysis/"
data_dir<- "analysis/Merge/"

if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}

```


## Load in seurat objects, DE results
```{r}
## Load in tenx/ss2 merged seurat objects
files<- list.files(data_dir)
seurat_list<- lapply(files, function(f){
   readRDS(paste0(data_dir, f))
})
names(seurat_list)<- gsub("_seurat.Rds", "", files)


## DE results
ss2_de<- readRDS("../01a_SS2/Immune/analysis/program_de/DE_Untreated.v.PD1_byProgram.Rds")
tenx_de<- readRDS("../01b_tenX/analysis/program_de/DE_Untreated.v.PD1_byProgram.Rds")

## Colors 
tcell_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Tcell.programs_rank8sharedByunion_reAnnot.Rds"))
myeloid_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Myeloid_program_colors.Rds"))
myeloid_colors<- myeloid_colors$rank6
immune_colors<- c(tcell_colors, myeloid_colors)

colors_treatment<- c(PD1="orange", Untreated="purple")
colors_samples<-c( BT1857="skyblue", BT2062="gold", BT2080="salmon",
                   BT1478="purple", BT1745="blue", BT1857="cyan",
                    BT1910="orange", BT1935="khaki")
colors_sequencing<- c(SS2="magenta", TenX="green")


```


## Venn diagram of overlaps
Note here that it doesn't take into account logFC direction
```{r}
pvalue_thresh<- .9

ss2_de_filtered<- lapply(ss2_de, function(x){x[x$p_val_adj <=pvalue_thresh,]})
tenx_de_filtered<- lapply(tenx_de, function(x){x[x$p_val_adj <=pvalue_thresh,]})

## Venn diagram of overlaps
all_plots<- lapply(names(ss2_de_filtered), function(x){
  ss2<- ss2_de_filtered[[x]]
  tenx<- tenx_de_filtered[[x]]
  
  programs<- unique(ss2$program)[unique(ss2$program) %in% unique(tenx$program)]
  program_plots<- lapply(programs, function(y){
    ggVennDiagram::ggVennDiagram(list(ss2=ss2[ss2$program==y, "gene"],
                                      tenx=tenx[tenx$program==y, "gene"]))+
      scale_fill_gradient(low="white", high="red")+
      scale_color_manual(values=rep("black", 4))+
      ggtitle(y)
  })
  cowplot::plot_grid(plotlist = program_plots, ncol=length(programs))
})
cowplot::plot_grid(plotlist = all_plots, ncol=1)
ggsave(paste0(figure_dir, "Venn_DEGoverlap_padjThresh", pvalue_thresh, ".png"), width=12, height=8)
```

## Get shared DEGs
```{r}
pvalue_thresh<- .9

ss2_de_filtered<- lapply(ss2_de, function(x){x[x$p_val_adj <=pvalue_thresh,]})
tenx_de_filtered<- lapply(tenx_de, function(x){x[x$p_val_adj <=pvalue_thresh,]})

shared_degs<- lapply(names(ss2_de_filtered), function(x){
  ss2<- ss2_de_filtered[[x]]
  tenx<- tenx_de_filtered[[x]]
  
  ## Subset to columns of interest
  ss2<- ss2[,c("avg_log2FC", "p_val_adj", "program", "gene")]
  tenx<- tenx[,c("avg_log2FC", "p_val_adj", "program", "gene")]
  
  ## rename columns- will merge by gene
  colnames(ss2)<-ifelse(!colnames(ss2) %in% c("gene", "program"), paste0("SS2_", colnames(ss2)), colnames(ss2))
  colnames(tenx)<-ifelse(!colnames(tenx) %in% c("gene", "program"), paste0("TenX_", colnames(tenx)), colnames(tenx))
  
  ## Merge by gene and program
  de_merged<- merge(ss2, tenx, by=c("gene", "program"))
  
  ## Add column whether logFC matches
  de_merged$sameLogFCDirection<- ifelse(sign(de_merged$SS2_avg_log2FC) == sign(de_merged$TenX_avg_log2FC), "Same", "Different")
  return(de_merged)
    
})
names(shared_degs)<- names(ss2_de_filtered)

saveRDS(shared_degs, file=paste0(analysis_dir, "shared_degs_tenx.ss2_pvalueThresh", pvalue_thresh, ".Rds"))
```

## Plot logFC in ss2 vs tenx for each shared gene
```{r}
pvalue_thresh<- 0.9
shared_degs<- readRDS(paste0(analysis_dir, "shared_degs_tenx.ss2_pvalueThresh", pvalue_thresh, ".Rds"))

all_plots<-lapply(names(shared_degs), function(celltype){
  celltype_degs<- shared_degs[[celltype]]
  
  minLogFC<- min(celltype_degs$SS2_avg_log2FC, celltype_degs$TenX_avg_log2FC)
  maxLogFC<- max(celltype_degs$SS2_avg_log2FC, celltype_degs$TenX_avg_log2FC)
  
  celltype_degs$mean_padj<- apply(celltype_degs, 1, function(x){mean(as.numeric(x["SS2_p_val_adj"]), 
                                                                     as.numeric(x["TenX_p_val_adj"]))})
  
  ## linear regression for r2/pvalue
  model<- summary(lm(SS2_avg_log2FC ~ TenX_avg_log2FC, celltype_degs))
  
  ## Plot
  ggplot(celltype_degs, aes(x=SS2_avg_log2FC, y=TenX_avg_log2FC, color=mean_padj))+
    geom_point()+
    theme_bw()+
    xlim(minLogFC, maxLogFC )+
    ylim(minLogFC, maxLogFC)+
    scale_color_gradientn(colours = rainbow(10))+
    ggtitle(celltype, 
            subtitle = paste0("R2: ", round(model$r.squared, 2),
                              "\npvalue: ", formatC(model$coefficients["TenX_avg_log2FC", "Pr(>|t|)"])))+
    geom_smooth(method="lm", color="black", linetype="dashed", se=FALSE)
    
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "Correlation_logFC_pvalueThresh", pvalue_thresh, ".png"), width=15, height=4)
```

## Heatmap of shared DEGs
```{r}
pvalue_thresh<- 0.1
shared_degs<- readRDS(paste0(analysis_dir, "shared_degs_tenx.ss2_pvalueThresh", pvalue_thresh, ".Rds"))
shared_degs<- lapply(shared_degs, function(x){x[x$sameLogFCDirection=="Same",]})

samples<- c("BT1478" , "BT1745","BT1857.SS2","BT1857.TenX","BT1910" ,"BT1935", "BT2062","BT2080")

all_plots<- lapply(names(seurat_list), function(x){
  tmp<- seurat_list[[x]]
  markers_use<-shared_degs[[x]]
  
  ## subset to programs with degs
  tmp<- subset(tmp, Program_projected %in% unique(markers_use$program))
  markers_use<- markers_use[markers_use$gene %in% rownames(tmp),]
  markers_use<- markers_use[order(markers_use$program, markers_use$SS2_avg_log2FC),]
  
  ## Fix samples with underscores
  tmp$sample<- gsub("_", "\\.", tmp$sample)
  
  samples_use<-samples[samples %in% unique(tmp$sample)]
  
  ## Create pseudobulking variable
  tmp$pb<- paste0(tmp$sample, "_", gsub("\\.", "\n", gsub("_", "\n", tmp$Program_projected)))
  pb<- pseudobulk_byVariable(tmp, tmp@meta.data, "pb")
  
  
  ## center for each sample separately
  #center_bySample<- lapply(unique(tmp$sample), function(sample){
  #  pb_tmp<- pb[,grepl(sample, colnames(pb))]
  #  if(class(pb_tmp)=="numeric"){
  #    pb_tmp<- as.data.frame(pb_tmp)
  #    colnames(pb_tmp)<- colnames(pb)[grepl(sample, colnames(pb))]
  #  }
  #  pb_list_tmp<- NormCenter(pb_tmp)
  #  pb_center_tmp<- pb_list_tmp$center_data
  #  return(pb_center_tmp)
  #})
  #center_bySample<- do.call("cbind", center_bySample)
  cm_list<- NormCenter(pb)
  center_bySample<- cm_list$center_data
  
  
  ## Plot
  myHeatmap(center_bySample, GOI=unique(markers_use$gene), min.value = -3, max.value = 3, 
                orderFactors =  gsub("\\.", "\n",gsub("_", "\n", unique(markers_use$program))),
            orderSubtypes =samples_use) +
    scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")+
    ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "Heatmap_SharedDEGs_pvalueThresh",pvalue_thresh, ".png"), width=25, height=5)
```

