---
title: "ProjectBroadAnnots.Rmd"
author: "Jenna LaBelle"
date: "10/07/22"
output: html_document
---

## Expression patterns of various genes of interest acros PD1 treated samples/untreated GBM

```{r}
library(Seurat) 
library(ggplot2)
library(ggpubr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/PD1_Treat/02_Merge_SS2.tenX/")

analysis_dir<- paste0(working_dir, "/analysis/clonal_analysis/")
figure_dir<- paste0(working_dir, "/figures/clonal_analysis/")
data_dir<- "analysis/Merge/"
full_cohort_dir<- "../../ImmuneCellAnalysis/"

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)
source(paste0(base_dir, script_dir, "seurat_transferAnchors_helper.R"))

```


## Load seurat objects, colors, tcr info
```{r}
## Load in tenx/ss2 merged seurat objects
files<- list.files(data_dir)
files<- files[!grepl("myeloid", files)]
seurat_list<- lapply(files, function(f){
   readRDS(paste0(data_dir, f))
})
names(seurat_list)<- gsub("_seurat.Rds", "", files)

## Colors 
tcell_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Tcell.programs_rank8sharedByunion_reAnnot.Rds"))
myeloid_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Myeloid_program_colors.Rds"))
myeloid_colors<- myeloid_colors$rank6
immune_colors<- c(tcell_colors, myeloid_colors)
colors_treatment<- c(PD1="orange", Untreated="purple")

## tenx/ss2 clonal info named differently- fix
seurat_list<- lapply(seurat_list, function(x){
  ## sometimes NA is character, sometimes it's special- change so that all are character
  x$clonal<- ifelse(is.na(x$clonal), "NA", x$clonal)
  x$clonotype_id<- ifelse(is.na(x$clonotype_id), "NA", x$clonotype_id)
  x$clonotype_size<- ifelse(is.na(x$clonotype_size), "NA", x$clonotype_size)
  
  ## if tcr was reconstructed but group size is NA- reassign to 1
  x$clonotype_size<- ifelse(x$tcr_reconstructed=="Yes" & x$clonotype_size=="NA", 1, x$clonotype_size)
  
  ## if clonotype size is avaialble but tcr was was assigned as not assembled- reassign to tcr assembled (just will be nonclonal)
  x$tcr_reconstructed<-  ifelse(is.na(x$clonotype_size) | x$clonotype_size=="NA", "No", "Yes")
  
  ## if tcr is reconstructed but group size =1 or NA: reassign as nonclonal
  x$clonal<- ifelse( (is.na(x$clonotype_size) | x$clonotype_size==1 | x$clonotype_size=="NA") & x$tcr_reconstructed=="Yes", "nonclonal", x$clonal)
  
  ## if tcr was not reconstructed- reassign clonal as NA
  x$clonal<- ifelse(x$tcr_reconstructed=="No", "NA", x$clonal)
  
  
  return(x)
})

```

## Percentage of clonally expanded T cells in treated vs untreated
separate for 10x/ss2
```{r}
ss2_list<- lapply(seurat_list, function(x){subset(x, Sequencing=="SS2")})
tenx_list<- lapply(seurat_list, function(x){subset(x, Sequencing=="TenX")})

## SS2
all_plots<- lapply(names(ss2_list), function(celltype){
  seurat<- ss2_list[[celltype]]
  clonal_meta<- seurat@meta.data[,c("clonal", "tcr_reconstructed", "clonotype_id", "clonotype_size", "Treatment", "sample")]
  
  ## subset to cells with tcr reconstructed
  clonal_meta<- clonal_meta[clonal_meta$tcr_reconstructed=="Yes",]
  
  ## proportion of clonal for treated/untreated
  clonal_prop<- as.data.frame(table(clonal_meta$sample, clonal_meta$clonal))
  colnames(clonal_prop)<- c("Sample", "Clonal", "NCells")
  clonal_prop<- clonal_prop %>% group_by(Sample) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()
  clonal_prop$Treatment<-as.character(plyr::mapvalues(clonal_prop$Sample, clonal_meta$sample, clonal_meta$Treatment, warn_missing = FALSE))
  clonal_prop$Treatment<- factor(clonal_prop$Treatment, levels=c("Untreated", "PD1"))
  
  ggplot(clonal_prop, aes(x=Sample, y=PerCells, fill=Clonal))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=c(clonal="orange", nonclonal="navy"))+
    theme_classic()+
    ggtitle(celltype)+
    theme(axis.text = element_text(face="bold", color="black"),
          strip.text = element_text(face="bold", color="black"),
          axis.text.x = element_text(angle=45, hjust=1))+
    facet_grid(cols=vars(Treatment), space="free", scale="free")+
    xlab("")+ylab("Proportion of clonally expanded T cells")
  
})
cowplot::plot_grid(plotlist = all_plots, ncol=2)
ggsave(paste0(figure_dir,"barchart_SS2_proportionClonallyExpanded.png"), width=8, height=3)

## TenX
all_plots<- lapply(names(tenx_list), function(celltype){
  seurat<- tenx_list[[celltype]]
  clonal_meta<- seurat@meta.data[,c("clonal", "tcr_reconstructed", "clonotype_id", "clonotype_size", "Treatment", "sample")]
  
  ## subset to cells with tcr reconstructed
  clonal_meta<- clonal_meta[clonal_meta$tcr_reconstructed=="Yes",]
  
  ## proportion of clonal for treated/untreated
  clonal_prop<- as.data.frame(table(clonal_meta$sample, clonal_meta$clonal))
  colnames(clonal_prop)<- c("Sample", "Clonal", "NCells")
  clonal_prop<- clonal_prop %>% group_by(Sample) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()
  clonal_prop$Treatment<-as.character(plyr::mapvalues(clonal_prop$Sample, clonal_meta$sample, clonal_meta$Treatment, warn_missing = FALSE))
  clonal_prop$Treatment<- factor(clonal_prop$Treatment, levels=c("Untreated", "PD1"))
  
  ggplot(clonal_prop, aes(x=Sample, y=PerCells, fill=Clonal))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=c(clonal="orange", nonclonal="navy"))+
    theme_classic()+
    ggtitle(celltype)+
    theme(axis.text = element_text(face="bold", color="black"),
          strip.text = element_text(face="bold", color="black"),
          axis.text.x = element_text(angle=45, hjust=1))+
    facet_grid(cols=vars(Treatment), space="free", scale="free")+
    xlab("")+ylab("Proportion of clonally expanded T cells")
  
})
cowplot::plot_grid(plotlist = all_plots, ncol=2)
ggsave(paste0(figure_dir,"barchart_TenX_proportionClonallExpanded.png"), width=8, height=3)
```

## clonal group size by treatment
only for tenx- really not enough cells for ss2
```{r}
tenx_list<- lapply(seurat_list, function(x){subset(x, Sequencing=="TenX")})

## tenx- boxplot
all_plots<- lapply(names(tenx_list), function(celltype){
  seurat<- tenx_list[[celltype]]
  clonal_meta<- seurat@meta.data[,c("clonal", "tcr_reconstructed", "clonotype_id", "clonotype_size", "Treatment", "sample")]
  
  ## subset to cells with tcr reconstructed
  clonal_meta<- clonal_meta[clonal_meta$tcr_reconstructed=="Yes",]
  clonal_meta$clonotype_size<- as.numeric(clonal_meta$clonotype_size)

  clonal_meta$Treatment<- factor(clonal_meta$Treatment, levels=c("Untreated", "PD1"))  
  ggplot(clonal_meta, aes(x=Treatment, y=log(clonotype_size), fill=Treatment))+
    geom_boxplot(outlier.size=0)+
    geom_jitter(width=0.1, size=0.1)+
    scale_fill_manual(values=colors_treatment)+
    stat_compare_means(method="t.test")+
    theme_classic()+
    ggtitle(celltype)+
    theme(axis.text = element_text(face="bold", color="black"),
          strip.text = element_text(face="bold", color="black"),
          axis.text.x = element_text(angle=45, hjust=1))
  
})
cowplot::plot_grid(plotlist = all_plots, ncol=2)
ggsave(paste0(figure_dir,"boxplot_TenX_groupSize.png"), width=8, height=4)

## NOT RUN
## Barchart by binned group size
all_plots<- lapply(names(tenx_list), function(celltype){
  seurat<- tenx_list[[celltype]]
  clonal_meta<- seurat@meta.data[,c("clonal", "tcr_reconstructed", "clonotype_id", "clonotype_size", "Treatment", "sample")]
  
  ## subset to cells with tcr reconstructed
  clonal_meta<- clonal_meta[clonal_meta$tcr_reconstructed=="Yes",]
  clonal_meta$clonotype_size<- as.numeric(clonal_meta$clonotype_size)
  
  clonal_meta<- clonal_meta %>% mutate(size_bin = cut(clonotype_size, breaks=c(0, 2, 5, 10, 15, 20,50,75, max(clonal_meta$clonotype_size))))
  clonal_meta$firstNumber<- as.numeric(gsub("\\(", "", gsub(",.*", "", as.character(clonal_meta$size_bin))))
  clonal_meta$size_bin<- gsub("\\(", "", gsub(",", "-", gsub("]", "", clonal_meta$size_bin)))
  bin_size_order<- clonal_meta$size_bin; names(bin_size_order)<- clonal_meta$firstNumber
  clonal_meta$size_bin<- factor(clonal_meta$size_bin, levels=unique(clonal_meta$size_bin[order(clonal_meta$firstNumber)] ))
  clonal_meta$Treatment<- factor(clonal_meta$Treatment, levels=c("Untreated", "PD1"))  
  
  ## number of cells in each bin
  clonal_prop<- as.data.frame(table(clonal_meta$Treatment, clonal_meta$size_bin))
  colnames(clonal_prop)<- c("Treatment", "SizeBin", "NCells")
  clonal_prop<- clonal_prop %>% group_by(Treatment) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()


  ggplot(clonal_prop, aes(x=Treatment, y=PerCells, fill=rev(SizeBin)))+
    geom_bar(stat="identity")+
    geom_jitter(width=0.1, size=0.1)+
    #scale_fill_manual(values=colors_treatment)+
    stat_compare_means(method="t.test")+
    theme_classic()+
    ggtitle(celltype)+
    theme(axis.text = element_text(face="bold", color="black"),
          strip.text = element_text(face="bold", color="black"),
          axis.text.x = element_text(angle=45, hjust=1))
  
})
cowplot::plot_grid(plotlist = all_plots, ncol=2)
ggsave(paste0(figure_dir,"boxplot_TenX_groupSize.png"), width=8, height=3)
```

## Vln of clone size 
```{r}
ss2_list<- lapply(seurat_list, function(x){subset(x, Sequencing=="SS2")})
tenx_list<- lapply(seurat_list, function(x){subset(x, Sequencing=="TenX")})

## Tenx- by treatment
all_plots<- lapply(c("cd4", "cd8"), function(celltype){
  seurat<- tenx_list[[celltype]]
  seurat<- subset(seurat, clonotype_size != "NA")
  seurat$clonotype_size<- as.numeric(seurat$clonotype_size)
  VlnPlot(seurat, group.by = "Treatment", features="clonotype_size", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=colors_treatment)+
    ggtitle(paste0("Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
})
cowplot::plot_grid(plotlist=all_plots, ncol=1)
ggsave(paste0(figure_dir, "Vln_clonotypeSize_byTreatment_TenX.png"), width=4, height=8)

## Tenx- by program
all_plots<- lapply(c("cd4", "cd8"), function(celltype){
  seurat<- tenx_list[[celltype]]
  seurat<- subset(seurat, clonotype_size != "NA")
  seurat$clonotype_size<- as.numeric(seurat$clonotype_size)
  VlnPlot(seurat, group.by = "Program_projected", features="clonotype_size", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=immune_colors)+
    ggtitle(paste0("Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
})
cowplot::plot_grid(plotlist=all_plots, ncol=1)
ggsave(paste0(figure_dir, "Vln_clonotypeSize_byProgram_TenX.png"), width=6, height=8)

## tenx- by program, treated/untreated separately
all_plots<- lapply(c("cd4", "cd8"), function(celltype){
  seurat<- tenx_list[[celltype]]
  seurat<- subset(seurat, clonotype_size != "NA")
  seurat$clonotype_size<- as.numeric(seurat$clonotype_size)
  p1<- VlnPlot(subset(seurat, Treatment=="Untreated"), group.by = "Program_projected", features="clonotype_size", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=immune_colors)+
    ggtitle(paste0("Untreated Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
  
  p2<- VlnPlot(subset(seurat, Treatment=="PD1"), group.by = "Program_projected", features="clonotype_size", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=immune_colors)+
    ggtitle(paste0("PD1 Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
  p1+p2
})
cowplot::plot_grid(plotlist=all_plots, ncol=2)
ggsave(paste0(figure_dir, "Vln_clonotypeSize_byProgram.Treatment_TenX.png"), width=12, height=8)

## tenx- by program, treated/untreated separately, normalized by program size
all_plots<- lapply(c("cd4", "cd8"), function(celltype){
  seurat<- tenx_list[[celltype]]
  seurat<- subset(seurat, clonotype_size != "NA")
  seurat$clonotype_size<- as.numeric(seurat$clonotype_size)
  
  ## Normalize by size of program
  nCells_program<- as.data.frame(table(seurat$Program_projected))
  seurat$NCells<- as.numeric(as.character(plyr::mapvalues(seurat$Program_projected, 
                                                              nCells_program$Var1, nCells_program$Freq, warn_missing = FALSE)))
  seurat$clonotype_size_norm<- seurat$clonotype_size / seurat$NCells * 1000
  
  p1<- VlnPlot(subset(seurat, Treatment=="Untreated"), group.by = "Program_projected", features="clonotype_size_norm", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=immune_colors)+
    ggtitle(paste0("Untreated Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
  
  p2<- VlnPlot(subset(seurat, Treatment=="PD1"), group.by = "Program_projected", features="clonotype_size_norm", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=immune_colors)+
    ggtitle(paste0("PD1 Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
  p1+p2
})
cowplot::plot_grid(plotlist=all_plots, ncol=2)
ggsave(paste0(figure_dir, "Vln_clonotypeSizeNorm_byProgram.Treatment_TenX.png"), width=12, height=8)




## SS2- by treatment
all_plots<- lapply(c("cd4", "cd8"), function(celltype){
  seurat<- ss2_list[[celltype]]
  seurat<- subset(seurat, clonotype_size != "NA")
  seurat$clonotype_size<- as.numeric(seurat$clonotype_size)
  VlnPlot(seurat, group.by = "Treatment", features="clonotype_size", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=colors_treatment)+
    ggtitle(paste0("Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
})
cowplot::plot_grid(plotlist=all_plots, ncol=1)
ggsave(paste0(figure_dir, "Vln_clonotypeSize_byTreatment_SS2.png"), width=4, height=8)

## ss2- by program
all_plots<- lapply(c("cd4", "cd8"), function(celltype){
  seurat<- ss2_list[[celltype]]
  seurat<- subset(seurat, clonotype_size != "NA")
  seurat$clonotype_size<- as.numeric(seurat$clonotype_size)
  VlnPlot(seurat, group.by = "Program_projected", features="clonotype_size", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=immune_colors)+
    ggtitle(paste0("Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
})
cowplot::plot_grid(plotlist=all_plots, ncol=1)
ggsave(paste0(figure_dir, "Vln_clonotypeSize_byProgram_SS2.png"), width=6, height=8)

## SS2- by program, treated/untreated separately
all_plots<- lapply(c("cd4", "cd8"), function(celltype){
  seurat<- ss2_list[[celltype]]
  seurat<- subset(seurat, clonotype_size != "NA")
  seurat$clonotype_size<- as.numeric(seurat$clonotype_size)
  p1<- VlnPlot(subset(seurat, Treatment=="Untreated"), group.by = "Program_projected", features="clonotype_size", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=immune_colors)+
    ggtitle(paste0("Untreated Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
  
  p2<- VlnPlot(subset(seurat, Treatment=="PD1"), group.by = "Program_projected", features="clonotype_size", sort=TRUE)+NoLegend()+
    scale_fill_manual(values=immune_colors)+
    ggtitle(paste0("PD1 Group size: ", celltype))+
    xlab("")+ ylab("clonotype group size")
  p1+p2
})
cowplot::plot_grid(plotlist=all_plots, ncol=2)
ggsave(paste0(figure_dir, "Vln_clonotypeSize_byProgram.Treatment_SS2.png"), width=12, height=8)
```


NOT RUN
## Plot group size of clonotype (including clones where n=1; i.e., TCR reconstructed but not expanded) in a histogram
```{r}
seurat_list<- seurat_list

## Plot clonotype sizes (raw)
all_plots<- lapply(seurat_list, function(seurat){
  meta<- seurat@meta.data
  
  ## Subset to cells with TCR
  meta<- meta[meta$tcr_reconstructed=="Yes",]
  meta$clonotype_size<- as.numeric(meta$clonotype_size)
  
  ## Plot in hist, by treatment
  ggplot(meta, aes(x=clonotype_size, fill=Treatment))+
    geom_histogram(bins=4)+
    ggtitle(unique(meta$detailed_annot))+
    theme_bw()+
    scale_fill_manual(values=colors_treatment)+
    xlab("Clonotype size")
})
cowplot::plot_grid(plotlist=all_plots)

## Plot clonotype sizes as a proportion of total TCRs
## % of all TCRs with clonotype size 1, etc
all_plots<- lapply(seurat_list, function(seurat){
  meta<- seurat@meta.data
  
  ## Subset to cells with TCR
  meta<- meta[meta$tcr_reconstructed=="Yes",]

  ## Get number/proportion of all cells (with TCR) in each clonotype size
  meta$clonotype_size<- as.numeric(meta$clonotype_size)
  bd<- as.data.frame(table(meta$clonotype_size, meta$Treatment))
  colnames(bd)<- c("ClonotypeSize", "Treatment", "NumberCells")
  bd<- bd %>% group_by(Treatment) %>% mutate(PerCells=NumberCells/sum(NumberCells)) %>% as.data.frame()
  
  ## Plot
  ggplot(bd, aes(x=ClonotypeSize, y=PerCells, fill=Treatment))+
    geom_bar(stat="identity")+
    theme_bw()+
    xlab("Clonotype size")+
    ylab("Percentage of cells in clonotype size")+
    scale_fill_manual(values=colors_treatment)+
    ggtitle(unique(meta$detailed_annot))
})
cowplot::plot_grid(plotlist=all_plots)

## Boxplot of clonotype size by program
all_plots<- lapply(seurat_list, function(seurat){
  meta<- seurat@meta.data
  
  ## Subset to cells with TCR
  meta<- meta[meta$tcr_reconstructed=="Yes",]
  meta$clonotype_size<- as.numeric(meta$clonotype_size)
  

  ## Plot
  ggplot(meta, aes(x=Program_projected, y=clonotype_size, fill=Treatment))+
    #geom_boxplot()+
    geom_bar(stat="identity", position="dodge")+
    theme_bw()+
    xlab("Clonotype size")+
    ylab("Percentage of cells in clonotype size")+
    scale_fill_manual(values=colors_treatment)+
    ggtitle(unique(meta$detailed_annot))
})
cowplot::plot_grid(plotlist=all_plots)
```

```{r}

  
## TCR data- load for PD1, BT1857, BT1478, 10X (treated and untreated)
  ## SmartSeq2
  ## BT1857 and BT1478 data load separately bc the latter processed by Orr, former by me/Li
    ## PD1- raw cell_data.csv from tracer available
    pd1_tcr_dir<- "../SS2/Immune/data/tracer_output/"
    pd1_tcr_samples<- list.files(pd1_tcr_dir)
    all_tcr<- lapply(pd1_tcr_samples, function(sample){
      tmp<- read.csv(paste0(pd1_tcr_dir, sample, "/cell_data.csv"))
      tmp$sample<- sample
      return(tmp)
    })
    pd1_tcr<- do.call(rbind, all_tcr)
    
    ## BT1478 (whole cohort, subset to just this sample)
    fullCohort_tcr_dir<-"../../ImmuneCellAnalysis/data/data_fromOrr"
    load(paste0(working_dir, fullCohort_tcr_dir, "2020_12_10_clonalTcells_cluster.Rda"))
    bt1478_cells_with_tcr<- tcell.reconstructed$BT1478
    bt1478_clonotypes<- clonotypes.all[clonotypes.all$experiment=="BT1478",]
    rm(clonal); rm(clonotypes.all); rm(gcdata); rm(clonotypes.cells)
    
    ## BT1857- raw cell_data.csv from tracer available
    newSamples_tcr_dir<- "../../ImmuneCellAnalysis/01_Preprocessing/01_Tracer_Preprocess/data/BT1857/output/filtered_TCRAB_summary/"
    bt1857_tcr<- read.csv(paste0(newSamples_tcr_dir, "cell_data.csv"))
    bt1857_tcr$sample<- "BT1857"
    
   ## Add BT1857/pd1 TCR info together + subset to "productive" TCRs (either A or B is not "")
    pd1_bt1857_tcr<- rbind(pd1_tcr, bt1857_tcr)
    rownames(pd1_bt1857_tcr)<- pd1_bt1857_tcr$cell_name
    pd1_bt1857_tcr<-pd1_bt1857_tcr[pd1_bt1857_tcr$A_productive != "" | pd1_bt1857_tcr$B_productive != "",]
    pd1_bt1857_tcr<- pd1_bt1857_tcr[,c("clonotype_id", "clonotype_size")]
    rownames(pd1_bt1857_tcr)<- gsub( "_", "\\.",rownames(pd1_bt1857_tcr))
  
  ## tenX
  pd1_tenx_tcr<- read.csv("../tenX/data/multi/per_sample_outs/BT2062/vdj_t/filtered_contig_annotations.csv")
  bt1857_tenx_tcr<- read.csv("../../ImmuneCellAnalysis/tenX/unTreatedAlone/data/VDJ/multi/per_sample_outs/BT1857/vdj_t/filtered_contig_annotations.csv")
  pd1_tenx_tcr$raw_clonotype_id_sample<- paste0("BT2062_", pd1_tenx_tcr$raw_clonotype_id)
  bt1857_tenx_tcr$raw_clonotype_id_sample<- paste0("BT1857_", bt1857_tenx_tcr$raw_clonotype_id)
  tenx_tcr<- rbind(pd1_tenx_tcr, bt1857_tenx_tcr)
```
