---
title: "BroadAnnotation"
author: "Jenna LaBelle"
date: "11/04/21"
output: html_document
---

## Purpose:
Broad level annotation (Tcell vs Myeloid, CD4 vs CD8) 


## Approach:
Expression of common marker genes, perform cluster-wise annotation. Have option to use harmony-integrated or non-integrated seurat objects (no integration used for final analysis for assigning tcell/myeloid)

```{r}
library(Seurat) 

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/01_Preprocessing/02_Counts_Preprocess/")

integration<- "" ## HarmonyInt/ or ""
analysis_dir<- paste0(working_dir, "/analysis/broad_annotation/", integration)
figure_dir<- paste0(working_dir, "/figures/broad_annotation/", integration)

if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```


## Load in processed seurat object 
## Can either be for all samples, or ped only samples (no MUV90/91/92, or MUV62)
```{r}
cohort<- "pedOnly_nomuv63"
figure_dir<- paste0(figure_dir, cohort, "/")
analysis_dir<-  paste0(analysis_dir, cohort, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

immune_seurat<- readRDS(paste0("analysis/merge_samples/",integration, "tcell.myeloid_seurat_", cohort, ".Rds"))
```


##################################
## Identify Tcell/Myeloid cells ##
##################################

## Basic plot to determine myeloid/tcell
## Note: these figures were also generated for seurat object BEFORE removing cluster 11 (low quality)
## These pre-removal figures are in figures/broadAnnotation/WithCluster11/
## Figures in figures/merge_samples/broadAnnotation do NOT include cluster 11
```{r}
FeaturePlot(immune_seurat, features=c("CD4", "CD8A", "CD3G","CD2", "CD247", "CD14", "CSF1R"), ncol= 4)
ggsave(paste0(figure_dir, "UMAP_Tcell.Myeloid_Markers.png"), width=16, height=8)

VlnPlot(immune_seurat, features=c("CD4", "CD8A", "CD3G", "CSF1R","CD2", "CD247", "CD14"), sort=FALSE, ncol= 1)
ggsave(paste0(figure_dir, "Vln_Tcell.Myeloid_Markers.png"), width=8, height=20)

DimPlot(immune_seurat, label=TRUE)+NoLegend()
ggsave(paste0(figure_dir, "UMAP_seurat.png"), width=6, height=6)

# Glial cells
genes.astrocytes <- c("APOE", "SLC1A2", "SLC1A3", "ALDOC", "GFAP", "CLU", "GPM6B", "TUBA1A", "NGFRAP1", "CD9", "GFAP", "CKB", "PTGDS", "S100B", "STMN1")
immune_seurat <- AddModuleScore(immune_seurat, list(genes.astrocytes), ctrl = 10, name = c("astrocyte"))
FeaturePlot(immune_seurat, c("astrocyte1"))
ggsave(paste0(figure_dir, "UMAP_astrocyte.png"), width = 6, height = 6)

# Cyototxicity 
genes.cytotoxicity <- c("PRF1", "GZMB", "GZMA", "GZMH", "NKG7", "GNLY")
immune_seurat <- AddModuleScore(immune_seurat, list(genes.cytotoxicity), ctrl = 10, name = c("cytotoxicity"))
FeaturePlot(immune_seurat, c("cytotoxicity1"))
ggsave(paste0(figure_dir, "_UMAP_cytotoxicity.png"), width = 8, height = 8)

# Antigen presenting cells
genes.APC <- c("CD74", "CD86", "LYZ", "MS4A7", "FCER1G", "FCGR3A", "CD14", "CBFA2T3", "CD163", "CD33", "ID2")
FeaturePlot(immune_seurat, genes.APC,pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_APC.png"), width = 12, height = 9)

# T cells
genes.Tcell <- c("CD3D", "CD3E", "CD4","CD8A", "CD28")
FeaturePlot(immune_seurat, genes.Tcell,  pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell.png"), width =12, height = 9)

# Th17
genes.Th17 <- c("IL17A", "IL17F", "IL23R", "IL22", "RORC", "PLZF")
FeaturePlot(immune_seurat, genes.Th17,  pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell_Th17.png"), width = 6, height = 9)

# Treg cells
genes.Treg <- c("CD4","CD25","FOXP3","SMAD3","STAT5A", "STAT5B")
FeaturePlot(immune_seurat, genes.Treg,  pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell_Treg.png"), width = 6, height = 9)

# Th1 cells
genes.Th1 <- c("CD4","CXCR3","TBX21","STAT1","STAT6")
FeaturePlot(immune_seurat, genes.Th1,  pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell_Th1.png"), width = 6, height = 9)

# Inhibitory T cell markers (HAVCR2 = TIM-3)
genes.inhibit <- c("PDCD1", "CTLA4", "HAVCR2", "LAG3", "TIGIT")
FeaturePlot(immune_seurat, genes.inhibit,  pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell_inhibit.png"), width = 6, height = 9)

# T cell Cytokines
genes <- c("IL2", "IFNG", "TNF", "IL10", "TGFB1", "TGFB2")
FeaturePlot(immune_seurat, genes,  pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell_cytokine.png"), width = 6, height = 9)

# T-cell cytotoxicity
genes <- c("GZMA", "GZMB", "PRF1")
FeaturePlot(immune_seurat, genes,  pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell_cytotoxic.png"), width = 8, height = 8)

# T cell activation/regulatory
genes <- c("HLA-DRA", "ICOS", "TNFRSF4", "TNFRSF18")
FeaturePlot(immune_seurat, genes,  pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell_activation.png"), width = 6, height = 6)

# T cell memory
genes <- c("SELL", "IL7R", "CD44")
FeaturePlot(immune_seurat, genes, pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell_memory.png"), width = 6, height = 6)

# T cell TFs
genes <- c("TBX21", "STAT1", "RUNX3", "EOMES", "NFATC2", "PRDM1", "FOXP3", "STAT5A", "STAT5B", "RORA", "RORC", "STAT3")
FeaturePlot(immune_seurat, genes,  pt.size = 1)
ggsave(paste0(figure_dir, "UMAP_Tcell_tfs.png"), width = 12, height = 9)


immune_seurat$astrocyte1<-NULL; immune_seurat$cytotoxicity1<-NULL
```

## WITHOUT INTEGRATION
## Based on above graphs, assign each cluster to T cell or Myeloid
```{r}
## Need to increase resolution- one cluster appears to be mostly myeloid, with 1 small cluster of cd8+ t cells
## cluster 11 in all_samples, cluster 10 in pedOnly, cluster 1/9 in pedOnly_nomuv63
immune_seurat<- FindClusters(immune_seurat, resolution = 1.6, algorithm = 4, random.seed = 100)
seurat_clusters<- as.numeric(as.character(unique(immune_seurat$seurat_clusters)))
seurat_clusters<- seurat_clusters[order(seurat_clusters)]

##  For pedOnly, pedOnly_nomuv63: After increasing resolution, one cluster (15 for pedOnly, 12, for pedOnly_nomuv63, 17 for allSamples) is still split into 2 small tcell/myeloid cells

## For these cells: rerun clustering, assign that way
pedOnly_unclearcluster<- subset(immune_seurat, seurat_clusters==17) 
pedOnly_unclearcluster$sampleid<- "dummy_sample"
pedOnly_unclearcluster<- RunFullSeurat_Immune(pedOnly_unclearcluster@assays$RNA@counts, pedOnly_unclearcluster$sampleid)
pedOnly_unclearcluster$broad_annot<- ifelse(pedOnly_unclearcluster$seurat_clusters==1, "Tcell", "Myeloid") ## changed from 2 for pedOnly_novmu63
p1<- DimPlot(pedOnly_unclearcluster, label=TRUE)+NoLegend()+
  ggtitle("In full cohort, these cells assigned to single cluster\nRerun clustering, assign to correct broad annot")
p2<- FeaturePlot(pedOnly_unclearcluster, features=c("CD3G", "CSF1R"))
p3<- DimPlot(pedOnly_unclearcluster, group.by="broad_annot", label=TRUE)+NoLegend()+ggtitle("")
p1+p2+p3
ggsave(paste0(figure_dir, "forUnclearCluster_RerunCluster_AssignMyeloid.Tcell.png"), width=12, height=4)

## Assign each cluster to tcell/myeloid
  ## allSamples:
  names(seurat_clusters)<- c("Myeloid", "Tcell", "Tcell", "Myeloid", "Tcell", "Myeloid", "Tcell", "Tcell",
                             "Tcell", "Tcell", "Myeloid", "Tcell", "Tcell", "Myeloid", "Myeloid", "Myeloid", "Reannotate")
  ## pedOnly:
  ##names(seurat_clusters)<- c("Tcell", "Tcell", "Myeloid", "Myeloid", "Tcell", "Tcell", "Tcell",
                            # "Tcell", "Tcell", "Myeloid", "Tcell", "Myeloid", "Myeloid", "Tcell", "Tcell", "Myeloid",
                           #  "Myeloid", "Myeloid")

  ## pedOnly_nomuv63
  names(seurat_clusters)<- c("Tcell", "Tcell", "Myeloid", "Myeloid","Tcell", "Tcell", "Tcell", "Tcell",
                             "Myeloid", "Tcell", "Tcell", "Tcell", "Myeloid", "Myeloid", "Tcell", "Myeloid", "Myeloid")
  
  ## pedOnly_nomuv63.withmuv91
  #names(seurat_clusters)<- c("Myeloid", "Tcell", "Myeloid", "Tcell", "Tcell", "Tcell", "Myeloid", "Tcell", 
                            # "Tcell", "Tcell", "Tcell", "Tcell", "Myeloid", "Myeloid", "Tcell", 
                             #"Myeloid", "Myeloid","Tcell")
  
    
immune_seurat$broad_annot<- as.character(plyr::mapvalues(immune_seurat$seurat_clusters, seurat_clusters,
                                                         names(seurat_clusters)))

## For ped only/pedOnly_muv63/allSamples: for cluster 15/12/17 cells, reassign annotation based on above analysis
immune_seurat$broad_annot<- ifelse(colnames(immune_seurat) %in% colnames(pedOnly_unclearcluster),
                                   as.character(plyr::mapvalues(colnames(immune_seurat), 
                                                   colnames(pedOnly_unclearcluster), 
                                                   pedOnly_unclearcluster$broad_annot)),
                                   immune_seurat$broad_annot)
immune_seurat$seurat_clusters<- ifelse(colnames(immune_seurat) %in% 
                                         colnames(pedOnly_unclearcluster)[pedOnly_unclearcluster$broad_annot=="Myeloid"],
                                       length(seurat_clusters)+1, immune_seurat$seurat_clusters) ## manually adding another cluster
immune_seurat@active.ident<- factor(immune_seurat$seurat_clusters)

## Confirm with dot plot- each cluster should express only CD3G or CSF1R 
p1<- DotPlot(immune_seurat, features=c("CD3G", "CSF1R"))
p2<- DimPlot(immune_seurat, label=TRUE)+NoLegend()
p3<- DimPlot(immune_seurat, group.by = "broad_annot", label=TRUE) + NoLegend()
p1+p2+p3
ggsave(paste0(figure_dir, "Myeloid.Tcell_BroadAnnotation_CD3G.CSF1R.png"), width=18, height=6)
```

## WITH INTEGRATION
```{r}
seurat_clusters<- as.numeric(as.character(unique(immune_seurat$seurat_clusters)))
seurat_clusters<- seurat_clusters[order(seurat_clusters)]

names(seurat_clusters)<- c("Myeloid", "Tcell", "Tcell", "Tcell", "Tcell", "Myeloid", "Myeloid",
                           "Tcell", "Myeloid")
immune_seurat$broad_annot<- as.character(plyr::mapvalues(immune_seurat$seurat_clusters, seurat_clusters,
                                                         names(seurat_clusters)))


## Confirm with dot plot- each cluster should express only CD3G or CSF1R 
p1<- DotPlot(immune_seurat, features=c("CD3G", "CSF1R"))
p2<- DimPlot(immune_seurat, label=TRUE)+NoLegend()
p3<- DimPlot(immune_seurat, group.by = "broad_annot", label=TRUE) + NoLegend()
p1+p2+p3
ggsave(paste0(figure_dir, "Myeloid.Tcell_BroadAnnotation_CD3G.CSF1R.png"), width=18, height=6)
```


## Split by Myeloid/T cell, rerun seurat pipeline for each separately
```{r}
## Split by broad annotation
myeloid_seurat<- subset(immune_seurat, broad_annot=="Myeloid")
tcell_seurat<- subset(immune_seurat, broad_annot=="Tcell")

## Save metadata for adding to seurat
myeloid_meta<- myeloid_seurat@meta.data; tcell_meta<- tcell_seurat@meta.data

## Run full seurat pipeline on myeloid/tcell separately
if(integration=="HarmonyInt/"){
  library(harmony)
  myeloid_seurat<- RunFullSeurat(myeloid_seurat@assays$RNA@counts, samples=myeloid_seurat$sampleid, RunHarmony = TRUE)
  tcell_seurat<- RunFullSeurat(tcell_seurat@assays$RNA@counts, samples=tcell_seurat$sampleid, RunHarmony = TRUE)
}else{
  myeloid_seurat<- RunFullSeurat_Immune(myeloid_seurat@assays$RNA@counts, samples=myeloid_seurat$sampleid)
  tcell_seurat<- RunFullSeurat_Immune(tcell_seurat@assays$RNA@counts, samples=tcell_seurat$sampleid)
}


## Add metadata back
myeloid_seurat<- AddMetaData(myeloid_seurat, metadata=myeloid_meta[,!colnames(myeloid_meta) %in%
                                                                     colnames(myeloid_seurat@meta.data)])
tcell_seurat<- AddMetaData(tcell_seurat, metadata=tcell_meta[,!colnames(tcell_meta) %in% 
                                                               colnames(tcell_seurat@meta.data)])

## Save 
saveRDS(myeloid_seurat, file=paste0(analysis_dir, "myeloid_seurat_preReAnnotation.Rds"))
saveRDS(tcell_seurat, file=paste0(analysis_dir, "tcell_seurat_preReAnnotation.Rds"))
saveRDS(immune_seurat, file=paste0(analysis_dir, "tcell.myeloid_seurat_preReAnnotation.Rds"))
```




###############################################
## T cell filtering based on Nathan/Orr paper##
###############################################


##remove any NK/astrocytes,  unclear CD4/CD8 cycling cells 
```{r}
tcell_seurat<- readRDS(paste0(analysis_dir, "tcell_seurat_preReAnnotation.Rds"))
myeloid_seurat<- readRDS(paste0(analysis_dir, "myeloid_seurat_preReAnnotation.Rds"))

## Astrocytes- no clear astroytes
tcell_seurat <- AddModuleScore(tcell_seurat, list(c("APOE", "SLC1A2", "SLC1A3", "ALDOC", "GFAP", "CLU", "GPM6B", "TUBA1A", "NGFRAP1", 
                                                      "CD9", "GFAP", "CKB", "PTGDS", "S100B", "STMN1")), 
                                ctrl = 10, name = c("astrocyte"))
FeaturePlot(tcell_seurat, c("astrocyte1"))
tcell_seurat$astrocyte1<- NULL


## NK- slight NK cluster 4, but likely "NK-like" as NK signature is not strong
tcell_seurat <- AddModuleScore(tcell_seurat, list(c("KIR2DL1", "KIR2DL3", "KIR2DS4")), 
                                ctrl = 10, name = c("NK"))
FeaturePlot(tcell_seurat, c("NK1"))
tcell_seurat$NK1<- NULL

```

## Identify any doublets- by NMF on all T cells
## NMF run in: /NMF/All_Tcells/NMF.Rmd
```{r}
## Read in cell assignments for rank6
cell_assignemnts<-readRDS(paste0("../../02a_Tcells/01a_Ped/01_NMF/analysis/DoubletID/", integration, "/", cohort,
                                 "/cell_assignment_rank6.Rds"))
tcell_seurat$NMF_allTcell_rank6<- plyr::mapvalues(colnames(tcell_seurat), names(cell_assignemnts), cell_assignemnts,
                                                  warn_missing = FALSE)

## For each cluster, determine whether each marker gene set is "expressed"
## "Expressed" = mean(ln(TP100K+1))>1
marker_genes<- list(tcell_markers= c("CD3D", "CD3E", "CD3G", "CD247"),
                    astrocyte_markers= c("GFAP","APOE"),
                    myeloid_markers= c("C1QB", "C1QC"),
                    NK_markers= c("KIR2DL1", "KIR2DL3", "KIR2DS4"))

mean_marker_expr_all<- lapply(unique(tcell_seurat$NMF_allTcell_rank6), function(x){
  print(x)
  ## Subset to NMF program
  seurat_tmp<- subset(tcell_seurat, NMF_allTcell_rank6==x)
  
  ## Cycle through each marker set, determining mean expression
  mean_marker_expr<- lapply(names(marker_genes), function(y_name){
    print(y_name); y<- marker_genes[[y_name]]
    goi_expr<- seurat_tmp@assays$RNA@counts[y,]
    goi_expr<- as.data.frame(t(goi_expr))
    goi_ln<- apply(goi_expr, 2, function(z){log(z/10 + 1)})
    goi_mean<- mean(rowMeans(goi_ln))
    goi_PF<- goi_mean>1
    return(goi_PF)
  })
  names(mean_marker_expr)<- names(marker_genes)
  mean_marker_expr<- as.data.frame(mean_marker_expr)
  rownames(mean_marker_expr)<- x
  print("------")
  return(mean_marker_expr)
})
mean_marker_expr_df<- do.call("rbind", mean_marker_expr_all)

## no integration results: one program expresses myeloid markers (and for some, also doesn't express t cell markers)
  ## NMF2 in pedOnly, NMF5 in pedOnly_nomuv63, NMF3 in pedOnly_nomuv63.withmuv91, NMF3 in allSamples
  ## No doublet clusters - all programs express either t cell or myeloid alone
  ## No NK programs
  ## For allSamples, "special" program expresses both tcell/myeloid. All others only express myeloid.
  ## For all cohorts, a subset of cells in "special" program express Tcell markers, a subset express myeloid

## Integration results: all programs express t cell markers, but one program (NMF2) also expresses myeloid markers
unclearNMF<- "2" ## 2 for pedOnly noInt, 5 for pedOnly_nomuv63 noInt, 2 for pedOnly_nomuv63 with Int

## Double check that NMF2/5/3 doesn't express CD3G + expresses CSF1R
VlnPlot(tcell_seurat, group.by="NMF_allTcell_rank6", features=c("CD3G", "CSF1R","C1QB", "C1QC"), ncol = 2)
ggsave(paste0(figure_dir, "TcellNMF_CD3.CSF1R_NMF5.is.myeloid.png"), width=10, height=8)

## Unclear NMF has very low (but not 0) CD3G expression. Are cells expressing both, or are there some Tcells/some myeloid in NMF2?
seurat_tmp<- subset(tcell_seurat,NMF_allTcell_rank6==unclearNMF ) 
marker_df<- as.data.frame(t(seurat_tmp@assays$RNA@counts[c("CSF1R", "CD3G", "CD3E", "CD3D"),]))
marker_df<- marker_df[rowSums(marker_df)!=0,] ## removes cells with 0 expression for all (but keeps a few (8) with very low expression, still will be removed downstream)
marker_df<- apply(marker_df, 2, function(x){log(x/10 + .1)})
marker_expr<- as.data.frame(apply(marker_df, 2, function(x){x>1}))
marker_expr$MyeloidExpr<- ifelse(marker_expr$CSF1R, "Myeloid", "NotMyeloid")
marker_expr$TcellExpr<- ifelse(marker_expr$CD3G | marker_expr$CD3E | marker_expr$CD3D, "Tcell", "NotTcell")
table(marker_expr$TcellExpr, marker_expr$MyeloidExpr)

## 4 cell subsets for unclear NMF: NEITHER expressed (34/34/7/6) / MYELOID expressed (40/29/3/24) / TCELL expressed (42/50/123/68) / BOTH expressed (12/14/21/16) (for x/x/x/allSamples)
## Will remove NEITHER/BOTH, add MYELOID/TCELL to their seurat 
neither_cells<- c(colnames(seurat_tmp)[!colnames(seurat_tmp)  %in% rownames(marker_df)],
                  rownames(marker_expr)[marker_expr$MyeloidExpr=="NotMyeloid" & marker_expr$TcellExpr == "NotTcell"])
both_cells<- rownames(marker_expr)[marker_expr$MyeloidExpr=="Myeloid" & marker_expr$TcellExpr == "Tcell"]
myeloid_cells<- rownames(marker_expr)[marker_expr$MyeloidExpr=="Myeloid" & marker_expr$TcellExpr == "NotTcell"]
tcell_cells<- rownames(marker_expr)[marker_expr$MyeloidExpr=="NotMyeloid" & marker_expr$TcellExpr == "Tcell"]

  ## Merge current myeloid cells with 40 new myeloid cells, run seurat pipeline
  new_myeloid<- subset(tcell_seurat, cells=myeloid_cells); new_myeloid$broad_annot<- "Myeloid"
  new_myeloid$re_annotation<- "TcellNMF"; myeloid_seurat$re_annotation<- "OriginallyMyeloid" ## for looking at where the "reannotated" cells end up in full myeloid
  sum(rownames(myeloid_seurat)==rownames(new_myeloid)); nrow(myeloid_seurat); nrow(new_myeloid)
  myeloid_cm<- cbind(myeloid_seurat@assays$RNA@counts,new_myeloid@assays$RNA@counts)
  myeloid_meta<- rbind(myeloid_seurat@meta.data[,colnames(myeloid_seurat@meta.data) %in%
                                                  colnames(new_myeloid@meta.data)],
                       new_myeloid@meta.data[,colnames(new_myeloid@meta.data) %in% colnames(myeloid_seurat@meta.data)])
  
  if(integration=="HarmonyInt/"){
    myeloid_seurat<- RunFullSeurat(myeloid_cm, RunHarmony = TRUE, samples =  myeloid_meta$sampleid)
  }else{
      myeloid_seurat<- RunFullSeurat_Immune(myeloid_cm, myeloid_meta$sampleid)
  }
  myeloid_seurat<- AddMetaData(myeloid_seurat, myeloid_meta[,!colnames(myeloid_meta) %in%
                                                              colnames(myeloid_seurat@meta.data)])

  ## Remove BOTH/NEITHER/MYELOID cells from tcell seurat, rerun seurat pipeline
  tcell_seurat<- subset(tcell_seurat, 
                      cells=colnames(tcell_seurat)[!colnames(tcell_seurat) %in% 
                                                     c(neither_cells, both_cells, myeloid_cells)])
  tcell_seurat$re_annotation<- ifelse(colnames(tcell_seurat) %in% tcell_cells, "TcellNMF",
                                      "OtherTcellNMF") 
  tcell_meta<- tcell_seurat@meta.data
  if(integration=="HarmonyInt/"){
    tcell_seurat<- RunFullSeurat(tcell_seurat@assays$RNA@counts, RunHarmony = TRUE,
                                 samples =  tcell_meta$sampleid)
  }else{
      tcell_seurat<- RunFullSeurat_Immune(tcell_seurat@assays$RNA@counts, tcell_meta$sampleid)
  }
  tcell_seurat<-AddMetaData(tcell_seurat, tcell_meta[,!colnames(tcell_meta) %in% colnames(tcell_seurat@meta.data)])

## Plot where "Tcell NMF2" cells end up after being merged with tcell/myeloid seurat objects
p1<-DimPlot(myeloid_seurat, group.by="re_annotation")+
  scale_color_manual(values=c(OriginallyMyeloid="grey", TcellNMF="red"))+
  ggtitle(paste0("Myeloid cells\nreannotated by Tcell NMF", unclearNMF))

p2<-DimPlot(tcell_seurat, group.by="re_annotation")+
  scale_color_manual(values=c(OtherTcellNMF="grey", TcellNMF="red"))+
  ggtitle(paste0("T cells\nTcell NMF", unclearNMF, "vs other NMF"))
p1+p2
ggsave(paste0(figure_dir, "Myeloid.Tcell_cellReannotatedByTcellNMF", unclearNMF, ".png"), width=12, height=5)

## Plot myeloid/tcell markers in myeloid/tcell, comparing NMF2 cells vs others
VlnPlot(myeloid_seurat, group.by="re_annotation", features="CSF1R")+NoLegend()
ggsave(paste0(figure_dir, "Myeloid_CSF1R_originalMyeloid_vs_reannotatedByTcellNMF", unclearNMF,".png"), 
       width=3, height=4)

VlnPlot(tcell_seurat, group.by = "re_annotation", features=c("CD3G", "CD3E", "CD3D"))+NoLegend()
ggsave(paste0(figure_dir, "Tcell_CD3G.E.D_TcellNMF", unclearNMF, "_vs_otherNMF.png"), width=9, height=4)
    
## Save updated myeloid/tcell
saveRDS(tcell_seurat, file=paste0(analysis_dir, "tcell_seurat.Rds"))
saveRDS(myeloid_seurat, file=paste0(analysis_dir, "myeloid_seurat.Rds"))
```

## Remove any cycling cells that express both CD4/CD8 OR neither CD4/CD8
```{r}
## Subset to cycling
cycling_seurat<-subset(tcell_seurat, cycling_predicted=="cycling")

## determine CD4/CD8 "expression": ln(TP100K+1)>1
cd4.8_expr<- as.data.frame(cycling_seurat@assays$RNA@counts)[c("CD4", "CD8A"),]
cd4.8_expr<- as.data.frame(t(cd4.8_expr))
cd4.8_ln<- apply(cd4.8_expr, 2, function(x){log(x/10 + 1)})
cd4.8_binary<- as.data.frame(cd4.8_expr>1)

## cells that express both/neither (9 cells)
express_both<- cd4.8_binary[cd4.8_binary$CD4 & cd4.8_binary$CD8A,]
express_neither<- cd4.8_binary[!cd4.8_binary$CD4 & !cd4.8_binary$CD8A,]
cells_remove<- c(rownames(express_both), rownames(express_neither))

## Remove from tcell_seurat
cells_keep<- colnames(tcell_seurat)[!colnames(tcell_seurat) %in% cells_remove]
tcell_seurat<-subset(tcell_seurat, cells=cells_keep)

## Rerun clustering with subsetted T cells
tcell_meta<- tcell_seurat@meta.data
if(integration=="HarmonyInt/"){
  tcell_seurat<- RunFullSeurat(tcell_seurat@assays$RNA@counts,RunHarmony = TRUE, samples=tcell_meta$sampleid)
}else{
  tcell_seurat<- RunFullSeurat_Immune(tcell_seurat@assays$RNA@counts, tcell_seurat$sampleid)
}
tcell_seurat<- AddMetaData(tcell_seurat, tcell_meta[,!colnames(tcell_meta) %in% colnames(tcell_seurat@meta.data)])

saveRDS(tcell_seurat, file=paste0(analysis_dir, "tcell_seurat.Rds"))
```

## For full tcell/myeloid seurat: remove BOTH/NEITHER cells, fix broad annotation, rerun pipeline, and save
```{r}
immune_seurat<- subset(immune_seurat, cells=c(colnames(tcell_seurat), colnames(myeloid_seurat)))
immune_meta<- immune_seurat@meta.data
if(integration=="HarmonyInt/"){
  immune_seurat<- RunFullSeurat(immune_seurat@assays$RNA@counts, RunHarmony=TRUE, samples=immune_seurat$sampleid)
}else{
  immune_seurat<- RunFullSeurat_Immune(immune_seurat@assays$RNA@counts, immune_seurat$sampleid)
}
immune_seurat<- AddMetaData(immune_seurat, immune_meta[,!colnames(immune_meta) %in% colnames(immune_seurat@meta.data)])

## fix broad annotation
updated_annot<- c(tcell_seurat$broad_annot, myeloid_seurat$broad_annot)
names(updated_annot)<- c(colnames(tcell_seurat),colnames(myeloid_seurat))
immune_seurat$broad_annot<- plyr::mapvalues(colnames(immune_seurat), names(updated_annot), updated_annot)

saveRDS(immune_seurat, file=paste0(analysis_dir, "tcell.myeloid_seurat.Rds"))
```


## Compare how well "with integration" and "without integration" annotations match
```{r}
cohort<- "pedOnly_nomuv63"

## Read in seurat objects
tcell_harmony<- readRDS(paste0("analysis/broad_annotation/HarmonyInt/", cohort, "/tcell_seurat.Rds"))
tcell_noInt<- readRDS(paste0("analysis/broad_annotation/", cohort, "/tcell_seurat.Rds"))

myeloid_harmony<- readRDS(paste0("analysis/broad_annotation/HarmonyInt/", cohort, "/myeloid_seurat.Rds"))
myeloid_noInt<- readRDS(paste0("analysis/broad_annotation/", cohort, "/myeloid_seurat.Rds"))

## Read in Tcell NMF annotations (from no integration)
cell_assignemnts<-readRDS(paste0("../../NMF/All_Tcells/analysis/DoubletID/", cohort, "/cell_assignment_rank6.Rds"))

## Tcells
  ## Number of tcells in each, number that match
  print(paste0("Tcells with integration: ", ncol(tcell_harmony))) ## 1785
  print(paste0("Tcells without integration: ", ncol(tcell_noInt))) ## 1724
  print(paste0("Shared Tcells: ", sum(colnames(tcell_harmony) %in% colnames(tcell_noInt)))) ## 1688
  
  ## Where do mismatch cells fall on umap? Do they express CD3?
  misMatchTcells<- colnames(tcell_harmony)[!colnames(tcell_harmony) %in% colnames(tcell_noInt)]
  tcell_harmony$NotTcellInOriginal<- ifelse(colnames(tcell_harmony) %in% misMatchTcells, "NotInOriginal",
                                              "InOriginal")
  p1<-DimPlot(tcell_harmony, group.by = "NotTcellInOriginal")+
    scale_color_manual(values=c(InOriginal="grey", NotInOriginal="red"))
  p2<- VlnPlot(tcell_harmony, group.by = "NotTcellInOriginal", features=c("CD3G", "CSF1R"))
  p1+p2
  ggsave(paste0(figure_dir, "UMAP_Tcells_InOriginalWithoutIntegration_butNotInCurrent.png"), width=12, height=5)
  
## Myeloid
  ## Number of myeloids in each, number that match
  print(paste0("myeloid with integration: ", ncol(myeloid_harmony))) ## 1080
  print(paste0("myeloid without integration: ", ncol(myeloid_noInt))) ## 1043
  print(paste0("Shared myeloids: ", sum(colnames(myeloid_harmony) %in% colnames(myeloid_noInt)))) ## 1027
  
  ## Where do mismatch cells fall on umap? Do they express CD3?
  misMatchmyeloids<- colnames(myeloid_harmony)[!colnames(myeloid_harmony) %in% colnames(myeloid_noInt)]
  myeloid_harmony$NotMyeloidInOriginal<- ifelse(colnames(myeloid_harmony) %in% misMatchmyeloids, "NotInOriginal",
                                              "InOriginal")
  p1<-DimPlot(myeloid_harmony, group.by = "NotMyeloidInOriginal")+
    scale_color_manual(values=c(InOriginal="grey", NotInOriginal="red"))
  p2<- VlnPlot(myeloid_harmony, group.by = "NotMyeloidInOriginal", features=c("CD3G", "CSF1R"))
  p1+p2
  ggsave(paste0(figure_dir, "UMAP_myeloid_InOriginalWithoutIntegration_butNotInCurrent.png"), width=12, height=5)
  
  ## From NMF for harmony integration, seem to have gained cycling cells? Do cells found in new but not in old score highly for cycling?
  ## seems to not be the case
  mismatchMyeloid_seurat<-subset(myeloid_harmony, cells=misMatchmyeloids)
  table(mismatchMyeloid_seurat$cycling_predicted)

```

