---
title: "DetailedAnnotation"
author: "Jenna LaBelle"
date: "11/04/21"
output: html_document
---

## Purpose:
Assign more detailed annotations to myeloid/tcells (CD4/CD8, DC/BCells). HOWEVER- CD4/CD8 annotations generated here are NOT used downstream. These annotations are done later, on program-wise basis.

## Approach:
Visualize marker genes for detailed cell types, assign on cluster-wise basis. For CD4/CD8 (again, these annotations not actually used downstream), sub-clustering and cell-wise annotation is required for some clusters/cells. Also have the option to use harmony or non-harmony integrated seurat objects.

```{r}
library(Seurat) 

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/01_Preprocessing/02_Counts_Preprocess/")

integration<- "" ## HarmonyInt/ or ""
analysis_dir<- paste0(working_dir, "/analysis/detailed_annotation/", integration, "/")
figure_dir<- paste0(working_dir, "/figures/detailed_annotation/", integration, "/")
data_dir<- paste0("../Counts_Preprocess/analysis/broad_annotation/", integration, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in processed seurat objects
```{r}
cohort<- "pedOnly_nomuv63"

analysis_dir<- paste0(analysis_dir, cohort, "/")
figure_dir<- paste0(figure_dir, cohort, "/")
for(x in c(analysis_dir, figure_dir)){if(!dir.exists(x)){dir.create(x)}}
data_dir<- paste0(data_dir, cohort, "/")

immune_seurat<- readRDS(paste0(data_dir,"tcell.myeloid_seurat.Rds"))
tcell_seurat<- readRDS(paste0(data_dir,   "tcell_seurat.Rds"))
myeloid_seurat<- readRDS(paste0(data_dir, "myeloid_seurat.Rds"))

```


#############################
## Myeloid cell annotation ##
#############################

For now, just identifying B cells and dendritic cells
May also identify/annotation macrophages and microglia here
```{r}
## for harmony integration, pedOnly_nomuv63: need to increase resolution in order to split DC cluster
myeloid_seurat<- FindClusters(myeloid_seurat, resolution = 1)

bcell_markers<- c("CD19", "MS4A1")
dendritic_markers<- c("FLT3", "CD207", "CD1A", "CD1C", "FCER1A",
                      "NOTCH2")

FeaturePlot(myeloid_seurat, features=bcell_markers, ncol = 2)
ggsave(paste0(figure_dir, "Myeloid_BcellMarkers.png"), width=8, height=4)

FeaturePlot(myeloid_seurat, features=dendritic_markers, ncol = 3)
ggsave(paste0(figure_dir, "Myeloid_DendriticMarkers.png"), width=12, height=8)

## annotate clusters based on these markers
#cluster_annot<- c(9, 10) ## pedOnly
#cluster_annot<-c(7,8) ## pedOnly_nomuv63
#cluster_annot<- c(9,8) ## pedOnly_nomuv63.withmuv91
##cluster_annot<- c(8, 6) ## pedOnly_nomuv63, with harmony integration
cluster_annot<- c(8,7) ## allSamples
names(cluster_annot)<- c("DC", "BCell")
myeloid_seurat$detailed_annot<- ifelse(myeloid_seurat$seurat_clusters %in% cluster_annot,
                                            plyr::mapvalues(as.character(myeloid_seurat$seurat_clusters), 
                                                            cluster_annot, names(cluster_annot)),
                                            "Myeloid")
saveRDS(myeloid_seurat, file=paste0(analysis_dir, "myeloid_seurat.Rds"))

## Remove DC/Bcells and rerun seurat
myeloid_seurat_noDC.B<- subset(myeloid_seurat, detailed_annot=="Myeloid")
myeloid_seurat_noDC.B_meta<- myeloid_seurat_noDC.B@meta.data

if(integration=="HarmonyInt/"){
 myeloid_seurat_noDC.B<- RunFullSeurat(myeloid_seurat_noDC.B@assays$RNA@counts,RunHarmony = TRUE,
                                             samples=myeloid_seurat_noDC.B$sampleid) 
}else{
  myeloid_seurat_noDC.B<- RunFullSeurat_Immune(myeloid_seurat_noDC.B@assays$RNA@counts,
                                             myeloid_seurat_noDC.B$sampleid)
}

myeloid_seurat_noDC.B<- AddMetaData(myeloid_seurat_noDC.B, 
                                    myeloid_seurat_noDC.B_meta[,!colnames(myeloid_seurat_noDC.B_meta) %in%
                                                                 colnames(myeloid_seurat_noDC.B@meta.data)])
saveRDS(myeloid_seurat_noDC.B, file=paste0(analysis_dir, "myeloid_seurat_noDC.B.Rds"))

```

############################
## Identify CD4/CD8 cells ##
############################


```{r}
## Decrease resolution
if(integration=="HarmonyInt/"){
  tcell_seurat<- FindClusters(tcell_seurat, resolution = .6)
}else{
  tcell_seurat<- FindClusters(tcell_seurat,  resolution = .7, algorithm = 4, random.seed = 100) ## .6 res used for pedOnly, .7 for pedOnly_nomuv63. Same cluster number
}
FeaturePlot(tcell_seurat, features=c("CD3G", "CD4", "CD8A", "CD8B"), ncol = 2)
ggsave(paste0(figure_dir, "UMAP_Tcell_CD3.4.8.png"), width=6, height=6)

DimPlot(tcell_seurat, label=TRUE)+NoLegend()
VlnPlot(tcell_seurat, features= c("CD3G", "CD4", "CD8A", "CD8B"), ncol=4)
ggsave(paste0(figure_dir, "Vln_Tcell_CD3.4.8.png"), width=12, height=3)

## 6 is cycling, 1/4 are cd8, 2 is cd4, 3/5 are not clear (For pedOnly AND pedOnly_nomuv63)
## 6 is cycling, 1 is CD4, 0/3/5 are CD8, 2/4 are unclear (for harmony + pedOnly_nomuv63)
```

## Tcells in some clusters are not clearly cd4/cd8
## Further analysis/plotting to determine annotation for these cells
```{r}
unclear_clusters<- c(2,4)

## Marker genes for unclearclusters
all_markers<- FindAllMarkers(tcell_seurat, min.pct = 0.2)
all_markers_filt<- all_markers[all_markers$p_val_adj<0.05 & all_markers$avg_log2FC>0.5,]
write.csv(all_markers_filt, file=paste0(analysis_dir, "Tcells_seuratClusters_markerGenes_Cluster", 
                                        paste(unclear_clusters, collapse="."),
                                        "Unclear.csv"))

unclearCluster_markers<- all_markers_filt[all_markers_filt$cluster %in% unclear_clusters,]
unclearCluster_markers<- unclearCluster_markers[order(unclearCluster_markers$avg_log2FC, decreasing = TRUE),]
p1<-DoHeatmap(tcell_seurat, features=unclearCluster_markers$gene[1:35])
p2<-VlnPlot(tcell_seurat, features=c("CD4", "CD8A", "CD3G"))
p1+p2
ggsave(paste0(figure_dir, "Tcells_seuratClusters_cluster", paste(unclear_clusters,
                                                                 collapse="."),"Markers_UnclearCD4.CD8.Cluster.png"),
       width=10, height=5)

## Subset to just unclear clusters, Further  plotting to help annotate
unclearAnnot_seurat<- subset(tcell_seurat, seurat_clusters %in% unclear_clusters) 
FeaturePlot(unclearAnnot_seurat, features=c("CD4", "CD8A", "CD3G"), ncol = 3)
FeaturePlot(unclearAnnot_seurat, 
            features=c("CD3D", "CD4", "IL7R", "CD69", "CD40LG", "KLRB1", "ZBTB16", "PTGER2", "PTGER4"), 
                    pt.size = 1)


## assign annotations cell-wise: by highest expressort
## issue with this approach: some cells don't express EITHER CD4 or CD8
cd4.8_expr<- as.data.frame(t(unclearAnnot_seurat@assays$RNA@data[c("CD4", "CD8A"),]))
cd4.8_expr$MaxExpr<- apply(cd4.8_expr, 1, function(x){names(x)[which.max(x)]})
cd4.8_expr$MaxExpr<- ifelse(cd4.8_expr$CD4==0 & cd4.8_expr$CD8A==0, "Neither", cd4.8_expr$MaxExpr)
unclearAnnot_seurat$MaxCD4.CD8<- plyr::mapvalues(colnames(unclearAnnot_seurat),
                                                 rownames(cd4.8_expr), cd4.8_expr$MaxExpr)
DimPlot(unclearAnnot_seurat, group.by = "MaxCD4.CD8")+scale_color_manual(values=c(CD4="red", CD8A="green", Neither="grey"))

## plot CD4/CD8 expression on single scale
## negative values = more CD4 / positive = more CD8 / zero = expresses both CD4/CD8
tmp_seurat<-subset(unclearAnnot_seurat, MaxCD4.CD8 != "Neither")
cd4.8_expr<- as.data.frame(t(tmp_seurat@assays$RNA@data[c("CD4", "CD8A"),]))
cd4.8_expr$CD4<- cd4.8_expr$CD4 * -1
cd4.8_expr$SlidingScale<- cd4.8_expr$CD4 + cd4.8_expr$CD8A
tmp_seurat$Sliding_CD4.CD8_Scale<- as.numeric(plyr::mapvalues(colnames(tmp_seurat),
                                                                       rownames(cd4.8_expr),
                                                                       cd4.8_expr$SlidingScale))
FeaturePlot(tmp_seurat, features = "Sliding_CD4.CD8_Scale")+
  scale_color_gradient2(low="red", mid="grey", high="green", midpoint=0)

## rerun seurat, see if cells cluster by CD4/CD8 expression then

  ## Seurat pipeline function requires >1 cell per sample or it will break
  ## 2 samples have only 1 cell. Temporarily merge these 2 cells into another random sample, then rename after
  unclearAnnot_meta<- unclearAnnot_seurat@meta.data
  unclearAnnot_meta$originalSample<- unclearAnnot_meta$sampleid
  nSamples<- table(unclearAnnot_meta$sampleid)
  unclearAnnot_meta$sampleid<- ifelse(unclearAnnot_meta$sampleid %in% names(nSamples)[nSamples==1],
                                      names(nSamples[nSamples>1][1]), unclearAnnot_meta$sampleid)

if(integration=="HarmonyInt/"){
  unclearAnnot_rerun<- RunFullSeurat(unclearAnnot_seurat@assays$RNA@counts,RunHarmony = TRUE, 
                                     samples=unclearAnnot_meta$sampleid)
}  else{
  unclearAnnot_rerun<- RunFullSeurat_Immune(unclearAnnot_seurat@assays$RNA@counts, unclearAnnot_meta$sampleid)
}

unclearAnnot_rerun<- AddMetaData(unclearAnnot_rerun, 
                                 metadata=unclearAnnot_meta[,!colnames(unclearAnnot_meta) %in%
                                                              colnames(unclearAnnot_rerun@meta.data) ])
unclearAnnot_rerun$sampleid<- unclearAnnot_rerun$originalSample

if(integration=="HarmonyInt/"){
  unclearAnnot_rerun<- FindClusters(unclearAnnot_rerun, resolution = 2.75)
}else{
  unclearAnnot_rerun<- FindClusters(unclearAnnot_rerun,  resolution =.85, algorithm = 4, random.seed = 100) # res 1 for pedOnly, res1 for pedOnly_nomuv63.withmuv91
}


p1<- DimPlot(unclearAnnot_rerun, label=TRUE)+NoLegend()
p2<- FeaturePlot(unclearAnnot_rerun, features=c("CD4", "CD8A"), ncol = 2)
p3<-VlnPlot(unclearAnnot_rerun, features=c("CD4", "CD8A"))
p1+p3
ggsave(paste0(figure_dir, "Tcells_unclearCluster", paste(unclear_clusters,
                                                                 collapse="."),
              "_rerunSeurat_CD4.CD8Expr_res.2.75.png"), width=10, height=4)


  ## Clusters 2/4 are CD4, Clusters 1 are CD8, cluster 3 is neither? (for pedOnly)
  ## clusters 1/3 are CD4, clusters 2 is CD8, cluster 4 is unclear (for pedOnly_nomuv63)
  ## clusters 3/5 are CD4, clusters 1/4 are CD8 (for pedOnly_nomuv63.withmuv91), cluster 2 is unclear
  ## clusters 4/5/10/12 are CD8 clusters 0/1/2/3/6/7/9/13 are CD4, clusters 8/11 are unclear

  ## for cluster 3 (pedOnly): no marker genes. Remove these cells (114 cells)
  ## for cluster 3 (pedOnly_nomuv63): there are only 3 marker genes. CD8A/B is marker for cluster 2. remove cells (63 cells)
  ## for cluster 2 (pedOnly_nomuv63.withmuv91): only 1 marker gene, ribosomal
  all_markers<- FindAllMarkers(unclearAnnot_rerun, min.pct = 0.2)
  all_markers_filtered<- all_markers[all_markers$p_val_adj<0.1 & all_markers$avg_log2FC>0,]
  unclear_cluster<- subset(unclearAnnot_rerun, seurat_clusters %in% c(8,11))
  
  ## Assign CD4/CD8 by subcluster- 
  cd4_clusters<- c(0,1,2,3,6,7,9,13); cd8_clusters<- c(4,5,10,12)
  unclearAnnot_rerun$MaxCD4.CD8<- ifelse(unclearAnnot_rerun$seurat_clusters %in% cd4_clusters, "CD4", 
                                        ifelse(unclearAnnot_rerun$seurat_clusters %in% cd8_clusters, "CD8", 
                                                "Unclear_Removed"))
  

## To annotate:
## By clustering for clear cd4/cd8 clusters
## By clustering for unclear cluster- rerun pipeline, assign cd4/cd8 by cluster
## Within unclear cluster, one sub-cluster (3 in pedOnly, 4 in pedOnly_nomub63) is still unclear. will remove these cells.
unclearCluster_Annotations<- unclearAnnot_rerun$MaxCD4.CD8
names(unclearCluster_Annotations)<- colnames(unclearAnnot_rerun)

p2
```

## Alternative to the above: score cells for multiple genes, instead of just CD4/CD8
## Doesn't clarify annotation, not used downstream
```{r}
cd4.8_genes<- list(CD8=c("CD8A", "CD8B"), CD4=c("CD4", "IL7R", "CD69", "CD40LG", "KLRB1", "ZBTB16"))

tcell_seurat<-AddModuleScore(tcell_seurat, features=list(cd4.8_genes[["CD8"]]), name="CD8_Score__")
tcell_seurat<-AddModuleScore(tcell_seurat, features=list(cd4.8_genes[["CD4"]]), name="CD4_Score__")
colnames(tcell_seurat@meta.data)<- gsub("__1", "", colnames(tcell_seurat@meta.data))

DotPlot(tcell_seurat, features=c("CD8_Score", "CD4_Score"))
FeaturePlot(tcell_seurat, features=c("CD8_Score", "CD4_Score"))
```


## For cycling: cycling drives clustering, but still are underlying cd4/cd8
## Assign as CD4/CD8 based on max expression
```{r}
cycling_seurat<- subset(tcell_seurat, seurat_clusters =="6")
cd4.cd8_expr<- as.data.frame(t(cycling_seurat@assays$RNA@counts[c("CD4", "CD8A"),]))
cd4.cd8_expr$Max<- apply(cd4.cd8_expr, 1, function(x){names(x)[which.max(x)]})
cycling_annotation<- cd4.cd8_expr$Max; names(cycling_annotation)<- colnames(cycling_seurat)

```

## Add final CD4/CD8 annotations
## Based on: marker expression in "clear" clusters + reclustering of "unclear" + reclustering of cycling
```{r}
## clear cd4/cd8 clusters
cd4_clusters<- c(1)
cd8_clusters<- c(0,3,5) ## 1,4 for pedOnly_nomuv63

## For 2 unclear clusters + cycling
cd4_cells<-c(names(unclearCluster_Annotations)[unclearCluster_Annotations=="CD4"],
             names(cycling_annotation)[cycling_annotation=="CD4"])
cd8_cells<-c(names(unclearCluster_Annotations)[unclearCluster_Annotations=="CD8"],
             names(cycling_annotation)[cycling_annotation=="CD8A"])
unclear_cells<- c(names(unclearCluster_Annotations)[unclearCluster_Annotations=="Unclear_Removed"])

## If cell is in "clear" cluster, use that annotation. If not, use cell-wise annotation
tcell_seurat$detailed_annot<- ifelse(tcell_seurat$seurat_clusters %in% cd4_clusters, "CD4",
                          ifelse(tcell_seurat$seurat_clusters %in% cd8_clusters, "CD8",
                          ifelse(colnames(tcell_seurat) %in% cd4_cells, "CD4",
                          ifelse(colnames(tcell_seurat) %in% cd8_cells, "CD8", 
                          ifelse(colnames(tcell_seurat) %in% unclear_cells, "Unclear", "issue")))))

## Add CD4/CD8/Cycling annotation
  tcell_seurat$detailed_annot_withCycling<- ifelse(tcell_seurat$cycling_predicted=="cycling", "Cycling",
                                               tcell_seurat$detailed_annot)
  
## Remove "Unclear" (114 cells for pedOnly, 63 for pedOnly_nomuv63, 88 for pedOnly_nomuv63.withmuv91, 57 in harmonyInt)
tcell_seurat<- subset(tcell_seurat, detailed_annot != "Unclear")

## Confirm that each cluster expresses CD3 + EITHER CD4 or CD8 (or at least one is clearly higher)
tcell_seurat$cd4.8_cluster<-paste0("cluster", tcell_seurat$seurat_clusters, "_", tcell_seurat$detailed_annot)
DotPlot(tcell_seurat, group.by="cd4.8_cluster", features=c("CD3G", "CD4", "CD8A"))
ggsave(paste0(figure_dir, "CD4.CD8_finalClusterAnnotations_expression.png"), width=5, height=5)
tcell_seurat$cd4.8_cluster<- NULL

## Rerun pipeline on finalized set of cells
tcell_meta<- tcell_seurat@meta.data
if(integration=="HarmonyInt/"){
 tcell_seurat<- RunFullSeurat(tcell_seurat@assays$RNA@counts,RunHarmony = TRUE, samples= tcell_seurat$sampleid) 
}else{
  tcell_seurat<- RunFullSeurat_Immune(tcell_seurat@assays$RNA@counts, tcell_seurat$sampleid)
}
tcell_seurat<- AddMetaData(tcell_seurat, tcell_meta[,!colnames(tcell_meta) %in% colnames(tcell_seurat@meta.data)])


saveRDS(tcell_seurat, file=paste0(analysis_dir, "tcell_seurat.Rds"))
```

## Split Tcell by CD4/CD8, rerun seurat pipeline for each separately
```{r}
cd4_seurat<- subset(tcell_seurat,detailed_annot=="CD4")
cd8_seurat<- subset(tcell_seurat, detailed_annot=="CD8")

cd4_meta<- cd4_seurat@meta.data; cd8_meta<- cd8_seurat@meta.data

if(integration=="HarmonyInt/"){
  cd4_seurat<- RunFullSeurat(cd4_seurat@assays$RNA@counts,RunHarmony = TRUE, samples= cd4_seurat$sampleid)
  cd8_seurat<- RunFullSeurat(cd8_seurat@assays$RNA@counts,RunHarmony = TRUE, samples=  cd8_seurat$sampleid)
}else{
  cd4_seurat<- RunFullSeurat_Immune(cd4_seurat@assays$RNA@counts, cd4_seurat$sampleid)
  cd8_seurat<- RunFullSeurat_Immune(cd8_seurat@assays$RNA@counts, cd8_seurat$sampleid)
}

cd4_seurat<- AddMetaData(cd4_seurat, metadata=cd4_meta[,!colnames(cd4_meta) %in% colnames(cd4_seurat@meta.data)])
cd8_seurat<- AddMetaData(cd8_seurat, metadata=cd8_meta[,!colnames(cd8_meta) %in% colnames(cd8_seurat@meta.data)])

saveRDS(cd4_seurat, file=paste0(analysis_dir, "tcell_cd4_seurat.Rds"))
saveRDS(cd8_seurat, file=paste0(analysis_dir, "tcell_cd8_seurat.Rds"))
```


## Add on CD4/CD8/DC/Bcell annotation to Tcell+Myeloid seurat obj
```{r}
## Add on CD4/CD8
immune_seurat$detailed_annot<- ifelse(colnames(immune_seurat) %in% colnames(cd4_seurat), "CD4", 
                                               ifelse(colnames(immune_seurat) %in% colnames(cd8_seurat), "CD8",
                                                      immune_seurat$broad_annot))
immune_seurat$detailed_annot<- gsub("Tcell", "Tcell_UnclearCD4.CD8", immune_seurat$detailed_annot)

## Add on DC/Bcell
immune_seurat$detailed_annot<- ifelse(colnames(immune_seurat) %in% colnames(myeloid_seurat),
                                      plyr::mapvalues(colnames(immune_seurat), 
                                                      colnames(myeloid_seurat), myeloid_seurat$detailed_annot),
                                      immune_seurat$detailed_annot)


saveRDS(immune_seurat, file=paste0(analysis_dir, "tcell.myeloid_seurat.Rds"))
```


## For non-harmony integrated ONLY: integrate with harmony, then add on non-harmony-integrated annotations
This only performed for tcells, tcell+myeloid, myeloid+DC+Bcells- not for CD8/CD4/Myeloid without DC/Bcells
Similar analysis performed for cd4/cd8/myeloid without DC,Bcells in their NMF folders- after adding all NMF annotations, etc
```{r}
integration<- "" 
cohort<- "pedOnly_nomuv63"
analysis_dir<- paste0(working_dir, "/analysis/detailed_annotation/", integration, "/")
analysis_dir<- paste0(analysis_dir, cohort, "/")

## Read in unintegrated seurat objects
myeloid_seurat<- readRDS(paste0(analysis_dir, "myeloid_seurat.Rds"))
tcell_seurat<- readRDS(paste0(analysis_dir, "tcell_seurat.Rds"))
immune_seurat<- readRDS(paste0(analysis_dir, "tcell.myeloid_seurat.Rds"))


## Run harmony integration
  ## Tcells 
  tcell_meta<- tcell_seurat@meta.data
  tcell_seurat<- RunFullSeurat(tcell_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=tcell_seurat$sampleid)
  tcell_seurat<- AddMetaData(tcell_seurat, metadata = tcell_meta[,!colnames(tcell_meta) %in% colnames(tcell_seurat@meta.data)])
  
  ## myeloid
  myeloid_meta<- myeloid_seurat@meta.data
  myeloid_seurat<- RunFullSeurat(myeloid_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=myeloid_seurat$sampleid)
  myeloid_seurat<- AddMetaData(myeloid_seurat, metadata = myeloid_meta[,!colnames(myeloid_meta) %in% colnames(myeloid_seurat@meta.data)])
  
  ## immune (all)
  immune_meta<- immune_seurat@meta.data
  immune_seurat<- RunFullSeurat(immune_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=immune_seurat$sampleid)
  immune_seurat<- AddMetaData(immune_seurat, metadata =immune_meta[,!colnames(immune_meta) %in% colnames(immune_seurat@meta.data)])

saveRDS(tcell_seurat, file=paste0(analysis_dir, "tcell_seurat_harmony.Rds"))
saveRDS(myeloid_seurat, file=paste0(analysis_dir, "myeloid_seurat_harmony.Rds"))
saveRDS(immune_seurat, file=paste0(analysis_dir, "tcell.myeloid_seurat_harmony.Rds"))
```


## Compare how well "with integration" and "without integration" annotations match
```{r}
cohort<- "pedOnly_nomuv63"

## Read in seurat objects
cd4_harmony<- readRDS(paste0("analysis/detailed_annotation/HarmonyInt/", cohort, "/tcell_cd4_seurat.Rds"))
cd4_noInt<- readRDS(paste0("analysis/detailed_annotation/", cohort, "/tcell_cd4_seurat.Rds"))

cd8_harmony<- readRDS(paste0("analysis/detailed_annotation/HarmonyInt/", cohort, "/tcell_cd8_seurat.Rds"))
cd8_noInt<- readRDS(paste0("analysis/detailed_annotation/", cohort, "/tcell_cd8_seurat.Rds"))


## CD4
  ## Number of tcells in each, number that match
  print(paste0("CD4 with integration: ", ncol(cd4_harmony))) ## 683
  print(paste0("CD4 without integration: ", ncol(cd4_noInt))) ## 701
  print(paste0("Shared CD4: ", sum(colnames(cd4_harmony) %in% colnames(cd4_noInt)))) ## 531
  
  ## Where do mismatch cells fall on umap? Do they express CD4?
  misMatchCD4<- colnames(cd4_harmony)[!colnames(cd4_harmony) %in% colnames(cd4_noInt)]
  cd4_harmony$NotCD4InOriginal<- ifelse(colnames(cd4_harmony) %in% misMatchCD4, "NotInOriginal",
                                              "InOriginal")
  p1<-DimPlot(cd4_harmony, group.by = "NotCD4InOriginal")+
    scale_color_manual(values=c(InOriginal="grey", NotInOriginal="red"))
  p2<- VlnPlot(cd4_harmony, group.by = "NotCD4InOriginal", features=c("CD4", "CD8A"))
  p1+p2
  ggsave(paste0(figure_dir, "UMAP_CD4_InOriginalWithoutIntegration_butNotInCurrent.png"), width=12, height=5)
  
## CD8
  ## Number of tcells in each, number that match
  print(paste0("CD8 with integration: ", ncol(cd8_harmony))) ## 1045
  print(paste0("CD8 without integration: ", ncol(cd8_noInt))) ## 960
  print(paste0("Shared CD8: ", sum(colnames(cd8_harmony) %in% colnames(cd8_noInt)))) ## 843
  
  ## Where do mismatch cells fall on umap? Do they express cd8?
  misMatchcd8<- colnames(cd8_harmony)[!colnames(cd8_harmony) %in% colnames(cd8_noInt)]
  cd8_harmony$NotCD8InOriginal<- ifelse(colnames(cd8_harmony) %in% misMatchcd8, "NotInOriginal",
                                              "InOriginal")
  p1<-DimPlot(cd8_harmony, group.by = "NotCD8InOriginal")+
    scale_color_manual(values=c(InOriginal="grey", NotInOriginal="red"))
  p2<- VlnPlot(cd8_harmony, group.by = "NotCD8InOriginal", features=c("CD4", "CD8A"))
  p1+p2
  ggsave(paste0(figure_dir, "UMAP_CD8_InOriginalWithoutIntegration_butNotInCurrent.png"), width=12, height=5)

```