---
title: "MergeSamples_QC_ToSeurat.Rmd"
author: "Jenna LaBelle"
date: "11/04/21"
output: html_document
---


```{r}
library(Seurat) 

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/Preprocessing/Counts_Preprocess/")

analysis_dir<- paste0(working_dir, "/analysis/compareSeuratPipelines/")
figure_dir<- paste0(working_dir, "/figures/compareSeuratPipelines/")
data_dir<- "../Counts_Preprocess/analysis/broad_annotation/"

if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in processed seurat objects
```{r}
cohort<- "pedOnly_nomuv63.withmuv91"

data_dir<- paste0(data_dir, cohort, "/")

immune_seurat<- readRDS(paste0(data_dir,"tcell.myeloid_seurat.Rds"))
tcell_seurat<- readRDS(paste0(data_dir,   "tcell_seurat.Rds"))
myeloid_seurat<- readRDS(paste0(data_dir, "myeloid_seurat.Rds"))

```

```{r}
cm=tcell_seurat@assays$RNA@counts
project=""
samples<- tcell_seurat$sampleid
meta<- tcell_seurat@meta.data
```

## Our standard pipeline, without integration
```{r}
## Create Seurat Obj
seurat_obj = CreateSeuratObject(cm, min.cells=0, min.features = 0,  project=project)

## Normalize data
seurat_obj <- NormalizeData(
  object = seurat_obj,
  normalization.method = "LogNormalize",
  scale.factor = 1E5
)

## Detection of variable genes across the single cells
seurat_obj <- FindVariableFeatures(
  object = seurat_obj,
  mean.function = ExpMean, 
  dispersion.function = LogVMR 
)

## Scaling the data and removing unwanted sources of variation
seurat_obj <- ScaleData(
  object = seurat_obj,
  do.scale = FALSE,
  do.center = TRUE
)

seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)
  

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = "pca", dims = 1:20)
seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = "pca",
                            dims = 1:20,
                            force.recalc = TRUE)  
seurat_obj<- FindClusters(seurat_obj, resolution = .8)
seurat_obj<- AddMetaData(seurat_obj, meta[,!colnames(meta) %in% colnames(seurat_obj@meta.data)])    

seurat_ourStandard_noInt<- seurat_obj

## Add cd4/cd8 score- subtract cd4 expr from cd8
cd4.cd8_expr<- as.data.frame(t(seurat_ourStandard_noInt@assays$RNA@data[c("CD4", "CD8A"),]))
cd4.cd8_expr$score<- cd4.cd8_expr$CD4 - cd4.cd8_expr$CD8A
seurat_ourStandard_noInt$cd4.cd8_score<- as.numeric(as.character(plyr::mapvalues(colnames(seurat_ourStandard_noInt),
                                                         rownames(cd4.cd8_expr),
                                                         cd4.cd8_expr$score)))

p1<- DimPlot(seurat_ourStandard_noInt)
p2<- DimPlot(seurat_ourStandard_noInt, group.by = "detailed_annot")+
  scale_color_manual(values=c(CD4="navy", 
                              CD8="cadetblue3"))
p3<- DimPlot(seurat_ourStandard_noInt, group.by="sampleid")+NoLegend()
p4<- FeaturePlot(seurat_ourStandard_noInt, features = "cd4.cd8_score")+scale_color_gradient2(low="blue", mid="white", high="red")
p1+p2+p3+p4
ggsave(paste0(figure_dir, "OurStandardPipeline_notInt_Tcells.png"), width=10, height=9)


```

## Our standard pipeline, with integration
```{r}
library(harmony)

## Create Seurat Obj
seurat_obj = CreateSeuratObject(cm, min.cells=0, min.features = 0,  project=project)

## Normalize data
seurat_obj <- NormalizeData(
  object = seurat_obj,
  normalization.method = "LogNormalize",
  scale.factor = 1E5
)

## Detection of variable genes across the single cells
seurat_obj <- FindVariableFeatures(
  object = seurat_obj,
  mean.function = ExpMean, 
  dispersion.function = LogVMR 
)

## Scaling the data and removing unwanted sources of variation
seurat_obj <- ScaleData(
  object = seurat_obj,
  do.scale = FALSE,
  do.center = TRUE
)

seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)
  
seurat_obj<- AddMetaData(seurat_obj, meta[,!colnames(meta) %in% colnames(seurat_obj@meta.data)])    
seurat_obj = RunHarmony(seurat_obj, "sampleid", theta = 2, 
                                     max.iter.harmony = 50, plot_convergence = TRUE)
seurat_obj <- RunUMAP(seurat_obj, reduction = "harmony", dims = 1:20)
seurat_obj <- FindNeighbors(seurat_obj,
                          reduction = "harmony",
                          dims = 1:20,
                          force.recalc = TRUE)   
seurat_obj<- FindClusters(seurat_obj, resolution = .8)

seurat_ourStandard_harmony<- seurat_obj
  
## Add cd4/cd8 score- subtract cd4 expr from cd8
cd4.cd8_expr<- as.data.frame(t(seurat_ourStandard_harmony@assays$RNA@data[c("CD4", "CD8A"),]))
cd4.cd8_expr$score<- cd4.cd8_expr$CD4 - cd4.cd8_expr$CD8A
seurat_ourStandard_harmony$cd4.cd8_score<- as.numeric(as.character(plyr::mapvalues(colnames(seurat_ourStandard_harmony),
                                                         rownames(cd4.cd8_expr),
                                                         cd4.cd8_expr$score)))

p1<- DimPlot(seurat_ourStandard_harmony)
p2<- DimPlot(seurat_ourStandard_harmony, group.by = "detailed_annot")+
  scale_color_manual(values=c(CD4="navy", 
                              CD8="cadetblue3"))
p3<- DimPlot(seurat_ourStandard_harmony, group.by="sampleid")+NoLegend()
p4<- FeaturePlot(seurat_ourStandard_harmony, features = "cd4.cd8_score")+scale_color_gradient2(low="blue", mid="white", high="red")
p1+p2+p3+p4
ggsave(paste0(figure_dir, "OurStandardPipeline_harmony_Tcells.png"), width=10, height=9)


```

## Orr's pipeline
```{r}
seurat_obj<-CreateSeuratObject(counts = cm, min.cells = 0, min.features = 0, project = "")
seurat_obj$sampleid<- samples

## Remove samples without at least 1 cell
nSamples<-table(seurat_obj$sampleid)
seurat_obj<- subset(seurat_obj, sampleid %in% names(nSamples[nSamples>1]))

## Normalize- scale factor different than ours
seurat_obj <- NormalizeData(seurat_obj, normalization.method = "LogNormalize", scale.factor = 10000)

## Split object by sample, select integration features, set these as varible features
seurat_obj_list<- SplitObject(seurat_obj, split.by = "sampleid")
var.genes <- SelectIntegrationFeatures(seurat_obj_list, 
                                       nfeatures = 2000, verbose = TRUE, fvf.nfeatures = 2000, 
                                       selection.method = "vst")
VariableFeatures(seurat_obj) <- var.genes

## Scale
seurat_obj <- ScaleData(seurat_obj, split.by = "sampleid", features = VariableFeatures(seurat_obj), 
                    do.center = T, do.scale = F)

## PCA/neighbors/clusters/umap
seurat_obj <- RunPCA(seurat_obj, features = VariableFeatures(seurat_obj), npcs = 40)
seurat_obj <- RunUMAP(seurat_obj, dims = 1:20, reduction = "pca", n.neighbors = 15,  ## used n.neigbors=30 in original
                      min.dist = 0.5, spread = 1, metric = "euclidean", seed.use = 1)  
seurat_obj <- FindNeighbors(seurat_obj, reduction = "pca", dims = 1:20, k.param = 20) ## n neighbors = 30 in original
seurat_obj <- FindClusters(seurat_obj, resolution = 1, algorithm = 4, random.seed = 100)  ## used louvain not leiden in original

seurat_obj<- AddMetaData(seurat_obj, meta[,!colnames(meta) %in% colnames(seurat_obj@meta.data)])    

seurat_orr<- seurat_obj

## Add cd4/cd8 score- subtract cd4 expr from cd8
cd4.cd8_expr<- as.data.frame(t(seurat_orr@assays$RNA@data[c("CD4", "CD8A"),]))
cd4.cd8_expr$score<- cd4.cd8_expr$CD4 - cd4.cd8_expr$CD8A
seurat_orr$cd4.cd8_score<- as.numeric(as.character(plyr::mapvalues(colnames(seurat_orr),
                                                         rownames(cd4.cd8_expr),
                                                         cd4.cd8_expr$score)))

p1<- DimPlot(seurat_orr)
p2<- DimPlot(seurat_orr, group.by = "detailed_annot")+
  scale_color_manual(values=c(CD4="navy", 
                              CD8="cadetblue3"))
p3<- DimPlot(seurat_orr, group.by="sampleid")+NoLegend()
p4<- FeaturePlot(seurat_orr, features = "cd4.cd8_score")+scale_color_gradient2(low="blue", mid="white", high="red")
p1+p2+p3+p4
ggsave(paste0(figure_dir, "OrrPipeline_Tcells.png"), width=10, height=9)


```

## Our standard pipeline, modified so that all settings match Orr's EXCEPT for variable features
```{r}
## Create Seurat Obj
seurat_obj = CreateSeuratObject(cm, min.cells=0, min.features = 0,  project=project)

## Normalize data
seurat_obj <- NormalizeData(
  object = seurat_obj,
  normalization.method = "LogNormalize",
  scale.factor = 10000 ## changed from 100,000
)

## Detection of variable genes across the single cells
seurat_obj <- FindVariableFeatures(
  object = seurat_obj,
  mean.function = ExpMean, 
  dispersion.function = LogVMR 
)

## Scaling the data and removing unwanted sources of variation
seurat_obj <- ScaleData(
  object = seurat_obj,
  do.scale = FALSE,
  do.center = TRUE
)

seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 40, ## changed from 100
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)
  

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = "pca", dims = 1:20, metric = "euclidean", n.neighbors = 15)
seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = "pca",
                            dims = 1:20,
                            force.recalc = TRUE)  
seurat_obj<- FindClusters(seurat_obj, resolution = .8, algorithm = 4)
seurat_obj<- AddMetaData(seurat_obj, meta[,!colnames(meta) %in% colnames(seurat_obj@meta.data)])    

seurat_ourStandard_noInt_matchOrr<- seurat_obj
  
p1<- DimPlot(seurat_ourStandard_noInt_matchOrr)
p2<- DimPlot(seurat_ourStandard_noInt_matchOrr, group.by = "detailed_annot")+
  scale_color_manual(values=c(CD4="navy", 
                              CD8="cadetblue3"))
p3<- DimPlot(seurat_ourStandard_noInt_matchOrr, group.by="sampleid")+NoLegend()
p1+p2+p3
ggsave(paste0(figure_dir, "OurStandardPipeline_notInt_matchOrrExceptForVariableFeatures_Tcells.png"), width=15, height=4.5)


```

## Our standard pipeline, without integration- change variable features approach to match Orr
```{r}
## Create Seurat Obj
seurat_obj = CreateSeuratObject(cm, min.cells=0, min.features = 0,  project=project)

## Normalize data
seurat_obj <- NormalizeData(
  object = seurat_obj,
  normalization.method = "LogNormalize",
  scale.factor = 1E5
)

## Split object by sample, select integration features, set these as varible features
seurat_obj<- AddMetaData(seurat_obj, meta[,!colnames(meta) %in% colnames(seurat_obj@meta.data)]) 
seurat_obj_list<- SplitObject(seurat_obj, split.by = "sampleid")
var.genes <- SelectIntegrationFeatures(seurat_obj_list, 
                                       nfeatures = 2000, verbose = TRUE, fvf.nfeatures = 2000, 
                                       selection.method = "vst")
VariableFeatures(seurat_obj) <- var.genes

## Scaling the data and removing unwanted sources of variation
seurat_obj <- ScaleData(
  object = seurat_obj,
  do.scale = FALSE,
  do.center = TRUE
)

seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)
  

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = "pca", dims = 1:20)
seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = "pca",
                            dims = 1:20,
                            force.recalc = TRUE)  
seurat_obj<- FindClusters(seurat_obj, resolution = .8)
  

seurat_ourStandard_noInt_matchVFOrr<- seurat_obj
  
p1<- DimPlot(seurat_ourStandard_noInt_matchVFOrr)
p2<- DimPlot(seurat_ourStandard_noInt_matchVFOrr, group.by = "detailed_annot")+
  scale_color_manual(values=c(CD4="navy", 
                              CD8="cadetblue3"))
p3<- DimPlot(seurat_ourStandard_noInt_matchVFOrr, group.by="sampleid")+NoLegend()
p1+p2+p3
ggsave(paste0(figure_dir, "OurStandardPipeline_notInt_matchOrrVariableFeatures_Tcells.png"), width=15, height=4.5)


```

## How many variable features match between our pipeline/orr
```{r}
sum(VariableFeatures(seurat_orr) %in% VariableFeatures(seurat_ourStandard_noInt))
length(VariableFeatures(seurat_orr))
length(VariableFeatures(seurat_ourStandard_noInt))
```

