---
title: "ImmuneRegev_MidlineHistoneMutant_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "C:/Users/jenna/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
project_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/Immune/")
working_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/TumorCellAnnotation/",
                     "BySubtype/Midline/Midline_HistoneMutantWithNewSample/BT1873Only/")

library(Seurat, lib.loc = "C:/Users/jenna/OneDrive/Documents/R/win-library/4.0/Seurat/V4") 

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
stat_helper = paste0(base_dir, script_dir, "common_stat_test.R")
source(preprocessing)
source(stat_helper)

nmf_analysis_folder = "analysis/nmf/Merge/"
liger_analysis_folder = "analysis/liger/"
seurat_analysis_folder = "analysis/seurat/"
seurat_fig_folder = "figures/seurat/"
for (folder in c(seurat_analysis_folder, seurat_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder,recursive = TRUE)
  }
}

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

subtypes<- read.csv(paste0(project_dir, "Subtypes.csv"))
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj_Cluster6Removed.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}

seurat_obj<-seurat_obj_annot
```

## Load counts and metadata for initiating seurat object
```{r load-data}
load(paste0(project_dir, "SplitBySubtype/WithNewSamples/analysis/midlineHistoneMut_cm_list.Robj"))
samples = readRDS(paste0(project_dir, "SplitBySubtype/WithNewSamples/analysis/midlineHistoneMut_samples.rds"))
tumor_stat = readRDS(paste0(project_dir, "SplitBySubtype/WithNewSamples/analysis/midlineHistoneMut_cluster6CellsNormal_call_tumor.rds"))

cm_list<-midlineHistoneMut_cm_list

sum(colnames(cm_list$raw_data)==names(samples)); length(samples)
sum(names(samples)==names(tumor_stat))

## Subset to just BT1873 cells
samples<- samples[samples == "BT1873vent"]
cm_list<-lapply(cm_list, function(x){x[,colnames(x) %in% names(samples)]})
tumor_stat<-tumor_stat[names(tumor_stat) %in% names(samples)]

sum(colnames(cm_list$raw_data)==names(samples)); length(samples)
sum(names(samples)==names(tumor_stat))
```

## Subset cm to remove noise/normal samples
```{r subset-tumor-cells}
cm = cm_list$raw_data; rm(cm_list); gc()
cm = cm[,tumor_stat=="Malignant"]
samples = samples[tumor_stat=="Malignant"]

any(colnames(cm) != names(samples))
```

## Initiate seurat obj and add metadata 
```{r seurat_obj_meta_data}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", min.cells=0, min.genes=0, scale.factor=1E5, do.scale=F, do.center=T)

## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples

sample_df<- as.data.frame(samples); sample_df$cell<- rownames(sample_df)
colnames(sample_df)<- c("sample", "cell")
subtype_df<- merge(sample_df, subtypes, by="sample")

seurat_obj@meta.data$subtype<- subtype_df$subtype
seurat_obj@meta.data$AdditionalSubtype<- subtype_df$Diagnosis
```

## OPTIONAL: plot highly variable genes
```{r}
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## PCA and PC selection
```{r PCA_selection}
## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)

```


## Clustering and UMAP
## Without new sample (BT1873) there were 9 clusters (prior to removing cluster 6 which is likely normal)
## See that same "normal" cluster with the new sample. Prior to removing it, want to aim for 9 clusters
## Resolution of 0.6 gives 9 clusters (with normal cluster included)

## Without new samples and AFTER removing cluster 6: also 9 clusters, 8 clusters if you want less resolution on OPC-like types
## With new samples: resolution of 0.8 gives 9 clusters
```{r clustering_tSNE}
## Run UMAP and clustering using reduced dimensions from Harmony
## You may play around with resolution parameter to tune the clustering result
resolution = .6
reduction_type = "pca"
dims = 20

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = reduction_type, dims = 1:dims)


seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = reduction_type,
                            dims = 1:dims,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = resolution)


r0.9<-DimPlot(seurat_obj, reduction="umap", pt.size = 1.5) + ggtitle("Resolution_0.9") 
r0.5<-DimPlot(seurat_obj, reduction="umap", pt.size = 1.5)+ ggtitle("Resolution_0.5")
r1.2<-DimPlot(seurat_obj, reduction = "umap", pt.size = 1.5)+ ggtitle("Resolution_1.2")

r0.5+r0.9+r1.2
ggsave("UMAP_clustering_VaryResolution.jpg", path=seurat_fig_folder, width=20, height=6)

## Clustering results 
table(Idents(seurat_obj))


## Save the seurat object
save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```

## UMAP plot colored by clustering and samples
## Can color tSNE with other metadata features (e.g. primary/recurrence, treatment)
```{r UMAP_plots}
## By clustering
DimPlot(object = seurat_obj, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_Resolution1.jpg", path=seurat_fig_folder, width=8, height=8)

```



## DEGs for each cluster 
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.25)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05 & all_markers$avg_logFC > .25, ]

## Cluster 1 has very few DEGs- should probably be removed or merged (resolution 1)


## Subset the top 30 DEGs for each cluster
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC) %>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(seurat_analysis_folder,
                                       "all_markers_Theta.1.2.Robj"))
write.csv(top_marker_genes, file=paste0(seurat_analysis_folder,
                                        "Top30Markers_Theta.1.2.csv"))


## Heatmap for initial inspection
DoHeatmap(seurat_obj, 
          features=all_markers_filtered$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs_Theta.1.2.jpg", path=seurat_fig_folder)
```

## Quick annotation by OPC/OC/AC marker genes 
```{r}
## OPC-like
FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OPC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## OC-like cells
FeaturePlot(seurat_obj,
            c("BCAS1", "SIRT2", "RGR", 
              "GPR17", "NFASC", "MBP",
              "TF", "MOG", "PLP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## AC-like cells 
FeaturePlot(seurat_obj,
            c("APOE", "AQP4", "AGT", 
              "SPARCL1","ATP1A2", "GFAP",
              "S100B", "ALDOC", "CLU"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("AC_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## Neuronal-like cells
FeaturePlot(seurat_obj,
            c("STMN2", "STMN1", "STMN4", 
              "TUBB3", "INA", "ELAVL4",
              "SNAP25", "GABRG2", "ST18"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("Neuron_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## FOS/JUN
FeaturePlot(seurat_obj,
            c("FOS", "FOSB",
              "JUN", "JUNB"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("FOS.JUN_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## pre-OPC markers
FeaturePlot(seurat_obj,
            c("SOX1", "ASCL1"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("PreOPC_markers.jpg", path=seurat_fig_folder, width=9, height=4.5)

## mature OPC markers
FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "OLIG2"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("PreOPC_markers.jpg", path=seurat_fig_folder, width=9, height=4.5)

FeaturePlot(seurat_obj,
            c("A2B5", "PDGFRA", "NG2", "O4", "o1", "CNPase", "GalC", "MBP", "MAG", "MOG"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("PreOPC_markers.jpg", path=seurat_fig_folder, width=9, height=4.5)

## inhibitory ligands identified in CellChat
FeaturePlot(seurat_obj,
            c("CLEC2D", "WNT5A", "PVR", "CDH1", "SPP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("InhibitoryLigand_markers.jpg", path=seurat_fig_folder, width=9, height=6)
```

## Violin plot
```{r}
VlnPlot(seurat_obj, 
        c("BCAS1", "SIRT2", "RGR", 
          "GPR17", "NFASC", "MBP", 
          "TF", "MOG", "PLP1"),
        pt.size = 0)
ggsave("OC_markers_ViolinPlot.jpg", path=seurat_fig_folder, width=9, height=9)

VlnPlot(seurat_obj, 
       c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
        pt.size = 0)
ggsave("OPC_markers_ViolinPlot.jpg", path=seurat_fig_folder, width=9, height=9)

VlnPlot(seurat_obj, group.by = "sample",
       c("CLEC2D", "WNT5A", "PVR", "CDH1", "SPP1"),
        pt.size = 0, sort = TRUE)
ggsave("InhibitoryLigand_markers_ViolinPlot_samples.jpg", path=seurat_fig_folder, width=9, height=6)
```

## ORA test 
```{r over-representation-test}
library(clusterProfiler)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
allOE_genes <- as.character(gene_symbol_to_ensembl_id(rownames(seurat_obj@assays$RNA@counts),dataset)$ensembl_gene_id)
save(allOE_genes, file="analysis/seurat/allOE_genes.Robj")

save(all_markers_filtered, file="analysis/seurat/all_markers_filtered_temp.Robj")

## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
sigOE_genes_list<-list()

all_markers_filtered<-all_markers_filtered[all_markers_filtered$cluster==0,]
all_markers_filtered$cluster<- as.character(all_markers_filtered$cluster)
for (i in unique(all_markers_filtered$cluster)){
  sigOE_genes = as.character(gene_symbol_to_ensembl_id(subset(all_markers_filtered,
                                                              cluster==i,gene)[,],
                                                       dataset)$ensembl_gene_id)
  print(head(sigOE_genes))
  sigOE_genes_list[[i]]<-sigOE_genes
  
} 
go_result = list()
for (i in names(sigOE_genes_list)){
  sigOE_genes<-sigOE_genes_list[[i]]
  go_result[[paste0("c", i)]] = go_analysis(sigOE_genes, allOE_genes)
}


## Plot top pathways for all clusters, display on one plot
clusters<-names(go_result)
allDotPlots_20<-list()
for (c in 1:length(clusters)){
  cluster=clusters[c]
  print(cluster)
  if (nrow(go_result[[cluster]]$cluster_summary)>0){
    print(cluster)
    allDotPlots_20[[c]]<-dotplot(go_result[[cluster]]$ego, 
                                showCategory=15, 
                                font.size = 15, 
                                title = cluster, 
                                label_format=10) +
      scale_y_discrete(labels=function(x) str_wrap(x, width=60))+
      theme(plot.title = element_text(hjust = 0.5, face = "bold", 
                                      color="black", size = 28),
            axis.title = element_text(face = "bold", color="black"), 
            axis.text.x = element_text(angle = 45, hjust = 1, color="black"),
            axis.text.y = element_text( color="black",face = "bold"))
  }
  
}

allDotPlots_20<- allDotPlots_20[!(unlist(lapply(allDotPlots_20, function(x){is.null(x)})))]
cowplot::plot_grid(plotlist = allDotPlots_20, ncol=2)
ggsave(paste0("GO_OVA_AllClusters_top20_Res0.5.jpg"), path=seurat_fig_folder, width=18, height=25)
save(go_result, file=paste0(seurat_analysis_folder,  "go_result_Resolution0.5.Robj"))

```


## ORA with midline genesets
## Compare with metaprograms from previously defined metaprograms from ependymoma
```{r}
all_markers_list<-split(all_markers_filtered, all_markers_filtered$cluster )
all_markers_list<-lapply(all_markers_list, function(x){x$gene})
names(all_markers_list)<-paste0("Cluster", names(all_markers_list))

previous_dir<- paste0(project_dir, "../../PublishedDatasets/FilbinPublishedDatasets/Metaprograms/")
midline_genesets<-readRDS(paste0(previous_dir,"midlinePreviousPrograms_OC.AC.OPC.MES1.rds"))


tmp2 = sapply(all_markers_list, function(x){
    sapply(midline_genesets, function(y){
        fisher_test(x, y, unlist(union(midline_genesets, all_markers_list)))
    })
})
tmp2 = data.frame(tmp2)
tmp3<- apply(tmp2, 2, function(x) {p.adjust(x) < 0.1})
tmp4<-apply(tmp2, 2, function(x) {p.adjust(x)})

write.table(tmp4, paste0(seurat_analysis_folder,
                        "ScoreSeuratClustersUsingPreviousNMF_Theta0.9_padj.txt"))
```




## Add a second metagene classification without cell cycle programs
```{r}
nmf_score = readRDS(paste0("analysis/nmf/all_subtypes/new/nmf_score_updated", NameOfMetaprogramList,  ".rds"))
seurat_obj@meta.data$signaure_1_without_cc = nmf_score$signature_1
```

## Make pseudobulk
```{r}
## Split by cell type
metagene_program_names = sort(unique(nmf_score$signature_1))
pseudobulk = NULL
aggr_method = "mean"
for (metagene in metagene_program_names){
  cm = seurat_obj@assays$RNA@counts[,nmf_score$signature_1 == metagene & nmf_score$score_1 >= 1]
  if (aggr_method == "mean"){
    pseudobulk = cbind(pseudobulk, rowMeans(cm))
  }
  if (aggr_method == "median"){
    pseudobulk = cbind(pseudobulk, rowMedians(cm))
  }
}
colnames(pseudobulk) = metagene_program_names
rownames(pseudobulk) = rownames(seurat_obj@assays$RNA@counts)
saveRDS(pseudobulk, file=paste0(seurat_analysis_folder, 
                                "pseudobulk_cm_unique_only_", aggr_method, NameOfMetaprogramList,".RDS"))

## All cell types together- single pseudobulk for each sample
## Sum of counts across each sample
pseudobulk<- data.frame(gene=rownames(seurat_obj))
mySamples<- unique(seurat_obj$sample)
for (i in mySamples){
  seurat_tmp<- subset(seurat_obj, sample==i)
  cm_tmp<-seurat_tmp@assays$RNA@counts
  sumGene<- as.data.frame(rowSums(cm_tmp))
  colnames(sumGene)<- i
  sumGene$gene<- rownames(sumGene)
  pseudobulk<- merge(pseudobulk, sumGene, by="gene")
}

rownames(pseudobulk)<- pseudobulk$gene
pseudobulk<- pseudobulk[,-1]

save(pseudobulk, file="../data/PseudobulkCounts/MidlineHistoneMutant_Pseudobulk.Robj")
```

## Metagene program specific genes as overall DEGs
```{r}
seurat_obj = SetIdent(seurat_obj, value = seurat_obj@meta.data$signaure_1_without_cc)

all_markers_nmf = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_nmf = all_markers[all_markers$p_val_adj < 0.05, ]
saveRDS(all_markers, paste0(seurat_analysis_folder, "deg.rds"))
##sapply(names(nmf_marker_genes_final), function(x) table(nmf_marker_genes_final[[x]] %in% gene_signatures[[x]]))
signature_genes = unique(all_markers$gene)
saveRDS(signature_genes, paste0(seurat_analysis_folder, "signature_genes.rds"))
```




