---
title: "Visualize_reannotate_programs"
output: html_document
---

## Purpose:
NMF programs were generated based on 1) de novo NMF separately for ped/adult Tcells 2) correlating programs 3) keeping "specific" programs separate, but merging "shared" programs --> score --> assign max -->regenerate gene lists. These programs are annotated based on correlation with original ped cd4/cd8 programs (no longer in use, archived). Some of these annotations make sense, but others (like "CD4_Inhibitory") don't. Here, basic visualizations of these programs, comparing ped/adult proportions, GO terms, marker genes- all used to help re-annotate programs as needed


## Approach:
First, just want to do basic visualizations (proportion of programs in cd4/cd8), then compare differences in ped/adult proportions within CD4/CD8. Then various visualizations to help annotate: GO terms, marker genes, etc. Re-annotate programs as needed, then update seurat objects + marker gene lists as needed to reflect this


```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(harmony)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/", 
                     "03_Visualize.Reannotate_programs/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/Reannotate_programs/")
figure_dir_tmp<- paste0(working_dir, "/figures/Reannotate_programs/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in ped tcell seurat object. Contains NMF assignments (annots added in deNovoPedTcell.v.deNovoAdultTcell, based on correlation with adult T cell + correlation with ped cd4/cd8 programs)
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../02_detailed_annot/analysis/identifyCD4.CD8_byNMF/", "rank",rank_use, "_sharedBy", sharedBy, "/" )

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
  
## ped/adult tcell seurat objects
tcell_seurat<- readRDS(paste0("../../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/Tcells_Ped.Adult_Merge/",
                              "adult_ped_tcell_merged_seurat_harmony.Rds"))

cd4_seurat<- readRDS(paste0(data_dir, "ped.adult_seurat_cd4_harmony.Rds"))
cd8_seurat<-readRDS(paste0(data_dir, "ped.adult_seurat_cd8_harmony.Rds"))

## Read in marker genes
program_markers<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                                 "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", sharedBy, ".Rds"))

## Read in colors to use
program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, ".Rds"))
celltype_colors<- readRDS(paste0("../../../plot_colors/Tcell.celltypes.Rds"))
```


## UMAP of programs
```{r}
DimPlot(cd4_seurat, group.by = program_name)+scale_color_manual(values=colors_use)
ggsave(paste0(figure_dir, "CD4_byProgram.png"), width=7, height=5)

DimPlot(cd8_seurat, group.by = program_name)+scale_color_manual(values=colors_use)
ggsave(paste0(figure_dir, "CD8_byProgram.png"), width=7, height=5)
```

## Breakdown of programs within cd4/cd8
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy)
annot_name<- paste0("detailed_annot_byNMFrank", rank_use, "_mergedBy", sharedBy)


## Proportion of program by CD4/CD8
meta<-data.frame(program=tcell_seurat@meta.data[[program_name]],
                 celltype=tcell_seurat@meta.data[[annot_name]],
                 sample=tcell_seurat$sampleid,
                 age=tcell_seurat$Age)

bd<- as.data.frame(table(meta$program, meta$celltype))
colnames(bd)<- c("Program", "CellType", "Ncells")
bd<- bd[bd$CellType!="Unclear",]

  ## Cell type by program
  bd<- bd %>% group_by(CellType) %>% mutate(PerCells=Ncells/sum(Ncells)) %>% as.data.frame()
  ggplot(bd, aes(x=CellType, y=PerCells, fill=Program))+
    geom_bar(stat="identity", position="stack")+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))+
    ylab("Proportion of cells")+xlab("")
  ggsave(paste0(figure_dir, "CD4.CD8breakdown_byProgram.png"), width=5, height=5)
  
  ## Program by cell type
  bd<- bd %>% group_by(Program) %>% mutate(PerCells=Ncells/sum(Ncells)) %>% as.data.frame()
  ggplot(bd, aes(x=Program, y=PerCells, fill=CellType))+
    geom_bar(stat="identity", position="stack")+
    scale_fill_manual(values=celltype_colors)+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))+
    ylab("Proportion of cells")+xlab("")
  ggsave(paste0(figure_dir, "ProgramBreakdown_byCD4.CD8.png"), width=5, height=3)

## Proportion of program by CD4/CD8, grouped by age
bd_age<- meta %>% group_by(celltype, program, age) %>% summarise(nCells=n())  %>% as.data.frame()
bd_age<- bd_age[bd_age$celltype!="Unclear",]

  ## Cell type by program
  bd_age<- bd_age %>%  group_by(celltype,  age) %>% mutate(PerCells=nCells/sum(nCells)) %>% as.data.frame()
  ggplot(bd_age, aes(x=age, y=PerCells, fill=program))+
    geom_bar(stat="identity", position="stack")+
    scale_fill_manual(values=colors_use)+
    facet_grid(cols=vars(celltype))+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))+
    ylab("Proportion of cells")+xlab("")
  ggsave(paste0(figure_dir, "CD4.CD8breakdown_byProgram.byAge.png"), width=7, height=5)
  
  ## Program by cell type
  bd_age<- bd_age %>%  group_by(program,  age) %>% mutate(PerCells=nCells/sum(nCells)) %>% as.data.frame()
  bd_age$program<- gsub("\\.", "\n", gsub("EffectorMemory", "Effector\nMemory", bd_age$program))
  ggplot(bd_age, aes(x=age, y=PerCells, fill=celltype))+
    geom_bar(stat="identity", position="stack")+
    scale_fill_manual(values=celltype_colors)+
    facet_grid(cols=vars(program))+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))+
    ylab("Proportion of cells")+xlab("")
  ggsave(paste0(figure_dir, "ProgramBreakdown_byCD4.CD8.byAge.png"), width=8, height=3)
```

## GO terms
```{r}
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert input using biomart
ensembl = useMart("ensembl",dataset="hsapiens_gene_ensembl")
allOE_genes<-getBM(attributes=c('ensembl_gene_id','hgnc_symbol'), 
                   filters = 'hgnc_symbol', 
                   values = rownames(tcell_seurat@assays$RNA@counts), 
                   mart = ensembl)

## Subset to top 100 markers- all have >100 except for stress (only 65)
top_markers<- program_markers %>% group_by(cluster) %>% top_n(n=100, wt=avg_log2FC) %>% as.data.frame()

## Convert marker genes to ensembl IDs too, removing any genes that don't map to ens
top_markers$gene_ens<- as.character(plyr::mapvalues(top_markers$gene, allOE_genes$hgnc_symbol, 
                                                    allOE_genes$ensembl_gene_id, warn_missing = FALSE))
top_markers<-top_markers[top_markers$gene != top_markers$gene_ens,]


## run OR analysis (GO)
go_result = list()
for (i in unique(top_markers$cluster)){
  print(i)
  genes<- top_markers[top_markers$cluster==i, "gene_ens"]
  if(length(genes)!=1){
    go_result[[paste0( i)]] = go_analysis(genes, allOE_genes$ensembl_gene_id)
    }
} 
all_plots<- plotGO(go_result, n_terms = 15)
cowplot::plot_grid(plotlist = all_plots, ncol=3)

ggsave(paste0(figure_dir, "GO_NMF.png"), width=30, height=18)
saveRDS(go_result, file=paste0(analysis_dir, "go_result_NMF.Rds"))
```

## Plot marker genes
```{r}
## CD8 by dysfunction/Cytotoxic markers
  ## Vln
  program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy)
  VlnPlot(cd8_seurat, features=c("ENTPD1", "ITGAE", "TNFRSF9", "TNFRSF18", "PDCD1", "CTLA4", "LAG3", "GZMK", "IFIT1"), 
          ncol=3, sort=TRUE,
          group.by = program_name,
          cols = program_colors)
  ggsave(paste0(figure_dir, "CD8_markerGenes.dysfunctional.cyto.png"), width=9, height=9)

  ## Dot plot: by program + age
  cd8_seurat$tmp<- paste0(cd8_seurat@meta.data[[program_name]], "_", cd8_seurat$Age)
  DotPlot(cd8_seurat, features=c("ENTPD1", "ITGAE", "TNFRSF9", "TNFRSF18", "PDCD1", "CTLA4", "LAG3", 
                                 "GZMK", "IFIT1", "TOP2A", "CCR7", "GNLY"),
          group.by = "tmp")+
    theme(axis.text.x = element_text(angle=45, hjust=1))+
    ylab("") + xlab("")
  ggsave(paste0(figure_dir, "CD8_dotPlot_markerGenes.dysfunctional.cyto.png"), width=7, height=5)
  
## CD4 by dysfunction/Cytotoxic markers
  ## Dot plot: by program + age
  cd4_seurat$tmp<- paste0(cd4_seurat@meta.data[[program_name]], "_", cd4_seurat$Age)
  DotPlot(cd4_seurat, features=c("ENTPD1", "ITGAE", "TNFRSF9", "TNFRSF18", "PDCD1", "CTLA4", "LAG3", 
                                 "GZMK", "IFIT1", "TOP2A", "CCR7", "FOXP3"),
          group.by = "tmp")+
    theme(axis.text.x = element_text(angle=45, hjust=1))+
    ylab("") + xlab("")
  ggsave(paste0(figure_dir, "CD4_dotPlot_markerGenes.dysfunctional.cyto.png"), width=7, height=5)
```

## Add module score for cyto/dysfunctional markers
Based on CD8 paper (Nature 2020)
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy)

all_markers<- list(cytotoxic= c("CX3CR1", "PRF1", "GZMA", "GZMB", "GZMH",  "FGFBP2","KLRG1", "FCGR3A"),
                   naive=c("TCF7", "CCR7", "SELL", "LEF1"),
                   predysfunctional=c("GZMK", "PDCD1", "CXCR3"),
                   dysfunctional= c("HAVCR2", "GZMB","IFNG", "CXCL13", "TIGIT", "LAYN", "LAG3", 
                                    "PDCD1", "CTLA4", "ITGAE", "ENTPD1")) 

## Add score to cd8/cd4
cd8_seurat<- AddModuleScore(cd8_seurat, all_markers, name = names(all_markers))
cd4_seurat<- AddModuleScore(cd4_seurat, all_markers, name = names(all_markers))
rename_cols<-paste0(names(all_markers), 1:4); names(rename_cols)<-names(all_markers)
colnames(cd8_seurat@meta.data)<- plyr::mapvalues(colnames(cd8_seurat@meta.data), rename_cols, names(rename_cols))
colnames(cd4_seurat@meta.data)<- plyr::mapvalues(colnames(cd4_seurat@meta.data), rename_cols, names(rename_cols))

## Add on max score/program from 4 gene lists- ScoreTooLow if less than 0.5
cd8_seurat$MaxNature2020Program<- apply(cd8_seurat@meta.data[, names(all_markers)], 1, 
                                      function(x){names(x)[which.max(x)]})
cd4_seurat$MaxNature2020Program<- apply(cd4_seurat@meta.data[, names(all_markers)], 1, 
                                      function(x){names(x)[which.max(x)]})

cd8_seurat$MaxNature2020Score<- apply(cd8_seurat@meta.data[, names(all_markers)], 1, 
                                      function(x){x[which.max(x)]})
cd4_seurat$MaxNature2020Score<- apply(cd4_seurat@meta.data[, names(all_markers)], 1, 
                                      function(x){x[which.max(x)]})

cd8_seurat$MaxNature2020Program<- ifelse(cd8_seurat$MaxNature2020Score<0.5, "ScoreTooLow", cd8_seurat$MaxNature2020Program)
cd4_seurat$MaxNature2020Program<- ifelse(cd4_seurat$MaxNature2020Score<0.5, "ScoreTooLow", cd4_seurat$MaxNature2020Program)

## Plot scores
  ## CD8
  all_plots<- lapply(names(all_markers), function(x){
    FeaturePlot(cd8_seurat, features=x)+scale_color_gradient2(low="blue", mid="white", high="red")
  })
  cowplot::plot_grid(plotlist=all_plots)
  ggsave(paste0(figure_dir, "CD8_markerScoresFromNature2020.png"), width=8, height=6)
  
  ## CD4
  all_plots<- lapply(names(all_markers), function(x){
  FeaturePlot(cd4_seurat, features=x)+scale_color_gradient2(low="blue", mid="white", high="red")
  })
  cowplot::plot_grid(plotlist=all_plots)
  ggsave(paste0(figure_dir, "CD4_markerScoresFromNature2020.png"), width=8, height=6)

## Plot scores by program
VlnPlot(cd8_seurat, features=names(all_markers), 
        group.by = program_name,  cols = program_colors, ncol = 2, sort=TRUE, pt.size = .01)
ggsave(paste0(figure_dir, "CD8_markerScoresFromNature2020_byProgram.png"), width=8, height=6)

VlnPlot(cd4_seurat, features=names(all_markers), 
        group.by = program_name,  cols = program_colors, ncol = 2, sort=TRUE, pt.size = .01)
ggsave(paste0(figure_dir, "CD4_markerScoresFromNature2020_byProgram.png"), width=8, height=6)

## Dotplot of scores by program, split by age- same pattern for ped/adult?
  ## CD8
  cd8_seurat$tmp<- paste0(cd8_seurat@meta.data[[program_name]], "_", cd8_seurat$Age)
  DotPlot(cd8_seurat, features=names(all_markers), group.by = "tmp")+
    theme(axis.text.x = element_text(angle=45, hjust=1))+
    ylab("")+xlab("")
  ggsave(paste0(figure_dir, "CD8_markerScoresFromNature2020_byProgram.Age.png"), width=6, height=5)
  
  ## CD
  cd4_seurat$tmp<- paste0(cd4_seurat@meta.data[[program_name]], "_", cd4_seurat$Age)
  DotPlot(cd4_seurat, features=names(all_markers), group.by = "tmp")+
    theme(axis.text.x = element_text(angle=45, hjust=1))+
    ylab("")+xlab("")
  ggsave(paste0(figure_dir, "CD4_markerScoresFromNature2020_byProgram.Age.png"), width=6, height=5)
  
## Breakdown of max geneset by program
  plot_bd<- function(bd, colors){
    p<- ggplot(bd, aes(x=tmp, y=perCells, fill=MaxNature2020Program))+
      geom_bar(stat="identity", position="stack")+
      theme_classic()+
      scale_fill_manual(values=colors)+
      theme(axis.text.x = element_text(angle=45, hjust=1),
            axis.text = element_text(face="bold", color="black"),
            axis.title = element_text(face="bold", color="black"))+
      ylab("Proportion of cells") + xlab("")
    return(p)
  }
  colors<- c(cytotoxic="olivedrab", dysfunctional="red", predysfunctional="orange", naive="skyblue", ScoreTooLow="grey")
  
  ## CD8
  cd8_seurat$tmp<- cd8_seurat@meta.data[[program_name]]
  bd<- cd8_seurat@meta.data[,c("tmp", "MaxNature2020Program", "Age")] %>% 
    group_by(tmp, MaxNature2020Program) %>% summarise(nCells=n()) %>% as.data.frame() %>%
    group_by(tmp) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()
  plot_bd(bd, colors)
  ggsave(paste0(figure_dir, "CD8_Nature2020Geneset_bdByProgram.png"), width=5, height=3)
  
    ## No ScoreToLow
    bd<- bd %>% filter(MaxNature2020Program!="ScoreTooLow") %>% group_by(tmp) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()
    plot_bd(bd, colors[names(colors)!="ScoreTooLow"])
    ggsave(paste0(figure_dir, "CD8_Nature2020Geneset_bdByProgram_NoScoreTooLow.png"), width=5, height=3)
    
  ## CD4
  cd4_seurat$tmp<- cd4_seurat@meta.data[[program_name]]
  bd<- cd4_seurat@meta.data[,c("tmp", "MaxNature2020Program", "Age")] %>% 
    group_by(tmp, MaxNature2020Program) %>% summarise(nCells=n()) %>% as.data.frame() %>%
    group_by(tmp) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()
  plot_bd(bd, colors)
  ggsave(paste0(figure_dir, "CD4_Nature2020Geneset_bdByProgram.png"), width=5, height=3)
  
    ## No ScoreToLow
    bd<- bd %>% filter(MaxNature2020Program!="ScoreTooLow") %>% group_by(tmp) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()
    plot_bd(bd, colors[names(colors)!="ScoreTooLow"])
    ggsave(paste0(figure_dir, "CD4_Nature2020Geneset_bdByProgram_NoScoreTooLow.png"), width=5, height=3)
    
    
   
saveRDS(cd4_seurat, file=paste0(data_dir, "ped.adult_seurat_cd4_harmony.Rds"))
saveRDS(cd8_seurat, file=paste0(data_dir, "ped.adult_seurat_cd8_harmony.Rds"))
```


## Plot by number of clonal cells
```{r}
cd8_seurat$clonal<- gsub("not clonal", "nonclonal", cd8_seurat$clonal)
cd4_seurat$clonal<- gsub("not clonal", "nonclonal", cd4_seurat$clonal)


## CD8: ## Proportion of clonal cells in each program
meta<-data.frame(program=cd8_seurat@meta.data[[program_name]],
                 clonal=cd8_seurat$clonal,
                 sample=cd8_seurat$sampleid,
                 age=cd8_seurat$Age)
  
  ## Adult/ped together
  bd_age<- meta %>% group_by(clonal, program) %>% summarise(nCells=n())  %>% as.data.frame()
  bd_age<- bd_age %>%  group_by(program) %>% mutate(PerCells=nCells/sum(nCells)) %>% as.data.frame()
  ggplot(bd_age, aes(x=program, y=PerCells, fill=clonal))+
    geom_bar(stat="identity", position="stack")+
    scale_fill_manual(values=c(clonal="red", nonclonal="grey"))+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))+
    ylab("Proportion of cells")+xlab("")
  ggsave(paste0(figure_dir, "ProgramBreakdown_byClonal.png"), width=5, height=3)


  
  ## By age
  bd_age<- meta %>% group_by(clonal, program, age) %>% summarise(nCells=n())  %>% as.data.frame()
  bd_age<- bd_age %>%  group_by(program,  age) %>% mutate(PerCells=nCells/sum(nCells)) %>% as.data.frame()
  bd_age$program<- gsub("\\.", "\n", gsub("EffectorMemory", "Effector\nMemory", bd_age$program))
  ggplot(bd_age, aes(x=age, y=PerCells, fill=clonal))+
    geom_bar(stat="identity", position="stack")+
    scale_fill_manual(values=c(clonal="red", nonclonal="grey"))+
    facet_grid(cols=vars(program))+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))+
    ylab("Proportion of cells")+xlab("")
  ggsave(paste0(figure_dir, "ProgramBreakdown_byClonal.byAge.png"), width=8, height=3)
  
## CD4: ## Proportion of clonal cells in each program
meta<-data.frame(program=cd4_seurat@meta.data[[program_name]],
                 clonal=cd4_seurat$clonal,
                 sample=cd4_seurat$sampleid,
                 age=cd4_seurat$Age)
  
  ## Adult/ped together
  bd_age<- meta %>% group_by(clonal, program) %>% summarise(nCells=n())  %>% as.data.frame()
  bd_age<- bd_age %>%  group_by(program) %>% mutate(PerCells=nCells/sum(nCells)) %>% as.data.frame()
  ggplot(bd_age, aes(x=program, y=PerCells, fill=clonal))+
    geom_bar(stat="identity", position="stack")+
    scale_fill_manual(values=c(clonal="red", nonclonal="grey"))+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))+
    ylab("Proportion of cells")+xlab("")
  ggsave(paste0(figure_dir, "CD4_ProgramBreakdown_byClonal.png"), width=5, height=3)


  
  ## By age
  bd_age<- meta %>% group_by(clonal, program, age) %>% summarise(nCells=n())  %>% as.data.frame()
  bd_age<- bd_age %>%  group_by(program,  age) %>% mutate(PerCells=nCells/sum(nCells)) %>% as.data.frame()
  bd_age$program<- gsub("\\.", "\n", gsub("EffectorMemory", "Effector\nMemory", bd_age$program))
  ggplot(bd_age, aes(x=age, y=PerCells, fill=clonal))+
    geom_bar(stat="identity", position="stack")+
    scale_fill_manual(values=c(clonal="red", nonclonal="grey"))+
    facet_grid(cols=vars(program))+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1),
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"))+
    ylab("Proportion of cells")+xlab("")
  ggsave(paste0(figure_dir, "CD4_ProgramBreakdown_byClonal.byAge.png"), width=8, height=3)
```



## Find markers between 3 most unclear programs (separately for CD4/CD8): Cyto CD4, Inhibitory CD8, Inhibitory CD4 (maybe EM)
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy)

cd8_seurat$tmp<- cd8_seurat@meta.data[[program_name]]
cd8_subset<- subset(cd8_seurat, tmp %in% c("Cytotoxic.CD4", "Inhibitory.CD8", "Inhibitory.CD4", "EffectorMemory"))
cd8_subset<- SetIdent(cd8_subset, value=cd8_subset$tmp)
all_markers<-FindAllMarkers(cd8_subset, min.pct = 0.2)
all_markers_filtered<- all_markers[all_markers$p_val_adj<0.05 & all_markers$avg_log2FC>0,]
top_markers<- all_markers_filtered %>% group_by(cluster) %>% top_n(n=20, wt=avg_log2FC) %>% as.data.frame()
rescale<- ScaleData(cd8_subset, do.scale = FALSE, do.center = TRUE, features=rownames(cd8_subset))
DoHeatmap(rescale, group.by = "tmp", features=top_markers$gene)+scale_fill_gradient2(low="blue", mid="white", high="red")
ggsave(paste0(figure_dir, "CD8_Heatmap_CytoCD4.InhCD8.CD4.EM.png"), width=7, height=12)
```


## Based on the above, update NMF annotations as needed
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy)
new_program_name<- paste0(program_name, "_reAnnot")

current_programs<- unique(as.character(tcell_seurat@meta.data[[program_name]]))

## Rank8, shared by union
names(current_programs)<-c("Memory","Inhibitory", "Inhibitory.Cycling", "IFN.Response", "Stress", "Effector.Memory",
                           "Predysfunctional", "Treg", "Cytotoxic")

## Update seurat objects
tcell_seurat@meta.data[[new_program_name]]<- as.character(plyr::mapvalues(tcell_seurat@meta.data[[program_name]],
                                                                          current_programs,
                                                                          names(current_programs)))
cd8_seurat@meta.data[[new_program_name]]<- as.character(plyr::mapvalues(cd8_seurat@meta.data[[program_name]],
                                                                          current_programs,
                                                                          names(current_programs)))
cd4_seurat@meta.data[[new_program_name]]<- as.character(plyr::mapvalues(cd4_seurat@meta.data[[program_name]],
                                                                          current_programs,
                                                                          names(current_programs)))

## Update marker lists
program_markers$cluster<- as.character(plyr::mapvalues(program_markers$cluster, current_programs, names(current_programs)))

## Update program colors
program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, ".Rds"))
names(program_colors)<- plyr::mapvalues(names(program_colors), current_programs, names(current_programs))

## Update scores
all_scores<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/",
                            "scores_merged.ped.adultRank", rank, "_specific.sharedBy", sharedBy, ".Rds"))
colnames(all_scores)<-plyr::mapvalues(colnames(all_scores), current_programs, names(current_programs))
all_scores$MaxProgram<- plyr::mapvalues(all_scores$MaxProgram, current_programs, names(current_programs))


## Update seurat list split by each program
seurat_list<- readRDS(paste0("../02_detailed_annot/analysis/identifyCD4.CD8_byNMF/rank", rank_use,
                             "_sharedBy", sharedBy, "/seurat_list_splitByNMFrank", rank_use, ".Rds"))
names(seurat_list)<- plyr::mapvalues(names(seurat_list), current_programs, names(current_programs))

## Save
saveRDS(tcell_seurat, file=paste0(analysis_dir, "ped.adult_seurat_tcell_harmony.Rds"))
saveRDS(cd4_seurat, file=paste0(analysis_dir, "ped.adult_seurat_cd4_harmony.Rds"))
saveRDS(cd8_seurat, file=paste0(analysis_dir, "ped.adult_seurat_cd8_harmony.Rds"))

saveRDS(program_markers, paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                                 "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", sharedBy, "_reAnnot.Rds"))

saveRDS(program_colors, file=paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, "_reAnnot.Rds"))

saveRDS(all_scores, file=paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/",
                            "scores_merged.ped.adultRank", rank, "_specific.sharedBy", sharedBy, "_reAnnot.Rds"))

saveRDS(seurat_list, file=paste0("../02_detailed_annot/analysis/identifyCD4.CD8_byNMF/rank", rank_use,
                             "_sharedBy", sharedBy, "/seurat_list_splitByNMFrank", rank_use, ".Rds"))


```

## Remove cycling from CD8, rerun clustering- does Inhibitory.Cycling still cluster separately?
```{r}
cycling_genes<- read.table(paste0(base_dir,"Dropbox (Partners HealthCare)/FilbinLab/data_analysis/",
                                "Marker_genes/regev_lab_cell_cycle_genes.txt"))

cm<- cd8_seurat@assays$RNA@counts
cm<- cm[!rownames(cm) %in% cycling_genes$V1,]

cd8_seurat_noCycling<-RunFullSeurat(cm, RunHarmony = TRUE, samples=cd8_seurat$sampleid)
cd8_seurat_noCycling<- AddMetaData(cd8_seurat_noCycling, 
                                   cd8_seurat@meta.data[,!colnames(cd8_seurat@meta.data) %in%
                                                          colnames(cd8_seurat_noCycling@meta.data)])

DimPlot(cd8_seurat_noCycling, group.by = "tmp") +
  scale_color_manual(values=program_colors)
saveRDS(cd8_seurat_noCycling, file=paste0(analysis_dir, "cd8_seurat_harmony_no.cycling.genes.Rds"))
```

## Split by ped/adult
```{r}
## Pediatric
ped_cd8_seurat<- subset(cd8_seurat, Age=="Ped")
ped_cd4_seurat<- subset(cd4_seurat, Age=="Ped")
ped_tcell_seurat<- subset(tcell_seurat, Age=="Ped")

ped_cd8_meta<- ped_cd8_seurat@meta.data
ped_cd4_meta<- ped_cd4_seurat@meta.data
ped_tcell_meta<- ped_tcell_seurat@meta.data

ped_cd8_seurat<- RunFullSeurat(ped_cd8_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=ped_cd8_meta$sampleid)
ped_cd4_seurat<- RunFullSeurat(ped_cd4_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=ped_cd4_meta$sampleid)
ped_tcell_seurat<- RunFullSeurat(ped_tcell_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=ped_tcell_meta$sampleid)

ped_cd8_seurat<- AddMetaData(ped_cd8_seurat, ped_cd8_meta[,!colnames(ped_cd8_meta) %in% colnames(ped_cd8_seurat@meta.data)])
ped_cd4_seurat<- AddMetaData(ped_cd4_seurat, ped_cd4_meta[,!colnames(ped_cd4_meta) %in% colnames(ped_cd4_seurat@meta.data)])
ped_tcell_seurat<- AddMetaData(ped_tcell_seurat, ped_tcell_meta[,!colnames(ped_tcell_meta) %in% colnames(ped_tcell_seurat@meta.data)])

## Adult
adult_cd8_seurat<- subset(cd8_seurat, Age=="Adult")
adult_cd4_seurat<- subset(cd4_seurat, Age=="Adult")
adult_tcell_seurat<- subset(tcell_seurat, Age=="Adult")

adult_cd8_meta<- adult_cd8_seurat@meta.data
adult_cd4_meta<- adult_cd4_seurat@meta.data
adult_tcell_meta<- adult_tcell_seurat@meta.data

adult_cd8_seurat<- RunFullSeurat(adult_cd8_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=adult_cd8_meta$sampleid)
adult_cd4_seurat<- RunFullSeurat(adult_cd4_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=adult_cd4_meta$sampleid)
adult_tcell_seurat<- RunFullSeurat(adult_tcell_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=adult_tcell_meta$sampleid)

adult_cd8_seurat<- AddMetaData(adult_cd8_seurat, 
                               adult_cd8_meta[,!colnames(adult_cd8_meta) %in% colnames(adult_cd8_seurat@meta.data)])
adult_cd4_seurat<- AddMetaData(adult_cd4_seurat, 
                               adult_cd4_meta[,!colnames(adult_cd4_meta) %in% colnames(adult_cd4_seurat@meta.data)])
adult_tcell_seurat<- AddMetaData(adult_tcell_seurat, 
                                 adult_tcell_meta[,!colnames(adult_tcell_meta) %in% colnames(adult_tcell_seurat@meta.data)])

## Save
saveRDS(ped_cd8_seurat, file=paste0(analysis_dir, "ped_seurat_cd8_harmony.Rds"))
saveRDS(ped_cd4_seurat, file=paste0(analysis_dir, "ped_seurat_cd4_harmony.Rds"))
saveRDS(ped_tcell_seurat, file=paste0(analysis_dir, "ped_seurat_tcell_harmony.Rds"))
saveRDS(adult_cd8_seurat, file=paste0(analysis_dir, "adult_seurat_cd8_harmony.Rds"))
saveRDS(adult_cd4_seurat, file=paste0(analysis_dir, "adult_seurat_cd4_harmony.Rds"))
saveRDS(adult_tcell_seurat, file=paste0(analysis_dir, "adult_seurat_tcell_harmony.Rds"))
```

## Breakdowns for scCoda-
```{r}
sccoda_dir<- paste0("../../../scCoda/data/",cohort, "/")
if(!dir.exists(sccoda_dir)){dir.create(sccoda_dir)}

## CD8
meta<- cd8_seurat@meta.data

  ## Program by Age
  bd<- as.data.frame.matrix(table(meta$sampleid, meta$NMF_rank8_annot_mergedByunion_reAnnot))
  
  ## In a sample, can't have more than 50% of cell types be 0
  #nPrograms<- ncol(bd)
  #bd$nPrograms<- apply(bd, 1, function(x){sum(x>1)})
  #bd$PF<- bd$nPrograms>(nPrograms/2)
  #bd<- bd[bd$PF,]; bd$PF<-NULL; bd$nPrograms<-NULL
  
  ## Add on age group to compare proportions
  bd$Subtype<- plyr::mapvalues(rownames(bd), meta$sampleid, meta$Age, warn_missing = FALSE)
  bd$Sample<- rownames(bd)
  bd$Subtype<- gsub("-", "_", bd$Subtype)
  
  write.csv(bd, paste0(sccoda_dir,"Ped.Adult_CD8_NMFPrograms.csv"), row.names = FALSE)
  
## CD4
meta<- cd4_seurat@meta.data

  ## Program by Age
  bd<- as.data.frame.matrix(table(meta$sampleid, meta$NMF_rank8_annot_mergedByunion_reAnnot))
  
  ## In a sample, can't have more than 50% of cell types be 0
  #nPrograms<- ncol(bd)
  #bd$nPrograms<- apply(bd, 1, function(x){sum(x>1)})
  #bd$PF<- bd$nPrograms>(nPrograms/2)
  #bd<- bd[bd$PF,]; bd$PF<-NULL; bd$nPrograms<-NULL
  
  ## Add on age group to compare proportions
  bd$Subtype<- plyr::mapvalues(rownames(bd), meta$sampleid, meta$Age, warn_missing = FALSE)
  bd$Sample<- rownames(bd)
  bd$Subtype<- gsub("-", "_", bd$Subtype)
  
  write.csv(bd, paste0(sccoda_dir,"Ped.Adult_CD4_NMFPrograms.csv"), row.names = FALSE)

```