---
title: "Visualize_programs"
output: html_document
---

## Purpose:
After re-annotating programs based on markers, GO, reference genesets, etc in 01_Visualize.Reannotate_programs, visualize updated programs (all Tcells together) here


## Approach:
Using updated annotations, plot nice plots for heatmap, UMAP, density, etc


```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(harmony)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/", 
                     "03_Visualize.Reannotate_programs/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/NMF_plots/AllTcells/")
figure_dir_tmp<- paste0(working_dir, "/figures/NMF_plots/AllTcells/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in ped tcell seurat object. Contains NMF assignments (annots added in deNovoPedTcell.v.deNovoAdultTcell, based on correlation with adult T cell + correlation with ped cd4/cd8 programs)
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../03_Visualize.Reannotate_programs/analysis/Reannotate_programs/", "rank",rank_use, "_sharedBy", sharedBy, "/" )

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
  
## ped/adult tcell seurat object
tcell_seurat<- readRDS(paste0(data_dir, "ped.adult_seurat_tcell_harmony.Rds"))

## Read in marker genes
program_markers<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                                 "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", sharedBy, "_reAnnot.Rds"))

## Read in scores
all_scores<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/",
                            "scores_merged.ped.adultRank", rank_use, "_specific.sharedBy", sharedBy, "_reAnnot.Rds"))

## Read in colors to use
program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, "_reAnnot.Rds"))
celltype_colors<- readRDS(paste0("../../../plot_colors/Tcell.celltypes.Rds"))
```

## Subset to top markers for each nmf- top 10
```{r}
marker_genes<- program_markers
nMarkers<- 5

## If any genes are top markers for multiple clusters- remove, then find top 10 again
top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = nMarkers) %>% as.data.frame()
if(sum(duplicated(top_markers$gene))!=0){
  duplicated<- top_markers$gene[duplicated(top_markers$gene)]
  print(paste0("Shared top markers: ", paste(duplicated, collapse=",")))
  marker_genes<- marker_genes[!marker_genes$gene %in% duplicated,]
  top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = nMarkers) %>% as.data.frame()
}
all_markers<- top_markers
all_markers$cluster<- as.character(all_markers$cluster)
```



## Plot heatmaps of top marker gene expression
```{r}
rank<- "8"
add_lines<- TRUE
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
grouping_variable<- "detailed_annot" ## variable to pseudobulk by. "Subtype", "Age", "detailed_annot"

top_markers<-all_markers
program_order<- c("Cytotoxic\nNKreceptor", "Predysfunctional", "Inhibitory", "Treg", "Cycling", "IFN\nResponse", "Memory","Effector\nMemory", "Stress")

## Order markers
top_markers$cluster<- factor(top_markers$cluster, levels=gsub("\n", "\\.", program_order))
top_markers<- top_markers[order(top_markers$cluster),]
top_markers$cluster<- as.character(top_markers$cluster)

## Pseudobulk
if(grouping_variable=="Subtype"){
  tcell_seurat$SubtypeEdit<- gsub("Hemispheric-BCOR", "BCOR", gsub("Hemispheric-H3G34R/V", "H3G34R/V", gsub("Midline-H3K27M", "H3K27M",
                           gsub("GBM", "Adult.GBM",gsub("IDH-G", "Adult.IDH-G", tcell_seurat$NiceSubtype)))))
  tcell_seurat$PB_variable<- paste0(tcell_seurat$SubtypeEdit, "_",
                                gsub("\\.", "\n", tcell_seurat@meta.data[[program_name]]))
  figure_append<- "_bySubtype"
}else if(grouping_variable=="Age"){
  tcell_seurat$PB_variable<- paste0(tcell_seurat$Age, "_", 
                                gsub("\\.", "\n", tcell_seurat@meta.data[[program_name]]))
  figure_append<- "_byAge"
} else if(grouping_variable== "detailed_annot"){
  tmp<- paste0(grouping_variable, "_byNMFrank", rank_use, "_mergedBy", sharedBy)
  tcell_seurat$PB_variable<- paste0(tcell_seurat@meta.data[[tmp]], "_", 
                                gsub("\\.", "\n", tcell_seurat@meta.data[[program_name]]))
  figure_append<- "_byDetailedAnnot"
}

## Pseudobulk
pb<- pseudobulk_byVariable(tcell_seurat@assays$RNA@counts, tcell_seurat@meta.data, "PB_variable")

## Remove "Unclear" (for detailed_annot)
pb<-pb[,!grepl("Unclear", colnames(pb))]

## center and score
pb_list<- NormCenter(pb)
pb_center<- pb_list$center_data


## Plot
p<- myHeatmap(pb_center, GOI=top_markers$gene, min.value = -3, max.value = 3, 
              orderFactors =  gsub("\\.", "\n", program_order)) +
  scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")

## Add horizontal lines around each set of nmf genes
if(add_lines){
  tmp<- split(top_markers, f=top_markers$cluster)
  lastGene<- lapply(tmp, function(x){
    x$tmp<- 1:nrow(x)
    x<- x[x$gene %in% rownames(pb), ]
    return(x[max(x$tmp), "gene"])
  })
  linePositions<- as.data.frame(pb_center[top_markers$gene,])
  linePositions$tmp<- 1:nrow(linePositions)
  linePositions<- linePositions[unlist(lastGene),"tmp"]
  linePositions<- linePositions[1:length(linePositions)-1]+ 0.5
  
  p+geom_hline(yintercept = linePositions, linetype="dashed",size=1)
}else{p}

  
ggsave(paste0(figure_dir, "AllTcells_Heatmap_TopNMFgenes_annot_rank", rank,figure_append, "_5Markers.pdf"), device="pdf", width=16, height=18)


```

## UMAP with annotations
```{r}
rank<- "8"
sharedBy<- "union"
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
program_colors_use<- program_colors

DimPlot(tcell_seurat, group.by = program_name)+
  scale_color_manual(values=program_colors_use)+
  ggtitle("")
ggsave(paste0(figure_dir, "AllTcells_UMAP_annot.png"), width=6, height=4)

## Plot each program one at a time
seurat_tmp<- tcell_seurat
all_plots<- lapply(as.character(unique(tcell_seurat@meta.data[[program_name]])), function(x){
  seurat_tmp$tmp<- ifelse(seurat_tmp@meta.data[[program_name]]==x,
                          x, "Other")
  p<-DimPlot(seurat_tmp, group.by = "tmp")+scale_color_manual(values=c(program_colors_use, Other="grey"))+ggtitle(x)+NoLegend()
  return(p)
})
cowplot::plot_grid(plotlist=all_plots, ncol = 2)
ggsave(paste0(figure_dir, "AllTcells_UMAP_indv.png"), width=6, height=13)
```

## Density plots
```{r}
## Read in scores for programs
rank<- "8"
sharedBy<- "union"
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")

## Create df for plotting
df<-tcell_seurat@meta.data[,c("sampleid", "Age", program_name)]
sum(rownames(all_scores)==rownames(df)); nrow(all_scores); nrow(df)
df<- cbind(df, all_scores[,!grepl("Max", colnames(all_scores))] )

## get number of cells per subtype, add to df
nCells<- as.data.frame(table(tcell_seurat$Age))
df$NCells<- plyr::mapvalues(tcell_seurat$Age, nCells$Var1, nCells$Freq)
df$JustAge<- df$Age
df$Age<- paste0(df$Age, "\n", "nCells=",df$NCells)


colors_use<- c(Ped="red", Adult="blue")
names(colors_use)<- plyr::mapvalues(names(colors_use), df$JustAge, df$Age, warn_missing=FALSE)



## Plot each nmf program
all_plots<- list()
for(i in as.character(unique(tcell_seurat@meta.data[[program_name]]))){
  print(i)
  df_tmp<- df[,c(i, "Age")]
  colnames(df_tmp)<- c("score", "Age")
  p<-ggplot(df_tmp, aes(x=score, fill=Age))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.position = "none",
          plot.title = element_text(size=20, face="bold", hjust=0.5))

  all_plots[[i]]<-p
}

## add legend separately
legend<-ggplot(df_tmp, aes(x=score, fill=Age))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.key.size = unit(1.5, 'cm'),
          legend.text = element_text(size=20))
all_plots$legend<-  as_ggplot(get_legend(legend))
cowplot::plot_grid(plotlist = all_plots, ncol=3)

ggsave(paste0(figure_dir, "AllTcells_DensityOfNMFByAge.png"), width=10, height=8)

```


## Programs by age/subtype
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
tcell_seurat$tmp<- tcell_seurat@meta.data[[program_name]]

## By Age
df<- tcell_seurat@meta.data %>% group_by(tmp, Age) %>% summarise(nCells=n()) %>% 
  group_by(Age) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()

ggplot(df, aes(x=Age, y=perCells, fill=tmp))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=program_colors[names(program_colors) %in% unique(df$tmp)])+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"),
        axis.title = element_text(face="bold", color="black"))+
  ylab("Proportion of cells")+xlab("")
ggsave(paste0(figure_dir, "AllTcells_Breakdown_AgeByProgram.png"), width=4, height=4)

## By Subtype, grouped by age, with n cells by subtype added on
df<- tcell_seurat@meta.data %>% group_by(tmp, Age, NiceSubtype) %>% summarise(nCells=n()) %>% 
  group_by(NiceSubtype) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()
nCells<- table(tcell_seurat$NiceSubtype)
df$NiceSubtype<- paste0(df$NiceSubtype, 
                        "\nn=",
                        as.character(plyr::mapvalues(df$NiceSubtype, names(nCells), nCells)))

ggplot(df, aes(x=NiceSubtype, y=perCells, fill=tmp))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=program_colors[names(program_colors) %in% unique(df$tmp)])+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"),
        axis.title = element_text(face="bold", color="black"))+
  ylab("Proportion of cells")+xlab("")+
  facet_grid(cols=vars(Age), scales = "free_x", space="free_x")
ggsave(paste0(figure_dir, "AllTcells_Breakdown_Subtype.AgeByProgram.png"), width=7, height=4)
```

## Proportion of programs by age (integer, actual value)
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
tcell_seurat$tmp<- tcell_seurat@meta.data[[program_name]]

## Proportion of program by sample
df<- tcell_seurat@meta.data %>% group_by(tmp,sampleid) %>% summarise(nCells=n()) %>% 
  group_by(sampleid) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()

## Add on age
df$age<- as.numeric(as.character(plyr::mapvalues(df$sampleid, tcell_seurat$sampleid, tcell_seurat$age_numeric,
                                                 warn_missing = FALSE)))

## Plot
all_plots<-lapply(unique(tcell_seurat$tmp), function(x){
  df_tmp<- df[df$tmp==x,]
  ggplot(df_tmp, aes(x=age, y=perCells, color=tmp))+
    geom_point()+
    scale_color_manual(values=program_colors)+
    theme_bw()+
    theme(legend.position = "none",
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"),
          title = element_text(face="bold", size=12))+
    ggtitle(x)+
    xlab("Age") + ylab(paste0("Proportion\n", x, " cells"))+
    geom_smooth(se=FALSE, method="lm", color="black", linetype="dashed")
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "AllTcells_NumericAge.vs.ProgramProportion.png"), width=10, height=10)
```

## Dotplot of GOI
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
tcell_seurat$tmp<- tcell_seurat@meta.data[[program_name]]
tcell_seurat$tmp<- gsub("Inhibitory.Cycling", "Cycling", tcell_seurat$tmp)

DotPlot(tcell_seurat, group.by = "tmp", features = c("TOP2A","UBE2C", "GNLY", "PRF1","LMNA", "ANXA1","IFIT1", "IFIT3", 
                                                     "CTSH", "CAPG", "CCR7","SELL", "GZMK","PDCD1", "HSPA6","HSPH1", "FOXP3", "TIGIT"))+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  coord_flip()+
  ylab("")+xlab("")
ggsave(paste0(figure_dir, "DotPlot_GOI.png"), width=5.5, height=6)

```

## GO terms
```{r}
## GO term ORA
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
ensembl = useMart("ensembl",dataset="hsapiens_gene_ensembl")
allOE_genes<-getBM(attributes='ensembl_gene_id', 
                   filters = 'hgnc_symbol', 
                   values = rownames(tcell_seurat@assays$RNA@counts), 
                  mart = ensembl)
  
## Run GO: Convert marker genes to ensembl IDs too, then run OR analysis (GO)
go_result = list()
for (i in unique(program_markers$cluster)){
  print(i)
  genes<- program_markers[program_markers$cluster==i, "gene"]
  if(length(genes)!=1){
    sigOE_genes = getBM(attributes='ensembl_gene_id', 
                        filters = 'hgnc_symbol', 
                        values = genes,
                        mart = ensembl)$ensembl_gene_id
    go_result[[i]] = go_analysis(sigOE_genes, allOE_genes)
    }
} 

## Plot and save
go_plots<- plotGO(go_result = go_result, n_terms = 15)
cowplot::plot_grid(plotlist=go_plots, ncol=2)
ggsave(paste0(figure_dir, "GO_NMF.png"), width=18, height=35)
saveRDS(go_result, file=paste0(analysis_dir, "go_result.Rds"))
```

