---
title: "Visualize_programs"
output: html_document
---

## Purpose:
After re-annotating programs based on markers, GO, reference genesets, etc in 01_Visualize.Reannotate_programs, visualize updated programs (CD8 only) here


## Approach:
Using updated annotations, plot nice plots for heatmap, UMAP, density, etc


```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(harmony)
library(ggpubr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/", 
                     "03_Visualize.Reannotate_programs/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/NMF_plots/CD8/")
figure_dir_tmp<- paste0(working_dir, "/figures/NMF_plots/CD8/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in data
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../03_Visualize.Reannotate_programs/analysis/Reannotate_programs/", "rank",rank_use, "_sharedBy", sharedBy, "/" )

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
  
## ped/adult tcell seurat object
cd8_seurat<- readRDS(paste0(data_dir, "ped.adult_seurat_cd8_harmony.Rds"))


## Read in marker genes
program_markers<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                                 "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", sharedBy, "_reAnnot.Rds"))


## Read in scores
all_scores<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/",
                            "scores_merged.ped.adultRank", rank_use, "_specific.sharedBy", sharedBy, "_reAnnot.Rds"))

## Read in colors to use
program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, "_reAnnot.Rds"))
celltype_colors<- readRDS(paste0("../../../plot_colors/Tcell.celltypes.Rds"))
```

# Need to re-annotate BT1478 to non-BCOR; hemispheric, maybe GBM? Set to GBM for now, may update
```{r}
updateBCOR<- TRUE

figure_dir<- ifelse(removeBCOR, paste0(figure_dir, "updateBCOR/"), figure_dir_tmp)
#analysis_dir<- ifelse(removeBCOR, paste0(analysis_dir, "updateBCOR/"), data_dir_tmp)

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
#if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

if(updateBCOR){
  cd8_seurat$NiceSubtype<- ifelse(cd8_seurat$sample=="BT1478", "Hemispheric-HistoneWT", cd8_seurat$NiceSubtype)

}
```

## Subset to top markers for each nmf- top 10
```{r}
marker_genes<- program_markers

## If any genes are top markers for multiple clusters- remove, then find top 10 again
top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
if(sum(duplicated(top_markers$gene))!=0){
  duplicated<- top_markers$gene[duplicated(top_markers$gene)]
  print(paste0("Shared top markers: ", paste(duplicated, collapse=",")))
  marker_genes<- marker_genes[!marker_genes$gene %in% duplicated,]
  top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
}
all_markers<- top_markers
all_markers$cluster<- as.character(all_markers$cluster)
```



## Plot heatmaps of top marker gene expression
```{r}
rank<- "8"
sharedBy<- "union"
add_lines<- TRUE
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
bySubtype<- TRUE

top_markers<-all_markers[all_markers$cluster!="Treg",]

## Pseudobulk
if(bySubtype){
  cd8_seurat$SubtypeEdit<- gsub("Hemispheric-BCOR", "BCOR", gsub("Hemispheric-H3G34R/V", "H3G34R/V", gsub("Midline-H3K27M", "H3K27M",
                           gsub("GBM", "Adult.GBM",gsub("IDH-G", "Adult.IDH-G", cd8_seurat$NiceSubtype)))))
  cd8_seurat$PB_variable<- paste0(cd8_seurat$SubtypeEdit, "_",
                                gsub("\\.", "\n", cd8_seurat@meta.data[[program_name]]))
  figure_append<- "_bySubtype"
}else{
  cd8_seurat$PB_variable<- paste0(cd8_seurat$Age, "_", 
                                gsub("\\.", "\n", cd8_seurat@meta.data[[program_name]]))
  figure_append<- ""
}

pb<- pseudobulk_byVariable(cd8_seurat@assays$RNA@counts, cd8_seurat@meta.data, "PB_variable")

## center and score
pb_list<- NormCenter(pb)
pb_center<- pb_list$center_data

## Plot
p<- myHeatmap(pb_center, GOI=top_markers$gene, min.value = -3, max.value = 3, 
              orderFactors =  gsub("\\.", "\n", unique(top_markers$cluster))) +
  scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")

## Add horizontal lines around each set of nmf genes
if(add_lines){
  tmp<- split(top_markers, f=top_markers$cluster)
  lastGene<- lapply(tmp, function(x){
    x$tmp<- 1:nrow(x)
    x<- x[x$gene %in% rownames(pb), ]
    return(x[max(x$tmp), "gene"])
  })
  linePositions<- as.data.frame(pb_center[top_markers$gene,])
  linePositions$tmp<- 1:nrow(linePositions)
  linePositions<- linePositions[unlist(lastGene),"tmp"]
  linePositions<- linePositions[1:length(linePositions)-1]+ 0.5
  
  p+geom_hline(yintercept = linePositions, linetype="dashed",size=1)
}else{p}

  
ggsave(paste0(figure_dir, "CD8_Heatmap_TopNMFgenes_annot_rank", rank,figure_append, ".png"), width=17, height=11)


```

## UMAP with annotations
```{r}
rank<- "8"
sharedBy<- "union"
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
program_colors_use<- program_colors[names(program_colors) %in% unique(cd8_seurat@meta.data[[program_name]])]

DimPlot(cd8_seurat, group.by = program_name)+
  scale_color_manual(values=program_colors_use)+
  ggtitle("")
ggsave(paste0(figure_dir, "CD8_UMAP_annot.pdf"), device="pdf", width=6, height=4)

## Plot each program one at a time
seurat_tmp<- cd8_seurat
all_plots<- lapply(as.character(unique(cd8_seurat@meta.data[[program_name]])), function(x){
  seurat_tmp$tmp<- ifelse(seurat_tmp@meta.data[[program_name]]==x,
                          x, "Other")
  p<-DimPlot(seurat_tmp, group.by = "tmp")+scale_color_manual(values=c(program_colors_use, Other="grey"))+ggtitle(x)+NoLegend()
  return(p)
})
cowplot::plot_grid(plotlist=all_plots, ncol = 2)
ggsave(paste0(figure_dir, "CD8_UMAP_indv.png"), width=6, height=10)
```

## Density plots
```{r}
## Read in scores for programs
rank<- "8"
sharedBy<- "union"
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")

## Create df for plotting
df<-cd8_seurat@meta.data[,c("sampleid", "Age", program_name)]
scores_use<- all_scores[rownames(all_scores) %in% rownames(df),]
sum(rownames(scores_use) == rownames(df)); nrow(scores_use); nrow(df)
df<- cbind(df, scores_use[,!grepl("Max", colnames(scores_use))] )

## get number of cells per subtype, add to df
nCells<- as.data.frame(table(cd8_seurat$Age))
df$NCells<- plyr::mapvalues(cd8_seurat$Age, nCells$Var1, nCells$Freq)
df$JustAge<- df$Age
df$Age<- paste0(df$Age, "\n", "nCells=",df$NCells)


colors_use<- c(Ped="red", Adult="blue")
names(colors_use)<- plyr::mapvalues(names(colors_use), df$JustAge, df$Age, warn_missing=FALSE)



## Plot each nmf program
all_plots<- list()
for(i in as.character(unique(cd8_seurat@meta.data[[program_name]]))){
  print(i)
  df_tmp<- df[,c(i, "Age")]
  colnames(df_tmp)<- c("score", "Age")
  p<-ggplot(df_tmp, aes(x=score, fill=Age))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.position = "none",
          plot.title = element_text(size=20, face="bold", hjust=0.5))

  all_plots[[i]]<-p
}

## add legend separately
legend<-ggplot(df_tmp, aes(x=score, fill=Age))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.key.size = unit(1.5, 'cm'),
          legend.text = element_text(size=20))
all_plots$legend<-  as_ggplot(get_legend(legend))
cowplot::plot_grid(plotlist = all_plots, ncol=3)

ggsave(paste0(figure_dir, "CD8_DensityOfNMFByAge.png"), width=10, height=6)

```

## Programs by age/subtype
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
cd8_seurat$tmp<- cd8_seurat@meta.data[[program_name]]

## By Age
df<- cd8_seurat@meta.data %>% group_by(tmp, Age) %>% summarise(nCells=n()) %>% 
  group_by(Age) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()

ggplot(df, aes(x=Age, y=perCells, fill=tmp))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=program_colors[names(program_colors) %in% unique(df$tmp)])+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"),
        axis.title = element_text(face="bold", color="black"))+
  ylab("Proportion of cells")+xlab("")
ggsave(paste0(figure_dir, "CD8_Breakdown_AgeByProgram.png"), width=4, height=4)


## By Age- under 4, 4-25, over 25
cd8_seurat$age_numeric<- as.numeric(cd8_seurat$age_numeric)
cd8_seurat$Age_group<- ifelse(cd8_seurat$age_numeric<4, "under4",
                              ifelse(cd8_seurat$age_numeric>=4 & cd8_seurat$age_numeric <= 25, "age4-25", "over25"))
df<- cd8_seurat@meta.data %>% group_by(tmp, Age_group) %>% summarise(nCells=n()) %>% 
  group_by(Age_group) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()
df$Age_group<- factor(df$Age_group, levels=c("under4", "age4-25", "over25"))

ggplot(df, aes(x=Age_group, y=perCells, fill=tmp))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=program_colors[names(program_colors) %in% unique(df$tmp)])+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"),
        axis.title = element_text(face="bold", color="black"))+
  ylab("Proportion of cells")+xlab("")
ggsave(paste0(figure_dir, "CD8_Breakdown_Age.under4.ByProgram.png"), width=4, height=4)

## By Subtype, grouped by age, with n cells by subtype added on
df<- cd8_seurat@meta.data %>% group_by(tmp, Age, NiceSubtype) %>% summarise(nCells=n()) %>% 
  group_by(NiceSubtype) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()
nCells<- table(cd8_seurat$NiceSubtype)
df$NiceSubtype<- paste0(df$NiceSubtype, 
                        "\nn=",
                        as.character(plyr::mapvalues(df$NiceSubtype, names(nCells), nCells)))

ggplot(df, aes(x=NiceSubtype, y=perCells, fill=tmp))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=program_colors[names(program_colors) %in% unique(df$tmp)])+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"),
        axis.title = element_text(face="bold", color="black"))+
  ylab("Proportion of cells")+xlab("")+
  facet_grid(cols=vars(Age), scales = "free_x", space="free_x")
ggsave(paste0(figure_dir, "CD8_Breakdown_Subtype.AgeByProgram.png"), width=7, height=4)
```

## Proportion of programs by age (integer, actual value)
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
cd8_seurat$tmp<- cd8_seurat@meta.data[[program_name]]

## Proportion of program by sample
df<- cd8_seurat@meta.data %>% group_by(tmp,sampleid) %>% summarise(nCells=n()) %>% 
  group_by(sampleid) %>% mutate(perCells=nCells/sum(nCells)) %>% as.data.frame()

## Add on age
df$age<- as.numeric(as.character(plyr::mapvalues(df$sampleid, cd8_seurat$sampleid, cd8_seurat$age_numeric,
                                                 warn_missing = FALSE)))

## Plot
all_plots<-lapply(unique(cd8_seurat$tmp), function(x){
  df_tmp<- df[df$tmp==x,]
  
  ## model
  model<- summary(lm(age~perCells, df_tmp))
  r2<- round(model$r.squared,2)
  pvalue=round(model$coefficients["perCells","Pr(>|t|)" ],4)
  
  ## Plot    
  ggplot(df_tmp, aes(x=age, y=perCells, color=tmp))+
    geom_point()+
    scale_color_manual(values=program_colors)+
    theme_bw()+
    theme(legend.position = "none",
          axis.text = element_text(face="bold", color="black"),
          axis.title = element_text(face="bold", color="black"),
          title = element_text(face="bold", size=12))+
    ggtitle(x, subtitle =  paste0("R2: ", r2, ", pvalue: ", pvalue))+
    xlab("Age") + ylab(paste0("Proportion\n", x, " cells"))+
    geom_smooth(se=FALSE, method="lm", color="black", linetype="dashed")
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "CD8_NumericAge.vs.ProgramProportion.png"), width=10, height=10)
```


## Plot GOI+ Nature2020 genesets (and individual GOI from these genesets) for each program
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")


## Order programs
cd8_seurat$tmp<- factor(cd8_seurat@meta.data[[program_name]], levels=c( "Memory","Cycling", "Predysfunctional", "Cytotoxic.NKreceptor",
                                                                       "Stress", "Effector.Memory", "IFN.Response"))

## Nature2020 scores + subset of GOI
DotPlot(cd8_seurat, features=c("naive", "dysfunctional",  "predysfunctional","cytotoxic",
                               "ENTPD1", "ITGAE", "TNFRSF9", "CCL4L1", "PDCD1", "CTLA4", "LAG3", "GZMK", "GNLY"),
        group.by = "tmp")+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"))+
  ylab("") + xlab("")+
  coord_flip()
ggsave(paste0(figure_dir, "CD8_DotPlot_Nature2020.GOI_byProgram.png"), width=6, height=6)

## Inhibitory GOI
DotPlot(cd8_seurat, features=c( "ITGAE", "TNFRSF9",  "PDCD1", "CTLA4", "LAG3", "TIGIT", "KLRB1", "HLA-DRA"),
        group.by = "tmp")+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"))+
  ylab("") + xlab("")+
  coord_flip()
ggsave(paste0(figure_dir, "CD8_DotPlot_Inhibitory.GOI_byProgram.png"), width=6, height=5)
```



## Plot scores for Nature2020 genesets (cytotoxic, dysfunctional, naive, predysfunctional), colored by program
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")

cd8_seurat$tmp<- cd8_seurat@meta.data[[program_name]]

df<- cd8_seurat@meta.data[,c("tmp", c("naive", "dysfunctional",  "predysfunctional","cytotoxic"))]

## option 1: cyto/naive on 1 scale, predys/dys on the other scale
df$scale1<- df$cytotoxic-df$naive
df$scale2<- df$dysfunctional - df$predysfunctional
ggplot(df, aes(x=scale1, y=scale2, color=tmp))+
  geom_point()+
  xlab("high naive ----------> high cytotoxic")+
  ylab("high predysfunctional ---------->  high dysfunctional")+
  scale_color_manual(values=program_colors)

 ## color separately for each program
  all_plots<- lapply(unique(df$tmp), function(x){
    df_tmp<- df
    df_tmp$tmp<- ifelse(df_tmp$tmp==x, x, "Other")
    ggplot(df_tmp, aes(x=scale1, y=scale2, color=tmp, alpha=tmp))+
      geom_point()+
      theme_bw()+
      xlab("high naive ----------> high cytotoxic")+
      ylab("high predysfunctional ---------->  high dysfunctional")+
      scale_color_manual(values=c(program_colors, Other="grey"))+
      scale_alpha_manual(values=c(Other=0.2))+
      theme(legend.position = "none",
            axis.text = element_text(face="bold", color="black"),
            axis.title = element_text(face="bold", color="black", size=7),
            title = element_text(face="bold", size=12))+
      ggtitle(x)
  })
  cowplot::plot_grid(plotlist=all_plots, ncol=3)
  ggsave(paste0(figure_dir, "CD8_Nature2020Genesets.NaivCytoDysPredys_scoresByCell_indv.png"), width=9, height = 9)

## Option 2: no naive score, just cyto on 1 scale, predys/dys on other scale
df$scale1<- df$cytotoxic
df$scale2<- df$dysfunctional - df$predysfunctional
ggplot(df, aes(x=scale1, y=scale2, color=tmp))+
  geom_point()+
  xlab("low cytotoxic ----------> high cytotoxic")+
  ylab("high predysfunctional ---------->  high dysfunctional")+
  scale_color_manual(values=program_colors)
  
  ## color separately for each program
  all_plots<- lapply(unique(df$tmp), function(x){
    df_tmp<- df
    df_tmp$tmp<- ifelse(df_tmp$tmp==x, x, "Other")
    ggplot(df_tmp, aes(x=scale1, y=scale2, color=tmp, alpha=tmp))+
      geom_point()+
      theme_bw()+
      xlab("low cytotoxic ----------> high cytotoxic")+
      ylab("high predysfunctional ---------->  high dysfunctional")+
      scale_color_manual(values=c(program_colors, Other="grey"))+
      scale_alpha_manual(values=c(Other=0.2))+
      theme(legend.position = "none",
            axis.text = element_text(face="bold", color="black"),
            axis.title = element_text(face="bold", color="black", size=7),
            title = element_text(face="bold", size=12))+
      ggtitle(x)
  })
  cowplot::plot_grid(plotlist=all_plots, ncol=3)
  ggsave(paste0(figure_dir, "CD8_Nature2020Genesets.CytoDysPredys_scoresByCell_indv.png"), width=9, height = 9)
  
## Option 3: only predysfunction/cytotoxic
df$scale2<- df$cytotoxic
df$scale1<- df$predysfunctional
ggplot(df, aes(x=scale1, y=scale2, color=tmp))+
  geom_point()+
  theme_bw()+
  ylab("low cytotoxic ----------> high cytotoxic")+
  xlab("low predysfunctional ---------->  high predysfunctional")+
  scale_color_manual(values=program_colors)+
  theme(axis.text = element_text(face="bold", color="black"),
        axis.title = element_text(face="bold", color="black", size=7),
        title = element_text(face="bold", size=12))

  ## color separately for each program
  all_plots<- lapply(unique(df$tmp), function(x){
    df_tmp<- df
    df_tmp$tmp<- ifelse(df_tmp$tmp==x, x, "Other")
    ggplot(df_tmp, aes(x=scale1, y=scale2, color=tmp, alpha=tmp))+
      geom_point()+
      theme_bw()+
      ylab("low cytotoxic ----------> high cytotoxic")+
      xlab("low predysfunctional ---------->  high predysfunctional")+
      scale_color_manual(values=c(program_colors, Other="grey"))+
      scale_alpha_manual(values=c(Other=0.2))+
      theme(legend.position = "none",
            axis.text = element_text(face="bold", color="black"),
            axis.title = element_text(face="bold", color="black", size=7),
            title = element_text(face="bold", size=12))+
      ggtitle(x)
  })
  cowplot::plot_grid(plotlist=all_plots, ncol=3)
  ggsave(paste0(figure_dir, "CD8_Nature2020Genesets.CytoPredys_scoresByCell_indv.png"), width=9, height = 9)
  
  
```

## Plot by clonotype- within clonal cells, is clonotype shared more frequently between cyto/predys (within a sample) compared to other pairs?
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")

## Read in clonotype info
  ## Pediatric- from Orr
  load(paste0("../../../data/data_FromOrr/", "2020_12_10_clonalTcells_cluster.Rda"))
  ped_clonotypes<- clonotypes.all; rm(clonal); rm(clonotypes.all); rm(clonotypes.cells); rm(tcell.reconstructed); rm(gcdata)
  ped_clonotypes$cell_name<- gsub("MUV", "MUV0", ped_clonotypes$cell_name)
  ped_clonotype_list<- ped_clonotypes$clonal_group; names(ped_clonotype_list)<- ped_clonotypes$cell_name
  
  ## Adult
  load("../../../data/Adult/tcells_seurat_obj.Robj")
  adult_clonotypes<- seurat_obj$clonal_group; names(adult_clonotypes)<-colnames(seurat_obj)
  adult_clonotypes<- na.omit(adult_clonotypes)
  
  all_clonotypes<- c(ped_clonotype_list, adult_clonotypes)
  

## Subset to just clonal cells
clonal_cd8<- subset(cd8_seurat, clonal=="clonal")

## Add clonotype info
clonal_cd8$clonotype<- plyr::mapvalues(colnames(clonal_cd8), names(all_clonotypes), all_clonotypes, warn_missing = FALSE)

## A couple (26) are clonal but don't have clonotype info. Just remove
clonal_meta<- clonal_cd8@meta.data
clonal_meta$program<-clonal_meta[[program_name]]
clonal_meta<- clonal_meta[clonal_meta$clonotype != rownames(clonal_meta),]
clonal_meta<- clonal_meta[,c("sampleid", "program", "clonotype")]

## Rename Memory and Inhibitory to distinguish them from EffectorMemory and Predysfunctional
clonal_meta$program<-ifelse(clonal_meta$program=="Memory", gsub("Memory", "MemoryOnly", clonal_meta$program),
                     ifelse(clonal_meta$program=="Inhibitory", gsub("Inhibitory", "InhibitoryOnly", clonal_meta$program),
                            clonal_meta$program))

## Complete list of all possible pairwise combinations of NMF programs
programs<- as.character(unique(clonal_meta$program))
all_combinations<-with(subset(expand.grid(programs,programs),Var1!=Var2),paste0(Var1,'_',Var2))
all_combinations<- all_combinations[order(all_combinations)]

## Add unique ID: clonotype + sample (for rare case where multiple samples are in same clonotype)
clonal_meta$ID<- paste0(clonal_meta$sampleid, "_", clonal_meta$clonotype)

## For each unique ID (clonotype), determine which pairs of NMF programs occur
## There may be more than 2 programs in a clonotype of course, but all pairs are counted separately

  ## Group by clonotype, remove any clonotypes with just 1 program, get list of all programs within that clonotype
  sharedClonotypePrograms<- clonal_meta %>% group_by(ID) %>% mutate(numberUnique=length(unique(program))) %>%
    filter(numberUnique>1) %>%
    summarise(sharedPrograms=paste(unique(program),collapse =  "_")) %>% as.data.frame()
  clonotype_programs<- sharedClonotypePrograms$sharedPrograms
  
  ## Cycle through all pairwise combinations, adding this sample name + number of occurences
  all_occurences<- lapply(all_combinations, function(x){
    programs_to_count<- unlist(strsplit(x, split="_"))
    in_clonotype<- (grepl(programs_to_count[1], clonotype_programs, fixed=TRUE) & 
                      grepl(programs_to_count[2], clonotype_programs, fixed = TRUE))
    return(sum(in_clonotype))
  })
  names(all_occurences)<- all_combinations
  total_in_clonotypes<- as.data.frame(t(as.data.frame(all_occurences)))
  
  total_in_clonotypes$Pair<- rownames(total_in_clonotypes)
  colnames(total_in_clonotypes)<-c("Total", "Pair")
  
## Split names back into separate columns for cellTypeA, cellTypeB 
total_in_clonotypes$P1<- sapply(total_in_clonotypes$Pair, function(x){unlist(strsplit(x, split="_"))[1]})
total_in_clonotypes$P2<- sapply(total_in_clonotypes$Pair, function(x){unlist(strsplit(x, split="_"))[2]})

## Normalize by the number of clonotypes that the pair occurs in- use the minimum
## For example: Pair A_B occurs 10 times. CelltypeA is in 5 clonotypes, CellTypeB is in 2 clonotypes- use 2 to normalize
  ## Get number of clonotypes each program occurs in
  nClonotypes<- clonal_meta %>% group_by(program) %>% summarise(nClonotypes=length(unique(ID))) %>% as.data.frame()
  
  ## For each pair, determine the MINIMUM number of clonotypes one member of the pair occurs in
  ## This value is the MAXIMUM number of clonotypes that the pair can occur in
  total_in_clonotypes$minClonotype<- apply(total_in_clonotypes, 1, function(x){
    min(as.numeric(as.character(plyr::mapvalues(x["P1"], nClonotypes$program, nClonotypes$nClonotypes,
                                                warn_missing = FALSE))),
        as.numeric(as.character(plyr::mapvalues(x["P2"], nClonotypes$program, nClonotypes$nClonotypes,
                                                warn_missing = FALSE))))
  })
  
  ## Convert number of clonotypes per pair from NUMBER of occurences to PERCENTAGE of total maximum
  total_in_clonotypes$ProportionOfClonotypes<- total_in_clonotypes$Total/total_in_clonotypes$minClonotype
  
## Plot the number of occurences of each program pair within a clonotype
ggplot(total_in_clonotypes, aes(x=P1, y=P2, fill=ProportionOfClonotypes))+
  geom_tile()+
  theme_classic()+
  scale_fill_gradient(low="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(face="bold", color="black"))+
  ylab("")+xlab("")+
  ggtitle("Programs shared within same clonotype\n(sample-wise)")
ggsave(paste0(figure_dir, "CD8_NumberOfOccurences_programPairs_withinClonotype_normByPropClonotypes.png"),width=6, height=3.5)


```


