---
title: "Visualize_reannotate_programs"
output: html_document
---

## Purpose:
NMF programs were generated based on 1) de novo NMF separately for ped/adult Tcells 2) correlating programs 3) keeping "specific" programs separate, but merging "shared" programs --> score --> assign max -->regenerate gene lists. These programs are annotated based on correlation with original ped cd4/cd8 programs (no longer in use, archived). Some of these annotations make sense, but others (like "CD4_Inhibitory") don't. Here, basic visualizations of these programs, comparing ped/adult proportions, GO terms, marker genes- all used to help re-annotate programs as needed


## Approach:
First, just want to do basic visualizations (proportion of programs in cd4/cd8), then compare differences in ped/adult proportions within CD4/CD8. Then various visualizations to help annotate: GO terms, marker genes, etc. Re-annotate programs as needed, then update seurat objects + marker gene lists as needed to reflect this


```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(harmony)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/02_Ped.Adult/", 
                     "03_Visualize.Reannotate_programs/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/VisualizeProgramsIndv/")
figure_dir_tmp<- paste0(working_dir, "/figures/VisualizeProgramsIndv/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```


## Read in ped tcell seurat object. Contains NMF assignments (annots added in deNovoPedTcell.v.deNovoAdultTcell, based on correlation with adult T cell + correlation with ped cd4/cd8 programs)
```{r}
rank_use<-8
sharedBy<- "union" 
data_dir<- paste0("../03_Visualize.Reannotate_programs/analysis/Reannotate_programs/", "rank",rank_use, "_sharedBy", sharedBy, "/" )

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
  
## ped/adult tcell seurat objects
cd4_seurat<- readRDS(paste0(data_dir, "ped.adult_seurat_cd4_harmony.Rds"))
cd8_seurat<-readRDS(paste0(data_dir, "ped.adult_seurat_cd8_harmony.Rds"))
tcell_seurat<- readRDS(paste0(data_dir, "ped.adult_seurat_tcell_harmony.Rds"))

## Read in marker genes
program_markers<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                                 "new_markers_NMFrank", rank_use, "_bySeurat_sharedBy", sharedBy, "_reAnnot.Rds"))

## Read in scores
all_scores<- readRDS(paste0("../01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/",
                            "scores_merged.ped.adultRank", rank_use, "_specific.sharedBy", sharedBy, "_reAnnot.Rds"))

## Read in colors to use
program_colors<- readRDS(paste0("../../../plot_colors/Tcell.programs_rank", rank_use, "sharedBy", sharedBy, "_reAnnot.Rds"))
celltype_colors<- readRDS(paste0("../../../plot_colors/Tcell.celltypes.Rds"))


## list of Nature 2020 markers
all_markers<- list(cytotoxic= c("CX3CR1", "PRF1", "GZMA", "GZMB", "GZMH",  "FGFBP2","KLRG1", "FCGR3A"),
                   naive=c("TCF7", "CCR7", "SELL", "LEF1"),
                   predysfunctional=c("GZMK", "PDCD1", "CXCR3"),
                   dysfunctional= c("HAVCR2", "GZMB","IFNG", "CXCL13", "TIGIT", "LAYN", "LAG3", 
                                    "PDCD1", "CTLA4", "ITGAE", "ENTPD1")) 

cycling_genes<- read.table(paste0(base_dir,"Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/",
                                "Marker_genes/regev_lab_cell_cycle_genes.txt"))
```


## Preprocessing- only run once
## Cycle through each seurat list for each program, adding on detailed_annot from full seurat and splitting to CD4/CD8
```{r}
## Read in list used for detailed annotation (includes updated annotations)
seurat_list<- readRDS(paste0("../02_detailed_annot/analysis/IdentifyCD4.CD8_byNMF/rank", rank_use, "_sharedBy", sharedBy,"/seurat_list_splitByNMFrank8.Rds"))

annot_name<- paste0("detailed_annot_byNMFrank", rank_use, "_mergedBy", sharedBy)
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")

## For renaming genelist after AddModuleScore
rename_cols<-paste0(names(all_markers), 1:4); names(rename_cols)<-names(all_markers)



## Cycle through each seurat list for each program, adding on detailed_annot from full seurat and splitting to CD4/CD8
new_seurat_list<- lapply(names(seurat_list), function(program){
  x<- seurat_list[[program]]
  x@meta.data[[program_name]]<-program
  print(unique(x@meta.data[[program_name]]))
  x$detailed_annotByNMF<- plyr::mapvalues(colnames(x), colnames(tcell_seurat), tcell_seurat@meta.data[[annot_name]],
                                          warn_missing = FALSE)
  x<- AddModuleScore(x, all_markers, name=names(all_markers))
  colnames(x@meta.data)<- plyr::mapvalues(colnames(x@meta.data), rename_cols, names(rename_cols))
  if(sum(x$detailed_annotByNMF=="CD4")>0){
    cd4_tmp<- subset(x, detailed_annot=="CD4"); cd4_meta<- cd4_tmp@meta.data
    cd4_tmp<- RunFullSeurat(cd4_tmp@assays$RNA@counts, RunHarmony = TRUE, samples=cd4_meta$sampleid)
    cd4_tmp<- AddMetaData(cd4_tmp, cd4_meta[,!colnames(cd4_meta) %in% colnames(cd4_tmp@meta.data)])
  }else{cd4_tmp<-NULL}
  
  if(sum(x$detailed_annotByNMF=="CD8")>0){
    cd8_tmp<- subset(x, detailed_annot=="CD8"); cd8_meta<- cd8_tmp@meta.data
    cd8_tmp<- RunFullSeurat(cd8_tmp@assays$RNA@counts, RunHarmony = TRUE, samples=cd8_meta$sampleid)
    cd8_tmp<- AddMetaData(cd8_tmp, cd8_meta[,!colnames(cd8_meta) %in% colnames(cd8_tmp@meta.data)])
  }else{cd8_tmp<-NULL}
  new_list<- list(cd4=cd4_tmp, cd8=cd8_tmp)
  return(new_list)
  
})
names(new_seurat_list)<-names(seurat_list)
saveRDS(new_seurat_list, file=paste0(analysis_dir, "seurat_list_splitByNMF.CD4.CD8.Rds"))
```

## Read in processed seurat list, split by both program and cd4/cd8
```{r}
new_seurat_list<- readRDS(paste0(analysis_dir, "seurat_list_splitByNMF.CD4.CD8.Rds"))
```

## Plot Nature 2020 markers individually for each program: CD8
```{r}
## Plot each cd8 seurat module socres
cd8_seurat_list<- lapply(new_seurat_list, function(x){x$cd8})
cd8_seurat_list<- cd8_seurat_list[unlist(lapply(cd8_seurat_list, function(x){!is.null(x)}))]
cd8_plots<- lapply(cd8_seurat_list, function(x){
  program<- unique(x@meta.data[[program_name]])
  all_plots<- lapply(names(all_markers), function(y){
    FeaturePlot(x, features=y)+scale_color_gradient2(low="blue", mid="white", high="red")+
      ggtitle(paste0(program, "\n", y))
  })
  p<- cowplot::plot_grid(plotlist=all_plots, ncol=2)
  return(p)
})
cowplot::plot_grid(plotlist=cd8_plots, ncol=2)
ggsave(paste0(figure_dir, "CD8_ByProgram_Nature2020ProgramScores.png"), width=12, height=24)

## To support CD8 scores- expression of expected marker genes in each program
cd8_plots<- lapply(cd8_seurat_list, function(x){
  program<- unique(x@meta.data[[program_name]])
  all_plots<- lapply(c("GNLY", "CCR7", "CCL4", "PDCD1"), function(y){
    FeaturePlot(x, features=y)+scale_color_gradient(low="lightgrey", high="red")+
      ggtitle(paste0(program, "\n", y))
  })
  p<- cowplot::plot_grid(plotlist=all_plots, ncol=2)
  return(p)
})
cowplot::plot_grid(plotlist=cd8_plots, ncol=2)
ggsave(paste0(figure_dir, "CD8_ByProgram_MarkerGenes.png"), width=12, height=24)

## To support CD8 scores- ORIGINAL (pre-merging) NMF annotations
cd8_plots<- lapply(cd8_seurat_list, function(x){
  DimPlot(x, group.by = "NMF_rank8_annot")+ggtitle(unique(x@meta.data[[program_name]]))
})
cowplot::plot_grid(plotlist=cd8_plots, ncol=3)
ggsave(paste0(figure_dir, "CD8_ByProgram_OriginalRank8Annots.png"), width=12, height=8)
```


## Plot Nature 2020 markers individually for each program: CD4
```{r}
## Plot each cd4 seurat module socres
cd4_seurat_list<- lapply(new_seurat_list, function(x){x$cd4})
cd4_seurat_list<- cd4_seurat_list[unlist(lapply(cd4_seurat_list, function(x){!is.null(x)}))]
cd4_plots<- lapply(cd4_seurat_list, function(x){
  program<- unique(x@meta.data[[program_name]])
  all_plots<- lapply(names(all_markers), function(y){
    FeaturePlot(x, features=y)+scale_color_gradient2(low="blue", mid="white", high="red")+
      ggtitle(paste0(program, "\n", y))
  })
  p<- cowplot::plot_grid(plotlist=all_plots, ncol=2)
  return(p)
})
cowplot::plot_grid(plotlist=cd4_plots, ncol=2)
ggsave(paste0(figure_dir, "CD4_ByProgram_Nature2020ProgramScores.png"), width=12, height=24)

```


## Subset CD8 to just cytotoxic/inhibitory programs, rerun clustering. Any pattern in markers?
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot_mergedBy", sharedBy, "_reAnnot")
subset_to<- c("Cytotoxic.NKreceptor", "Predysfunctional", "Cycling")
remove_cycling<- FALSE
analysis_name<- "Cyto.Predys.Cyc"

## Load in or generate file- re-clustered seurat with programs in subset_to
file<-paste0(analysis_dir, "cd8_seurat_",analysis_name, ".Rds")
if(file.exists(file)){
  cd8_cyto.inh<- readRDS(paste0(analysis_dir, "cd8_seurat_",analysis_name, ".Rds"))
} else{
  ## Subset to programs of interest
  cd8_seurat$tmp<- cd8_seurat@meta.data[[program_name]]
  cd8_cyto.inh<- subset(cd8_seurat, tmp %in% subset_to)
  cd8_cyto.inh_meta<- cd8_cyto.inh@meta.data
  
  ## Remove cycling genes if desired
  cm<- cd8_cyto.inh@assays$RNA@counts
  if(remove_cycling){cm<- cm[!rownames(cm) %in% cycling_genes$V1,]}else{cm<- cm}
  
  ## rerun clustering
  cd8_cyto.inh<- RunFullSeurat(cm, RunHarmony = TRUE, samples=cd8_cyto.inh$sampleid)
  cd8_cyto.inh<-AddMetaData(cd8_cyto.inh, cd8_cyto.inh_meta[,!colnames(cd8_cyto.inh_meta) %in% colnames(cd8_cyto.inh@meta.data)])
  
  saveRDS(cd8_cyto.inh, file=paste0(analysis_dir, "cd8_seurat_",analysis_name, ".Rds"))
}


## Plot by program
DimPlot(cd8_cyto.inh, group.by = program_name, reduction="umap")+
  scale_color_manual(values=program_colors[names(program_colors) %in% unique(cd8_cyto.inh@meta.data[[program_name]])])
ggsave(paste0(figure_dir, "CD8_",analysis_name, "_UMAP_programs.png"), width=6, height=4.5)

colors_use<- program_colors[names(program_colors) %in% unique(cd8_cyto.inh@meta.data[[program_name]])]
colors_use<- c(Cycling="navy", Predysfunctional="lightsalmon3", Cytotoxic.NKreceptor="greenyellow")
DimPlot(cd8_cyto.inh, group.by = program_name, reduction="pca")+
  scale_color_manual(values=colors_use)
ggsave(paste0(figure_dir, "CD8_",analysis_name, "_PCA_programs.png"), width=6, height=4.5)

## Plot by scores for Nature2020 genesets
FeaturePlot(cd8_cyto.inh, features=names(all_markers), reduction = "pca")
ggsave(paste0(figure_dir, "CD8_",analysis_name, "_PCA_Nature2020.dysfunction.cyto.genesets.png"), width=9, height=9)

## Plot by genes of interest
FeaturePlot(cd8_cyto.inh, features=c("GNLY","GZMB","PRF1",
                                     "PDCD1", "CTLA4", "LAG3","GZMK", "TIGIT","ITGAE","TNFRSF9",  "EOMES" ,  "GZMK", 
                                     "CCL4","CCL4L1", "CCL4L2", 
                                     "KLRB1", "SELPLG",  "ANXA1", "PTGER2",
                                     "ENTPD1") ,
            reduction="pca")
ggsave(paste0(figure_dir, "CD8_",analysis_name, "_PCA_dysfunction.cyto.markers.png"), width=12, height=15)

## Loading genes for PC2? Based on expression pattern, high CCL4L1 expression- others?
pca_loadings<- as.data.frame(Loadings(cd8_cyto.inh, reduction = "pca")[, 1:2])

  ## Top genes contributing to PC1/PC2
  pca_loadings_pos<- lapply(colnames(pca_loadings), function(x){
    tmp<- pca_loadings[,x]
    names(tmp)<- rownames(pca_loadings)
    tmp<- tmp[order(abs(tmp), decreasing = TRUE)]
    return(tmp)
  })
  names(pca_loadings_pos)<- colnames(pca_loadings)
  
  ## Plot PC2 genes
  FeaturePlot(cd8_cyto.inh, features=names(pca_loadings_pos$PC_2)[1:9],reduction = "pca")
  ggsave(paste0(figure_dir, "PC2.loading_CD8_",analysis_name, "_PCA_dysfunction.cyto.markers.png"), 
         width=12, height=9)
  
  ## Plot PC1 genes
  FeaturePlot(cd8_cyto.inh, features=names(pca_loadings_pos$PC_1)[1:9],reduction = "pca")
  ggsave(paste0(figure_dir, "PC1.loading_CD8_",analysis_name, "_PCA_dysfunction.cyto.markers.png"), 
         width=12, height=9)
  
## When removing cycling genes: what are markers of inhibitory.Cycling now?
#cd8_cyto.inh<- SetIdent(cd8_cyto.inh, value = cd8_cyto.inh$tmp)
#all_markers<- FindAllMarkers(cd8_cyto.inh, min.pct = 0.2)
#all_markers_filtered<- all_markers[all_markers$p_val_adj<0.05 & all_markers$avg_log2FC>0.2,]
#top_n<- all_markers_filtered %>% group_by(cluster) %>% top_n(n=20, wt=avg_log2FC) %>% as.data.frame()


```


