---
title: "PrepSeuratObjectsForCellChat"
output: html_document
---

```{r}
library(CellChat)
library(Seurat) 

options(stringsAsFactors = FALSE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/PD1_Treat/SS2/Tumor/03_CellChat/")

analysis_dir<- paste0(working_dir, "/analysis/")
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
```

## Read in seurat objects: ped tcells, myeloid cells, tumor cells
```{r}
tumor_seurat<- readRDS("../02_Annotate/analysis/seurat/seurat_obj.Rds")
cd4_seurat<- readRDS("../../../comparisonToUntreated/analysis/ss_cd4_Harmony_seurat.Rds")
cd8_seurat<- readRDS("../../../comparisonToUntreated/analysis/ss_cd8_Harmony_seurat.Rds")
myeloid_seurat<- readRDS("../../../comparisonToUntreated/analysis/ss_myeloid_Harmony_seurat.Rds")

## Remove non-treated from immune seurat objects
cd4_seurat<- subset(cd4_seurat, Treatment=="PD1")
cd8_seurat<- subset(cd8_seurat, Treatment=="PD1")
myeloid_seurat<- subset(myeloid_seurat, Treatment=="PD1")
```


## Merge seurat objects together
```{r}
## Add common column for: 
## 1) high level cell type (tcell, myeloid, tumor)
cd4_seurat$highLevel_annot<- "Tcell"
cd8_seurat$highLevel_annot<- "Tcell"
myeloid_seurat$highLevel_annot<- "Myeloid"
tumor_seurat$highLevel_annot<- "Tumor"

## 2) detailed - (CD4/CD8)
## For myeloid: may replace this with microglia/macrophage/monocyte distinction
cd4_seurat$detailed_annotForCellChat<- "CD4"
cd8_seurat$detailed_annotForCellChat<- "CD8"
myeloid_seurat$detailed_annotForCellChat<- "Myeloid"
tumor_seurat$detailed_annotForCellChat<- "Tumor"

## 3) program
cd4_seurat$immunePrograms_tumorCellTypes<- paste0("CD4_", cd4_seurat$Program_projected)
cd8_seurat$immunePrograms_tumorCellTypes<- paste0("CD8_", cd8_seurat$Program_projected)
myeloid_seurat$immunePrograms_tumorCellTypes<- myeloid_seurat$Program_projected
tumor_seurat$immunePrograms_tumorCellTypes<- tumor_seurat$Program_projected_manualOPC

## Add on any other colums needed before merging
tumor_seurat$sampleid<- tumor_seurat$sample

## Check that scaling is the same for all
head(colSums(myeloid_seurat@assays$RNA@counts))/head(colSums(cd4_seurat@assays$RNA@counts))
head(colSums(myeloid_seurat@assays$RNA@counts))/head(colSums(tumor_seurat@assays$RNA@counts))

## Match metadata to clean up merged seurat
all_colnames<-c(colnames(myeloid_seurat@meta.data),colnames(cd4_seurat@meta.data), colnames(cd8_seurat@meta.data), colnames(tumor_seurat@meta.data))
commonColumns<- table(all_colnames); commonColumns<- names(commonColumns)[commonColumns==4]

tumor_seurat@meta.data<- tumor_seurat@meta.data[,commonColumns]
myeloid_seurat@meta.data<- myeloid_seurat@meta.data[,commonColumns]
cd4_seurat@meta.data<- cd4_seurat@meta.data[,commonColumns]
cd8_seurat@meta.data<- cd8_seurat@meta.data[,commonColumns]

## Merge together
merged_seurat<- merge(tumor_seurat, c(myeloid_seurat, cd4_seurat, cd8_seurat))

## Run full seurat processing- not needed for cellChat, but nice for a high-level look
merged_seurat_meta<- merged_seurat@meta.data
merged_seurat<- RunFullSeurat(merged_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=merged_seurat_meta$sampleid)
merged_seurat<- AddMetaData(merged_seurat, merged_seurat_meta[,!colnames(merged_seurat_meta) %in% colnames(merged_seurat@meta.data)])

saveRDS(merged_seurat, file=paste0(analysis_dir, "seurat_tumor.tcell.myeloid.Rds"))
```


## Split into 1 seurat object per sample
```{r}
all_seurat<- readRDS(paste0(analysis_dir, "seurat_tumor.tcell.myeloid.Rds"))

## Number of cells in each cell type per sample, save
## This includes detailed annot (CD4, CD8, etc) as well as Tcell/myeloid programs
meta<- all_seurat@meta.data

  ## detailed annot
  nCells_bySample<- as.data.frame(table(meta$detailed_annotForCellChat, meta$sample))
  colnames(nCells_bySample)<-c("CellType", "Sample", "NCells")
  
  ## By program
  nProgramCells_bySample<- as.data.frame(table(meta$immunePrograms_tumorCellTypes, meta$sample))
  colnames(nProgramCells_bySample)<-c("CellType", "Sample", "NCells")

nCells_bySample<- rbind(nCells_bySample, nProgramCells_bySample[!nProgramCells_bySample$CellType %in% nCells_bySample$CellType,])  
saveRDS(nCells_bySample, file=paste0(analysis_dir, "nCellTypes_bySample.Rds"))

## Cycle through each sample, split seurat object
all_seurat_list<- lapply(unique(meta$sample), function(x){
  seurat_tmp<- subset(all_seurat, sample==x)
  print(paste0(x, " has ", ncol(seurat_tmp)," cells"))
  return(seurat_tmp)
})
names(all_seurat_list)<- unique(meta$sample)

saveRDS(all_seurat_list, file=paste0(analysis_dir, "seurat_list_bySample.Rds"))
```


## Write out comparison file- sender/receiver text file
file_with_comparisons should be a text file with 3 columns that follow the structure: comparison_name sender receiver
Can run all simultaneously, but takes too long to run- over 3 days
Split into multiple comparison files, just so they can run simultaneously

## By program
```{r}
all_seurat<- readRDS(paste0(analysis_dir, "seurat_tumor.tcell.myeloid.Rds"))
nCells_bySample<- readRDS(paste0(analysis_dir, "nCellTypes_bySample.Rds"))


## Get list of all cell types by high level cell type
cd8_types<- unique(all_seurat$immunePrograms_tumorCellTypes[all_seurat$detailed_annotForCellChat=="CD8"])
cd4_types<- unique(all_seurat$immunePrograms_tumorCellTypes[all_seurat$detailed_annotForCellChat=="CD4"])
myeloid_types<- unique(all_seurat$immunePrograms_tumorCellTypes[all_seurat$detailed_annotForCellChat=="Myeloid"])
tumor_types<- unique(all_seurat$immunePrograms_tumorCellTypes[all_seurat$detailed_annotForCellChat=="Tumor"])

## Get list of cell types with at least 11 cells- required for cell chat input
enough_cells<- nCells_bySample[nCells_bySample$NCells>10 & nCells_bySample$CellType %in% all_seurat$immunePrograms_tumorCellTypes,]  
cd8_types<- cd8_types[cd8_types %in% enough_cells$CellType]
cd4_types<- cd4_types[cd4_types %in% enough_cells$CellType]
myeloid_types<- myeloid_types[myeloid_types %in% enough_cells$CellType]
tumor_types<- tumor_types[tumor_types %in% enough_cells$CellType]

## Between high level cell type lists, get all pairwise combinations
cd4_myeloid<- expand.grid(Sender=as.list(cd4_types), Receiver=as.list(myeloid_types))
cd8_myeloid<- expand.grid(Sender=as.list(cd8_types), Receiver=as.list(myeloid_types))
cd4_cd8<- expand.grid(Sender=as.list(cd4_types), Receiver=as.list(cd8_types))
tumor_cd4<- expand.grid(Sender=as.list(tumor_types), Receiver=as.list(cd4_types))
tumor_cd8<- expand.grid(Sender=as.list(tumor_types), Receiver=as.list(cd8_types))
tumor_myeloid<- expand.grid(Sender=as.list(tumor_types), Receiver=as.list(myeloid_types))


## All together:
all_comparisons<- do.call("rbind", list(cd4_myeloid, cd8_myeloid, cd4_cd8, tumor_cd4, tumor_cd8, tumor_myeloid))
all_comparisons<- as.data.frame(apply(all_comparisons, 2, function(x){as.character(x)}))

  ## Add column with analysis name
  all_comparisons$name<- paste0(all_comparisons$Sender, "_to_", all_comparisons$Receiver)
  all_comparisons<- all_comparisons[,c("name", "Sender", "Receiver")]


write.table(all_comparisons, file=paste0(analysis_dir, "all_comparisons.txt"), col.names = FALSE, quote = FALSE, row.names = FALSE)


## Separated to speed up run time (these files used)
## not used currently
comparison_list<- list(tcell_myeloid=tcell_myeloid, myeloid_tcell=myeloid_tcell, 
                       tumor_clonal=tumor_clonal, tumor_nonclonal=tumor_nonclonal, tumor_myeloid=tumor_myeloid)
comparison_list<- lapply(comparison_list, function(x){
  x$name<- paste0(x$Sender, "_to_", x$Receiver)
  x<- x[,c("name", "Sender", "Receiver")]
  x$Sender<- as.character(x$Sender);x$Receiver<- as.character(x$Receiver)
  return(x)
})
for(i in names(comparison_list)){
  tmp<- comparison_list[[i]]
  write.table(tmp, file=paste0(analysis_dir, "/byCellType/comparisons_", i, ".txt"), col.names = FALSE, quote = FALSE, row.names = FALSE)
}

```


## By program: tumor cell types, myeloid programs, cd4 programs, cd8 programs
not used currently
```{r}
all_seurat<- readRDS(paste0(analysis_dir, "seurat_tumor.tcell.myeloid.Rds"))

## Add on CD4/CD8 distinction
all_seurat$detailed_annot<- ifelse(grepl("CD4", all_seurat$immunePrograms_tumorCellTypes), "CD4",
                                   ifelse(grepl("CD8", all_seurat$immunePrograms_tumorCellTypes), "CD8",
                                          ifelse(grepl("Unclear", all_seurat$immunePrograms_tumorCellTypes), "UnclearTcell",
                                                 all_seurat$highLevel_annot)))

## Remove "unclear" cells, inhibitory CD8
all_seurat<- subset(all_seurat, detailed_annot != "UnclearTcell")
all_seurat<- subset(all_seurat, immunePrograms_tumorCellTypes != "CD8_Inhibitory")

## Remove any programs/cell types without at least 10 cells in 1 sample
nCells<- as.data.frame.matrix(table(all_seurat$immunePrograms_tumorCellTypes, all_seurat$sample))
nSamples_pf<- apply(nCells, 1, function(x){sum(x>10)})
cellTypes_keep<- names(nSamples_pf)[nSamples_pf!=0]
all_seurat<- subset(all_seurat, immunePrograms_tumorCellTypes %in% cellTypes_keep)

## Get list of all programs by high level cell type
cd4_types<- unique(all_seurat$immunePrograms_tumorCellTypes[grepl("CD4", all_seurat$immunePrograms_tumorCellTypes)])
cd8_types<- unique(all_seurat$immunePrograms_tumorCellTypes[grepl("CD8", all_seurat$immunePrograms_tumorCellTypes)])
myeloid_types<- unique(all_seurat$immunePrograms_tumorCellTypes[all_seurat$detailed_annot=="Myeloid"])
tumor_types<- unique(all_seurat$immunePrograms_tumorCellTypes[all_seurat$highLevel_annot=="Tumor"])


## Between high level cell type lists, get all pairwise combinations
cd4_myeloid<- expand.grid(Sender=as.list(cd4_types), Receiver=as.list(myeloid_types))
cd8_myeloid<- expand.grid(Sender=as.list(cd8_types), Receiver=as.list(myeloid_types))
tumor_cd4<- expand.grid(Sender=as.list(tumor_types), Receiver=as.list(cd4_types))
tumor_cd8<- expand.grid(Sender=as.list(tumor_types), Receiver=as.list(cd8_types))
tumor_myeloid<- expand.grid(Sender=as.list(tumor_types), Receiver=as.list(myeloid_types))
all_list<- list(cd4_myeloid=cd4_myeloid, cd8_myeloid=cd8_myeloid, tumor_cd4=tumor_cd4, tumor_cd8=tumor_cd8, tumor_myeloid=tumor_myeloid)

## For each cell type list, cycle through each comparison. Remove any without at least 1 sample with 10 cells in each
nCells_allTypes<- as.data.frame.matrix(table(all_seurat$immunePrograms_tumorCellTypes, all_seurat$sample))
all_list_filtered<- lapply(all_list, function(x){
  ## Cycle through each pair in the comparisons table- check whether at least 1 sample "passes" (has 10 cells in each)
  pair_pf<- unlist(lapply(1:nrow(x), function(pair){
    tmp<- nCells_allTypes[unlist(x[pair,]),]
    pf<- sum(apply(tmp, 2, function(y){sum(y>10)==2})) ## number of samples with > 10 cells in BOTH cell types in pair
    if(pf!=0){return("pass")}else{return("fail")}
  }))
  
  ## Only return comparisons that "pass"
  return(x[pair_pf=="pass",])
})

## Add column with analysis name to each comparison table, then write out
all_list_filtered<- lapply(names(all_list_filtered), function(x){
  all_comparisons<- all_list_filtered[[x]]
  
  ## Convert columns to character- converted to list during pairwise creation
  all_comparisons$Sender<- as.character(all_comparisons$Sender)
  all_comparisons$Receiver<- as.character(all_comparisons$Receiver)
  
  ## Add on "comparison name"
  all_comparisons$name<- paste0(all_comparisons$Sender, "_to_", all_comparisons$Receiver)
  all_comparisons<- all_comparisons[,c("name", "Sender", "Receiver")]
  
  ## Write out to file for input into cellchat
  write.table(all_comparisons, file=paste0(analysis_dir, "byProgram/comparisons_byPrograms_",x, ".txt"), col.names = FALSE, quote = FALSE, row.names = FALSE)
})



```

```{r}
## How many cells are actually input into cellchat for each interaction? 
## To simplify analysis, remove an interactions that do not have at least 1 sample with >10 cells in both interactor
## These interactions will not be run in cell chat anyways
## Cycle through each sample, determine whether it has enough cells to run each interaction
cell_counts<- lapply(all_seurat_list, function(x){
  nCells<- as.data.frame(table(x$detailed_annotForCellChat))
  comp_counts<- lapply(comparisons$V1, function(comp){
    tmp<- comparisons[comparisons$V1==comp,]
    PF<- sum(nCells[nCells$Var1==tmp$V2, "Freq"]>10 & nCells[nCells$Var1==tmp$V3, "Freq"]>10)==1
    if(PF){
      total_counts<- nCells[nCells$Var1==tmp$V2, "Freq" ] + nCells[nCells$Var1==tmp$V3, "Freq" ] 
    }else{total_counts<-0}
    return(total_counts)
  })
  names(comp_counts)<- comparisons$V1
  comp_counts<- lapply(comp_counts, function(x){ifelse(length(x)==0, 0, x)})
  comp_counts<- as.data.frame(unlist(comp_counts))
  colnames(comp_counts)<-unique(x$sample)
  return(comp_counts)
})
cell_counts<- do.call("cbind", cell_counts)

## Number of samples that would be run for each interaction
nSamples_interaction<- apply(cell_counts,1, function(x){sum(x>0)})

## Interactions with 0 samples that would not be run- remove here for simplicity
interactions_remove<- names(nSamples_interaction)[nSamples_interaction==0]
comparisons<- comparisons[!comparisons$V1 %in% interactions_remove,]
write.table(all_comparisons, file=paste0(analysis_dir, "all_comparisons.txt"), col.names = FALSE, quote = FALSE, row.names = FALSE)
```

## For testing purposes: subsample from program to determine what effect cell count has on results
Specifically: IFN_TAM shows high LR --> CD4, but if you divide by total cell count, this effect disappears
Is the normalization removing technical bias or hiding real biological difference?
To test: run cellchat on IFN_TAM + CD4 programs, subsampling # of IFN cells
If no big difference in # LRs across subsampling: no need to normalize (i.e., was hiding biological difference)
If clearly decreasing # LRs: cell count affects # LR detected, need to normalize (i.e., was removing technical bias)
```{r}
## Read in seurat object
all_seurat<- readRDS(paste0(analysis_dir, "seurat_tumor.tcell.myeloid.Rds"))

## Subset to IFN_TAM + CD4 programs
cd4_programs<- unique(all_seurat$immunePrograms_tumorCellTypes[grepl("CD4", all_seurat$immunePrograms_tumorCellTypes)])
seurat_use<- subset(all_seurat, immunePrograms_tumorCellTypes %in% c(cd4_programs, "IFN_TAM"))

## 
```

