---
title: "R Notebook"
output: html_document
---

## Preprocess matrices to seurat object, then do broad annotation (myeloid, tcell) + detailed (cd4, cd8) and save these seurat objects

```{r}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/PD1_Treat/tenX/")
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"


preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
library(Matrix)

analysis_dir<- "analysis/preprocessing/"
figure_dir<- "figures/qc/"
data_dir<- "data/"

if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}

marker_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Marker_genes/")
hkgenes <- read.table(paste0(marker_dir, "tirosh_house_keeping.txt"), skip = 2)
hkgenes <- as.vector(hkgenes$V1)
```

## Read in matrixes- 1 per sample
```{r}
matrix_dirs<- list.files(data_dir)


## Remove TCR/VDJ directories- analyzed separately
matrix_dirs<- matrix_dirs[!grepl("TCR", matrix_dirs)]
matrix_dirs<- matrix_dirs[!grepl("VDJ", matrix_dirs)]

all_matrix<- lapply(matrix_dirs, function(x){
  matrix_dir<- paste0(data_dir, x)
  print(matrix_dir)
  barcode.path <- paste0(matrix_dir, "/filtered_feature_bc_matrix/barcodes.tsv.gz")
  features.path <- paste0(matrix_dir, "/filtered_feature_bc_matrix/features.tsv.gz")
  matrix.path <- paste0(matrix_dir, "/filtered_feature_bc_matrix/matrix.mtx.gz")
  mat <- readMM(file = matrix.path)
  feature.names = read.delim(features.path,
                             header = FALSE,
                             stringsAsFactors = FALSE)
  barcode.names = read.delim(barcode.path,
                             header = FALSE,
                             stringsAsFactors = FALSE)
  colnames(mat) = barcode.names$V1
  rownames(mat) = feature.names$V2
  return(mat)
})
names(all_matrix)<- matrix_dirs

```

## Make simple metadata for each matrix- cells + sample
```{r}
all_meta<- lapply(names(all_matrix), function(x){
  matrix<- all_matrix[[x]]
  meta<- data.frame(cell=colnames(matrix), sample=rep(x, ncol(matrix)))
  return(meta)
})
```

## Save cm and meta
```{r}
meta<- do.call("rbind", all_meta)
cm<- all_matrix$GEX

saveRDS(cm, file=paste0(analysis_dir, "raw_cm.Rds"))
saveRDS(meta, file=paste0(analysis_dir, "raw_meta.Rds"))
```

## Create seurat object, do initial QC
```{r}
cm<- readRDS(paste0(analysis_dir, "raw_cm.Rds"))
meta<- readRDS(paste0(analysis_dir, "raw_meta.Rds"))

seurat_raw<- RunFullSeurat(cm=cm, RunHarmony = FALSE, samples=meta$sample)


## Identify # of HK genes
hkgenes.notfound.indices <- which(!hkgenes %in% rownames(seurat_raw))  
if (length(hkgenes.notfound.indices)) {
  message = paste0(paste0(hkgenes[hkgenes.notfound.indices], collapse = " "), " housekeeping genes were not found in sample ", i)
  print(message)
}

# Get number of expressed housekeeping genes.
hkgenes.found <- hkgenes[-hkgenes.notfound.indices]  # remove hkgenes that were not found
seurat_raw[["nHK_RNA"]] <- Matrix::colSums(seurat_raw[['RNA']]@data[hkgenes.found, ] > 0)

# Get number of expressed housekeeping genes.
hkgenes.found <- hkgenes[-hkgenes.notfound.indices]  # remove hkgenes that were not found
seurat_raw[["nHK_RNA"]] <- Matrix::colSums(seurat_raw[['RNA']]@data[hkgenes.found, ] > 0)

## Get percentage of mito genes
seurat_raw[["percent_mito"]] <- PercentageFeatureSet(seurat_raw, pattern = "^MT-")
  


## QC plots
p1<-DimPlot(seurat_raw, group.by="sample", label=TRUE)+NoLegend()
p2<- VlnPlot(seurat_raw, features = "nFeature_RNA", group.by="sample")+NoLegend()
p3<- VlnPlot(seurat_raw, features="percent_mito", group.by="sample")+NoLegend()
p4<- VlnPlot(seurat_raw, features="nHK_RNA", group.by="sample")+NoLegend()


p1+p2+p3+p4
ggsave(paste0(figure_dir, "nFeature.Count_BySample.png"), width=8, height=10)
saveRDS(seurat_raw, file=paste0(analysis_dir, "raw_seurat.Rds"))
```

## Plot histogram of nFeature/nHK/nMitochondrial to determine good cutoff
```{r}
df<- seurat_raw@meta.data
p1<- ggplot(df, aes(x=nFeature_RNA, fill=sample))+geom_histogram()+ggtitle("nFeatures")
p2<- ggplot(df, aes(x=nHK_RNA, fill=sample))+ geom_histogram()+ggtitle("n HK genes")+theme(legend.position = "none")
p3<- ggplot(df, aes(x=nCount_RNA, fill=sample))+ geom_histogram() + ggtitle("nCounts")
p4<- ggplot(df, aes(x=percent_mito, fill=sample))+ geom_histogram() + ggtitle("percent mito")+theme(legend.position = "none")
p1+p2+p3+p4
ggsave(paste0(figure_dir, "Histogram_Features.Counts.HK.Mito_noTCR.png"), width=10, height=8)
```




## Filter cells by nFeature + nHK
```{r}
seurat_raw<- readRDS(paste0(analysis_dir, "raw_seurat.Rds"))

## Filter 
seurat_obj_filt <- subset(seurat_raw, subset = nFeature_RNA > 500  & nHK_RNA > 20)
seurat_raw$PF<- ifelse(colnames(seurat_raw) %in% colnames(seurat_obj_filt), "Pass", "Fail")

df<- seurat_raw@meta.data
df<- as.data.frame(table(df$sample, df$PF))
colnames(df)<- c("Sample", "PF", "nCells")
ggplot(df, aes(x=Sample, y=nCells, fill=PF))+
  geom_bar(stat="identity", position="stack")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("nFeature_RNA > 500\nnHK_RNA > 20")+
  xlab("")
ggsave(paste0(figure_dir, "PF_BySample_nFeature.nHK.png"), width=3, height=3)

## recluster without low quality cells
seurat_obj_filt_meta<- seurat_obj_filt@meta.data
seurat_obj_filt<- RunFullSeurat(seurat_obj_filt@assays$RNA@counts, RunHarmony = FALSE, samples=seurat_obj_filt_meta$sample)

saveRDS(seurat_obj_filt, file=paste0(analysis_dir, "seurat_obj.Rds"))

```

## PD1 has a few AC, OC- remove prior to projection
```{r}
seurat_obj<- readRDS(paste0(analysis_dir, "seurat_obj.Rds"))


## Reduce resolution to identify markers for major clusters
seurat_obj<-FindClusters(seurat_obj, resolution = 0.15)
all_markers<- FindAllMarkers(seurat_obj, min.pct = 0.4, logfc.threshold = 0.5)
all_markers_filtered<- all_markers[all_markers$p_val_adj<0.05,]
top_markers<- all_markers_filtered %>% group_by(cluster) %>% top_n(n=10, wt=avg_log2FC) %>% as.data.frame()

## Assign cluster 6 to AC, cluster r to Unknown (not T cells or)
seurat_obj$AC<- ifelse(seurat_obj$seurat_clusters==6, "AC", ifelse(seurat_obj$seurat_clusters==5, "Unknown", "Other"))
FeaturePlot(seurat_obj, features=c("CD3G", "CSF1R", "GFAP")) + DimPlot(seurat_obj, group.by = "AC")
ggsave(paste0(figure_dir, "UMAP_removeAC.Unknown.png"), width=7, height=6)


## Remove AC/Unknown
seurat_obj<- subset(seurat_obj, AC =="Other")
pd1_meta<- seurat_obj@meta.data
seurat_obj<- RunFullSeurat(seurat_obj@assays$RNA@counts, RunHarmony = FALSE, samples=pd1_meta$sample)
seurat_obj<- AddMetaData(seurat_obj, pd1_meta[,!colnames(pd1_meta) %in% colnames(seurat_obj@meta.data)])

saveRDS(seurat_obj, file="analysis/preprocessing/seurat_obj_noAC.Unknown.Rds")
```
