---
title: ""
author: "Jenna LaBelle"
date: "10/07/22"
output: html_document
---

## Use programs from full cohort to score adult pd1 10x

```{r}
library(Seurat) 

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/PD1_Treat/03_comparisonToAdult/")

analysis_dir<- paste0(working_dir, "/analysis/score_programs/")
figure_dir<- paste0(working_dir, "/figures/score_programs/")
data_dir<- paste0(working_dir, "analysis/preprocess_adult/")
full_cohort_dir<- "../../ImmuneCellAnalysis/"

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)
source(paste0(base_dir, script_dir, "seurat_transferAnchors_helper.R"))

```

```{r}
## Helper function to score a seurat object for a set of markers
my_scoreGeneset<- function(seurat, markers){
  ## Get top markers to score- same # for all programs
  min_n_markers<- markers %>% group_by(cluster) %>% summarise(nMarkers=n()) %>% filter(nMarkers==min(nMarkers)) %>% as.data.frame()
  markers<- markers %>% group_by(cluster)  %>% top_n(n=min_n_markers$nMarkers, wt=avg_log2FC) %>% as.data.frame()
  markers<- split(markers, f=markers$cluster)
  markers<- lapply(markers, function(x){x$gene})
  
  ## Norm/center
  cm_list<- NormCenter(seurat@assays$RNA@counts)
  cm_mean<- rowMeans(log2(cm_list$raw_data + 1))
  
  ## score
  scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, markers)))
  
  ## Get max scoring program
  scores$MaxScoringProgram<- apply(scores, 1, function(x){names(x)[which.max(x)]})
  scores$MaxScoreProgram<- apply(scores[,colnames(scores) != "MaxScoringProgram"], 1, function(x){x[which.max(x)]})
  
  ## Add to seurat
  seurat<- AddMetaData(seurat, scores[,c("MaxScoringProgram", "MaxScoreProgram")])
  return(seurat)
  }
```

## Load seurat objects: published adult PD1 and full immune cohort (both myeloid and t cells)

```{r}
## Load adult pd1 treated
pd1_cd4<- readRDS(paste0(data_dir, "cd4_seurat.Rds"))
pd1_cd8<- readRDS(paste0(data_dir, "cd8_seurat.Rds"))
pd1_myeloid<- readRDS(paste0(data_dir, "myeloid_seurat.Rds"))

## rename treatment- untreated/treated
pd1_cd4$Treatment<- factor(ifelse(pd1_cd4$condition=="GBM.PD1", "Treated", "Untreated"), levels=c("Untreated", "Treated"))
pd1_cd8$Treatment<- factor(ifelse(pd1_cd8$condition=="GBM.PD1", "Treated", "Untreated"), levels=c("Untreated", "Treated"))
pd1_myeloid$Treatment<- factor(ifelse(pd1_myeloid$condition=="GBM.PD1", "Treated", "Untreated"), levels=c("Untreated", "Treated"))


## Colors 
tcell_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Tcell.programs_rank8sharedByunion_reAnnot.Rds"))
myeloid_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Myeloid_program_colors.Rds"))
myeloid_colors<- myeloid_colors$rank6
immune_colors<- c(tcell_colors, myeloid_colors)
colors_treatment<- c(PD1="orange", Untreated="purple")

## Full cohort marker genes
tcell_markers<- readRDS(paste0(full_cohort_dir, 
                             "02a_Tcells/02_Ped.Adult/01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/", 
                             "new_markers_NMFrank8_bySeurat_sharedByunion_reAnnot.Rds"))
myeloid_markers<- read.csv(paste0(full_cohort_dir,
                               "02b_Myeloid/01a_NMF/analysis/de.novoNMF/pedOnly_nomuv63/", 
                               "top_DE_genes_10Xannotated_NMF_rank6.csv"))

  ## Top markers per program
  top_tcell_markers<- tcell_markers %>% group_by(cluster) %>% top_n(n=5, wt=avg_log2FC) %>% as.data.frame()
  top_myeloid_markers<- myeloid_markers %>% group_by(cluster) %>% top_n(n=5, wt=avg_log2FC) %>% as.data.frame()

```




## score, assign max
Note: too computationally intensive to do for myeloid. Scored on O2, read scores in here, add max scoring program/score
```{r}
pd1_cd4<- my_scoreGeneset(pd1_cd4,  tcell_markers[!tcell_markers$cluster %in% c("Cytotoxic.NKreceptor", "Predysfunctional"),])
pd1_cd8<- my_scoreGeneset(pd1_cd8,  tcell_markers[!tcell_markers$cluster %in% c("Treg", "Inhibitory"),])
#pd1_myeloid<- my_scoreGeneset(pd1_myeloid,  myeloid_markers)

saveRDS(pd1_cd4, file=paste0(data_dir, "cd4_seurat.Rds"))
saveRDS(pd1_cd8, file=paste0(data_dir, "cd8_seurat.Rds"))
#saveRDS(pd1_myeloid, file=paste0(data_dir, "myeloid_seurat.Rds"))
```

## Plot UMAP of programs

```{r}
tenx_seurat<- list(CD4=pd1_cd4, CD8=pd1_cd8, Myeloid=pd1_myeloid)

## Plot UMAP
all_plots<- lapply(names(tenx_seurat), function(x){
  tmp<- tenx_seurat[[x]]
  DimPlot(tmp, group.by="MaxScoringProgram", pt.size = .5, label=FALSE, label.size = 6)+ 
    scale_color_manual(values=immune_colors[names(immune_colors) %in% tmp$MaxScoringProgram])+
    ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "UMAP_MaxScoringProgram.png"), width=18, height=5)
```

## Plot proportion of programs
```{r}
tenx_seurat<- list(CD4=pd1_cd4, CD8=pd1_cd8, Myeloid=pd1_myeloid)

## By treatment
all_plots<- lapply(names(tenx_seurat), function(x){
  tmp<- tenx_seurat[[x]]
  my_barchart_programProp(tmp,grouping_variable = "Treatment", coloring_variable = "MaxScoringProgram", colors=immune_colors)+ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "ProgramProportion_All_byTreatment.png"), width=12, height=4)


## By primary/recurrent/treated
all_plots<- lapply(names(tenx_seurat), function(x){
  tmp<- tenx_seurat[[x]]
  my_barchart_programProp(tmp,grouping_variable = "condition", coloring_variable = "MaxScoringProgram", colors=immune_colors)+ggtitle(x)+
    theme(axis.text.x = element_text(angle=45, hjust=1))
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "ProgramProportion_All_byTreatment.withCondition.png"), width=12, height=4)


## By treatment + sample
all_plots<- lapply(names(tenx_seurat), function(x){
  tmp<- tenx_seurat[[x]]
  my_barchart_programProp(tmp,grouping_variable="sample", wrapping_variable = "Treatment", coloring_variable = "MaxScoringProgram", colors=immune_colors)+ggtitle(x)+
      theme(axis.text.x = element_blank())
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "ProgramProportion_All_bySample.Treatment.png"), width=24, height=4)
```





## Pseudobulked heatmap of expression of marker genes
```{r}
marker_list<- list(CD4=top_tcell_markers[!top_tcell_markers$cluster %in% c("Predysfunctional", "Cytotoxic.NKreceptor"),], 
                   CD8=top_tcell_markers[!top_tcell_markers$cluster %in% c("Treg", "Inhibitory"),],
                   Myeloid=top_myeloid_markers)

sample_center_method<- "bySample" ## approach for centering- allSamples or bySample

all_plots<- lapply(names(tenx_seurat), function(x){
  tmp<- tenx_seurat[[x]]
  markers_use<-marker_list[[x]]
  markers_use<- markers_use[markers_use$cluster %in% unique(tmp$MaxScoringProgram),]
  markers_use<- markers_use[markers_use$gene %in% rownames(tmp),]
  
  ## Create pseudobulking variable
  tmp$pb<- paste0(tmp$Treatment, "_", gsub("\\.", "\n", gsub("_", "\n", tmp$MaxScoringProgram)))
  pb<- pseudobulk_byVariable(tmp, tmp@meta.data, "pb")
  
  
  ## center for each sample separately
  if(sample_center_method=="bySample"){
      center_bySample<- lapply(unique(tmp$Treatment), function(sample){
    pb_tmp<- pb[,grepl(sample, colnames(pb))]
    if(class(pb_tmp)=="numeric"){
      pb_tmp<- as.data.frame(pb_tmp)
      colnames(pb_tmp)<- colnames(pb)[grepl(sample, colnames(pb))]
    }
    pb_list_tmp<- NormCenter(pb_tmp)
    pb_center_tmp<- pb_list_tmp$center_data
    return(pb_center_tmp)
  })
  center_bySample<- do.call("cbind", center_bySample)
  }
  
  ## or- center across all samples
  if(sample_center_method=="allSamples"){
    center_bySample<- NormCenter(pb)
    center_bySample<- center_bySample$center_data
  }

  ## Plot
  myHeatmap(center_bySample, GOI=markers_use$gene, min.value = -3, max.value = 3, 
                orderFactors =  gsub("\\.", "\n",gsub("_", "\n", unique(markers_use$cluster))),
            orderSubtypes =c("Untreated", "Treated") ) +
    scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")+
    ggtitle(x)
})
cowplot::plot_grid(plotlist=all_plots, ncol=3)
ggsave(paste0(figure_dir, "Heatmap_ExpectedMarkerGenes_All_centerBySample.png"), width=30, height=9)
```



