---
title: "NMF_merge_template"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}
library(NMF)

base_dir = "C:/Users/jenna/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/Immune/Midline_HistoneMutant/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
plot_helper = paste0(base_dir, script_dir, "Plotting_helper_functions.R")
stat_helper = paste0(base_dir, script_dir, "common_stat_test.R")
source(preprocessing)
source(nmf_helper)
source(plot_helper)
source(stat_helper)

nmf_folder = "analysis/nmf/"

nmf_analysis_folder = "analysis/nmf/Merge/"
if (!dir.exists(nmf_analysis_folder)){
  dir.create(nmf_analysis_folder)
}

nmf_fig_folder = "figures/nmf/Merge/"
if (!dir.exists(nmf_fig_folder)){
  dir.create(nmf_fig_folder)
}
```

# Load and preprocess raw data

## Load cm, tumor_id, samples and genes_to_keep
```{r load-precomputed-data}
load("../analysis/midlineHistoneMut_cm_list.Robj")
tumor_stat = readRDS("../analysis/inferCNV/seurat/midlineHistoneMut_call_tumor.rds")
samples = readRDS("../analysis/midlineHistoneMut_samples.rds")
genes_to_keep = readRDS(paste0(nmf_folder, "top_10000_ode.rds"))

cm_list<-midlineHistoneMut_cm_list

sum(colnames(cm_list$raw_data)==names(samples)); length(samples)
sum(names(samples)==names(tumor_stat))
```

## Load precomputed cm_mean, cm_center, basis and samples in the list 
## Only do this after you have ran through this code before
## When you want to re-compute metaprograms 
```{r}
fname = paste0(nmf_folder, "nmf_raw_data.Robj")
if(file.exists(fname)){
  load(fname)
  
  cm_mean = nmf_raw_data$cm_mean
  cm_center = nmf_raw_data$cm_center
  nmf_basis_concat = nmf_raw_data$basis
  samples_tumor = nmf_raw_data$samples
  
  nmf_score = nmf_raw_data[["nmf_score"]] = 
  nmf_factor_cor = nmf_raw_data[["nmf_factor_cor"]]
  nmf_basis_concat = nmf_raw_data[["basis"]]

  rm(nmf_raw_data);gc()
}
```

## Normalize and center cm 
## Still only keep malignant cells and all genes that pass the filter 
```{r subset tumor cells}
## Subset raw and log transformed data
cm = cm_list$raw_data[, tumor_stat=="Malignant"]
cm_norm = cm_list$norm_data[, tumor_stat=="Malignant"]
cm_mean = log2(rowMeans(cm)+1)
##cm_mean = rowMeans(cm_norm)
## Remake centered values 
cm_center = cm_norm - rowMeans(cm_norm)
samples_tumor = samples[tumor_stat=="Malignant"]
rm(cm_list); gc()
```

## Save computed cm, normalized/centered cm and samples in a list 
## For computing metaprograms again later if needed
## Saving these intermediate steps occupies quite some space 
```{r}
nmf_raw_data = list()
nmf_raw_data[["cm"]] = cm
nmf_raw_data[["cm_norm"]] = cm_norm
nmf_raw_data[["cm_center"]] = cm_center 
nmf_raw_data[["cm_mean"]] = cm_mean
nmf_raw_data[["samples"]] = samples_tumor 
```

## Load computed nmf_obj_list and nmf_gene_list
```{r load-computed-NMF-factors}
load(paste0(nmf_folder, "nmf_obj_list.Robj"))
load(paste0(nmf_folder, "nmf_gene_list.Robj"))
```

# Cluster individual NMF factors and merge them into metaprograms

## Compute nmf scores for each nmf factor in the nmf_gene_list
```{r nmf_score}
## Compute signature score for each NMF factor (30 representative genes) in all cells 
nmf_score = scoreNmfGenes(cm_center, cm_mean, nmf_gene_list)
nmf_score = t(nmf_score)
```

## Concatenate basis matrix of each sample together 
```{r}
nmf_basis_concat = concatNmfFactor(nmf_obj_list, genes_to_keep)
nmf_raw_data[["basis"]] = nmf_basis_concat
```

## Hierarchical clustering of nmf factors with correlation among nmf scores 
## The resulting heatmap and dendrogram will help you aggregate individual NMF factors into metaprograms
```{r NMF_factor_merge_2}
## NMF factor correlation based on cell score 
nmf_factor_hc = clusterNmfFactors(nmf_score)
nmf_factor_cor = nmf_factor_hc$cor_coef[nmf_factor_hc$hc_obj$order, nmf_factor_hc$hc_obj$order]

## Heatmap of correlations
hm_colors = rev((brewer.pal(n=9, name="RdBu")))
hm_colors = colorRampPalette(colors = hm_colors)
pheatmap(nmf_factor_cor, color = hm_colors(100), 
         cluster_rows = F, cluster_cols = F, 
         annotation_names_row = F, annotation_names_col =T,
         show_rownames = F, show_colnames = T,
         filename = paste0(nmf_fig_folder, "nmf_factors_pairwise_cor.png"),
         width = 12, height = 8)

## Dendrogram of hierarchical clustering 
jpeg(filename=paste0(nmf_fig_folder, "nmf_factors_hc.png"), width = 1200, height = 800)
plot(nmf_factor_hc$hc_obj)
dev.off()
```



## Merge nmf factors into metagene programs with correlation among gene weights or nmf scores 
## We can go over your heatmap and dendrograms to see how to aggregate them
```{r nmf_factor_merge}
## Concatenate all basis vectors with gene weights
nmf_raw_data[["nmf_score"]] = nmf_score
nmf_raw_data[["nmf_factor_cor"]] = nmf_factor_cor 
save(nmf_raw_data, file=paste0("analysis/nmf/", "nmf_raw_data.Robj"))
```





```{r}
NameOfMetaprogramList<-"Merge2"
## NMF metaprograms
## Adapt this to your dataset
nmf_meta_programs = list()


nmf_meta_programs[["NMF1"]]<- colnames(nmf_factor_cor)[1:6]
nmf_meta_programs[["NMF2"]]<- colnames(nmf_factor_cor)[7:9]
nmf_meta_programs[["NMF3"]]<- colnames(nmf_factor_cor)[10:15]
nmf_meta_programs[["NMF4"]]<- colnames(nmf_factor_cor)[16:19]
#nmf_meta_programs[["NMF5"]]<- colnames(nmf_factor_cor)[21:22]
nmf_meta_programs[["NMF5"]]<- colnames(nmf_factor_cor)[23:30]
nmf_meta_programs[["NMF6"]]<- colnames(nmf_factor_cor)[31:35]



save(nmf_meta_programs, file = paste0(nmf_analysis_folder, "nmf_meta_programs", NameOfMetaprogramList, ".Robj"))
```



## Compuate average cor
## OPTIONAL, may exclude cluster of NMF factors (metaprograms) with low average correlations
```{r}
tmp = sapply(nmf_meta_programs, computeAverageCor, nmf_factor_cor)
```

## Compute top30 marker genes and diagnostic statistics for merged metagene programs 
```{r metaprogram_score_stats}
program_names = names(nmf_meta_programs)
nmf_marker_genes_final = list()
nmf_programs_stats = list()

for (i in seq(length(program_names))){
  ## Metaprogram marker genes 
  nmf_marker_genes_final[[program_names[[i]]]] = mergeNmfGenes(nmf_basis_concat, 
                                                    nmf_meta_programs[[i]], n = 30)
  ## Metaprogram statistics
  nmf_programs_stats[[program_names[[i]]]] = calculateNmfProgramStats(nmf_basis_concat, 
                                                                      nmf_meta_programs[[i]])
}

save(nmf_marker_genes_final, file=paste0(nmf_analysis_folder, "nmf_marker_genes_final", NameOfMetaprogramList, ".Robj"))
save(nmf_programs_stats, file=paste0(nmf_analysis_folder, "nmf_program_stats", NameOfMetaprogramList, ".Robj"))
```

```{r}
## Check overlap of gene programs 
for (i in 1:(length(nmf_marker_genes_final)-1)){
  for (j in (i+1):length(nmf_marker_genes_final)){
    tmp = intersect(nmf_marker_genes_final[[i]], nmf_marker_genes_final[[j]])
    if (length(tmp) >= 10){
      print("========================")
      print(names(nmf_marker_genes_final)[i])
      print(names(nmf_marker_genes_final)[j])
      print(tmp)
      print(length(tmp))
    }
  }
}
```


# Assign metaprogram to each cell 

## Compute metagene scores for metagene programs 
```{r final_NMF_scores}
nmf_score_final = scoreNmfGenes(cm_center, cm_mean, nmf_marker_genes_final)
nmf_score_final = t(nmf_score_final) 

## Compute and save metagene scores and 1-3 highest metagene programs and scores 

num_meta_program = length(nmf_marker_genes_final)
nmf_score_final_t = data.frame(nmf_score_final)

for (i in 1:3){
  nmf_score_final_t = metagene_score_signature(nmf_score_final_t, num_meta_program, i)
}

saveRDS(nmf_score_final_t, file=paste0(nmf_analysis_folder, "nmf_score", NameOfMetaprogramList, ".rds"))
```

## Exclude cell cycle scores and recompute/save metagene scores and 1-3 highest metagene programs and scores 
## May need to adapt this if you only have a single cell cycle program (not splited into S and G2M phase)
```{r}
nmf_score_final_tmp = data.frame(nmf_score_final)
## s_score = nmf_score_final_tmp$S_phase
## g2m_score = nmf_score_final_tmp$G2M_phase
cellcycle_score<- nmf_score_final_tmp$CellCycle
nmf_score_final_without_cc = nmf_score_final_tmp[,1:num_meta_program]
## nmf_score_final_without_cc = subset(nmf_score_final_without_cc, select=-c(S_phase, G2M_phase))
nmf_score_final_without_cc = subset(nmf_score_final_without_cc, select=-c(CellCycle))
num_meta_program = ncol(nmf_score_final_without_cc)

for (i in 1:3){
  nmf_score_final_without_cc = metagene_score_signature(nmf_score_final_without_cc, num_meta_program, i)
}

## nmf_score_final_without_cc$S_phase = s_score
## nmf_score_final_without_cc$G2M_phase = g2m_score
nmf_score_final_without_cc$CellCycle = cellcycle_score
## nmf_score_final_without_cc$CC_binary = apply(nmf_score_final_without_cc[,c("S_phase", "G2M_phase")], 1,
                                             ## function(x) max(x)>=1)
nmf_score_final_without_cc$CC_binary = sapply(nmf_score_final_without_cc[,c("CellCycle")],
                                             function(x) max(x)>=1)
saveRDS(nmf_score_final_without_cc, file=paste0(nmf_analysis_folder, "nmf_score_updated", NameOfMetaprogramList, ".rds"))
```

## Save metaprogram genes as a txt file
```{r}
tmp = matrix(unlist(nmf_marker_genes_final), ncol=length(nmf_marker_genes_final))
colnames(tmp) = names(nmf_marker_genes_final)
tmp = data.frame(tmp)
write.table(tmp, file=paste0(nmf_analysis_folder, "metaprogram_genes", NameOfMetaprogramList, ".txt"),
            sep="\t", quote = F, row.names = F, col.names = T)
```

# Visulization of metaprograms (heatmap and histogram)

## Hierarchical clustering of metagene programs and cells using metagene scores
```{r hc_on_nmf_scores}
## HC on this 
cell_dist = 1-cor(t(nmf_score_final), method="pearson")
hc_cell = hclust(as.dist(cell_dist), method="ward.D2")
metagene_dist = 1-cor(nmf_score_final, method="pearson")
hc_metagene = hclust(as.dist(metagene_dist), method="ward.D2")
```

## Heatmap for metagene scores
```{r heatmap}
metagene_order = hc_metagene$labels[hc_metagene$order]
plotMetaScore(nmf_score_final, nmf_score_final_t$signature_1, metagene_order)
```

## Heatmap for metagene marker genes 
```{r}
plotMetaGeneExpr(cm_center, nmf_score_final, 
                 nmf_score_final_t$signature_1, metagene_order)
```

## Hierarchical clustering on metagene programs 
```{r HC_on_metagenes}
num_metagenes = length(nmf_marker_genes_final)
tmp = nmf_score_final_t[,1:num_metagenes]
pairwise_cor = cor(tmp[,1:num_metagenes])
pairwise_dist = 1 - pairwise_cor
hc = hclust(dist(pairwise_dist), method = "ward.D2")

jpeg(paste0(nmf_fig_folder, "nmf_program_dendrogram",NameOfMetaprogramList, ".jpg"), width = 1200, height = 1800, res = 300)
plot(hc)
dev.off()
```

## Histogram of metagene scores
```{r histogram_of_nmf_scores}
tmp = nmf_score_final_t[,1:num_metagenes]
tmp = melt(tmp)
colnames(tmp) = c("metagene", "score")
ggplot(data=tmp, aes(x=score, fill=metagene)) + geom_histogram() + theme_bw() + facet_wrap(.~metagene)
ggsave(paste0(nmf_fig_folder, "metagene_score_histogram.jpg"), width=12, height=12)
```

# Preliminary annotation of metaprograms 

## ORA with DIPG and GBM genesets
## Compare with metaprograms from previously defined metaprograms from DIPG/GBM
```{r}
DIPG_GBM_genesets = readRDS("../../Marker_genes/Brain_tumor_metaprograms/2019_GBM/all_metaprograms.rds")
ordered_clusters<- order(unique(all_markers_filtered$cluster))
tmp = sapply(ordered_clusters, function(x){
    sapply(DIPG_GBM_genesets, function(y){
        fisher_test(all_markers_filtered[all_markers_filtered$cluster == x, "gene"],
                    y, rownames(seurat_obj@assays$RNA@counts))
    })
})

tmp = data.frame(tmp)


tmp2 = sapply(nmf_marker_genes_final, function(x){
    sapply(DIPG_GBM_genesets, function(y){
        fisher_test(x, y, unlist(union(DIPG_GBM_genesets, nmf_marker_genes_final)))
    })
})
tmp2 = data.frame(tmp2)
apply(tmp2, 2, function(x) {p.adjust(x) < 0.05})
```

## GO-term overrespresentation test 
```{r over-representation-test}
library(clusterProfiler)

## Gene annotations 
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert by biomart
allOE_genes <- as.character(gene_symbol_to_ensembl_id(rownames(cm_center), dataset)$ensembl_gene_id)
allsigOE_genes<- lapply(nmf_marker_genes_final, function(x){
  print(head(x))
  as.character(gene_symbol_to_ensembl_id(x, dataset)$ensembl_gene_id)})
saveRDS(allsigOE_genes, file=paste0(nmf_analysis_folder, "sigOEGenes_Ensembl", NameOfMetaprogramList, ".rds"))

go_result = list()
nmf_factor_names = names(nmf_marker_genes_final)
for (i in 1:length(nmf_marker_genes_final)){
  sigOE_genes = allsigOE_genes[[i]]
  print(sigOE_genes)
  go_result[[nmf_factor_names[i]]] = go_analysis(sigOE_genes, allOE_genes)
} 


## Plot top pathways for all clusters, display on one plot
allDotPlots_20<-list()
for (c in 1:length(nmf_factor_names)){
  program=nmf_factor_names[c]
  print(program)
  allDotPlots_20[[c]]<-dotplot(go_result[[program]]$ego, showCategory=20, font.size = 12, title = program)
}

cowplot::plot_grid(plotlist = allDotPlots_20, ncol=3)
ggsave(paste0("GO_OVA_NMFFactors_Top20",NameOfMetaprogramList, ".jpg"), path=nmf_fig_folder, width=30, height=30)
save(go_result, file=paste0(nmf_analysis_folder, "go_result_metagene_DEGs", NameOfMetaprogramList, ".Robj"))


```