---
title: "NMF_template"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(NMF)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/BySubtype/Midline/Midline_HistoneMutant/")


preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")

source(preprocessing)
source(nmf_helper)

nmf_folder = "analysis/nmf/"
if (!dir.exists(nmf_folder)){
  dir.create(nmf_folder)
}

nmf_fig_folder = "figures/nmf/"
if (!dir.exists(nmf_fig_folder)){
  dir.create(nmf_fig_folder)
}
```

## Set global options for nmf 
```{r set_up_new_lists}
## Set global options for nmf 

## This sets the number of cores for computing NMF (P6 means 6 cores)
## NMF is super computationally heavy, so I usually use all cores; 
## also the computer still uses all cores even I change P6 to P4...
## 'snmf/r' is the algorithm Volker suggests to use and we keep that
## I turned off shared.memory. If it is left on, it may sometimes cause memory leak (likely a bug)
nmf.options(verbose=TRUE, p="vP6", default.algorithm="snmf/r", shared.memory=FALSE)

## ranges of ranks to check out if you decide to computationally estimate the optimal ranks to use (OPTIONAL)
rank_range = 2:8
## Number of NMF repeats for estimating optimal rank (OPTIONAL)
nrun_multi = 30
## Number of NMF repeats for when using the determined/optimal rank for final NMF calculation
nrun_single = 100
## A list to store preprocessed count matrix for each sample 
cm_sample = list()
## A list to store NMF results with differentt ranks for each sample (OPTIONAL)
nmf_estim_r_obj_list = list()
## A list to store the optimal rank for each sample (OPTIONAL)
rank_list = list()
## A list to store the final NMF result object for each sample 
nmf_obj_list = list()
## A list to store the final top 30 weighted genes for each sample 
nmf_gene_list = list()
```

## Load cm, tumor_stat and samples; compute and save overdispersed genes and top 10K variable genes 
```{r load-precomputed-data}
## Just use filtered count matrix 
## Both cells and genes need to be filtered (implemented in the next block)
## Feel free to use other ways to subset genes/cells 
input_dir<- "../../../Preprocessing/SplitBySubtype/analysis/"
load(paste0(input_dir, "midlineHistoneMut_cm_list.Robj"))
samples = readRDS(paste0(input_dir,"midlineHistoneMut_samples.rds"))
tumor_stat = readRDS(paste0(input_dir,"midlineHistoneMut_cluster6CellsNormal_call_tumor.rds"))
cm_list<-midlineHistoneMut_cm_list

## Compute highly variable genes for tumor cells only 
genes_to_keep<-hvgSeurat(cm_list$raw_data[, tumor_stat=="Malignant"], variableFeatures=10000)

## Alternative- pagoda
#ode = hvgPagoda(cm_list$raw_data[, tumor_stat])
#genes_to_keep = rownames(ode)[1:10000]

saveRDS(genes_to_keep, file=paste0(nmf_folder, "top_10000_ode_cluster6CellsNormal.rds"))

## If the analysis is already done
##load("analysis/nmf_obj_list.Robj")
##load("analysis/nmf_gene_list.Robj")
```

## Subset tumor cells and normalize/center cm 
```{r subset_tumor_cells}
## Keep only tumor cells and top 10000 highly variable genes for raw and log transformed data
cm = cm_list$raw_data[rownames(cm_list$raw_data) %in% genes_to_keep,
                      tumor_stat=="Malignant"]
cm_norm = cm_list$norm_data[rownames(cm_list$raw_data) %in% genes_to_keep,
                      tumor_stat=="Malignant"]
cm_mean = log2(rowMeans(cm)+1)

## Re-centering the data for tumor cells only 
cm_center = cm_norm - rowMeans(cm_norm)
samples_tumor = samples[tumor_stat=="Malignant"]

rm(cm_list); gc()
```

## Estimate optimal # of nmf factors 
## This is OPTIONAL, as it is not very effective and super slow 
## I used to do this, but realize computationally determined optimal rank might not be biologically optimal 
## So I switched to specify a pre-determined rank, based on the number of clusters from Seurat's clustering algorithm 
## In the EPN paper, we compared using a rank of 6 vs 10 and got pretty similar results, indicating a rank of 5-10 probably yields similar results
## Most tumors we have worked so far only show 5-10 subpopulations in total, and usually 3-6 in each patient, so I would suggest starting with a rank of 5 or 6, which is also faster than a rank of 10. 
```{r estim-r}
## For each patient (1) filter out genes with 0 in all cells 
## (2) convert negative values to 0
for (current_sample in unique(samples_tumor)){
  cm_sample[[current_sample]] = nmf_df_preprocessing(cm_center[,samples_tumor == current_sample])
}

## OPTIONAL, computationally estimate the optimal rank 
# for (current_sample in unique(samples_tumor)){
#   # perform 30 runs for each value of r in range 2:10
#   message(paste0("Current sample: ", current_sample, "Starting time: ", Sys.time()))
#   nmf_estim_r_obj_list[[current_sample]] <- nmf(cm_sample[[current_sample]], rank_range, nrun=nrun_multi, seed=123456)
#   message(paste0("Ending time: ", Sys.time()))
# }
# save(nmf_estim_r_obj_list, file=paste0(nmf_folder, "nmf_estim_r_obj_list.Robj"))
# ##load("analysis/nmf/nmf_estim_r_obj_list.Robj")

## Manual assignment of ranks for each sample 
for (current_sample in unique(samples_tumor)){
  rank_list[[current_sample]] = 5
}
save(rank_list, file=paste0(nmf_folder, "nmf_rank_list_cluster6CellsNormal.Robj"))
```

## NMF with optimal k and compute top 30 marker genes for each nmf factor 
```{r nmf-fitting-genes}
## NMF with optimal rank 
for (current_sample in names(rank_list)){
  message(paste0("Current sample: ", current_sample, "; Starting time: ", Sys.time(), "\n"))
  nmf_obj_list[[current_sample]] = nmf(cm_sample[[current_sample]],
                                       rank=rank_list[[current_sample]], 
                                       nrun=nrun_single, seed=123456)
  message(paste0("Ending time: ", Sys.time(), "\n"))
}
save(nmf_obj_list, file=paste0(nmf_folder, "nmf_obj_list_cluster6CellsNormal.Robj"))

cat("Computing NMF genes...")
for (current_sample in names(rank_list)){
  ##nmf_gene_list[[current_sample]] = findNmfCorrelatedGenes(basis(nmf_obj_list[[current_sample]]), current_sample, return_list=F)
  nmf_gene_list = append(nmf_gene_list,
                         findNmfCorrelatedGenes(basis(nmf_obj_list[[current_sample]]),
                                                current_sample, return_list=T))
}
save(nmf_gene_list, file=paste0(nmf_folder, "nmf_gene_list_cluster6CellsNormal.Robj"))
```

## NMF scores for each nmf factor 
```{r nmf_score}
## Compute signature score for each NMF factor (30 representative genes) in all cells 
nmf_score = scoreNmfGenes(cm_center, cm_mean, nmf_gene_list)
```

## SD for each nmf factor 
## A previous paper exclude NMF factors with low variation among samples (SD < 0.8)
## But we don't really exclude any NMF factors based on that
```{r SD of cell score}
## SD of cell score for each nmf factor within each sample 
nmf_samples = sapply(rownames(nmf_score), function(x) unlist(strsplit(x, split = "_", fixed = T))[1])
nmf_sd = NULL
for (sample in unique(samples_tumor)){
  tmp = nmf_score[nmf_samples == sample,which(samples_tumor == sample)]
  nmf_sd =  c(nmf_sd, apply(tmp, 1, sd))
}
table(nmf_sd > 0.8)
```


