---
title: "ImmuneRegev_PreviousMetaprogram scoring"
output: html_notebook
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "C:/Users/jenna/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/Immune/Midline_HistoneMutant")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)


seurat_analysis_folder = "analysis/seurat/"
seurat_fig_folder = "figures/seurat/"
nmf_analysis_folder = "analysis/nmf/Merge/"

## Set colors for use later- better for discriminating against groups
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))



## Function for Norm/center count matrix, score signature, then add scores to seurat object
ScoreSig_Seurat<- function(seurat_obj, metaprograms){
  ## Norm and center cm, get mean counts for each gene
  cm_list<-NormCenter(seurat_obj@assays$RNA@counts)
  cm_center<-cm_list$center_data
  cm_mean<-rowMeans(log2(cm_list$raw_data+1))
  
  ## Score, assign each cell to its highest scoring metaprogram
  scores<-lapply(metaprograms,function(x){scoreSignature(cm_center, cm_mean, x)})
  scores<-as.data.frame(scores)
  
  annotation<-colnames(scores)[apply(scores,1,which.max)]
  names(annotation)<-rownames(scores)
  secondHighest<- apply(scores, 1, function(x){names(sort(x, decreasing = TRUE))[2]})
  
  
  maxScore<-apply(scores, 1, max)
  
  
  ## Add metaprogram annotation to seurat object
  seurat_obj@meta.data$PreviousMetaprogram<-annotation
  seurat_obj@meta.data$MaxMetaprogramScore<-maxScore
  seurat_obj@meta.data$SecondHighestMetaprogram<-secondHighest
  
  return(seurat_obj)
}
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj_Cluster6Removed.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}
```


## Load previous metaprograms, subset to needed programs
```{r}
## DIPG/GBM programs
metaprograms_DIPG_GBM_2019<-readRDS(paste0("../../../Marker_genes/Brain_tumor_metaprograms/2019_GBM/",
                                           "all_metaprograms.rds"))
  
metaprograms_DIPG_GBM_2018<-readRDS(paste0("../../../Marker_genes/Brain_tumor_metaprograms/2018_DIPG/",
                                           "all_metaprograms_jenna.rds"))

## Merge DIPG/GBM 2018/2019 together
metaprograms_DIPG_GBM_All<- c(metaprograms_DIPG_GBM_2019[c(1:6,8)],
                              metaprograms_DIPG_GBM_2018[3:6])
names(metaprograms_DIPG_GBM_All)<-c(names(metaprograms_DIPG_GBM_2019)[c(1:6,8)],
                                    names(metaprograms_DIPG_GBM_2018[3:6]))


## Create list of programs for midline samples: AC/OC/OPC from 2018 + MES1 from 2019
midline_programs<- list(AC_like=metaprograms_DIPG_GBM_2018$`AC-like_2018`,
                        OC_like=metaprograms_DIPG_GBM_2018$`OC-like_2018`,
                        OPC_like=metaprograms_DIPG_GBM_2018$`OPC-like_2018`,
                        OPC_like_variable=metaprograms_DIPG_GBM_2018$`OPC-like-variable_2018`,
                        MES1=metaprograms_DIPG_GBM_2019$MES1_2019,
                        Cycling=metaprograms_DIPG_GBM_2018$CellCycle_2018)
saveRDS(midline_programs, file=paste0(nmf_analysis_folder, "midlinePreviousPrograms_OC.AC.OPC.MES1.rds"))
 
```


## Score each seurat count matrix for each sample using its respective metaprogram
```{r}
seurat_obj<- ScoreSig_Seurat(seurat_obj, midline_programs)

## Plot UMAP, coloring by highest scoring program
```{r}
DimPlot(object = seurat_obj, reduction = "umap", pt.size = 1, 
        group.by = "PreviousMetaprogram", cols = col_vector[10:28])
ggsave(paste0(seurat_fig_folder, "MidlineHistoneMutant_PreviousMetaprograms_AC.OC.OPC.MES1.png"))
```


# Pie chart of breakdown of each cluster, programs identified
```{r}
group_colors<- col_vector[1:length(midline_programs)]
names(group_colors)<- names(midline_programs)
  
 
  
## midline
  table_midline<-as.data.frame.matrix(table(seurat_obj@meta.data$PreviousMetaprogram, 
                   seurat_obj@meta.data$seurat_clusters))
  
  all_pie_charts<-list()
  for (cluster in colnames(table_midline)){
    df<- as.data.frame(table_midline[,as.character(cluster)])
    colnames(df)<-"Cluster"
    df$Program<-rownames(table_midline)
    df<-df[df$Cluster>3,] ## at least 3 cells
    p<-ggplot(df, aes(x="", y=Cluster, fill=Program))+
      geom_bar(stat="identity") + 
      coord_polar("y", start=0)+
      ggtitle(paste0("Cluster", cluster))+
      scale_fill_manual(values=group_colors)
    all_pie_charts[[paste0("Cluster", cluster)]]<-p
  }
  
  cowplot::plot_grid(plotlist = all_pie_charts, ncol=2)
  ggsave(paste0(seurat_fig_folder, "PieChartBreakdown_PreviousProgramsOC.AC.OPC.MES1_midline.png"),
         width=10,height=10)
```



## Save seurat objects

```{r}
save(seurat_obj, file= paste0(seurat_analysis_folder, "seurat_obj_Cluster6Removed.Robj"))
```

