---
title: "_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/", 
                     "Pedi_Immune/TumorCellAnnotation/BySubtype/Hemispheric/HistoneWT/BT1745_BT1857/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
stat_helper = paste0(base_dir, script_dir, "common_stat_test.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(stat_helper)
source(nmf_helper)

nmf_analysis_folder = "analysis/nmf/Merge/"
seurat_analysis_folder = "analysis/seurat/"
seurat_fig_folder = "figures/seurat/"
for (folder in c(seurat_analysis_folder, seurat_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder, recursive = TRUE)
  }
}

liger_analysis_folder<- paste0(working_dir, "analysis/liger/")

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
library(Seurat)

n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

subtypes<- read.csv("../../../../../NotUploaded/Subtypes.csv")
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}
```


## Load counts and metadata for initiating seurat object
```{r load-data}
data_dir<- "../../../../Preprocessing/SplitBySubtype/WithNewSamples/analysis/"
load(paste0(data_dir, "hemiHistoneWT_cm_list.Robj"))
samples = readRDS(paste0(data_dir, "hemiHistoneWT_samples.rds"))
tumor_stat = readRDS(paste0(data_dir, "hemiHistoneWT_call_tumor.rds"))

cm_list<-hemiHistoneWT_cm_list

samples<- samples[samples %in% c("BT1745", "BT1857")]
cm_list<- lapply(cm_list, function(x){x[,colnames(x) %in% names(samples)]})
tumor_stat<- tumor_stat[names(tumor_stat) %in% names(samples)]

sum(colnames(cm_list$raw_data)==names(samples)); length(samples)
sum(names(samples)==names(tumor_stat))
```

## Subset cm to remove noise/normal samples
```{r subset-tumor-cells}
cm = cm_list$raw_data; rm(cm_list); gc()
cm = cm[,tumor_stat=="Malignant"]
samples = samples[tumor_stat=="Malignant"]

any(colnames(cm) != names(samples))
```

## Initiate seurat obj and add metadata 
```{r seurat_obj_meta_data}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", min.cells=0, min.genes=0, scale.factor=1E5, do.scale=F, do.center=T)

## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples

sample_df<- as.data.frame(samples); sample_df$cell<- rownames(sample_df); colnames(sample_df)<- c("sample", "cell")
subtype_df<- merge(sample_df, subtypes, by="sample")

seurat_obj@meta.data$subtype<- subtype_df$subtype
seurat_obj@meta.data$AdditionalSubtype<- subtype_df$Diagnosis
```

## OPTIONAL: plot highly variable genes
```{r}
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## PCA and PC selection
```{r PCA_selection}
## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)

```

## Harmony data integration 
```{r run_harmony}
library(harmony)
library(magrittr)
library(cowplot)

## Run Harmony
seurat_obj = RunHarmony(seurat_obj, "sample", theta = 2, 
                        max.iter.harmony = 50, plot_convergence = TRUE)

## Simple dim and vlnplot to examine integration 
options(repr.plot.height = 6, repr.plot.width = 12)

p1 <- DimPlot(object = seurat_obj, reduction = "pca", pt.size = .1, group.by = "sample")
p2<- DimPlot(object = seurat_obj, reduction = "harmony", pt.size = .1, group.by = "sample")
p3 <- VlnPlot(object = seurat_obj, features = "PC_1", pt.size = .1, group.by = "sample")
p4 <- VlnPlot(object = seurat_obj, features = "harmony_1", pt.size = .1, group.by = "sample")

plot_grid(p1,p2, p3, p4)
ggsave("Harmony_diagnostics_theta2.jpg", path=seurat_fig_folder, width=12, height=8)


## DimHeatmap
DimHeatmap(object = seurat_obj, reduction = "harmony", cells = 500, dims = 1:6, balanced = TRUE)
```

## Clustering and UMAP
```{r clustering_tSNE}
## Run UMAP and clustering using reduced dimensions from Harmony
## You may play around with resolution parameter to tune the clustering result
resolution = .8
reduction_type = "harmony"
dims = 20

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = reduction_type, dims = 1:dims)


seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = reduction_type,
                            dims = 1:dims,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = resolution)


r0.8<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5) + ggtitle("Resolution_0.8")
r0.6<-DimPlot(seurat_obj, reduction="umap", pt.size = 2.5)+ ggtitle("Resolution_0.6")
r1.4<-DimPlot(seurat_obj, reduction = "umap", pt.size = 2.5)+ ggtitle("Resolution_1.4")

r0.4+r0.8+r1.4 #resolution of 0.8 may be best
ggsave("UMAP_clustering_VaryResolution.jpg", path=seurat_fig_folder, width=20, height=6)

## Clustering results 
table(Idents(seurat_obj))


## Save the seurat object
save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```

## UMAP plot colored by clustering and samples
## Can color tSNE with other metadata features (e.g. primary/recurrence, treatment)
```{r UMAP_plots}
## By clustering
DimPlot(object = seurat_obj, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 3, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_Resolution0.8.jpg", path=seurat_fig_folder, width=8, height=8)

## By sample
DimPlot(object = seurat_obj, reduction="umap", group.by = "sample", 
        label = TRUE, pt.size = 3, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_Resolution0.8.jpg", path=seurat_fig_folder, width=8, height=8)

```



## DEGs for each cluster 
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.5)
all_markers_filtered = all_markers[all_markers$p_val_adj<0.05 & all_markers$avg_logFC>0.5,]


## Subset the top 30 DEGs for each cluster
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC) %>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(seurat_analysis_folder, "all_markers.Robj"))
write.csv(top_marker_genes, file=paste0(seurat_analysis_folder, "Top30Markers_Theta2.csv"))

## Heatmap for initial inspection
DoHeatmap(seurat_obj, 
          features=all_markers_filtered$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs.jpg", path=seurat_fig_folder)


```

## Quick annotation by OPC/OC/AC marker genes 
```{r}
## OPC-like
FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OPC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## OC-like cells
FeaturePlot(seurat_obj,
            c("BCAS1", "SIRT2", "RGR", 
              "GPR17", "NFASC", "MBP",
              "TF", "MOG", "PLP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## AC-like cells 
FeaturePlot(seurat_obj,
            c("APOE", "AQP4", "AGT", 
              "SPARCL1","ATP1A2", "GFAP",
              "S100B", "ALDOC", "CLU"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("AC_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## Neuronal-like cells
FeaturePlot(seurat_obj,
            c("STMN2", "STMN1", "STMN4", 
              "TUBB3", "INA", "ELAVL4",
              "SNAP25", "GABRG2", "ST18"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("Neuron_markers.jpg", path=seurat_fig_folder, width=9, height=9)

##CLEC2D
FeaturePlot(seurat_obj,
            c("CLEC2D"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("CLEC2D_KLRB1_markers.jpg", path=seurat_fig_folder, width=9, height=4.5)
```

## ORA test 
```{r over-representation-test}
library(clusterProfiler)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
allOE_genes <- as.character(gene_symbol_to_ensembl_id(rownames(seurat_obj@assays$RNA@counts),dataset)$ensembl_gene_id)
save(allOE_genes, file="analysis/seurat/allOE_genes.Robj")


## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
sigOE_genes_list<-list()
for (i in unique(all_markers_filtered$cluster)){
  sigOE_genes = as.character(gene_symbol_to_ensembl_id(subset(all_markers_filtered,
                                                              cluster==i,gene)[,],
                                                       dataset)$ensembl_gene_id)
  print(head(sigOE_genes))
  sigOE_genes_list[[i]]<-sigOE_genes
  
} 
go_result = list()
for (i in names(sigOE_genes_list)){
  sigOE_genes<-sigOE_genes_list[[i]]
  go_result[[paste0("c", i)]] = go_analysis(sigOE_genes, allOE_genes)
}


## Plot top pathways for all clusters, display on one plot
clusters<-names(go_result)
allDotPlots_20<-list()
for (c in 1:length(clusters)){
  cluster=clusters[c]
  print(cluster)
  if (nrow(go_result[[cluster]]$cluster_summary)>0){
    print(cluster)
    allDotPlots_20[[c]]<-dotplot(go_result[[cluster]]$ego, 
                                showCategory=20, 
                                font.size = 15, 
                                title = cluster, 
                                label_format=10) +
      scale_y_discrete(labels=function(x) str_wrap(x, width=60))+
      theme(plot.title = element_text(hjust = 0.5, face = "bold", 
                                      color="black", size = 28),
            axis.title = element_text(face = "bold", color="black"), 
            axis.text.x = element_text(angle = 45, hjust = 1, color="black"),
            axis.text.y = element_text( color="black",face = "bold"))
  }
  
}

allDotPlots_20<- allDotPlots_20[!(unlist(lapply(allDotPlots_20, function(x){is.null(x)})))]
cowplot::plot_grid(plotlist = allDotPlots_20, ncol=2)
ggsave(paste0("GO_OVA_AllClusters_top20_Cluster7TopBottom.jpg"), path=seurat_fig_folder, width=18, height=12)
save(go_result, file=paste0(seurat_analysis_folder,  "go_result.Robj"))
```

## ORA with hemispheric genesets
## Compare with metaprograms from previously defined metaprograms from ependymoma
```{r}
all_markers_list<-split(all_markers_filtered, all_markers_filtered$cluster )
all_markers_list<-lapply(all_markers_list, function(x){x$gene})
names(all_markers_list)<-paste0("Cluster", names(all_markers_list))

hemi_genesets<-readRDS(paste0("../", nmf_analysis_folder,
                                 "hemisphericPreviousPrograms_2019_OC.rds"))

tmp2 = sapply(all_markers_list, function(x){
    sapply(hemi_genesets, function(y){
        fisher_test(x, y, unlist(union(hemi_genesets, all_markers_list)))
    })
})
tmp2 = data.frame(tmp2)
tmp3<- apply(tmp2, 2, function(x) {p.adjust(x) < 0.1})
tmp4<-apply(tmp2, 2, function(x) {p.adjust(x)})

write.table(tmp4, paste0(seurat_analysis_folder,
                        "ScoreSeuratClusters_NeftelPrograms_padj.txt"))

```

# Score cells based on Neftel GBM programs
```{r}
hemi_genesets<-readRDS(paste0("../", nmf_analysis_folder,
                                 "hemisphericPreviousPrograms_2019_OC.rds"))
gene_list<-hemi_genesets

## Norm and center my data
cm_center<- NormCenter(seurat_obj@assays$RNA@counts)$center_data
cm_mean<- log2(rowMeans(seurat_obj@assays$RNA@counts)+1)

## score my data for each signature
scores<- scoreNmfGenes(cm_center, cm_mean, gene_list)
scores = t(scores)

topScore<- apply(scores[,1:length(gene_list)], 1, 
                  function(x) sort(x, decreasing = T)[1])
topSig<-apply(scores[,1:length(gene_list)], 1, 
                  function(x) names(sort(x, decreasing = T))[1])

## Add to seurat metadata, plot
seurat_obj@meta.data$TopSigPrevious<- topSig
DimPlot(seurat_obj, group.by="TopSigPrevious", pt.size = 2) + NoAxes()
ggsave(paste0(seurat_fig_folder, "ScoredBasedOnNeftelPrograms.png"))

## Add 2nd top score to seurat metadata, plot
secondTopSig<-apply(scores[,1:length(gene_list)], 1, 
                  function(x) names(sort(x, decreasing = T))[2])
seurat_obj@meta.data$SecondTopSigPrevious<- secondTopSig
DimPlot(seurat_obj, group.by="SecondTopSigPrevious", pt.size = 2) + NoAxes()
ggsave(paste0(seurat_fig_folder, "ScoredBasedOnNeftelPrograms_SecondHighest.png"))

## Grey out all cells except cells that scored highest for MES1, then show their second highest
df<- seurat_obj@meta.data[,c("TopSigPrevious", "SecondTopSigPrevious" )]
df$cell<-rownames(df)
mes1TopSig<- df[df$TopSigPrevious=="MES1_2019", c("SecondTopSigPrevious", "cell")]
mes1_keep<- mes1TopSig$SecondTopSigPrevious; names(mes1_keep)<- mes1TopSig$cell

mes1_greyOut<- colnames(seurat_obj)[!(colnames(seurat_obj) %in% names(mes1_keep))]
greyOut<- rep("grey", length(mes1_greyOut)); names(greyOut)<-mes1_greyOut

all_greyOut<- c(mes1_keep, greyOut)
all_greyOut<- all_greyOut[order(names(all_greyOut))]

seurat_obj@meta.data$GreyOutSecondTopMes1<- all_greyOut
DimPlot(seurat_obj, group.by="GreyOutSecondTopMes1", cols = c("pink", "yellow", "grey",
                                                              "cyan", "blue", "magenta"))
ggsave(paste0(seurat_fig_folder, "SecondTopProgram_Mes1CellsOnly.png"))

## Add 3rd top score to seurat metadata, plot
thirdTopSig<-apply(scores[,1:length(gene_list)], 1, 
                  function(x) names(sort(x, decreasing = T))[3])
seurat_obj@meta.data$ThirdTopSigPrevious<- thirdTopSig


## Grey out all cells except cells that scored highest for MES1, then show their second highest
df<- seurat_obj@meta.data[,c("TopSigPrevious", "ThirdTopSigPrevious" )]
df$cell<-rownames(df)
mes1TopSig<- df[df$TopSigPrevious=="MES1_2019", c("ThirdTopSigPrevious", "cell")]
mes1_keep<- mes1TopSig$ThirdTopSigPrevious; names(mes1_keep)<- mes1TopSig$cell

mes1_greyOut<- colnames(seurat_obj)[!(colnames(seurat_obj) %in% names(mes1_keep))]
greyOut<- rep("grey", length(mes1_greyOut)); names(greyOut)<-mes1_greyOut

all_greyOut<- c(mes1_keep, greyOut)
all_greyOut<- all_greyOut[order(names(all_greyOut))]

seurat_obj@meta.data$GreyOutThirdTopMes1<- all_greyOut
DimPlot(seurat_obj, group.by="GreyOutThirdTopMes1", cols = c("pink", "yellow", "grey",
                                                              "cyan", "blue", "magenta",
                                                             "yellow", "orange"))
ggsave(paste0(seurat_fig_folder, "ThirdTopProgram_Mes1CellsOnly.png"))
```

## Assign clusters to cell type
## Primarily based on marker genes, slightly on Neftel programs
```{r}
clusters<- as.character(unique(seurat_obj@active.ident))
names(clusters)<- c("AC-like","Cycling", "OPC-like")

seurat_obj$ManualAnnot<- plyr::mapvalues(seurat_obj$seurat_clusters,
                                         from=clusters, to=names(clusters),
                                         warn_missing = FALSE)

save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```






## Make pseudobulk
```{r}
metagene_program_names = sort(unique(nmf_score$signature_1))
pseudobulk = NULL
aggr_method = "mean"
for (metagene in metagene_program_names){
  cm = seurat_obj@assays$RNA@counts[,nmf_score$signature_1 == metagene & nmf_score$score_1 >= 1]
  if (aggr_method == "mean"){
    pseudobulk = cbind(pseudobulk, rowMeans(cm))
  }
  if (aggr_method == "median"){
    pseudobulk = cbind(pseudobulk, rowMedians(cm))
  }
}
colnames(pseudobulk) = metagene_program_names
rownames(pseudobulk) = rownames(seurat_obj@assays$RNA@counts)
saveRDS(pseudobulk, file=paste0(seurat_analysis_folder, 
                                "pseudobulk_cm_unique_only_", aggr_method, NameOfMetaprogramList,".RDS"))
```

## Metagene program specific genes as overall DEGs
```{r}
seurat_obj = SetIdent(seurat_obj, value = seurat_obj@meta.data$signaure_1_without_cc)

all_markers_nmf = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_nmf = all_markers[all_markers$p_val_adj < 0.05, ]
saveRDS(all_markers, paste0(seurat_analysis_folder, "deg.rds"))
##sapply(names(nmf_marker_genes_final), function(x) table(nmf_marker_genes_final[[x]] %in% gene_signatures[[x]]))
signature_genes = unique(all_markers$gene)
saveRDS(signature_genes, paste0(seurat_analysis_folder, "signature_genes.rds"))
```

## Updated annotation: same annotations used for CellChat
## UMAP and sample breakdown
```{r}
seurat_obj$CellChatAnnot<-seurat_obj$ManualAnnot

## Set colors so they're shared across all subtypes
load("../../../cellTypeColors.Robj")
cellTypes<- unique(seurat_obj$CellChatAnnot)
colorsUse<-colorsUse
save(colorsUse, file="../../../cellTypeColors.Robj")

DimPlot(object = seurat_obj, reduction="umap", group.by = "CellChatAnnot", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend() + scale_color_manual(values=colorsUse)
ggsave("UMAP_clustering_CellChatAnnot.jpg", path=seurat_fig_folder, width=6, height=6)


annotBreakdown<- as.data.frame(table(seurat_obj$sample, seurat_obj$CellChatAnnot))
colnames(annotBreakdown)<- c("Sample", "CellType", "NCells")

plot_list<-list()
for(i in unique(annotBreakdown$Sample)){
  df<- annotBreakdown[annotBreakdown$Sample==i,]
  p<-ggplot(df, aes(x="", y=NCells, fill=CellType))+
    geom_bar(stat="identity", color="white")+
    coord_polar("y", start=0)+
    theme_void()+
    ggtitle(i)+theme(plot.title = element_text(size=15, face="bold", hjust=0.5))+
    scale_fill_manual(values=colorsUse)
  plot_list[[i]]<- p
}
cowplot::plot_grid(plotlist=plot_list, ncol=1)
ggsave(paste0(seurat_fig_folder, "CellBreakdown.png"), width=3, height=3)
```