---
title: "R Notebook"
output: html_document
---

```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02a_Tcells/01b_Adult/", 
                     "02_ClonalOverexpression/")
library(Seurat)
library(ggplot2)
library(ggpubr)
library(Rtreemix)

figure_dir_tmp<- "figures/"
analysis_dir_tmp<- "analysis/"
data_dir<- "fromO2/"

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(nmf_helper)
clonalOverexpression<- paste0(base_dir, script_dir, "CellChat_HelperFunctions.R")
source(clonalOverexpression)

```

## Set figure dir/analysis dir
```{r}
rank_use<-8
sharedBy<- "union" 

figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "_sharedBy", sharedBy, "/")
```

## For each analysis:
1) read in results
2) Plot expression/CV z scores, labeling GOI + outliers
3) Plot expression of outliers in clonal/nonclonal cells to confirm
4) Save top 100 outliers for comparison
```{r}
all_files<- list.files(data_dir, pattern=".Rds")
all_files<- all_files[grepl("WithBackground", all_files)]
analysis_names<-gsub("WithBackgroundx100_ClonalVsNonclonal_Expression.CV_Zscores_adult_cd8_", "", gsub(".Rds", "", all_files))


GOI<- c("KLRB1", "GZMB", "NKG7", "CCL5", "IL7R", "CCR7", "KLRK1", "ITGAE", "CCL4", "KLRD1", "GPR56", "PTGER2", "SELPLG")
GOI<- c("KLRB1","PTGER2", "SELPLG")
outliers_plot<- 20

all_outliers<- list()
for(analysis in analysis_names){
  print(analysis)
  file<- paste0(data_dir, "WithBackgroundx100_ClonalVsNonclonal_Expression.CV_Zscores_adult_cd8_", analysis, ".Rds")
  
  ## 1. Read in CNC res + seurat object
  cnc_res<- readRDS(file)
  cnc_res<- cnc_res[!is.na(cnc_res$Expr_background_zscore) & !is.na(cnc_res$CV_background_zscore),]
  seurat_list<- readRDS(paste0(analysis_dir, "adult_cd8_seurat_cnc_preprocessed_",  analysis, ".Rds"))
  
  ## 2. Plot difference in expression vs difference in CV 
    ## GOI to color/label
    cnc_res$GOI<- ifelse(rownames(cnc_res) %in% GOI, "yes", "no")
    cnc_res$label<-ifelse(rownames(cnc_res) %in% GOI, rownames(cnc_res), "")
      
    ## assign outliers- expression + cv filter, then top n by expression
    outliers_all<- cnc_res[cnc_res$Expr_background_zscore>1.5 & cnc_res$CV_background_zscore < -1.8,] ## was -1.2
    outliers_all<- outliers_all[order(outliers_all$Expr_background_zscore, decreasing =TRUE),]
    outliers<- rownames(outliers_all)[1:outliers_plot]
    
    ## Label as GOI, outlier
    cnc_res$GOI<- ifelse(rownames(cnc_res) %in% GOI, "GOI", ifelse(rownames(cnc_res) %in% outliers, "outlier", "no"))
    cnc_res$label<-ifelse(rownames(cnc_res) %in% c(outliers, GOI), rownames(cnc_res), "")
    
  ggplot(cnc_res, aes(x=CV_background_zscore, y=Expr_background_zscore, color=GOI))+
        geom_point(aes(alpha=GOI))+
        theme_classic()+
        #scale_color_manual(values=c(yes="red", no="grey71"))+
        scale_color_manual(values=c(GOI="red",outlier="black", no="grey71"))+
        geom_text(aes(label=label, alpha=GOI, size=GOI),  nudge_x = 0.4, nudge_y=-0.1)+
        scale_alpha_manual(values=c(GOI=1, outlier=.7, no=0.2))+
        scale_size_manual(values=c(GOI=3.5, outlier=2, no=0))+
        theme(legend.position = "none",
              axis.title=element_text(face="bold", color="black"),
              axis.text = element_text(face="bold", color="black"))+
        xlim(-6.5,6.5)+
        ggtitle(analysis)+
        ylab("Expression background Z score")+
        xlab("CV background Z score")
    
    ggsave(paste0(figure_dir, analysis, "_ExprCVZscores_GOI.outliers_editedFigure.pdf"),device="pdf", width=6, height=4.5)
    
  ## 3. Plot top 20 genes in DotPlot
    ## merge clonal/nonclonal seurat objects
    cnc_seurat<- merge(seurat_list$clonal, seurat_list$nonclonal)
    DotPlot(cnc_seurat, features=unique(c(GOI, outliers)), group.by = "clonal")+
      coord_flip()+
      ylab("") + xlab("")+
      ggtitle(analysis)+
      theme(axis.text.x = element_text(angle=45, hjust=1))
    ggsave(paste0(figure_dir, analysis, "_DotPlot_GOI.outliers.png"), width=5, height=8)
    
  ## 4. Save top outliers- up to 50
    outliers_all$rank<- 1:nrow(outliers_all)
    top_outliers<- na.omit(outliers_all[1:100,c("Expr_Clonal", "Expr_NonClonal", 
                                       "CV_clonal", "CV_NonClonal",
                                       "Expr_background_zscore", "CV_background_zscore", "rank")])
    top_outliers$analysis<- analysis
    top_outliers$gene<- rownames(top_outliers)
    all_outliers[[analysis]]<- top_outliers
}
all_outliers<- do.call("rbind", unname(all_outliers))

## Add column denoting how many analyses each gene is an outlier in. export
nOcc<- as.data.frame(table(all_outliers$gene))
all_outliers$nAnalysisFoundIn<- as.character(plyr::mapvalues(all_outliers$gene, nOcc$Var1, nOcc$Freq))
write.csv(all_outliers, file=paste0(analysis_dir, "top100.outliers_allAnalyses.csv"))

```


## Venn diagram of top 100 outliers across all analyses
## Not run now- can run if IDH-mut analysis is completed
```{r}
all_outliers<- read.csv(paste0(analysis_dir, "top100.outliers_allAnalyses.csv"))

## Split to list by analysis
outlier_list<- split(all_outliers, f=all_outliers$analysis)

## separate lists- all/subtypes. Subset to # of outliers in smallest group
  
  ## all/subtypes
  subtype_outlier_list<- outlier_list[names(outlier_list) 
                                      %in% c(names(outlier_list)[!names(outlier_list) %in% names(age_outlier_list)])]
  min_length<- min(unlist(lapply(subtype_outlier_list, nrow)))
  subtype_outlier_list<- lapply(subtype_outlier_list, function(x){
    tmp<- x[order(x$rank),]
    tmp<- tmp[1:min_length,]
    return(tmp$gene)
  })


## Venn diagram: age
ggVennDiagram::ggVennDiagram(age_outlier_list, label="count")+
  scale_fill_gradient(low="white", high = "red")+
  scale_color_manual(values=rep("black", 3))+
  ggtitle(paste0("top ", length(age_outlier_list$all), " outliers"))
ggsave(paste0(figure_dir, "Venn_Outliers_all.vs.under4.vs.over4.png"), width=4, height=4)


## Venn diagram: subtype
ggVennDiagram::ggVennDiagram(subtype_outlier_list, label="count")+
  scale_fill_gradient(low="white", high = "red")+
  scale_color_manual(values=rep("black", 6))+
  ggtitle(paste0("top ", length(subtype_outlier_list$Ependymoma), " outliers"))
ggsave(paste0(figure_dir, "Venn_Outliers_subtypes.png"), width=5, height=5)
```



#################################
## Nothing run past this point ## 
#################################




## Alternative to background zscore approach: differential expression
```{r}
## Merge clonal/nonclonal seurat objects
seurat_obj<- merge(seurat_list$clonal, seurat_list$nonclonal)


## Set ident to clonality, find markers
seurat_obj<- SetIdent(seurat_obj, value=seurat_obj$clonal)
all_markers<- FindAllMarkers(seurat_obj, min.pct = 0.2)
all_markers_filtered<- all_markers[all_markers$p_val<0.05 & all_markers$avg_log2FC>0,]

DotPlot(seurat_obj, features =c(outliers[1:2], "PTGER2", "KLRB1"))
  coord_flip()
```




## Volcano plot
## Plotting zscores as pvalue (-log10) vs logFC
```{r}
GOI<- c("KLRB1", "GZMB", "NKG7", "CCL5", "IL7R", "CCR7", "KLRK1", "ITGAE", "CCL4", "KLRD1", "GPR56")
outliers<- c("GNLY", "WBP11", "CCL4L2", "CCL4L1", "PRDM1", "SRP9", "PGK1", "SAMSN1", "EIF3C", "TMEM241", 
             "KDELR2", "ILKAP", "CIR1", "SNRPD2", "WIPF1", "CLNS1A", "CLIC3")

## pvalues
  ## Calculate based on z score
  vplot_df<- df_final
  vplot_df$zscore_pvalue<- ifelse(vplot_df$Expr_Clonal>vplot_df$Expr_NonClonal, 
                           pnorm(q=vplot_df$Expr_background_zscore, lower.tail=FALSE),
                           pnorm(q=vplot_df$Expr_background_zscore, lower.tail=TRUE))

                         

## Log FC
vplot_df$Expr_log2FC<- log2(vplot_df$Expr_Clonal/vplot_df$Expr_NonClonal) 
vplot_df<- na.omit(vplot_df)



## volcano with ggplot
  ## select pvalue to use
  vplot_df$pvalue<- vplot_df$zscore_pvalue
  
  ## add adjusted pvalue. will color points that are also sig with MTC
  vplot_df$adj_pvalue<- p.adjust(vplot_df$pvalue, method = "fdr")
  vplot_df$Significant_AfterMultipleTestCorrection<- ifelse(vplot_df$adj_pvalue<0.1 & vplot_df$Expr_log2FC>0, "yes", "no")
  padj_genes<- rownames(vplot_df)[vplot_df$Significant_AfterMultipleTestCorrection=="yes"]
  
vplot_df$log10pvalue<- -log10(vplot_df$pvalue)
sig_genes<- rownames(vplot_df)[vplot_df$pvalue<0.05 & vplot_df$Expr_log2FC>1.5]

vplot_df$GOI<- ifelse(rownames(vplot_df) %in% GOI, "GOI", 
                      ifelse(vplot_df$Significant_AfterMultipleTestCorrection=="yes", "Putative GOI\nsignificant after MTC", 
                             ifelse(rownames(vplot_df) %in% sig_genes, "Putative GOI", "no")))
vplot_df$label<- ifelse(rownames(vplot_df) %in% c(GOI, sig_genes, padj_genes), rownames(vplot_df), "")


ggplot(vplot_df, aes(y=log10pvalue, x=Expr_log2FC, color=GOI, alpha=GOI, size=GOI))+
  geom_point()+
  scale_color_manual(values=c(`Known GOI`="red", 
                              `Putative GOI`="black", 
                              `Putative GOI\nsignificant after MTC`="blue", 
                              PTGER2="darkgreen",
                              no="grey71"))+
  geom_text(aes(label=label), size=1.8, nudge_x = 0.12, nudge_y=-0.05)+
  scale_size_manual(values=c(`Known GOI`=1.5, 
                              `Putative GOI`=1.5, 
                              `Putative GOI\nsignificant after MTC`=1.5, 
                             PTGER2=1.5,
                              no=1.5))+
  scale_alpha_manual(values=c(`Known GOI`=1, 
                              `Putative GOI`=1, 
                              `Putative GOI\nsignificant after MTC`=1,
                              PTGER2=1,
                              no=0.2))+
  theme_classic()+
  geom_hline(yintercept = 1.3, color="red", linetype="dotted")+
  geom_vline(xintercept = 1.5, color="red", linetype="dotted")+
  geom_vline(xintercept = -1.5, color="red", linetype="dotted")+
  xlim(-3,3)
ggsave(paste0(figure_dir, "VolcanoPlot_FDRInfo_", analysis_append, ".png"), width=12, height =6)

write.csv(vplot_df, file=paste0(analysis_dir, "LogFC_pvalues_WithFDR_", analysis_name, ".csv"))
```


## Pathway analysis using top genes
```{r}
load(paste0(analysis_dir, "adult_cd8_seurat_cnc_preprocessed.Robj"))
load(paste0(analysis_dir, "adult_gbm_cd8_seurat_cnc_preprocessed.Robj"))

seurat_list<- adult_seurat_cnc
analysis_name<-"test_norm_noClonoMerge_PedCD8"


library(clusterProfiler)
library(biomaRt)
library(stringr)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
ensembl = useMart("ensembl",dataset="hsapiens_gene_ensembl")
allOE_genes<-getBM(attributes='ensembl_gene_id', 
                   filters = 'hgnc_symbol', 
                   values = rownames(seurat_list$clonal@assays$RNA@counts), 
                   mart = ensembl)


## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
genes<- rownames(cnc_res[cnc_res$Expr_background_zscore>2,])
sigOE_genes = getBM(attributes='ensembl_gene_id', 
                        filters = 'hgnc_symbol', 
                        values = genes,
                        mart = ensembl)$ensembl_gene_id
go_results = go_analysis(sigOE_genes, allOE_genes$ensembl_gene_id)
 



## Plot top pathways for all clusters, display on one plot
dotplot(go_results$ego, 
        showCategory=30, 
        font.size = 15, 
        label_format=10) +
     scale_y_discrete(labels=function(x) str_wrap(x, width=60))+
      theme(plot.title = element_text(hjust = 0.5, face = "bold", 
                                      color="black", size = 28),
            axis.title = element_text(face = "bold", color="black"), 
            axis.text.x = element_text(angle = 45, hjust = 1, color="black"),
            axis.text.y = element_text( color="black",face = "bold"))



ggsave(paste0("GO_ORA_top30_", analysis_name, ".png"), path=figure_dir, width=9, height=12)
save(go_result, file=paste0(seurat_analysis_folder, "go_result_NoInt_BetweenSamples.Robj"))
```


## Histogram of zscores for expr and cv
```{r}

expr_hist<- ggplot(cnc_res, aes(x=Expr_background_zscore))+ 
  geom_histogram() + 
  ggtitle("expression zscores: compared to background")

cv_hist<- ggplot(cnc_res, aes(x=CV_background_zscore))+ 
  geom_histogram() + 
  ggtitle("cv zscores: compared to background")

expr_hist + cv_hist
```


## Exploratory- can delete later
```{r}
## Expect potential correlation b/w SELPLG various exhaustion markers- test
GOI<- c("SELPLG", "PDCD1", "HAVCR2", "LAG3", "CD160", "BTLA", "IL7R", "IL2RA")

## subset to predysfunctional, cytotoxic

seurat_use<- subset(cd8_seurat, NMF_rank8_annot_mergedByunion_reAnnot =="Inhibitory.Cycling")

goi_expr_full<- as.data.frame(t(as.data.frame(seurat_use@assays$RNA@data[GOI,])))
goi_expr_full$age<- plyr::mapvalues(rownames(goi_expr_full), colnames(cd8_seurat), cd8_seurat$Age, warn_missing = FALSE)
goi_expr_full$clonal<- plyr::mapvalues(rownames(goi_expr_full), colnames(cd8_seurat), cd8_seurat$clonal, warn_missing = FALSE)

all_plots<- lapply(GOI[GOI != "SELPLG"], function(x){
  goi_expr<- goi_expr_full
  goi_expr$y<- goi_expr[[x]]
  ggplot(goi_expr, aes(x=SELPLG, y=y, color=clonal))+
    geom_point()+
    theme_classic()+
    facet_grid(cols = vars(age))+
    ylab(paste0(x, " logTPM"))+
    xlab("SELPLG logTPM")+
    scale_color_manual(values=c(clonal="red", nonclonal="grey"))+
    geom_smooth(method="lm", se=FALSE, color="black")+
    ggtitle(paste0(x, " vs SELPLG"))
})
cowplot::plot_grid(plotlist = all_plots, ncol=4)
ggsave(paste0(figure_dir, "SELPLGexpr.vs.exhaustionMarkers.png"), width=18, height=6)

## same as above, remove 0 expressors
all_plots<- lapply(GOI[GOI != "SELPLG"], function(x){
  goi_expr<- goi_expr_full
  goi_expr$y<- goi_expr[[x]]
  goi_expr<- goi_expr[goi_expr$y>0 & goi_expr$SELPLG>0,]
  ggplot(goi_expr, aes(x=SELPLG, y=y, color=clonal))+
    geom_point()+
    theme_classic()+
    facet_grid(cols = vars(age))+
    ylab(paste0(x, " logTPM"))+
    xlab("SELPLG logTPM")+
    scale_color_manual(values=c(clonal="red", nonclonal="grey"))+
    geom_smooth(method="lm", se=FALSE, color="black")+
    ggtitle(paste0(x, " vs SELPLG"))
})
cowplot::plot_grid(plotlist = all_plots, ncol=4)
ggsave(paste0(figure_dir, "SELPLGexpr.vs.exhaustionMarkers_no0Counts.png"), width=18, height=6)
```


