---
title: "TCR"
author: "Jenna LaBelle"
date: "10/07/22"
output: html_document
---

## Plot TCR results, clonotype expansion. Add to seurat object

```{r}
library(Seurat) 

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/tenX/Immune/")

analysis_dir<- paste0(working_dir, "/analysis/TCR/")
figure_dir<- paste0(working_dir, "/figures/TCR/")
full_cohort_dir<- "../../"

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

sample_colors<- c(BT1857="grey30", BT2082="olivedrab", BT2083="violetred")
```

## Helper functions
```{r}
## Helper function to plot v/j gene usage
my_plotVJgeneUsage<- function(seurat, plot_v_or_j, n_genes_plot){
  ## Get df with all cells with V/J gene
  tmp<- seurat@meta.data
  tmp<- tmp[rownames(tmp)!=tmp[plot_v_or_j],]
  tmp<- as.data.frame(table(tmp[plot_v_or_j]))
  
  ## Freq of V/J gene
  tmp<- tmp[order(tmp$Freq, decreasing = TRUE),]
  
  ## Subset to top n V/J genes, order
  tmp<- tmp[1:n_genes_plot,]
  tmp$Var1<- factor(tmp$Var1, levels=tmp$Var1)
  
  ## Plot
  ggplot(tmp, aes(x=Var1, y=Freq))+
    geom_bar(stat="identity")+
    theme_classic()+
    theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"))+
    xlab(plot_v_or_j)+
    ylab("Occurence of gene")
}

## Helper function to get # of clonotypes per program, normalized by total number of cells per program
my_getNormNClonotype<- function(seurat, program_column="Program_projected", clonotype_id_column="raw_clonotype_id"){
  tmp<- seurat@meta.data
  tmp<- tmp[rownames(tmp)!=tmp[clonotype_id_column],] ## getting rid of cells without tcr info
  tmp<-as.data.frame(table(tmp[,clonotype_id_column], tmp[,program_column]))
  colnames(tmp)<- c("Clonotype", "Program", "Ncells")
  
  ## Number of unique clonotypes present in each program
  n_clonotypes_perProgram<- sapply(unique(as.character(tmp$Program)), function(program){
    program_tmp<- tmp[tmp$Program==program & tmp$Ncells>0,]
    return(length(unique(program_tmp$Clonotype)))
  })
  names(n_clonotypes_perProgram)<- unique(as.character(tmp$Program))
  
  
  ## Normalize number of clonotypes by number of cells
  n_CC_perProgram<- data.frame(Program=names(n_clonotypes_perProgram),
                               n_clonotypes=n_clonotypes_perProgram,
                               n_cells= as.numeric(table(seurat$Program_projected)))
  n_CC_perProgram$n_clonotypes_norm<- n_CC_perProgram$n_clonotypes/n_CC_perProgram$n_cells
  return(n_CC_perProgram)
}

## Helper function to get largest clonotype per program
my_getMaxClonotype<- function(seurat, program_column="Program_projected", clonotype_id_column="raw_clonotype_id"){
  tmp<- seurat@meta.data
  tmp<- tmp[rownames(tmp)!=tmp[clonotype_id_column],] ## getting rid of cells without tcr info
  tmp<-as.data.frame(table(tmp[,clonotype_id_column], tmp[,program_column]))
  colnames(tmp)<- c("Clonotype", "Program", "Ncells")

  
  ## get largest clonotype for each program
  max_clonotype<- tmp %>% group_by(Program) %>% summarise(maxClonotypeSize=max(Ncells)) %>% as.data.frame()
  return(max_clonotype)
}
```

## Load data: tcell seurat objects, TCR results, colors
NOTE: vdj to be run on BT2080, will incorporate when finished (+ hopefully after rerunning 10x seq on bt2080)
```{r}
## Seurat
seurat_obj<- readRDS("analysis/preprocessing/seurat_obj_noTumor.Rds")

## TCR
samples<- c("BT1857", "BT2082", "BT2083")
tcr_list<- lapply(samples, function(x){
  tcr<-read.csv(paste0("data/",x,"/VDJ/multi/per_sample_outs/", x, "/vdj_t/filtered_contig_annotations.csv"))
  tcr$sample<- x
  
  ## for bt1857, add as "E167CD35resub"- that's how all tcells are named (only CD3+ used for VDJ)
  tcr$barcode<- gsub("-", "_", tcr$barcode)
  tcr$barcode<- ifelse(tcr$sample=="BT1857", paste0("E167CD35resub_", tcr$barcode),
                                paste0(x, "_", tcr$barcode))
  
  ## Add on sample id to clonotype (can just use sample id, just for distinguishing b/w 3 samples)
  tcr$raw_clonotype_id<- paste0(x, "_", tcr$raw_clonotype_id)
  return(tcr)
})
tcr<- do.call("rbind", tcr_list)


## Colors 
tcell_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Tcell.programs_rank8sharedByunion_reAnnot.Rds"))
myeloid_colors<- readRDS(paste0(full_cohort_dir, "/plot_colors/Myeloid_program_colors.Rds"))
myeloid_colors<- myeloid_colors$rank6

saveRDS(tcr, file=paste0(analysis_dir, "TCR_info.Rds"))
```


## Add VDJ info to seurat objects
```{r}
tcr<- readRDS(paste0(analysis_dir, "TCR_info.Rds"))

## Add clonotype id/vj gene
seurat_obj$raw_clonotype_id<-ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                                    as.character(plyr::mapvalues(colnames(seurat_obj), 
                                                                 tcr$barcode, tcr$raw_clonotype_id, warn_missing = FALSE)),"NA")
seurat_obj$v_gene<- ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                           as.character(plyr::mapvalues(colnames(seurat_obj), 
                                                        tcr$barcode, tcr$v_gene, warn_missing = FALSE)), "NA")
seurat_obj$j_gene<- ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                           as.character(plyr::mapvalues(colnames(seurat_obj), 
                                                        tcr$barcode, tcr$j_gene, warn_missing = FALSE)),"NA")



## Add clonotype size + whether clonally expanded
c_size<- as.data.frame(table(tcr$raw_clonotype_id))

seurat_obj$tcr_reconstructed<- ifelse(colnames(seurat_obj) %in% tcr$barcode, "Yes", "No")

seurat_obj$clonotype_size<- ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                                   as.numeric(as.character(plyr::mapvalues(seurat_obj$raw_clonotype_id, 
                                                                           c_size$Var1, c_size$Freq, warn_missing = FALSE))), "NA")
seurat_obj$clonal<- ifelse(seurat_obj$clonotype_size != "NA", 
                           ifelse(as.numeric(seurat_obj$clonotype_size>1), "clonal",
                                             "nonclonal"),  "NA")
seurat_obj$clonotype_size_max20<- ifelse(seurat_obj$clonotype_size>20 & seurat_obj$clonotype_size != "NA", 20, seurat_obj$clonotype_size)

saveRDS(seurat_obj, file="analysis/preprocessing/seurat_obj_noTumor.Rds")
```

## Plot V/J gene usage
```{r}
seurat_obj<- readRDS("analysis/preprocessing/seurat_obj_noTumor.Rds")


my_plotVJgeneUsage(subset(seurat_obj,clonotype_size != "NA"), "v_gene", 20) + ggtitle("top V gene usage")
ggsave(paste0(figure_dir, "TopV_geneUsage.png"), width=10, height=4)

my_plotVJgeneUsage(subset(seurat_obj,clonotype_size != "NA"), "j_gene", 20)+ ggtitle("top J gene usage")
ggsave(paste0(figure_dir, "TopJ_geneUsage.png"), width=10, height=4)

```

## Plot clonal expansion by sample
```{r}
clonal_meta<- seurat_obj@meta.data[,c("clonal", "tcr_reconstructed", "raw_clonotype_id","clonotype_size",  "sample")]

## subset to cells with tcr reconstructed
clonal_meta<- clonal_meta[clonal_meta$tcr_reconstructed=="Yes",]

## proportion of clonal for each sample
clonal_prop<- as.data.frame(table(clonal_meta$sample, clonal_meta$clonal))
colnames(clonal_prop)<- c("Sample", "Clonal", "NCells")
clonal_prop<- clonal_prop %>% group_by(Sample) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()

ggplot(clonal_prop, aes(x=Sample, y=PerCells, fill=Clonal))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=c(clonal="orange", nonclonal="navy"))+
  theme_classic()+
  theme(axis.text = element_text(face="bold", color="black"),
        strip.text = element_text(face="bold", color="black"),
        axis.text.x = element_text(angle=45, hjust=1))
  xlab("")+ylab("Proportion of clonally expanded T cells")
  ggsave(paste0(figure_dir, "Proportion_clonallyExpanded_bySample.png"), width = 4, height=4)
```


## Plot histogram of clonotype size by sample
```{r}
clonal_meta<- seurat_obj@meta.data[,c("clonal", "tcr_reconstructed", "raw_clonotype_id","clonotype_size",  "sample")]

## subset to cells with tcr reconstructed
clonal_meta<- clonal_meta[clonal_meta$tcr_reconstructed=="Yes",]
clonal_meta$clonotype_size<- as.numeric(clonal_meta$clonotype_size)

ggplot(clonal_meta, aes(x=clonotype_size, fill=sample))+
  geom_histogram()+
  theme_bw()+
  scale_fill_manual(values=sample_colors)
ggsave(paste0(figure_dir, "Hist_clonotypeSize_bySample.png"), width=4, height=3)
```

## Plot dotplot of clonotype size by sample
```{r}
clonal_meta<- seurat_obj@meta.data[,c("clonal", "tcr_reconstructed", "raw_clonotype_id","clonotype_size",  "sample")]

## subset to cells with tcr reconstructed
clonal_meta<- clonal_meta[clonal_meta$tcr_reconstructed=="Yes",]
clonal_meta$clonotype_size<- as.numeric(clonal_meta$clonotype_size)

## Number of cells in each clonotype size by sample
clonal_prop<- as.data.frame(table(clonal_meta$clonotype_size, clonal_meta$sample))
colnames(clonal_prop)<-c("Clonotype_size", "Sample", "NCells")
clonal_prop$Clonotype_size<- as.numeric(clonal_prop$Clonotype_size)


ggplot(clonal_prop, aes(x=Clonotype_size, y=NCells, color=Sample))+
  geom_point()+
  theme_bw()+
  scale_color_manual(values=sample_colors)
ggsave(paste0(figure_dir, "dotPlot_clonotypeSize_bySample.png"), width=6, height=3)
```

## Exploratory- potentially use downsampling approach to control for differences in cell count
More cells sequenced --> artificially inflate clonotype size
Can't just divide by n cells seq- will penalize highly sequenced samples
Instead- try randomly sampling tcrs for each sample so that n=lowest # of Tcrs
```{r}
tcr<- readRDS(paste0(analysis_dir, "TCR_info.Rds"))

## Get minimum tcr to downsample to
n_tcr<- table(tcr$sample)
min_tcr<- min(n_tcr)

## randomly sample without replacement for all samples, then recalculate clonotype size
samples<- c("BT1857", "BT2082", "BT2083")

tcr_downsample<- lapply(samples, function(sample){
  tcr_sample<- tcr[tcr$sample == sample,]
  
  ## Randomly select tcrs
  set.seed(42)
  tcr_barcodes_ds<- sample(tcr_sample$contig_id, size=min_tcr)
  tcr_ds<- tcr_sample[tcr_sample$contig_id %in% tcr_barcodes_ds,]
  
  ## re-calculate clonotype size (note that even the min TCR n sample needs to be included here- clonotype size calculation)
  c_size<- table(tcr_ds$raw_clonotype_id)
  tcr_ds$clonotype_size<- plyr::mapvalues(tcr_ds$raw_clonotype_id, names(c_size), unname(c_size), warn_missing = FALSE)
  
  return(tcr_ds)
})
tcr_ds<-do.call("rbind", tcr_downsample)

## Add downsampled results to seurat object
  tcr<- tcr_ds
  
  ## Add clonotype id/vj gene
  seurat_obj$raw_clonotype_id_ds<-ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                                      as.character(plyr::mapvalues(colnames(seurat_obj), 
                                                                   tcr$barcode, tcr$raw_clonotype_id, warn_missing = FALSE)),"NA")
  seurat_obj$v_gene_ds<- ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                             as.character(plyr::mapvalues(colnames(seurat_obj), 
                                                          tcr$barcode, tcr$v_gene, warn_missing = FALSE)), "NA")
  seurat_obj$j_gene_ds<- ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                             as.character(plyr::mapvalues(colnames(seurat_obj), 
                                                          tcr$barcode, tcr$j_gene, warn_missing = FALSE)),"NA")
  
  
  
  ## Add clonotype size + whether clonally expanded
  c_size<- as.data.frame(table(tcr$raw_clonotype_id))
  
  seurat_obj$tcr_reconstructed_ds<- ifelse(colnames(seurat_obj) %in% tcr$barcode, "Yes", "No")
  
  seurat_obj$clonotype_size_ds<- ifelse(colnames(seurat_obj) %in% tcr$barcode, 
                                     as.numeric(as.character(plyr::mapvalues(seurat_obj$raw_clonotype_id_ds, 
                                                                             c_size$Var1, c_size$Freq, warn_missing = FALSE))), "NA")
  seurat_obj$clonal_ds<- ifelse(seurat_obj$clonotype_size_ds != "NA", 
                             ifelse(as.numeric(seurat_obj$clonotype_size_ds>1), "clonal",
                                               "nonclonal"),  "NA")
  seurat_obj$clonotype_size_max20_ds<- ifelse(seurat_obj$clonotype_size_ds>20 & seurat_obj$clonotype_size_ds != "NA", 20,
                                              seurat_obj$clonotype_size_ds)
  
  
## Plot same as before with downsampled clonotype information
  
  clonal_meta<- seurat_obj@meta.data[,c("clonal_ds", "tcr_reconstructed_ds", "raw_clonotype_id_ds","clonotype_size_ds",  "sample")]
  colnames(clonal_meta)<-gsub("_ds", "", colnames(clonal_meta))

  ## subset to cells with tcr reconstructed
  clonal_meta<- clonal_meta[clonal_meta$tcr_reconstructed=="Yes",]
  
  ## proportion of clonal for each sample
  clonal_prop<- as.data.frame(table(clonal_meta$sample, clonal_meta$clonal))
  colnames(clonal_prop)<- c("Sample", "Clonal", "NCells")
  clonal_prop<- clonal_prop %>% group_by(Sample) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()
  
  ggplot(clonal_prop, aes(x=Sample, y=NCells, fill=Clonal))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=c(clonal="orange", nonclonal="navy"))+
    theme_classic()+
    theme(axis.text = element_text(face="bold", color="black"),
          strip.text = element_text(face="bold", color="black"),
          axis.text.x = element_text(angle=45, hjust=1))
    xlab("")+ylab("Proportion of clonally expanded T cells")
    ggsave(paste0(figure_dir, "Proportion_clonallyExpanded_bySample_Downsampled.png"), width = 4, height=4)
  
  #saveRDS(seurat_obj, file="analysis/preprocessing/seurat_obj_noTumor.Rds")
```

