---
title: "Compare de novo Tcell NMF to full cohort"
author: "Jenna LaBelle"
output: html_document
---

Correlate scores for de novo tcell 10X programs with programs from full cohort

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(grid)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/tenX/Immune/")

analysis_dir<- paste0(working_dir, "/analysis/correlate_deNovo10X_fullCohort/Tcells/")
figure_dir<- paste0(working_dir, "/figures/correlate_deNovo10X_fullCohort/Tcells/")
data_dir <- paste0(working_dir, "analysis/nmf/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)


```

```{r}
## Save pheatmap as pdf- doesn't save correctly with standard method after you add on other layers for coloring axes
save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
   stopifnot(!missing(x))
   stopifnot(!missing(filename))
   pdf(filename, width=width, height=height)
   grid::grid.newpage()
   grid::grid.draw(x$gtable)
   dev.off()
}
```

## Load in tenx seurat object and full cohort programs
```{r}
## Seurat- 10X
tcell_seurat<- readRDS("analysis/manual_broad_annot/tcell_seurat.Rds")

## Seurat- full cohort
full_tcell_seurat<- readRDS(paste0("../../",
                           "02a_Tcells/02_Ped.Adult/03_Visualize.Reannotate_programs/analysis/Reannotate_programs/rank8_sharedByunion/",
                           "ped_seurat_tcell_harmony.Rds"))

## Markers
full_markers<-readRDS("../../02a_Tcells/02_Ped.Adult/01_compare_NMF/analysis/newMarkers_deNovoPed.AdultTcell/new_markers_NMFrank8_bySeurat_sharedByunion_reAnnot.Rds")

## colors
tcell_colors<- readRDS("../../plot_colors/Tcell.programs_rank8sharedByunion_reAnnot.Rds")

```


## Score 10x Tcellfor full cohort programs
```{r}
## norm/center 10x
cm_list<- NormCenter(tcell_seurat@assays$RNA@counts)
cm_mean<- row_means (log2(cm_list$raw_data + 1))

## split full markers to list
full_marker_list<- split(full_markers, f=full_markers$cluster)
full_marker_list<- lapply(full_marker_list, function(x){x$gene})

## score
full_scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, full_marker_list)))
  
saveRDS(full_scores, file=paste0(analysis_dir, "tenx_scores_fullPrograms.Rds"))  
```


## Score SS2 for tcell/myeloid programs from full cohort
Note that I've definitely done this previously.. but easier to just redo and know I have the right marker set. I'm great at this.
```{r}
## Tcell
  ## norm/center 10x
  cm_list_ss2<- NormCenter(full_tcell_seurat@assays$RNA@counts)
  mean_ss2<- rowMeans(log2(cm_list_ss2$raw_data + 1))
  
  ## split full markers to list
  full_marker_list<- split(full_markers, f=full_markers$cluster)
  full_marker_list<- lapply(full_marker_list, function(x){x$gene})
  
  ## score
  full_scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, mean, full_marker_list)))
  
saveRDS(full_scores, file=paste0(analysis_dir, "SS2_scores_fullPrograms.Rds"))  
```

## for each rank, load marker genes, score TENX for de novo 10x nmf, then correlate with full cohort
```{r}
analysis_dir_use<- paste0(analysis_dir, "scoreTenX/")
figure_dir_use<- paste0(figure_dir, "scoreTenX/")
for(i in c(analysis_dir_use, figure_dir_use)){if(!dir.exists(i)){dir.create(i)}}

## Load scores for 10x for full program
full_scores<- readRDS(paste0(analysis_dir, "tenx_scores_fullPrograms.Rds"))

## For each rank- load in markers, score 10x tcells
ranks<- 6:10
denovo_scores<- lapply(ranks, function(r){
  ## Load markers
  markers<- read.csv(paste0(data_dir, "Tcell/top_DE_genes_NMF_rank", r, ".csv"))
  
  ## Split to list
  markers<- split(markers, f=markers$cluster)
  markers<- lapply(markers, function(x){x$gene})
  names(markers)<- paste0("Denovo_NMF", names(markers))
  
  ## score
  scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, markers)))
  return(scores)
  })
names(denovo_scores)<- paste0("Denovo_NMF", ranks)

## For each rank- Merge full scores + de novo scores, correlate
for(r in ranks){
  print(r)
  denovo_scores<- denovo_scores[[paste0("Denovo_NMF", r)]]
  scores<- cbind(denovo_scores, full_scores)
  
  ## NMF factor correlation based on cell score 
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Add on column for color
  scores_factor_cor<- as.data.frame(scores_factor_cor)
  scores_factor_cor$color<-gsub("_.*", "", rownames(scores_factor_cor))
  scores_factor_cor$color<-ifelse(grepl("Denovo", scores_factor_cor$color), "red", "black")
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  
  p<-pheatmap(scores_factor_cor[,colnames(scores_factor_cor) != "color"], color = hm_colors(100), 
         cluster_rows = F, cluster_cols = F, 
         annotation_names_row = F, annotation_names_col =T,
         show_rownames = T, show_colnames = F,
         #filename = paste0(figure_dir , "CorHeatmap.png"),
         width = 10, height = 6)
  
  cols=scores_factor_cor[order(match(rownames(scores_factor_cor), p$gtable$grobs[[2]]$label)), ]$color
  p$gtable$grobs[[2]]$gp=gpar(col=cols)
  
  png(filename=paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort.png"), width = 700, height = 400)
  p
  dev.off()
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir_use, "HC_DenovoRank", r, "_fullCohort.png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
  
}
```

## for each rank, load marker genes, score SS2 for de novo 10x nmf, then correlate with full cohort
```{r}
analysis_dir_use<- paste0(analysis_dir, "scoreSS2/")
figure_dir_use<- paste0(figure_dir, "scoreSS2/")
for(i in c(analysis_dir_use, figure_dir_use)){if(!dir.exists(i)){dir.create(i)}}

## Load scores for 10x for full program
full_scores<- readRDS(paste0(analysis_dir_use, "ss2_scores_fullPrograms.Rds"))

## For each rank- load in de novo 10x markers, score ss2 tcells
ranks<- 6:10
denovo_scores_list<- lapply(ranks, function(r){
  ## Load markers
  markers<- read.csv(paste0(data_dir, "Tcell/top_DE_genes_NMF_rank", r, ".csv"))
  
  ## Split to list
  markers<- split(markers, f=markers$cluster)
  markers<- lapply(markers, function(x){x$gene})
  names(markers)<- paste0("Denovo_NMF", names(markers))
  
  ## score
  scores<- as.data.frame(t(scoreNmfGenes(cm_list_ss2$center_data, mean_ss2, markers)))
  return(scores)
  })
names(denovo_scores_list)<- paste0("Denovo_NMF", ranks)

## For each rank- Merge full scores + de novo scores, correlate
for(r in ranks){
  print(r)
  denovo_scores<- denovo_scores_list[[paste0("Denovo_NMF", r)]]
  scores<- cbind(denovo_scores, full_scores)
  
  ## NMF factor correlation based on cell score 
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Add on column for color
  scores_factor_cor<- as.data.frame(scores_factor_cor)
  scores_factor_cor$color<-gsub("_.*", "", rownames(scores_factor_cor))
  scores_factor_cor$color<-ifelse(grepl("Denovo", scores_factor_cor$color), "red", "black")
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  
  p<-pheatmap(scores_factor_cor[,colnames(scores_factor_cor) != "color"], color = hm_colors(100), 
         cluster_rows = F, cluster_cols = F, 
         annotation_names_row = F, annotation_names_col =T,
         show_rownames = T, show_colnames = F,
         #filename = paste0(figure_dir , "CorHeatmap.png"),
         width = 10, height = 6)
  
  cols=scores_factor_cor[order(match(rownames(scores_factor_cor), p$gtable$grobs[[2]]$label)), ]$color
  p$gtable$grobs[[2]]$gp=gpar(col=cols)
  
  save_pheatmap_pdf(p, paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort.pdf"), width=10, height=6)
  
  #png(filename=paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort.png"), width = 700, height = 400)
  #p
  #dev.off()
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir_use, "HC_DenovoRank", r, "_fullCohort.png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
  
}
```

## For ranks selected for follow up: annotate based on correlation with full cohort
```{r}
ranks<- 9:10

## Assign annotations
rank_annot<- list(NMF9=c("Stress", "Cycling", "Predysfunctional", "IFN.Response", "Cytotoxic.NKreceptor_Alt",
                                 "Treg", "Memory", "Activated_EM", "Cytotoxic.NKreceptor"),
                  NMF10=c("Cycling", "Stress", "Effector.Memory", "IFN.Response", "Cytotoxic.NKreceptor_Alt",
                                  "Memory", "Treg", "Cytotoxic.NKreceptor", "Predysfunctional", "Activated_EM"))
rank_annot<- lapply(rank_annot, function(x){names(x)<- paste0("NMF", 1:length(x)); return(x)})

## Annotate: markers, seurat cell assignments
for(r in ranks){
  annots<- rank_annot[[paste0("NMF", r)]]
  
  ## annotate markers
  markers<- read.csv(paste0(data_dir, "Tcell/top_DE_genes_NMF_rank", r, ".csv"))
  markers$cluster<- as.character(plyr::mapvalues(paste0("NMF", markers$cluster),
                                                  names(annots), annots))
  write.csv(markers, file=paste0(data_dir, "Tcell/top_DE_genes_NMF_rank", r, "_annotated.csv"))
  
  ## annotate seurat assignments
  nmf_column<- paste0("NMF_rank", r)
  tcell_seurat[[paste0(nmf_column, "_annot")]]<-as.character(plyr::mapvalues(paste0("NMF", as.character(tcell_seurat[[nmf_column]][,1])),
                                                  names(annots), annots))
}

saveRDS(tcell_seurat, file="analysis/manual_broad_annot/tcell_seurat.Rds")
```

## Basic plots- UMAP, barchart by sample, expression of expected GOI
```{r}
tcell_seurat<- readRDS("analysis/manual_broad_annot/tcell_seurat.Rds")
ranks<- 9:10
tcell_colors<- c(tcell_colors,"Cytotoxic.NKreceptor_Alt"="green", Activated_EM="red")

## Umap by program
all_plots<- lapply(ranks, function(r){
  tcell_seurat$tmp<- tcell_seurat@meta.data[[paste0("NMF_rank", r, "_annot")]]
  DimPlot(tcell_seurat, group.by = "tmp")+
    scale_color_manual(values=tcell_colors[names(tcell_colors) %in% tcell_seurat$tmp])+
    ggtitle(paste0("rank ", r))
})
cowplot::plot_grid(plotlist = all_plots, ncol=2)
ggsave(paste0(figure_dir, "UMAP_byAnnotProgram_ranks", paste(ranks, collapse = "."), ".png"), width=12, height=5)

## Proportion of program by sample
all_plots<- lapply(ranks, function(r){
  tcell_seurat$tmp<- tcell_seurat@meta.data[[paste0("NMF_rank", r, "_annot")]]
  my_barchart_programProp(tcell_seurat, coloring_variable = "tmp", colors=tcell_colors)+
    ggtitle(paste0("rank ", r))
})
cowplot::plot_grid(plotlist = all_plots, ncol=2)
ggsave(paste0(figure_dir, "Barchart_proportionProgram_bySample_ranks", paste(ranks, collapse = "."), ".png"), width=8, height=4)

## Expression of expected markers by sample
top_markers<- full_markers %>% group_by(cluster) %>% top_n(n=5, wt=avg_log2FC) %>% as.data.frame()
top_markers$cluster<- gsub("Inhibitory", "Activated_EM", top_markers$cluster)
top_markers$cluster<- factor(top_markers$cluster, levels=sort(unique(top_markers$cluster)))
top_markers<- top_markers[order(top_markers$cluster),]

all_plots<- lapply(ranks, function(r){
    tcell_seurat$tmp<- tcell_seurat@meta.data[[paste0("NMF_rank", r, "_annot")]]
    tcell_seurat$tmp<- factor(tcell_seurat$tmp, levels=sort(unique(tcell_seurat$tmp)))
    DotPlot(tcell_seurat, group.by = "tmp", features=top_markers$gene)+
      theme(axis.text.x = element_text(angle=45, hjust=1))+
      scale_color_gradient2(low="blue", mid="white", high="red")+
      ggtitle(paste0("rank", r))+
      xlab("")+ylab("")
})
cowplot::plot_grid(plotlist = all_plots, ncol=1)
ggsave(paste0(figure_dir, "DotPlot_expectedMarkers", paste(ranks, collapse = "."),".png"), width=14, height=8)
```

