---
title: "Compare de novo myeloid NMF to full cohort"
author: "Jenna LaBelle"
output: html_document
---

Correlate scores for de novo myeloid 10X programs with programs from full cohort

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(grid)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/tenX/Immune/")

analysis_dir<- paste0(working_dir, "/analysis/correlate_deNovo10X_fullCohort/Myeloid/")
figure_dir<- paste0(working_dir, "/figures/correlate_deNovo10X_fullCohort/Myeloid/")
data_dir <- paste0(working_dir, "analysis/nmf/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)


```

## Helper functions
```{r}
## correlate scores and plot heatmap
## Input: dataframe of scores, rank used for naming
my_correlationHeatmap<- function(scores,r, figure_dir=figure_dir_use){
  ## NMF factor correlation based on cell score 
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Add on column for color
  scores_factor_cor<- as.data.frame(scores_factor_cor)
  scores_factor_cor$color<-gsub("_.*", "", rownames(scores_factor_cor))
  scores_factor_cor$color<-ifelse(grepl("Denovo", scores_factor_cor$color), "red",
                                  ifelse(grepl("Adult10X",scores_factor_cor$color),"green", "black"))
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  
  p<-pheatmap(scores_factor_cor[,colnames(scores_factor_cor) != "color"], color = hm_colors(100), 
         cluster_rows = F, cluster_cols = F, 
         annotation_names_row = F, annotation_names_col =T,
         show_rownames = T, show_colnames = F,
         #filename = paste0(figure_dir , "CorHeatmap.png"),
         width = 10, height = 6)
  
  cols=scores_factor_cor[order(match(rownames(scores_factor_cor), p$gtable$grobs[[2]]$label)), ]$color
  p$gtable$grobs[[2]]$gp=gpar(col=cols)
  
  
  return(list(p=p, factor_cor=scores_factor_hc))
}

## Save pheatmap as pdf- doesn't save correctly with standard method after you add on other layers for coloring axes
save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
   stopifnot(!missing(x))
   stopifnot(!missing(filename))
   pdf(filename, width=width, height=height)
   grid::grid.newpage()
   grid::grid.draw(x$gtable)
   dev.off()
}
```

## Load in tenx seurat object and full cohort programs
```{r}
## Seurat- 10X
myeloid_seurat<- readRDS("analysis/manual_broad_annot/myeloid_seurat.Rds")


## Seurat- full cohort
full_myeloid_seurat<- readRDS(paste0("../../",
                               "01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/pedOnly_nomuv63/",
                              "myeloid_seurat_noDC.B_harmony.Rds"))


## Markers
full_markers<- read.csv("../../02b_Myeloid/01a_NMF/analysis/de.novoNMF/pedOnly_nomuv63/top_DE_genes_10Xannotated_NMF_rank6.csv")

## adult 10x markers and split
adult_markers_top100<- readRDS(paste0("../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/Preprocess_AdultMyeloid/", 
                                      "GBM.10X_recurrent_myeloid_markers_top100.Rds"))


## colors
myeloid_colors<- readRDS("../../plot_colors/Myeloid_program_colors.Rds")
myeloid_colors<- myeloid_colors$rank6
```


## Score 10x myeloid for: 
full cohort programs
de novo programs
```{r}
ranks<- c(6,7,9)

## norm/center 10x
cm_list<- NormCenter(myeloid_seurat@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data + 1))

## Full cohort markers  
  ## split full markers to list
  full_marker_list<- split(full_markers, f=full_markers$cluster)
  full_marker_list<- lapply(full_marker_list, function(x){x$gene})
  
  ## score
  full_scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, full_marker_list)))

## de novo programs  
  ## For each rank- load in markers, score 10x Myeloid
  denovo_scores_list<- lapply(ranks, function(r){
    ## Load markers
    markers<- read.csv(paste0(data_dir, "Myeloid/top_DE_genes_NMF_rank", r, ".csv"))
    
    ## Split to list
    markers<- split(markers, f=markers$cluster)
    markers<- lapply(markers, function(x){x$gene})
    names(markers)<- paste0("Denovo_NMF", names(markers))
    
    ## score
    scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, markers)))
    return(scores)
    })
  names(denovo_scores_list)<- paste0("Denovo_NMF", ranks)
  
saveRDS(full_scores, file=paste0(analysis_dir, "tenx_scores_fullPrograms.Rds"))  
saveRDS(denovo_scores_list, file=paste0(analysis_dir, "tenx_scores_denovoPrograms_rank",paste(ranks, collapse = "."), ".Rds"))  
```


## Score SS2 for myeloid programs from full cohort + 10x adult GBM programs + de novo programs
Note that I've definitely done this previously.. but easier to just redo and know I have the right marker set. I'm great at this.
```{r}
ranks<- c(6,7,9)

## norm/center 10x
cm_list_ss2<- NormCenter(full_myeloid_seurat@assays$RNA@counts)
mean_ss2<- rowMeans (log2(cm_list_ss2$raw_data + 1))

## myeloid programs from full cohort
  ## split full markers to list
  full_marker_list<- split(full_markers, f=full_markers$cluster)
  full_marker_list<- lapply(full_marker_list, function(x){x$gene})
  
  ## score
  full_scores<- as.data.frame(t(scoreNmfGenes(cm_list_ss2$center_data, mean_ss2, full_marker_list)))
  
## myeloid programs from adult 10X gbm
  ## split full markers to list
  adult_markers<- split(adult_markers_top100, f=adult_markers_top100$cluster)
  adult_markers<- lapply(adult_markers, function(x){x$gene})
  
  ## score
  full_scores_adult10X<- as.data.frame(t(scoreNmfGenes(cm_list_ss2$center_data, mean_ss2, adult_markers)))

## de novo 10x programs
denovo_scores_list<- lapply(ranks, function(r){
  ## Load markers
  markers<- read.csv(paste0(data_dir, "Myeloid/top_DE_genes_NMF_rank", r, ".csv"))
  
  ## Split to list
  markers<- split(markers, f=markers$cluster)
  markers<- lapply(markers, function(x){x$gene})
  names(markers)<- paste0("Denovo_NMF", names(markers))
  
  ## score
  scores<- as.data.frame(t(scoreNmfGenes(cm_list_ss2$center_data, mean_ss2, markers)))
  return(scores)
  })
names(denovo_scores_list)<- paste0("Denovo_NMF", ranks)

saveRDS(full_scores, file=paste0(analysis_dir, "SS2_scores_fullPrograms.Rds"))  
saveRDS(full_scores_adult10X, file=paste0(analysis_dir, "SS2_scores_adult10XGBM.Rds"))  
saveRDS(denovo_scores_list, file=paste0(analysis_dir, "SS2_scores_denovoPrograms_rank",paste(ranks, collapse = "."), ".Rds"))  
```

## for each rank, load marker genes, score TENX for de novo 10x nmf, then correlate with full cohort
```{r}
ranks<- c(10,12,13)

analysis_dir_use<- paste0(analysis_dir, "scoreTenX/")
figure_dir_use<- paste0(figure_dir, "scoreTenX/")
for(i in c(analysis_dir_use, figure_dir_use)){if(!dir.exists(i)){dir.create(i)}}

## Load scores for 10x for full program
full_scores<- readRDS(paste0(analysis_dir, "tenx_scores_fullPrograms.Rds"))
denovo_scores_list<- readRDS(paste0(analysis_dir, "tenx_scores_denovoPrograms_rank",paste(ranks, collapse = "."), ".Rds"))

## For each rank- Merge full scores + de novo scores, correlate
for(r in ranks){
  print(r)
  denovo_scores<- denovo_scores_list[[paste0("Denovo_NMF", r)]]
  scores<- cbind(denovo_scores, full_scores)
  
  p<-my_correlationHeatmap(scores, r)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir, "HC_DenovoRank", r, "_fullCohort.png"), width = 400, height = 600)
  plot(p$factor_cor$hc_obj)
  dev.off()
  
  ## heatmap
  png(filename=paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort.png"), width = 700, height = 400)
  p$p
  dev.off()
  
}
```

## for each rank, load marker genes, score SS2 for de novo 10x nmf, then correlate with full cohort
```{r}
ranks<- c(10,12,13)

analysis_dir_use<- paste0(analysis_dir, "scoreSS2/")
figure_dir_use<- paste0(figure_dir, "scoreSS2/")
for(i in c(analysis_dir_use, figure_dir_use)){if(!dir.exists(i)){dir.create(i)}}

## Load scores for ss2 for full program
full_scores<- readRDS(paste0(analysis_dir, "ss2_scores_fullPrograms.Rds"))
denovo_scores_list<- readRDS(paste0(analysis_dir, "SS2_scores_denovoPrograms_rank",paste(ranks, collapse = "."), ".Rds"))


## For each rank- Merge full scores + de novo scores, correlate
for(r in ranks){
  print(r)
  denovo_scores<- denovo_scores_list[[paste0("Denovo_NMF", r)]]
  scores<- cbind(denovo_scores, full_scores)
  
  p<-my_correlationHeatmap(scores, r)
  cp<- p$p
  
  save_pheatmap_pdf(cp, paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort.pdf"), width=10, height=6)
  
  #pdf(paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort.pdf"), width = 700, height = 400)
  #grid::grid.newpage()
  #grid::grid.draw(cp$gtable)
  #dev.off()
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir_use, "HC_DenovoRank", r, "_fullCohort.png"), width = 400, height = 600)
  plot(p$factor_cor$hc_obj)
  dev.off()
  
}

## Same, but remove AC program (FOR RANK 13)

## For each rank- Merge full scores + de novo scores, correlate
for(r in ranks){
  print(r)
  denovo_scores<- denovo_scores_list[[paste0("Denovo_NMF", r)]]
  denovo_scores<- denovo_scores[,colnames(denovo_scores) != "Denovo_NMF10"]
  scores<- cbind(denovo_scores, full_scores)
  
  p<-my_correlationHeatmap(scores, r)
  cp<- p$p
  
  save_pheatmap_pdf(cp, paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort_noAC.pdf"), width=10, height=6)
  
  #pdf(paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort.pdf"), width = 700, height = 400)
  #grid::grid.newpage()
  #grid::grid.draw(cp$gtable)
  #dev.off()
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir_use, "HC_DenovoRank", r, "_fullCohort.png"), width = 400, height = 600)
  plot(p$factor_cor$hc_obj)
  dev.off()
  
}
```

## Correlate with adult 10X gbm- ss2 scored for new 10x programs + adult 10x programs
```{r}
ranks<- c(10,12,13)

## set analysis/figure dir
analysis_dir_use<- paste0(analysis_dir, "scoreSS2/")
figure_dir_use<- paste0(figure_dir, "scoreSS2/")
for(i in c(analysis_dir_use, figure_dir_use)){if(!dir.exists(i)){dir.create(i)}}

## Load scores for ss2 for adult 10x programs
full_scores<- readRDS(paste0(analysis_dir, "SS2_scores_adult10XGBM.Rds"))
denovo_scores_list<- readRDS(paste0(analysis_dir, "SS2_scores_denovoPrograms_rank",paste(ranks, collapse = "."), ".Rds"))

## For each rank- Merge full scores + de novo scores, correlate
for(r in ranks){
  print(r)
  denovo_scores<- denovo_scores_list[[paste0("Denovo_NMF", r)]]
  scores<- cbind(denovo_scores, full_scores)
  
  p<-my_correlationHeatmap(scores, r)
  
  png(filename=paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort_adult10XGBM.png"), width = 700, height = 400)
  p$p
  dev.off()
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir_use, "HC_DenovoRank", r, "_fullCohort_adult10XGBM.png"), width = 400, height = 600)
  plot(p$factor_cor$hc_obj)
  dev.off()
  
}
```

## Correlate with adult 10X gbm + full cohort programs- ss2 scored for new 10x programs + adult 10x programs+ full cohort programs
```{r}
ranks<- c(10,12,13)

## set analysis/figure dir
analysis_dir_use<- paste0(analysis_dir, "scoreSS2/")
figure_dir_use<- paste0(figure_dir, "scoreSS2/")
for(i in c(analysis_dir_use, figure_dir_use)){if(!dir.exists(i)){dir.create(i)}}

## Load scores 
full_scores<- readRDS(paste0(analysis_dir, "ss2_scores_fullPrograms.Rds"))
denovo_scores_list<- readRDS(paste0(analysis_dir, "SS2_scores_denovoPrograms_rank",paste(ranks, collapse = "."), ".Rds"))
tenxAdultGBM_scores<- readRDS(paste0(analysis_dir, "SS2_scores_adult10XGBM.Rds"))

## rename columns for coloring
colnames(tenxAdultGBM_scores)<- paste0("Adult10X_", colnames(tenxAdultGBM_scores))

## For each rank- Merge full scores + de novo scores, correlate
for(r in ranks){
  print(r)
  denovo_scores<- denovo_scores_list[[paste0("Denovo_NMF", r)]]
  scores<- cbind(denovo_scores, full_scores)
  scores<- cbind(scores, tenxAdultGBM_scores)
  
  p<-my_correlationHeatmap(scores, r)
  
  png(filename=paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort_adult10XGBM.fullCohortPrograms.png"), width = 700, height = 400)
  p$p
  dev.off()
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir_use, "HC_DenovoRank", r, "_fullCohort_adult10XGBM.fullCohortPrograms.png"), width = 400, height = 600)
  plot(p$factor_cor$hc_obj)
  dev.off()
  
}

## For rank 13- remove AC, Merge full scores + de novo scores, correlate
  r<- 13
  denovo_scores<- denovo_scores_list[[paste0("Denovo_NMF13")]]
  denovo_scores<- denovo_scores[,!grepl("NMF10", colnames(denovo_scores))]
  scores<- cbind(denovo_scores, full_scores)
  scores<- cbind(scores, tenxAdultGBM_scores)
  
  p<-my_correlationHeatmap(scores, r)
  
  png(filename=paste0(figure_dir_use, "Cor_DenovoRank", r, "_fullCohort_adult10XGBM.fullCohortPrograms_noAC.png"), width = 700, height = 400)
  p$p
  dev.off()
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir_use, "HC_DenovoRank", r, "_fullCohort_adult10XGBM.fullCohortPrograms_noAC.png"), width = 400, height = 600)
  plot(p$factor_cor$hc_obj)
  dev.off()

```




## For ranks selected for follow up: annotate based on correlation with full cohort
```{r}
ranks_use<-c(9,13)
ranks_all<-c(6,7,9)

## Assign annotations
rank_annot<- list(NMF9=c("Inflammatory_Microglia", "Hypoxic_TAM1", "Hypoxic_TAM2", "IFN_TAM", "Monocytes1", 
                         "IFN_Mo_TAM", "CD3pos_Myeloid","Monocytes2", "SEPP1_Mo_TAM"),
                  NMF13=c("Inflammatory1", "CD3pos_Myeloid", "IFN_Mo_TAM", "Microglia", "SEPP1_Mo_TAM",
                          "Hypoxic_TAM1", "Monocytes1", "Inflammatory2", "IFN_TAM1", "AC", "IFN_TAM2",
                          "Hypoxic_TAM2", "Monocytes2"))
rank_annot<- lapply(rank_annot, function(x){names(x)<- paste0("NMF", 1:length(x)); return(x)})

## Annotate: markers, seurat cell assignments, scores
for(r in ranks_use){
  annots<- rank_annot[[paste0("NMF", r)]]
  
  ## annotate markers
  markers<- read.csv(paste0(data_dir, "Myeloid/top_DE_genes_NMF_rank", r, ".csv"))
  markers$cluster<- as.character(plyr::mapvalues(paste0("NMF", markers$cluster),
                                                  names(annots), annots))
  write.csv(markers, file=paste0(data_dir, "Myeloid/top_DE_genes_NMF_rank", r, "_annotated.csv"))
  
  ## annotate seurat assignments
  nmf_column<- paste0("NMF_rank", r)
  myeloid_seurat[[paste0(nmf_column, "_annot")]]<-as.character(plyr::mapvalues(paste0("NMF", as.character(myeloid_seurat[[nmf_column]][,1])),
                                                  names(annots), annots))

}

## annotate scores
  ## 10x
  denovo_scores_list<- readRDS(paste0(analysis_dir, "tenx_scores_denovoPrograms_rank",paste(ranks_all, collapse = "."), ".Rds"))
  denovo_scores_list<- denovo_scores_list[names(denovo_scores_list) %in% paste0("Denovo_NMF", ranks_use)]
  denovo_scores_list<- lapply(names(denovo_scores_list), function(x) {
    annots<- rank_annot[[gsub("Denovo_", "", x)]]
    scores<- denovo_scores_list[[x]]
    colnames(scores)<- gsub("Denovo_", "", colnames(scores))
    colnames(scores)<- plyr::mapvalues(colnames(scores), names(annots), annots)
    return(scores)
  })
  names(denovo_scores_list)<- paste0("Denovo_NMF", ranks_use)
  saveRDS(denovo_scores_list, file=paste0(analysis_dir, "tenx_scores_denovoPrograms_rank",paste(ranks_use, collapse = "."), "_annot.Rds"))
  
  ## SS2 full cohort
  denovo_scores_list<- readRDS(paste0(analysis_dir, "SS2_scores_denovoPrograms_rank",paste(ranks_all, collapse = "."), ".Rds"))
  denovo_scores_list<- denovo_scores_list[names(denovo_scores_list) %in% paste0("Denovo_NMF", ranks_use)]
  denovo_scores_list<- lapply(names(denovo_scores_list), function(x) {
    annots<- rank_annot[[gsub("Denovo_", "", x)]]
    scores<- denovo_scores_list[[x]]
    colnames(scores)<- gsub("Denovo_", "", colnames(scores))
    colnames(scores)<- plyr::mapvalues(colnames(scores), names(annots), annots)
    return(scores)
  })
  names(denovo_scores_list)<- paste0("Denovo_NMF", ranks_use)
  saveRDS(denovo_scores_list, file=paste0(analysis_dir, "SS2_scores_denovoPrograms_rank",paste(ranks_use, collapse = "."), "_annot.Rds"))

## Add annotation column with same programs (e.g., inflammatory 1 and 2) merged
for(r in ranks_use){
  m<-paste0("NMF_rank", r, "_annot") 
  m_simple<- paste0("NMF_rank", r, "_annot_simplified")
  myeloid_seurat@meta.data[[m_simple]]<-ifelse(myeloid_seurat@meta.data[[m]] != "SEPP1_Mo_TAM", 
                                               gsub("[1-2]", "", myeloid_seurat@meta.data[[m]]),
                                               myeloid_seurat@meta.data[[m]])
}  
  
saveRDS(myeloid_seurat, file="analysis/manual_broad_annot/myeloid_seurat.Rds")

```



## Using NMF rank 13, identified small AC subset. Remove these and rerun clustering
```{r}
myeloid_seurat<- readRDS("analysis/manual_broad_annot/myeloid_seurat.Rds")

## Remove AC
myeloid_seurat<- subset(myeloid_seurat, NMF_rank13_annot != "AC")

## Rerun clustering
myeloid_meta<- myeloid_seurat@meta.data
myeloid_seurat<- RunFullSeurat(myeloid_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=myeloid_meta$sample)
myeloid_seurat<- AddMetaData(myeloid_seurat, myeloid_meta[,!colnames(myeloid_meta) %in% colnames(myeloid_seurat@meta.data)])

saveRDS(myeloid_seurat, file="analysis/manual_broad_annot/myeloid_seurat_removeAC.byNMF.Rds")
```


## RANK 13 ONLY
## Supplemental plots to support annotation of Mo TAM (NMF3) 
```{r}
myeloid_seurat<- readRDS("analysis/manual_broad_annot/myeloid_seurat.Rds")
ranks<- c(10,12,13)
denovo_scores_list<- readRDS(paste0(analysis_dir, "SS2_scores_denovoPrograms_rank",paste(ranks, collapse = "."), ".Rds"))
tenxAdultGBM_scores<- readRDS(paste0(analysis_dir, "SS2_scores_adult10XGBM.Rds"))

## Mo TAM correlation
denovo_scores<- denovo_scores_list$Denovo_NMF13

  ## Correlate nmf 3 scores with all tenx adult gbm scores
  cors<- as.data.frame(apply(tenxAdultGBM_scores, 2, function(x){cor(x, denovo_scores$Denovo_NMF3)}))
  colnames(cors)<- "CorrelationWithMoTAM"
  cors$DenovoProgram<- rownames(cors)
  cors$Y<-"Y"
  
  ## plot
  ggplot(cors, aes(y=DenovoProgram, x="Y", color=CorrelationWithMoTAM, size=CorrelationWithMoTAM))+
    geom_point()+
    scale_colour_gradientn(colors=rev(rainbow(4)))+
    theme_classic()+
    xlab("") +ylab("")+
    theme(axis.text = element_text(face="bold", color="black"))
  ggsave(paste0(figure_dir, "NMFrank13_corr10XAdult.NMF3_toShowMoTAM.pdf"), device="pdf", width=4, height=5)
  
```