---
title: "AnnotateByNMF"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Using shared/specific Tcell programs, annotate each program separately as CD4/CD8

## Approach:
Split seurat object by program, recluster. Annotate cluster-wise when possible, perform re-clustering and cell-wise annotation when needed. Also compare to "actual" annotations (by expression) to assess accuracy

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(harmony)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/tenX/Immune/")

analysis_dir_tmp<- paste0(working_dir, "/analysis/identifyCD4.CD8_byNMF/")
figure_dir_tmp<- paste0(working_dir, "/figures/identifyCD4.CD8_byNMF/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

immune_colors<- readRDS("../../plot_colors/ImmuneCell.celltypes.Rds")
```

## Read in ped tcell seurat object. Contains NMF assignments (annots added in deNovoPedTcell.v.deNovoAdultTcell, based on correlation with adult T cell + correlation with ped cd4/cd8 programs)
```{r}
## ped/adult tcell seurat
tcell_seurat<- readRDS("analysis/manual_broad_annot/tcell_seurat.Rds")

## figure/analysis dir based on rank + how shared programs are merged
rank_use<-9
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use,  "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
```


## For initial estimate of CD4/CD8: classify each cell cell-wise, plot proportions by program
```{r}
variable_name<- paste0("NMF_rank", rank_use, "_annot")

## Make df with CD4/CD8 expression
cd4.cd8_expr<- as.data.frame(t(tcell_seurat@assays$RNA@data[c("CD8A", "CD8B", "CD4"),]))
cd4.cd8_expr$CD8<- apply(cd4.cd8_expr, 1, function(x){max(x["CD8A"], x["CD8B"])})

## Remove cells that are 0 for both
cd4.cd8_expr<- cd4.cd8_expr[rowSums(cd4.cd8_expr)>0,] ## 2,154 (52%)

## Assign each cell based on whether CD4 or CD8 expression is higher
cd4.cd8_expr$annotate<- ifelse(cd4.cd8_expr$CD4>cd4.cd8_expr$CD8, "CD4",
                               ifelse(cd4.cd8_expr$CD8 > cd4.cd8_expr$CD4, "CD8", "equal"))

## Remove cells with equal cd4/cd8 expression (can't be annotated)
cd4.cd8_expr<- cd4.cd8_expr[cd4.cd8_expr$annotate!="equal",] ## 64 cells removed

## Add on program info
cd4.cd8_expr$program<- plyr::mapvalues(rownames(cd4.cd8_expr), colnames(tcell_seurat),
                                       tcell_seurat@meta.data[[variable_name]],
                                       warn_missing = FALSE)

## Get proportion of cell types by program
df<- as.data.frame(table(cd4.cd8_expr$program, cd4.cd8_expr$annotate))
colnames(df)<-c("Program", "CellType", "Ncells")
df<-df %>% group_by(Program) %>% mutate(perCells=Ncells/sum(Ncells)) %>% as.data.frame()

## Order
df$CellType<- factor(df$CellType, levels=c("CD4", "CD8"))

## Plot
ggplot(df, aes(x=Program, y=perCells, fill=CellType))+
  geom_bar(stat="identity", position="stack")+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("") + ylab("Proportion of cells")+
  geom_hline(yintercept = 0.5, linetype="dashed")
ggsave(paste0(figure_dir, "CD4.CD8.predictions_byCellwiseExpression.png"), width=5, height=3)

saveRDS(cd4.cd8_expr, file=paste0(analysis_dir, "cd4.cd8_expression_values.Rds"))
```

## Split by NMF, re-run clustering for each
```{r}
rank_use<-9
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use,  "/")
variable_name<- paste0("NMF_rank", rank_use, "_annot")

tcell_seurat$tmp<- tcell_seurat@meta.data[[variable_name]]
all_seurat<- lapply(unique(tcell_seurat$tmp), function(x){
  print(x)
  seurat_tmp<- subset(tcell_seurat, tmp==x)
  meta_tmp<- seurat_tmp@meta.data
  seurat_tmp<- RunFullSeurat(cm=seurat_tmp@assays$RNA@counts, RunHarmony = TRUE, samples=meta_tmp$sample)
  seurat_tmp<- AddMetaData(seurat_tmp, meta_tmp[,!colnames(meta_tmp) %in% colnames(seurat_tmp@meta.data)])
  return(seurat_tmp)
})
names(all_seurat)<- unique(tcell_seurat$tmp)



## Plot CD4/CD8 expression for each
all_plots<-lapply(names(all_seurat), function(x){
  seurat_tmp<- all_seurat[[x]]
  p1<-FeaturePlot(seurat_tmp, features=c("CD4", "CD8A"))
  p2<- DimPlot(seurat_tmp, label=TRUE)+NoLegend()+ggtitle(x)
  p3<- VlnPlot(seurat_tmp, features=c("CD4", "CD8A"))
  cowplot::plot_grid(plotlist=list(p2, p1, p3), ncol=3)
})
names(all_plots)<-names(all_seurat)
p<-cowplot::plot_grid(plotlist=all_plots, ncol=1) 

ggsave(plot=p, paste0(figure_dir, "CD4.CD8_expr_byNMFProgram.png"), width=15, height=25)

saveRDS(all_seurat, file=paste0(analysis_dir, "seurat_list_splitByNMF.Rds"))
```



## Annotate cells for each program cluster-wise
Three main types of programs:
1. For some programs (NK, Tregs), all cells are clearly either CD4/CD8
2. For some programs, the clustering above clearly splits cells into CD4/CD8
3. For some programs, while there is clearly CD4/CD8 patterning, need to set the resolution manually to split cluster-wise
4. For some programs, need to do some sub-clustering

## Read in seurat object list (above)- 1 seurat object for each program
```{r}
rank_use<- 9
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use,  "/")


seurat_list<- readRDS(paste0(analysis_dir, "seurat_list_splitByNMF.Rds"))

```

## Option 1: all cells either CD4/CD8
```{r}
programs<- c("Treg", "Cytotoxic.NKreceptor", "Cytotoxic.NKreceptor_Alt", "Predysfunctional")
names(programs)<-c("CD4", "CD8", "CD8", "CD8")
for(i in 1:length(programs)){
  program<- programs[i]
  print(program)
  tmp<- seurat_list[[program]]
  tmp$detailed_annot_byProgram<- names(program)
  seurat_list[[program]]<- tmp
}

```

## Option 2: program contains CD4/CD8, current resolution splits cleanly
```{r}
## Programs that can be annotated at current resolution- annotations based on figure "CD4.CD8_expr_byNMFProgram.png"
programs<-list(Memory=list(CD4=c(0:2), CD8=c(3)), Cycling=list(CD4=c(0), CD8=c(1:2)))

all_plots<-list()
for(program in names(programs)){
  print(program)
  tmp<- seurat_list[[program]]
  program_annot<- programs[[program]]
  tmp$detailed_annot_byProgram<- ifelse(tmp$seurat_clusters %in% program_annot$CD4, "CD4", "CD8")
  p<- DimPlot(tmp, group.by = "detailed_annot_byProgram") + VlnPlot(tmp,
                                                                    group.by = "detailed_annot_byProgram", 
                                                                    features=c("CD4", "CD8A") )
  all_plots[[program]]<-p
  seurat_list[[program]]<- tmp
}
cowplot::plot_grid(plotlist=all_plots)
```

## Option 3: program needs to be manually re-clustered
```{r}
## IFN Response
seurat_tmp<- seurat_list[["IFN.Response"]]
seurat_tmp<- FindClusters(seurat_tmp, resolution = 2)
cd4_clusters<- c(0:2, 4:5, 7,9)
cd8_clusters<-c(3, 6,8)
seurat_tmp$detailed_annot_byProgram<- ifelse(seurat_tmp$seurat_clusters %in% cd4_clusters, "CD4",
                                             ifelse(seurat_tmp$seurat_clusters %in% cd8_clusters, "CD8",
                                                    "issue"))
DimPlot(seurat_tmp, group.by = "detailed_annot_byProgram") + VlnPlot(seurat_tmp,
                                                                  group.by = "detailed_annot_byProgram", 
                                                                  features=c("CD4", "CD8A") )
seurat_list$IFN.Response<- seurat_tmp

## Stress
seurat_tmp<- seurat_list$Stress

  ## Remove hsp genes- masking cd4/cd8 differences
  hsp_genes<- rownames(seurat_tmp)[grepl("^HSP", rownames(seurat_tmp))]
  seurat_no_hsp<- RunFullSeurat(seurat_tmp@assays$RNA@counts[!rownames(seurat_tmp) %in% hsp_genes,],
                             RunHarmony = TRUE, samples = seurat_tmp$sample)
  seurat_no_hsp<- FindClusters(seurat_no_hsp, resolution =.7)
  p1<- FeaturePlot(seurat_no_hsp, features=c("CD4", "CD8A")) 
  p2<- VlnPlot(seurat_no_hsp, features=c("CD4", "CD8A"))
  p3<- DimPlot(seurat_no_hsp, label=TRUE) +NoLegend()
  cowplot::plot_grid(plotlist=list(p1, p2, p3), ncol=1)
  ggsave(paste0(figure_dir, "Stress_NoHSPgenes_annotation.png"), width=6, height=9)
  
  cd4_clusters<-  c(2)
  cd8_clusters<-c(0:1, 4:5)
  unclear_clusters<- 3 
  
  seurat_no_hsp$detailed_annot_byProgram<- ifelse(seurat_no_hsp$seurat_clusters %in% cd4_clusters, "CD4",
                                         ifelse(seurat_no_hsp$seurat_clusters %in% cd8_clusters, "CD8",
                                                ifelse(seurat_no_hsp$seurat_clusters %in% unclear_clusters, "unclear",
                                                       "issue")))
  
  
  ## Add these annotations back to original seurat object (with hsp genes)
  seurat_tmp$detailed_annot_byProgram<- plyr::mapvalues(colnames(seurat_tmp), colnames(seurat_no_hsp),
                                                        seurat_no_hsp$detailed_annot_byProgram)  
  DimPlot(seurat_tmp, group.by = "detailed_annot_byProgram") + VlnPlot(seurat_tmp,
                                                                group.by = "detailed_annot_byProgram", 
                                                                features=c("CD4", "CD8A") )

  
  seurat_list$Stress<- seurat_tmp
```

## Option 4: sub-clustering required
```{r}
## Activated EM
seurat_tmp<- seurat_list$Activated_EM
seurat_tmp<- FindClusters(seurat_tmp, resolution = 1.5)

p1<- FeaturePlot(seurat_tmp, features=c("CD4", "CD8A")) 
p2<- VlnPlot(seurat_tmp, features=c("CD4", "CD8A"))
p3<- DimPlot(seurat_tmp, label=TRUE) +NoLegend()
cowplot::plot_grid(plotlist=list(p1, p2, p3), ncol=1)

cd4_clusters<- c(0:3,5:7,9)
unclear_clusters<- c(4,8)

seurat_tmp$detailed_annot_byProgram<- ifelse(seurat_tmp$seurat_clusters %in% unclear_clusters, "Subcluster",
                                              ifelse(seurat_tmp$seurat_clusters %in% cd4_clusters, "CD4",
                                              "issue"))

  ## Subcluster 
  subcluster_seurat<- subset(seurat_tmp, detailed_annot_byProgram == "Subcluster")
  subcluster_seurat<- RunFullSeurat(subcluster_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=subcluster_seurat$sample)
  subcluster_seurat$detailed_annot_byProgram<- ifelse(subcluster_seurat$seurat_clusters==1, "CD8", "CD4")  
  
  ## Add onto rest of inhibitory CD8 annotations
  seurat_tmp$detailed_annot_byProgram<- ifelse(colnames(seurat_tmp) %in% colnames(subcluster_seurat),
                                                   plyr::mapvalues(colnames(seurat_tmp), 
                                                                   colnames(subcluster_seurat),
                                                                   subcluster_seurat$detailed_annot_byProgram),
                                                   seurat_tmp$detailed_annot_byProgram)
  
 DimPlot(seurat_tmp, group.by = "detailed_annot_byProgram") + VlnPlot(seurat_tmp,
                                                                group.by = "detailed_annot_byProgram", 
                                                                features=c("CD4", "CD8A") )
  
seurat_list$Activated_EM<- seurat_tmp
    

```

## Save list with updated annotations
```{r}
rank_use<- 9
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use,  "/")

saveRDS(seurat_list, file=paste0(analysis_dir, "seurat_list_splitByNMF.Rds"))
```

## Plot total cd4/cd8 annot by program- compare to estimation at beginning of code
```{r}
rank_use<-9
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use, "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use,  "/")
seurat_list<- readRDS(paste0(analysis_dir, "seurat_list_splitByNMF.Rds"))

all_meta_list<- lapply(names(seurat_list), function(x){
  seurat_tmp<- seurat_list[[x]]
  tmp<-data.frame(annot=seurat_tmp$detailed_annot_byProgram, Program=x)
  return(tmp)
})
all_meta_withUnclear<- do.call("rbind", all_meta_list)

all_meta<- all_meta_withUnclear[all_meta_withUnclear$annot!="unclear",]

## Breakdown of percent of cd4/cd8 cells per program
df_bd<- as.data.frame(table(all_meta$annot, all_meta$Program))
colnames(df_bd)<-c("CellType", "Program", "NCells")
df_bd<- df_bd %>% group_by(Program) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()

## Plot
ggplot(df_bd, aes(x=Program, y=PerCells, fill=CellType))+
  geom_bar(stat="identity", position="stack")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("") + ylab("Proportion of cells")+
  geom_hline(yintercept = 0.5, linetype="dashed")
ggsave(paste0(figure_dir, "CD4.CD8.annotationBy_NMF.png"), width=5, height=3)

saveRDS(all_meta_withUnclear, file=paste0(analysis_dir, "all_annotations_byNMFProgram.Rds"))
```


## Add CD4/CD8 annotation onto full tcell seurat object
```{r}
rank_use<-9
nmf_annotations<- readRDS(paste0(analysis_dir, "all_annotations_byNMFProgram.Rds"))
variable_name<- paste0("detailed_annot_byNMFrank", rank_use)

## Add to tcell seurat object
tcell_seurat@meta.data[[variable_name]]<- plyr::mapvalues(colnames(tcell_seurat), rownames(nmf_annotations), nmf_annotations$annot)


saveRDS(tcell_seurat, file="analysis/manual_broad_annot/tcell_seurat.Rds")
```

## Split into cd4/cd8, re-run clustering
```{r}
rank_use<-9
variable_name<- paste0("detailed_annot_byNMFrank", rank_use)
tcell_seurat$tmp<- tcell_seurat@meta.data[[variable_name]]

## CD4
  ## Subset to CD4
  cd4_seurat<-subset(tcell_seurat, tmp=="CD4")
  cd4_meta<- cd4_seurat@meta.data
  
  ## Rerun clustering
  cd4_seurat<- RunFullSeurat(cd4_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=cd4_meta$sample)
  cd4_seurat<- AddMetaData(cd4_seurat, cd4_meta[,!colnames(cd4_meta) %in% colnames(cd4_seurat@meta.data)])
  
  ## Quick vis to check for any clear CD8- some in cluster 5. but also expresses CD4. kept
  DimPlot(cd4_seurat, label=TRUE, group.by = "seurat_clusters")
  VlnPlot(cd4_seurat, features=c("CD4", "CD8A"), ncol=1)
  FeaturePlot(cd4_seurat, features=c("CD4", "CD8A"), ncol=1) + DimPlot(cd4_seurat, label=TRUE)
  
  
## CD8
  ## Subset to CD4
  cd8_seurat<-subset(tcell_seurat, tmp=="CD8")
  cd8_meta<- cd8_seurat@meta.data
  
  ## Rerun clustering
  cd8_seurat<- RunFullSeurat(cd8_seurat@assays$RNA@counts, RunHarmony = TRUE, samples=cd8_meta$sample)
  cd8_seurat<- AddMetaData(cd8_seurat, cd8_meta[,!colnames(cd8_meta) %in% colnames(cd8_seurat@meta.data)])
  
  ## Quick vis to check for any clear CD8- some in cluster 2. sub-clustering shows small cd8 cluster, mostly in stress, but don't cluster in stress seurat. kept.
  DimPlot(cd8_seurat, label=TRUE, group.by = "seurat_clusters")
  VlnPlot(cd8_seurat, features=c("CD4", "CD8A"), ncol=1)
  FeaturePlot(cd8_seurat, features=c("CD4", "CD8A"), ncol=1) + DimPlot(cd8_seurat, label=TRUE) 


saveRDS(cd4_seurat, file=paste0(analysis_dir, "cd4_seurat.Rds"))
saveRDS(cd8_seurat, file=paste0(analysis_dir, "cd8_seurat.Rds"))
```

## Quick visualizations to confirm
```{r}
FeaturePlot(cd4_seurat, features=c("CD4", "CD8A", "GZMK", "TOP2A", "CCR7", "FOXP3"), ncol=3)
ggsave(paste0(figure_dir, "CD4_MarkerGeneExpression.png"), width=9, height=6)

FeaturePlot(cd8_seurat, features=c("CD4", "CD8A", "GZMK", "TOP2A", "CCR7", "FOXP3"), ncol=3)
ggsave(paste0(figure_dir, "CD8_MarkerGeneExpression.png"), width=9, height=6)
```


## Add finalized annotations to full immune seurat object 
```{r}
immune_seurat<- readRDS("analysis/preprocessing/seurat_obj_noTumor.Rds")
immune_seurat$detailed_annot<- ifelse(colnames(immune_seurat) %in% colnames(tcell_seurat),
                                      as.character(plyr::mapvalues(colnames(immune_seurat),
                                                                   colnames(tcell_seurat),
                                                                   tcell_seurat$detailed_annot_byNMFrank9)),
                                      "Myeloid")
saveRDS(immune_seurat, file="analysis/preprocessing/seurat_obj_noTumor.Rds")
```

## re-score cells split into cd4/cd8 for programs
```{r}
## Read in cd4/cd8 seurat objects
cd4_seurat<- readRDS(paste0(analysis_dir, "cd4_seurat.Rds"))
cd8_seurat<- readRDS(paste0(analysis_dir, "cd8_seurat.Rds"))

## Read in marker genes
rank_use<-9
program_markers<- read.csv(paste0("analysis/nmf/Tcell/top_DE_genes_NMF_rank", rank_use, "_annotated.csv"))
program_markers<- split(program_markers, f=program_markers$cluster)
program_markers<- lapply(program_markers, function(x){x$gene})

## Subset marker genes to cd4/cd8 specific
cd4_markers<- program_markers[!names(program_markers) %in% c("Cytotoxic.NKreceptor","Cytotoxic.NKreceptor_Alt",
                                                             "Predysfunctional")]
cd8_markers<- program_markers[!names(program_markers) %in% c("Treg")]

## Score CD4
cm_list<- NormCenter(cd4_seurat@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data+1))
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, cd4_markers)))
saveRDS(scores, file=paste0(analysis_dir, "cd4_program_scores.Rds"))

## Score CD8
cm_list<- NormCenter(cd8_seurat@assays$RNA@counts)
cm_mean<- rowMeans(log2(cm_list$raw_data+1))
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, cd8_markers)))
saveRDS(scores, file=paste0(analysis_dir, "cd8_program_scores.Rds"))


```

