---
title: "Visualize_programs"
output: html_document
---

## Purpose:
After re-annotating programs based on correlation with full cohort, visualize updated programs (CD8 only) here


## Approach:
Using updated annotations, plot nice plots for heatmap, UMAP, density, etc


```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)
library(harmony)
library(ggpubr)


base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/tenX/Immune/")
analysis_dir_tmp<- paste0(working_dir, "/analysis/NMF_plots/CD4/")
figure_dir_tmp<- paste0(working_dir, "/figures/NMF_plots/CD4/")

if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp,recursive = TRUE)}
if(!dir.exists(analysis_dir_tmp)){dir.create(analysis_dir_tmp, recursive = TRUE)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

```

## Read in data
```{r}
rank_use<-9
data_dir<- paste0("analysis/identifyCD4.CD8_byNMF/", "rank",rank_use, "/" )

## figure/analysis dir based on rank + how shared programs are merged
figure_dir<- paste0(figure_dir_tmp, "rank", rank_use,  "/")
analysis_dir<- paste0(analysis_dir_tmp, "rank", rank_use, "/")

if(!dir.exists(figure_dir)){dir.create(figure_dir,recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}
  
## ped/adult tcell seurat object
cd4_seurat<- readRDS(paste0(data_dir, "cd4_seurat.Rds"))


## Read in marker genes
program_markers<- read.csv(paste0("analysis/nmf/Tcell/top_DE_genes_NMF_rank", rank_use, "_annotated.csv"))


## Read in scores
all_scores<- readRDS(paste0(data_dir, "cd4_program_scores.Rds"))

## Read in colors to use
program_colors<- readRDS(paste0("../../plot_colors/Tcell.programs_rank8sharedByunion_reAnnot.Rds"))
celltype_colors<- readRDS(paste0("../../plot_colors/Tcell.celltypes.Rds"))
program_colors<- c(program_colors, Cytotoxic.NKreceptor_Alt="palegreen", Activated_EM="red")
```


## Subset to top markers for each nmf- top 10
```{r}
marker_genes<- program_markers

## If any genes are top markers for multiple clusters- remove, then find top 10 again
top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
if(sum(duplicated(top_markers$gene))!=0){
  duplicated<- top_markers$gene[duplicated(top_markers$gene)]
  print(paste0("Shared top markers: ", paste(duplicated, collapse=",")))
  marker_genes<- marker_genes[!marker_genes$gene %in% duplicated,]
  top_markers<- marker_genes %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 10) %>% as.data.frame()
}
all_markers<- top_markers
all_markers$cluster<- as.character(all_markers$cluster)
```



## Plot heatmaps of top marker gene expression
```{r}
rank<- "9"
add_lines<- TRUE
program_name<-paste0("NMF_rank", rank_use, "_annot")

top_markers<-all_markers[!all_markers$cluster %in% c("Cytotoxic.NKreceptor", "Cytotoxic.NKreceptor_Alt", "Predysfunctional"),]

## Pseudobulk
cd4_seurat$PB_variable<- paste0(cd4_seurat$sample, "_", 
                                gsub("\\.", "\n", gsub("_", "\n",cd4_seurat@meta.data[[program_name]])))

pb<- pseudobulk_byVariable(cd4_seurat@assays$RNA@counts, cd4_seurat@meta.data, "PB_variable")

## center and score
pb_list<- NormCenter(pb)
pb_center<- pb_list$center_data

## Plot
p<- myHeatmap(pb_center, GOI=top_markers$gene, min.value = -3, max.value = 3, 
              orderFactors =  gsub("\\.", "\n",gsub("_", "\n", unique(top_markers$cluster)))) +
  scale_fill_gradient2(high="red", low="blue", mid="white", name="Centered log2(TPM)")

## Add horizontal lines around each set of nmf genes
if(add_lines){
  tmp<- split(top_markers, f=top_markers$cluster)
  lastGene<- lapply(tmp, function(x){
    x$tmp<- 1:nrow(x)
    x<- x[x$gene %in% rownames(pb), ]
    return(x[max(x$tmp), "gene"])
  })
  linePositions<- as.data.frame(pb_center[top_markers$gene,])
  linePositions$tmp<- 1:nrow(linePositions)
  linePositions<- linePositions[unlist(lastGene),"tmp"]
  linePositions<- linePositions[1:length(linePositions)-1]+ 0.5
  
  p+geom_hline(yintercept = linePositions, linetype="dashed",size=1)
}else{p}

  
ggsave(paste0(figure_dir, "CD4_Heatmap_TopNMFgenes_annot_rank", rank_use,".png"), width=14, height=11)


```

## Dotplot of selected marker genes
```{r}
rank<- "9"
add_lines<- TRUE
program_name<-paste0("NMF_rank", rank_use, "_annot")

cd4_seurat@meta.data[[program_name]]<- factor(cd4_seurat@meta.data[[program_name]],
                                              levels=c("Treg", "Activated_EM", "Cycling", 
                                                        "Memory",
                                                       "IFN.Response", "Stress"))
DotPlot(cd4_seurat, group.by =program_name,
        features=c("FOXP3", "CTLA4", "CD40LG", "TIMP1", "KLRB1","TOP2A", "UBE2C",
                   "LMNA", "CCR7", "SELL", "IFIT3","IFIT1",
                     "HSPA6", "HSPA1A"))+
  coord_flip()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  scale_color_gradient2(low="blue", mid="white", high="red")
ggsave(paste0(figure_dir, "DotPlot_MarkersOfInterest_byProgram.png"), width=6, height=8)
```

## UMAP with annotations
```{r}
rank<- "9"
program_name<-paste0("NMF_rank", rank_use, "_annot")
program_colors_use<- program_colors[names(program_colors) %in% unique(cd4_seurat@meta.data[[program_name]])]

DimPlot(cd4_seurat, group.by = program_name)+
  scale_color_manual(values=program_colors_use)+
  ggtitle("")
ggsave(paste0(figure_dir, "CD4_UMAP_annot.pdf"), device="pdf", width=6, height=4)

## Plot each program one at a time
seurat_tmp<- cd4_seurat
all_plots<- lapply(as.character(unique(cd4_seurat@meta.data[[program_name]])), function(x){
  seurat_tmp$tmp<- ifelse(seurat_tmp@meta.data[[program_name]]==x,
                          x, "Other")
  p<-DimPlot(seurat_tmp, group.by = "tmp")+scale_color_manual(values=c(program_colors_use, Other="grey"))+ggtitle(x)+NoLegend()
  return(p)
})
cowplot::plot_grid(plotlist=all_plots, ncol = 2)
ggsave(paste0(figure_dir, "CD4_UMAP_indv.png"), width=6, height=10)
```


## Programs by sample
```{r}
program_name<-paste0("NMF_rank", rank_use, "_annot")
cd4_seurat$tmp<- cd4_seurat@meta.data[[program_name]]

my_barchart_programProp(cd4_seurat, coloring_variable = "tmp", colors=program_colors)
ggsave(paste0(figure_dir, "Barchart_programProportion_bySample.png"), width=5, height=4)
```





