---
title: "R Notebook"
output: html_document
---

## Purpose:
Nice plots (mainly UMAP) of immune cells- myeloid/tcell together and separate, by cell type, by sample, subtype, etc

## Approach:
Cell type annotations for tcells (CD4/CD8) based on "NMF approach"- briefly, run tcell nmf --> identify programs --> split seurat by program --> identify cd4/cd8 cluster-wise for each program. This analysis performed for ped/adult simultaneously. This analysis performed in 02a_Tcells/02_Ped.Adult/02_detailed_annot/IdentifyCD4.CD8_byNMF.Rmd


```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/03_Basic_Visualizations/")

library(Seurat)
library(ggplot2)
library(ggpubr)
library(randomcoloR)


figure_dir_tmp<- "figures/BasicVis/"; if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp)}
tcell_data_dir<- "../02a_Tcells/02_Ped.Adult/03_Visualize.Reannotate_programs/analysis/Reannotate_programs/"
myeloid_data_dir<- "../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/"

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)
```

## Read in processed seurat objects
```{r}
rank_use<- "8"; sharedBy="union"; cohort<- "pedOnly_nomuv63"
tcell_data_dir<- paste0(tcell_data_dir, "rank", rank_use, "_sharedBy", sharedBy, "/")
myeloid_data_dir<- paste0(myeloid_data_dir, cohort, "/")

figure_dir<- paste0(figure_dir_tmp, cohort, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)}


## Read in seurat objects
immune_seurat<- readRDS(paste0(myeloid_data_dir, "tcell.myeloid_seurat_harmony.Rds"))
myeloid_seurat<- readRDS(paste0(myeloid_data_dir, "myeloid_seurat_harmony.Rds"))
myeloid_seurat_noDC.Bcell<- readRDS(paste0(myeloid_data_dir, "myeloid_seurat_noDC.B_harmony.Rds"))
tcell_seurat<- readRDS(paste0(tcell_data_dir, "ped_seurat_tcell_harmony.Rds"))
cd4_seurat<- readRDS(paste0(tcell_data_dir, "ped_seurat_cd4_harmony.Rds"))
cd8_seurat<- readRDS(paste0(tcell_data_dir, "ped_seurat_cd8_harmony.Rds"))
adult_tcell<- readRDS(paste0(tcell_data_dir, "adult_seurat_tcell_harmony.Rds"))
ped.adult_tcell<-readRDS(paste0(tcell_data_dir, "ped.adult_seurat_tcell_harmony.Rds"))

## Read in colors
immune_colors<- readRDS("../plot_colors/ImmuneCell.celltypes.Rds")
subtype_colors<-  readRDS("../plot_colors/Subtype.Rds")
```

# Need to re-annotate BT1478 to non-BCOR; hemispheric, maybe GBM? Set to GBM for now, may update
```{r}
updateBCOR<- TRUE

figure_dir<- ifelse(updateBCOR, paste0(figure_dir, "updateBCOR/"), figure_dir_tmp)

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}


if(updateBCOR){
  tcell_seurat$NiceSubtype<- ifelse(tcell_seurat$sample=="BT1478", "Hemispheric-HistoneWT", tcell_seurat$NiceSubtype)
  ped.adult_tcell$NiceSubtype<- ifelse(ped.adult_tcell$sample=="BT1478", "Hemispheric-HistoneWT", ped.adult_tcell$NiceSubtype)
  myeloid_seurat_noDC.Bcell$NiceSubtype<- ifelse( myeloid_seurat_noDC.Bcell$sample=="BT1478", "Hemispheric-HistoneWT",
                                                   myeloid_seurat_noDC.Bcell$NiceSubtype)
}
```

## Plotting themes
```{r}
umap_theme<- function(p, colors=immune_colors){
  p+NoLegend()+
  scale_color_manual(values=colors[names(colors) %in% unique(seurat_plot$plotting)])+
  theme(legend.position = "bottom",
        legend.text = element_text(face="bold"),
      plot.title = element_blank(),
      axis.text = element_text(face="bold", colour = "black"),
      axis.title = element_text(face="bold", colour = "black"))+
  ylab("UMAP2")+xlab("UMAP1")
}

bargraph_bd_theme<- function(){
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black"),
        axis.text = element_text(color="black", face="bold"),
        axis.title =  element_text(color="black", face="bold"),
        axis.text.y = element_text(color="black"),
        legend.position = "bottom",
        legend.title=element_blank())
  }


```

## Additions to seurat objects for nicer plotting
```{r}
tcell_seurat$NiceSubtype_Merge<-gsub("-H3G34R/V", "", gsub("-HistoneWT", "", gsub("-H3K27M", "", gsub("-BCOR", "", tcell_seurat$NiceSubtype))))
tcell_seurat$detailed_annot_withCycling_byNMF<- ifelse(tcell_seurat$detailed_annot_withCycling=="Cycling", "Cycling", 
                                                       tcell_seurat$detailed_annot_byNMFrank8_mergedByunion)

adult_tcell$detailed_annot_withCycling<- ifelse(adult_tcell$detailed_annot_withCyc.Treg.Cyto=="Cycling", "Cycling",
                                                adult_tcell$detailed_annot_byNMFrank8_mergedByunion)
```


## Tcell/Myeloid together, colored by major annot
```{r}
## By T cell/Myeloid
seurat_plot<- immune_seurat
seurat_plot$plotting<- seurat_plot$broad_annot
DimPlot(seurat_plot, group.by = "plotting", label = TRUE, label.size = 10) %>% umap_theme()
ggsave(paste0(figure_dir, "TCellMyeloid.png"), width=6, height=5.5)

## With CD4/CD8 distinction
seurat_plot<- subset(immune_seurat, detailed_annot_byNMF != "Unclear")
seurat_plot$plotting<- gsub("DC", "Myeloid", gsub("BCell", "Myeloid", seurat_plot$detailed_annot_byNMF))

DimPlot(seurat_plot, group.by = "plotting", label=TRUE, label.size = 8) %>% umap_theme()
ggsave(paste0(figure_dir, "CD4CD8Myeloid.png"), width=6, height=5.5)

## With CD4/CD8/DC/Bcell distinction
seurat_plot<- subset(immune_seurat, detailed_annot_byNMF != "Unclear")
seurat_plot$plotting<- seurat_plot$detailed_annot_byNMF
DimPlot(seurat_plot, group.by = "plotting", label=TRUE, label.size = 5) %>% umap_theme()
ggsave(paste0(figure_dir, "CD4CD8.BcellDC.png"), width=6, height=5.5)

## Add on monocyte annotation (confirmed with NMF programs)- just to explain far-clustering myeloid cluster
## Need to increase resolution first
seurat_plot<- FindClusters(immune_seurat, resolution = 1.5)
seurat_plot<- subset(seurat_plot, detailed_annot_byNMF != "Unclear")
seurat_plot$plotting<- ifelse(seurat_plot$seurat_clusters=="10", "Monocyte",seurat_plot$detailed_annot_byNMF)
immune_colors<- c(immune_colors, Monocyte="orange")
DimPlot(seurat_plot, group.by = "plotting", label=TRUE, label.size = 5) %>% umap_theme()
ggsave(paste0(figure_dir, "CD4CD8.BcellDCMonocyte.png"), width=6, height=5.5)

## With DC/Bcell/monocyte, but no CD4/CD8
seurat_plot$plotting<- gsub("CD4", "Tcell", gsub("CD8", "Tcell", seurat_plot$plotting))
DimPlot(seurat_plot, group.by = "plotting", label=TRUE, label.size = 5) %>% umap_theme()
ggsave(paste0(figure_dir, "CD4CD8.BcellDCMonocyte_NoCD4.CD8.pdf"), device="pdf", width=6, height=5.5)

## To support myeloid/tcell annotation: show myeloid marker genes
FeaturePlot(immune_seurat, features=c( "CSF1R", "CD3G", "CD8A", "CD4"), ncol = 4, cols = c("grey", "red"))
ggsave(paste0(figure_dir, "MyeloidTcellMarkers.png"), width=24, height=5.5)

## To support myeloid, t cell, monocyte, bcell, and dc annotation: show top marker genes
  ## UMAP
  cowplot::plot_grid(plotlist=lapply(c("CD3G", "CSF1R", "FCN1", "CD207", "MS4A1" ), function(x){
    FeaturePlot(immune_seurat, features=x, pt.size = 0.2)+
      scale_color_gradient(low="grey80", high="red")
  })) 
  ggsave(paste0(figure_dir, "Tcell.Myeloid.Monocyte.DC.BC_markers_UMAP.pdf"), device="pdf", width=11, height=6)
  
  ## VlnPlot
  cowplot::plot_grid(plotlist=lapply(c("CD3G", "CSF1R", "FCN1", "CD207", "MS4A1" ), function(x){
    VlnPlot(seurat_plot, group.by = "plotting", features=x, sort=FALSE, ncol=1)+ NoLegend()+
      scale_fill_manual(values=immune_colors)+
      xlab("")
  }), ncol=5) 
  ggsave(paste0(figure_dir, "Tcell.Myeloid.Monocyte.DC.BC_markers_Vln.pdf"), device="pdf", width=11, height=4)

## Breakdown of myeloid/tcells by subtype
df_bd<- as.data.frame(table(immune_seurat$NiceSubtype, immune_seurat$broad_annot))
colnames(df_bd)<- c("Subtype", "CellType", "NCells")
df_bd <- df_bd %>% group_by(Subtype) %>% mutate(CellProportion=NCells/sum(NCells)) %>% as.data.frame()

ggplot(df_bd, aes(x=Subtype, y=CellProportion, fill=CellType))+
  geom_bar(position = "stack", stat="identity") + 
  theme_classic()+
  bargraph_bd_theme()+
  xlab("")+ ylab("Proportion of cell type")+
  scale_fill_manual(values=c(immune_colors[names(immune_colors) %in% unique(df_bd$CellType)]))
ggsave(paste0(figure_dir, "Immune_broadAnnot_BreakdownAcrossAllSubtypes.pdf"), device="pdf", width=4, height=4)
```


## Tcells: CD4/CD8/Cycling 
```{r}
## CD4/CD8/Cycling
## cycling distinction added on (based on non-NMF based approach, but won't change cycling cells)
seurat_plot<- subset(tcell_seurat,detailed_annot_byNMFrank8_mergedByunion != "Unclear")
seurat_plot$plotting<- seurat_plot$detailed_annot_withCycling_byNMF
DimPlot(seurat_plot, group.by = "plotting", label = TRUE, label.size = 8) %>% umap_theme()
ggsave(paste0(figure_dir, "TCell_CD4.CD8.Cycling.png"), width=6, height=5.5)


## Treg markers: to show that they don't form a distinct cluster
FeaturePlot(tcell_seurat, features=c("FOXP3", "IL2RA", "CTLA4"), ncol = 1, cols=c("grey", "red"))
ggsave(paste0(figure_dir, "Tcell_TregMarkers.png"), width=6, height=18)

## CD4/CD8 only, no cycling
seurat_plot<- subset(tcell_seurat,detailed_annot_byNMFrank8_mergedByunion != "Unclear")
seurat_plot$plotting<- seurat_plot$detailed_annot_byNMFrank8_mergedByunion
DimPlot(seurat_plot, group.by = "plotting", label = TRUE, label.size = 8) %>% umap_theme()
ggsave(paste0(figure_dir, "TCell_CD4.CD8.png"), width=6, height=5.5)
```

## Myeloid cells: by B cells/DC/Other
```{r}
seurat_plot<- myeloid_seurat;seurat_plot$plotting<- seurat_plot$detailed_annot
DimPlot(seurat_plot, group.by = "plotting", label = TRUE, label.size = 8) %>% umap_theme()
ggsave(paste0(figure_dir, "Myeloid_WithDC.BCells.png"), width=6, height=5.5)
```

## By sample
```{r}
## Both T cell /myeloid
colors_use<- unique(immune_seurat$sampleid)
colors_use<- distinctColorPalette(length(unique(immune_seurat$sampleid)))
names(colors_use)<-unique(immune_seurat$sampleid)
seurat_plot<- immune_seurat; seurat_plot$plotting<- seurat_plot$sampleid
DimPlot(seurat_plot, group.by = "plotting", label = FALSE) %>% umap_theme(colors=colors_use)
ggsave(paste0(figure_dir, "TCellMyeloid_Sample.pdf"), device="pdf", width=6, height=6.5)

## T cells
seurat_plot<- tcell_seurat; seurat_plot$plotting<- seurat_plot$sampleid
DimPlot(seurat_plot, group.by = "plotting", label = FALSE) %>% umap_theme(colors=colors_use)
ggsave(paste0(figure_dir, "Tcells_Sample.png"), width=6.5, height=5.5)

## Myeloid
seurat_plot<- myeloid_seurat; seurat_plot$plotting<- seurat_plot$sampleid
DimPlot(seurat_plot, group.by = "plotting", label = FALSE) %>% umap_theme(colors=colors_use)
ggsave(paste0(figure_dir, "Myeloid_Sample.png"), width=6.5, height=5.5)

## Myeloid, without dc/bcells
seurat_plot<- myeloid_seurat_noDC.Bcell; seurat_plot$plotting<- seurat_plot$sampleid
DimPlot(seurat_plot, group.by = "plotting", label = FALSE) %>% umap_theme(colors=colors_use)
ggsave(paste0(figure_dir, "Myeloid_Sample_noDC.Bcell.png"), width=6.5, height=5.5)

```


## By subtype
```{r}
## By detailed subtype
  ## Tcell
  seurat_plot<- tcell_seurat; seurat_plot$plotting<- seurat_plot$NiceSubtype
  DimPlot(seurat_plot, group.by = "plotting", label = FALSE) %>% umap_theme(colors=subtype_colors) +
    theme(legend.position = "right")
  ggsave(paste0(figure_dir, "Tcell_Subtype.png"), width=7, height=5)
  
  ## Myeloid
  seurat_plot<- myeloid_seurat_noDC.Bcell; seurat_plot$plotting<- seurat_plot$NiceSubtype
  DimPlot(seurat_plot, group.by = "plotting", label = FALSE) %>% umap_theme(colors=subtype_colors) +
    theme(legend.position = "right")
  ggsave(paste0(figure_dir, "Myeloid.noBC.DC_Subtype.png"), width=7, height=5)


## By broad subtype
colors_use<- c(Ependymoma="magenta3",
               `Hemispheric`="deepskyblue1",
               `Midline`="chartreuse3")

  ## Tcell
  seurat_plot<- tcell_seurat; seurat_plot$plotting<- seurat_plot$NiceSubtype_Merge
  DimPlot(seurat_plot, group.by = "plotting", label = FALSE) %>% umap_theme(colors=colors_use) 
  ggsave(paste0(figure_dir, "Tcell_BroadSubtype.png"), width=6, height=5)

  ## Myeloid
  seurat_plot<- myeloid_seurat_noDC.Bcell; seurat_plot$plotting<- seurat_plot$NiceSubtype_Merge
  DimPlot(seurat_plot, group.by = "plotting", label = FALSE) %>% umap_theme(colors=colors_use) 
  ggsave(paste0(figure_dir, "Myeloid.noBC.DC_BroadSubtype.png"), width=6, height=5)


```

## Breakdown of cell proportions by subtype
```{r}
## Tcell, with cycling
df_bd<- as.data.frame(table(tcell_seurat$NiceSubtype, tcell_seurat$detailed_annot_withCycling_byNMF))
colnames(df_bd)<- c("Subtype", "CellType", "NCells")
df_bd$CellType<- as.character(df_bd$CellType)
df_bd<- df_bd[df_bd$CellType!="Unclear",]
df_bd<- df_bd %>% group_by(Subtype) %>% mutate(CellProportion=NCells/sum(NCells)) %>% as.data.frame()

ggplot(df_bd, aes(x=Subtype, y=CellProportion, fill=CellType))+
  geom_bar(position = "stack", stat="identity") + 
  theme_classic()+
  bargraph_bd_theme()+
  xlab("")+ ylab("Proportion of cell type")+
  scale_fill_manual(values=c(immune_colors[names(immune_colors) %in% unique(df_bd$CellType)]))
ggsave(paste0(figure_dir, "Tcell_BreakdownAcrossAllSubtypes_WithCycling.png"), width=4, height=4)


## Tcell, by major subtype
df_bd<- as.data.frame(table(tcell_seurat$NiceSubtype_Merge, tcell_seurat$detailed_annot_withCycling_byNMF))
colnames(df_bd)<- c("Subtype", "CellType", "NCells")
df_bd$CellType<- as.character(df_bd$CellType)
df_bd<- df_bd[df_bd$CellType!="Unclear",]
df_bd<- df_bd %>% group_by(Subtype) %>% mutate(CellProportion=NCells/sum(NCells)) %>% as.data.frame()

ggplot(df_bd, aes(x=Subtype, y=CellProportion, fill=CellType))+
  geom_bar(position = "stack", stat="identity") + 
  theme_classic()+
  bargraph_bd_theme()+
  xlab("")+ ylab("Proportion of cell type")+
  scale_fill_manual(values=c(immune_colors[names(immune_colors) %in% unique(df_bd$CellType)]))
ggsave(paste0(figure_dir, "Tcell_BreakdownAcrossMajorSubtypes.png"), width=4, height=4)

## Myeloid
df_bd<- as.data.frame(table(myeloid_seurat$NiceSubtype, myeloid_seurat$detailed_annot))
colnames(df_bd)<- c("Subtype", "CellType", "NCells")
df_bd$CellType<- as.character(df_bd$CellType)
df_bd<- df_bd %>% group_by(Subtype) %>% mutate(CellProportion=NCells/sum(NCells)) %>% as.data.frame()

ggplot(df_bd, aes(x=Subtype, y=CellProportion, fill=CellType))+
  geom_bar(position = "stack", stat="identity") + 
  theme_classic()+
  bargraph_bd_theme()+
  xlab("")+ ylab("Proportion of cell type")+
  scale_fill_manual(values=c(immune_colors[names(immune_colors) %in% unique(df_bd$CellType)]))
ggsave(paste0(figure_dir, "Myeloid_BreakdownAcrossAllSubtypes.png"), width=4, height=4)

## Myeloid- breakdown cell type by program
df_bd<- as.data.frame(table(myeloid_seurat$NiceSubtype, myeloid_seurat$detailed_annot))
colnames(df_bd)<- c("Subtype", "CellType", "NCells")
df_bd$CellType<- as.character(df_bd$CellType)

  ## Normalize so that all subtypes are weighted the same, normalizing for total size
  nCells<- table(myeloid_seurat$NiceSubtype)
  nCells_norm<- max(nCells)/nCells
  df_bd$Norm_factor<-as.numeric(as.character(plyr::mapvalues(df_bd$Subtype, names(nCells_norm), nCells_norm)))
  df_bd$NCells_norm<- df_bd$NCells * df_bd$Norm_factor

  df_bd<- df_bd %>% group_by(CellType) %>% mutate(CellProportion=NCells_norm/sum(NCells_norm)) %>% as.data.frame()

ggplot(df_bd, aes(x=CellType, y=CellProportion, fill=Subtype))+
  geom_bar(position = "stack", stat="identity")+
  theme_classic()+
  bargraph_bd_theme()+
  xlab("")+ ylab("Proportion of cell types")+
  scale_fill_manual(values=subtype_colors)
ggsave(paste0(figure_dir, "Myeloid_BreakdownOfCellTypes_AcrossAllSubtypes.png"), width=4, height=3)
```

## clonal vs nonclonal t cells
```{r}
df<- tcell_seurat@meta.data

df_bd<- as.data.frame(table(df$sampleid, df$clonal))
colnames(df_bd)<- c("Sample", "Clonality", "NCells")

## Convert to proportion of cells
df_bd <- df_bd %>% group_by(Sample) %>% mutate(PerCells=NCells/sum(NCells)) %>% as.data.frame()

## Add subtype
df_bd$Subtype<- plyr::mapvalues(df_bd$Sample, tcell_seurat$sampleid, tcell_seurat$NiceSubtype, warn_missing = FALSE)
df_bd$Subtype<- gsub("-", "\n", gsub("Hemispheric", "Hemi", gsub("Histone", "Hist", df_bd$Subtype)))

## Plot
ggplot(df_bd, aes(x=Sample, y=PerCells, fill=Clonality))+
  geom_bar(stat="identity", position="stack")+
  facet_grid(cols=vars(Subtype), scale="free_x", space="free_x")+
  theme_classic()+
  bargraph_bd_theme()+
  ylab("Proportion of Tcells")+
  xlab("")+
  scale_fill_manual(values=c(clonal="red", nonclonal="grey"))
ggsave(paste0(figure_dir, "Tcells_clonal.vs.nonclonal.png"), width=10, height=3)
```


##  T cell types, with adult
```{r}
## Plot and save adult umap with major annotations
seurat_plot<- subset(ped.adult_tcell, Final_detailed_annot != "Unclear")
seurat_plot$plotting<- seurat_plot$Final_detailed_annot
DimPlot(seurat_plot, group.by = "plotting",label=TRUE, label.size = 5) %>% umap_theme(immune_colors)
ggsave(paste0(figure_dir, "Ped.Adult_TCellMajorCellTypes.pdf"), device="pdf", width=5, height=5.5)

## Bar plot of adult major annots + ped major annots. Grouped by subtype
  ## Ped
  df_bd<- as.data.frame(table(tcell_seurat$NiceSubtype, tcell_seurat$detailed_annot_withCycling_byNMF))
  colnames(df_bd)<- c("Subtype", "CellType", "NCells")
  df_bd<- df_bd[df_bd$CellType!="NA",]
  df_bd<- df_bd %>% group_by(Subtype) %>% mutate(CellProportion=NCells/sum(NCells)) %>% as.data.frame()
  ped_df_bd<- df_bd
  ped_df_bd$Age<- "Ped"

  ## Adult
  adult_df_bd<- as.data.frame(table(adult_tcell$NiceSubtype, adult_tcell$detailed_annot_withCycling))
  colnames(adult_df_bd)<- c("Subtype", "CellType", "NCells")
  adult_df_bd<- adult_df_bd[adult_df_bd$CellType!="NA",]
  adult_df_bd<- adult_df_bd %>% group_by(Subtype) %>% mutate(CellProportion=NCells/sum(NCells)) %>% as.data.frame()
  adult_df_bd$Age<- "Adult"

df_bd<- rbind(ped_df_bd, adult_df_bd)

ggplot(df_bd, aes(x=Subtype, y=CellProportion, fill=CellType))+
  geom_bar(position = "stack", stat="identity")+
  theme_classic()+
  bargraph_bd_theme()+
  xlab("")+ ylab("Percentage of cell types")+
  scale_fill_manual(values=immune_colors[names(immune_colors) %in% unique(df_bd$CellType)])+
  facet_grid(cols=vars(Age), scale="free_x", space="free_x")
ggsave(paste0(figure_dir, "CellTypeBreakdownAcrossSubtypes_WithCycling_WithAdult.png"), width=5, height=4)
```




## Saw poor quality after FACS for MUV92/94, ok for MUV079
## Does the quality here reflect this?
```{r}
## T cells
df<-tcell_seurat@meta.data
df$FACS_Qual<- ifelse(df$sampleid %in% c("MUV092", "MUV094"), "LowQual", 
                      ifelse(df$sampleid=="MUV079", "HighQual", "Unknown"))
ggplot(df, aes(x=reorder(sampleid, nFeature_RNA, FUN=median),y=nFeature_RNA, fill=FACS_Qual))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("")+
  scale_fill_manual(values=c(LowQual="red", HighQual="green", Unknown="grey"))+
  ggtitle("T cells: nGenes detected\nQuality by FACS")

VlnPlot(tcell_seurat, group.by = "sampleid", features="nFeature_RNA")

## Myeloid
df<-myeloid_seurat@meta.data
df$FACS_Qual<- ifelse(df$sampleid %in% c("MUV092", "MUV094"), "LowQual", 
                      ifelse(df$sampleid=="MUV079", "HighQual", "Unknown"))
ggplot(df, aes(x=reorder(sampleid, nFeature_RNA, FUN=median),y=nFeature_RNA, fill=FACS_Qual))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab("")+
  scale_fill_manual(values=c(LowQual="red", HighQual="green", Unknown="grey"))+
  ggtitle("T cells: nGenes detected\nQuality by FACS")

```

## For comparison/visualization purposes only (not used anywhere in manuscript): how does UMAP look without ANY integration?
```{r}
## Read in immune seurat
immune_seurat<- readRDS(paste0(data_dir, "tcell.myeloid_seurat.Rds"))

## Re-process using our standard pipeline but without any integration
immune_meta<- immune_seurat@meta.data
immune_seurat<- RunFullSeurat(immune_seurat@assays$RNA@counts, RunHarmony = FALSE, samples=immune_meta$sampleid)
immune_seurat<- AddMetaData(immune_seurat, immune_meta[,!colnames(immune_meta) %in% colnames(immune_seurat@meta.data)])

## plot
immune_seurat_tmp<- subset(immune_seurat, detailed_annot != "Tcell_UnclearCD4.CD8")
DimPlot(immune_seurat_tmp, group.by = "detailed_annot", label=TRUE, label.size = 5)+
  scale_color_manual(values=c(CD4="navy", 
                              CD8="cadetblue3",
                              DC="violetred",
                              BCell="olivedrab",
                              Myeloid="gold3"))+
    theme(legend.position = "bottom",
        plot.title = element_blank())+
  ylab("UMAP2")+xlab("UMAP1")+ NoLegend()
```

