---
title: "R Notebook"
output: html_document
---

## Purpose:
Nice plots (mainly UMAP) of immune cells- myeloid/tcell together and separate, by cell type, by sample, subtype, etc

## Approach:
Cell type annotations for tcells (CD4/CD8) based on "NMF approach"- briefly, run tcell nmf --> identify programs --> split seurat by program --> identify cd4/cd8 cluster-wise for each program. This analysis performed for ped/adult simultaneously. This analysis performed in 02a_Tcells/02_Ped.Adult/02_detailed_annot/IdentifyCD4.CD8_byNMF.Rmd


```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/03_Basic_Visualizations/")

library(Seurat)
library(ggplot2)
library(ggpubr)
library(randomcoloR)


figure_dir_tmp<- "figures/PedAdult.MyeloidTcell/"; if(!dir.exists(figure_dir_tmp)){dir.create(figure_dir_tmp)}
tcell_data_dir<- "../02a_Tcells/02_Ped.Adult/03_Visualize.Reannotate_programs/analysis/Reannotate_programs/"
myeloid_data_dir<- "../02b_Myeloid/03_CompareToAdult/analysis/Merge_AssessIntegration/"

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)
```

## Read in processed seurat objects
```{r}
rank_use<- "8"; sharedBy="union"; cohort<- "pedOnly_nomuv63"; adultWithMUV<- FALSE
tcell_data_dir<- paste0(tcell_data_dir, "rank", rank_use, "_sharedBy", sharedBy, "/")
myeloid_data_dir<- paste0(myeloid_data_dir, cohort, "/adult_WithMUV.", adultWithMUV, "/" )

figure_dir<- paste0(figure_dir_tmp, cohort, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir)}


## Read in seurat objects
myeloid_seurat<- readRDS(paste0(myeloid_data_dir, "myeloid_seurat_Ped.AdultGBM.IDHMut_merged.Rds"))
tcell_seurat<- readRDS(paste0(tcell_data_dir, "ped.adult_seurat_tcell_harmony.Rds"))


## Read in colors
immune_colors<- readRDS("../plot_colors/ImmuneCell.celltypes.Rds")
subtype_colors<-  readRDS("../plot_colors/Subtype.Rds")
```

## Plotting themes
```{r}
umap_theme<- function(p, colors=immune_colors){
  p+NoLegend()+
  scale_color_manual(values=colors[names(colors) %in% unique(seurat_plot$plotting)])+
  theme(legend.position = "bottom",
        legend.text = element_text(face="bold"),
      plot.title = element_blank(),
      axis.text = element_text(face="bold", colour = "black"),
      axis.title = element_text(face="bold", colour = "black"))+
  ylab("UMAP2")+xlab("UMAP1")
}




```

## Additions to seurat objects for nicer plotting
```{r}
tcell_seurat$NiceSubtype_Merge<-gsub("-H3G34R/V", "", gsub("-HistoneWT", "", gsub("-H3K27M", "", gsub("-BCOR", "", tcell_seurat$NiceSubtype))))
tcell_seurat$detailed_annot_withCycling_byNMF<- ifelse(tcell_seurat$detailed_annot_withCycling=="Cycling", "Cycling", 
                                                       tcell_seurat$detailed_annot_byNMFrank8_mergedByunion)


```


## Merge myeloid/tcell together
```{r}
library(harmony)

## Add on broad annotation
myeloid_seurat$broad_annot<- "Myeloid"
tcell_seurat$broad_annot<- "Tcell"

## Merge
myeloid_seurat$sample<- myeloid_seurat$sampleid
immune_seurat<- merge(myeloid_seurat, tcell_seurat)
immune_meta<- immune_seurat@meta.data

## Rerun processing
immune_seurat<- RunFullSeurat(immune_seurat@assays$RNA@counts, samples=immune_meta$sample, RunHarmony = TRUE)
immune_seurat<- AddMetaData(immune_seurat, immune_meta[,!colnames(immune_meta) %in% colnames(immune_seurat@meta.data)])
```


## Plot by broad annotation and age
```{r}
## By T cell/Myeloid
seurat_plot<- immune_seurat
seurat_plot$plotting<- seurat_plot$broad_annot
DimPlot(seurat_plot, group.by = "plotting", label = TRUE, label.size = 10) %>% umap_theme()
ggsave(paste0(figure_dir, "TCellMyeloid.png"), width=6, height=5.5)

## With CD4/CD8 distinction
immune_seurat$Final_detailed_annot<- ifelse(immune_seurat$broad_annot=="Myeloid", "Myeloid", immune_seurat$Final_detailed_annot)
seurat_plot<- subset(immune_seurat, Final_detailed_annot != "Unclear")
seurat_plot$plotting<- seurat_plot$Final_detailed_annot

DimPlot(seurat_plot, group.by = "plotting", label=TRUE, label.size = 8) %>% umap_theme()
ggsave(paste0(figure_dir, "CD4CD8Myeloid.png"), width=6, height=5.5)


## By age
seurat_plot<- immune_seurat
seurat_plot$plotting<- seurat_plot$Age
DimPlot(seurat_plot, group.by = "plotting", label=FALSE, label.size = 8) %>% umap_theme(colors=c(Ped="red", Adult="navy"))
ggsave(paste0(figure_dir, "byAge.png"), width=6, height=5.5)

## By dataset
seurat_plot<- immune_seurat
seurat_plot$dataset<- ifelse(seurat_plot$broad_annot=="Tcell", paste0("Tcell_", seurat_plot$Age),
                               ifelse(seurat_plot$Age=="Ped", "Myeloid_Ped",
                                      paste0("Myeloid_Adult.", seurat_plot$NiceSubtype)))
seurat_plot$plotting<- seurat_plot$dataset
DimPlot(seurat_plot, group.by = "plotting", label=FALSE, label.size = 8) %>%
  umap_theme(colors=c(Myeloid_Ped="red", Myeloid_Adult.GBM="navy", Myeloid_Adult.IDHmut="cyan",
                      Tcell_Adult="skyblue", Tcell_Ped="firebrick"))
  
ggsave(paste0(figure_dir, "byDataset.png"), width=6, height=5.5)
```


