---
title: "NMF.Rmd"
author: "Jenna LaBelle"
date: "11/11/2021"
output: html_document
---

## Purpose:
Run de novo NMF on merged adult/ped myeloid cells. Compare these programs to ped to annotate. Used to help compare proportions of programs in ped vs adult

## Approach:
Run de novo NMF on merged ped/adult myeloid cells. Correlate with ped programs to help annotate.

```{r}
library(Seurat) 
library(clusterProfiler)
library(biomaRt)
library(stringr)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02b_Myeloid/01b_Ped.Adult.NMF/")


analysis_dir<- paste0(working_dir, "/analysis/Ped.Adult_NMF/")
figure_dir<- paste0(working_dir, "/figures/Ped.Adult_NMF/")

if(!dir.exists(figure_dir)){dir.create(figure_dir)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(preprocessing)
source(nmf_helper)

library(reticulate) 
use_python("/Users/jlabelle/miniconda3/bin/python", required = T)  # specify python in my conda env for scrublet
use_condaenv("base", conda = "/Users/jlabelle/miniconda3/condabin/conda")  # reticulate library to load the conda environnment for running
library(Seurat)

source(paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/", "/Scripts/plotutils.R"))
source(paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/", "/Scripts/seurat3utils.R"))
source(paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/", "/Scripts/color.R"))

tests <- c("wilcox", "t", "roc", "MAST", "bimod")
test <- tests[2]
```

## Read in merged ped/adult seurat object. Merged in Merge_AssessIntegration.Rmd
```{r}
cohort<-"pedOnly_nomuv63"
adult_withMUV<- FALSE

analysis_dir<- paste0(analysis_dir, cohort, "/", "AdultWithMUV.", adult_withMUV, "/")
figure_dir<- paste0(figure_dir, cohort, "/","AdultWithMUV.", adult_withMUV, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

ped_rank<- 6
myeloid_seurat<-readRDS(paste0("../03_CompareToAdult/analysis/Merge_AssessIntegration/",
                               cohort,"/adult_withMUV.FALSE/myeloid_seurat_Ped.AdultGBM.IDHMut_merged.Rds" ))


## load in ped marker genes
ped_genes<- read.csv(paste0("../01a_NMF/analysis/de.novoNMF/", cohort, "/top_DE_genes_10Xannotated_NMF_rank", ped_rank, ".csv"))
ped_list<- split(ped_genes, f=ped_genes$cluster); ped_list<- lapply(ped_list, function(x){x$gene})

## alternative: adult 10x recurrent programs
adult_markers_top100<- readRDS(paste0("../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/", 
                                      "Preprocess_AdultMyeloid/", 
                                      "GBM.10X_recurrent_myeloid_markers_top100.Rds"))
adult_myeloid_markers<- split(adult_markers_top100, f=adult_markers_top100$cluster)
adult_myeloid_markers<- lapply(adult_myeloid_markers, function(x){x$gene})
names(adult_myeloid_markers)<- paste0("Adult_", names(adult_myeloid_markers))

markers_use<- ped_list
analysis_name<- "originalPed"
```

## Clustering using non-negative matrix factorization (NMF)
Cluster data using NMF with a range of ranks. The rank is the number of metagenes and the number of clusters. The choice of the rank is heuristic. For each clustering, we also calculate DE genes.

## This section run on O2
```{r}
# Vector of ranks for which to calculate NMF.
rank <- 5:7

# Identify variable genes within this cell subset.
var.genes <- SelectIntegrationFeatures(SplitObject(myeloid_seurat, split.by = "sampleid"), nfeatures = 2000, verbose = TRUE,
                                       fvf.nfeatures = 2000, selection.method = "vst")

VariableFeatures(myeloid_seurat) <- var.genes

# Number of variable genes
print(length(VariableFeatures(myeloid_seurat)))

# Perform 10 NMF runs for each value of rank r.
x <- as.matrix(myeloid_seurat[['RNA']][VariableFeatures(myeloid_seurat), ])
x <- x[which(rowSums(x) > 0), ]  # remove any rows filled with 0 (just in case these exist)

estim.r <- nmf(x, rank = rank, method = "brunet", nrun = 10, seed = 123456, .options = "tvP")
saveRDS(estim.r, file=paste0(analysis_dir, "estimR_rank5-7.Rds"))
```

## Read in estim.R and finish processing
```{r NMF, eval = F}
estim.r<- readRDS(paste0(analysis_dir, "estimR_rank8-13.Rds"))
rank <- 8:13

png(paste0(figure_dir, "_NMF_plot_rank8-13.png"), width = 2000, height = 1000, res = 100)
plot(estim.r)
dev.off()

# Iterate over different ranks and look at clustering.
for (i in seq(rank)) {
  r <- rank[i]
  
  # Plot coefficient values of metagenes for each cell.
  png(paste0(figure_dir, "_NMF_coefmap_rank", r, ".png"), width = 2000, height = 1000, res = 100)
  coefmap(estim.r$fit[[i]], Colv=T)
  dev.off()

  # Show genes contributing the most to each metagene (factor). 
  # Remember, this is limited to the variable genes if that was the input to NMF.
  for (j in seq(r)) {
    top.features <- names(sort(estim.r$fit[[i]]@fit@W[, j], decreasing = T)[1:10])
    print(paste(top.features, collapse = " "))
  }
  s <- extractFeatures(estim.r$fit[[i]], 10L)
  for (j in seq(r)) {
    top.features <- rownames(estim.r$fit[[i]]@fit@W)[s[[j]]]
    print(paste(top.features, collapse = " "))
  }

  # Plot NMF residuals.
  png(paste0(figure_dir, "_NMF_residuals_rank", r, ".png"), width = 2000, height = 1000, res = 100)
  plot(estim.r$fit[[i]])
  dev.off()

  # Check coefficient matrix has cell names in same order as Seurat object myeloid_seurat.
  H <- estim.r$fit[[i]]@fit@H
  all.equal(as.integer(as.character(predict(estim.r$fit[[i]], what = "samples"))), as.integer(apply(H, 2, which.max)))
  all.equal(colnames(H), colnames(myeloid_seurat))
  all.equal(rownames(estim.r$fit[[i]]@fit@W), VariableFeatures(myeloid_seurat))
  
  # Cluster cells based on maximum coefficient for metagene.
  myeloid_seurat[[paste0("NMF_rank", r)]] <- predict(estim.r$fit[[i]], what = "samples")
  Idents(myeloid_seurat) <- paste0("NMF_rank", r)

  # UMAP visualization and myeloid and myeloid expression across clusters and find DE genes.
  DimPlot(myeloid_seurat, reduction = "umap", cols = material.heat(n_distinct(Idents(myeloid_seurat))), label = T)
  ggsave(paste0(figure_dir,  "UMAP_NMF_rank", r, ".png"), width = 6, height = 6)
  VlnPlot(myeloid_seurat, features = c("CSF1R", "CD3G"), cols = material.heat(n_distinct(Idents(myeloid_seurat))), 
          ncol = 1, pt.size = 0.5) 
  ggsave(paste0(figure_dir, "vln_NMF_rank", r, ".png"), width = 8, height = 8)
  
  # Find differentially expressed genes for each cluster.
  myeloid_seurat.markers <- FindAllMarkers(myeloid_seurat, min.pct = 0.1, logfc.threshold = 0.5, 
                                           max.cells.per.ident = 1500, test.use = test, 
                                           return.thresh = 0.1 / nrow(myeloid_seurat), verbose = T)
  top10genes <- myeloid_seurat.markers %>% filter(avg_log2FC > 0) %>% group_by(cluster) %>% top_n(10, avg_log2FC)
  top100genes <- myeloid_seurat.markers %>% filter(avg_log2FC > 0) %>% group_by(cluster) %>% top_n(100, avg_log2FC)
  
  # Write cluster markers to csv file.
  write.csv(top100genes, file = paste0(analysis_dir,  "top_DE_genes_NMF_rank", r, ".csv"))
  myeloid_seurat <- ScaleData(myeloid_seurat, features = c(top10genes$gene), do.center = T, do.scale = F)
  DoHeatmap(myeloid_seurat, features = top10genes$gene) + scale_fill_distiller(palette = "RdYlBu")
  ggsave(paste0(figure_dir,  "DE_genes_expression_rank", r, ".png"), width = 18, height = 24)
}


saveRDS(myeloid_seurat, file=paste0("../03_CompareToAdult/analysis/Merge_AssessIntegration/",
                               cohort,"/adult_withMUV.FALSE/myeloid_seurat_Ped.AdultGBM.IDHMut_merged.Rds" ))
```

## Correlate with ped programs to help annotate merged  programs
```{r}
## Norm/center current cm
cm_list<- NormCenter(myeloid_seurat@assays$RNA@counts)
cm_mean<- log2(rowMeans(cm_list$raw_data)+1)

## score current (merged) cm for ped programs
ped_scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, markers_use)))

ped.adult_merged_scores<- list()
## for merged rank 5/6, score, then correlate scores with ped program scores
ranks<- 6
for(rank in ranks){
  print(rank)
  
  ## Read in current marker genes- rank 6. Convert to list
  current_markers<- read.csv(paste0(analysis_dir, "top_DE_genes_NMF_rank", rank, ".csv"))
  current_markers<- split(current_markers, f=current_markers$cluster)
  current_markers<- lapply(current_markers, function(x){x$gene})
  names(current_markers)<- paste0("NewNMF_", names(current_markers))
  
  ## Score current cm 
  ped.adult_scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, current_markers)))
  
  ## Merge current ped/adult scores with scores for ped programs
  scores<- cbind(ped.adult_scores, ped_scores)
  ped.adult_merged_scores[[rank]]<- scores
  
  ## Correlate and plot heatmap
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  pheatmap(scores_factor_cor, color = hm_colors(100), 
           cluster_rows = F, cluster_cols = F, 
           annotation_names_row = F, annotation_names_col =T,
           show_rownames = T, show_colnames = F,
           filename = paste0(figure_dir, "Cor_NewPed.AdultNMFrank", rank, "_", analysis_name, ".png"),
           width = 8, height = 4)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir,"HC_NewPed.AdultNMFrank", rank, "_PedPrograms.png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
  
}


saveRDS(ped.adult_merged_scores, file=paste0(analysis_dir, "scores_Ped.AdultMerged_rank8-13.Rds"))
```

## Correlate with ped programs + 10x recurrent programs to help annotate merged  programs
```{r}
## Norm/center current cm
cm_list<- NormCenter(myeloid_seurat@assays$RNA@counts)
cm_mean<- log2(rowMeans(cm_list$raw_data)+1)

## score current (merged) cm for ped programs
ped_scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, markers_use)))

## score current (merged) cm for 10x recurrent programs
gbm10x_scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, adult_myeloid_markers)))

## Combine ped/gbm scores, renamed for clarity
colnames(ped_scores)<- paste0("PedProgram_", colnames(ped_scores))
colnames(gbm10x_scores)<- paste0("GBM10X_", colnames(gbm10x_scores))
ped.gbm10x_scores<- cbind(ped_scores, gbm10x_scores)

## for merged rank 4-7, score, then correlate scores with ped program scores
ped.adult_merged_scores<- list()
ranks<- 8:12
for(rank in ranks){
  print(rank)
  
  ## Read in current marker genes. Convert to list
  current_markers<- read.csv(paste0(analysis_dir, "top_DE_genes_NMF_rank", rank, ".csv"))
  current_markers<- split(current_markers, f=current_markers$cluster)
  current_markers<- lapply(current_markers, function(x){x$gene})
  names(current_markers)<- paste0("NewNMF_", names(current_markers))
  
  ## Score current cm 
  ped.adult_scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, current_markers)))
  
  ## Merge current ped/adult scores with scores for ped programs
  scores<- cbind(ped.adult_scores, ped.gbm10x_scores)
  ped.adult_merged_scores[[rank]]<- scores
  
  ## Correlate and plot heatmap
  scores_factor_hc = clusterNmfFactors(scores)
  scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]
  
  ## Heatmap of correlations
  hm_colors = rev((brewer.pal(n=9, name="RdBu")))
  hm_colors = colorRampPalette(colors = hm_colors)
  pheatmap(scores_factor_cor, color = hm_colors(100), 
           cluster_rows = F, cluster_cols = F, 
           annotation_names_row = F, annotation_names_col =T,
           show_rownames = T, show_colnames = F,
           filename = paste0(figure_dir, "Cor_Ped.GBM10x_NewPed.AdultNMFrank", rank, "_", analysis_name, ".png"),
           width = 8, height = 4)
  
  ## Dendrogram of hierarchical clustering 
  jpeg(filename=paste0(figure_dir,"HC_Ped.GBM10x_NewPed.AdultNMFrank", rank, "_PedPrograms.png"), width = 400, height = 600)
  plot(scores_factor_hc$hc_obj)
  dev.off()
  
}


#saveRDS(ped.adult_merged_scores, file=paste0(analysis_dir, "scores_Ped.AdultMerged_rank5-7.Rds"))
```



## For pedOnly_nomuv63: Potential adult-specific program NMF3: annotation attempt
```{r}
rank<- 6
markers<- read.csv(paste0(analysis_dir, "top_DE_genes_NMF_rank", rank, ".csv"))

## Heatmap
nmf3_markers<- markers[markers$cluster==3,]; nmf3_markers<- nmf3_markers[order(nmf3_markers$avg_log2FC, decreasing = TRUE),]
DoHeatmap(myeloid_seurat, features=nmf3_markers$gene[1:25], group.by = paste0("NMF_rank", rank))+
  scale_fill_gradient2(low="blue", mid="white", high="red")
ggsave(paste0(figure_dir, "Heatmap_rank", rank, "NMF2_AdultSpecific.png"), width=4, height=3)

## GO term ORA
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Read in df for converting symbol --> ensembl
all_genes_Ens.Sym<- readRDS(paste0("/Users/jlabelle/Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Marker_genes/", 
                                   "allGenes_Ensembl.Symbol.Rds"))

## Convert input genes to ensembl. Not all genes found in all_genes_Ens.Sym bc not all Symbols match Ensembl
## These genes removed
allOE_genes<- rownames(myeloid_seurat@assays$RNA@counts)[rownames(myeloid_seurat@assays$RNA@counts) %in%
                                                       all_genes_Ens.Sym$hgnc_symbol]
allOE_genes<- plyr::mapvalues(allOE_genes,all_genes_Ens.Sym$hgnc_symbol, all_genes_Ens.Sym$ensembl_gene_id, warn_missing = FALSE)
  
## Run GO: Convert marker genes to ensembl IDs too, then run OR analysis (GO)
go_result = list()
for (i in unique(markers$cluster)){
  print(i)
  genes<- markers[markers$cluster==i, "gene"]
  if(length(genes)!=1){
    sigOE_genes = genes[genes %in% all_genes_Ens.Sym$hgnc_symbol]
    sigOE_genes<- plyr::mapvalues(sigOE_genes, all_genes_Ens.Sym$hgnc_symbol, 
                                  all_genes_Ens.Sym$ensembl_gene_id, warn_missing = FALSE)
    go_result[[i]] = go_analysis(sigOE_genes, allOE_genes)
    }
} 
names(go_result)<- paste0("NMF", 1:length(go_result))

## Plot and save
go_plots<- plotGO(go_result = go_result, n_terms = 15)
cowplot::plot_grid(plotlist=go_plots, ncol=3)
ggsave(paste0(figure_dir, "GO_NMFrank", rank, "_NMF6.png"), width=24, height=12)
```
## Density plots- comparing to mg/mo/monocyte. To help annotate NMF3
## In progress
```{r}
colors_use<- c(Adult_Monocytes="turquoise3", Adult_Mg_TAM="blue", Adult_Mo_TAM="red", Adult_prol_TAM="gold3")

## Read in mg/mo/monocyte markers
adult_markers_top100<- readRDS(paste0("../../01_Preprocessing/03_PreprocessAdultDatasets/", 
                                      "analysis/Preprocess_AdultMyeloid/", 
                                      "GBM.10X_recurrent_mg.mo.monocyte.prol_markers_top100.Rds"))
adult_myeloid_markers<- split(adult_markers_top100, f=adult_markers_top100$cluster)
adult_myeloid_markers<- lapply(adult_myeloid_markers, function(x){x$gene})
names(adult_myeloid_markers)<- paste0("Adult_", names(adult_myeloid_markers))

## Norm/center current cm
cm_list<- NormCenter(myeloid_seurat@assays$RNA@counts)
cm_mean<- log2(rowMeans(cm_list$raw_data)+1)

## score current (merged) cm for programs, add max to seurat object
scores<- as.data.frame(t(scoreNmfGenes(cm_list$center_data, cm_mean, markers_use)))
scores$MaxProgram<- apply(scores, 1, function(x){names(x)[which.max(x)]})


## Create df for plotting
myeloid_df<-myeloid_seurat@meta.data[,c("sampleid", "MgMoProlMonocyte_annot", paste0("NMF_rank", rank_use, "_annotByAdult10X"))]
sum(rownames(scores)==rownames(myeloid_df)); nrow(scores); nrow(myeloid_df)
myeloid_df<- cbind(myeloid_df, scores[,!grepl("Max", colnames(scores))] )


myeloid_df$MgMoProlMonocyte_annot<- factor(myeloid_df$MgMoProlMonocyte_annot, 
                                          levels=c("Adult_prol_TAM", "Adult_Monocytes", "Adult_Mg_TAM", "Adult_Mo_TAM"))

## Plot each nmf program
all_plots<- list()
programs<- as.character(unique(myeloid_seurat@meta.data[[paste0("NMF_rank", rank_use, "_annotByAdult10X")]]))
for(i in programs){
  df<- myeloid_df[,c(as.character(i), "MgMoProlMonocyte_annot")]
  colnames(df)<- c("score", "CellType")
  p<-ggplot(df, aes(x=score, fill=CellType))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.position = "none",
          plot.title = element_text(size=20, face="bold", hjust=0.5))

  all_plots[[i]]<-p
}

## add legend separately
legend<-ggplot(df, aes(x=score, fill=CellType))+
    geom_density(alpha=0.6)+
    ggtitle(i)+
    scale_fill_manual(values=colors_use)+
    theme_classic()+
    theme(legend.title = element_blank(),
          legend.key.size = unit(1.5, 'cm'),
          legend.text = element_text(size=20))
all_plots$legend<-  as_ggplot(get_legend(legend))
all_plots<- all_plots[c(1:3,8,4:7)]
cowplot::plot_grid(plotlist = all_plots, ncol=4)

ggsave(paste0(figure_dir, "DensityOfNMFByMgMoProlMonocytes_rank", rank_use,".png"), width=20, height=8)

```


## Gene list options- for scoring ped/adult, comparing proportions

## rank6- seems to be the best for now, may change downstream
```{r}
rank<- 6
markers<- read.csv(paste0(analysis_dir, "top_DE_genes_NMF_rank", rank, ".csv"))
marker_list<- split(markers, f=markers$cluster); marker_list<- lapply(marker_list, function(x){x$gene})

## pedOnly_nomuv63
  nmf_programs<- names(marker_list)
  names(nmf_programs)<-c("Monocytes",  "Hypoxic_TAM", "NMF3", "Stress_Response","IFN_TAM", "Inflammatory")
  option1_markers<- marker_list
  names(option1_markers)<- plyr::mapvalues(names(option1_markers), nmf_programs, names(nmf_programs))
  
  markers$cluster<-  plyr::mapvalues(markers$cluster, nmf_programs, names(nmf_programs))
  top50<- markers %>% group_by(cluster) %>% top_n(n=50, wt=avg_log2FC) %>% as.data.frame()
  
  options_markers<- list(KeepAll=option1_markers)
  write.csv(top50, file=paste0(analysis_dir, "myeloidProgram_mergedAdult.ped_markers_rank6_top50.csv"))
  saveRDS(options_markers, file=paste0(analysis_dir, "Adult.Ped_Markers_option1.Rds"))

  ## pedOnly_nomuv63.withmuv91
  ## 1: Rename NMF2 to "Phago_IFN", use all programs as-is
  nmf_programs<- names(marker_list)
  names(nmf_programs)<-c("Inflammatory", "Phago_IFN", "Stress_Response", "Mo_TAM", "Monocytes", "NMF6")
  option1_markers<- marker_list
  names(option1_markers)<- plyr::mapvalues(names(option1_markers), nmf_programs, names(nmf_programs))
  
  ## 2: Remove NMF2, add in full Ped IFN and ped Phago programs
  option2_markers<- option1_markers
  option2_markers$Phago_IFN<-NULL
  option2_markers$Phago_Lipid<- ped_list$Phago_Lipid_Mg_TAM
  option2_markers$IFN<- ped_list$IFN_Mg_TAM
  
  ## 3: Remove NMF2, add in Ped IFN and phago programs, but ONLY for genes that are also in NMF2
  option3_markers<- option2_markers
  option3_markers$Phago_Lipid<- option3_markers$Phago_Lipid[option3_markers$Phago_Lipid %in% marker_list$`2`] ## 15 genes
  option3_markers$IFN<- option3_markers$IFN[option3_markers$IFN %in% marker_list$`2`] ## 18 genes
  
  options_markers<- list(KeepNMF2=option1_markers,
                         FullPhagoIFN=option2_markers,
                         SubsetPhagoIFN=option3_markers)
  saveRDS(options_markers, file=paste0(analysis_dir, "Adult.Ped_Markers_options1-3.Rds"))
```

