---
title: "AdultGBM10X_AnnotatePed.Rmd"
output: html_document
date: "11/11/2021"
---

## Purpose:
Use adult GBM 10X dataset to help annotate merged Ped/Adult Myeloid NMF programs

## Approach:
Score ped and adult myeloid cells for both adult GBM 10x programs + ped/adult merged NMF programs, correlate to help annotate ped/adult merged NMF programs


## Settings/input as of 4/4/22:
integration=""
cohort=pedOnly_nomuv63
rank=6

```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/02b_Myeloid/01b_Ped.Adult.NMF/")
data_dir_tmp<- "../../01_Preprocessing/02_Counts_Preprocess/analysis/detailed_annotation/"


preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)

integration<- "" ## HarmonyInt/ or ""
figure_dir_tmp<- paste0("figures/GBM.10X_AnnotatePedAdult_recurrent/", integration)
analysis_dir_tmp<- paste0("analysis/GBM.10X_AnnotatePedAdult_recurrent/", integration)

for(i in c(figure_dir_tmp, analysis_dir_tmp)){
  if(!dir.exists(i)){dir.create(i, recursive = TRUE)}
}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(nmf_helper)
```

## Read in merged ped/adult seurat object. Merged in Merge_AssessIntegration.Rmd
```{r}
cohort<-"pedOnly_nomuv63"
adult_withMUV<- FALSE

analysis_dir<- paste0(analysis_dir_tmp, cohort, "/", "AdultWithMUV.", adult_withMUV, "/")
figure_dir<- paste0(figure_dir_tmp, cohort, "/","AdultWithMUV.", adult_withMUV, "/")
if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)}
if(!dir.exists(analysis_dir)){dir.create(analysis_dir, recursive = TRUE)}

myeloid_seurat<-readRDS(paste0("../03_CompareToAdult/analysis/Merge_AssessIntegration/",
                               cohort,"/adult_withMUV.FALSE/myeloid_seurat_Ped.AdultGBM.IDHMut_merged.Rds" ))

## Center/mean ped.adult myeloid
pedAdult_cm_list<- NormCenter(myeloid_seurat@assays$RNA@counts)
pedAdult_cm_mean<- rowMeans(log2(myeloid_seurat@assays$RNA@counts + 1))
```

## load in marker genes
```{r}
rank_use<- 11

## load in ped/adult marker genes
markers<- read.csv(paste0("analysis/Ped.Adult_NMF/", cohort, "/", "AdultWithMUV.", adult_withMUV, "/", "top_DE_genes_NMF_rank", rank_use, ".csv"))
marker_list<- split(markers, f=markers$cluster)
marker_list<- lapply(marker_list, function(x){x$gene})

## adult 10x recurrent programs
adult_markers_top100<- readRDS(paste0("../../01_Preprocessing/03_PreprocessAdultDatasets/analysis/", 
                                      "Preprocess_AdultMyeloid/", 
                                      "GBM.10X_recurrent_myeloid_markers_top100.Rds"))
adult_myeloid_markers<- split(adult_markers_top100, f=adult_markers_top100$cluster)
adult_myeloid_markers<- lapply(adult_myeloid_markers, function(x){x$gene})
names(adult_myeloid_markers)<- paste0("Adult_", names(adult_myeloid_markers))

markers_use<- marker_list
analysis_name<- "mergedPedAdult"


```





## score for both pedAdultMerged markers /adult 10x
## Correlate
```{r}
## score
scores<- as.data.frame(t(scoreNmfGenes(pedAdult_cm_list$center_data, pedAdult_cm_mean,  
                                                   c(adult_myeloid_markers, markers_use))))


## NMF factor correlation based on cell score 
scores_factor_hc = clusterNmfFactors(scores)
scores_factor_cor = scores_factor_hc$cor_coef[scores_factor_hc$hc_obj$order, scores_factor_hc$hc_obj$order]

## Heatmap of correlations
hm_colors = rev((brewer.pal(n=9, name="RdBu")))
hm_colors = colorRampPalette(colors = hm_colors)
pheatmap(scores_factor_cor, color = hm_colors(100), 
         cluster_rows = F, cluster_cols = F, 
         annotation_names_row = F, annotation_names_col =T,
         show_rownames = T, show_colnames = F,
         filename = paste0(figure_dir, "Cor_PedAdultNMFrank", rank_use, "_AdultCellTypes.png"),
         width = 8, height = 4)

## Dendrogram of hierarchical clustering 
jpeg(filename=paste0(figure_dir, "HC_PedAdultNMFrank", rank_use, "_AdultCellTypes.png"), width = 400, height = 600)
plot(scores_factor_hc$hc_obj)
dev.off()

## Save scores
saveRDS(scores, file=paste0(analysis_dir, "scores_pedAdultNMFrank", rank_use, "_adult10x.Rds"))
```


## Plot max 10X program
```{r}
## Subset to just adult programs, assign max program
rank<-10
scores<- readRDS(paste0(analysis_dir, "scores_pedAdultNMFrank", rank_use, "_adult10x.Rds"))
scores<- scores[,grepl("Adult_", colnames(scores))]
scores$MaxProgram<- apply(scores, 1, function(x){names(x)[which.max(x)]})
scores$MaxScore<- apply(scores[,colnames(scores) != "MaxProgram"], 1, function(x){x[which.max(x)]})

## add to seurat object
sum(rownames(scores)==colnames(myeloid_seurat)); nrow(scores); ncol(myeloid_seurat)
myeloid_seurat$Max10XProgram<- scores$MaxProgram; myeloid_seurat$Max10XScore<-scores$MaxScore

## Plot, highlighting prol_tam
myeloid_seurat$prol_tam<- ifelse(myeloid_seurat$Max10XProgram=="Adult_prol_TAM",
                                 ifelse(myeloid_seurat$Max10XScore>0.5, "highScore_prolTAM", "lowScore_prolTAM"), "Other")
p1<- DimPlot(myeloid_seurat, group.by="Max10XProgram")
p2<- DimPlot(myeloid_seurat, group.by = "prol_tam", cols = c(highScore_prolTAM="red", lowScore_prolTAM="blue", Other="grey"))
p1+p2
ggsave(paste0(figure_dir, "Max10XProgram_ProlTAM.png"), width=12, height=5.5)

myeloid_seurat$tmp<- ifelse(myeloid_seurat$prol_tam=="highScore_prolTAM", "prol_TAM", "Other")
DimPlot(myeloid_seurat, group.by="tmp", cols=c(prol_TAM="red", Other="grey"), label=TRUE, label.size = 6)+
  ggtitle("")+NoLegend()
ggsave(paste0(figure_dir, "UMAP_prolTAM.png"), width=6, height=5.5)

## Seems to be prol_tam cluster-may use these annotations downstream. save myeloid seurat object with this distinction.
saveRDS(myeloid_seurat, paste0(data_dir, "myeloid_seurat_noDC.B.Rds"))
```


## Rename ped gene lists/seurat annotations based on correlation
```{r}
annot_rank_list<- list(
  rank6=c("Monocytes", "Mo_TAM", "NMF3", "Mg_TAM", "IFN_TAM", "Inflammatory"),
  rank7=c("NMF1", "Monocytes", "prol_TAM", "Mo_TAM", "IFN_TAM", "NMF6", "Mg_TAM"),
  rank10=c("Monocytes", "IFN_TAM", "SEPP1_TAM", "MgTAM.4", "NMF5", "Mo_TAM", "prol_TAM", "Transitory_Mo_TAM", "MgTAM.9", "MgTAM.10")
)
annot_rank_list<- lapply(annot_rank_list, function(x){
  names(x)<- paste0("NMF", 1:length(x))
  return(x)
})

  

saveRDS(annot_rank_list, file=paste0(analysis_dir, "PedAdultMergedNMF_RenamedByGBM10XCorrelation.Rds"))

```

## After determining annotations based on correlation with adult 10x programs, rename NMF programs here
## Changed to using recurrent programs instead of primary on 1/24/22
```{r}
rank_use<-10
ND.R<- "_recurrent" ## "_recurrent" to use annotations based on correlation with recurrent 10X programs, "" to use primary tumor programs
cohort<- "pedOnly_nomuv63"

## read in scores + markers  to be renamed
scores<- readRDS(paste0(analysis_dir, "scores_pedAdultNMFrank", rank_use, "_adult10x.Rds"))
markers<- read.csv(paste0("analysis/Ped.Adult_NMF/", cohort, "/", "AdultWithMUV.", adult_withMUV, "/", "top_DE_genes_NMF_rank", rank_use, ".csv"))

## Set annotations to use (above)
annots<- annot_rank_list[[paste0("rank", rank_use)]]

  
## Reassign nmf names based on 10x annotations
  ## marker genes
  markers$cluster<- plyr::mapvalues(paste0("NMF", markers$cluster),  names(annots) ,annots)

  ## seurat
  myeloid_seurat@meta.data[[paste0("NMF_rank", rank_use, "_annotByAdult10X")]]<- plyr::mapvalues(myeloid_seurat@meta.data[[paste0("NMF_rank", rank_use)]],
                                                                                          gsub("NMF", "", names(annots)) ,annots)
  myeloid_seurat@meta.data[[paste0("NMF_rank", rank_use, "byScore_annotByAdult10X")]]<- plyr::mapvalues(myeloid_seurat@meta.data[[paste0("NMF_rank", rank_use)]],
                                                                                     gsub("NMF", "", names(annots)) ,annots)
  
  ## scores
  scores<- scores[,!grepl("Adult_", colnames(scores))]
  colnames(scores)<- plyr::mapvalues(colnames(scores), gsub("NMF", "", names(annots)) ,annots)
  scores$MaxProgram<- apply(scores, 1, function(x){names(x)[which.max(x)]})
  scores$MaxScore<- apply(scores[,colnames(scores) != "MaxProgram"], 1, function(x){x[which.max(x)]})
  
  ## Save marker genes with annotations
  write.csv(markers, file=paste0(analysis_dir,  "top_DE_genes_10Xannotated_NMF_rank", rank_use, ".csv"))


## save seurat with 10x annotations
saveRDS(myeloid_seurat, file=paste0("../03_CompareToAdult/analysis/Merge_AssessIntegration/",
                               cohort,"/adult_withMUV.FALSE/myeloid_seurat_Ped.AdultGBM.IDHMut_merged.Rds" ))

## save scores with 10x annotations
saveRDS(scores, file=paste0(analysis_dir, "NMF_scores_annotByAdult10X_rank", rank_use, ".Rds"))


```

## Save colors for all follow up
```{r}
## pedOnly_nomuv63- annotated with 10X primary
color_annots<- list(rank6=c(SEPP1_Mo_TAM ="orange", Inflammatory="red", IFN_Mg_TAM="purple", Stress_Response="olivedrab4",
                            Monocytes="turquoise3", Mo_TAM="navy"))

## pedOnly_novmu63- annotated with 10X recurrent
color_annots<- list(rank6=c(SEPP1_Mo_TAM ="orange", Inflammatory="red", IFN_TAM="purple", Stress_Response="olivedrab4",
                            Monocytes="turquoise3", Hypoxic_TAM="navy"))

## pedOnly_novmu63, with harmony integration- annotated with 10X recurrent
color_annots<- list(rank7=c(SEPP1_Hi_Mo_TAM ="orange", Inflammatory="red", IFN_TAM="purple", Stress_Response="olivedrab4",
                            Monocytes="turquoise3", Hypoxic_TAM="navy",prol_TAM="grey22" ))

saveRDS(color_annots, paste0(analysis_dir, "annot_colors.Rds"))
```

## Exploratory- can delete
```{r}
df<- as.data.frame(table(myeloid_seurat$NiceSubtype, myeloid_seurat$NMF_rank10_annotByAdult10X))
colnames(df)<- c("Age", "Program", "Ncells")
df<- df %>% group_by(Age) %>% mutate(perCells=Ncells/sum(Ncells)) %>% as.data.frame()

ggplot(df, aes(x=Age, y=perCells, fill=Program))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

