---
title: "inferCNV_clustering"
output: html_document
---

Cluster tumor data and integrate with inferCNV results. Use combination of inferCNV and clustering with normal cells to assign final malignant/normal cell annotations

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and functions
```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/Immune/PD1_Treat/01_Preprocessing/")

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)

inferCNV_analysis_folder = "analysis/inferCNV/seurat/"
inferCNV_fig_folder = "figures/inferCNV/seurat/"
for (folder in c(inferCNV_analysis_folder, inferCNV_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}
```

## Load data
```{r load-data}
## Load cm, samples, and qc data
cm = readRDS("analysis/inferCNV/cm_exp_ctrl.rds")
samples = readRDS("analysis/inferCNV/samples_exp_ctrl.rds")
any(colnames(cm) != rownames(samples))
qc_filtered_data<- readRDS("analysis/qc_filtered_data.rds")

## If seurat_obj already generated 
seurat_filename = paste0(inferCNV_analysis_folder,
                         "seurat_obj.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}

## Load cnv status inferred by inferCNV
cnv_stat = readRDS(paste0("analysis/inferCNV/HC/", "call_cnv_w_ctrl.rds"))
obs_ref = readRDS(paste0("analysis/inferCNV/", "obs_ref.rds"))
hc_res<- readRDS(paste0("analysis/inferCNV/HC/", "cnv_hc_res.rds"))

```

## Preprocessing with Seurat
```{r seurat_clustering}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", min.cells=0, min.genes=0,scale.factor=1E5, do.scale=F, do.center=T)

## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples
seurat_obj@meta.data$type = obs_ref
seurat_obj@meta.data$cnv = cnv_stat

## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 50,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)

## PC selection
seurat_obj <- JackStraw(
  object = seurat_obj, 
  dims = 50,
  num.replicate = 100, 
  verbose = TRUE
)
seurat_obj<- ScoreJackStraw(seurat_obj, dims=1:50)
JackStrawPlot(seurat_obj, dims=1:50); ggsave(filename=paste0(inferCNV_fig_folder, "Jackstraw.jpg"),
                                             width=12, height=8)
ElbowPlot(object = seurat_obj, ndims = 50); ggsave(filename=paste0(inferCNV_fig_folder, "Jackstraw_ElbowPlot.jpg"),width=12, height=8)

## Clustering
numPC = 20
resolution =1
seurat_obj <- FindNeighbors(object=seurat_obj,
                            dims=1:numPC,
                            reduction="pca")
seurat_obj<- FindClusters(object=seurat_obj, 
                          resolution=resolution)


## UMAP
table(seurat_obj@active.ident)
seurat_obj <- RunUMAP(
  object = seurat_obj, 
  dims= 1:numPC, 
  do.fast = TRUE
)

save(seurat_obj, file = paste0(inferCNV_analysis_folder, "seurat_obj.Robj"))
```

## UMAP plots
```{r tsne_plots}
## Colored by clustering
DimPlot(object = seurat_obj, label = T, pt.size = 2, label.size = 8) + NoLegend()
ggsave("UMAP_clustering.jpg", path=inferCNV_fig_folder, width = 8, height = 6, dpi = 300)

## Colored by samples
DimPlot(object = seurat_obj, label = T, pt.size = 2, label.size = 8, group.by = "sample") + NoLegend()
ggsave("UMAP_sample.jpg", path=inferCNV_fig_folder, width = 8, height = 6, dpi = 300)

## Colored by obs vs ref
DimPlot(object = seurat_obj, label = T, pt.size = 2, label.size = 8, group.by = "type") + NoLegend()
ggsave("UMAP_type.jpg", path=inferCNV_fig_folder, width = 8, height = 5, dpi = 300)

## Colored by CNV status
DimPlot(object = seurat_obj, label = T, pt.size = 1, label.size = 8, group.by = "cnv") + NoLegend()
ggsave("UMAP_cnv.jpg", path=inferCNV_fig_folder, width = 8, height = 6, dpi = 300)
```

## Add HC cluster # results, plot UMAP
```{r}
ref_cells<- names(obs_ref)[obs_ref=="ref"]
ref_hc_res<- rep("Ref", length(ref_cells)); names(ref_hc_res)<- ref_cells
hc_res<- c(hc_res, ref_hc_res)
hc_res<- hc_res[order(match(names(hc_res),colnames(seurat_obj)))]
sum(names(hc_res)==colnames(seurat_obj)); ncol(seurat_obj); length(hc_res)

seurat_obj$hc_res<- hc_res

## Subset to 1 sample + ref, plot its clusters
plot_list<-list()
for (sample_use in c("BT1910", "BT1935")){
  seurat_tmp<- subset(seurat_obj, sample==sample_use | type=="ref")
  colors_use<- c(brewer.pal(6, "Set1"), "grey"); names(colors_use)<- c(1:6, "Ref")
  p<- DimPlot(seurat_tmp, group.by = "hc_res", pt.size = 1)+
    scale_color_manual(values=colors_use)+
    ggtitle(sample_use)
  plot_list[[sample_use]]<- p
}
cowplot::plot_grid(plotlist=plot_list)
ggsave(paste0(inferCNV_fig_folder, "UMAP_HC_cluster.png"), width=10, height=5)
```

## DEGs
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05, ]
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC)%>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(inferCNV_analysis_folder, "all_markers.Robj"))

## Heatmap
DoHeatmap(seurat_obj, 
          features=all_markers_filtered$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs.jpg", path=inferCNV_fig_folder)
```

## Expresssions of common normal cell marker
```{r deg_for_normal_cells}
FeaturePlot(seurat_obj, 
            features = c("MBP", "PLP1", "TF", ## Oligodendrocyte
                         "OLIG1", "PDGFRA", "APOD", ## OPC
                         "CTGF", "IFITM1", "MGP", ## Endothelial cells,
                         "CD14", "CD3E", "CD4" ## Immune cells
                         ), #CD45 marker for all immune cells-distinguish neural from immune
            cols = c("lightgrey", "red"), ncol=3,
            slot="data")
ggsave("normal_marker_genes_umap_data.jpg", path=inferCNV_fig_folder, width=9, height=6)
```

## Map gene counts to UMAP
```{r}
FeaturePlot(seurat_obj, "nFeature_RNA", cols = c("lightgrey", "purple"))
ggsave("nFeature_RNA_UMAP.jpg", path=inferCNV_fig_folder, width=9, height=12)
```

## Map mito counts to UMAp
```{r}
mito.genes <- grep(pattern = "^MT-", x = rownames(seurat_obj@assays$RNA@data), 
                   value = TRUE)
percent.mito <- Matrix::colSums(seurat_obj@assays$RNA@counts[mito.genes,])/
  Matrix::colSums(seurat_obj@assays$RNA@counts)
seurat_obj <- AddMetaData(seurat_obj, 
                          metadata = percent.mito, 
                          col.name = "percent_mito")
FeaturePlot(seurat_obj, "percent_mito", cols = c("lightgrey", "red"))
ggsave("mito_genes.jpg", path=inferCNV_fig_folder, width=6, height=6)
```

## Map alignment rate onto tSNE
```{r}
qc_raw_data = readRDS("analysis/qc_raw_data.rds")
seurat_obj@meta.data$alignment_rate = qc_raw_data[colnames(seurat_obj), 
                                                  "align_rate"]
tmp = is.na(seurat_obj@meta.data$alignment_rate)
seurat_obj@meta.data$alignment_rate[tmp] = 0.5
FeaturePlot(seurat_obj, "alignment_rate", cols = c("lightgrey", "red"))
ggsave("align_rate.jpg", path=inferCNV_fig_folder, width=6, height=6)
```

## OPTIONAL: Additional QC on CNV
```{r}
## cluster vs cnv: which cluster has the highest % of cells without CNV 
tmp = table(seurat_obj@meta.data$seurat_clusters, seurat_obj@meta.data$cnv)
tmp2 = round(tmp/rowSums(tmp),2)
tmp2[order(tmp2[,1], decreasing = T),]

## sample vs clustering: which sample belongs to normal clusters
normal_clust = c( "0", "1", "3", "4", "6")
tmp3 = table(seurat_obj@meta.data$sample, seurat_obj@meta.data$seurat_cluster)
tmp4 = round(tmp3/rowSums(tmp3),2)
tmp4[,normal_clust]
rowSums(tmp4[,normal_clust])

## sample vs cnv: which sample has the highest % fo CNV negative cells 
tmp5 = table(seurat_obj@meta.data$sample, seurat_obj@meta.data$cnv)
tmp6 = round(tmp5/rowSums(tmp5),2)

## Compare % of cells for each sample in normal clusters vs CNV negative cells
rowSums(tmp4[,normal_clust])
tmp6[,1]
```

## Check intrapatient batch effect
```{r}
## subset seurat object to just obs samples, then add plate metadata
seurat_onlyObs<-subset(seurat_obj, type=="obs" )
plate<-qc_filtered_data$plate; names(plate)<- rownames(qc_filtered_data)
seurat_onlyObs@meta.data$plate<- plate
samples_plot<- c("BT1910", "BT1935")

## Iteratively plot the spread of cells across all plates for each sample individually
plots<-list()
for (sample in samples_plot){
  plots[[sample]]<-DimPlot(seurat_onlyObs, group.by="plate", reduction="umap",
         cells = colnames(seurat_onlyObs)[seurat_onlyObs@meta.data$sample == sample], 
         label.size = 8, pt.size = 2, label = T)
}

## Save each plot
for (i in 1:length(plots)){
  filename=paste0(inferCNV_fig_folder, "IntrpatientBatchEffect_", unique(samples)[i], ".jpg")
  jpeg(filename,width=600,height=600)
  print(plots[[i]])
  dev.off()
}
```

## Malignant vs non-malignant cell calling 
```{r call_tumor}
#############Strategy###############
## All cells in normal clusters and without CNV = Normal
## All cells in samples without cnv are Malignant
## All cells NOT in normal clusters and with CNV = Malignant 
## The rest = Low-quality 
####################################
normal_clusters = normal_clust
samples_without_cnv = c()
call_tumor <- function(x){
  ## cells in normal cluster are normal cells 
  if (seurat_obj@meta.data$seurat_cluster[x] %in% normal_clusters &
      !seurat_obj@meta.data$cnv[x]){
    return("Normal")
  }else if (seurat_obj@meta.data$sample[x] %in% samples_without_cnv){
    return("Malignant")
  ## All others classified based on having CNV or not 
  }else if (!(seurat_obj@meta.data$seurat_cluster[x] %in% normal_clusters) &
            seurat_obj@meta.data$cnv[x]){
     return("Malignant") 
  }else{
    return("Noise")
  }
}
tumor_stat_w_ctrl = sapply(seq(length(colnames(seurat_obj))), call_tumor)
names(tumor_stat_w_ctrl) = colnames(seurat_obj)
any(names(cnv_stat) != names(tumor_stat_w_ctrl))

## Manually assign cell(s) to malig
manual_rename<- seurat_obj@meta.data[seurat_obj$sample=="BT1910" &
                                       seurat_obj$seurat_clusters=="5" &
                                       seurat_obj$cnv=="FALSE",]
tumor_stat_w_ctrl[rownames(manual_rename)]<- "Malignant"

## Remove normal controls 
tumor_stat = tumor_stat_w_ctrl[(seurat_obj@meta.data$sample != "mg" & 
                                  seurat_obj@meta.data$sample != "od")]
saveRDS(tumor_stat, file = paste0(inferCNV_analysis_folder, "call_tumor.rds"))

## UMAP colored by malignancy
seurat_obj@meta.data$tumor = tumor_stat_w_ctrl
DimPlot(object = seurat_obj, label = F, pt.size = 1, 
        label.size = 8, group.by = "tumor", reduction="umap",
        cols = c("lightgrey", gg_color_hue(2))) + NoAxes()

ggsave("UMAP_tumor.jpg", path=inferCNV_fig_folder, width = 8, height = 6, dpi = 300)
save(seurat_obj, file = "analysis/inferCNV/seurat_obj.Robj")
```


## UMAP with only tumor samples and high quality cells
```{r}
## Generate df for only tumor samples
df = cbind.data.frame(seurat_obj@reductions$umap@cell.embeddings,
                      sample=as.character(seurat_obj@meta.data$sample),
                      annotation=as.character(seurat_obj@meta.data$annotation))

df = df[seurat_obj@meta.data$type == "obs", ]

## Remove low quality cells
df = df[df$annotation != "Noise",]
df$annotation = ifelse(df$annotation == "Malignant", df$sample, df$annotation)

mapped_colors = c("BT1478"= "blue4", "BT1516" = "cyan", 
                  "MUV065" = "green4","MUV066"="purple", "MUV084" = "violetred",
                  "ImmuneCell"= "grey1")

ggplot(df, aes(x=UMAP_1, y=UMAP_2, color=annotation)) + 
    geom_point(size=4, alpha=0.5) +
    ggtitle("") + xlab("UMAP-1") + ylab("UMAP-2") + labs(color="") +
    scale_color_manual(values=mapped_colors) +
    theme(axis.title = element_text(face="bold", size=24),
          axis.text.x = element_text(size=24),
          axis.text.y = element_text(size=24),
          legend.text = element_text(size=16))


ggsave("UMAP_annotation_final.png", path = inferCNV_fig_folder, width = 10, height = 8)
```

## OPTIONAL: Annotation based on clustering only
```{r}
clustering = seurat_obj@meta.data$seurat_cluster
malig_status = ifelse(clustering %in% 5:7, F, T)
names(malig_status) = colnames(seurat_obj)
malig_status = malig_status[seurat_obj@meta.data$type == "obs"]
saveRDS(malig_status, paste0(inferCNV_analysis_folder, "call_tumor_clust_only.rds"))
```

## Plot proportion of normal/noisy/malig cells
```{r}
df<- as.data.frame(tumor_stat)
df$Sample<- plyr::mapvalues(rownames(df), names(samples),samples, warn_missing = FALSE)
df_bd<- as.data.frame(table(df$tumor_stat, df$Sample))
colnames(df_bd)<- c("CellType", "Sample", "NCells")
df_bd$CellType<- gsub("Normal", "Normal: OC", df_bd$CellType)
df_bd<- df_bd %>% group_by(Sample) %>% mutate(CellPer=NCells/sum(NCells)) %>% as.data.frame()

ggplot(df_bd, aes(x=Sample, y=CellPer, fill=CellType))+
  geom_bar(position="stack", stat="identity")+
  xlab("")+ylab("Cell Percentage")+
  theme_classic()+
  theme(axis.text.x = element_text(angle=45, hjust=1, face="bold", color="black"))
ggsave(paste0(inferCNV_fig_folder, "ProporitionOfNormalCells.png"), width=4, height=4)
```
