---
title: "R Notebook"
output: html_document
---

Following cellchat run separately for each pairwise interaction of interest (run on O2), read in results and generate single dataframe for visualization

Note that, originally, this analysis was run on O2 as a single batch for all comparisons. It was later updated so that a separate batch is submitted for each comparison of interest (to reduce run time). For program-wise analysis, both methods were used, but the "final" analysis used is the approach using a separate batch for each comparison of interest- processing of both approaches is shown here for completeness, but the approach using a "single batch" is moved to ARCHIVE. For program-wise analysis downsampled, only the "separate batch" method is used. For celltype-wise analysis, only the "single batch" approach is used.

```{r}
library(CellChat)
library(NMF)
library(circlize)
library(patchwork)
library(Seurat) 
library(ggh4x)

options(stringsAsFactors = FALSE)

base_dir = "/Users/jlabelle/"
script_dir = paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/")
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/04a_Integrate_Immune.Tumor/CellChat/")
analysis_dir<- paste0(working_dir, "/analysis/")
figure_dir<- paste0(working_dir, "/figures/")
data_dir<- "fromO2/"

cellchatHelper<- paste0(script_dir, "CellChat_HelperFunctions.R")
source(cellchatHelper)


## Load in list of all cell types across all subtypes
nCellTypes<- readRDS(paste0(analysis_dir, "nCellTypes_bySample.Rds"))

library(RColorBrewer)
myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))

## Load in subtype info
subtypes<- read.csv("../../../notUploaded/Subtypes.csv")

## read in cell type info
#cellTypes_bySubtype<- readRDS("../../Preprocessing/analysis/WithNewSamples/cellTypes_bySubtype.rds")
```

## By CellType (CD4, CD8, tumor cell types, myeloid)
## Load in each analysis, extract results, and merge
"single batch" approach
```{r}
## Settings for extracting results, converting to simple df
PathwaysRemove<-c("HLA", "COL", "LAMININ", "MHC-I")
sigThreshold<- 1
senderCells<- unique(nCellTypes$CellType)
receiverCells<- unique(nCellTypes$CellType)

## List of results
files<- list.files(paste0(data_dir, "byCellType/")) ## adding "byProgram"- other cellChat analyses already processed
files<- files[!grepl("byProgram", files)]

## Ran myeloid --> tcell and tcell --> myeloid- this produces same results
## Remove 1 here
files<- files[files!="myeloid_tcell"]

## Cycle through each output,
all_output<- lapply(files, function(x){
  print(x)
  cc_output<- readRDS(paste0(data_dir,"byCellType/", x))
  sigLRs<- Extract_res_allComparisons(cc_output, PathwaysRemove, sigThreshold, senderCells, receiverCells, subtypes)
  sigLRs$analysis<- gsub("all_cellChat_interact_", "", gsub(".Rds", "", x))
  return(sigLRs)
})
all_sigLRs<- do.call("rbind", all_output)
saveRDS(all_sigLRs, file=paste0(analysis_dir, "byCellType/all_LRs.Rds")) ## adding "byProgram" or "byDetailedAnnot"
write.csv(all_sigLRs, file=paste0(analysis_dir, "byCellType/all_LRs.csv")) ## adding "byProgram" or "byDetailedAnnot"
```




## By Program (CD4, CD8, myeloid programs, tumor celll types)
"separate batches"- used downstream
```{r}
analysis_dir_use<- paste0(analysis_dir, "byProgram/all_cells/")
data_dir_use<- paste0(data_dir, "byProgram/")

files<-list.files(data_dir_use)
files<- files[grepl("cellChat_LRs", files)]

all_res<- lapply(files, function(file){
  readRDS(paste0(data_dir_use, file))
})
all_res<- do.call("rbind", all_res)

## Add on subtype
all_res$Subtype<- as.character(plyr::mapvalues(all_res$sample, subtypes$sample, subtypes$MySubtypes,warn_missing = FALSE))

## rename columns to match other cellchat results
all_res<- all_res %>% dplyr::rename(Prob=prob, sender=source, receiver=target, pvalue=pval) %>% as.data.frame()

saveRDS(all_res,paste0(analysis_dir_use, "all_LRs.Rds"))
write.csv(all_res,paste0(analysis_dir_use, "all_LRs.csv"))
```


## By program (downsampled)
```{r}
analysis_dir_use<- paste0(analysis_dir, "byProgram/downsampled/")
data_dir_use<- paste0(data_dir, "byProgram_downsampled/")

files<-list.files(data_dir_use)
files<- files[grepl("cellChat_LRs", files)]

all_res<- lapply(files, function(file){
  readRDS(paste0(data_dir_use, file))
})
all_res<- do.call("rbind", all_res)

## Add on subtype
all_res$Subtype<- as.character(plyr::mapvalues(all_res$sample, subtypes$sample, subtypes$MySubtypes,warn_missing = FALSE))

## rename columns to match other cellchat results
all_res<- all_res %>% dplyr::rename(Prob=prob, sender=source, receiver=target, pvalue=pval) %>% as.data.frame()

saveRDS(all_res,paste0(analysis_dir_use, "all_LRs.Rds"))
write.csv(all_res,paste0(analysis_dir_use, "all_LRs.csv"))
```


## NO LONGER IN USE
## By Program (CD4, CD8, myeloid programs, tumor celll types)
## Load in each analysis, extract results, and merge
"single batch"- not used downstream
```{r}
analysis_dir_use<- paste0(analysis_dir, "byProgram/ARCHIVE/")
data_dir_use<- paste0(data_dir, "ARCHIVE/")

## Settings for extracting results, converting to simple df
PathwaysRemove<-c("HLA", "COL", "LAMININ", "MHC-I")
sigThreshold<- 1
senderCells<- unique(nCellTypes$CellType)
receiverCells<- unique(nCellTypes$CellType)

## List of results
files<- list.files(data_dir_use, pattern="byProgram") ## adding "byProgram"- other cellChat analyses already processed

## Cycle through each output,
all_output<- lapply(files, function(x){
  print(x)
  cc_output<- readRDS(paste0(data_dir, x))
  sigLRs<- Extract_res_allComparisons(cc_output, PathwaysRemove, sigThreshold, senderCells, receiverCells, subtypes)
  sigLRs$analysis<- gsub("all_cellChat_interact_", "", gsub(".Rds", "", x))
  return(sigLRs)
})
all_sigLRs<- do.call("rbind", all_output)
saveRDS(all_sigLRs, file=paste0(analysis_dir_use, "all_LRs.Rds")) 
write.csv(all_sigLRs, file=paste0(analysis_dir_use, "all_LRs.csv"))
```

