---
title: "R Notebook"
output: html_document
---

```{r}
library(CellChat)
library(NMF)
library(circlize)
library(patchwork)
library(Seurat) 
library(ggh4x)

options(stringsAsFactors = FALSE)

base_dir = "/Users/jlabelle/"
script_dir = paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/")
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/04a_Integrate_Immune.Tumor/CellChat/")
analysis_dir<- paste0(working_dir, "/analysis/")
figure_dir<- paste0(working_dir, "/figures/")

cellchatHelper<- paste0(script_dir, "CellChat_HelperFunctions.R")
source(cellchatHelper)

setwd(working_dir)

## Load in list of all cell types across all subtypes
nCellTypes<- readRDS(paste0(analysis_dir, "nCellTypes_bySample.Rds"))


library(RColorBrewer)
myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
```

## Load in LR results for all analyses (already merged)
```{r}
allLRs<- readRDS(paste0(analysis_dir, "all_LRs.Rds"))

## Subset to just sig- by pvalue and prob
sigLRs<- allLRs[allLRs$pvalue<=0.1 & allLRs$Prob>0,]

## Add spacing to subtypes for better plotting
sigLRs$Subtype<- gsub("Histone", "\nHistone\n", gsub("Ependymoma", "EPN", sigLRs$Subtype))


```

## 3 layers of cell type classification
"major" cell type = Tcell, Myeloid, or Tumor
"minor" cell type = CD4, CD8, AC-like, etc
"clonality" = only for CD8 T cells

## Set minor cell types within major cell type, add onto sigLRs
```{r}
tcell_types<- c("CD4_nonclonal", "CD8_nonclonal", "CD8_clonal")
myeloid_types<- "Myeloid"
tumor_types<- unique(sigLRs$sender)[!unique(sigLRs$sender) %in% c(Tcell_types, Myeloid_types)]

sigLRs$sender_major<- ifelse(sigLRs$sender %in% tumor_types, "Tumor",
                             ifelse(sigLRs$sender %in% tcell_types, "Tcell", "Myeloid"))
sigLRs$receiver_major<- ifelse(sigLRs$receiver %in% tumor_types, "Tumor",
                             ifelse(sigLRs$receiver %in% tcell_types, "Tcell", "Myeloid"))
```


## General plot types:
1. all LRs between 2 major cell types, regardless of clonality or sender cell 
2. All LRs between 1 major cell type and Tcells (or vice versa), regardless of minor cell type, by clonality
3. All LRs between 1 major cell type and minor cell types, without clonality
4. All LRs between 2 major cell types' minor cell types- heatmap colored by % of samples that LR is found in
5. Summary plots for LRs of interest


## 1. all LRs between 2 major cell types, regardless of clonality or sender cell 
```{r}
## Set major sender/receiver cell types
majorSender<- "Tcell"
majorReceiver<- "Myeloid"
sigLRs_plot<-sigLRs[sigLRs$sender_major==majorSender & sigLRs$receiver_major==majorReceiver,]

## If using Tcell, remove clonality info
sigLRs_plot$sender<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$sender))
sigLRs_plot$receiver<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$receiver))

## Add column for unique ID- by sample/ligand/receptor
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",sigLRs_plot$receptor)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by- usually will be subtype -  and plot
sigLRs_unique$FacetWrap1<- sigLRs_unique$Subtype

PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c("Subtype"))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir, "Sender.", majorSender, "_Receiver.", majorReceiver, ".png"), width=width, height=height)
```

## 2. All LRs between 1 major cell type and Tcells (or vice versa), regardless of minor cell type, by clonality
```{r}
## Set major sender/receiver cell types
majorSender<- "Tumor"
majorReceiver<- "CD4" ## receiver has to be Tcell, CD4, or CD8

## Split by clonality
sigLRs_plot<- sigLRs
sigLRs_plot$receiver_clonality<- ifelse(grepl("_clonal", sigLRs_plot$receiver), "clonal", 
                               ifelse(grepl("_nonclonal", sigLRs_plot$receiver), "nonclonal", "NA"))

sigLRs_plot$sender<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$sender))
sigLRs_plot$receiver<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$receiver))

## Subset to all Tcells, CD4, or CD8
if(majorReceiver=="Tcell"){
  sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver_major==majorReceiver,]
}else{
  sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver==majorReceiver,]
}


## Add column for unique ID- by sample/ligand/receptor
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",sigLRs_plot$receptor, "_",sigLRs_plot$receiver_clonality)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by- usually will be subtype -  and plot
sigLRs_unique$FacetWrap2<- sigLRs_unique$Subtype
sigLRs_unique$FacetWrap1<- sigLRs_unique$Sample

PlotPvalueProb(sigLRs_unique,xAxis = "receiver_clonality", labelXAxis = TRUE, facetWrap = c("Subtype", "Sample"))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir, "Sender.", majorSender, "_Receiver.Clonal",majorReceiver, ".png"), width=width, height=height)
```

## 3. All LRs between 1 major cell type and minor cell types, without clonality
For now, only using this for Tcells (CD4, CD8). May expand to include tumor cell types, or additional programs within Tcell/myeloid
```{r}
## Set major sender/receiver cell types
majorSender<- "Tcell"
majorReceiver<- "Tumor" 

## Remove clonality
sigLRs_plot<- sigLRs
sigLRs_plot$sender<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$sender))
sigLRs_plot$receiver<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$receiver))

## Subset to sender/receiver major cell type
sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver_major==majorReceiver,]

## Add column for unique ID- by sample/ligand/receptor/receiver cell type
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",
                              sigLRs_plot$receptor, "_",sigLRs_plot$receiver)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by and plot
sigLRs_unique$FacetWrap2<- sigLRs_unique$receiver
sigLRs_unique$FacetWrap1<- sigLRs_unique$Subtype

PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c("Subtype", "Receiver"),
               ncolForFacetWrap = length(unique(paste0(sigLRs_unique$FacetWrap1, sigLRs_unique$FacetWrap2))))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir, "Sender.", majorSender, "_Receiver.All",majorReceiver, ".png"), width=width, height=height)
```


## 4. All LRs between 2 major cell types' minor cell types- heatmap colored by % of samples that LR is found in
```{r}
## Set major sender/receiver cell types
majorSender<- "Tumor"
majorReceiver<- "Myeloid" 

## Subset to sender/receiver major cell type
sigLRs_plot<- sigLRs
sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver_major==majorReceiver,]

## For each unique sender/receiver combination, determine the total number of samples tested
sigLRs_plot$interaction_id<- paste0(sigLRs_plot$sender, "_", sigLRs_plot$receiver)
sigLRs_plot<- sigLRs_plot %>% group_by(interaction_id) %>% 
  mutate(nSamples_inInteraction=length(unique(Sample))) %>% as.data.frame()

## For each unique LR/interaction, determine the total number of samples identified
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$ligand, "_", sigLRs_plot$receptor, "_", sigLRs_plot$interaction_id)
sigLRs_plot<- sigLRs_plot %>% group_by(UniqueID) %>% mutate(nSamples_inLR=length(unique(Sample))) %>% as.data.frame()

## For each unique LR/interaction, divide n samples with LR/ n samples in interacting pair
sigLRs_unique<- sigLRs_plot[!duplicated(sigLRs_plot$UniqueID),]
sigLRs_unique$perSamples_inLR<- sigLRs_unique$nSamples_inLR/sigLRs_unique$nSamples_inInteraction

## Add on ID for ligand/receptor
sigLRs_unique$LR<- paste0(sigLRs_unique$ligand, "_", sigLRs_unique$receptor)

ggplot(sigLRs_unique, aes(x=sender, y=LR, fill=perSamples_inLR))+
  geom_tile()+
  theme_linedraw() + 
  facet_grid(cols = vars(receiver), scales = "free_x", space="free_x")+
  scale_fill_gradient(low="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text=element_text(color="black", face="bold"),
        axis.title = element_text(color="black", face="bold"))+
  ylab("") + xlab("Sender cell type")

height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$sender)) * .25
ggsave(paste0(figure_dir, "HeatmapBySampleProportion_Sender.", majorSender, "_Receiver.", majorReceiver, ".png"),
       width=width, height=height)
```

## 5. Summary plots for LRs of interest
```{r}
## Set major sender/receiver cell types +  ligand/receptors of interest
majorSender<- "Tumor"
majorReceiver<- "Myeloid" ## Tcell, CD4, CD8, or Myeloid
LR_list<- list(Tumor.Tcell.CD4.CD8=c("PVR_TIGIT","PROS1_AXL","CXCL16_CXCR6","CDH1_KLRG1","ANGPTL2_LILRB3", "CLEC2D_KLRB1"),
               Tumor.Myeloid=c("CCL5_CCR1","PROS1_AXL","ANGPTL2_LILRB3"))
LRs<- unlist(unname(LR_list[grepl(majorReceiver, names(LR_list))]))

## If using Tcell, remove clonality info
sigLRs_plot<- sigLRs
sigLRs_plot$sender<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$sender))
sigLRs_plot$receiver<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$receiver))


## Subset to all Tcells, CD4, or CD8
if(majorReceiver=="Tcell"){
  sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver_major==majorReceiver,]
}else{
  sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver==majorReceiver,]
}

## Subset to LR pairs of interest
sigLRs_plot$LR<- paste0(sigLRs_plot$ligand, "_", sigLRs_plot$receptor)
sigLRs_plot<- sigLRs_plot[sigLRs_plot$LR %in% LRs,]

## If using Tcell, remove clonality info
sigLRs_plot$sender<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$sender))
sigLRs_plot$receiver<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$receiver))

## Add column for unique ID- by sample/ligand/receptor
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",sigLRs_plot$receptor)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by- usually will be subtype -  and plot
sigLRs_unique$FacetWrap1<- sigLRs_unique$Subtype

PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c("Subtype"))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir, "LR.summary_Sender.", majorSender, "_Receiver.", majorReceiver, ".png"), width=width, height=height)
```


