---
title: "R Notebook"
output: html_document
---

```{r}
library(CellChat)
library(NMF)
library(circlize)
library(patchwork)
library(Seurat) 
library(ggh4x)

options(stringsAsFactors = FALSE)

base_dir = "/Users/jlabelle/"
script_dir = paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/")
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/04a_Integrate_Immune.Tumor/CellChat/")
analysis_dir<- paste0(working_dir, "/analysis/byProgram/")
figure_dir<- paste0(working_dir, "/figures/byProgram/")

cellchatHelper<- paste0(script_dir, "CellChat_HelperFunctions.R")
source(cellchatHelper)

setwd(working_dir)

## Load in list of all cell types across all subtypes
nCellTypes<- readRDS( "analysis/nCellTypes_bySample.Rds")


library(RColorBrewer)
myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
```


## Load in LR results for all analyses (already merged)
```{r}
allLRs<- readRDS(paste0(analysis_dir, "all_LRs.Rds")) 

## Subset to just sig- by pvalue and prob
sigLRs<- allLRs[allLRs$pvalue<=0.1 & allLRs$Prob>0,]

## Add spacing to subtypes for better plotting
sigLRs$Subtype<- gsub("Histone", "\nHistone\n", gsub("Ependymoma", "EPN", sigLRs$Subtype))


```



## 2 layers of cell type classification:
"major" cell type = CD4, CD8, Myeloid, or Tumor
"minor" cell type = Tumor cell type (AC-like, OPC-like), myeloid program, CD4 program, CD8 program
  

## Set minor cell types within major cell type, add onto sigLRs
```{r}
cd4_types<- unique(sigLRs$sender)[grepl("CD4", unique(sigLRs$sender))]
cd8_types<- unique(sigLRs$sender)[grepl("CD8", unique(sigLRs$sender))]
myeloid_types<- c("IFN_TAM", "Monocytes", "Inflammatory", "Hypoxic_TAM", "Stress_Response", "SEPP1_Mo_TAM")
tumor_types<- unique(sigLRs$sender)[!unique(sigLRs$sender) %in% c(cd4_types, cd8_types, myeloid_types)]

sigLRs$sender_major<- ifelse(sigLRs$sender %in% tumor_types, "Tumor",
                      ifelse(sigLRs$sender %in% myeloid_types, "Myeloid",
                      ifelse(sigLRs$sender %in% cd4_types, "CD4", "CD8")))
sigLRs$receiver_major<- ifelse(sigLRs$receiver %in% tumor_types, "Tumor",
                        ifelse(sigLRs$receiver %in% myeloid_types, "Myeloid",
                        ifelse(sigLRs$receiver %in% cd4_types, "CD4", "CD8")))

## Add \n to minor cell types for better plotting
sigLRs$sender<- gsub("_", "\n", sigLRs$sender)
sigLRs$receiver<- gsub("_", "\n", sigLRs$receiver)
```


## General plot types:
1. all LRs between 2 major cell types, regardless of minor cell types
2. All LRs between 1 major cell type and minor cell types
3. All LRs between 2 major cell types' minor cell types- heatmap colored by % of samples that LR is found in
4. Number of interactions between all pairwise combinations of 2 sets of minor cell types
5. Summary plots for LRs of interest


## 1. all LRs between 2 major cell types, regardless of minor cell types
```{r}
## Set major sender/receiver cell types
majorSender<- "Tumor"
majorReceiver<- "CD8"
sigLRs_plot<-sigLRs[sigLRs$sender_major==majorSender & sigLRs$receiver_major==majorReceiver,]

## Add column for unique ID- by sample/ligand/receptor
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",sigLRs_plot$receptor)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by- usually will be subtype -  and plot
sigLRs_unique$FacetWrap1<- sigLRs_unique$Subtype

PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c("Subtype"))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir_use, "Sender.", majorSender, "_Receiver.", majorReceiver, ".png"), width=width, height=height)
```


## 2. All LRs between 1 major cell type and minor cell types
```{r}
## Set major sender/receiver cell types
majorSender<- "Tumor"
majorReceiver<- "CD8" 

## Subset to sender/receiver major cell type
sigLRs_plot<-sigLRs[sigLRs$sender_major==majorSender & sigLRs$receiver_major==majorReceiver,]

## Add column for unique ID- by sample/ligand/receptor/receiver cell type
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",
                              sigLRs_plot$receptor, "_",sigLRs_plot$receiver)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by and plot
sigLRs_unique$FacetWrap1<- sigLRs_unique$receiver
#sigLRs_unique$FacetWrap1<- sigLRs_unique$Subtype

#PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c("Subtype", "Receiver"),
               #ncolForFacetWrap = length(unique(paste0(sigLRs_unique$FacetWrap1, sigLRs_unique$FacetWrap2))))+
 # theme(axis.text = element_text(face="bold"))+
  #ggtitle("")+
  #theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  #ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))

PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c( "Receiver"),
               ncolForFacetWrap = length(unique(paste0(sigLRs_unique$FacetWrap1))))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir_use, "Sender.", majorSender, "_Receiver.All",majorReceiver, ".png"), width=width, height=height)
```


## 3. All LRs between 2 major cell types' minor cell types- heatmap colored by % of samples that LR is found in
```{r}
## Set major sender/receiver cell types
majorSender<- "Tumor"
majorReceiver<- "Myeloid" 

## Subset to sender/receiver major cell type
sigLRs_plot<- sigLRs
sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver_major==majorReceiver,]

## Remove \n from sender cell types for better plotting
sigLRs_plot$sender<- gsub("\n", "_", sigLRs_plot$sender)

## For each unique sender/receiver combination, determine the total number of samples tested
sigLRs_plot$interaction_id<- paste0(sigLRs_plot$sender, "_", sigLRs_plot$receiver)
sigLRs_plot<- sigLRs_plot %>% group_by(interaction_id) %>% 
  mutate(nSamples_inInteraction=length(unique(Sample))) %>% as.data.frame()

## For each unique LR/interaction, determine the total number of samples identified
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$ligand, "_", sigLRs_plot$receptor, "_", sigLRs_plot$interaction_id)
sigLRs_plot<- sigLRs_plot %>% group_by(UniqueID) %>% mutate(nSamples_inLR=length(unique(Sample))) %>% as.data.frame()

## For each unique LR/interaction, divide n samples with LR/ n samples in interacting pair
sigLRs_unique<- sigLRs_plot[!duplicated(sigLRs_plot$UniqueID),]
sigLRs_unique$perSamples_inLR<- sigLRs_unique$nSamples_inLR/sigLRs_unique$nSamples_inInteraction

## Add on ID for ligand/receptor
sigLRs_unique$LR<- paste0(sigLRs_unique$ligand, "_", sigLRs_unique$receptor)

ggplot(sigLRs_unique, aes(x=sender, y=LR, fill=perSamples_inLR))+
  geom_tile()+
  theme_linedraw() + 
  facet_grid(cols = vars(receiver), scales = "free_x", space="free_x")+
  scale_fill_gradient(low="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text=element_text(color="black", face="bold"),
        axis.title = element_text(color="black", face="bold"))+
  ylab("") + xlab("Sender cell type")

height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$sender)) * .25
ggsave(paste0(figure_dir_use, "HeatmapBySampleProportion_Sender.", majorSender, "_Receiver.", majorReceiver, ".png"),
       width=width, height=height)
```

## 4. Number of interactions between all pairwise combinations of 2 sets of minor cell types
```{r}
## Set major sender/receiver cell types
majorCellType1<- c("CD4", "CD8")
majorCellType2<- "Tumor" 

## Set normalization method-"noNorm", "bySample", "byCell"
norm_method<- "noNorm"

## Subset to sender/receiver major cell type
sigLRs_plot<- sigLRs
sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major %in% 
                           c(majorCellType1, majorCellType2) & sigLRs_plot$receiver_major %in% c(majorCellType1, majorCellType2) ,]

## Remove \n from cell types for better plotting
sigLRs_plot$sender<- gsub("\n", "_", sigLRs_plot$sender)
sigLRs_plot$receiver<- gsub("\n", "_", sigLRs_plot$receiver)

## For each unique sender/receiver combination, determine the total number of sig LRs
sigLRs_plot$interaction_id<- paste0(sigLRs_plot$sender, "__", sigLRs_plot$receiver)
sigLRs_nLR<- sigLRs_plot %>% group_by(interaction_id) %>% summarise(nLRs=length(unique(LR))) %>% as.data.frame()


## Split sender/receiver back up
sigLRs_nLR$sender<- sapply(sigLRs_nLR$interaction_id, function(x){unlist(strsplit(x, "__"))[[1]]})
sigLRs_nLR$receiver<- sapply(sigLRs_nLR$interaction_id, function(x){unlist(strsplit(x, "__"))[[2]]})

## Normalize based on set method
  ## Divide by number of samples tested in pair
  if(norm_method=="bySample"){
    ## For each unique sender/receiver combination, determine the total number of samples tested, add to number of LRs
    sigLRs_nSamples<-  sigLRs_plot %>% group_by(interaction_id) %>%
      summarise(nSamples_inInteraction=length(unique(Sample))) %>% as.data.frame()
    sigLRs_nLR$nSamples<- as.numeric(plyr::mapvalues(sigLRs_nLR$interaction_id, sigLRs_nSamples$interaction_id, sigLRs_nSamples$nSamples_inInteraction))
    sigLRs_nLR$nLRs_norm<- sigLRs_nLR$nLRs/sigLRs_nLR$nSamples
  }

  ## Divide by number of cells tested in pair
  if(norm_method=="byCell"){
    ## Remove sample/cell type with < 10- not input into cellchat
    nCells_bySample_filt<- nCells_bySample[nCells_bySample$NCells>10,]
    
    nCells_total<- nCells_bySample %>% group_by(CellType) %>% summarise(nCells=sum(NCells)) %>% as.data.frame()
    
    ## Cycle through each interaction, determining the number of cells that were input into cellchat
    ## Note that this needs to be done on sample-wise basis
    cells_perPair<- apply(sigLRs_nLR, 1, function(x){
      nCells_pair<- nCells_bySample_filt[nCells_bySample_filt$CellType %in% c(x["sender"],x["receiver"]),]
      nCells_pair_filt<- nCells_pair %>% group_by(Sample) %>% summarise(PF=sum(NCells>10)) %>% as.data.frame()
      samples_used<- as.character(nCells_pair_filt[nCells_pair_filt$PF==2, "Sample"])
      nCells_pair<- nCells_pair[nCells_pair$Sample %in% samples_used,]
      nCells_total<- sum(nCells_pair$NCells)
      return(nCells_total)
    })
    sigLRs_nLR$nCells<- cells_perPair
    sigLRs_nLR$nLRs_norm<- sigLRs_nLR$nLRs/sigLRs_nLR$nCells
  }

  ## No norm
  if(norm_method=="noNorm"){
    sigLRs_nLR$nLRs_norm<- sigLRs_nLR$nLRs
  }

## Get vector of both cell types used
majorCells1<- unique(sigLRs_plot$sender[sigLRs_plot$sender_major %in% majorCellType1])
majorCells2<- unique(sigLRs_plot$sender[sigLRs_plot$sender_major %in% majorCellType2])
  
  
## Add on major cell type
sigLRs_nLR$major_sender<- ifelse(sigLRs_nLR$sender %in% majorCells1, majorCellType1, majorCellType2 )
sigLRs_nLR$major_receiver<- ifelse(sigLRs_nLR$receiver %in%majorCells1, majorCellType1, majorCellType2 )
sigLRs_nLR$major_sr<- paste0(sigLRs_nLR$major_sender, "\nto\n", sigLRs_nLR$major_receiver)


## Switch sender/receiver for one- just for plotting purposes
sigLRs_nLR$sender.plot<- ifelse(sigLRs_nLR$major_receiver==majorCellType1, sigLRs_nLR$receiver, sigLRs_nLR$sender)
sigLRs_nLR$receiver.plot<- ifelse(sigLRs_nLR$major_receiver==majorCellType1, sigLRs_nLR$sender, sigLRs_nLR$receiver)

## Order by major cell type
sigLRs_nLR$sender.plot<- factor(sigLRs_nLR$sender.plot, levels=c(majorCells1, majorCells2))
sigLRs_nLR$receiver.plot<- factor(sigLRs_nLR$receiver.plot, levels=c(majorCells1, majorCells2))

## Plot
ggplot(sigLRs_nLR, aes(x=sender.plot, y=receiver.plot, fill=nLRs_norm))+
  geom_tile()+
  #theme_linedraw() + 
  theme_classic()+
  scale_fill_gradient(low="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text=element_text(color="black", face="bold"),
        axis.title = element_text(color="black", face="bold"),
        strip.text = element_text(face="bold", size=12))+
  ylab("") + xlab("")+
  facet_grid(cols=vars(major_sr))

ggsave(paste0(figure_dir, "HeatmapByTotalLR_Norm", norm_method,"_",majorCellType1,".", majorCellType2, ".png"),
       width=6, height=4)

```


## Alternative to above: chord diagram
```{r}
## Alternative: chord diagram
majorCells1<- gsub("_", "\n",gsub("\\.", "\n",majorCells1))
majorCells2<- gsub("_", "\n",gsub("\\.", "\n",majorCells2))

  ## Set colors- shades of blue and green
  colfunc1 <- colorRampPalette(c("olivedrab", "cyan"))
  colfunc2<- colorRampPalette(c("orange", "violetred"))
  colors_use<-c(colfunc1(length(majorCells1)),colfunc2(length(majorCells2)))
  names(colors_use)<- c(majorCells1, majorCells2)
  
  ## Subset to needed columns
  tmp<- sigLRs_nLR[,c( "sender", "receiver","nLRs_norm")]
  tmp$sender<- gsub( "_","\n",gsub("\\.", "\n", tmp$sender))
  tmp$receiver<- gsub("_", "\n", gsub("\\.", "\n",tmp$receiver))
  
  ## Plot
  png(paste0(figure_dir_use, "ChordByTotalLR_Norm", norm_method,"_",majorCellType1,".", majorCellType2, ".png"), width=400, height=400) ## for png, use size of 1000
  chordDiagram(tmp,annotationTrack = c("grid", "name"), order=c(majorCells1, majorCells2),
                       grid.col = colors_use)
  dev.off()
```

## 5. Summary plots for LRs of interest
```{r}
## Set major sender/receiver cell types +  ligand/receptors of interest
majorSender<- "Tumor"
majorReceiver<- "Myeloid" ## Tcell, CD4, CD8, or Myeloid
LR_list<- list(Tumor.Tcell.CD4.CD8=c("PVR_TIGIT","PROS1_AXL","CXCL16_CXCR6","CDH1_KLRG1","ANGPTL2_LILRB3", "CLEC2D_KLRB1"),
               Tumor.Myeloid=c("CCL5_CCR1","PROS1_AXL","ANGPTL2_LILRB3"))
LRs<- unlist(unname(LR_list[grepl(majorReceiver, names(LR_list))]))

## If using Tcell, remove clonality info
sigLRs_plot<- sigLRs
sigLRs_plot$sender<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$sender))
sigLRs_plot$receiver<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$receiver))


## Subset to all Tcells, CD4, or CD8
if(majorReceiver=="Tcell"){
  sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver_major==majorReceiver,]
}else{
  sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver==majorReceiver,]
}

## Subset to LR pairs of interest
sigLRs_plot$LR<- paste0(sigLRs_plot$ligand, "_", sigLRs_plot$receptor)
sigLRs_plot<- sigLRs_plot[sigLRs_plot$LR %in% LRs,]

## If using Tcell, remove clonality info
sigLRs_plot$sender<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$sender))
sigLRs_plot$receiver<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$receiver))

## Add column for unique ID- by sample/ligand/receptor
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",sigLRs_plot$receptor)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by- usually will be subtype -  and plot
sigLRs_unique$FacetWrap1<- sigLRs_unique$Subtype

PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c("Subtype"))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir, "LR.summary_Sender.", majorSender, "_Receiver.", majorReceiver, ".png"), width=width, height=height)
```

## Exploratory- can delete
```{r}
## Interaction pairs with at least 2 samples
sigLRs$pair<- paste0(sigLRs$sender, "__", sigLRs$receiver)
nSamples_perPair<- sigLRs %>% group_by(pair) %>% summarise(nSamples=length(unique(Sample))) %>% as.data.frame()
sigLRs_multiple<- sigLRs[sigLRs$pair %in% nSamples_perPair$pair[nSamples_perPair$nSamples > 1],]

## Get # of LRs for each pair by sample
nLR_bySample<- sigLRs_multiple %>% group_by(pair, Sample) %>% summarise(nLR=n()) %>% as.data.frame()

## Add on number of total cells in that pair
nLR_bySample$sender<- sapply(nLR_bySample$pair, function(x){unlist(strsplit(x, "__"))[[1]]})
nLR_bySample$receiver<- sapply(nLR_bySample$pair, function(x){unlist(strsplit(x, "__"))[[2]]})
nCells_inPair<- sapply(1:nrow(nLR_bySample), function(x){
  tmp<- nLR_bySample[x,]
  cells<- sum(nCellTypes$NCells[nCellTypes$Sample==tmp$Sample & nCellTypes$CellType==tmp$sender],
              nCellTypes$NCells[nCellTypes$Sample==tmp$Sample & nCellTypes$CellType==tmp$receiver])
  return(cells)
})
nLR_bySample$nCellsInput<- nCells_inPair

## Plot for a few pairs
nLR_plot<- nLR_bySample[nLR_bySample$pair %in% nSamples_perPair$pair[nSamples_perPair$nSamples>4],]
all_plots<- lapply(unique(nLR_plot$pair), function(x){
  tmp<- nLR_plot[nLR_plot$pair==x,]
  ggplot(tmp, aes(x=nLR, y=nCellsInput))+
    geom_point()+
    geom_smooth(method="lm", se=FALSE)+
    ggtitle(gsub("__", " to ", x))
})
cowplot::plot_grid(plotlist=all_plots, ncol=4)
ggsave(paste0(figure_dir, "testing_nLRs.vs.cellInput.png"), width=12, height=12)

```

