---
title: "Plot cellchat"
output: html_document
---

## Plot cell chat results between immune programs, tumor cell types

```{r}
library(CellChat)
library(NMF)
library(circlize)
library(patchwork)
library(Seurat) 
library(ggh4x)
library(ggalluvial)

options(stringsAsFactors = FALSE)

base_dir = "/Users/jlabelle/"
script_dir = paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/")
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/04a_Integrate_Immune.Tumor/CellChat/")
analysis_dir<- paste0(working_dir, "/analysis/byProgram/")
figure_dir<- paste0(working_dir, "/figures/byProgram/")

cellchatHelper<- paste0(script_dir, "CellChat_HelperFunctions.R")
source(cellchatHelper)

setwd(working_dir)

## Load in list of all cell types across all subtypes
nCellTypes<- readRDS( "analysis/nCellTypes_bySample.Rds")


library(RColorBrewer)
myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
```


## Load in LR results for all analyses (already merged)
```{r}
allLRs<- readRDS(paste0(analysis_dir, "all_LRs.Rds")) 

## Subset to just sig- by pvalue and prob
sigLRs<- allLRs[allLRs$pvalue<=0.1 & allLRs$Prob>0,]

## Add spacing to subtypes for better plotting
sigLRs$Subtype<- gsub("Histone", "\nHistone\n", gsub("Ependymoma", "EPN", sigLRs$Subtype))


```



## 2 layers of cell type classification:
"major" cell type = CD4, CD8, Myeloid, or Tumor
"minor" cell type = Tumor cell type (AC-like, OPC-like), myeloid program, CD4 program, CD8 program
  

## Set minor cell types within major cell type, add onto sigLRs
```{r}
cd4_types<- unique(sigLRs$sender)[grepl("CD4", unique(sigLRs$sender))]
cd8_types<- unique(sigLRs$sender)[grepl("CD8", unique(sigLRs$sender))]
myeloid_types<- c("IFN_TAM", "Monocytes", "Inflammatory", "Hypoxic_TAM", "Stress_Response", "SEPP1_Mo_TAM")
tumor_types<- unique(sigLRs$sender)[!unique(sigLRs$sender) %in% c(cd4_types, cd8_types, myeloid_types)]

sigLRs$sender_major<- ifelse(sigLRs$sender %in% tumor_types, "Tumor",
                      ifelse(sigLRs$sender %in% myeloid_types, "Myeloid",
                      ifelse(sigLRs$sender %in% cd4_types, "CD4", "CD8")))
sigLRs$receiver_major<- ifelse(sigLRs$receiver %in% tumor_types, "Tumor",
                        ifelse(sigLRs$receiver %in% myeloid_types, "Myeloid",
                        ifelse(sigLRs$receiver %in% cd4_types, "CD4", "CD8")))

## Add \n to minor cell types for better plotting
sigLRs$sender<- gsub("_", "\n", sigLRs$sender)
sigLRs$receiver<- gsub("_", "\n", sigLRs$receiver)
```


## General plot types:
1. all LRs between 2 major cell types, regardless of minor cell types
2. All LRs between 1 major cell type and minor cell types
3. All LRs between 2 major cell types' minor cell types- heatmap colored by % of samples that LR is found in
4. Number of interactions between all pairwise combinations of 2 sets of minor cell types
5. Summary plots for LRs of interest


## 1. all LRs between 2 major cell types, regardless of minor cell types
```{r}
## Set major sender/receiver cell types
majorSender<- "Tumor"
majorReceiver<- "CD8"
sigLRs_plot<-sigLRs[sigLRs$sender_major==majorSender & sigLRs$receiver_major==majorReceiver,]

## Add column for unique ID- by sample/ligand/receptor
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",sigLRs_plot$receptor)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by- usually will be subtype -  and plot
sigLRs_unique$FacetWrap1<- sigLRs_unique$Subtype

PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c("Subtype"))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir_use, "Sender.", majorSender, "_Receiver.", majorReceiver, ".png"), width=width, height=height)
```


## 2. All LRs between 1 major cell type and minor cell types
```{r}
## Set major sender/receiver cell types
majorSender<- "Tumor"
majorReceiver<- "CD8" 

## Subset to sender/receiver major cell type
sigLRs_plot<-sigLRs[sigLRs$sender_major==majorSender & sigLRs$receiver_major==majorReceiver,]

## Add column for unique ID- by sample/ligand/receptor/receiver cell type
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",
                              sigLRs_plot$receptor, "_",sigLRs_plot$receiver)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by and plot
sigLRs_unique$FacetWrap1<- sigLRs_unique$receiver
#sigLRs_unique$FacetWrap1<- sigLRs_unique$Subtype

#PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c("Subtype", "Receiver"),
               #ncolForFacetWrap = length(unique(paste0(sigLRs_unique$FacetWrap1, sigLRs_unique$FacetWrap2))))+
 # theme(axis.text = element_text(face="bold"))+
  #ggtitle("")+
  #theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  #ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))

PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c( "Receiver"),
               ncolForFacetWrap = length(unique(paste0(sigLRs_unique$FacetWrap1))))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir_use, "Sender.", majorSender, "_Receiver.All",majorReceiver, ".png"), width=width, height=height)
```


## 3. All LRs between 2 major cell types' minor cell types- heatmap colored by % of samples that LR is found in
```{r}
## Set major sender/receiver cell types
majorSender<- "Tumor"
majorReceiver<- "Myeloid" 

## Subset to sender/receiver major cell type
sigLRs_plot<- sigLRs
sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver_major==majorReceiver,]

## Remove \n from sender cell types for better plotting
sigLRs_plot$sender<- gsub("\n", "_", sigLRs_plot$sender)

## For each unique sender/receiver combination, determine the total number of samples tested
sigLRs_plot$interaction_id<- paste0(sigLRs_plot$sender, "_", sigLRs_plot$receiver)
sigLRs_plot<- sigLRs_plot %>% group_by(interaction_id) %>% 
  mutate(nSamples_inInteraction=length(unique(Sample))) %>% as.data.frame()

## For each unique LR/interaction, determine the total number of samples identified
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$ligand, "_", sigLRs_plot$receptor, "_", sigLRs_plot$interaction_id)
sigLRs_plot<- sigLRs_plot %>% group_by(UniqueID) %>% mutate(nSamples_inLR=length(unique(Sample))) %>% as.data.frame()

## For each unique LR/interaction, divide n samples with LR/ n samples in interacting pair
sigLRs_unique<- sigLRs_plot[!duplicated(sigLRs_plot$UniqueID),]
sigLRs_unique$perSamples_inLR<- sigLRs_unique$nSamples_inLR/sigLRs_unique$nSamples_inInteraction

## Add on ID for ligand/receptor
sigLRs_unique$LR<- paste0(sigLRs_unique$ligand, "_", sigLRs_unique$receptor)

ggplot(sigLRs_unique, aes(x=sender, y=LR, fill=perSamples_inLR))+
  geom_tile()+
  theme_linedraw() + 
  facet_grid(cols = vars(receiver), scales = "free_x", space="free_x")+
  scale_fill_gradient(low="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text=element_text(color="black", face="bold"),
        axis.title = element_text(color="black", face="bold"))+
  ylab("") + xlab("Sender cell type")

height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$sender)) * .25
ggsave(paste0(figure_dir_use, "HeatmapBySampleProportion_Sender.", majorSender, "_Receiver.", majorReceiver, ".png"),
       width=width, height=height)
```

## 4. Number of interactions between all pairwise combinations of 2 sets of minor cell types
Plotting options: heatmap, chord diagram, sankey plot

## Preprocess
```{r}
## Set major sender/receiver cell types
majorCellType1<- c("CD4", "CD8")
majorCellType2<- "Myeloid" 

## Set normalization method-"noNorm", "bySample", "byCell"
norm_method<- "noNorm"

## Subset to sender/receiver major cell type
sigLRs_plot<- sigLRs
sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major %in% 
                           c(majorCellType1, majorCellType2) & sigLRs_plot$receiver_major %in% c(majorCellType1, majorCellType2) ,]

## Remove \n from cell types for better plotting
sigLRs_plot$sender<- gsub("\n", "_", sigLRs_plot$sender)
sigLRs_plot$receiver<- gsub("\n", "_", sigLRs_plot$receiver)

## For each unique sender/receiver combination, determine the total number of sig LRs
sigLRs_plot$interaction_id<- paste0(sigLRs_plot$sender, "__", sigLRs_plot$receiver)
sigLRs_nLR<- sigLRs_plot %>% group_by(interaction_id) %>% summarise(nLRs=length(unique(LR))) %>% as.data.frame()


## Split sender/receiver back up
sigLRs_nLR$sender<- sapply(sigLRs_nLR$interaction_id, function(x){unlist(strsplit(x, "__"))[[1]]})
sigLRs_nLR$receiver<- sapply(sigLRs_nLR$interaction_id, function(x){unlist(strsplit(x, "__"))[[2]]})

## Normalize based on set method
  ## Divide by number of samples tested in pair
  if(norm_method=="bySample"){
    ## For each unique sender/receiver combination, determine the total number of samples tested, add to number of LRs
    sigLRs_nSamples<-  sigLRs_plot %>% group_by(interaction_id) %>%
      summarise(nSamples_inInteraction=length(unique(Sample))) %>% as.data.frame()
    sigLRs_nLR$nSamples<- as.numeric(plyr::mapvalues(sigLRs_nLR$interaction_id, sigLRs_nSamples$interaction_id, sigLRs_nSamples$nSamples_inInteraction))
    sigLRs_nLR$nLRs_norm<- sigLRs_nLR$nLRs/sigLRs_nLR$nSamples
  }

  ## Divide by number of cells tested in pair
  if(norm_method=="byCell"){
    ## Remove sample/cell type with < 10- not input into cellchat
    nCells_bySample_filt<- nCells_bySample[nCells_bySample$NCells>10,]
    
    nCells_total<- nCells_bySample %>% group_by(CellType) %>% summarise(nCells=sum(NCells)) %>% as.data.frame()
    
    ## Cycle through each interaction, determining the number of cells that were input into cellchat
    ## Note that this needs to be done on sample-wise basis
    cells_perPair<- apply(sigLRs_nLR, 1, function(x){
      nCells_pair<- nCells_bySample_filt[nCells_bySample_filt$CellType %in% c(x["sender"],x["receiver"]),]
      nCells_pair_filt<- nCells_pair %>% group_by(Sample) %>% summarise(PF=sum(NCells>10)) %>% as.data.frame()
      samples_used<- as.character(nCells_pair_filt[nCells_pair_filt$PF==2, "Sample"])
      nCells_pair<- nCells_pair[nCells_pair$Sample %in% samples_used,]
      nCells_total<- sum(nCells_pair$NCells)
      return(nCells_total)
    })
    sigLRs_nLR$nCells<- cells_perPair
    sigLRs_nLR$nLRs_norm<- sigLRs_nLR$nLRs/sigLRs_nLR$nCells
  }

  ## No norm
  if(norm_method=="noNorm"){
    sigLRs_nLR$nLRs_norm<- sigLRs_nLR$nLRs
  }

## Get vector of both cell types used
majorCells1<- unique(sigLRs_plot$sender[sigLRs_plot$sender_major %in% majorCellType1])
majorCells2<- unique(sigLRs_plot$sender[sigLRs_plot$sender_major %in% majorCellType2])
  
  
## Add on major cell type
sigLRs_nLR$major_sender<- ifelse(sigLRs_nLR$sender %in% majorCells1, majorCellType1, majorCellType2 )
sigLRs_nLR$major_receiver<- ifelse(sigLRs_nLR$receiver %in%majorCells1, majorCellType1, majorCellType2 )
sigLRs_nLR$major_sr<- paste0(sigLRs_nLR$major_sender, "\nto\n", sigLRs_nLR$major_receiver)
```

## Heatmap
```{r}
sigLRs_plot<- sigLRs_nLR

## Switch sender/receiver for one- just for plotting purposes
sigLRs_plot$sender.plot<- ifelse(sigLRs_plot$major_receiver==majorCellType1, sigLRs_plot$receiver, sigLRs_plot$sender)
sigLRs_plot$receiver.plot<- ifelse(sigLRs_plot$major_receiver==majorCellType1, sigLRs_plot$sender, sigLRs_plot$receiver)

## Order by major cell type
sigLRs_plot$sender.plot<- factor(sigLRs_plot$sender.plot, levels=c(majorCells1, majorCells2))
sigLRs_plot$receiver.plot<- factor(sigLRs_plot$receiver.plot, levels=c(majorCells1, majorCells2))

## Plot
ggplot(sigLRs_plot, aes(x=sender.plot, y=receiver.plot, fill=nLRs_norm))+
  geom_tile()+
  #theme_linedraw() + 
  theme_classic()+
  scale_fill_gradient(low="white", high="red")+
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text=element_text(color="black", face="bold"),
        axis.title = element_text(color="black", face="bold"),
        strip.text = element_text(face="bold", size=12))+
  ylab("") + xlab("")+
  facet_grid(cols=vars(major_sr))

ggsave(paste0(figure_dir, "HeatmapByTotalLR_Norm", norm_method,"_",majorCellType1,".", majorCellType2, ".png"),
       width=6, height=4)

```


## chord diagram
```{r}
sigLRs_plot<- sigLRs_nLR

## Alternative: chord diagram
majorCells1<- gsub("_", "\n",gsub("\\.", "\n",majorCells1))
majorCells2<- gsub("_", "\n",gsub("\\.", "\n",majorCells2))

  ## Set colors- shades of blue and green
  colfunc1 <- colorRampPalette(c("olivedrab", "cyan"))
  colfunc2<- colorRampPalette(c("orange", "violetred"))
  colors_use<-c(colfunc1(length(majorCells1)),colfunc2(length(majorCells2)))
  names(colors_use)<- c(majorCells1, majorCells2)
  
  ## Subset to needed columns
  tmp<- sigLRs_plot[,c( "sender", "receiver","nLRs_norm")]
  tmp$sender<- gsub( "_","\n",gsub("\\.", "\n", tmp$sender))
  tmp$receiver<- gsub("_", "\n", gsub("\\.", "\n",tmp$receiver))
  
  ## Plot
  png(paste0(figure_dir_use, "ChordByTotalLR_Norm", norm_method,"_",majorCellType1,".", majorCellType2, ".png"), width=400, height=400) ## for png, use size of 1000
  chordDiagram(tmp,annotationTrack = c("grid", "name"), order=c(majorCells1, majorCells2),
                       grid.col = colors_use)
  dev.off()
```


## Sankey plot
```{r}
sigLRs_plot<- sigLRs_nLR

## Order cell types by major cell type
ordered_celltypes<- gsub("\n", "_", c(cd4_types, cd8_types, myeloid_types, tumor_types))
sigLRs_plot$receiver.plot<- factor(sigLRs_plot$receiver, levels=ordered_celltypes[ordered_celltypes %in% sigLRs_plot$receiver])
sigLRs_plot$sender.plot<- factor(sigLRs_plot$sender, levels=ordered_celltypes[ordered_celltypes %in% sigLRs_plot$sender])

sigLRs_plot$major_sender<- gsub("CD4", "Immune", gsub("CD8", "Immune",gsub("Myeloid2", "Immune", sigLRs_plot$major_sender)))

## Plot
  ggplot(data=sigLRs_plot, aes(axis1=sender, axis2=receiver,  y=nLRs_norm))+
    geom_alluvium(aes(fill = nLRs_norm)) +
    geom_stratum() +
    theme_classic()+
    geom_text(stat = "stratum",
              aes(label = after_stat(stratum))) +
    scale_x_discrete(limits = c("Sender", "Receiver"),
                     expand = c(0.15, 0.05), position="top") +
    #scale_fill_gradientn(colors=c("blue", "green", "yellow", "orange", "red"))+
    #scale_fill_viridis_c()+
    paletteer::scale_fill_paletteer_c("viridis::plasma")+
    ylab("")+
    theme(axis.text.y=element_blank(), 
          axis.ticks.y=element_blank(),
          axis.text.x = element_text(face="bold", color="black", size=12),
          strip.text = element_blank())  +
    facet_grid(rows=vars(major_sender), scale="free_y", space="free_y")

ggsave(paste0(figure_dir, "Sankey_ByTotalLR_Norm", norm_method,"_",paste(majorCellType1, collapse="."),".", majorCellType2, ".png"),
       width=8, height=16)


```


sankey plot testing
```{r}


df<- sigLRs_plot[,c("sender", "receiver", "major_sender", "major_receiver")]
df_list<- split(df, f=df$sender_major)
df_list<- lapply(df_list, function(x){x[,!colnames(x) %in% c("sender_major", "receiver_major")]})
colnames(df_list[[1]])<- c("CellA1", "CellB")
colnames(df_list[[2]])<- c("CellB", "CellA2")

test<- lapply(unique(df_list[[1]]$CellB), function(x){
  df1<- df_list[[1]]
  df2<- df_list[[2]]
  
  df1<- df1[df1$CellB==x,]
  df2<- df2[df2$CellB==x,]
  
  ## get largest and smallest- CellB as sender or as receiver?
  if(nrow(df1)>nrow(df2)){max_df<- df1; min_df<- df2}else{max_df<- df2;min_df<-df1}
  row_diff<- nrow(max_df)-nrow(min_df)
  
  print(x)
  if(sum(grepl("CellA1", colnames(max_df)))>0){
    print("More LRs in tcell sender")
  }else{print("More LRs in tcell receiver")}
  
  ## For min df- add on NAs to make up difference with major
  min_df_add<- data.frame(CellA=rep(NA, row_diff), CellB=rep(x, row_diff))
  min_df<- min_df[,order(colnames(min_df))]
  colnames(min_df_add)<- colnames(min_df)
  min_df<- rbind(min_df, min_df_add)
  
  ## Order columns in min/max the same way
  min_df_cellA<- colnames(min_df)[grepl("CellA", colnames(min_df))]
  min_df<-min_df[order(min_df[[min_df_cellA]]),]
  
  max_df_cellA<- colnames(max_df)[grepl("CellA", colnames(max_df))]
  max_df<-max_df[order(max_df[[max_df_cellA]]),]
  
  ## Add on number to CellB- just for merging purposes
  max_df$CellB<- paste0(1:nrow(max_df),"remove", max_df$CellB)
  min_df$CellB<- paste0(1:nrow(min_df), "remove", min_df$CellB)
  
  ## Merge
  df<- merge(max_df, min_df, by="CellB")
  
  ## Remove numbering from CellB
  df$CellB<- gsub('([0-9])+(remove)', "", df$CellB)
  df<-df[,c("CellA1", "CellB", "CellA2")]
  
  return(df)
})
test<- do.call("rbind", test)

## for ggaluvial
test$ID<- paste0(test$CellA1, "__", test$CellB, "__", test$CellA2)
test2<- as.data.frame(table(test$ID))
test2$CellA1<- sapply(test2$Var1, function(x){unlist(strsplit(as.character(x), split="__"))[1]})
test2$CellB<- sapply(test2$Var1, function(x){unlist(strsplit(as.character(x), split="__"))[2]})
test2$CellA2<- sapply(test2$Var1, function(x){unlist(strsplit(as.character(x), split="__"))[3]})

## Determine number of LRs for each pair- use to color plot
CellA_sender<- as.data.frame(table(paste0(test$CellA1, "__", test$CellB)))
CellB_sender<- as.data.frame(table(paste0(test$CellB, "__", test$CellA2)))

## Prep for sankey
df<- test %>% make_long(CellA1, CellB, CellA2) 
df$ID<- paste0(df$node, "__", df$next_node)
df$nLR<- ifelse(df$x=="CellA1", plyr::mapvalues(df$ID, CellA_sender$Var1, CellA_sender$Freq, warn_missing = FALSE),
         ifelse(df$x=="CellB", plyr::mapvalues(df$ID, CellB_sender$Var1, CellB_sender$Freq, warn_missing = FALSE),
                0))
df$nLR<- as.numeric(as.character(df$nLR))

df$color<- ifelse(df$nLR>30, "high", "low")
df$color<- ifelse(is.na(df$next_node) | is.na(df$next_x) | is.na(df$next_node), NA, df$color)

nWithColor<- 2
df$color<- c(rep("red", nWithColor), rep(NA, nrow(df)-nWithColor))
df$color<-c("red", NA, NA,"red",rep(NA,2),"red", rep(NA,2), rep(NA, 699))
df$color<- rep(c("red", NA, NA),236 )

ggplot(df, aes(x = x, 
               next_x = next_x, 
               node = node, 
               next_node = next_node,
               #fill = test,
               label=node)) +
  geom_sankey(aes(fill=factor(color))) +
  geom_sankey_label()+
  theme_sankey(base_size = 16)
    #theme(legend.position = "none")
      #scale_fill_gradient2(low="blue", mid="white", high="red", midpoint = mean(df$nLR))


```


```{r}
# Library
library(networkD3)
library(dplyr)
 
# A connection data frame is a list of flows with intensity for each flow
links <- data.frame(
  source=c("group_A","group_A", "group_B", "group_C", "group_C", "group_E"), 
  target=c("group_C","group_D", "group_E", "group_F", "group_G", "group_H"), 
  value=c(2,3, 2, 3, 1, 3)
  )
links<- sigLRs_nLR[,c("sender", "receiver", "nLRs_norm")] 

# From these flows we need to create a node data frame: it lists every entities involved in the flow
nodes <- data.frame(
  name=c(as.character(links$sender), 
  as.character(links$receiver)) %>% unique()
)
 
# With networkD3, connection must be provided using id, not using real name like in the links dataframe.. So we need to reformat it.
links$IDsender <- match(links$sender, nodes$name)-1 
links$IDreceiver <- match(links$receiver, nodes$name)-1

# Add a 'group' column to each connection:
links$group <- as.factor(c("type_a","type_a","type_a","type_b","type_b","type_b"))
 
# Add a 'group' column to each node. Here I decide to put all of them in the same group to make them grey
nodes$group <- as.factor(c("my_unique_group"))
 
# Give a color for each group:
my_color <- 'd3.scaleOrdinal() .domain(["type_a", "type_b", "my_unique_group"]) .range(["#69b3a2", "steelblue", "grey"])'
 
# Make the Network
p <- sankeyNetwork(Links = links, Nodes = nodes, Source = "IDsource", Target = "IDtarget", 
                   Value = "value", NodeID = "name", 
                   colourScale=my_color, LinkGroup="group", NodeGroup="group")

p

p <- sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsender", Target = "IDreceiver",
              Value = "nLRs_norm", NodeID = "name", 
              sinksRight=FALSE)
p
```

## 5. Summary plots for LRs of interest
```{r}
## Set major sender/receiver cell types +  ligand/receptors of interest
majorSender<- "Tumor"
majorReceiver<- "Myeloid" ## Tcell, CD4, CD8, or Myeloid
LR_list<- list(Tumor.Tcell.CD4.CD8=c("PVR_TIGIT","PROS1_AXL","CXCL16_CXCR6","CDH1_KLRG1","ANGPTL2_LILRB3", "CLEC2D_KLRB1"),
               Tumor.Myeloid=c("CCL5_CCR1","PROS1_AXL","ANGPTL2_LILRB3"))
LRs<- unlist(unname(LR_list[grepl(majorReceiver, names(LR_list))]))

## If using Tcell, remove clonality info
sigLRs_plot<- sigLRs
sigLRs_plot$sender<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$sender))
sigLRs_plot$receiver<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$receiver))


## Subset to all Tcells, CD4, or CD8
if(majorReceiver=="Tcell"){
  sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver_major==majorReceiver,]
}else{
  sigLRs_plot<-sigLRs_plot[sigLRs_plot$sender_major==majorSender & sigLRs_plot$receiver==majorReceiver,]
}

## Subset to LR pairs of interest
sigLRs_plot$LR<- paste0(sigLRs_plot$ligand, "_", sigLRs_plot$receptor)
sigLRs_plot<- sigLRs_plot[sigLRs_plot$LR %in% LRs,]

## If using Tcell, remove clonality info
sigLRs_plot$sender<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$sender))
sigLRs_plot$receiver<- gsub("_nonclonal", "", gsub("_clonal", "", sigLRs_plot$receiver))

## Add column for unique ID- by sample/ligand/receptor
sigLRs_plot$UniqueID<- paste0(sigLRs_plot$Sample, "_", sigLRs_plot$ligand, "_",sigLRs_plot$receptor)

## remove duplicate interactions
sigLRs_unique<- merge(aggregate(Prob ~ UniqueID, sigLRs_plot, max), sigLRs_plot)
sigLRs_unique<- sigLRs_unique[!duplicated(sigLRs_unique$UniqueID),] ## removes any total duplicates (identical prob too)

## Set variable to facet wrap by- usually will be subtype -  and plot
sigLRs_unique$FacetWrap1<- sigLRs_unique$Subtype

PlotPvalueProb(sigLRs_unique,xAxis = "Sample", labelXAxis = TRUE, facetWrap = c("Subtype"))+
  theme(axis.text = element_text(face="bold"))+
  ggtitle("")+
  theme(panel.grid.major.x = element_blank() ,panel.grid.major.y = element_line( size=.1, color="black" ) )+
  ggtitle(paste0("Sender: ", majorSender, "\nReceiver: ", majorReceiver))


## get appropriate width/height
height<- 3+length(unique(sigLRs_unique$LR))*.1
width<- 5+length(unique(sigLRs_unique$Sample)) * .25

## save
ggsave(paste0(figure_dir, "LR.summary_Sender.", majorSender, "_Receiver.", majorReceiver, ".png"), width=width, height=height)
```



