---
title: "Plot cellchat"
output: html_document
---

## Plot cell chat results between immune programs, tumor cell types

```{r}
library(CellChat)
library(NMF)
library(circlize)
library(patchwork)
library(Seurat) 
library(ggh4x)
library(ggalluvial)

options(stringsAsFactors = FALSE)

base_dir = "/Users/jlabelle/"
script_dir = paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/")
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/ImmuneCellAnalysis/04a_Integrate_Immune.Tumor/CellChat/")
analysis_dir<- paste0(working_dir, "/analysis/byProgram/")
figure_dir<- paste0(working_dir, "/figures/byProgram_Simplified/")

if(!dir.exists(figure_dir)){dir.create(figure_dir)}

cellchatHelper<- paste0(script_dir, "CellChat_HelperFunctions.R")
source(cellchatHelper)

setwd(working_dir)

## Load in list of all cell types across all subtypes
nCellTypes<- readRDS( "analysis/nCellTypes_bySample.Rds")


library(RColorBrewer)
myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
```

## Settings
```{r}
## Remove BCOR?
remove_bcor<- TRUE

## Remove epn?
remove_epn<- TRUE

## Just epn (for supplemental)
just_epn<- FALSE

## Set new figure dir
if(remove_bcor & remove_epn & !just_epn){
  figure_dir_use<- paste0(figure_dir, "removeBCOR.EPN/")
  if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}
} else if(just_epn){
  figure_dir_use<- paste0(figure_dir, "onlyEPN/")
  if(!dir.exists(figure_dir_use)){dir.create(figure_dir_use)}
}
print(paste0("FIGURE DIRECTORY IN USE: ", figure_dir_use))
```

## Load in LR results for all analyses (already merged)
```{r}
allLRs<- readRDS(paste0(analysis_dir, "all_LRs.Rds")) 

## Subset to just sig- by pvalue and prob
sigLRs<- allLRs[allLRs$pvalue<=0.1 & allLRs$Prob>0,]

## Add spacing to subtypes for better plotting
sigLRs$Subtype<- gsub("Histone", "\nHistone\n", gsub("Ependymoma", "EPN", sigLRs$Subtype))

## Remove bcor/epn if needed
if(remove_bcor){
  sigLRs<- sigLRs[sigLRs$Subtype != "hemiBCOR",]
}

if(remove_epn & !just_epn){
  sigLRs<- sigLRs[sigLRs$Subtype != "EPN",]
}

if(just_epn){
  sigLRs<- sigLRs[sigLRs$Subtype == "EPN",]
}

```



## 2 layers of cell type classification:
"major" cell type = CD4, CD8, Myeloid, or Tumor
"minor" cell type = Tumor cell type (AC-like, OPC-like), myeloid program, CD4 program, CD8 program
  

## Set minor cell types within major cell type, add onto sigLRs
```{r}
cd4_types<- unique(sigLRs$sender)[grepl("CD4", unique(sigLRs$sender))]
cd8_types<- unique(sigLRs$sender)[grepl("CD8", unique(sigLRs$sender))]
myeloid_types<- c("IFN_TAM", "Monocytes", "Inflammatory", "Hypoxic_TAM", "Stress_Response", "SEPP1_Mo_TAM")
tumor_types<- unique(sigLRs$sender)[!unique(sigLRs$sender) %in% c(cd4_types, cd8_types, myeloid_types)]

sigLRs$sender_major<- ifelse(sigLRs$sender %in% tumor_types, "Tumor",
                      ifelse(sigLRs$sender %in% myeloid_types, "Myeloid",
                      ifelse(sigLRs$sender %in% cd4_types, "CD4", "CD8")))
sigLRs$receiver_major<- ifelse(sigLRs$receiver %in% tumor_types, "Tumor",
                        ifelse(sigLRs$receiver %in% myeloid_types, "Myeloid",
                        ifelse(sigLRs$receiver %in% cd4_types, "CD4", "CD8")))

## Add \n to minor cell types for better plotting
sigLRs$sender<- gsub("_", "\n", sigLRs$sender)
sigLRs$receiver<- gsub("_", "\n", sigLRs$receiver)
```

## Simpliifed river plots needed:
tumor/myeloid/cd4/cd8
simplified tumor (ac, opc, oc, cycling)/myeloid

## Tumor/myeloid/cd4/cd8
```{r}
## Simplify sender to major cell type
sigLRs_tmp<- sigLRs
sigLRs_tmp$sender<- sigLRs$sender_major
sigLRs_tmp$receiver<- sigLRs$receiver_major


my_sankeyPlot(process_LR_forSankeyPlot(sigLRs_tmp, 
                                       c("Myeloid", "CD4", "CD8"), 
                                       c("Tumor"), 
                                       "noNorm"), 
              facet_grid = FALSE) 
ggsave(paste0(figure_dir_use, "Sankey_MajorCellTypes.png"), width=5, height=3)
```


simplified tumor (ac, opc, oc, cycling)/myeloid
```{r}
## Simplify tumor cell type
sigLRs_tmp<- sigLRs
sigLRs_tmp$sender<- gsub("GIPC-like", "Other", gsub("Neuronal-like", "Other", gsub("Ribo-Active", "Other", sigLRs_tmp$sender)))
sigLRs_tmp$receiver<- gsub("GIPC-like", "Other", gsub("Neuronal-like", "Other", gsub("Ribo-Active", "Other", sigLRs_tmp$receiver)))

## Simplify sender to major cell type (for immune only)
sigLRs_tmp$sender<- ifelse(sigLRs_tmp$sender_major != "Tumor", sigLRs$sender_major, sigLRs_tmp$sender)
sigLRs_tmp$receiver<- ifelse(sigLRs_tmp$receiver_major != "Tumor", sigLRs$receiver_major, sigLRs_tmp$receiver)



## Set order to plot by
sigLRs_plot<-process_LR_forSankeyPlot(sigLRs_tmp, 
                                       c("Myeloid", "CD4", "CD8"), 
                                       c("Tumor"), 
                                       "noNorm")
sigLRs_plot$sender<-factor(sigLRs_plot$sender, levels=c("CD4", "CD8", "Myeloid", "Cycling", "AC-like", "OPC-like", "OC-like", "Other"))

my_sankeyPlot(sigLRs_plot, 
              facet_grid = FALSE) 
ggsave(paste0(figure_dir_use, "Sankey_MajorCellTypes_tumorPrograms.png"), width=6, height=8)
```


