---
title: "preprocess CD162 sorted T cells"
author: "Jenna LaBelle"
date: "11/04/21"
output: html_document
---

QC and initial processing for CD162+/- T cells
```{r}
library(Seurat) 

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/",
                     "data_analysis/Projects/git_repos/Pedi_Immune/CD162_Sort/")

analysis_dir<- paste0(working_dir, "/analysis/")
figure_dir<- paste0(working_dir, "/figures/QC/")

if(!dir.exists(figure_dir)){dir.create(figure_dir, recursive = TRUE)};if(!dir.exists(analysis_dir)){dir.create(analysis_dir)}

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
nmf_helper = paste0(base_dir, script_dir, "NMF_helper_function.R")
source(nmf_helper)

## Read in housekeeping genes
marker_dir<- paste0(base_dir, "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Shared/Marker_genes/")
hkgenes <- read.table(paste0(marker_dir, "tirosh_house_keeping.txt"), skip = 2)
hkgenes <- as.vector(hkgenes$V1)
```

## Outputs
```{r}
raw_seurat<- paste0(analysis_dir, "gcdata_unfiltered.Rds")
metrics.out.file<- paste0(analysis_dir, "qc_metrics.txt")
filtered_seurat<- paste0(analysis_dir, "gcdata_filtered.Rds")
```

## Read in tpm/qc
```{r}
cm<- read.csv("data/cm_tpm_rsem.csv", row.names="gene_id")


```


## Initialize Seurat object with TPM data
```{r}
## Create samplesheet with metadata on each sample
samplesheet<- data.frame(sampleid="BT2074",
                         tumor=c("BT2074"),
                         condition="Fresh",
                         sort="Hold", ## will add this after creating seurat object
                         steroids="NA",
                         sex=c( "NA"),
                         age=c("NA"),
                         source=c("NA"),
                         group=c("NA"),
                         prefix="BT2074")

## Create intitial seurat object
gcdata <-CreateSeuratObject(counts = cm, min.cells = 0, min.features = 0, project = "BT2074")

# Add metadata from the samplesheet to each Seurat object.
gcdata <- AddMetaData(object = gcdata, 
                      metadata = data.frame(row.names = colnames(gcdata), 
                                            "sampleid" = rep(samplesheet$sampleid, ncol(gcdata)), 
                                            "tumor" = rep(samplesheet$tumor, ncol(gcdata)), 
                                            "condition" = rep(samplesheet$condition, ncol(gcdata)), 
                                            "steroids" = rep(samplesheet$steroids, ncol(gcdata)), 
                                            "sex" = rep(samplesheet$sex, ncol(gcdata)), 
                                            "age" = rep(samplesheet$age, ncol(gcdata)), 
                                            "source" = rep(samplesheet$source, ncol(gcdata)), 
                                            "group" = rep(samplesheet$group, ncol(gcdata)), 
                                            "prefix" = rep(samplesheet$prefix, ncol(gcdata))))

## Add on sort info
gcdata$sort<- sapply(colnames(gcdata), function(x){
  gsub("BT2074", "",unlist(lapply(strsplit(x, split="_"), function(y){y[1]})))
})

n.cells.beforeqc<- as.data.frame(table(gcdata$orig.ident))
colnames(n.cells.beforeqc)<- c("sampleid", "n.cells.before")

# Save current progress.
saveRDS(gcdata, file = raw_seurat)
```


## Basic QC- matches same QC as rest of cohort, by Orr
## Quality control metrics

Count number of UMIs per cell (these are TPM in SmartSeq2), number of genes expressed per cell, number of human housekeeping genes expressed per cell, and fraction mitochondrial gene expression per cell.

```{r quality, eval = T}
# Within each sample, get number of expressed housekeeping genes detected per cell barcode and
# fraction of mitochondrial gene expression.

  # Indices of hkgenes not present in gcdata.
  hkgenes.notfound.indices <- which(!hkgenes %in% rownames(gcdata))  
  if (length(hkgenes.notfound.indices)) {
    message = paste0(paste0(hkgenes[hkgenes.notfound.indices], collapse = " "), " housekeeping genes were not found in sample ")
    print(message)
  }
  
  # Get number of expressed housekeeping genes.
  hkgenes.found <- hkgenes[-hkgenes.notfound.indices]  # remove hkgenes that were not found
  gcdata[["nHK_RNA"]] <- Matrix::colSums(gcdata[['RNA']]@data[hkgenes.found, ] > 0)
  
  # Mitochondrial genes can be identified in USCS Table Browser using chrM search.
  gcdata[["percent_mito"]] <- PercentageFeatureSet(gcdata, pattern = "^MT-")
  
  print (paste0(sum(gcdata[["percent_mito"]] > 10), " of ", ncol(gcdata), " cells have greater than 10% mitochondrial gene expression"))



# Plot number of genes per cell, percent mitochondrial gene expression, and number of housekeeping genes. 
# This is done on a Seurat object merged across samples.
VlnPlot(gcdata, features = c("nFeature_RNA", "nHK_RNA"), ncol = 1, pt.size = 0.5) 
ggsave(paste0(figure_dir, "vln_nGene_nHK.png"), width = 6, height = 6)
```

## Filter cells
Remove cells that have a high percentage mitochondrial gene expression, or too few or too many genes detected. Remove cells from E30, E88, E119. This is all stored in Seurat @data slot. 
```{r filtercells, eval = T}
# Filter cells based on nGene, nHK, and percent.mito and store in @data slot 
gcdata <- subset(gcdata, subset = nFeature_RNA > 500 & nFeature_RNA < 5000 & nHK_RNA > 20)
```

## Quality control metrics after filtering
These are a few quality control metrics after the final filtering of genes and cells. We look at number of cells, number of genes per cell, and fraction of reads mapping to different genomic regions.
```{r qc_postfilter, eval = T}
# Number of cells in each experiment after filtering.
table(Idents(gcdata))

# Group cells by their experimental sample and find number of cells and median genes per sample after filtering. 
qc <- gcdata@meta.data %>% group_by(orig.ident) %>% summarise(n.cells.after = n(), med.nGene = median(nFeature_RNA)) 
qc$sampleid<- qc$orig.ident

# Include number of cells per sample before filtering and sort samples by their original order.
qc <- merge(n.cells.beforeqc, qc, by = "sampleid") 
rownames(qc) <- qc$sampleid
print(qc)
write.table(qc, metrics.out.file, sep = "\t", quote = F)

# Total number of cells before and after filtering
sum(qc$n.cells.before)
sum(qc$n.cells.after)

# Save current progress.
saveRDS(gcdata, file=filtered_seurat)
```

## Plot PF cells
```{r}
pf_cells<- as.data.frame(table(gcdata$orig.ident))
colnames(pf_cells)<- c("Sample", "Pass")
pf_cells$Fail<- n.cells.beforeqc$n.cells.before - pf_cells$Pass
pf_cells<- melt(pf_cells, id="Sample")
colnames(pf_cells)<- c("Sample", "PF", "Ncells")
pf_cells$PF<- factor(pf_cells$PF, levels = c("Fail", "Pass"))

ggplot(pf_cells, aes(x=Sample, y=Ncells, fill=PF))+
  geom_bar(stat="identity", position="stack")
ggsave(paste0(figure_dir, "PF_bySample.png"), width=4, height=4)
```

## Run full seurat pipeline
```{r}
gcdata_meta<- gcdata@meta.data
gcdata<- RunFullSeurat(gcdata@assays$RNA@counts, samples=gcdata$sampleid, RunHarmony = FALSE, dims=18, pca_dims = 18, n.neighbors = 17)
gcdata<- AddMetaData(gcdata, gcdata_meta[,!colnames(gcdata_meta) %in% colnames(gcdata@meta.data)])

saveRDS(gcdata, file=filtered_seurat)
```


## Plot SELPLG
```{r}
gcdata<- readRDS(filtered_seurat)

VlnPlot(gcdata, group.by = "orig.ident", features=c("SELPLG", "CD3G", "CD3E", "CD3D"), ncol=4)
ggsave(paste0(figure_dir, "GOI_pos.vs.neg.png"), width=7, height=3)

DoHeatmap(gcdata,group.by = "orig.ident", features="SELPLG")+
  scale_fill_gradient2(low="blue", mid="white", high="red")
ggsave(paste0(figure_dir, "Heatmap_goi_pos.vs.neg.png"), width=8, height=3)
```


