---
title: "ImmuneRegev_MidlineHistoneMutant_clustering"
output: html_document
---

```{r read-data}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/BySubtype/Midline/Midline_HistoneMutant/WithNewSample/")

library(Seurat, lib.loc = "C:/Users/jenna/OneDrive/Documents/R/win-library/4.0/Seurat/V4") 

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
stat_helper = paste0(base_dir, script_dir, "common_stat_test.R")
source(preprocessing)
source(stat_helper)

nmf_analysis_folder = "analysis/nmf/Merge/"
liger_analysis_folder = "analysis/liger/"
seurat_analysis_folder = "analysis/seurat/"
seurat_fig_folder = "figures/seurat/"
for (folder in c(seurat_analysis_folder, seurat_fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder,recursive = TRUE)
  }
}

## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

subtypes<- read.csv("../../../../../NotUploaded/Subtypes.csv")
```

## Load pre-computed seurat object 
```{r load-seurat-obj}
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj_Cluster6Removed.Robj")
if(file.exists(seurat_filename)){
  load(seurat_filename)
}

seurat_obj<-seurat_obj_annot
```

## Load counts and metadata for initiating seurat object
```{r load-data}
input_dir<- "../../../../Preprocessing/SplitBySubtype/analysis/"
load(paste0(input_dir, "midlineHistoneMut_cm_list.Robj"))
samples = readRDS(paste0(input_dir,"midlineHistoneMut_samples.rds"))
tumor_stat = readRDS(paste0(input_dir,"midlineHistoneMut_cluster6CellsNormal_call_tumor.rds"))
cm_list<-midlineHistoneMut_cm_list

sum(colnames(cm_list$raw_data)==names(samples)); length(samples)
sum(names(samples)==names(tumor_stat))
```

## Subset cm to remove noise/normal samples
```{r subset-tumor-cells}
cm = cm_list$raw_data; rm(cm_list); gc()
cm = cm[,tumor_stat=="Malignant"]
samples = samples[tumor_stat=="Malignant"]

any(colnames(cm) != names(samples))
```

## Initiate seurat obj and add metadata 
```{r seurat_obj_meta_data}
## Create seurat project
seurat_obj = preprocessSeuratObject(cm, project="", min.cells=0, min.genes=0, scale.factor=1E5, do.scale=F, do.center=T)

## Add samples and subtype as meta.data 
seurat_obj@meta.data$sample = samples

sample_df<- as.data.frame(samples); sample_df$cell<- rownames(sample_df)
colnames(sample_df)<- c("sample", "cell")
subtype_df<- merge(sample_df, subtypes, by="sample")

seurat_obj@meta.data$subtype<- subtype_df$subtype
seurat_obj@meta.data$AdditionalSubtype<- subtype_df$Diagnosis
```

## OPTIONAL: plot highly variable genes
```{r}
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## PCA and PC selection
```{r PCA_selection}
## PCA
seurat_obj <- RunPCA(
  object = seurat_obj, 
  features = seurat_obj@assays$RNA@var.features, 
  npcs = 100,
  verbose = TRUE, 
  ndims.print  = 1:5, 
  nfeatures.print = 5
)

```

## Harmony data integration 
```{r run_harmony}
library(harmony)
library(magrittr)
library(cowplot)

## Run Harmony
seurat_obj = RunHarmony(seurat_obj, "sample", theta = 2, 
                        max.iter.harmony = 50, plot_convergence = TRUE)

## Simple dim and vlnplot to examine integration 
options(repr.plot.height = 6, repr.plot.width = 12)

p1 <- DimPlot(object = seurat_obj, reduction = "pca", pt.size = .1, group.by = "sample")
p2<- DimPlot(object = seurat_obj, reduction = "harmony", pt.size = .1, group.by = "sample")
p3 <- VlnPlot(object = seurat_obj, features = "PC_1", pt.size = .1, group.by = "sample")
p4 <- VlnPlot(object = seurat_obj, features = "harmony_1", pt.size = .1, group.by = "sample")

plot_grid(p1,p2, p3, p4)
ggsave("Harmony_diagnostics.jpg", path=seurat_fig_folder, width=12, height=8)


## DimHeatmap
DimHeatmap(object = seurat_obj, reduction = "harmony", cells = 500, dims = 1:6, balanced = TRUE)
```

## Clustering and UMAP
## Without new sample (BT1873) there were 9 clusters (prior to removing cluster 6 which is likely normal)
## See that same "normal" cluster with the new sample. Prior to removing it, want to aim for 9 clusters
## Resolution of 0.6 gives 9 clusters (with normal cluster included)

## Without new samples and AFTER removing cluster 6: also 9 clusters, 8 clusters if you want less resolution on OPC-like types
## With new samples: resolution of 0.8 gives 9 clusters
```{r clustering_tSNE}
## Run UMAP and clustering using reduced dimensions from Harmony
## You may play around with resolution parameter to tune the clustering result
resolution = .5
reduction_type = "harmony"
dims = 20

## Run UMAP
seurat_obj <- RunUMAP(seurat_obj, reduction = reduction_type, dims = 1:dims)


seurat_obj <- FindNeighbors(seurat_obj,
                            reduction = reduction_type,
                            dims = 1:dims,
                            force.recalc = TRUE)   %>% 
              FindClusters(resolution = resolution)


r0.8<-DimPlot(seurat_obj, reduction="umap", pt.size = 1.5) + ggtitle("Resolution_0.8") 
r0.5<-DimPlot(seurat_obj, reduction="umap", pt.size = 1.5)+ ggtitle("Resolution_0.5")
r1.2<-DimPlot(seurat_obj, reduction = "umap", pt.size = 1.5)+ ggtitle("Resolution_1.2")

r0.5+r0.8+r1.2
ggsave("UMAP_clustering_VaryResolution.jpg", path=seurat_fig_folder, width=20, height=6)

## Clustering results 
table(Idents(seurat_obj))


## Save the seurat object
save(seurat_obj, file = paste0(seurat_analysis_folder, "seurat_obj_Cluster6Removed.Robj"))
```

## UMAP plot colored by clustering and samples
## Can color tSNE with other metadata features (e.g. primary/recurrence, treatment)
```{r UMAP_plots}
## By clustering
DimPlot(object = seurat_obj, reduction="umap", group.by = "seurat_clusters", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_Resolution0.5.jpg", path=seurat_fig_folder, width=8, height=8)


## By clustering- clusters 0/2 merged
DimPlot(object = seurat_obj, reduction="umap", group.by = "MergedClusters", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_Theta0.9_Clusters1.2Merged.jpg", path=seurat_fig_folder, width=8, height=8)

## By samples
DimPlot(object = seurat_obj, reduction="umap", group.by = "sample", label = FALSE, 
        pt.size = 2, label.size = 8) + NoAxes() 
ggsave("UMAP_sample.jpg", path=seurat_fig_folder, width=10, height=8)

## By subtype
DimPlot(object = seurat_obj, reduction="umap", group.by = "subtype", label = FALSE, 
        pt.size = 2, label.size = 8) + NoAxes() 
ggsave("UMAP_subtype.jpg", path=seurat_fig_folder, width=10, height=8)


```


## Plot breakdown of each cluster by sample
```{r}
cluster_by_sample<- table(seurat_obj@meta.data$sample, seurat_obj@meta.data$seurat_clusters)
cluster_by_sample<- melt(cluster_by_sample)
colnames(cluster_by_sample)<- c("Sample", "Cluster", "NCells")
cluster_by_sample$Cluster<- paste0("Cluster", cluster_by_sample$Cluster)

ggplot(cluster_by_sample, aes(x=Cluster, y=NCells, fill=Sample))+ 
  geom_bar(position="stack", stat="identity")+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Seurat clustering: \nbreakdown of samples per cluster")
ggsave(paste0(seurat_fig_folder, "SeuratClusterSampleDistr_Res0.5.png"), width=5, height=3)
```


## Cluster6 may be normal- cluster closely (but not with) OC in inferCNV
## CNVs for cells in this inferCNV cluster were double checked, iffy for MUV86/86
## Originally were called as malignant, now thinking they might not be
## Remove these cells from the seurat object --> rerun PCA/harmony/UMAP
```{r}
## Save metadata denoting cluster 6- use to subset tumor_call object in SubsetBySubtype
SeuratCluster6_PossiblyNormal<-seurat_obj@meta.data[seurat_obj@meta.data$seurat_clusters=="6",]
saveRDS(SeuratCluster6_PossiblyNormal, file= "analysis/SeuratCluster6_PossiblyNormal.Rds")
```

## DEGs for each cluster 
```{r differential_gene_expressions}
all_markers = FindAllMarkers(seurat_obj, only.pos = TRUE, 
                             logfc.threshold = .25, min.pct = 0.3)
all_markers_filtered = all_markers[all_markers$p_val_adj < 0.05 & all_markers$avg_logFC > .25, ]

## Cluster 8 has very few DEGs- should probably be removed or merged (resolution 0.9)


## Clusters 1/5 are very similar- DEGs between? (resolution 0.9)
## Add column with cluster info (based on logFC), then convert all logFC to positive
cluster1v5<- FindMarkers(seurat_obj, ident.1 = 1, ident.2 = 5, min.pct = 0.25)
cluster1v5_filtered<- cluster1v5[cluster1v5$p_val_adj<0.05 | abs(cluster1v5$avg_logFC)>1,]
cluster1v5_filtered$cluster<- ifelse(cluster1v5_filtered$avg_logFC>0, "1", "5")
cluster1v5_filtered$avg_logFC<- abs(cluster1v5_filtered$avg_logFC)
cluster1v5_filtered$gene<- rownames(cluster1v5_filtered)
saveRDS(cluster1v5_filtered, file="analysis/seurat/cluster1v5Markers_Resolution0.9.rds")
write.csv(cluster1v5_filtered, file=paste0(seurat_analysis_folder,
                                        "FilteredMarkers_Cluster1vsCluster5.csv"))



## Subset the top 30 DEGs for each cluster
num_marker_genes = 30
top_marker_genes<- all_markers_filtered %>% 
                   group_by(cluster) %>% 
                   top_n(num_marker_genes, avg_logFC) %>%
                   as.data.frame()
save(all_markers_filtered, file=paste0(seurat_analysis_folder,
                                       "all_markers_Theta0.5.Robj"))
write.csv(top_marker_genes, file=paste0(seurat_analysis_folder,
                                        "Top30Markers_Theta0.5.csv"))


## Heatmap for initial inspection
DoHeatmap(seurat_obj, 
          features=all_markers_filtered$gene) + NoLegend() + NoAxes()
ggsave("Heatmap_all_degs_Theta0.6.jpg", path=seurat_fig_folder)
```

## Quick annotation by OPC/OC/AC marker genes 
```{r}
## OPC-like
FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OPC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## OC-like cells
FeaturePlot(seurat_obj,
            c("BCAS1", "SIRT2", "RGR", 
              "GPR17", "NFASC", "MBP",
              "TF", "MOG", "PLP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("OC_markers.jpg", path=seurat_fig_folder, width = 9, height=9)

## AC-like cells 
FeaturePlot(seurat_obj,
            c("APOE", "AQP4", "AGT", 
              "SPARCL1","ATP1A2", "GFAP",
              "S100B", "ALDOC", "CLU"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("AC_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## Neuronal-like cells
FeaturePlot(seurat_obj,
            c("STMN2", "STMN1", "STMN4", 
              "TUBB3", "INA", "ELAVL4",
              "SNAP25", "GABRG2", "ST18"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("Neuron_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## FOS/JUN
FeaturePlot(seurat_obj,
            c("FOS", "FOSB",
              "JUN", "JUNB"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("FOS.JUN_markers.jpg", path=seurat_fig_folder, width=9, height=9)

## pre-OPC markers
FeaturePlot(seurat_obj,
            c("SOX1", "ASCL1"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("PreOPC_markers.jpg", path=seurat_fig_folder, width=9, height=4.5)

## mature OPC markers
FeaturePlot(seurat_obj,
            c("PDGFRA", "OLIG1", "OLIG2"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("PreOPC_markers.jpg", path=seurat_fig_folder, width=9, height=4.5)

FeaturePlot(seurat_obj,
            c("A2B5", "PDGFRA", "NG2", "O4", "o1", "CNPase", "GalC", "MBP", "MAG", "MOG"),
            cols = c("lightgrey", "red"),
            ncol = 2)
ggsave("PreOPC_markers.jpg", path=seurat_fig_folder, width=9, height=4.5)

## inhibitory ligands identified in CellChat
FeaturePlot(seurat_obj,
            c("CLEC2D", "WNT5A", "PVR", "CDH1", "SPP1"),
            cols = c("lightgrey", "red"),
            ncol = 3)
ggsave("InhibitoryLigand_markers.jpg", path=seurat_fig_folder, width=9, height=6)
```

## Violin plot
```{r}
VlnPlot(seurat_obj, 
        c("BCAS1", "SIRT2", "RGR", 
          "GPR17", "NFASC", "MBP", 
          "TF", "MOG", "PLP1"),
        pt.size = 0)
ggsave("OC_markers_ViolinPlot.jpg", path=seurat_fig_folder, width=9, height=9)

VlnPlot(seurat_obj, 
       c("PDGFRA", "OLIG1", "APOD",
              "CCND1", "CD24", "ASCL1",
              "ETV1", "SEZ6L", "HES6"),
        pt.size = 0)
ggsave("OPC_markers_ViolinPlot.jpg", path=seurat_fig_folder, width=9, height=9)

VlnPlot(seurat_obj, group.by = "sample",
       c("CLEC2D", "WNT5A", "PVR", "CDH1", "SPP1"),
        pt.size = 0, sort = TRUE)
ggsave("InhibitoryLigand_markers_ViolinPlot_samples.jpg", path=seurat_fig_folder, width=9, height=6)
```

## ORA test 
```{r over-representation-test}
library(clusterProfiler)

##
OrgDb = "org.Hs.eg.db"
dataset = "hsapiens_gene_ensembl"

## Convert using biomart
allOE_genes <- as.character(gene_symbol_to_ensembl_id(rownames(seurat_obj@assays$RNA@counts),dataset)$ensembl_gene_id)
save(allOE_genes, file="analysis/seurat/allOE_genes.Robj")

save(all_markers_filtered, file="analysis/seurat/all_markers_filtered_temp.Robj")

## Convert marker genes to ensembl IDs too, then run OR analysis (GO)
sigOE_genes_list<-list()

all_markers_filtered<-all_markers_filtered[all_markers_filtered$cluster==0,]
all_markers_filtered$cluster<- as.character(all_markers_filtered$cluster)
for (i in unique(all_markers_filtered$cluster)){
  sigOE_genes = as.character(gene_symbol_to_ensembl_id(subset(all_markers_filtered,
                                                              cluster==i,gene)[,],
                                                       dataset)$ensembl_gene_id)
  print(head(sigOE_genes))
  sigOE_genes_list[[i]]<-sigOE_genes
  
} 
go_result = list()
for (i in names(sigOE_genes_list)){
  sigOE_genes<-sigOE_genes_list[[i]]
  go_result[[paste0("c", i)]] = go_analysis(sigOE_genes, allOE_genes)
}


## Plot top pathways for all clusters, display on one plot
clusters<-names(go_result)
allDotPlots_20<-list()
for (c in 1:length(clusters)){
  cluster=clusters[c]
  print(cluster)
  if (nrow(go_result[[cluster]]$cluster_summary)>0){
    print(cluster)
    allDotPlots_20[[c]]<-dotplot(go_result[[cluster]]$ego, 
                                showCategory=15, 
                                font.size = 15, 
                                title = cluster, 
                                label_format=10) +
      scale_y_discrete(labels=function(x) str_wrap(x, width=60))+
      theme(plot.title = element_text(hjust = 0.5, face = "bold", 
                                      color="black", size = 28),
            axis.title = element_text(face = "bold", color="black"), 
            axis.text.x = element_text(angle = 45, hjust = 1, color="black"),
            axis.text.y = element_text( color="black",face = "bold"))
  }
  
}

allDotPlots_20<- allDotPlots_20[!(unlist(lapply(allDotPlots_20, function(x){is.null(x)})))]
cowplot::plot_grid(plotlist = allDotPlots_20, ncol=2)
ggsave(paste0("GO_OVA_AllClusters_top20_Res0.5.jpg"), path=seurat_fig_folder, width=18, height=25)
save(go_result, file=paste0(seurat_analysis_folder,  "go_result_Resolution0.5.Robj"))

```


## ORA with ependymoma genesets
## Compare with metaprograms from previously defined metaprograms from ependymoma
```{r}
all_markers_list<-split(all_markers_filtered, all_markers_filtered$cluster )
all_markers_list<-lapply(all_markers_list, function(x){x$gene})
names(all_markers_list)<-paste0("Cluster", names(all_markers_list))

midline_genesets<-readRDS(paste0(nmf_analysis_folder,"midlinePreviousPrograms_OC.AC.OPC.MES1.rds"))


tmp2 = sapply(all_markers_list, function(x){
    sapply(midline_genesets, function(y){
        fisher_test(x, y, unlist(union(midline_genesets, all_markers_list)))
    })
})
tmp2 = data.frame(tmp2)
tmp3<- apply(tmp2, 2, function(x) {p.adjust(x) < 0.1})
tmp4<-apply(tmp2, 2, function(x) {p.adjust(x)})

write.table(tmp4, paste0(seurat_analysis_folder,
                        "ScoreSeuratClustersUsingPreviousNMF_Theta0.9_padj.txt"))

## Alternatively: fisher test with de novo NMF programs
NMF_denovo<- read.delim("analysis/nmf/Merge/metaprogram_genesMerge4_cluster6CellsNormal.txt")
NMF_denovo<-as.list(NMF_denovo)

tmp2 = sapply(all_markers_list, function(x){
    sapply(NMF_denovo, function(y){
        fisher_test(x, y, unlist(union(NMF_denovo, all_markers_list)))
    })
})
tmp2 = data.frame(tmp2)
tmp3<- apply(tmp2, 2, function(x) {p.adjust(x) < 0.1})
tmp4<-apply(tmp2, 2, function(x) {p.adjust(x)})

write.table(tmp4, paste0(seurat_analysis_folder,
                        "ScoreSeuratClustersUsingDenovoNMF_Theta0.9_padj.txt"))
```



# Add NMF results to the seurat object

## Map NMF scores to seurat object 
```{r add-NMF-score}
NameOfMetaprogramList<-"Merge4_cluster6CellsNormal"

nmf_score = readRDS(paste0("analysis/nmf/Merge/nmf_score", NameOfMetaprogramList, ".rds"))
seurat_obj = AddMetaData(seurat_obj, nmf_score)
load(paste0("analysis/nmf/Merge/nmf_marker_genes_final", NameOfMetaprogramList, ".Robj"))

DimPlot(object = seurat_obj, group.by = "signature_1", label = TRUE, 
         pt.size = 2, label.size = 8, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_nmf_signature",NameOfMetaprogramList, ".jpg"), 
       path=seurat_fig_folder, 
       width=14, height=12)

#DoHeatmap(seurat_obj, 
          #features=unlist(nmf_marker_genes_final)) + NoLegend()

#ggsave("nmf_marker_genes.jpg", path=seurat_fig_folder, width=12, height=8)
```

## Map LIGER iNMF scores to seurat object 
```{r add-NMF-score}
nmf_score = readRDS(paste0(liger_analysis_folder, "nmf_score_20Factors_No4", ".Rds"))
seurat_obj = AddMetaData(seurat_obj, nmf_score)

DimPlot(object = seurat_obj, group.by = "signature_1", label = TRUE, 
         pt.size = 2, label.size = 8, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_nmf_signature_LIGER_20Factors_No4", ".jpg"), 
       path=seurat_fig_folder, 
       width=14, height=12)

## Annotate factors
factors<-as.character(unique(seurat_obj@meta.data$signature_1))
annot_factors<- c("OPC-like", "AC-like", "Cycling", "OPC-like_Alt", "OC-like")
seurat_obj@meta.data$FactorAnnot<- plyr::mapvalues(seurat_obj@meta.data$signature_1,
                                                   from=factors, to=annot_factors)

DimPlot(object = seurat_obj, group.by = "FactorAnnot", label = TRUE, 
         pt.size = 2, label.size = 8, reduction="umap", cols = col_vector[10:28]) + NoAxes()
ggsave(paste0("UMAP_nmf_signature_LIGER_NoFactor7.9_ManualAnnot", ".jpg"), 
       path=seurat_fig_folder, 
       width=8, height=8)

## Set idents to NMF factors, then find markers between desired factors
seurat_obj_liger<-seurat_obj
Idents(seurat_obj_liger)<- "signature_1"
f18_v_f8<- FindMarkers(seurat_obj_liger, 
                       ident.1="factor18",
                       ident.2="factor8")
f18_v_f8_filtered<- f18_v_f8[f18_v_f8$p_val_adj<0.05 &
                                abs(f18_v_f8$avg_logFC)>1,]
write.csv(f18_v_f8_filtered, paste0(seurat_analysis_folder, "LIGER_factor18_vs_factor8.csv"))

#DoHeatmap(seurat_obj, 
          #features=unlist(nmf_marker_genes_final)) + NoLegend()

#ggsave("nmf_marker_genes.jpg", path=seurat_fig_folder, width=12, height=8)
```
```{r}
save(seurat_obj, file= paste0(seurat_analysis_folder, "seurat_obj.Robj"))
```

## Add "Shared" factor- only some cells with have shared factors
```{r}
load("../analysis/liger/TopLigerFactor_SharedScoreAbove2.Robj")
sharedScore_df<- semiShared_HighScores[semiShared_HighScores$cell %in% colnames(seurat_obj),]
sharedScore<- sharedScore_df$signature_1
names(sharedScore)<-sharedScore_df$cell

noSharedScore_cells<- colnames(seurat_obj)[!(colnames(seurat_obj) %in% names(sharedScore))]
noSharedScore<- rep("-", length(noSharedScore_cells))
names(noSharedScore)<- noSharedScore_cells

sharedScore_forSeurat<- c(sharedScore, noSharedScore)
sharedScore_forSeurat<- sharedScore_forSeurat[order(names(sharedScore_forSeurat))]

seurat_obj@meta.data$SharedFactor<-sharedScore_forSeurat
DimPlot(seurat_obj, group.by="SharedFactor", cols=c("grey", "magenta", "orange", "blue4", "blue", "orange", "yellow"), pt.size=2)
ggsave(paste0(seurat_fig_folder, "SharedPrograms_ScoreAbove2.png"))

```




## Plot OPC-like/OC scores for clusters 1/2/5/7
```{r}
clusters_1257<-seurat_obj@meta.data[seurat_obj@meta.data$seurat_clusters %in% c(1,2,5),]
clusters_1257<-clusters_1257[clusters_1257$signature_1 %in% c("OPC.like", "OC.like"),]

df3<-clusters_1257[,colnames(clusters_1257) %in% c("seurat_clusters", "OPC.like", "OC.like")]

ggplot(df3, aes(x=OC.like, y=OPC.like, color=seurat_clusters))+geom_point()
ggsave("figures/seurat/NMFOC.vs.NMFOPC_Clusters1.2.5_OnlyOPC.OCHighScores.png")
```


## Plot OPC-like/MES1 scores for clusters 1/2/5
```{r}
clusters_1256<-seurat_obj@meta.data[seurat_obj@meta.data$seurat_clusters %in%
                                      c(1,2,5,6),]
clusters_1256<-clusters_1256[clusters_1256$signature_1 %in% c("OPC.like", "MES1"),]

df2<-clusters_1256[,colnames(clusters_1256) %in% c("seurat_clusters", "OPC.like", "MES1")]

ggplot(df2, aes(x=MES1, y=OPC.like, color=seurat_clusters))+geom_point()
ggsave("figures/seurat/NMFMES.vs.NMFOPC_Clusters1.2.5.6.png")
```




## Add a second metagene classification without cell cycle programs
```{r}
nmf_score = readRDS(paste0("analysis/nmf/all_subtypes/new/nmf_score_updated", NameOfMetaprogramList,  ".rds"))
seurat_obj@meta.data$signaure_1_without_cc = nmf_score$signature_1
```

## Make pseudobulk
```{r}
## Split by cell type
metagene_program_names = sort(unique(nmf_score$signature_1))
pseudobulk = NULL
aggr_method = "mean"
for (metagene in metagene_program_names){
  cm = seurat_obj@assays$RNA@counts[,nmf_score$signature_1 == metagene & nmf_score$score_1 >= 1]
  if (aggr_method == "mean"){
    pseudobulk = cbind(pseudobulk, rowMeans(cm))
  }
  if (aggr_method == "median"){
    pseudobulk = cbind(pseudobulk, rowMedians(cm))
  }
}
colnames(pseudobulk) = metagene_program_names
rownames(pseudobulk) = rownames(seurat_obj@assays$RNA@counts)
saveRDS(pseudobulk, file=paste0(seurat_analysis_folder, 
                                "pseudobulk_cm_unique_only_", aggr_method, NameOfMetaprogramList,".RDS"))

## All cell types together- single pseudobulk for each sample
## Sum of counts across each sample
pseudobulk<- data.frame(gene=rownames(seurat_obj))
mySamples<- unique(seurat_obj$sample)
for (i in mySamples){
  seurat_tmp<- subset(seurat_obj, sample==i)
  cm_tmp<-seurat_tmp@assays$RNA@counts
  sumGene<- as.data.frame(rowSums(cm_tmp))
  colnames(sumGene)<- i
  sumGene$gene<- rownames(sumGene)
  pseudobulk<- merge(pseudobulk, sumGene, by="gene")
}

rownames(pseudobulk)<- pseudobulk$gene
pseudobulk<- pseudobulk[,-1]

save(pseudobulk, file="../data/PseudobulkCounts/MidlineHistoneMutant_Pseudobulk.Robj")
```

## Metagene program specific genes as overall DEGs
```{r}
seurat_obj = SetIdent(seurat_obj, value = seurat_obj@meta.data$signaure_1_without_cc)

all_markers_nmf = FindAllMarkers(seurat_obj, only.pos = TRUE, logfc.threshold = 1, min.pct = 0.5)
all_markers_nmf = all_markers[all_markers$p_val_adj < 0.05, ]
saveRDS(all_markers, paste0(seurat_analysis_folder, "deg.rds"))
##sapply(names(nmf_marker_genes_final), function(x) table(nmf_marker_genes_final[[x]] %in% gene_signatures[[x]]))
signature_genes = unique(all_markers$gene)
saveRDS(signature_genes, paste0(seurat_analysis_folder, "signature_genes.rds"))
```


## Integrating NMF/seurat annotations manually
## Two options: high level (OC/OPC/AC/MES/cycling) and low level (all clusters)

## LOW LEVEL:
## Using resolution of 0.9
## NMF annotations used, except not scoring NMF3/6 (Merge5)
## seurat annotations for clusters 1/5/7 used, overriding NMF scores
```{r}
## Make named list of nmf annotations
nmf_annot<-seurat_obj@meta.data$signature_1
names(nmf_annot)<- rownames(seurat_obj@meta.data)

## Make named vector of seurat annotations, but only for clusters 1/2/5/7
seurat_annot<- seurat_obj@meta.data$seurat_clusters
names(seurat_annot)<- rownames(seurat_obj@meta.data)
tmp<- seurat_annot[seurat_annot=="1" |
                                  seurat_annot=="5" |
                                  seurat_annot =="7"| seurat_annot=="2"]
seurat_annot_1257<- paste0("Cluster", tmp)
names(seurat_annot_1257)<-names(tmp)

merged_annot<-nmf_annot[!(names(nmf_annot) %in% names(seurat_annot_1257))]
merged_annot<-c(merged_annot, seurat_annot_1257)

merged_annot<- gsub("Cluster1", "OPC-like",
                    gsub("Cluster2", "OPC-like_Ribosomal",
                         gsub("Cluster5", "OPC.OC.Transitional",
                              gsub("Cluster7", "OC-like",merged_annot ))))

merged_annot<-merged_annot[order(names(merged_annot))]

seurat_obj@meta.data$ManualAnnotation<-merged_annot

DimPlot(object = seurat_obj, reduction="umap", group.by = "ManualAnnotation", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_ManualAnnotation_Detailed.jpg", path=seurat_fig_folder, width=8, height=8)
```


## HIGH LEVEL
## Using resolution of 0.8
## Clusters 1/2/0 merged manually
## NMF annotations used, except not scoring NMF1/3/6 (Merge6)
## seurat annotations for clusters 1/5 used, overriding NMF scores
```{r}
## Merge clusters 1/2/0/6/5
newIdents<-gsub("2", "1", gsub("0", "1", gsub("6", "1", 
                                              gsub("5", "1", seurat_obj@active.ident))))
names(newIdents)<- names(seurat_obj@active.ident)

seurat_obj<-SetIdent(seurat_obj, value=newIdents)
seurat_obj<-AddMetaData(seurat_obj, seurat_obj@active.ident, col.name = "MergedClusters")

## Can optionally stop here
clusters<-as.character(unique(seurat_obj@active.ident))
cluster_annot<-c("OPC-like", "AC-like", "SPhase", "G2M", "OC-like")
seurat_obj@meta.data$HighLevelAnnot<- plyr::mapvalues(seurat_obj@active.ident,
                                                      from=clusters,
                                                      to=cluster_annot)
DimPlot(object = seurat_obj, reduction="umap", group.by = "HighLevelAnnot", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_ManualAnnotation_HighLevel_NoMES.jpg", path=seurat_fig_folder, width=8, height=8)

## Make named list of NMF annotations
nmf_annot<-seurat_obj@meta.data$signature_1
names(nmf_annot)<- rownames(seurat_obj@meta.data)

## Make named list of seurat annotations, only for clusters 1 and 5
seurat_annot<- seurat_obj@meta.data$MergedClusters
names(seurat_annot)<- rownames(seurat_obj@meta.data)
tmp<- seurat_annot[seurat_annot=="1" |
                                  seurat_annot=="5"]
seurat_annot_15<- paste0("Cluster", tmp)
names(seurat_annot_15)<-names(tmp)

## Removed seurat annotated cells from nmf annotations, then merge together
merged_annot<-nmf_annot[!(names(nmf_annot) %in% names(seurat_annot_15))]
merged_annot<-c(merged_annot, seurat_annot_15)

## Rename seurat clusters as needed
merged_annot<- gsub("Cluster1", "OPC-like", gsub("Cluster5", "OC-like", merged_annot))

## Order same as seurat object, add to metadata, and plot
merged_annot<-merged_annot[order(names(merged_annot))]
seurat_obj@meta.data$ManualAnnotation<-merged_annot

DimPlot(object = seurat_obj, reduction="umap", group.by = "ManualAnnotation", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_ManualAnnotation_HighLevel.jpg", path=seurat_fig_folder, width=8, height=8)
```

## Annotate seurat clusters
## Here, using resolution of 0.5

```{r}
## Annotate clusters
clusters<- as.character(unique(seurat_obj@active.ident))
cluster_annot<- c("OPC-like_MES", "S-Phase", "OPC-like_Alt",  "AC-like","G2M", "OPC-like_Ribosomal", "OC-like")
newIdents<-plyr::mapvalues(seurat_obj@active.ident, from=clusters, to=cluster_annot)

seurat_obj_annot<-SetIdent(seurat_obj, value=newIdents)
seurat_obj_annot@meta.data$ManualAnnot<- newIdents
DimPlot(object = seurat_obj_annot, reduction="umap", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_Resolution0.5_ManualAnnot.jpg", path=seurat_fig_folder, width=8, height=8)

## Save
save(seurat_obj_annot, file=paste0(seurat_analysis_folder, "seurat_obj_Cluster6Removed.Robj"))
```

## Export marker genes for all cell types
```{r}
seurat_obj_annot<- SetIdent(seurat_obj_annot, value=seurat_obj_annot$ManualAnnot)
ManualCellAnnot_Markers<-FindAllMarkers(seurat_obj_annot, min.pct=.5)
ManualCellAnnot_Markers_filt<-ManualCellAnnot_Markers[ManualCellAnnot_Markers$p_val_adj<0.05
                                                      & ManualCellAnnot_Markers$avg_log2FC >0.5,]
save(ManualCellAnnot_Markers_filt, file="analysis/ManualCellAnnot_Markers.Robj")
```


## Compare annotations from original analysis
## Remove new sample, read in original seurat, map annot to current seurat obj
```{r}
seurat_backup<- seurat_obj
seurat_tmp<- subset(seurat_obj, sample != "BT1873vent")
seurat_tmp$Cell<- colnames(seurat_tmp)

load("../analysis/seurat/seurat_obj_Cluster6Removed.Robj")
seurat_obj_annot$Cell<- colnames(seurat_obj_annot)
seurat_obj_annot$ManualAnnot<- as.character(seurat_obj_annot$ManualAnnot)

## Add previous annots to current
seurat_tmp$OriginalAnnot<- plyr::mapvalues(seurat_tmp$Cell,
                       from=seurat_obj_annot$Cell,
                       to=seurat_obj_annot$ManualAnnot)
seurat_tmp<- subset(seurat_tmp, OriginalAnnot %in% seurat_obj_annot$ManualAnnot)
DimPlot(object = seurat_tmp, reduction="umap", group.by="OriginalAnnot", 
        label = TRUE, pt.size = 2, label.size = 8) + NoAxes() + NoLegend()
ggsave("UMAP_clustering_ComparisonToOriginalAnnot.jpg", path=seurat_fig_folder, width=8, height=8)
```



## For origianl samples, use "original" annotations. For new sample, use projected annotations
## Projections from Projection_OriginalToWithNewSample
```{r}
## Load projected seurat object
## Contains projected annots for all cells, but will only use those for the new sample
load(paste0(seurat_analysis_folder,"AnnotationsProjectedFromOriginal_seurat.Robj"))
projected_seurat<- current_seurat_obj; rm(current_seurat_obj)

## Subset to just new sample
projected_seurat<- subset(projected_seurat, sample=="BT1873vent")
projected_annot<- projected_seurat$Predictions; names(projected_annot)<- colnames(projected_seurat)

## Make list of annotations: projected for new, original for original samples
original_annot<- seurat_tmp$OriginalAnnot; names(original_annot)<- colnames(seurat_tmp)
all_annot<- c(projected_annot, original_annot)

## Add to seurat object
seurat_obj$Annot_ProjectedAndOriginal<- plyr::mapvalues(colnames(seurat_obj),
                                                        from=names(all_annot),
                                                        to=all_annot)
seurat_obj<- subset(seurat_obj, Annot_ProjectedAndOriginal %in% all_annot)
DimPlot(seurat_obj, group.by="Annot_ProjectedAndOriginal", label=TRUE, label.size = 6)+
  NoLegend()
ggsave(paste0(seurat_fig_folder, "OriginalSamples.OriginalAnnot_NewSample.ProjectedAnnot.png"),
              width=7, height=7)
seurat_obj<- SetIdent(seurat_obj, value=seurat_obj$Annot_ProjectedAndOriginal)
option3_markers<- FindAllMarkers(seurat_obj, min.pct = .5)
option3_filtered_markers<- option3_markers[option3_markers$p_val_adj<0.05 &
                                             option3_markers$avg_log2FC>0.5,]
```

## Split adult sample(s) into separate seurat object, remove from other samples
```{r}
## Read in seurat object with all samples
seurat_filename = paste0(seurat_analysis_folder, "seurat_obj_Cluster6Removed.Robj")
load(seurat_filename)

## Set adult sample(s)
adult_samples<-"MUV091"
ped_samples<- unique(seurat_obj$sample[!seurat_obj$sample %in% adult_samples])

## Create adult/ped seurat object
adult_seurat<- subset(seurat_obj, sample %in%adult_samples)
seurat_obj<- subset(seurat_obj, sample %in% ped_samples)

## Save- ped will be used for downstream, adult only for minimal comparison purposes
save(seurat_obj, file=paste0(seurat_analysis_folder, "seurat_obj_Cluster6Removed.Robj"))
save(adult_seurat, file=paste0(seurat_analysis_folder, "seurat_obj_adultSamples.Robj"))
```