---
title: "R Notebook"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/BySubtype/Midline/Midline_HistoneMutant/WithNewSample/")
library(Seurat) 

preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
stat_helper = paste0(base_dir, script_dir, "common_stat_test.R")
source(preprocessing)
source(stat_helper)


seurat_analysis_folder = "analysis/seurat/"
fig_folder = "figures/LabelTransferOptimization/"
for (folder in c(seurat_analysis_folder, fig_folder)){
  if (!dir.exists(folder)){
    dir.create(folder)
  }
}


## Set colors for use later- better for discriminating against groups
library(sparseMatrixStats)
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

subtypes<- read.csv(paste0(project_dir, "Subtypes.csv"))
```

## Load in original and current seurat objs
```{r}
load(paste0(seurat_analysis_folder, "seurat_obj_Cluster6Removed.Robj"))
current_seurat_obj<- seurat_obj_annot; rm(seurat_obj_annot)

load("../analysis/seurat/seurat_obj_Cluster6Removed.Robj")
original_seurat_obj<-seurat_obj_annot; rm(seurat_obj_annot)
```

## Wrapper function for transfering labels
```{r}
runTransferAnchors<- function(seurat_list, query_seurat_obj, ref_seurat_obj,
                              FTA_reduction, FTA_reference.reduction,
                              TD_refdata, TD_weight.reduction){
  
  transferAnchors<-FindTransferAnchors(reference = ref_seurat_obj, 
                                       query=query_seurat_obj,
                                       reduction=FTA_reduction,
                                       reference.reduction = FTA_reference.reduction)
  
  predictions <- TransferData(anchorset = transferAnchors, 
                              refdata = TD_refdata,
                              weight.reduction = TD_weight.reduction)
  
  predictions_final<- predictions$predicted.id; names(predictions_final)<- rownames(predictions)
  query_seurat_obj$Predictions<- predictions_final
  return(query_seurat_obj)
}
```

## Tests:
## Accuracy determined by % of cells that match between current/original (minus new sample)
1. query current seurat (MINUS new sample): reduction method optimization
2. query current seurat (MINUS new sample) with OPC-* collapsed: reduction method optimization
3. Compare "best" settings from #1/#2 to current seurat INCLUDING new sample
4. Compare "best" settings from #1/#2 to current seurat INCLUDING new sample, with OPC-* collapsed
5. Query full current seurat vs 1 sample (do for each sample independently). Use best reduction settings.

## Test 1: reduction method optimization
options for FindTransferAnchors: pp + pca, pp + harmony, rpca, cca (must run cca for transferData as well)
options for TransferData: pcaproject, cca

1. pp + pca / pp
2. pp + harmony / pp
3. cca / cca
4. rpca / pcaproject

```{r}
seurat_list<-list(current=subset(current_seurat_obj, sample != "BT1873vent"),
                  original=original_seurat_obj)

test1_results<-list()
## 1: pp+pca / pp
test1_results[["option1"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "pcaproject", FTA_reference.reduction = "pca",
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")
## 2: pp+harmony / pp
test1_results[["option2"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "pcaproject", FTA_reference.reduction = "harmony",
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")

## 3: cca / pp
test1_results[["option3"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "cca", FTA_reference.reduction = NULL,
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "cca")

## 4: rpca / pp
test1_results[["option4"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "rpca", FTA_reference.reduction = NULL,
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")

```

## Quantify accuracy of each of the above methods- what percent of cells match between original/current?
```{r}
percentAccurate<- lapply(test1_results, function(seurat_transfer){
  ## Subset to cells also found in original (only a few differences due to inferCNV)
  commonCells<- colnames(seurat_transfer)[colnames(seurat_transfer) %in% colnames(seurat_list$original)]
  seurat_transfer<-subset(seurat_transfer, cells=commonCells)
  seurat_original<- subset(seurat_list$original, cells=commonCells)
  print(sum(colnames(seurat_transfer) == colnames(seurat_original))); print(ncol(seurat_transfer)); print(ncol(seurat_original))
  
  perMatches<- round(sum(seurat_transfer$Predictions==seurat_original$ManualAnnot)/ncol(seurat_transfer),3)
  return(perMatches)
})

## bar plot of accuracy
df<-as.data.frame(t(as.data.frame(percentAccurate))); colnames(df)<- "PercentMatching"
df$settings<- c("pcaproject+pca\npcaproject",
                "pcaproject+harmony\npcaproject",
                "cca\ncca",
                "rpca\npcaproject")
ggplot(df, aes(x=settings, y=PercentMatching))+
  geom_bar(position="stack", stat="identity")+
  ylim(0,1)+
  xlab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"))+
  ggtitle("Percent accuracy in transfer vs original:\nReduction method optimization")+
  geom_hline(yintercept = max(df$PercentMatching), linetype="dotted", color="red")
ggsave(paste0(fig_folder, "ReductionMethodOptimization.png"), width=5, height=5)
```
 best is pp + pca, pp (88%)

## Test 2: reduction method optimization, with OPC-* collapsed
options for FindTransferAnchors: pp + pca, pp + harmony, rpca, cca (must run cca for transferData as well)
options for TransferData: pcaproject, cca

1. pp + pca / pp
2. pp + harmony / pp
3. cca / cca
4. rpca / pcaproject

```{r}
seurat_list<-list(current=subset(current_seurat_obj, sample != "BT1873vent"),
                  original=original_seurat_obj)

seurat_list$original$ManualAnnot<-gsub("OPC-like_Ribosomal", "OPC-like", 
                                  gsub("OPC-like_MES", "OPC-like",
                                  gsub("OPC-like_Alt", "OPC-like", seurat_list$original$ManualAnnot)))

test2_results<-list()
## 1: pp+pca / pp
test2_results[["option1"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "pcaproject", FTA_reference.reduction = "pca",
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")
## 2: pp+harmony / pp
test2_results[["option2"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "pcaproject", FTA_reference.reduction = "harmony",
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")

## 3: cca / pp
test2_results[["option3"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "cca", FTA_reference.reduction = NULL,
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "cca")

## 4: rpca / pp
test2_results[["option4"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "rpca", FTA_reference.reduction = NULL,
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")

```

## Quantify accuracy of each of the above methods- what percent of cells match between original/current?
```{r}
percentAccurate<- lapply(test2_results, function(seurat_transfer){
  ## Subset to cells also found in original (only a few differences due to inferCNV)
  commonCells<- colnames(seurat_transfer)[colnames(seurat_transfer) %in% colnames(seurat_list$original)]
  seurat_transfer<-subset(seurat_transfer, cells=commonCells)
  seurat_original<- subset(seurat_list$original, cells=commonCells)
  print(sum(colnames(seurat_transfer) == colnames(seurat_original))); print(ncol(seurat_transfer)); print(ncol(seurat_original))
  
  perMatches<- round(sum(seurat_transfer$Predictions==seurat_original$ManualAnnot)/ncol(seurat_transfer),3)
  return(perMatches)
})

## bar plot of accuracy
df<-as.data.frame(t(as.data.frame(percentAccurate))); colnames(df)<- "PercentMatching"
df$settings<- c("pcaproject+pca\npcaproject",
                "pcaproject+harmony\npcaproject",
                "cca\ncca",
                "rpca\npcaproject")
ggplot(df, aes(x=settings, y=PercentMatching))+
  geom_bar(position="stack", stat="identity")+
  ylim(0,1)+
  xlab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"))+
  ggtitle("Percent accuracy in transfer vs original:\nReduction method optimization, OPC collapsed")+
  geom_hline(yintercept = max(df$PercentMatching), linetype="dotted", color="red")
ggsave(paste0(fig_folder, "ReductionMethodOptimization_OPCcollapsed.png"), width=5, height=5)
```

best is pp + pca , pp- 95.5% (just barely) same as above

## Test 3: Reduction method optimization: Compare to current seurat INCLUDING new sample
```{r}
seurat_list<-list(current=current_seurat_obj,
                  original=original_seurat_obj)


test3_results<-list()
## 1: pp+pca / pp
test3_results[["option1"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "pcaproject", FTA_reference.reduction = "pca",
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")
## 2: pp+harmony / pp
test3_results[["option2"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "pcaproject", FTA_reference.reduction = "harmony",
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")

## 3: cca / pp
test3_results[["option3"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "cca", FTA_reference.reduction = NULL,
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "cca")

## 4: rpca / pp
test3_results[["option4"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "rpca", FTA_reference.reduction = NULL,
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")

```

## Quantify accuracy of each of the above methods- what percent of cells match between original/current?
```{r}
percentAccurate<- lapply(test3_results, function(seurat_transfer){
  ## Subset to cells also found in original (only a few differences due to inferCNV)
  commonCells<- colnames(seurat_transfer)[colnames(seurat_transfer) %in% colnames(seurat_list$original)]
  seurat_transfer<-subset(seurat_transfer, cells=commonCells)
  seurat_original<- subset(seurat_list$original, cells=commonCells)
  print(sum(colnames(seurat_transfer) == colnames(seurat_original))); print(ncol(seurat_transfer)); print(ncol(seurat_original))
  
  perMatches<- round(sum(seurat_transfer$Predictions==seurat_original$ManualAnnot)/ncol(seurat_transfer),3)
  return(perMatches)
})

## bar plot of accuracy
df<-as.data.frame(t(as.data.frame(percentAccurate))); colnames(df)<- "PercentMatching"
df$settings<- c("pcaproject+pca\npcaproject",
                "pcaproject+harmony\npcaproject",
                "cca\ncca",
                "rpca\npcaproject")
ggplot(df, aes(x=settings, y=PercentMatching))+
  geom_bar(position="stack", stat="identity")+
  ylim(0,1)+
  xlab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"))+
  ggtitle("Percent accuracy in transfer vs original:\nReduction method optimization")+
  geom_hline(yintercept = max(df$PercentMatching), linetype="dotted", color="red")
ggsave(paste0(fig_folder, "ReductionMethodOptimization_NewSampleIncluded.png"), width=5, height=5)
```

best is still pp,pca,pp at 88.5

## Test 2: reduction method optimization, with OPC-* collapsed, including new sample
options for FindTransferAnchors: pp + pca, pp + harmony, rpca, cca (must run cca for transferData as well)
options for TransferData: pcaproject, cca

1. pp + pca / pp
2. pp + harmony / pp
3. cca / cca
4. rpca / pcaproject

```{r}
seurat_list<-list(current=current_seurat_obj,
                  original=original_seurat_obj)

seurat_list$original$ManualAnnot<-gsub("OPC-like_Ribosomal", "OPC-like", 
                                  gsub("OPC-like_MES", "OPC-like",
                                  gsub("OPC-like_Alt", "OPC-like", seurat_list$original$ManualAnnot)))

test4_results<-list()
## 1: pp+pca / pp
test4_results[["option1"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "pcaproject", FTA_reference.reduction = "pca",
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")
## 2: pp+harmony / pp
test4_results[["option2"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "pcaproject", FTA_reference.reduction = "harmony",
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")

## 3: cca / pp
test4_results[["option3"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "cca", FTA_reference.reduction = NULL,
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "cca")

## 4: rpca / pp
test4_results[["option4"]]<-runTransferAnchors(seurat_list=seurat_list,
                                               query_seurat_obj=seurat_list$current,
                                               ref_seurat_obj = seurat_list$original,
                                               FTA_reduction = "rpca", FTA_reference.reduction = NULL,
                                               TD_refdata = seurat_list$original$ManualAnnot,
                                               TD_weight.reduction = "pcaproject")

```

## Quantify accuracy of each of the above methods- what percent of cells match between original/current?
```{r}
percentAccurate<- lapply(test4_results, function(seurat_transfer){
  ## Subset to cells also found in original (only a few differences due to inferCNV)
  commonCells<- colnames(seurat_transfer)[colnames(seurat_transfer) %in% colnames(seurat_list$original)]
  seurat_transfer<-subset(seurat_transfer, cells=commonCells)
  seurat_original<- subset(seurat_list$original, cells=commonCells)
  print(sum(colnames(seurat_transfer) == colnames(seurat_original))); print(ncol(seurat_transfer)); print(ncol(seurat_original))
  
  perMatches<- round(sum(seurat_transfer$Predictions==seurat_original$ManualAnnot)/ncol(seurat_transfer),3)
  return(perMatches)
})

## bar plot of accuracy
df<-as.data.frame(t(as.data.frame(percentAccurate))); colnames(df)<- "PercentMatching"
df$settings<- c("pcaproject+pca\npcaproject",
                "pcaproject+harmony\npcaproject",
                "cca\ncca",
                "rpca\npcaproject")
ggplot(df, aes(x=settings, y=PercentMatching))+
  geom_bar(position="stack", stat="identity")+
  ylim(0,1)+
  xlab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"))+
  ggtitle("Percent accuracy in transfer vs original:\nReduction method optimization, OPC collapsed")+
  geom_hline(yintercept = max(df$PercentMatching), linetype="dotted", color="red")
ggsave(paste0(fig_folder, "ReductionMethodOptimization_OPCcollapsed_WithNewSample.png"), width=5, height=5)
```


best is still pp,pca,pp at 95.7%

## Test5: Query full current seurat vs 1 sample (do for each sample independently). Use best reduction settings.
```{r}
samples<- unique(original_seurat_obj$sample)

test5_results<- list()
for(s in samples){
  print(s)
  seurat_list<- list(current=subset(current_seurat_obj, sample==s),
                     original=original_seurat_obj)
  test5_results[[s]]<- runTransferAnchors(seurat_list=seurat_list,
                                          query_seurat_obj=seurat_list$current,
                                          ref_seurat_obj = seurat_list$original,
                                          FTA_reduction = "pcaproject", FTA_reference.reduction = "pca",
                                          TD_refdata = seurat_list$original$ManualAnnot,
                                          TD_weight.reduction = "pcaproject")
  
}
```

## Quantify accuracy of each of the above methods- what percent of cells match between original/current?
```{r}
percentAccurate<- lapply(test5_results, function(seurat_transfer){
  ## Subset to cells also found in original (only a few differences due to inferCNV)
  commonCells<- colnames(seurat_transfer)[colnames(seurat_transfer) %in% colnames(seurat_list$original)]
  seurat_transfer<-subset(seurat_transfer, cells=commonCells)
  seurat_original<- subset(seurat_list$original, cells=commonCells)
  print(sum(colnames(seurat_transfer) == colnames(seurat_original))); print(ncol(seurat_transfer)); print(ncol(seurat_original))
  
  perMatches<- round(sum(seurat_transfer$Predictions==seurat_original$ManualAnnot)/ncol(seurat_transfer),3)
  return(perMatches)
})

## bar plot of accuracy
df<-as.data.frame(t(as.data.frame(percentAccurate))); colnames(df)<- "PercentMatching"
df$Sample<-rownames(df)

## Manually add a row with the percent matching rate when using all samples, same reduction, also w/o opc collapse
df[nrow(df)+1,]<- c(.88, "AllSamples")
df$PercentMatching<- as.numeric(df$PercentMatching)
df$Comparison<- ifelse(df$Sample=="AllSamples", "yes", "no")

ggplot(df, aes(x=Sample, y=PercentMatching, fill=Comparison))+
  geom_bar(position="stack", stat="identity")+
  ylim(0,1)+
  xlab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"),
        legend.position = "none")+
  ggtitle("Percent accuracy in transfer vs original:\nSingle samples vs all together")+
  geom_hline(yintercept = max(df$PercentMatching), linetype="dotted", color="red")+
  scale_fill_manual(values=c(yes="red", no="black"))
ggsave(paste0(fig_folder, "LabelTransferSepForEachSample.png"), width=5, height=5)
```


## Test 6: Query full current seurat vs 1 sample (do for each sample independently). Use best reduction settings.
OPC-* collapsed
```{r}
samples<- unique(original_seurat_obj$sample)

test6_results<- list()
for(s in samples){
  print(s)
  seurat_list<- list(current=subset(current_seurat_obj, sample==s),
                     original=original_seurat_obj)
  
  seurat_list$original$ManualAnnot<-gsub("OPC-like_Ribosomal", "OPC-like", 
                                  gsub("OPC-like_MES", "OPC-like",
                                  gsub("OPC-like_Alt", "OPC-like", seurat_list$original$ManualAnnot)))
  print(table(seurat_list$original$ManualAnnot))
  test6_results[[s]]<- runTransferAnchors(seurat_list=seurat_list,
                                          query_seurat_obj=seurat_list$current,
                                          ref_seurat_obj = seurat_list$original,
                                          FTA_reduction = "pcaproject", FTA_reference.reduction = "pca",
                                          TD_refdata = seurat_list$original$ManualAnnot,
                                          TD_weight.reduction = "pcaproject")
  
}
```

## Quantify accuracy of each of the above methods- what percent of cells match between original/current?
```{r}
percentAccurate<- lapply(test6_results, function(seurat_transfer){
  ## Subset to cells also found in original (only a few differences due to inferCNV)
  commonCells<- colnames(seurat_transfer)[colnames(seurat_transfer) %in% colnames(seurat_list$original)]
  seurat_transfer<-subset(seurat_transfer, cells=commonCells)
  seurat_original<- subset(seurat_list$original, cells=commonCells)
  print(sum(colnames(seurat_transfer) == colnames(seurat_original))); print(ncol(seurat_transfer)); print(ncol(seurat_original))
  
  perMatches<- round(sum(seurat_transfer$Predictions==seurat_original$ManualAnnot)/ncol(seurat_transfer),3)
  return(perMatches)
})

## bar plot of accuracy
df<-as.data.frame(t(as.data.frame(percentAccurate))); colnames(df)<- "PercentMatching"
df$Sample<-rownames(df)

## Manually add a row with the percent matching rate when using all samples, same reduction, also w/o opc collapse
df[nrow(df)+1,]<- c(.95, "AllSamples")
df$PercentMatching<- as.numeric(df$PercentMatching)
df$Comparison<- ifelse(df$Sample=="AllSamples", "yes", "no")

ggplot(df, aes(x=Sample, y=PercentMatching, fill=Comparison))+
  geom_bar(position="stack", stat="identity")+
  ylim(0,1)+
  xlab("")+
  theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"),
        legend.position = "none")+
  ggtitle("Percent accuracy in transfer vs original:\nSingle samples vs all together, OPC collapsed")+
  geom_hline(yintercept = max(df$PercentMatching), linetype="dotted", color="red")+
  scale_fill_manual(values=c(yes="red", no="black"))
ggsave(paste0(fig_folder, "LabelTransferSepForEachSample_OPCcollapsed.png"), width=5, height=5)
```

