---
title: "R Notebook"
output: html_document
---

```{r}
base_dir = "/Users/jlabelle/"
script_dir = "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Scripts/"
working_dir<- paste0(base_dir, 
                     "Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Projects/git_repos/Pedi_Immune/",
                     "TumorCellAnnotation/Preprocessing/Preprocess_BT1873.BT1857/")


preprocessing = paste0(base_dir, script_dir, "single_cell_preprocessing_helper_functions.R")
source(preprocessing)
```

```{r load-house-keeping-genes}
marker_dir = "/Dropbox (Partners HealthCare)/FilbinLab/data_analysis/Marker_genes/"
hk_genes_dir = paste0(base_dir, marker_dir, "tirosh_house_keeping.txt")
hk_genes = readLines(hk_genes_dir)
hk_genes = hk_genes[3:length(hk_genes)]
```

```{r merge_datasets}
data_dir<- "../../../data/"

## Load and merge all count matrix
cm_files = list.files(path = data_dir, pattern="cm_tpm*")
  
  ## Subset to new samples- BT1873 and BT1857
  cm_files<- cm_files[grep(paste(c("BT1873", "BT1857"), collapse="|"), cm_files, invert =FALSE)]

cm = read.csv(paste0(data_dir, cm_files[1]), row.names = 1, stringsAsFactors = F)
for (f in cm_files[2:length(cm_files)]){
  tmp = read.csv(paste0(data_dir, f), row.names = 1, stringsAsFactors = F)
  cm = cbind.data.frame(cm, tmp)
}

## load and merge all alignment qc- optional for fresh samples
qc_files = list.files(path = data_dir, pattern="qc_*")

  ## Subset to new samples- BT1873 and BT1857
  qc_files<- qc_files[grep(paste(c("BT1873", "BT1857"), collapse="|"), qc_files, invert = FALSE)]
  
qc_align=data.frame()
for (f in qc_files[1:length(qc_files)]){
  tmp = read.csv(paste0(data_dir, f), row.names = 1, stringsAsFactors = F)
  qc_align = rbind.data.frame(qc_align, tmp)
}

## Check if cm and qc matrix have the same # of cells 
any(colnames(cm) != rownames(qc_align))


## rename BT1873 to reflect pons/ven
subtypes[subtypes$sample=="BT1873", "sample"]<- paste0(subtypes[subtypes$sample=="BT1873", "sample"],
                                                       subtypes[subtypes$sample=="BT1873", "Location"])
subtypes$sample<- gsub("ventricle", "vent",subtypes$sample)

sum(subtypes$sample %in% qc_align$sample)
qc_align_subtypes<- merge(qc_align, subtypes, by="sample")
rownames(qc_align_subtypes)<- rownames(qc_align)
qc_align<-qc_align_subtypes

## Save merged datasets as rds files
saveRDS(cm, file="analysis/cm.rds")
saveRDS(qc_align, file="analysis/qc_align.rds")
rm(cm);rm(qc_align);gc()
```

```{r load-raw-data}
## Load counts and alignment-qc
cm <- readRDS("analysis/cm.rds")
qc_align <- readRDS("analysis/qc_align.rds")

## Check if cb matches
any(colnames(cm) != rownames(qc_align))

## Make a vector of sample names
samples = qc_align[,"sample"]; names(samples) = rownames(qc_align)
```

## Compute QC metrics on all cells 
```{r qc_on_raw_data}
qc_raw_data = calculateQcMetrics(cm, hk_genes)
qc_raw_data$align_rate = qc_align$nAligned/qc_align$nTotal
```

## Scatter plot of # of genes and mean expr of house keeping genes- optional, use chunk below to color by PF 
```{r}
#Create qc figure directory. figures directory should already be created
qc_fig_dir = "figures/"
if (!dir.exists(qc_fig_dir)){
  dir.create(qc_fig_dir)
}

pairs(qc_raw_data[,4:6])

ggplot(qc_raw_data, aes(x=gene, y=hk))+
  geom_point()+
  xlab("Total number of genes per cell")+
  ylab("Mean expression of house keeping genes")+
  geom_vline(xintercept=2000, color="red", linetype="dashed", size=1)+
  geom_hline(yintercept = 2.5, color="red", linetype="dashed", size=1)+
  theme_classic()+
  theme(axis.title = element_text(face="bold", size=24),
        axis.text = element_text(face="bold", size=20))

    
  
                          

ggsave(filename=paste0(qc_fig_dir, "genes_vs_hk.jpg"), width=12, height=8)
```

## Scatter plot of # of genes and mean expr of house keeping genes, colored by QC
```{r}
#Set cutoff values
num_gene_cutoff = 2000
hk_mean_expr_cutoff = 2.5

#Add column delineating pass/fail of cell
qc_raw_data$pass_qc = ifelse(qc_raw_data$gene >= num_gene_cutoff & qc_raw_data$hk >= hk_mean_expr_cutoff, T, F)
sum(rownames(qc_raw_data) == rownames(qc_align))
qc_raw_data$subtype<-qc_align$subtype

ggplot(data=qc_raw_data, aes(x=gene, y=hk, color=pass_qc)) + geom_point() + 
  geom_hline(yintercept = hk_mean_expr_cutoff, color="red", linetype="dashed", size=1) +
  geom_vline(xintercept = num_gene_cutoff, color="red", linetype="dashed", size=1) +
  xlab("Total number of genes\n") + ylab("Mean expression of \n house-keeping genes") +
  theme_classic() + theme(plot.title = element_text(hjust = 0.5, face="bold", size=24),
                          axis.title = element_text(face="bold", size=24, color="black"),
                          axis.text = element_text(face="bold", size=18, color="black"),
                          legend.title = element_text(face="bold", size=24, color="black"),
                          legend.text = element_text(size=20))
ggsave(filename=paste0(qc_fig_dir, "genes_vs_hk.jpg"), width=8, height=8)
```

## Histogram of # of genes and mean expr of house keeping genes 
```{r}
## Histogram of total number of genes 
ggplot(data=qc_raw_data, aes(x=sample, y=gene, fill=sample)) + 
  geom_violin() + geom_hline(yintercept = num_gene_cutoff, color="red", linetype="dashed", size=1) +
  ggtitle("Total number of detected genes\n") + xlab("") + ylab("Number of genes") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold", color="black", size = 28),
        axis.title = element_text(face = "bold", color="black", size = 24), 
        axis.text.x = element_text(angle = 45, hjust = 1, size=18, color="black", face="bold"),
        axis.text.y = element_text(size = 18, color="black",face = "bold"),
        legend.title = element_text(face = "bold",  color="black",size = 24),
        strip.text = element_text(size=24, face="bold"),
        legend.position = "none")
  #facet_wrap(.~ subtype, scales = "free_x") 
ggsave("total_genes.jpg", path=qc_fig_dir, width = 15, height = 6, units = "in")

## Histogram of average expr of house keeping genes
ggplot(data=qc_raw_data, aes(x=sample, y=hk, fill=sample)) + 
  geom_violin() + geom_hline(yintercept = hk_mean_expr_cutoff, color="red", linetype="dashed", size=1) +
  ggtitle("House keeping gene expression\n") + xlab("") + ylab("Log2 average expression of house keeping genes\n") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold", color="black", size = 28),
        axis.title = element_text(face = "bold", color="black", size = 24), 
        axis.text.x = element_text(angle = 45, hjust = 1, size=18, color="black"),
        axis.text.y = element_text(size = 24, color="black",face = "bold"),
        legend.title = element_text(face = "bold",  color="black",size = 24),
        strip.text = element_text(size=24, face="bold"),
        legend.position = "none")+
  facet_wrap(.~ subtype, scales = "free_x") 
ggsave("hk_Subtype.jpg", path=qc_fig_dir, width = 8, height = 10, units = "in")
```

## Preprocessing: filtering on cells and genes 
```{r preprocessing}
## Filter and normalize cm
cm_list = normalizeTPM(cm, 
                       min_genes=num_gene_cutoff, 
                       min_mean_hg = hk_mean_expr_cutoff,
                       hg_list=hk_genes, 
                       filter_gene_relaxed = T, 
                       gene_cutoff=log2(10), 
                       gene_cell_cutoff = 10)
## Subset sample
samples = samples[colnames(cm_list$raw_data)]

## Calculate and store qc metrics for filtered and unfiltered data 
qc_filtered_data = calculateQcMetrics(cm_list$raw_data, hk_genes)
```


## Save preprocessed results
```{r save-preprocessed-results}
save(cm_list, file="analysis/cm_list.Robj")
saveRDS(qc_align, file="analysis/qc_align.rds")
saveRDS(qc_raw_data, file="analysis/qc_raw_data.rds")
saveRDS(qc_filtered_data, file="analysis/qc_filtered_data.rds")
saveRDS(samples, file="analysis/samples.rds")

```

## Barplot of good/poor quality cells for each sample 
```{r barplot_of_qc}
## # of cells that pass QC in each sample 
crosstab = table(qc_raw_data$sample, qc_raw_data$pass_qc)
crosstab = data.frame(crosstab)
colnames(crosstab) = c("sample", "Good_quality", "Frequency")
crosstab<-merge(crosstab, subtypes, by="sample")
ggplot(crosstab, aes_string(x="sample", y="Frequency", fill="Good_quality")) + 
  geom_bar(stat="identity", color="black") + ggtitle("QC of each sample\n") + 
  xlab("") + ylab("N cells") +
  theme_classic() + 
   theme(plot.title = element_text(hjust = 0.5, face = "bold", color="black", size = 28),
        axis.title = element_text(face = "bold", color="black", size = 24), 
        axis.text.x = element_text(angle = 45, hjust = 1, size=18, color="black", face="bold"),
        axis.text.y = element_text(size = 24, color="black",face = "bold"),
        legend.title = element_text(face = "bold",  color="black",size = 24),
        strip.text = element_text(size=24, face="bold"))
  #facet_wrap(.~ subtype, scales = "free_x") 
ggsave("num_cells_pass_qc.jpeg", path=qc_fig_dir, width=15, height=6)

## % of cells that pass QC in each sample 
crosstab = table(qc_raw_data$sample, qc_raw_data$pass_qc)
crosstab = crosstab/rowSums(crosstab)
crosstab = data.frame(crosstab)
colnames(crosstab) = c("sample", "Good_quality", "Percentage")
crosstab<-merge(crosstab, subtypes, by="sample")
ggplot(crosstab, aes_string(x="sample", y="Percentage", fill="Good_quality")) + 
  geom_bar(stat="identity") + 
  geom_hline(yintercept = 0.5, color="red", linetype="dashed", size=1) +
  ggtitle("Percentage of good quality cells in each sample") + 
  #coord_flip() + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold", color="black", size = 28),
        axis.title = element_text(face = "bold", color="black", size = 24), 
        axis.text.x = element_text(angle = 45, hjust = 1, size=18, color="black"),
        axis.text.y = element_text(size = 24, color="black",face = "bold"),
        legend.title = element_text(face = "bold",  color="black",size = 24),
        strip.text = element_text(size=20, face="bold"))+
  facet_wrap(.~ subtype, scales = "free_x")
ggsave("percentage_cells_pass_qc_Subtype.jpeg", path=qc_fig_dir, width=10,height=8)
```


```{r}


```


## Process to seurat
```{r}
library(harmony)

## remove BT1873 pons- only 3 cells PF
load("analysis/cm_list.Robj")
samples<- readRDS("analysis/samples.rds")

cm<- cm_list$raw_data[, !grepl("pons", colnames(cm_list$raw_data))]
samples<-samples[samples != "BT1873pons"]
sum(names(samples)!=colnames(cm))

## Run seurat
seurat_obj<- RunFullSeurat(cm, RunHarmony = TRUE, samples=unname(samples))

## Subset to just BT1873
bt1873<- subset(seurat_obj, sample=="BT1873vent")
bt1873<- RunFullSeurat(bt1873@assays$RNA@counts, RunHarmony = FALSE, samples=bt1873$sample)
saveRDS(bt1873, file=paste0("analysis/seurat_obj_BT1873.Rds"))
```

